### 2015-05-09 paper/paper.tex
Initial commit

\input{header.tex}
\input{defs.tex}

\addbibresource{./managed.bib}

\title{Changing Flights In The Air\\ (Abortable Chain Agreement)}

\author{
  Rachid Guerraoui\\
  \texttt{rachid.guerraoui@epfl.ch}
  \and
  Giuliano Losa\\
  \texttt{giuliano.losa@epfl.ch}
}

\date{}

\begin{document}

\maketitle

\abstract{
    To achieve consistent performance, a distributed computing service must
    dynamically adapt to unpredictable crashes, communication failures, and
    changes in the availability of shared resources.
    Unfortunately, most existing algorithms are not adaptive: outside of their
    sweet spot, they have poor performance.

    We present a new abstraction, Abortable Generalized Consensus, which enables the
    correct-by-construction composition of heterogeneous State-Machine
    Replication algorithms.

    integrating state of the art techniques to optimize executions.  ACA is the most complete approach to date enabling the
    correct-by-construction composition of heterogeneous State-Machine
    Replication algorithms. We provide two specifications satisfying
    ACA and we illustrate their power by refining them to
    obtain a versatile adaptive algorithm that combines five existing
    algorithms into a single one. ACA is completely formalized and its
    underlying composition theorem has been mechanically checked by an interactive
    proof assistant.
}

\vspace{2cm}

\begin{center}
This paper is a regular submission.

\vspace{1cm}
This paper is a student paper.
\end{center}

\newpage

\section{Introduction}

The State-Machine Replication technique (abbreviated SMR) allows
building reliable services on top of unreliable hardware. The basic idea is to 
replicate a service over several servers, each executing the same
sequence of (deterministic) commands. 
Making SMR algorithms efficient has been a long-standing
challenge and a multitude of algorithms have been proposed. Examples
include Fast Paxos \cite{Lamport06FastPaxos}, Disk Paxos
\cite{GafniLamport03DiskPaxos}, Chain Replication
\cite{RenesseSchneider04ChainReplicationSupportingHighThroughputAvailability},
Ring Paxos
\cite{MarandiETAL10RingPaxosHighthroughputAtomicBroadcastProtocol},
Egalitarian Paxos
\cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments},
Multi-Coordinated Paxos
\cite{CamargosSchmidtPedone07MulticoordinatedPaxos}, Cheap Paxos
\cite{LamportMassa04CheapPaxos}, Vertical Paxos
\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication},
Paxos-MIC
\cite{HurfinMoiseNarzul11AdaptiveFastPaxosMakingQuickEverlasting},
Mencius
\cite{MaoJunqueiraMarzullo08MenciusBuildingEfficientReplicatedStateMachine},
and Fast Mencius \cite{WeiETAL13FastMenciusMenciusLowCommitLatency}.

Each of the existing algorithms is best suited to a particular range
of environments, characterized, e.g., by client load, network
topology, likelihood of faults, communication latency and bandwidth, processing
power of nodes, availability of disk storage and memory.
However, the environment faced by real systems evolves over time:
hardware is upgraded, new data-centers are built, the popularity of a
service changes, etc.  Consistent performance therefore comes with
adaptation to change.  Unfortunately, existing SMR algorithms cannot easily be
stopped and replaced by another algorithm on the fly. Moreover,
devising a new algorithm that can change behavior dynamically is hard:
dynamically switching from one SMR algorithm to another requires
synchronizing notoriously intricate algorithms.

We propose a new abstraction, \emph{Abortable Chain Agreement}
(abbreviated ACA), which subsumes SMR and whose implementations,
called \emph{ACA rounds}, can, by construction, pass the baton from
one to another on the fly and efficiently. Moreover, optimizing a replicated
service consisting of several ACA rounds becomes just a matter of devising a new
ACA round, which can then be plugged-in on the fly: any new ACA round
is compatible with every existing ACA round by construction.

Abortable Chain Agreement is the \emph{most complete framework to date} enabling
correct by construction composition of heterogeneous rounds.  
ACA is completely formalized and its underlying composition theorem has been
mechanically checked by an interactive proof assistant.  
Compared to ACA,
existing approaches
\cite{BarNoyETAL87ShiftingGearsChangingAlgorithmsFlyTo,GuerraouiETAL10Next700BftProtocols,GuerraouiKuncakLosa12SpeculativeLinearizability}
are limited. In particular, they do not allow the disorderly execution of
commuting and read-only commands, an important practical optimization
which is at the root of recent work on SMR like Generalized Consensus
\cite{Lamport05GeneralizeConsensus} and Generalized Lattice Agreement
\cite{FalerioETAL12GeneralizedLatticeAgreement}.

\begin{comment}
On top of the traditional roles of proposer, acceptor, and learners
\cite{lamport2001paxos} used in consensus algorithms, ACA introduces
the notion of \emph{switchers} and of \emph{scheduling policy}. The switchers
have the task of aborting a running ACA round and initializing a new
ACA round in a safe and live manner.  The \emph{scheduling policy}
decides when to abort an ACA round and choses the next ACA round to
execute. It must ensure that all switchers agree on the next ACA round
to initialize.
\end{comment}

We observe that, because of the FLP impossibility result
\cite{FischerLynchPaterson83ImpossibilityDistributedConsensusOneFaultyProcess},
most existing SMR algorithms are structured in a sequence of abortable
rounds which can easily be transformed into ACA rounds. To simplify
this task even further, we propose two specifications, \emph{conservative
ACA rounds} and \emph{fast ACA rounds}, which, as we show, can be concretized
to obtain ACA versions of the rounds of five existing algorithms: 
Classic Paxos, Fast Paxos, Chain Replication, Ring Paxos,
and Multi-Coordinated Paxos. ACA guarantees that the composition of the five ACA
rounds is correct by construction, forming a versatile adaptive replication
algorithm.

The rest of the paper is structured as follows.
\Cref{sec:related} surveys recent related work. \Cref{sec:ca} presents Chain
Agreement, which embodies the requirements for implementing SMR. \Cref{sec:aca}
presents Abortable Chain Agreement and its optimizations to treat commuting
and read-only commands efficiently. \Cref{sec:rounds} presents our specification
of conservative and fast ACA rounds, which implement ACA. In
\cref{sec:examples}, we show that the rounds of  
Classic Paxos, Fast Paxos, Chain Replication, Ring Paxos,
and Multi-Coordinated Paxos can easily be transformed into ACA implementations
by using our specifications of conservative and fast ACA rounds.
Finally, in \cref{sec:isaproofs}, we provide a complete formalization of ACA
and a mechanical proof of the \emph{composition theorem} (\cref{thm:comp}),
checked by the interactive proof assistant Isabelle/HOL \cite{NipkowPaulsonWenzel02IsabelleHOL}.

\begin{comment}
Abortable Chain Agreement builds on several recent advances in SMR
algorithms: ACA allows acceptors to only eventually reach agreement,
like in Generalized Paxos \cite{Lamport05GeneralizeConsensus},
boosting performance when most commands commute; ACA allows fast
execution of read commands like in Generalized Lattice Agreement
\cite{FalerioETAL12GeneralizedLatticeAgreement}; ACA implementations
can tolerate $f$ faults with only $f+1$ replicas and replace crashed replicas
upon changing round, as in Vertical Paxos  \cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}.  
\end{comment}

\section{Related Work}
\label{sec:related}

SMR is often implemented using a sequence of instances of Consensus.
There are numerous examples of consensus algorithms in the literature.
However, Abortable Chain Agreement is based on Generalized Consensus. 
In Generalized Consensus \cite{Lamport05GeneralizeConsensus}, the learners of a
SMR algorithm repeatedly learn new values belonging to a \emph{C-struct set}. A
C-struct can be seen as a data structure encoding  sequences of commands modulo
the ordering of commands that commute. The advantage of Generalized
Consensus over Consensus is that
two servers which receive commuting commands in different orders can still reach
agreement on C-structs without further synchronization. 
Our Abortable Chain Agreement can also use C-structs, with the same advantages
as Generalized Consensus. For simplicity, the presentation in this
paper uses the more concrete notion of \emph{command history} (see
\cref{sec:histories}).

Generalized Lattice Agreement \cite{FalerioETAL12GeneralizedLatticeAgreement} allows more efficient implementations of SMR when
commands are partitioned into read-only commands and write-only commands.
Abortable Chain Agreement supports the same optimizations as Generalized Lattice Agreement.
Notably, read-only commands are treated specially so as to optimize their
executions (see \cref{sec:readonly}).

Vertical Paxos algorithms 
\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication} are SMR algorithms
whose set of replica servers can be reconfigured without stopping to process new
commands. Crashed servers can therefore be replaced by new servers, ensuring the
long-term availability of a replicated service.
Vertical Paxos algorithms use an external reconfiguration master, itself
implemented as a replicated state machine, which determines the servers that should
be removed from or added to the configuration.
ACA shares some aspects of Vertical Paxos algorithms: the task of the
reconfiguration master is similar to the task of the scheduling policy
component of ACA (see \cref{sec:aca}).
However, in ACA, changing round can be used not only to replace crashed servers
but also to change the algorithm that the servers are running. 
Changing algorithm is not possible in Vertical Paxos.
In both Vertical Paxos and ACA, a reliable reconfiguration master, resp. a
reliable scheduling policy, allows tolerating $f$ crashes with $f+1$ servers.

The Abstract Framework \cite{GuerraouiETAL10Next700BftProtocols} proposes to
build Byzantine Fault-Tolerant algorithms as the succession of abortable rounds
called Abstract instances. Abstract instances must satisfy the Abstract
correctness properties. By construction, the composition of any number of
Abstract instances is a correct SMR implementation. However, the Abstract
Framework uses totally ordered sequence of commands, making difficult the
optimization of the execution of commuting and read-only commands that are
possible with ACA.  Moreover, the Abstract Framework does not identify the
crucial role of the scheduling policy for the long-term resilience of a service.
Abortable Chain Agreement combines the ideas of the Abstract Framework,
Generalized Paxos, Generalized Lattice Agreement, and Vertical Paxos in a
unified abstraction.

Finally, the Speculative Linearizability framework
\cite{GuerraouiKuncakLosa12SpeculativeLinearizability} was a first
attempt by the authors of the present paper at generalizing the
Abstract Framework and mechanically proving its properties with the
Isabelle/HOL interactive
proofs assistant.  Like the Abstract Framework, the
Speculative Linearizability framework is based on totally ordered
sequences of commands, with the same drawback for the execution of
commuting and read-only commands. Moreover, Abortable Chain Agreement is both
conceptually simpler and subsumes the Speculative Linearizability
property.  Like Speculative Linearizability, Abortable Chain Agreement
and its main property, the composition theorem (\cref{thm:comp}), are
formalized in Isabelle/HOL (see \cref{sec:isaproofs}).

Abortable Chain Agreement is the most complete framework to date for building
adaptive replication algorithms out of heterogeneous rounds, making state of the
art optimization techniques available to algorithm designers, which was not
possible before.

\section{Chain Agreement}
\label{sec:ca}

We consider a set of asynchronous servers communicating by message passing in an
asynchronous network. Servers are not necessarily sequential processes. We
describe a behavior of the system using an infinite sequence of states, where a
state is a function from variable names to values (i.e. a valuation
of variables).  We specify sets of allowed behaviors using an initial predicate,
a next-state relation (using unprimed and primed variables) and a fairness
condition, all in the style of TLA+ \cite{Lamport02SpecifyingSystems}. However,
for succinctness, our specification are rather informal, eluding non-essential
details.

Our specifications all contain by default a state variable $crashed$ and
$Crash\left( x \right)$ transition which is always enabled. A server $x$ faithfully
executes the algorithm that is assigned to it until its identifier is added to
the variable $crashed$ by a $Crash\left( x \right)$ transitions.  When the identifier of a
server is in the set $crashed$, all its actions are disabled.  In an execution,
a server that is never added to the set $crashed$ is said correct.

We consider a computing service exposing a set of commands $C$. The
service has a sequential specification consisting of an initial state
and a deterministic transition relation in which every command
atomically updates the state of the service and produces an output. We
assume that duplicate commands have no effect on the state of the
service and return the same output as their first occurrence.

In a traditional SMR algorithm, the servers can be seen has playing one or
several of three roles: \emph{proposers}, which propose new commands to execute,
\emph{acceptors}, which accept and order commands in a sequence, and
\emph{learners}, which learn about the growing sequence of commands to
execute \cite{lamport2001paxos}. Often, but not necessarily, every replica server plays the
three roles of proposer, acceptor, and learner: clients of the service
send their commands to some replica servers; replica servers submit
the received commands to the SMR protocol in the role of proposers;
replica servers respond to commands by executing the commands learned
in the role of learners; replica servers participate in command
ordering in the role of acceptors.  An SMR algorithm guarantees that
the replicas learn the same sequences of commands, possibly up to the
reordering of commuting commands. Therefore, every learner can respond to client
requests using its last learned sequence of commands, and clients will receive
the same responses as the ones produces by a centralized, non-replicated,
service.

Chain Agreement precisely specifies the allowed behaviors of the roles of \emph{proposers}
and \emph{learners}, abstracting over the behavior of acceptors.

\begin{comment}
We suppose that each server can play a number of \emph{roles}.
Chain Agreement specifies the allowed behaviors the roles of \emph{proposers}
and \emph{learners}. Proposers receive commands from the clients of the service
and propose them for execution by the replicated service. Each learner learns about
a growing sequence of commands. Thanks to the sequential specification of the
service, a sequence of commands determines an output for each the included commands. 
In a traditional SMR algorithm, servers can also play the role of
\emph{acceptors}, whose task is to collectively order commands and let the
learners know about the latest sequence of ordered commands.

The point of Chain Agreement is to implement a replicated service by having every learner learn about a growing sequence of commands, which
represent an execution of the sequential specification of the service.
Crucially, at all times, the sequences of commands learned by the
different learners must form a chain in the prefix order on sequences.
Moreover, for every command $c$ proposed by a correct proposer, every
correct learner must eventually learn a sequence containing $c$.  Once
a learner learns about a sequence containing a command $c$, it can
determine the output produced by $c$, using the sequential
specification of the service, and respond to the client that issued
the command $c$.
\end{comment}

To specify Chain Agreement, we need the notion of ordering of sequences and of
maximum sequence.  The prefix order on sequences is denoted by the operator
$\leq$ ($s_1 \leq s_2$ means that $s_1$ is a prefix of $s_2$). 
%Using the operator $\leq$, we define $\geq$, $<$, and $>$ as expected. 
Note that $\leq$ is a partial order on sequences. The maximum over a set of
sequences $S$ is denoted by $Max\left( S \right)$ and $Max\left( \aset{}
\right)$ is the empty sequence.

We specify Chain Agreement using the state variables $crashed$, $proposed$, containing the proposed commands,
and, for every learner $l$,  $learned\left[ l \right]$, a set of
sequences of commands. Initially, $crashed$, $proposed$ and, for every learner $l$,
$learned\left[ l \right]$ are empty sets. 
Remember that any transition involving a server $x$ is always disabled when $x\in crashed$.

The next-state relation consists of three types of transitions:
$Crash\left( x \right)$ transitions; $Propose\left( l, c
\right)$ transitions, for a learner $l$ and a command $c$, in which
$c$ is inserted in the set $proposed$; $Learn\left( l,cs \right)$
transitions, for a learner $l$ and a sequence of commands $cs$, in
which $cs$ is inserted in $learned\left[ l \right]$. 
A $Learn\left( l, cs \right)$ transition is guarded by the following
conditions:
\begin{compactitem}
    \item[\textbf{Agreement}:] The elements of the set $\aset{learned'\left[ l
      \right] : l \in Learners}$, the learned sequences after the transition,
      form a chain in the partial order $\leq$.
    \item[\textbf{Validity}:] Every learned sequence contains only
        proposed commands.  
    \item[\textbf{Irrevocability}:] The new learned sequence
        $cs$ is a strict extension of $Max\left(learned\left[ l
        \right]\right)$. 
        %\\ $cs > Max\left(learned\left[ l \right]\right)$.
\end{compactitem}
Finally, we add, for every learner $l$, one weak fairness condition
\cite{Lamport02SpecifyingSystems} on the transitions $Learn\left( l,
cs \right)$. The weak fairness conditions ensure that every correct
learner eventually learns about every proposed command.

Thanks to the properties of Chain Agreement, the clients, observing
their input commands and the received responses, see a
\emph{linearizable} execution
\cite{HerlihyWing90LinearizabilityCorrectnessConditionConcurrentObjects}.
Moreover, the service remains available as long as at least one
proposer and one learner are correct.

\section{Abortable Chain Agreement}
\label{sec:aca}

We now consider implementing Chain Agreement as a sequence of \emph{Abortable Chain
Agreement rounds}, where each round makes progress ordering requests or aborts
its execution and passes the baton to the next round in the sequence.
An algorithm composed of a sequence of ACA rounds is called an \emph{ACA
algorithm}.

On top of the traditional roles of proposer, acceptor, and learner, ACA
introduces the new role of \emph{switcher} and a new component, the
\emph{scheduling policy}. In ACA, the assignment of the roles of proposer,
acceptor, learner, and switcher can change in each rounds, allowing for
reconfiguration. 
In normal circumstances, the proposers, acceptors,
and learners of an ACA round work together to implement Chain Agreement.
However, if the environment becomes unfavorable to a running ACA round, the
scheduling policy can instruct the round to abort its execution and pass the
baton, using the switchers, to a new round.

A switcher of an aborting round determines, from the state of the
acceptors, an \emph{abort sequence} that it transfers to one or several
switchers of the new round.  Upon receiving an abort sequence, a switcher of the
new round initializes the acceptors of the new round, which can then
proceed implementing SMR, until the new round itself aborts and the
process is repeated. 
We assume that in the first round, which has no previous round, every 
switchers receives the empty abort sequence.

The scheduling policy determines when to abort an ACA round and which new ACA
round should take over. The scheduling policy must inform the switchers about the
identity of the next round and must guarantee that all switchers receive the
same information.  In this paper we abstract over the scheduling policy, assuming
that a round that aborts passes the baton a nondeterministically chosen new
round. A reliable scheduling-policy component can realistically be implemented
as a replicated state machine running on a large number of servers otherwise
dedicated to other tasks
\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}. We will see
 in \cref{sec:rounds} that, as in Vertical Paxos, a reliable scheduling policy component allows
tolerating $f$ faults with $f+1$ replicas, while maintaining the long-term
resilience of the service by replacing faulty servers upon round change.

The clients of an ACA algorithm can send their commands to the proposers of any round (in
practice, clients send their commands to proposers of the largest
active round) and servers can use sequences learned in any round
to determine outputs.

ACA specifies the allowed behaviors of the proposers, learners, and switchers
\emph{of a single round}. Similarly to
Chain Agreement, we use the state variables $crashed$, $proposed$,
and, for every learner $l$, $learned\left[ l \right]$.  Additionally,
we use the state variables $inits$, containing the abort sequences
received from the previous round, and $aborts$, containing the abort
sequences transfered to the next round.
Two consecutive ACA rounds are composed by letting the $aborts$ variable of the
first round be the $inits$ variable of the second round.

Initially, $crashed$, $proposed$, for every
learner $l$, $learned\left[ l \right]$, $inits$, and $aborts$ are all empty sets.
The next-state relation consists of five types of transitions. The
first three, $Crash\left( x \right)$, $Propose\left( l, c
\right)$, and $Learn\left( l, cs \right)$, for a learner, proposer, or
switcher $x$, for a learner $l$, and for a sequence of commands
$cs$, have the same effect as in Chain Agreement.
ACA introduces two new transitions, $Init\left(
s, cs \right)$ and $Abort\left( s, cs \right)$, for a switcher $s$ and a sequence of
commands $cs$. The transition $Init\left( s, cs \right)$ adds $cs$ to the set $inits$, and models a switcher receiving the
abort sequence $cs$ from the preceding round.
The transition $Abort\left( s, cs \right)$ adds $cs$ to the set $aborts$ and models a switcher transferring the abort sequence $cs$ to a switcher of the next round.
All the transition are guarded by the following conditions:
\begin{compactitem}
    \item[\textbf{Agreement}:] The elements of the set $\aset{learned'\left[ l
      \right] : l \in Learners}$ form a chain in the partial order.
    \item[\textbf{Validity}:] Every abort sequence or learned sequence
        is of the form $h \star cs$, where $cs$ contains only commands
        of the set $proposed$ and $h$ is the greatest lower bound of a
        nonempty subset of $inits$.
    \item[\textbf{Initialization}:] If $inits$ is empty, then no
        sequence can be learned and no switcher can abort.
    \item[\textbf{Irrevocability}:] In the case of a $Learn\left( l,cs
        \right)$ transition, the new learned sequence $cs$ is a strict
        extension of $Max\left(learned\left[ l \right]\right)$.
        %, $cs
        %\geq Max\left(learned\left[ l \right]\right)$.
    \item[\textbf{Safe Abort}:] Every abort sequence is larger in the
        partial order $\leq$ than every learned sequence.
\end{compactitem}
Note that Agreement and Irrevocability are the same as in Chain
Agreement.

Finally, for every learner $l$, we add a weak fairness condition on the
following state transition: 
\begin{equation}
    \exists  s \in Switchers, cs \in Sequence\left( C \right) : Learn\left( l, cs \right) \vee Abort\left( s,cs \right)
\end{equation}
The weak fairness conditions ensure that when a correct learner has not
learned a sequence containing all commands proposed in the round, then
it eventually learns a new command or, if there is at least one correct switcher, at least one switch value is transfered to the next round. 
Therefore, if a client has submitted a command to a proposer, eventually,
either the client gets a response from the current round, or
the next round is initialized with an abort sequence and the
command can be resubmitted to the next round.

\subsection{The Composition Theorem}

Consider the succession of $n$ rounds numbered $1$ to $n$.  To
distinguish the variables of each round, we annotate the local
variables of each round with the number of the round. For example, the
variables of the seventh round are written $proposed_7$, for every
learner $l$, $learned_7\left[ l \right]$, $inits_7$, and $aborts_7$.

Moreover, we assume that a command, which is proposed to a round and that
is not learned before the first abort sequence is transfered to the
next round, is proposed anew to the next round.

Define the state variable $proposed$ as the union over all rounds $i$
of the $proposed_i$ variables and, for every learner $l$, 
define the $learned\left[ l \right]$ variable as the union over all
rounds $i$ of the $learned_i\left[ l \right]$ variables.

\begin{thm}[Composition Theorem]
    \label{thm:comp}
    In every execution of a succession of ACA rounds, the sequence of valuations
    of the variable $proposed$ and, for every learner $l$, of the variables
    $learned\left[ l \right]$ satisfies Chain Agreement.
\end{thm}
\Cref{thm:comp} can be shown by induction on the number of rounds,
using the following lemma.
\begin{lem}
  \label{lem:comp}
    Consider two rounds numbered $i$ and $i+1$. In every execution of a
    succession of ACA rounds, the sequence of valuations of the variables $inits
    = inits_i$, $propose = proposed_i \cup propose_{i+1}$, for every learner
    $l$, $learned\left[ l \right] = learned_i\left[ l \right] \cup
    learned_{i+1}\left[ l \right]$and $aborts = aborts_{i+1}$ satisfies ACA.
\end{lem}
\begin{proof}
  A mechanically-checked proof in Isabelle/HOL appears in \cref{sec:isaproofs}.
  The proofs uses a version of the specification of ACA that has a more
  executable form that presented in this section. Moreover,
  instead of sequences of commands, the Isabelle/HOL specification uses the more
  general notion of command history, presented in the next subsection.
  The Isabelle/HOL formalization of ACA uses the theory of I/O automata, which
  is also formalized in the Isabelle/HOL. The proof uses a refinement
  mapping from the composition of two consecutive ACA instances to a single ACA
  instance.
\end{proof}

\Cref{thm:comp} guarantees that the successive composition of any
number of ACA rounds implements Chain Agreement.
Therefore, with ACA, one can easily build replicated services that
dynamically adapt to the current operating conditions of the system: a
replicated service has several types of rounds at its disposal, each being
tailored to particular conditions. Each time the conditions change,
the current ACA round aborts and passes the baton to a more 
appropriate type of round, as determined by the scheduling policy. Moreover, a new type
of ACA round can always be devised after the system is deployed and added on the fly to
the existing types of rounds.

We now describe two important practical optimizations of ACA.

\subsection{Histories}
\label{sec:histories}

ACA can be implemented more efficiently by substituting \emph{command histories} for sequences of commands, as in Generalized Consensus \cite{Lamport05GeneralizeConsensus}.

We say that two commands $c_1$ and $c_2$ \emph{commute} when, in every sequential
execution of the service, reversing the order of $c_1$ and $c_2$ does
not change any of the outputs, including those given to $c_1$ and
$c_2$.

A command history, abbreviated history, is a data structure that
represents a set of sequences which are the same modulo the reordering
of commuting commands, and in which duplicate commands do not appear.
All the sequences represented by a given history contain the same
commands and determine the same outputs for each command.  Therefore,
to implement a replicated service, it is sufficient that learners
learn command histories instead of sequences.  For a precise
definition of command histories, we refer the reader to
\cite{Lamport05GeneralizeConsensus}. Our Isabelle/HOL formalization
(\cref{sec:isaproofs}) abstracts
over the concrete representation of histories, assuming only their necessary
properties. Command histories were first
introduced in trace theory \cite{Mazurkiewicz84Semantics}.

A command history can be built by appending a sequence of commands
$cs$ to a history $h$, obtain the history $h \star cs$.  The major
advantage that histories have over sequences is that if a sequence
$cs_1$ is obtained by reordering the commuting commands of $cs_2$,
then $h \star cs_1 = h \star cs_2$. Therefore, in  implementations of
ACA, servers can agree on a history even though
they chose different orders for some commuting commands.

Command histories share the four properties of sequences that are
necessary and sufficient to substitute histories for sequences in the
definition of Chain Agreement and ACA.
\begin{compactitem}
    \item[\textbf{Distinctness}:] No duplicate command appears in a history. 
    \item[\textbf{Ordering}:] Command histories form a partial order:
        $h_1 \leq h_2$ when there exists a sequence $s$ such that $h_2
        = h_1 \star s$.
    \item[\textbf{GLB}:] Every finite set of command histories has a
        \emph{greatest lower bound}.
%    \item[\textbf{LUB}:] Every finite set of compatible command
%        histories has a \emph{least upper bound}, where a set of
%        command histories is compatible when they have a common upper
%        bound.
    \item[\textbf{Consistency}:] If $h_1 \leq h_2 \leq h_3$ and $h_3 =
        h_1 \star s$ then there exists $s'$ and $s''$ where every
        request of $s'$ and $s''$ appears in $s$ and such that $h_2 =
        h_1 \star s'$ and $h_3 = h_2 \star s''$.
\end{compactitem}
The optimized version of ACA is obtained by
simply substituting histories for sequences in the definition of ACA.
The Ordering, GLB, and Consistency properties of histories ensure that
\cref{thm:comp} still holds: the Isabelle/HOL proof of \cref{thm:comp} (see
\cref{sec:isaproofs}) relies only on the four properties above.

\subsection{Optimizing the Execution of Read-Only Commands}
\label{sec:readonly}

Chain Agreement can be further optimized by treating 
read-only commands specially, as in Generalized Lattice Agreement
\cite{FalerioETAL12GeneralizedLatticeAgreement}.

We consider the case in which a set of servers each play the four
roles of proposer, acceptor, learner, and switcher.
Clients sends their commands to any server, which proposes them as a
proposer, and records the command as pending.
%When a server learns a history that contains one of its pending
%commands, it uses the history to determine the response to the pending
%command and forwards the response to the appropriate client.

As observed in \cite{FalerioETAL12GeneralizedLatticeAgreement}, read-only
commands can be executed faster by directly reading from the last learned
history, instead of proposing the read commands and waiting for it to appear in
a future learned history.  Because learned histories form a chain, all the reads
return consistent values and this approach yields a \emph{serializable}
implementation of the service.

However linearizability is not satisfied if reads are served directly
from the last learned history: a lagging replica server could respond
to a read using a history which does not contain all completed update
commands. To make the executions linearizable without resorting to
proposing read-only commands as normal commands, we can employ the same
technique as in \cite{FalerioETAL12GeneralizedLatticeAgreement}: upon
receiving a read command, a server proposes a special no-op command
that commutes with all other commands; the server then responds to the
read command using the first learned history containing the no-op
command. This ensures that all the commands that had completed before
the read command was issued are present in the history.  Moreover,
since no-op commands commute with all other commands, they can be
treated more efficiently than regular read commands by the acceptors,
like, for example, in a fast round of Fast Paxos.

\section{Fast and Conservative Rounds}
\label{sec:rounds}

Most existing SMR algorithms can easily be obtained as the
composition of one or more ACA rounds.

The FLP impossibility result \cite{FischerLynchPaterson83ImpossibilityDistributedConsensusOneFaultyProcess}
implies that no algorithm can implement SMR to the letter in practice.
Instead, SMR algorithms guarantee safety (i.e. replica execute the
same sequence of commands) but may not be able to make progress (i.e.
order new commands). To make progress as likely as possible, most SMR
algorithms are structured in consecutive rounds (also called ballots,
views, or epochs). In each round, the algorithm tries to make progress
but cannot be guaranteed to succeed. When the algorithm detects that
progress is no more possible, a new round is started.  The new round
is initialized so as to preserve safety, usually by one of the
acceptors. For example, in Paxos, the leader of a newly started ballot
numbered $i$ determines a value that is \emph{safe at $i$} and
proposes it as first value to the other acceptors of the ballot $i$.
In existing SMR algorithms, new rounds are initialized ad-hoc, which
prevents mixing rounds of two different algorithms.  Instead,
ACA is an abstraction of rounds which establishes a common interface and contract for switching from
one round to the next, enabling switching from any ACA round to any other.

To simplify the task of expressing existing algorithms as the composition of one
or more ACA rounds, we propose two specifications, \emph{conservative ACA
rounds} and
\emph{fast ACA rounds}, which can easily be concretized to obtain executable rounds.
We show that ACA versions of Classic Paxos, Multi-Coordinated Paxos, Chain
Replication, and Ring Paxos can be obtained by concretizing the specification of
conservative ACA rounds, and that ACA versions of the fast rounds of Fast Paxos can be obtained by concretizing the specification of fast ACA rounds.

Fast and conservative ACA rounds are two specifications, at an
intermediate level of abstraction, which satisfy ACA. Both
specifications specify how the state of the acceptors should be
updated but abstract over the particular communication topology used
to implement the state accesses: the specifications use a global state
which is accessible by all servers in all roles. One can refine fast
or conservative rounds by implementing the state accesses and coordination
using the network, obtaining a concrete algorithm.

Conservative and fast ACA rounds are specified using the state variables $inits$, $proposed$, for every learner $l$, $learned\left[ l \right]$
and $aborts$, as in ACA, and, for every acceptor $a$, two new variables
$status\left[ a \right]$, which can be either ``idle'', ``ready'', or
``stopped'' and $hist\left[ a \right]$, the local history of the
acceptor $a$, which can be either a history or the special value $None$.
Initially, $inits$, $aborts$, and, for every learner
$l$, $learned\left[ l \right]$ are empty sets, and, for every acceptor $a$,
$status\left[ a \right]$  is ``idle'' and $hist\left[ a \right]$ is the special
value $None$.

The next-state relations of conservative and fast rounds are composed
of the following types of transitions (and the default $Crash\left( x
\right)$ transition).
\begin{compactitem}
    \item A $Propose\left( p,c \right)$ transition, for a proposer $p$
        and a command $c$, not guarded, inserts $c$ in the set
        $proposed$. 
    \item An $Init\left( s, h \right)$ transition, for a switcher $s$
        and a history $h$, not guarded, inserts $h$ in the set $inits$.
    \item A $WakeUp\left( a \right)$ transition, for an acceptor $a$,
        is enabled when $status\left[ a \right]$ is ``idle''. Its
        effect is to set $hist\left[ a \right]$ to a history found in
        $inits$ and to set $status\left[ a \right]$ to ``ready''.
    \item An $Accept\left( a \right)$ transition, for an acceptor $a$,
        is enabled when $status\left[ a \right]$ is ``ready'', and
        updates $hist\left[ a \right]$ to $hist\left[ a \right]\star
        c$, where $c$ is a proposed command. In conservative rounds,
        the transition is additionally guarded by the condition that
        the resulting set of local acceptor histories,
        $\aset{hist'\left[ a \right] : a \in Acceptors}$, forms a chain
        in the partial order $\leq$ (this is typically implemented with a
        leader).
    \item A $Learn\left( l, h \right)$ transition, for a learner $l$
      and a history $h$, enabled when $h > learned\left[ l \right]$ and is
      smaller than the GLB of the histories of a learn quorum of acceptors. The effect of
        the transition is to set $learned\left[ l \right]$ to $h$.
    \item A $Stop\left( a \right)$ transition, for an acceptor $a$,
        enabled when $status\left[ a \right]$ is ``ready'', sets
        $status\left[ a \right]$ to ``stopped'' (note that
        $status\left[ a \right] = $ ``stopped'' prevents any further
        ``accept'' transition of acceptor $a$).
%    \item REALLY NEEDED? A $TimeOut\left( a \right)$ transition, for an acceptor $a$,
%        is enabled when $status\left[ a \right]$ is ``idle'' and sets 
%        $status\left[ a \right]$ to ``stopped'' and $hist\left[ a
%        \right]$ to the special value $None$. 
    \item An $Abort\left( s, h \right)$ transition, for a switcher $s$
        and a history $h$, is enabled when there exists a abort quorum
        $R$ such that, for every member $a$ of $R$, $status\left[ a
        \right]$ is ``stopped'' and $h \in SafeAborts\left( R
        \right)$. Its effect is to insert  $h$ in the set $aborts$. 
\end{compactitem}

On top of the additional guard on the $Accept$ transition present only
in conservative rounds, conservative and fast ACA rounds differ in
their definition of learn quorums, abort quorums, and in their
definition of the $SafeAborts\left( R \right)$ operator.
To concisely define read and learn quorums, we define 
\begin{equation}
    Q\left( f \right) = \aset{as \subseteq Acceptors :
        Card\left(as\right) \geq \left\lfloor f\times Card\left(
    Acceptors \right)\right\rfloor + 1}.
\end{equation}

In conservative ACA rounds, abort and learn quorums are defined such
that the intersection between two learn quorums is nonempty and the
intersection between a learn quorum and an abort quorum is nonempty.
The definition leads to two interesting cases: $LearnQ = AbortQ =
Q\left( 1/2 \right)$; $LearnQ = \aset{Acceptors}$ and $AbortQ =
\aset{\aset{a} : a \in Acceptors}$. The operator $SafeAborts\left( R
\right)$ is defined as the singleton set containing the maximum
element over the local histories of the acceptors in $R$ which are not
$None$. Note that the maximum exists because conservative rounds
guarantee that the histories of the acceptors form a chain.

In fast ACA rounds, abort and learn quorums are defined such that
the intersection between any two learn quorums is nonempty and the
intersection between any two learn quorums and an abort quorum is
nonempty.  The definition leads to three interesting cases: $LearnQ =
AbortQ = Q\left( 2/3 \right)$; $LearnQ = Q\left( 1/2 \right)$ and
$AbortQ = Q\left( 3/4 \right)$; $LearnQ = \aset{Acceptors}$ and
$AbortQ = \aset{\aset{a : a \in Acceptors}}$.
The operator $SafeAborts\left( R \right)$ is defined as the set of
maximal elements over the set $\aset{GLB\left( \aset{hist\left[ a
\right] : a \in R \cap L \wedge hist\left[ a \right]\neq None} \right)
: L \in LearnQ}$.

The cases in which abort quorums are singleton sets, in both
conservative and fast rounds, allow tolerating
$f$ faults with $f+1$ servers: a round can abort as long as one
acceptor is correct and a new round, possibly using new correct
servers, can take over. Note that changing round cannot happen
without the scheduling policy instructing the switchers about the
identity of the next round, and thus tolerating $f$ faults with
$f+1$ servers relies on the assumption of a reliable scheduling policy.

Define, for every learner $l$, the history variable $\overline{learned}\left[ l \right]$ which is initialized to the empty set and which is
updated each time a new local history is accepted by the learner $l$ by
inserting the new local history in $\overline{learned}\left[ l \right]$ (see \cite{AbadiLamport91ExistenceRefinementMappings} for the
definition of a history variable).
\begin{thm}
  \label{thm:rounds}
    Both conservative and fast ACA rounds implement ACA under the refinement
    mapping that consists in substituting $\overline{learned\left[ l \right]}$
    for $learned\left[ l \right]$.
\end{thm}

\section{Examples of Concrete ACA Algorithms}
\label{sec:examples}

In this section we show how to modify the rounds of Classic Paxos,
Multi-Coordinated Paxos, Chain Replication, Ring Paxos, and Fast Paxos to obtain
ACA rounds. \Cref{thm:comp} then guarantee that all those different types of
rounds can be used in the same algorithm to yield an implementation of Chain
Agreement.

\subsection{Classic Paxos}

The rounds of the Classic Paxos algorithm can be modified so as to
obtain an implementation of conservative ACA rounds.  We define
Classic Paxos ACA rounds by specializing the specification of
conservative ACA rounds, giving a concrete protocol for disseminating
the proposed commands to the acceptors and enforcing that the learned
histories form a chain.

Each Classic Paxos round has a unique leader among the acceptors which
centrally orders the proposed commands, which are accepted by the
other acceptors in the order determined by the leader.  This protocol
ensures that every local acceptor history is a prefix of the local
history of the leader, therefore guaranteeing that the local acceptor histories
forming a chain.

Acceptor broadcast each newly accepted history to the learners, which
learn a history $h$ when they receive enough extensions of $h$ from the
acceptors.  When instructed by the scheduling policy to abort, a
switcher sends a stop message to the acceptors, which irrevocably stop
updating their local history and send back to the switcher an
acknowledgement containing their local history. A switcher
determines an abort history once it has received acknowledgments from
an abort quorum of acceptors $R$, using the operator $SafeAborts\left(
R \right)$.

The specification of Classic Paxos ACA rounds uses the variables of
the specification of conservative ACA rounds, namely the variables
$proposed$, $inits$, $aborts$, the arrays $learned\left[ l \right]$,
$status\left[ a \right]$, and $hist\left[ a \right]$, and two
additional arrays $accepted\left[ l \right]\left[ a \right]$ and
$stop\left[ s \right]\left[ a \right]$, where $l$ is a learner, $a$ is
an acceptor, and $s$ is a switcher.

There is also a network component that we do not describe explicitly.
However, the network allows any servers to communicate by sending and
receiving messages.

Initially, the variables $proposed$, $inits$, $aborts$, and, for every
learner $l$, $learned\left[ l \right]$ are empty sets. For every
acceptor $a$, $status\left[ a \right]$ is ``idle'' and $hist\left[ a
\right]$ is the special value $None$. For every learner $l$ and acceptor $a$,
$accepted\left[ l \right]\left[ a \right]$ is the special value
$None$. For every switcher $s$ and acceptor $a$, $stopped\left[ s
\right]\left[ a \right]$ is set to $false$.  There are initially no
messages in the network.

The next-state relation is composed of the following types of
transitions, where the definitions of learn quorums, abort quorums,
and of the $SafeAborts\left( R \right)$ operator are the ones of conservative
ACA rounds.
\begin{compactitem}
    \item A $Propose\left( p,c \right)$ transition, for a proposer $p$
        and a command $c$, not guarded, inserts $c$ in the set
        $proposed$ and sends the message $\aseq{\quo{prop},c}$ to
        the leader.
    \item An $Init\left( s, h \right)$ transition, for a switcher $s$
        and a history $h$, not guarded, inserts $h$ in the set $inits$ and
        sends the message $\aseq{\quo{init},h}$ to the leader.
    \item A $WakeUp\left( leader \right)$ transition, enabled when
        $status\left[ leader \right]$ is ``idle'' and the leader can
        receive a $\aseq{\quo{init},h}$ message. Its effect is to
        receive the message, set $hist\left[ leader \right]$ to $h$,
        to set $status\left[ leader \right]$ to ``ready'', and to
        broadcast the message $\aseq{\quo{leader-accept},h}$ to all
        other acceptors.
    \item An $Accept\left( leader \right)$ transition, enabled when
        $status\left[ leader \right]$ is ``ready'' and the leader can
        receive a $\aseq{\quo{prop},c}$ message. The effect of the
        action is to receive the message, to update $hist\left[ leader
        \right]$ to $hist\left[ leader \right]\star c$, and to
        broadcast the message $\aseq{\quo{ack},hist\left[ leader
        \right]\star c}$ to the learners, and to broadcast the message
        $\aseq{\quo{leader-accept},hist\left[ leader \right]\star c}$
        to all the other acceptors.
    \item An $Accept\left( a \right)$ transition, for an acceptor $a$
        which is not the leader, is enabled when $status\left[ a
        \right]$ is ``ready'' and $a$ can receive a
        $\aseq{\quo{leader-accept},h}$ message where $h > hist\left[ a
        \right]$. The effect of the action is to receive the message,
        to update $hist\left[ a \right]$ to $h$, and to broadcast the
        message $\aseq{\quo{ack},hist\left[ a \right]\star c}$ to the
        learners.
    \item A $RcvAck\left( l \right)$ transition, for a learner $l$, is
        enabled when $l$ can receive a message $\aseq{\quo{ack},h}$
        from an acceptor $a$. Its effect is to receive the message and
        to set $accepted\left[ l \right]\left[ a \right]$ to $h$.
    \item A $Learn\left( l, h \right)$ transition, for a learner $l$
        and a history $h$, is enabled when $h > Max\left(
        learned\left[ l \right] \right)$ and there is a learn quorum
        $Q$ of acceptors such that $h = GLB\left( \aset{accepted\left[
        l \right]\left[ a \right] : a \in Q }\right)$. Its effect is
        to insert $h$ in $learned\left[ l \right]$.
    \item A $Stop\left( s \right)$ transition, for a switcher $s$,
        always enabled, broadcasts a $\aseq{\quo{stop}}$ message to all
        the acceptors.  This transition models a switcher being
        instructed by the scheduling policy to abort the current round.
    \item A $Stop\left( a \right)$ transition, for an acceptor $a$,
        enabled when $a$ can receive a $\aseq{\quo{stop}}$ message
        from a switcher $s$. Its effect is to receive the message, set
        $status\left[ a \right]$ to ``stopped'', and to send the
        message $\aseq{\quo{stopped},hist\left[ a \right]}$ to the
        switcher $s$.
    \item A $RcvStopped\left( s \right)$ transition, for a switcher
        $s$, enabled when $s$ can receive a $\aseq{\quo{stopped},h}$
        message from an acceptor $a$. Its effect is to receive the
        message and to set $stopped\left[ s \right]\left[ a \right]$
        to $h$.
    \item An $Abort\left( s, h \right)$ transition, for a switcher $s$
        and a history $h$, is enabled when there exists an abort quorum
        $R$ such that, for every member $a$ of $R$, $stopped\left[ s
        \right]\left[ a \right]$ is not $false$ and $h \in
        SafeAborts\left( R \right)$, where $SafeAborts\left( R
        \right)$ is computed using the array $stopped\left[ a
        \right]$. Its effect is to insert  $h$ in the set $aborts$. 
\end{compactitem}
Compared to the rounds of the original Classic Paxos, Classic Paxos ACA
rounds add an extra level of indirection when changing round: in the
original rounds, the leader of the new round directly queries the
acceptors of the preceeding round, whereas in ACA rounds, the
switchers query the acceptors and then send their abort histories to
the leader. However, changing round is an infrequent event, therefore
this difference should not have much practical impact on performance.
\begin{thm}
  Classic Paxos ACA rounds satisfy Abortable Chain Agreement.
\end{thm}
\begin{proof}
  A Classic Paxos ACA round implements the specification of Abortable
  Chain Agreement under the refinement mapping consisting in
  projecting the state of the round onto the variables $proposed$,
  $inits$, $aborts$, and, for every learner $l$, $learned\left[ l
  \right]$.
\end{proof}

\subsection{Multi-Coordinated Paxos, Chain Replication, and Ring Paxos.}

Multi-Coordinated Paxos, Chain Replication, and Ring Paxos rounds have a similar
structure as Classic Paxos rounds but differ in the way the leader is
implemented and in the way that the histories of the leader are disseminated to
the other acceptors.

Multi-Coordinated Paxos avoids depending on a unique leader by using a
distributed leader implementation that can make progress as long as a
strict majority of the acceptors are correct, trading off resilience
for a higher time complexity. Conservative ACA
rounds can be implemented like with Classic Paxos, replacing the
leader by the distributed leader of Multi-Coordinated Paxos.

In Chain Replication, acceptors are arranged in a chain whose head is
the leader of the round. As in Classic Paxos, abort histories and
proposed commands are sent to the leader. However, instead of the
leader directly sending its local history to all the other acceptors,
the local history of the leader is
forwarded along the chain from one acceptor to the other until it
reaches the last acceptor, which sends the history to the learners. 
By spreading the load more evenly on the all the servers, Chain Replication
can achieve higher throughput that Classic Paxos.
Conservative ACA rounds can be implemented like with Classic Paxos,
replacing the broadcast of the leader by the propagation of the leader's history
along the chain of servers.

Ring Paxos optimizes Chain Replication by having the leader broadcast
commands to the acceptors using IP multicast and only sending hashes
of the commands along the chain. Ring Paxos round can be modified to implement
conservative ACA rounds in a similar way as for Chain Replication.

\subsection{Fast Paxos}

Fast Paxos is composed of two types of rounds: classic rounds, like in Classic
Paxos, and fast rounds.
The fast rounds can be transformed into
implementations of fast ACA rounds. Instead of sending their
commands to a leader, as in Classic Paxos, the proposers of fast
rounds directly broadcast their commands to all the acceptors. As long
as commands are received in the same order by a learn quorum of
acceptors or are received in different orders but commute, commands can be learned in two message-propagation delays instead of three in Classic Paxos.
However, if non-commuting commands are received in different orders by
several acceptors, the local history of the acceptors may not form a
chain anymore and the rounds must abort. The fast rounds of Fast
Paxos implement fast ACA rounds.

Fast Paxos ACA rounds are specified using the same set of variables as for
Classic Paxos, initialized to the same values.
The next-state relation is composed of the following types of
transitions, where the definitions of learn quorums, abort quorums,
and of the $SafeAborts\left( R \right)$ operator are the ones of fast
ACA rounds.
\begin{compactitem}
    \item A $Propose\left( p,c \right)$ transition, for a proposer $p$
        and a command $c$, not guarded, inserts $c$ in the set
        $proposed$ and broadcasts the message $\aseq{\quo{prop},c}$ to
        all the acceptors.
    \item An $Init\left( s, h \right)$ transition, for a switcher $s$
        and a history $h$, not guarded, inserts $h$ in the set $inits$ and
        broadcasts the message $\aseq{\quo{init},h}$ to all the acceptors.
    \item A $WakeUp\left( a \right)$ transition, for an acceptor $a$,
        is enabled when $status\left[ a \right]$ is ``idle'' and $a$
        can receive an $\aseq{\quo{init},h}$ message. Its effect is to
        receive the message, set $hist\left[ a \right]$ to $h$, and to
        set $status\left[ a \right]$ to ``ready''.
    \item An $Accept\left( a \right)$ transition, for an acceptor $a$,
        is enabled when $status\left[ a \right]$ is ``ready'' and $a$
        can receive a $\aseq{\quo{prop},c}$ message. The effect of the
        action is to receive the message, to update $hist\left[ a
        \right]$ to $hist\left[ a \right]\star c$, and to broadcast
        the message $\aseq{\quo{ack},hist\left[ a \right]\star c}$ to
        the learners.
    \item The $RcvAck\left( l \right)$ transitions, for a learner $l$,
        the $Learn\left( l, h \right)$ transitions, for a learner $l$
        and a history $h$, the $Stop\left( s \right)$ transitions, for
        a switcher $s$, $Stop\left( a \right)$ transitions, for an
        acceptor $a$, the $RcvStopped\left( s \right)$ transitions,
        for a switcher $s$, and the $Abort\left( s, h \right)$
        transitions, for a switcher $s$ and a history $h$, which are
        all the same as in Classic Paxos ACA rounds.
\end{compactitem}

As for Classic Paxos ACA rounds, Fast Paxos ACA rounds implement ACA
under the refinement mapping consisting projecting the state of the
round onto the variables $proposed$, $inits$, $aborts$, and, for every
learner $l$, $learned\left[ l \right]$.

\newpage

\printbibliography

\begin{appendix}

  \label{sec:isaproofs}

%  \includepdf[pages={1},pagecommand={\section{Mechanical Proof of the
%  Composition Theorem in Isabelle/HOL}\label{sec:isaproofs}},scale=1]{../IsabellePDFS/podctheories.pdf}
%  \includepdf[pages={2-},pagecommand={},scale=1]{../IsabellePDFS/podctheories.pdf}

\end{appendix}

\end{document}

    We present a new abstraction, Abortable Generalized Consensus, which enables the
    correct-by-construction composition of heterogeneous State-Machine
    Replication algorithms.

    integrating state of the art techniques to optimize executions.  ACA is the most complete approach to date enabling the
    correct-by-construction composition of heterogeneous State-Machine
    Replication algorithms. We provide two specifications satisfying
    ACA and we illustrate their power by refining them to
    obtain a versatile adaptive algorithm that combines five existing
    algorithms into a single one. ACA is completely formalized and its
    underlying composition theorem has been mechanically checked by an interactive
    proof assistant.
}

\vspace{2cm}

\begin{center}
This paper is a regular submission.

\vspace{1cm}
This paper is a student paper.
\end{center}

\newpage

\section{Introduction}

The State-Machine Replication technique (abbreviated SMR) allows
building reliable services on top of unreliable hardware. The basic idea is to 
replicate a service over several servers, each executing the same
sequence of (deterministic) commands. 
Making SMR algorithms efficient has been a long-standing
challenge and a multitude of algorithms have been proposed. Examples
include Fast Paxos \cite{Lamport06FastPaxos}, Disk Paxos
\cite{GafniLamport03DiskPaxos}, Chain Replication
\cite{RenesseSchneider04ChainReplicationSupportingHighThroughputAvailability},
Ring Paxos
\cite{MarandiETAL10RingPaxosHighthroughputAtomicBroadcastProtocol},
Egalitarian Paxos
\cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments},
Multi-Coordinated Paxos
\cite{CamargosSchmidtPedone07MulticoordinatedPaxos}, Cheap Paxos
\cite{LamportMassa04CheapPaxos}, Vertical Paxos
\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication},
Paxos-MIC
\cite{HurfinMoiseNarzul11AdaptiveFastPaxosMakingQuickEverlasting},
Mencius
\cite{MaoJunqueiraMarzullo08MenciusBuildingEfficientReplicatedStateMachine},
and Fast Mencius \cite{WeiETAL13FastMenciusMenciusLowCommitLatency}.

Each of the existing algorithms is best suited to a particular range
of environments, characterized, e.g., by client load, network
topology, likelihood of faults, communication latency and bandwidth, processing
power of nodes, availability of disk storage and memory.
However, the environment faced by real systems evolves over time:
hardware is upgraded, new data-centers are built, the popularity of a
service changes, etc.  Consistent performance therefore comes with
adaptation to change.  Unfortunately, existing SMR algorithms cannot easily be
stopped and replaced by another algorithm on the fly. Moreover,
devising a new algorithm that can change behavior dynamically is hard:
dynamically switching from one SMR algorithm to another requires
synchronizing notoriously intricate algorithms.

We propose a new abstraction, \emph{Abortable Chain Agreement}
(abbreviated ACA), which subsumes SMR and whose implementations,
called \emph{ACA rounds}, can, by construction, pass the baton from
one to another on the fly and efficiently. Moreover, optimizing a replicated
service consisting of several ACA rounds becomes just a matter of devising a new
ACA round, which can then be plugged-in on the fly: any new ACA round
is compatible with every existing ACA round by construction.

Abortable Chain Agreement is the \emph{most complete framework to date} enabling
correct by construction composition of heterogeneous rounds.  
ACA is completely formalized and its underlying composition theorem has been
mechanically checked by an interactive proof assistant.  
Compared to ACA,
existing approaches
\cite{BarNoyETAL87ShiftingGearsChangingAlgorithmsFlyTo,GuerraouiETAL10Next700BftProtocols,GuerraouiKuncakLosa12SpeculativeLinearizability}
are limited. In particular, they do not allow the disorderly execution of
commuting and read-only commands, an important practical optimization
which is at the root of recent work on SMR like Generalized Consensus
\cite{Lamport05GeneralizeConsensus} and Generalized Lattice Agreement
\cite{FalerioETAL12GeneralizedLatticeAgreement}.

\begin{comment}
On top of the traditional roles of proposer, acceptor, and learners
\cite{lamport2001paxos} used in consensus algorithms, ACA introduces
the notion of \emph{switchers} and of \emph{scheduling policy}. The switchers
have the task of aborting a running ACA round and initializing a new
ACA round in a safe and live manner.  The \emph{scheduling policy}
decides when to abort an ACA round and choses the next ACA round to
execute. It must ensure that all switchers agree on the next ACA round
to initialize.
\end{comment}

We observe that, because of the FLP impossibility result
\cite{FischerLynchPaterson83ImpossibilityDistributedConsensusOneFaultyProcess},
most existing SMR algorithms are structured in a sequence of abortable
rounds which can easily be transformed into ACA rounds. To simplify
this task even further, we propose two specifications, \emph{conservative
ACA rounds} and \emph{fast ACA rounds}, which, as we show, can be concretized
to obtain ACA versions of the rounds of five existing algorithms: 
Classic Paxos, Fast Paxos, Chain Replication, Ring Paxos,
and Multi-Coordinated Paxos. ACA guarantees that the composition of the five ACA
rounds is correct by construction, forming a versatile adaptive replication
algorithm.

The rest of the paper is structured as follows.
\Cref{sec:related} surveys recent related work. \Cref{sec:ca} presents Chain
Agreement, which embodies the requirements for implementing SMR. \Cref{sec:aca}
presents Abortable Chain Agreement and its optimizations to treat commuting
and read-only commands efficiently. \Cref{sec:rounds} presents our specification
of conservative and fast ACA rounds, which implement ACA. In
\cref{sec:examples}, we show that the rounds of  
Classic Paxos, Fast Paxos, Chain Replication, Ring Paxos,
and Multi-Coordinated Paxos can easily be transformed into ACA implementations
by using our specifications of conservative and fast ACA rounds.
Finally, in \cref{sec:isaproofs}, we provide a complete formalization of ACA
and a mechanical proof of the \emph{composition theorem} (\cref{thm:comp}),
checked by the interactive proof assistant Isabelle/HOL \cite{NipkowPaulsonWenzel02IsabelleHOL}.

\begin{comment}
Abortable Chain Agreement builds on several recent advances in SMR
algorithms: ACA allows acceptors to only eventually reach agreement,
like in Generalized Paxos \cite{Lamport05GeneralizeConsensus},
boosting performance when most commands commute; ACA allows fast
execution of read commands like in Generalized Lattice Agreement
\cite{FalerioETAL12GeneralizedLatticeAgreement}; ACA implementations
can tolerate $f$ faults with only $f+1$ replicas and replace crashed replicas
upon changing round, as in Vertical Paxos  \cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}.  
\end{comment}

\section{Related Work}
\label{sec:related}

SMR is often implemented using a sequence of instances of Consensus.
There are numerous examples of consensus algorithms in the literature.
However, Abortable Chain Agreement is based on Generalized Consensus. 
In Generalized Consensus \cite{Lamport05GeneralizeConsensus}, the learners of a
SMR algorithm repeatedly learn new values belonging to a \emph{C-struct set}. A
C-struct can be seen as a data structure encoding  sequences of commands modulo
the ordering of commands that commute. The advantage of Generalized
Consensus over Consensus is that
two servers which receive commuting commands in different orders can still reach
agreement on C-structs without further synchronization. 
Our Abortable Chain Agreement can also use C-structs, with the same advantages
as Generalized Consensus. For simplicity, the presentation in this
paper uses the more concrete notion of \emph{command history} (see
\cref{sec:histories}).

Generalized Lattice Agreement \cite{FalerioETAL12GeneralizedLatticeAgreement} allows more efficient implementations of SMR when
commands are partitioned into read-only commands and write-only commands.
Abortable Chain Agreement supports the same optimizations as Generalized Lattice Agreement.
Notably, read-only commands are treated specially so as to optimize their
executions (see \cref{sec:readonly}).

Vertical Paxos algorithms 
\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication} are SMR algorithms
whose set of replica servers can be reconfigured without stopping to process new
commands. Crashed servers can therefore be replaced by new servers, ensuring the
long-term availability of a replicated service.
Vertical Paxos algorithms use an external reconfiguration master, itself
implemented as a replicated state machine, which determines the servers that should
be removed from or added to the configuration.
ACA shares some aspects of Vertical Paxos algorithms: the task of the
reconfiguration master is similar to the task of the scheduling policy
component of ACA (see \cref{sec:aca}).
However, in ACA, changing round can be used not only to replace crashed servers
but also to change the algorithm that the servers are running. 
Changing algorithm is not possible in Vertical Paxos.
In both Vertical Paxos and ACA, a reliable reconfiguration master, resp. a
reliable scheduling policy, allows tolerating $f$ crashes with $f+1$ servers.

The Abstract Framework \cite{GuerraouiETAL10Next700BftProtocols} proposes to
build Byzantine Fault-Tolerant algorithms as the succession of abortable rounds
called Abstract instances. Abstract instances must satisfy the Abstract
correctness properties. By construction, the composition of any number of
Abstract instances is a correct SMR implementation. However, the Abstract
Framework uses totally ordered sequence of commands, making difficult the
optimization of the execution of commuting and read-only commands that are
possible with ACA.  Moreover, the Abstract Framework does not identify the
crucial role of the scheduling policy for the long-term resilience of a service.
Abortable Chain Agreement combines the ideas of the Abstract Framework,
Generalized Paxos, Generalized Lattice Agreement, and Vertical Paxos in a
unified abstraction.

Finally, the Speculative Linearizability framework
\cite{GuerraouiKuncakLosa12SpeculativeLinearizability} was a first
attempt by the authors of the present paper at generalizing the
Abstract Framework and mechanically proving its properties with the
Isabelle/HOL interactive
proofs assistant.  Like the Abstract Framework, the
Speculative Linearizability framework is based on totally ordered
sequences of commands, with the same drawback for the execution of
commuting and read-only commands. Moreover, Abortable Chain Agreement is both
conceptually simpler and subsumes the Speculative Linearizability
property.  Like Speculative Linearizability, Abortable Chain Agreement
and its main property, the composition theorem (\cref{thm:comp}), are
formalized in Isabelle/HOL (see \cref{sec:isaproofs}).

Abortable Chain Agreement is the most complete framework to date for building
adaptive replication algorithms out of heterogeneous rounds, making state of the
art optimization techniques available to algorithm designers, which was not
possible before.

\section{Chain Agreement}
\label{sec:ca}

We consider a set of asynchronous servers communicating by message passing in an
asynchronous network. Servers are not necessarily sequential processes. We
describe a behavior of the system using an infinite sequence of states, where a
state is a function from variable names to values (i.e. a valuation
of variables).  We specify sets of allowed behaviors using an initial predicate,
a next-state relation (using unprimed and primed variables) and a fairness
condition, all in the style of TLA+ \cite{Lamport02SpecifyingSystems}. However,
for succinctness, our specification are rather informal, eluding non-essential
details.

Our specifications all contain by default a state variable $crashed$ and
$Crash\left( x \right)$ transition which is always enabled. A server $x$ faithfully
executes the algorithm that is assigned to it until its identifier is added to
the variable $crashed$ by a $Crash\left( x \right)$ transitions.  When the identifier of a
server is in the set $crashed$, all its actions are disabled.  In an execution,
a server that is never added to the set $crashed$ is said correct.

We consider a computing service exposing a set of commands $C$. The
service has a sequential specification consisting of an initial state
and a deterministic transition relation in which every command
atomically updates the state of the service and produces an output. We
assume that duplicate commands have no effect on the state of the
service and return the same output as their first occurrence.

In a traditional SMR algorithm, the servers can be seen has playing one or
several of three roles: \emph{proposers}, which propose new commands to execute,
\emph{acceptors}, which accept and order commands in a sequence, and
\emph{learners}, which learn about the growing sequence of commands to
execute \cite{lamport2001paxos}. Often, but not necessarily, every replica server plays the
three roles of proposer, acceptor, and learner: clients of the service
send their commands to some replica servers; replica servers submit
the received commands to the SMR protocol in the role of proposers;
replica servers respond to commands by executing the commands learned
in the role of learners; replica servers participate in command
ordering in the role of acceptors.  An SMR algorithm guarantees that
the replicas learn the same sequences of commands, possibly up to the
reordering of commuting commands. Therefore, every learner can respond to client
requests using its last learned sequence of commands, and clients will receive
the same responses as the ones produces by a centralized, non-replicated,
service.

Chain Agreement precisely specifies the allowed behaviors of the roles of \emph{proposers}
and \emph{learners}, abstracting over the behavior of acceptors.

\begin{comment}
We suppose that each server can play a number of \emph{roles}.
Chain Agreement specifies the allowed behaviors the roles of \emph{proposers}
and \emph{learners}. Proposers receive commands from the clients of the service
and propose them for execution by the replicated service. Each learner learns about
a growing sequence of commands. Thanks to the sequential specification of the
service, a sequence of commands determines an output for each the included commands. 
In a traditional SMR algorithm, servers can also play the role of
\emph{acceptors}, whose task is to collectively order commands and let the
learners know about the latest sequence of ordered commands.

The point of Chain Agreement is to implement a replicated service by having every learner learn about a growing sequence of commands, which
represent an execution of the sequential specification of the service.
Crucially, at all times, the sequences of commands learned by the
different learners must form a chain in the prefix order on sequences.
Moreover, for every command $c$ proposed by a correct proposer, every
correct learner must eventually learn a sequence containing $c$.  Once
a learner learns about a sequence containing a command $c$, it can
determine the output produced by $c$, using the sequential
specification of the service, and respond to the client that issued
the command $c$.
\end{comment}

To specify Chain Agreement, we need the notion of ordering of sequences and of
maximum sequence.  The prefix order on sequences is denoted by the operator
$\leq$ ($s_1 \leq s_2$ means that $s_1$ is a prefix of $s_2$). 
%Using the operator $\leq$, we define $\geq$, $<$, and $>$ as expected. 
Note that $\leq$ is a partial order on sequences. The maximum over a set of
sequences $S$ is denoted by $Max\left( S \right)$ and $Max\left( \aset{}
\right)$ is the empty sequence.

We specify Chain Agreement using the state variables $crashed$, $proposed$, containing the proposed commands,
and, for every learner $l$,  $learned\left[ l \right]$, a set of
sequences of commands. Initially, $crashed$, $proposed$ and, for every learner $l$,
$learned\left[ l \right]$ are empty sets. 
Remember that any transition involving a server $x$ is always disabled when $x\in crashed$.

The next-state relation consists of three types of transitions:
$Crash\left( x \right)$ transitions; $Propose\left( l, c
\right)$ transitions, for a learner $l$ and a command $c$, in which
$c$ is inserted in the set $proposed$; $Learn\left( l,cs \right)$
transitions, for a learner $l$ and a sequence of commands $cs$, in
which $cs$ is inserted in $learned\left[ l \right]$. 
A $Learn\left( l, cs \right)$ transition is guarded by the following
conditions:
\begin{compactitem}
    \item[\textbf{Agreement}:] The elements of the set $\aset{learned'\left[ l
      \right] : l \in Learners}$, the learned sequences after the transition,
      form a chain in the partial order $\leq$.
    \item[\textbf{Validity}:] Every learned sequence contains only
        proposed commands.  
    \item[\textbf{Irrevocability}:] The new learned sequence
        $cs$ is a strict extension of $Max\left(learned\left[ l
        \right]\right)$. 
        %\\ $cs > Max\left(learned\left[ l \right]\right)$.
\end{compactitem}
Finally, we add, for every learner $l$, one weak fairness condition
\cite{Lamport02SpecifyingSystems} on the transitions $Learn\left( l,
cs \right)$. The weak fairness conditions ensure that every correct
learner eventually learns about every proposed command.

Thanks to the properties of Chain Agreement, the clients, observing
their input commands and the received responses, see a
\emph{linearizable} execution
\cite{HerlihyWing90LinearizabilityCorrectnessConditionConcurrentObjects}.
Moreover, the service remains available as long as at least one
proposer and one learner are correct.

\section{Abortable Chain Agreement}
\label{sec:aca}

We now consider implementing Chain Agreement as a sequence of \emph{Abortable Chain
Agreement rounds}, where each round makes progress ordering requests or aborts
its execution and passes the baton to the next round in the sequence.
An algorithm composed of a sequence of ACA rounds is called an \emph{ACA
algorithm}.

On top of the traditional roles of proposer, acceptor, and learner, ACA
introduces the new role of \emph{switcher} and a new component, the
\emph{scheduling policy}. In ACA, the assignment of the roles of proposer,
acceptor, learner, and switcher can change in each rounds, allowing for
reconfiguration. 
In normal circumstances, the proposers, acceptors,
and learners of an ACA round work together to implement Chain Agreement.
However, if the environment becomes unfavorable to a running ACA round, the
scheduling policy can instruct the round to abort its execution and pass the
baton, using the switchers, to a new round.

A switcher of an aborting round determines, from the state of the
acceptors, an \emph{abort sequence} that it transfers to one or several
switchers of the new round.  Upon receiving an abort sequence, a switcher of the
new round initializes the acceptors of the new round, which can then
proceed implementing SMR, until the new round itself aborts and the
process is repeated. 
We assume that in the first round, which has no previous round, every 
switchers receives the empty abort sequence.

The scheduling policy determines when to abort an ACA round and which new ACA
round should take over. The scheduling policy must inform the switchers about the
identity of the next round and must guarantee that all switchers receive the
same information.  In this paper we abstract over the scheduling policy, assuming
that a round that aborts passes the baton a nondeterministically chosen new
round. A reliable scheduling-policy component can realistically be implemented
as a replicated state machine running on a large number of servers otherwise
dedicated to other tasks
\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}. We will see
 in \cref{sec:rounds} that, as in Vertical Paxos, a reliable scheduling policy component allows
tolerating $f$ faults with $f+1$ replicas, while maintaining the long-term
resilience of the service by replacing faulty servers upon round change.

The clients of an ACA algorithm can send their commands to the proposers of any round (in
practice, clients send their commands to proposers of the largest
active round) and servers can use sequences learned in any round
to determine outputs.

ACA specifies the allowed behaviors of the proposers, learners, and switchers
\emph{of a single round}. Similarly to
Chain Agreement, we use the state variables $crashed$, $proposed$,
and, for every learner $l$, $learned\left[ l \right]$.  Additionally,
we use the state variables $inits$, containing the abort sequences
received from the previous round, and $aborts$, containing the abort
sequences transfered to the next round.
Two consecutive ACA rounds are composed by letting the $aborts$ variable of the
first round be the $inits$ variable of the second round.

Initially, $crashed$, $proposed$, for every
learner $l$, $learned\left[ l \right]$, $inits$, and $aborts$ are all empty sets.
The next-state relation consists of five types of transitions. The
first three, $Crash\left( x \right)$, $Propose\left( l, c
\right)$, and $Learn\left( l, cs \right)$, for a learner, proposer, or
switcher $x$, for a learner $l$, and for a sequence of commands
$cs$, have the same effect as in Chain Agreement.
ACA introduces two new transitions, $Init\left(
s, cs \right)$ and $Abort\left( s, cs \right)$, for a switcher $s$ and a sequence of
commands $cs$. The transition $Init\left( s, cs \right)$ adds $cs$ to the set $inits$, and models a switcher receiving the
abort sequence $cs$ from the preceding round.
The transition $Abort\left( s, cs \right)$ adds $cs$ to the set $aborts$ and models a switcher transferring the abort sequence $cs$ to a switcher of the next round.
All the transition are guarded by the following conditions:
\begin{compactitem}
    \item[\textbf{Agreement}:] The elements of the set $\aset{learned'\left[ l
      \right] : l \in Learners}$ form a chain in the partial order.
    \item[\textbf{Validity}:] Every abort sequence or learned sequence
        is of the form $h \star cs$, where $cs$ contains only commands
        of the set $proposed$ and $h$ is the greatest lower bound of a
        nonempty subset of $inits$.
    \item[\textbf{Initialization}:] If $inits$ is empty, then no
        sequence can be learned and no switcher can abort.
    \item[\textbf{Irrevocability}:] In the case of a $Learn\left( l,cs
        \right)$ transition, the new learned sequence $cs$ is a strict
        extension of $Max\left(learned\left[ l \right]\right)$.
        %, $cs
        %\geq Max\left(learned\left[ l \right]\right)$.
    \item[\textbf{Safe Abort}:] Every abort sequence is larger in the
        partial order $\leq$ than every learned sequence.
\end{compactitem}
Note that Agreement and Irrevocability are the same as in Chain
Agreement.

Finally, for every learner $l$, we add a weak fairness condition on the
following state transition: 
\begin{equation}
    \exists  s \in Switchers, cs \in Sequence\left( C \right) : Learn\left( l, cs \right) \vee Abort\left( s,cs \right)
\end{equation}
The weak fairness conditions ensure that when a correct learner has not
learned a sequence containing all commands proposed in the round, then
it eventually learns a new command or, if there is at least one correct switcher, at least one switch value is transfered to the next round. 
Therefore, if a client has submitted a command to a proposer, eventually,
either the client gets a response from the current round, or
the next round is initialized with an abort sequence and the
command can be resubmitted to the next round.

\subsection{The Composition Theorem}

Consider the succession of $n$ rounds numbered $1$ to $n$.  To
distinguish the variables of each round, we annotate the local
variables of each round with the number of the round. For example, the
variables of the seventh round are written $proposed_7$, for every
learner $l$, $learned_7\left[ l \right]$, $inits_7$, and $aborts_7$.

Moreover, we assume that a command, which is proposed to a round and that
is not learned before the first abort sequence is transfered to the
next round, is proposed anew to the next round.

Define the state variable $proposed$ as the union over all rounds $i$
of the $proposed_i$ variables and, for every learner $l$, 
define the $learned\left[ l \right]$ variable as the union over all
rounds $i$ of the $learned_i\left[ l \right]$ variables.

\begin{thm}[Composition Theorem]
    \label{thm:comp}
    In every execution of a succession of ACA rounds, the sequence of valuations
    of the variable $proposed$ and, for every learner $l$, of the variables
    $learned\left[ l \right]$ satisfies Chain Agreement.
\end{thm}
\Cref{thm:comp} can be shown by induction on the number of rounds,
using the following lemma.
\begin{lem}
  \label{lem:comp}
    Consider two rounds numbered $i$ and $i+1$. In every execution of a
    succession of ACA rounds, the sequence of valuations of the variables $inits
    = inits_i$, $propose = proposed_i \cup propose_{i+1}$, for every learner
    $l$, $learned\left[ l \right] = learned_i\left[ l \right] \cup
    learned_{i+1}\left[ l \right]$and $aborts = aborts_{i+1}$ satisfies ACA.
\end{lem}
\begin{proof}
  A mechanically-checked proof in Isabelle/HOL appears in \cref{sec:isaproofs}.
  The proofs uses a version of the specification of ACA that has a more
  executable form that presented in this section. Moreover,
  instead of sequences of commands, the Isabelle/HOL specification uses the more
  general notion of command history, presented in the next subsection.
  The Isabelle/HOL formalization of ACA uses the theory of I/O automata, which
  is also formalized in the Isabelle/HOL. The proof uses a refinement
  mapping from the composition of two consecutive ACA instances to a single ACA
  instance.
\end{proof}

\Cref{thm:comp} guarantees that the successive composition of any
number of ACA rounds implements Chain Agreement.
Therefore, with ACA, one can easily build replicated services that
dynamically adapt to the current operating conditions of the system: a
replicated service has several types of rounds at its disposal, each being
tailored to particular conditions. Each time the conditions change,
the current ACA round aborts and passes the baton to a more 
appropriate type of round, as determined by the scheduling policy. Moreover, a new type
of ACA round can always be devised after the system is deployed and added on the fly to
the existing types of rounds.

We now describe two important practical optimizations of ACA.

\subsection{Histories}
\label{sec:histories}

ACA can be implemented more efficiently by substituting \emph{command histories} for sequences of commands, as in Generalized Consensus \cite{Lamport05GeneralizeConsensus}.

We say that two commands $c_1$ and $c_2$ \emph{commute} when, in every sequential
execution of the service, reversing the order of $c_1$ and $c_2$ does
not change any of the outputs, including those given to $c_1$ and
$c_2$.

A command history, abbreviated history, is a data structure that
represents a set of sequences which are the same modulo the reordering
of commuting commands, and in which duplicate commands do not appear.
All the sequences represented by a given history contain the same
commands and determine the same outputs for each command.  Therefore,
to implement a replicated service, it is sufficient that learners
learn command histories instead of sequences.  For a precise
definition of command histories, we refer the reader to
\cite{Lamport05GeneralizeConsensus}. Our Isabelle/HOL formalization
(\cref{sec:isaproofs}) abstracts
over the concrete representation of histories, assuming only their necessary
properties. Command histories were first
introduced in trace theory \cite{Mazurkiewicz84Semantics}.

A command history can be built by appending a sequence of commands
$cs$ to a history $h$, obtain the history $h \star cs$.  The major
advantage that histories have over sequences is that if a sequence
$cs_1$ is obtained by reordering the commuting commands of $cs_2$,
then $h \star cs_1 = h \star cs_2$. Therefore, in  implementations of
ACA, servers can agree on a history even though
they chose different orders for some commuting commands.

Command histories share the four properties of sequences that are
necessary and sufficient to substitute histories for sequences in the
definition of Chain Agreement and ACA.
\begin{compactitem}
    \item[\textbf{Distinctness}:] No duplicate command appears in a history. 
    \item[\textbf{Ordering}:] Command histories form a partial order:
        $h_1 \leq h_2$ when there exists a sequence $s$ such that $h_2
        = h_1 \star s$.
    \item[\textbf{GLB}:] Every finite set of command histories has a
        \emph{greatest lower bound}.
%    \item[\textbf{LUB}:] Every finite set of compatible command
%        histories has a \emph{least upper bound}, where a set of
%        command histories is compatible when they have a common upper
%        bound.
    \item[\textbf{Consistency}:] If $h_1 \leq h_2 \leq h_3$ and $h_3 =
        h_1 \star s$ then there exists $s'$ and $s''$ where every
        request of $s'$ and $s''$ appears in $s$ and such that $h_2 =
        h_1 \star s'$ and $h_3 = h_2 \star s''$.
\end{compactitem}
The optimized version of ACA is obtained by
simply substituting histories for sequences in the definition of ACA.
The Ordering, GLB, and Consistency properties of histories ensure that
\cref{thm:comp} still holds: the Isabelle/HOL proof of \cref{thm:comp} (see
\cref{sec:isaproofs}) relies only on the four properties above.

\subsection{Optimizing the Execution of Read-Only Commands}
\label{sec:readonly}

Chain Agreement can be further optimized by treating 
read-only commands specially, as in Generalized Lattice Agreement
\cite{FalerioETAL12GeneralizedLatticeAgreement}.

We consider the case in which a set of servers each play the four
roles of proposer, acceptor, learner, and switcher.
Clients sends their commands to any server, which proposes them as a
proposer, and records the command as pending.
%When a server learns a history that contains one of its pending
%commands, it uses the history to determine the response to the pending
%command and forwards the response to the appropriate client.

As observed in \cite{FalerioETAL12GeneralizedLatticeAgreement}, read-only
commands can be executed faster by directly reading from the last learned
history, instead of proposing the read commands and waiting for it to appear in
a future learned history.  Because learned histories form a chain, all the reads
return consistent values and this approach yields a \emph{serializable}
implementation of the service.

However linearizability is not satisfied if reads are served directly
from the last learned history: a lagging replica server could respond
to a read using a history which does not contain all completed update
commands. To make the executions linearizable without resorting to
proposing read-only commands as normal commands, we can employ the same
technique as in \cite{FalerioETAL12GeneralizedLatticeAgreement}: upon
receiving a read command, a server proposes a special no-op command
that commutes with all other commands; the server then responds to the
read command using the first learned history containing the no-op
command. This ensures that all the commands that had completed before
the read command was issued are present in the history.  Moreover,
since no-op commands commute with all other commands, they can be
treated more efficiently than regular read commands by the acceptors,
like, for example, in a fast round of Fast Paxos.

\section{Fast and Conservative Rounds}
\label{sec:rounds}

Most existing SMR algorithms can easily be obtained as the
composition of one or more ACA rounds.

The FLP impossibility result \cite{FischerLynchPaterson83ImpossibilityDistributedConsensusOneFaultyProcess}
implies that no algorithm can implement SMR to the letter in practice.
Instead, SMR algorithms guarantee safety (i.e. replica execute the
same sequence of commands) but may not be able to make progress (i.e.
order new commands). To make progress as likely as possible, most SMR
algorithms are structured in consecutive rounds (also called ballots,
views, or epochs). In each round, the algorithm tries to make progress
but cannot be guaranteed to succeed. When the algorithm detects that
progress is no more possible, a new round is started.  The new round
is initialized so as to preserve safety, usually by one of the
acceptors. For example, in Paxos, the leader of a newly started ballot
numbered $i$ determines a value that is \emph{safe at $i$} and
proposes it as first value to the other acceptors of the ballot $i$.
In existing SMR algorithms, new rounds are initialized ad-hoc, which
prevents mixing rounds of two different algorithms.  Instead,
ACA is an abstraction of rounds which establishes a common interface and contract for switching from
one round to the next, enabling switching from any ACA round to any other.

To simplify the task of expressing existing algorithms as the composition of one
or more ACA rounds, we propose two specifications, \emph{conservative ACA
rounds} and
\emph{fast ACA rounds}, which can easily be concretized to obtain executable rounds.
We show that ACA versions of Classic Paxos, Multi-Coordinated Paxos, Chain
Replication, and Ring Paxos can be obtained by concretizing the specification of
conservative ACA rounds, and that ACA versions of the fast rounds of Fast Paxos can be obtained by concretizing the specification of fast ACA rounds.

Fast and conservative ACA rounds are two specifications, at an
intermediate level of abstraction, which satisfy ACA. Both
specifications specify how the state of the acceptors should be
updated but abstract over the particular communication topology used
to implement the state accesses: the specifications use a global state
which is accessible by all servers in all roles. One can refine fast
or conservative rounds by implementing the state accesses and coordination
using the network, obtaining a concrete algorithm.

Conservative and fast ACA rounds are specified using the state variables $inits$, $proposed$, for every learner $l$, $learned\left[ l \right]$
and $aborts$, as in ACA, and, for every acceptor $a$, two new variables
$status\left[ a \right]$, which can be either ``idle'', ``ready'', or
``stopped'' and $hist\left[ a \right]$, the local history of the
acceptor $a$, which can be either a history or the special value $None$.
Initially, $inits$, $aborts$, and, for every learner
$l$, $learned\left[ l \right]$ are empty sets, and, for every acceptor $a$,
$status\left[ a \right]$  is ``idle'' and $hist\left[ a \right]$ is the special
value $None$.

The next-state relations of conservative and fast rounds are composed
of the following types of transitions (and the default $Crash\left( x
\right)$ transition).
\begin{compactitem}
    \item A $Propose\left( p,c \right)$ transition, for a proposer $p$
        and a command $c$, not guarded, inserts $c$ in the set
        $proposed$. 
    \item An $Init\left( s, h \right)$ transition, for a switcher $s$
        and a history $h$, not guarded, inserts $h$ in the set $inits$.
    \item A $WakeUp\left( a \right)$ transition, for an acceptor $a$,
        is enabled when $status\left[ a \right]$ is ``idle''. Its
        effect is to set $hist\left[ a \right]$ to a history found in
        $inits$ and to set $status\left[ a \right]$ to ``ready''.
    \item An $Accept\left( a \right)$ transition, for an acceptor $a$,
        is enabled when $status\left[ a \right]$ is ``ready'', and
        updates $hist\left[ a \right]$ to $hist\left[ a \right]\star
        c$, where $c$ is a proposed command. In conservative rounds,
        the transition is additionally guarded by the condition that
        the resulting set of local acceptor histories,
        $\aset{hist'\left[ a \right] : a \in Acceptors}$, forms a chain
        in the partial order $\leq$ (this is typically implemented with a
        leader).
    \item A $Learn\left( l, h \right)$ transition, for a learner $l$
      and a history $h$, enabled when $h > learned\left[ l \right]$ and is
      smaller than the GLB of the histories of a learn quorum of acceptors. The effect of
        the transition is to set $learned\left[ l \right]$ to $h$.
    \item A $Stop\left( a \right)$ transition, for an acceptor $a$,
        enabled when $status\left[ a \right]$ is ``ready'', sets
        $status\left[ a \right]$ to ``stopped'' (note that
        $status\left[ a \right] = $ ``stopped'' prevents any further
        ``accept'' transition of acceptor $a$).
%    \item REALLY NEEDED? A $TimeOut\left( a \right)$ transition, for an acceptor $a$,
%        is enabled when $status\left[ a \right]$ is ``idle'' and sets 
%        $status\left[ a \right]$ to ``stopped'' and $hist\left[ a
%        \right]$ to the special value $None$. 
    \item An $Abort\left( s, h \right)$ transition, for a switcher $s$
        and a history $h$, is enabled when there exists a abort quorum
        $R$ such that, for every member $a$ of $R$, $status\left[ a
        \right]$ is ``stopped'' and $h \in SafeAborts\left( R
        \right)$. Its effect is to insert  $h$ in the set $aborts$. 
\end{compactitem}

On top of the additional guard on the $Accept$ transition present only
in conservative rounds, conservative and fast ACA rounds differ in
their definition of learn quorums, abort quorums, and in their
definition of the $SafeAborts\left( R \right)$ operator.
To concisely define read and learn quorums, we define 
\begin{equation}
    Q\left( f \right) = \aset{as \subseteq Acceptors :
        Card\left(as\right) \geq \left\lfloor f\times Card\left(
    Acceptors \right)\right\rfloor + 1}.
\end{equation}

In conservative ACA rounds, abort and learn quorums are defined such
that the intersection between two learn quorums is nonempty and the
intersection between a learn quorum and an abort quorum is nonempty.
The definition leads to two interesting cases: $LearnQ = AbortQ =
Q\left( 1/2 \right)$; $LearnQ = \aset{Acceptors}$ and $AbortQ =
\aset{\aset{a} : a \in Acceptors}$. The operator $SafeAborts\left( R
\right)$ is defined as the singleton set containing the maximum
element over the local histories of the acceptors in $R$ which are not
$None$. Note that the maximum exists because conservative rounds
guarantee that the histories of the acceptors form a chain.

In fast ACA rounds, abort and learn quorums are defined such that
the intersection between any two learn quorums is nonempty and the
intersection between any two learn quorums and an abort quorum is
nonempty.  The definition leads to three interesting cases: $LearnQ =
AbortQ = Q\left( 2/3 \right)$; $LearnQ = Q\left( 1/2 \right)$ and
$AbortQ = Q\left( 3/4 \right)$; $LearnQ = \aset{Acceptors}$ and
$AbortQ = \aset{\aset{a : a \in Acceptors}}$.
The operator $SafeAborts\left( R \right)$ is defined as the set of
maximal elements over the set $\aset{GLB\left( \aset{hist\left[ a
\right] : a \in R \cap L \wedge hist\left[ a \right]\neq None} \right)
: L \in LearnQ}$.

The cases in which abort quorums are singleton sets, in both
conservative and fast rounds, allow tolerating
$f$ faults with $f+1$ servers: a round can abort as long as one
acceptor is correct and a new round, possibly using new correct
servers, can take over. Note that changing round cannot happen
without the scheduling policy instructing the switchers about the
identity of the next round, and thus tolerating $f$ faults with
$f+1$ servers relies on the assumption of a reliable scheduling policy.

Define, for every learner $l$, the history variable $\overline{learned}\left[ l \right]$ which is initialized to the empty set and which is
updated each time a new local history is accepted by the learner $l$ by
inserting the new local history in $\overline{learned}\left[ l \right]$ (see \cite{AbadiLamport91ExistenceRefinementMappings} for the
definition of a history variable).
\begin{thm}
  \label{thm:rounds}
    Both conservative and fast ACA rounds implement ACA under the refinement
    mapping that consists in substituting $\overline{learned\left[ l \right]}$
    for $learned\left[ l \right]$.
\end{thm}

\section{Examples of Concrete ACA Algorithms}
\label{sec:examples}

In this section we show how to modify the rounds of Classic Paxos,
Multi-Coordinated Paxos, Chain Replication, Ring Paxos, and Fast Paxos to obtain
ACA rounds. \Cref{thm:comp} then guarantee that all those different types of
rounds can be used in the same algorithm to yield an implementation of Chain
Agreement.

\subsection{Classic Paxos}

The rounds of the Classic Paxos algorithm can be modified so as to
obtain an implementation of conservative ACA rounds.  We define
Classic Paxos ACA rounds by specializing the specification of
conservative ACA rounds, giving a concrete protocol for disseminating
the proposed commands to the acceptors and enforcing that the learned
histories form a chain.

Each Classic Paxos round has a unique leader among the acceptors which
centrally orders the proposed commands, which are accepted by the
other acceptors in the order determined by the leader.  This protocol
ensures that every local acceptor history is a prefix of the local
history of the leader, therefore guaranteeing that the local acceptor histories
forming a chain.

Acceptor broadcast each newly accepted history to the learners, which
learn a history $h$ when they receive enough extensions of $h$ from the
acceptors.  When instructed by the scheduling policy to abort, a
switcher sends a stop message to the acceptors, which irrevocably stop
updating their local history and send back to the switcher an
acknowledgement containing their local history. A switcher
determines an abort history once it has received acknowledgments from
an abort quorum of acceptors $R$, using the operator $SafeAborts\left(
R \right)$.

The specification of Classic Paxos ACA rounds uses the variables of
the specification of conservative ACA rounds, namely the variables
$proposed$, $inits$, $aborts$, the arrays $learned\left[ l \right]$,
$status\left[ a \right]$, and $hist\left[ a \right]$, and two
additional arrays $accepted\left[ l \right]\left[ a \right]$ and
$stop\left[ s \right]\left[ a \right]$, where $l$ is a learner, $a$ is
an acceptor, and $s$ is a switcher.

There is also a network component that we do not describe explicitly.
However, the network allows any servers to communicate by sending and
receiving messages.

Initially, the variables $proposed$, $inits$, $aborts$, and, for every
learner $l$, $learned\left[ l \right]$ are empty sets. For every
acceptor $a$, $status\left[ a \right]$ is ``idle'' and $hist\left[ a
\right]$ is the special value $None$. For every learner $l$ and acceptor $a$,
$accepted\left[ l \right]\left[ a \right]$ is the special value
$None$. For every switcher $s$ and acceptor $a$, $stopped\left[ s
\right]\left[ a \right]$ is set to $false$.  There are initially no
messages in the network.

The next-state relation is composed of the following types of
transitions, where the definitions of learn quorums, abort quorums,
and of the $SafeAborts\left( R \right)$ operator are the ones of conservative
ACA rounds.
\begin{compactitem}
    \item A $Propose\left( p,c \right)$ transition, for a proposer $p$
        and a command $c$, not guarded, inserts $c$ in the set
        $proposed$ and sends the message $\aseq{\quo{prop},c}$ to
        the leader.
    \item An $Init\left( s, h \right)$ transition, for a switcher $s$
        and a history $h$, not guarded, inserts $h$ in the set $inits$ and
        sends the message $\aseq{\quo{init},h}$ to the leader.
    \item A $WakeUp\left( leader \right)$ transition, enabled when
        $status\left[ leader \right]$ is ``idle'' and the leader can
        receive a $\aseq{\quo{init},h}$ message. Its effect is to
        receive the message, set $hist\left[ leader \right]$ to $h$,
        to set $status\left[ leader \right]$ to ``ready'', and to
        broadcast the message $\aseq{\quo{leader-accept},h}$ to all
        other acceptors.
    \item An $Accept\left( leader \right)$ transition, enabled when
        $status\left[ leader \right]$ is ``ready'' and the leader can
        receive a $\aseq{\quo{prop},c}$ message. The effect of the
        action is to receive the message, to update $hist\left[ leader
        \right]$ to $hist\left[ leader \right]\star c$, and to
        broadcast the message $\aseq{\quo{ack},hist\left[ leader
        \right]\star c}$ to the learners, and to broadcast the message
        $\aseq{\quo{leader-accept},hist\left[ leader \right]\star c}$
        to all the other acceptors.
    \item An $Accept\left( a \right)$ transition, for an acceptor $a$
        which is not the leader, is enabled when $status\left[ a
        \right]$ is ``ready'' and $a$ can receive a
        $\aseq{\quo{leader-accept},h}$ message where $h > hist\left[ a
        \right]$. The effect of the action is to receive the message,
        to update $hist\left[ a \right]$ to $h$, and to broadcast the
        message $\aseq{\quo{ack},hist\left[ a \right]\star c}$ to the
        learners.
    \item A $RcvAck\left( l \right)$ transition, for a learner $l$, is
        enabled when $l$ can receive a message $\aseq{\quo{ack},h}$
        from an acceptor $a$. Its effect is to receive the message and
        to set $accepted\left[ l \right]\left[ a \right]$ to $h$.
    \item A $Learn\left( l, h \right)$ transition, for a learner $l$
        and a history $h$, is enabled when $h > Max\left(
        learned\left[ l \right] \right)$ and there is a learn quorum
        $Q$ of acceptors such that $h = GLB\left( \aset{accepted\left[
        l \right]\left[ a \right] : a \in Q }\right)$. Its effect is
        to insert $h$ in $learned\left[ l \right]$.
    \item A $Stop\left( s \right)$ transition, for a switcher $s$,
        always enabled, broadcasts a $\aseq{\quo{stop}}$ message to all
        the acceptors.  This transition models a switcher being
        instructed by the scheduling policy to abort the current round.
    \item A $Stop\left( a \right)$ transition, for an acceptor $a$,
        enabled when $a$ can receive a $\aseq{\quo{stop}}$ message
        from a switcher $s$. Its effect is to receive the message, set
        $status\left[ a \right]$ to ``stopped'', and to send the
        message $\aseq{\quo{stopped},hist\left[ a \right]}$ to the
        switcher $s$.
    \item A $RcvStopped\left( s \right)$ transition, for a switcher
        $s$, enabled when $s$ can receive a $\aseq{\quo{stopped},h}$
        message from an acceptor $a$. Its effect is to receive the
        message and to set $stopped\left[ s \right]\left[ a \right]$
        to $h$.
    \item An $Abort\left( s, h \right)$ transition, for a switcher $s$
        and a history $h$, is enabled when there exists an abort quorum
        $R$ such that, for every member $a$ of $R$, $stopped\left[ s
        \right]\left[ a \right]$ is not $false$ and $h \in
        SafeAborts\left( R \right)$, where $SafeAborts\left( R
        \right)$ is computed using the array $stopped\left[ a
        \right]$. Its effect is to insert  $h$ in the set $aborts$. 
\end{compactitem}
Compared to the rounds of the original Classic Paxos, Classic Paxos ACA
rounds add an extra level of indirection when changing round: in the
original rounds, the leader of the new round directly queries the
acceptors of the preceeding round, whereas in ACA rounds, the
switchers query the acceptors and then send their abort histories to
the leader. However, changing round is an infrequent event, therefore
this difference should not have much practical impact on performance.
\begin{thm}
  Classic Paxos ACA rounds satisfy Abortable Chain Agreement.
\end{thm}
\begin{proof}
  A Classic Paxos ACA round implements the specification of Abortable
  Chain Agreement under the refinement mapping consisting in
  projecting the state of the round onto the variables $proposed$,
  $inits$, $aborts$, and, for every learner $l$, $learned\left[ l
  \right]$.
\end{proof}

\subsection{Multi-Coordinated Paxos, Chain Replication, and Ring Paxos.}

Multi-Coordinated Paxos, Chain Replication, and Ring Paxos rounds have a similar
structure as Classic Paxos rounds but differ in the way the leader is
implemented and in the way that the histories of the leader are disseminated to
the other acceptors.

Multi-Coordinated Paxos avoids depending on a unique leader by using a
distributed leader implementation that can make progress as long as a
strict majority of the acceptors are correct, trading off resilience
for a higher time complexity. Conservative ACA
rounds can be implemented like with Classic Paxos, replacing the
leader by the distributed leader of Multi-Coordinated Paxos.

In Chain Replication, acceptors are arranged in a chain whose head is
the leader of the round. As in Classic Paxos, abort histories and
proposed commands are sent to the leader. However, instead of the
leader directly sending its local history to all the other acceptors,
the local history of the leader is
forwarded along the chain from one acceptor to the other until it
reaches the last acceptor, which sends the history to the learners. 
By spreading the load more evenly on the all the servers, Chain Replication
can achieve higher throughput that Classic Paxos.
Conservative ACA rounds can be implemented like with Classic Paxos,
replacing the broadcast of the leader by the propagation of the leader's history
along the chain of servers.

Ring Paxos optimizes Chain Replication by having the leader broadcast
commands to the acceptors using IP multicast and only sending hashes
of the commands along the chain. Ring Paxos round can be modified to implement
conservative ACA rounds in a similar way as for Chain Replication.

\subsection{Fast Paxos}

Fast Paxos is composed of two types of rounds: classic rounds, like in Classic
Paxos, and fast rounds.
The fast rounds can be transformed into
implementations of fast ACA rounds. Instead of sending their
commands to a leader, as in Classic Paxos, the proposers of fast
rounds directly broadcast their commands to all the acceptors. As long
as commands are received in the same order by a learn quorum of
acceptors or are received in different orders but commute, commands can be learned in two message-propagation delays instead of three in Classic Paxos.
However, if non-commuting commands are received in different orders by
several acceptors, the local history of the acceptors may not form a
chain anymore and the rounds must abort. The fast rounds of Fast
Paxos implement fast ACA rounds.

Fast Paxos ACA rounds are specified using the same set of variables as for
Classic Paxos, initialized to the same values.
The next-state relation is composed of the following types of
transitions, where the definitions of learn quorums, abort quorums,
and of the $SafeAborts\left( R \right)$ operator are the ones of fast
ACA rounds.
\begin{compactitem}
    \item A $Propose\left( p,c \right)$ transition, for a proposer $p$
        and a command $c$, not guarded, inserts $c$ in the set
        $proposed$ and broadcasts the message $\aseq{\quo{prop},c}$ to
        all the acceptors.
    \item An $Init\left( s, h \right)$ transition, for a switcher $s$
        and a history $h$, not guarded, inserts $h$ in the set $inits$ and
        broadcasts the message $\aseq{\quo{init},h}$ to all the acceptors.
    \item A $WakeUp\left( a \right)$ transition, for an acceptor $a$,
        is enabled when $status\left[ a \right]$ is ``idle'' and $a$
        can receive an $\aseq{\quo{init},h}$ message. Its effect is to
        receive the message, set $hist\left[ a \right]$ to $h$, and to
        set $status\left[ a \right]$ to ``ready''.
    \item An $Accept\left( a \right)$ transition, for an acceptor $a$,
        is enabled when $status\left[ a \right]$ is ``ready'' and $a$
        can receive a $\aseq{\quo{prop},c}$ message. The effect of the
        action is to receive the message, to update $hist\left[ a
        \right]$ to $hist\left[ a \right]\star c$, and to broadcast
        the message $\aseq{\quo{ack},hist\left[ a \right]\star c}$ to
        the learners.
    \item The $RcvAck\left( l \right)$ transitions, for a learner $l$,
        the $Learn\left( l, h \right)$ transitions, for a learner $l$
        and a history $h$, the $Stop\left( s \right)$ transitions, for
        a switcher $s$, $Stop\left( a \right)$ transitions, for an
        acceptor $a$, the $RcvStopped\left( s \right)$ transitions,
        for a switcher $s$, and the $Abort\left( s, h \right)$
        transitions, for a switcher $s$ and a history $h$, which are
        all the same as in Classic Paxos ACA rounds.
\end{compactitem}

As for Classic Paxos ACA rounds, Fast Paxos ACA rounds implement ACA
under the refinement mapping consisting projecting the state of the
round onto the variables $proposed$, $inits$, $aborts$, and, for every
learner $l$, $learned\left[ l \right]$.

\newpage

\printbibliography

\begin{appendix}

  \label{sec:isaproofs}

%  \includepdf[pages={1},pagecommand={\section{Mechanical Proof of the
%  Composition Theorem in Isabelle/HOL}\label{sec:isaproofs}},scale=1]{../IsabellePDFS/podctheories.pdf}
%  \includepdf[pages={2-},pagecommand={},scale=1]{../IsabellePDFS/podctheories.pdf}

\end{appendix}

\end{document}

### 2015-05-09 paper/abstract1.tex
Work on the abstract

--- 2015-05-09 paper/paper.tex
@@
-\input{header.tex}
-\input{defs.tex}
-
-\addbibresource{./managed.bib}
-
-\title{Changing Flights In The Air\\ (Abortable Chain Agreement)}
-
-\author{
-  Rachid Guerraoui\\
-  \texttt{rachid.guerraoui@epfl.ch}
-  \and
-  Giuliano Losa\\
-  \texttt{giuliano.losa@epfl.ch}
-}
-
-\date{}
-
-\begin{document}
-
-\maketitle
-
 \abstract{
@@
 }
-
-\vspace{2cm}
-
-\begin{center}
-This paper is a regular submission.
-
-\vspace{1cm}
-This paper is a student paper.
-\end{center}
-
-\newpage
-
-\section{Introduction}
-
-The State-Machine Replication technique (abbreviated SMR) allows
-building reliable services on top of unreliable hardware. The basic idea is to 
-replicate a service over several servers, each executing the same
-sequence of (deterministic) commands. 
-Making SMR algorithms efficient has been a long-standing
-challenge and a multitude of algorithms have been proposed. Examples
-include Fast Paxos \cite{Lamport06FastPaxos}, Disk Paxos
-\cite{GafniLamport03DiskPaxos}, Chain Replication
-\cite{RenesseSchneider04ChainReplicationSupportingHighThroughputAvailability},
-Ring Paxos
-\cite{MarandiETAL10RingPaxosHighthroughputAtomicBroadcastProtocol},
-Egalitarian Paxos
-\cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments},
-Multi-Coordinated Paxos
-\cite{CamargosSchmidtPedone07MulticoordinatedPaxos}, Cheap Paxos
-\cite{LamportMassa04CheapPaxos}, Vertical Paxos
-\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication},
-Paxos-MIC
-\cite{HurfinMoiseNarzul11AdaptiveFastPaxosMakingQuickEverlasting},
-Mencius
-\cite{MaoJunqueiraMarzullo08MenciusBuildingEfficientReplicatedStateMachine},
-and Fast Mencius \cite{WeiETAL13FastMenciusMenciusLowCommitLatency}.
-
-Each of the existing algorithms is best suited to a particular range
-of environments, characterized, e.g., by client load, network
-topology, likelihood of faults, communication latency and bandwidth, processing
-power of nodes, availability of disk storage and memory.
-However, the environment faced by real systems evolves over time:
-hardware is upgraded, new data-centers are built, the popularity of a
-service changes, etc.  Consistent performance therefore comes with
-adaptation to change.  Unfortunately, existing SMR algorithms cannot easily be
-stopped and replaced by another algorithm on the fly. Moreover,
-devising a new algorithm that can change behavior dynamically is hard:
-dynamically switching from one SMR algorithm to another requires
-synchronizing notoriously intricate algorithms.
-
-We propose a new abstraction, \emph{Abortable Chain Agreement}
-(abbreviated ACA), which subsumes SMR and whose implementations,
-called \emph{ACA rounds}, can, by construction, pass the baton from
-one to another on the fly and efficiently. Moreover, optimizing a replicated
-service consisting of several ACA rounds becomes just a matter of devising a new
-ACA round, which can then be plugged-in on the fly: any new ACA round
-is compatible with every existing ACA round by construction.
-
-Abortable Chain Agreement is the \emph{most complete framework to date} enabling
-correct by construction composition of heterogeneous rounds.  
-ACA is completely formalized and its underlying composition theorem has been
-mechanically checked by an interactive proof assistant.  
-Compared to ACA,
-existing approaches
-\cite{BarNoyETAL87ShiftingGearsChangingAlgorithmsFlyTo,GuerraouiETAL10Next700BftProtocols,GuerraouiKuncakLosa12SpeculativeLinearizability}
-are limited. In particular, they do not allow the disorderly execution of
-commuting and read-only commands, an important practical optimization
-which is at the root of recent work on SMR like Generalized Consensus
-\cite{Lamport05GeneralizeConsensus} and Generalized Lattice Agreement
-\cite{FalerioETAL12GeneralizedLatticeAgreement}.
-
-\begin{comment}
-On top of the traditional roles of proposer, acceptor, and learners
-\cite{lamport2001paxos} used in consensus algorithms, ACA introduces
-the notion of \emph{switchers} and of \emph{scheduling policy}. The switchers
-have the task of aborting a running ACA round and initializing a new
-ACA round in a safe and live manner.  The \emph{scheduling policy}
-decides when to abort an ACA round and choses the next ACA round to
-execute. It must ensure that all switchers agree on the next ACA round
-to initialize.
-\end{comment}
-
-We observe that, because of the FLP impossibility result
-\cite{FischerLynchPaterson83ImpossibilityDistributedConsensusOneFaultyProcess},
-most existing SMR algorithms are structured in a sequence of abortable
-rounds which can easily be transformed into ACA rounds. To simplify
-this task even further, we propose two specifications, \emph{conservative
-ACA rounds} and \emph{fast ACA rounds}, which, as we show, can be concretized
-to obtain ACA versions of the rounds of five existing algorithms: 
-Classic Paxos, Fast Paxos, Chain Replication, Ring Paxos,
-and Multi-Coordinated Paxos. ACA guarantees that the composition of the five ACA
-rounds is correct by construction, forming a versatile adaptive replication
-algorithm.
-
-The rest of the paper is structured as follows.
-\Cref{sec:related} surveys recent related work. \Cref{sec:ca} presents Chain
-Agreement, which embodies the requirements for implementing SMR. \Cref{sec:aca}
-presents Abortable Chain Agreement and its optimizations to treat commuting
-and read-only commands efficiently. \Cref{sec:rounds} presents our specification
-of conservative and fast ACA rounds, which implement ACA. In
-\cref{sec:examples}, we show that the rounds of  
-Classic Paxos, Fast Paxos, Chain Replication, Ring Paxos,
-and Multi-Coordinated Paxos can easily be transformed into ACA implementations
-by using our specifications of conservative and fast ACA rounds.
-Finally, in \cref{sec:isaproofs}, we provide a complete formalization of ACA
-and a mechanical proof of the \emph{composition theorem} (\cref{thm:comp}),
-checked by the interactive proof assistant Isabelle/HOL \cite{NipkowPaulsonWenzel02IsabelleHOL}.
-
-\begin{comment}
-Abortable Chain Agreement builds on several recent advances in SMR
-algorithms: ACA allows acceptors to only eventually reach agreement,
-like in Generalized Paxos \cite{Lamport05GeneralizeConsensus},
-boosting performance when most commands commute; ACA allows fast
-execution of read commands like in Generalized Lattice Agreement
-\cite{FalerioETAL12GeneralizedLatticeAgreement}; ACA implementations
-can tolerate $f$ faults with only $f+1$ replicas and replace crashed replicas
-upon changing round, as in Vertical Paxos  \cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}.  
-\end{comment}
-
-\section{Related Work}
-\label{sec:related}
-
-SMR is often implemented using a sequence of instances of Consensus.
-There are numerous examples of consensus algorithms in the literature.
-However, Abortable Chain Agreement is based on Generalized Consensus. 
-In Generalized Consensus \cite{Lamport05GeneralizeConsensus}, the learners of a
-SMR algorithm repeatedly learn new values belonging to a \emph{C-struct set}. A
-C-struct can be seen as a data structure encoding  sequences of commands modulo
-the ordering of commands that commute. The advantage of Generalized
-Consensus over Consensus is that
-two servers which receive commuting commands in different orders can still reach
-agreement on C-structs without further synchronization. 
-Our Abortable Chain Agreement can also use C-structs, with the same advantages
-as Generalized Consensus. For simplicity, the presentation in this
-paper uses the more concrete notion of \emph{command history} (see
-\cref{sec:histories}).
-
-Generalized Lattice Agreement \cite{FalerioETAL12GeneralizedLatticeAgreement} allows more efficient implementations of SMR when
-commands are partitioned into read-only commands and write-only commands.
-Abortable Chain Agreement supports the same optimizations as Generalized Lattice Agreement.
-Notably, read-only commands are treated specially so as to optimize their
-executions (see \cref{sec:readonly}).
-
-Vertical Paxos algorithms 
-\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication} are SMR algorithms
-whose set of replica servers can be reconfigured without stopping to process new
-commands. Crashed servers can therefore be replaced by new servers, ensuring the
-long-term availability of a replicated service.
-Vertical Paxos algorithms use an external reconfiguration master, itself
-implemented as a replicated state machine, which determines the servers that should
-be removed from or added to the configuration.
-ACA shares some aspects of Vertical Paxos algorithms: the task of the
-reconfiguration master is similar to the task of the scheduling policy
-component of ACA (see \cref{sec:aca}).
-However, in ACA, changing round can be used not only to replace crashed servers
-but also to change the algorithm that the servers are running. 
-Changing algorithm is not possible in Vertical Paxos.
-In both Vertical Paxos and ACA, a reliable reconfiguration master, resp. a
-reliable scheduling policy, allows tolerating $f$ crashes with $f+1$ servers.
-
-The Abstract Framework \cite{GuerraouiETAL10Next700BftProtocols} proposes to
-build Byzantine Fault-Tolerant algorithms as the succession of abortable rounds
-called Abstract instances. Abstract instances must satisfy the Abstract
-correctness properties. By construction, the composition of any number of
-Abstract instances is a correct SMR implementation. However, the Abstract
-Framework uses totally ordered sequence of commands, making difficult the
-optimization of the execution of commuting and read-only commands that are
-possible with ACA.  Moreover, the Abstract Framework does not identify the
-crucial role of the scheduling policy for the long-term resilience of a service.
-Abortable Chain Agreement combines the ideas of the Abstract Framework,
-Generalized Paxos, Generalized Lattice Agreement, and Vertical Paxos in a
-unified abstraction.
-
-Finally, the Speculative Linearizability framework
-\cite{GuerraouiKuncakLosa12SpeculativeLinearizability} was a first
-attempt by the authors of the present paper at generalizing the
-Abstract Framework and mechanically proving its properties with the
-Isabelle/HOL interactive
-proofs assistant.  Like the Abstract Framework, the
-Speculative Linearizability framework is based on totally ordered
-sequences of commands, with the same drawback for the execution of
-commuting and read-only commands. Moreover, Abortable Chain Agreement is both
-conceptually simpler and subsumes the Speculative Linearizability
-property.  Like Speculative Linearizability, Abortable Chain Agreement
-and its main property, the composition theorem (\cref{thm:comp}), are
-formalized in Isabelle/HOL (see \cref{sec:isaproofs}).
-
-Abortable Chain Agreement is the most complete framework to date for building
-adaptive replication algorithms out of heterogeneous rounds, making state of the
-art optimization techniques available to algorithm designers, which was not
-possible before.
-
-\section{Chain Agreement}
-\label{sec:ca}
-
-We consider a set of asynchronous servers communicating by message passing in an
-asynchronous network. Servers are not necessarily sequential processes. We
-describe a behavior of the system using an infinite sequence of states, where a
-state is a function from variable names to values (i.e. a valuation
-of variables).  We specify sets of allowed behaviors using an initial predicate,
-a next-state relation (using unprimed and primed variables) and a fairness
-condition, all in the style of TLA+ \cite{Lamport02SpecifyingSystems}. However,
-for succinctness, our specification are rather informal, eluding non-essential
-details.
-
-Our specifications all contain by default a state variable $crashed$ and
-$Crash\left( x \right)$ transition which is always enabled. A server $x$ faithfully
-executes the algorithm that is assigned to it until its identifier is added to
-the variable $crashed$ by a $Crash\left( x \right)$ transitions.  When the identifier of a
-server is in the set $crashed$, all its actions are disabled.  In an execution,
-a server that is never added to the set $crashed$ is said correct.
-
-We consider a computing service exposing a set of commands $C$. The
-service has a sequential specification consisting of an initial state
-and a deterministic transition relation in which every command
-atomically updates the state of the service and produces an output. We
-assume that duplicate commands have no effect on the state of the
-service and return the same output as their first occurrence.
-
-In a traditional SMR algorithm, the servers can be seen has playing one or
-several of three roles: \emph{proposers}, which propose new commands to execute,
-\emph{acceptors}, which accept and order commands in a sequence, and
-\emph{learners}, which learn about the growing sequence of commands to
-execute \cite{lamport2001paxos}. Often, but not necessarily, every replica server plays the
-three roles of proposer, acceptor, and learner: clients of the service
-send their commands to some replica servers; replica servers submit
-the received commands to the SMR protocol in the role of proposers;
-replica servers respond to commands by executing the commands learned
-in the role of learners; replica servers participate in command
-ordering in the role of acceptors.  An SMR algorithm guarantees that
-the replicas learn the same sequences of commands, possibly up to the
-reordering of commuting commands. Therefore, every learner can respond to client
-requests using its last learned sequence of commands, and clients will receive
-the same responses as the ones produces by a centralized, non-replicated,
-service.
-
-Chain Agreement precisely specifies the allowed behaviors of the roles of \emph{proposers}
-and \emph{learners}, abstracting over the behavior of acceptors.
-
-\begin{comment}
-We suppose that each server can play a number of \emph{roles}.
-Chain Agreement specifies the allowed behaviors the roles of \emph{proposers}
-and \emph{learners}. Proposers receive commands from the clients of the service
-and propose them for execution by the replicated service. Each learner learns about
-a growing sequence of commands. Thanks to the sequential specification of the
-service, a sequence of commands determines an output for each the included commands. 
-In a traditional SMR algorithm, servers can also play the role of
-\emph{acceptors}, whose task is to collectively order commands and let the
-learners know about the latest sequence of ordered commands.
-
-The point of Chain Agreement is to implement a replicated service by having every learner learn about a growing sequence of commands, which
-represent an execution of the sequential specification of the service.
-Crucially, at all times, the sequences of commands learned by the
-different learners must form a chain in the prefix order on sequences.
-Moreover, for every command $c$ proposed by a correct proposer, every
-correct learner must eventually learn a sequence containing $c$.  Once
-a learner learns about a sequence containing a command $c$, it can
-determine the output produced by $c$, using the sequential
-specification of the service, and respond to the client that issued
-the command $c$.
-\end{comment}
-
-To specify Chain Agreement, we need the notion of ordering of sequences and of
-maximum sequence.  The prefix order on sequences is denoted by the operator
-$\leq$ ($s_1 \leq s_2$ means that $s_1$ is a prefix of $s_2$). 
-%Using the operator $\leq$, we define $\geq$, $<$, and $>$ as expected. 
-Note that $\leq$ is a partial order on sequences. The maximum over a set of
-sequences $S$ is denoted by $Max\left( S \right)$ and $Max\left( \aset{}
-\right)$ is the empty sequence.
-
-We specify Chain Agreement using the state variables $crashed$, $proposed$, containing the proposed commands,
-and, for every learner $l$,  $learned\left[ l \right]$, a set of
-sequences of commands. Initially, $crashed$, $proposed$ and, for every learner $l$,
-$learned\left[ l \right]$ are empty sets. 
-Remember that any transition involving a server $x$ is always disabled when $x\in crashed$.
-
-The next-state relation consists of three types of transitions:
-$Crash\left( x \right)$ transitions; $Propose\left( l, c
-\right)$ transitions, for a learner $l$ and a command $c$, in which
-$c$ is inserted in the set $proposed$; $Learn\left( l,cs \right)$
-transitions, for a learner $l$ and a sequence of commands $cs$, in
-which $cs$ is inserted in $learned\left[ l \right]$. 
-A $Learn\left( l, cs \right)$ transition is guarded by the following
-conditions:
-\begin{compactitem}
-    \item[\textbf{Agreement}:] The elements of the set $\aset{learned'\left[ l
-      \right] : l \in Learners}$, the learned sequences after the transition,
-      form a chain in the partial order $\leq$.
-    \item[\textbf{Validity}:] Every learned sequence contains only
-        proposed commands.  
-    \item[\textbf{Irrevocability}:] The new learned sequence
-        $cs$ is a strict extension of $Max\left(learned\left[ l
-        \right]\right)$. 
-        %\\ $cs > Max\left(learned\left[ l \right]\right)$.
-\end{compactitem}
-Finally, we add, for every learner $l$, one weak fairness condition
-\cite{Lamport02SpecifyingSystems} on the transitions $Learn\left( l,
-cs \right)$. The weak fairness conditions ensure that every correct
-learner eventually learns about every proposed command.
-
-Thanks to the properties of Chain Agreement, the clients, observing
-their input commands and the received responses, see a
-\emph{linearizable} execution
-\cite{HerlihyWing90LinearizabilityCorrectnessConditionConcurrentObjects}.
-Moreover, the service remains available as long as at least one
-proposer and one learner are correct.
-
-\section{Abortable Chain Agreement}
-\label{sec:aca}
-
-We now consider implementing Chain Agreement as a sequence of \emph{Abortable Chain
-Agreement rounds}, where each round makes progress ordering requests or aborts
-its execution and passes the baton to the next round in the sequence.
-An algorithm composed of a sequence of ACA rounds is called an \emph{ACA
-algorithm}.
-
-On top of the traditional roles of proposer, acceptor, and learner, ACA
-introduces the new role of \emph{switcher} and a new component, the
-\emph{scheduling policy}. In ACA, the assignment of the roles of proposer,
-acceptor, learner, and switcher can change in each rounds, allowing for
-reconfiguration. 
-In normal circumstances, the proposers, acceptors,
-and learners of an ACA round work together to implement Chain Agreement.
-However, if the environment becomes unfavorable to a running ACA round, the
-scheduling policy can instruct the round to abort its execution and pass the
-baton, using the switchers, to a new round.
-
-A switcher of an aborting round determines, from the state of the
-acceptors, an \emph{abort sequence} that it transfers to one or several
-switchers of the new round.  Upon receiving an abort sequence, a switcher of the
-new round initializes the acceptors of the new round, which can then
-proceed implementing SMR, until the new round itself aborts and the
-process is repeated. 
-We assume that in the first round, which has no previous round, every 
-switchers receives the empty abort sequence.
-
-The scheduling policy determines when to abort an ACA round and which new ACA
-round should take over. The scheduling policy must inform the switchers about the
-identity of the next round and must guarantee that all switchers receive the
-same information.  In this paper we abstract over the scheduling policy, assuming
-that a round that aborts passes the baton a nondeterministically chosen new
-round. A reliable scheduling-policy component can realistically be implemented
-as a replicated state machine running on a large number of servers otherwise
-dedicated to other tasks
-\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}. We will see
- in \cref{sec:rounds} that, as in Vertical Paxos, a reliable scheduling policy component allows
-tolerating $f$ faults with $f+1$ replicas, while maintaining the long-term
-resilience of the service by replacing faulty servers upon round change.
-
-The clients of an ACA algorithm can send their commands to the proposers of any round (in
-practice, clients send their commands to proposers of the largest
-active round) and servers can use sequences learned in any round
-to determine outputs.
-
-ACA specifies the allowed behaviors of the proposers, learners, and switchers
-\emph{of a single round}. Similarly to
-Chain Agreement, we use the state variables $crashed$, $proposed$,
-and, for every learner $l$, $learned\left[ l \right]$.  Additionally,
-we use the state variables $inits$, containing the abort sequences
-received from the previous round, and $aborts$, containing the abort
-sequences transfered to the next round.
-Two consecutive ACA rounds are composed by letting the $aborts$ variable of the
-first round be the $inits$ variable of the second round.
-
-Initially, $crashed$, $proposed$, for every
-learner $l$, $learned\left[ l \right]$, $inits$, and $aborts$ are all empty sets.
-The next-state relation consists of five types of transitions. The
-first three, $Crash\left( x \right)$, $Propose\left( l, c
-\right)$, and $Learn\left( l, cs \right)$, for a learner, proposer, or
-switcher $x$, for a learner $l$, and for a sequence of commands
-$cs$, have the same effect as in Chain Agreement.
-ACA introduces two new transitions, $Init\left(
-s, cs \right)$ and $Abort\left( s, cs \right)$, for a switcher $s$ and a sequence of
-commands $cs$. The transition $Init\left( s, cs \right)$ adds $cs$ to the set $inits$, and models a switcher receiving the
-abort sequence $cs$ from the preceding round.
-The transition $Abort\left( s, cs \right)$ adds $cs$ to the set $aborts$ and models a switcher transferring the abort sequence $cs$ to a switcher of the next round.
-All the transition are guarded by the following conditions:
-\begin{compactitem}
-    \item[\textbf{Agreement}:] The elements of the set $\aset{learned'\left[ l
-      \right] : l \in Learners}$ form a chain in the partial order.
-    \item[\textbf{Validity}:] Every abort sequence or learned sequence
-        is of the form $h \star cs$, where $cs$ contains only commands
-        of the set $proposed$ and $h$ is the greatest lower bound of a
-        nonempty subset of $inits$.
-    \item[\textbf{Initialization}:] If $inits$ is empty, then no
-        sequence can be learned and no switcher can abort.
-    \item[\textbf{Irrevocability}:] In the case of a $Learn\left( l,cs
-        \right)$ transition, the new learned sequence $cs$ is a strict
-        extension of $Max\left(learned\left[ l \right]\right)$.
-        %, $cs
-        %\geq Max\left(learned\left[ l \right]\right)$.
-    \item[\textbf{Safe Abort}:] Every abort sequence is larger in the
-        partial order $\leq$ than every learned sequence.
-\end{compactitem}
-Note that Agreement and Irrevocability are the same as in Chain
-Agreement.
-
-Finally, for every learner $l$, we add a weak fairness condition on the
-following state transition: 
-\begin{equation}
-    \exists  s \in Switchers, cs \in Sequence\left( C \right) : Learn\left( l, cs \right) \vee Abort\left( s,cs \right)
-\end{equation}
-The weak fairness conditions ensure that when a correct learner has not
-learned a sequence containing all commands proposed in the round, then
-it eventually learns a new command or, if there is at least one correct switcher, at least one switch value is transfered to the next round. 
-Therefore, if a client has submitted a command to a proposer, eventually,
-either the client gets a response from the current round, or
-the next round is initialized with an abort sequence and the
-command can be resubmitted to the next round.
-
-\subsection{The Composition Theorem}
-
-Consider the succession of $n$ rounds numbered $1$ to $n$.  To
-distinguish the variables of each round, we annotate the local
-variables of each round with the number of the round. For example, the
-variables of the seventh round are written $proposed_7$, for every
-learner $l$, $learned_7\left[ l \right]$, $inits_7$, and $aborts_7$.
-
-Moreover, we assume that a command, which is proposed to a round and that
-is not learned before the first abort sequence is transfered to the
-next round, is proposed anew to the next round.
-
-Define the state variable $proposed$ as the union over all rounds $i$
-of the $proposed_i$ variables and, for every learner $l$, 
-define the $learned\left[ l \right]$ variable as the union over all
-rounds $i$ of the $learned_i\left[ l \right]$ variables.
-
-\begin{thm}[Composition Theorem]
-    \label{thm:comp}
-    In every execution of a succession of ACA rounds, the sequence of valuations
-    of the variable $proposed$ and, for every learner $l$, of the variables
-    $learned\left[ l \right]$ satisfies Chain Agreement.
-\end{thm}
-\Cref{thm:comp} can be shown by induction on the number of rounds,
-using the following lemma.
-\begin{lem}
-  \label{lem:comp}
-    Consider two rounds numbered $i$ and $i+1$. In every execution of a
-    succession of ACA rounds, the sequence of valuations of the variables $inits
-    = inits_i$, $propose = proposed_i \cup propose_{i+1}$, for every learner
-    $l$, $learned\left[ l \right] = learned_i\left[ l \right] \cup
-    learned_{i+1}\left[ l \right]$and $aborts = aborts_{i+1}$ satisfies ACA.
-\end{lem}
-\begin{proof}
-  A mechanically-checked proof in Isabelle/HOL appears in \cref{sec:isaproofs}.
-  The proofs uses a version of the specification of ACA that has a more
-  executable form that presented in this section. Moreover,
-  instead of sequences of commands, the Isabelle/HOL specification uses the more
-  general notion of command history, presented in the next subsection.
-  The Isabelle/HOL formalization of ACA uses the theory of I/O automata, which
-  is also formalized in the Isabelle/HOL. The proof uses a refinement
-  mapping from the composition of two consecutive ACA instances to a single ACA
-  instance.
-\end{proof}
-
-\Cref{thm:comp} guarantees that the successive composition of any
-number of ACA rounds implements Chain Agreement.
-Therefore, with ACA, one can easily build replicated services that
-dynamically adapt to the current operating conditions of the system: a
-replicated service has several types of rounds at its disposal, each being
-tailored to particular conditions. Each time the conditions change,
-the current ACA round aborts and passes the baton to a more 
-appropriate type of round, as determined by the scheduling policy. Moreover, a new type
-of ACA round can always be devised after the system is deployed and added on the fly to
-the existing types of rounds.
-
-We now describe two important practical optimizations of ACA.
-
-\subsection{Histories}
-\label{sec:histories}
-
-ACA can be implemented more efficiently by substituting \emph{command histories} for sequences of commands, as in Generalized Consensus \cite{Lamport05GeneralizeConsensus}.
-
-We say that two commands $c_1$ and $c_2$ \emph{commute} when, in every sequential
-execution of the service, reversing the order of $c_1$ and $c_2$ does
-not change any of the outputs, including those given to $c_1$ and
-$c_2$.
-
-A command history, abbreviated history, is a data structure that
-represents a set of sequences which are the same modulo the reordering
-of commuting commands, and in which duplicate commands do not appear.
-All the sequences represented by a given history contain the same
-commands and determine the same outputs for each command.  Therefore,
-to implement a replicated service, it is sufficient that learners
-learn command histories instead of sequences.  For a precise
-definition of command histories, we refer the reader to
-\cite{Lamport05GeneralizeConsensus}. Our Isabelle/HOL formalization
-(\cref{sec:isaproofs}) abstracts
-over the concrete representation of histories, assuming only their necessary
-properties. Command histories were first
-introduced in trace theory \cite{Mazurkiewicz84Semantics}.
-
-A command history can be built by appending a sequence of commands
-$cs$ to a history $h$, obtain the history $h \star cs$.  The major
-advantage that histories have over sequences is that if a sequence
-$cs_1$ is obtained by reordering the commuting commands of $cs_2$,
-then $h \star cs_1 = h \star cs_2$. Therefore, in  implementations of
-ACA, servers can agree on a history even though
-they chose different orders for some commuting commands.
-
-Command histories share the four properties of sequences that are
-necessary and sufficient to substitute histories for sequences in the
-definition of Chain Agreement and ACA.
-\begin{compactitem}
-    \item[\textbf{Distinctness}:] No duplicate command appears in a history. 
-    \item[\textbf{Ordering}:] Command histories form a partial order:
-        $h_1 \leq h_2$ when there exists a sequence $s$ such that $h_2
-        = h_1 \star s$.
-    \item[\textbf{GLB}:] Every finite set of command histories has a
-        \emph{greatest lower bound}.
-%    \item[\textbf{LUB}:] Every finite set of compatible command
-%        histories has a \emph{least upper bound}, where a set of
-%        command histories is compatible when they have a common upper
-%        bound.
-    \item[\textbf{Consistency}:] If $h_1 \leq h_2 \leq h_3$ and $h_3 =
-        h_1 \star s$ then there exists $s'$ and $s''$ where every
-        request of $s'$ and $s''$ appears in $s$ and such that $h_2 =
-        h_1 \star s'$ and $h_3 = h_2 \star s''$.
-\end{compactitem}
-The optimized version of ACA is obtained by
-simply substituting histories for sequences in the definition of ACA.
-The Ordering, GLB, and Consistency properties of histories ensure that
-\cref{thm:comp} still holds: the Isabelle/HOL proof of \cref{thm:comp} (see
-\cref{sec:isaproofs}) relies only on the four properties above.
-
-\subsection{Optimizing the Execution of Read-Only Commands}
-\label{sec:readonly}
-
-Chain Agreement can be further optimized by treating 
-read-only commands specially, as in Generalized Lattice Agreement
-\cite{FalerioETAL12GeneralizedLatticeAgreement}.
-
-We consider the case in which a set of servers each play the four
-roles of proposer, acceptor, learner, and switcher.
-Clients sends their commands to any server, which proposes them as a
-proposer, and records the command as pending.
-%When a server learns a history that contains one of its pending
-%commands, it uses the history to determine the response to the pending
-%command and forwards the response to the appropriate client.
-
-As observed in \cite{FalerioETAL12GeneralizedLatticeAgreement}, read-only
-commands can be executed faster by directly reading from the last learned
-history, instead of proposing the read commands and waiting for it to appear in
-a future learned history.  Because learned histories form a chain, all the reads
-return consistent values and this approach yields a \emph{serializable}
-implementation of the service.
-
-However linearizability is not satisfied if reads are served directly
-from the last learned history: a lagging replica server could respond
-to a read using a history which does not contain all completed update
-commands. To make the executions linearizable without resorting to
-proposing read-only commands as normal commands, we can employ the same
-technique as in \cite{FalerioETAL12GeneralizedLatticeAgreement}: upon
-receiving a read command, a server proposes a special no-op command
-that commutes with all other commands; the server then responds to the
-read command using the first learned history containing the no-op
-command. This ensures that all the commands that had completed before
-the read command was issued are present in the history.  Moreover,
-since no-op commands commute with all other commands, they can be
-treated more efficiently than regular read commands by the acceptors,
-like, for example, in a fast round of Fast Paxos.
-
-\section{Fast and Conservative Rounds}
-\label{sec:rounds}
-
-Most existing SMR algorithms can easily be obtained as the
-composition of one or more ACA rounds.
-
-The FLP impossibility result \cite{FischerLynchPaterson83ImpossibilityDistributedConsensusOneFaultyProcess}
-implies that no algorithm can implement SMR to the letter in practice.
-Instead, SMR algorithms guarantee safety (i.e. replica execute the
-same sequence of commands) but may not be able to make progress (i.e.
-order new commands). To make progress as likely as possible, most SMR
-algorithms are structured in consecutive rounds (also called ballots,
-views, or epochs). In each round, the algorithm tries to make progress
-but cannot be guaranteed to succeed. When the algorithm detects that
-progress is no more possible, a new round is started.  The new round
-is initialized so as to preserve safety, usually by one of the
-acceptors. For example, in Paxos, the leader of a newly started ballot
-numbered $i$ determines a value that is \emph{safe at $i$} and
-proposes it as first value to the other acceptors of the ballot $i$.
-In existing SMR algorithms, new rounds are initialized ad-hoc, which
-prevents mixing rounds of two different algorithms.  Instead,
-ACA is an abstraction of rounds which establishes a common interface and contract for switching from
-one round to the next, enabling switching from any ACA round to any other.
-
-To simplify the task of expressing existing algorithms as the composition of one
-or more ACA rounds, we propose two specifications, \emph{conservative ACA
-rounds} and
-\emph{fast ACA rounds}, which can easily be concretized to obtain executable rounds.
-We show that ACA versions of Classic Paxos, Multi-Coordinated Paxos, Chain
-Replication, and Ring Paxos can be obtained by concretizing the specification of
-conservative ACA rounds, and that ACA versions of the fast rounds of Fast Paxos can be obtained by concretizing the specification of fast ACA rounds.
-
-Fast and conservative ACA rounds are two specifications, at an
-intermediate level of abstraction, which satisfy ACA. Both
-specifications specify how the state of the acceptors should be
-updated but abstract over the particular communication topology used
-to implement the state accesses: the specifications use a global state
-which is accessible by all servers in all roles. One can refine fast
-or conservative rounds by implementing the state accesses and coordination
-using the network, obtaining a concrete algorithm.
-
-Conservative and fast ACA rounds are specified using the state variables $inits$, $proposed$, for every learner $l$, $learned\left[ l \right]$
-and $aborts$, as in ACA, and, for every acceptor $a$, two new variables
-$status\left[ a \right]$, which can be either ``idle'', ``ready'', or
-``stopped'' and $hist\left[ a \right]$, the local history of the
-acceptor $a$, which can be either a history or the special value $None$.
-Initially, $inits$, $aborts$, and, for every learner
-$l$, $learned\left[ l \right]$ are empty sets, and, for every acceptor $a$,
-$status\left[ a \right]$  is ``idle'' and $hist\left[ a \right]$ is the special
-value $None$.
-
-The next-state relations of conservative and fast rounds are composed
-of the following types of transitions (and the default $Crash\left( x
-\right)$ transition).
-\begin{compactitem}
-    \item A $Propose\left( p,c \right)$ transition, for a proposer $p$
-        and a command $c$, not guarded, inserts $c$ in the set
-        $proposed$. 
-    \item An $Init\left( s, h \right)$ transition, for a switcher $s$
-        and a history $h$, not guarded, inserts $h$ in the set $inits$.
-    \item A $WakeUp\left( a \right)$ transition, for an acceptor $a$,
-        is enabled when $status\left[ a \right]$ is ``idle''. Its
-        effect is to set $hist\left[ a \right]$ to a history found in
-        $inits$ and to set $status\left[ a \right]$ to ``ready''.
-    \item An $Accept\left( a \right)$ transition, for an acceptor $a$,
-        is enabled when $status\left[ a \right]$ is ``ready'', and
-        updates $hist\left[ a \right]$ to $hist\left[ a \right]\star
-        c$, where $c$ is a proposed command. In conservative rounds,
-        the transition is additionally guarded by the condition that
-        the resulting set of local acceptor histories,
-        $\aset{hist'\left[ a \right] : a \in Acceptors}$, forms a chain
-        in the partial order $\leq$ (this is typically implemented with a
-        leader).
-    \item A $Learn\left( l, h \right)$ transition, for a learner $l$
-      and a history $h$, enabled when $h > learned\left[ l \right]$ and is
-      smaller than the GLB of the histories of a learn quorum of acceptors. The effect of
-        the transition is to set $learned\left[ l \right]$ to $h$.
-    \item A $Stop\left( a \right)$ transition, for an acceptor $a$,
-        enabled when $status\left[ a \right]$ is ``ready'', sets
-        $status\left[ a \right]$ to ``stopped'' (note that
-        $status\left[ a \right] = $ ``stopped'' prevents any further
-        ``accept'' transition of acceptor $a$).
-%    \item REALLY NEEDED? A $TimeOut\left( a \right)$ transition, for an acceptor $a$,
-%        is enabled when $status\left[ a \right]$ is ``idle'' and sets 
-%        $status\left[ a \right]$ to ``stopped'' and $hist\left[ a
-%        \right]$ to the special value $None$. 
-    \item An $Abort\left( s, h \right)$ transition, for a switcher $s$
-        and a history $h$, is enabled when there exists a abort quorum
-        $R$ such that, for every member $a$ of $R$, $status\left[ a
-        \right]$ is ``stopped'' and $h \in SafeAborts\left( R
-        \right)$. Its effect is to insert  $h$ in the set $aborts$. 
-\end{compactitem}
-
-On top of the additional guard on the $Accept$ transition present only
-in conservative rounds, conservative and fast ACA rounds differ in
-their definition of learn quorums, abort quorums, and in their
-definition of the $SafeAborts\left( R \right)$ operator.
-To concisely define read and learn quorums, we define 
-\begin{equation}
-    Q\left( f \right) = \aset{as \subseteq Acceptors :
-        Card\left(as\right) \geq \left\lfloor f\times Card\left(
-    Acceptors \right)\right\rfloor + 1}.
-\end{equation}
-
-In conservative ACA rounds, abort and learn quorums are defined such
-that the intersection between two learn quorums is nonempty and the
-intersection between a learn quorum and an abort quorum is nonempty.
-The definition leads to two interesting cases: $LearnQ = AbortQ =
-Q\left( 1/2 \right)$; $LearnQ = \aset{Acceptors}$ and $AbortQ =
-\aset{\aset{a} : a \in Acceptors}$. The operator $SafeAborts\left( R
-\right)$ is defined as the singleton set containing the maximum
-element over the local histories of the acceptors in $R$ which are not
-$None$. Note that the maximum exists because conservative rounds
-guarantee that the histories of the acceptors form a chain.
-
-In fast ACA rounds, abort and learn quorums are defined such that
-the intersection between any two learn quorums is nonempty and the
-intersection between any two learn quorums and an abort quorum is
-nonempty.  The definition leads to three interesting cases: $LearnQ =
-AbortQ = Q\left( 2/3 \right)$; $LearnQ = Q\left( 1/2 \right)$ and
-$AbortQ = Q\left( 3/4 \right)$; $LearnQ = \aset{Acceptors}$ and
-$AbortQ = \aset{\aset{a : a \in Acceptors}}$.
-The operator $SafeAborts\left( R \right)$ is defined as the set of
-maximal elements over the set $\aset{GLB\left( \aset{hist\left[ a
-\right] : a \in R \cap L \wedge hist\left[ a \right]\neq None} \right)
-: L \in LearnQ}$.
-
-The cases in which abort quorums are singleton sets, in both
-conservative and fast rounds, allow tolerating
-$f$ faults with $f+1$ servers: a round can abort as long as one
-acceptor is correct and a new round, possibly using new correct
-servers, can take over. Note that changing round cannot happen
-without the scheduling policy instructing the switchers about the
-identity of the next round, and thus tolerating $f$ faults with
-$f+1$ servers relies on the assumption of a reliable scheduling policy.
-
-Define, for every learner $l$, the history variable $\overline{learned}\left[ l \right]$ which is initialized to the empty set and which is
-updated each time a new local history is accepted by the learner $l$ by
-inserting the new local history in $\overline{learned}\left[ l \right]$ (see \cite{AbadiLamport91ExistenceRefinementMappings} for the
-definition of a history variable).
-\begin{thm}
-  \label{thm:rounds}
-    Both conservative and fast ACA rounds implement ACA under the refinement
-    mapping that consists in substituting $\overline{learned\left[ l \right]}$
-    for $learned\left[ l \right]$.
-\end{thm}
-
-\section{Examples of Concrete ACA Algorithms}
-\label{sec:examples}
-
-In this section we show how to modify the rounds of Classic Paxos,
-Multi-Coordinated Paxos, Chain Replication, Ring Paxos, and Fast Paxos to obtain
-ACA rounds. \Cref{thm:comp} then guarantee that all those different types of
-rounds can be used in the same algorithm to yield an implementation of Chain
-Agreement.
-
-\subsection{Classic Paxos}
-
-The rounds of the Classic Paxos algorithm can be modified so as to
-obtain an implementation of conservative ACA rounds.  We define
-Classic Paxos ACA rounds by specializing the specification of
-conservative ACA rounds, giving a concrete protocol for disseminating
-the proposed commands to the acceptors and enforcing that the learned
-histories form a chain.
-
-Each Classic Paxos round has a unique leader among the acceptors which
-centrally orders the proposed commands, which are accepted by the
-other acceptors in the order determined by the leader.  This protocol
-ensures that every local acceptor history is a prefix of the local
-history of the leader, therefore guaranteeing that the local acceptor histories
-forming a chain.
-
-Acceptor broadcast each newly accepted history to the learners, which
-learn a history $h$ when they receive enough extensions of $h$ from the
-acceptors.  When instructed by the scheduling policy to abort, a
-switcher sends a stop message to the acceptors, which irrevocably stop
-updating their local history and send back to the switcher an
-acknowledgement containing their local history. A switcher
-determines an abort history once it has received acknowledgments from
-an abort quorum of acceptors $R$, using the operator $SafeAborts\left(
-R \right)$.
-
-The specification of Classic Paxos ACA rounds uses the variables of
-the specification of conservative ACA rounds, namely the variables
-$proposed$, $inits$, $aborts$, the arrays $learned\left[ l \right]$,
-$status\left[ a \right]$, and $hist\left[ a \right]$, and two
-additional arrays $accepted\left[ l \right]\left[ a \right]$ and
-$stop\left[ s \right]\left[ a \right]$, where $l$ is a learner, $a$ is
-an acceptor, and $s$ is a switcher.
-
-There is also a network component that we do not describe explicitly.
-However, the network allows any servers to communicate by sending and
-receiving messages.
-
-Initially, the variables $proposed$, $inits$, $aborts$, and, for every
-learner $l$, $learned\left[ l \right]$ are empty sets. For every
-acceptor $a$, $status\left[ a \right]$ is ``idle'' and $hist\left[ a
-\right]$ is the special value $None$. For every learner $l$ and acceptor $a$,
-$accepted\left[ l \right]\left[ a \right]$ is the special value
-$None$. For every switcher $s$ and acceptor $a$, $stopped\left[ s
-\right]\left[ a \right]$ is set to $false$.  There are initially no
-messages in the network.
-
-The next-state relation is composed of the following types of
-transitions, where the definitions of learn quorums, abort quorums,
-and of the $SafeAborts\left( R \right)$ operator are the ones of conservative
-ACA rounds.
-\begin{compactitem}
-    \item A $Propose\left( p,c \right)$ transition, for a proposer $p$
-        and a command $c$, not guarded, inserts $c$ in the set
-        $proposed$ and sends the message $\aseq{\quo{prop},c}$ to
-        the leader.
-    \item An $Init\left( s, h \right)$ transition, for a switcher $s$
-        and a history $h$, not guarded, inserts $h$ in the set $inits$ and
-        sends the message $\aseq{\quo{init},h}$ to the leader.
-    \item A $WakeUp\left( leader \right)$ transition, enabled when
-        $status\left[ leader \right]$ is ``idle'' and the leader can
-        receive a $\aseq{\quo{init},h}$ message. Its effect is to
-        receive the message, set $hist\left[ leader \right]$ to $h$,
-        to set $status\left[ leader \right]$ to ``ready'', and to
-        broadcast the message $\aseq{\quo{leader-accept},h}$ to all
-        other acceptors.
-    \item An $Accept\left( leader \right)$ transition, enabled when
-        $status\left[ leader \right]$ is ``ready'' and the leader can
-        receive a $\aseq{\quo{prop},c}$ message. The effect of the
-        action is to receive the message, to update $hist\left[ leader
-        \right]$ to $hist\left[ leader \right]\star c$, and to
-        broadcast the message $\aseq{\quo{ack},hist\left[ leader
-        \right]\star c}$ to the learners, and to broadcast the message
-        $\aseq{\quo{leader-accept},hist\left[ leader \right]\star c}$
-        to all the other acceptors.
-    \item An $Accept\left( a \right)$ transition, for an acceptor $a$
-        which is not the leader, is enabled when $status\left[ a
-        \right]$ is ``ready'' and $a$ can receive a
-        $\aseq{\quo{leader-accept},h}$ message where $h > hist\left[ a
-        \right]$. The effect of the action is to receive the message,
-        to update $hist\left[ a \right]$ to $h$, and to broadcast the
-        message $\aseq{\quo{ack},hist\left[ a \right]\star c}$ to the
-        learners.
-    \item A $RcvAck\left( l \right)$ transition, for a learner $l$, is
-        enabled when $l$ can receive a message $\aseq{\quo{ack},h}$
-        from an acceptor $a$. Its effect is to receive the message and
-        to set $accepted\left[ l \right]\left[ a \right]$ to $h$.
-    \item A $Learn\left( l, h \right)$ transition, for a learner $l$
-        and a history $h$, is enabled when $h > Max\left(
-        learned\left[ l \right] \right)$ and there is a learn quorum
-        $Q$ of acceptors such that $h = GLB\left( \aset{accepted\left[
-        l \right]\left[ a \right] : a \in Q }\right)$. Its effect is
-        to insert $h$ in $learned\left[ l \right]$.
-    \item A $Stop\left( s \right)$ transition, for a switcher $s$,
-        always enabled, broadcasts a $\aseq{\quo{stop}}$ message to all
-        the acceptors.  This transition models a switcher being
-        instructed by the scheduling policy to abort the current round.
-    \item A $Stop\left( a \right)$ transition, for an acceptor $a$,
-        enabled when $a$ can receive a $\aseq{\quo{stop}}$ message
-        from a switcher $s$. Its effect is to receive the message, set
-        $status\left[ a \right]$ to ``stopped'', and to send the
-        message $\aseq{\quo{stopped},hist\left[ a \right]}$ to the
-        switcher $s$.
-    \item A $RcvStopped\left( s \right)$ transition, for a switcher
-        $s$, enabled when $s$ can receive a $\aseq{\quo{stopped},h}$
-        message from an acceptor $a$. Its effect is to receive the
-        message and to set $stopped\left[ s \right]\left[ a \right]$
-        to $h$.
-    \item An $Abort\left( s, h \right)$ transition, for a switcher $s$
-        and a history $h$, is enabled when there exists an abort quorum
-        $R$ such that, for every member $a$ of $R$, $stopped\left[ s
-        \right]\left[ a \right]$ is not $false$ and $h \in
-        SafeAborts\left( R \right)$, where $SafeAborts\left( R
-        \right)$ is computed using the array $stopped\left[ a
-        \right]$. Its effect is to insert  $h$ in the set $aborts$. 
-\end{compactitem}
-Compared to the rounds of the original Classic Paxos, Classic Paxos ACA
-rounds add an extra level of indirection when changing round: in the
-original rounds, the leader of the new round directly queries the
-acceptors of the preceeding round, whereas in ACA rounds, the
-switchers query the acceptors and then send their abort histories to
-the leader. However, changing round is an infrequent event, therefore
-this difference should not have much practical impact on performance.
-\begin{thm}
-  Classic Paxos ACA rounds satisfy Abortable Chain Agreement.
-\end{thm}
-\begin{proof}
-  A Classic Paxos ACA round implements the specification of Abortable
-  Chain Agreement under the refinement mapping consisting in
-  projecting the state of the round onto the variables $proposed$,
-  $inits$, $aborts$, and, for every learner $l$, $learned\left[ l
-  \right]$.
-\end{proof}
-
-\subsection{Multi-Coordinated Paxos, Chain Replication, and Ring Paxos.}
-
-Multi-Coordinated Paxos, Chain Replication, and Ring Paxos rounds have a similar
-structure as Classic Paxos rounds but differ in the way the leader is
-implemented and in the way that the histories of the leader are disseminated to
-the other acceptors.
-
-Multi-Coordinated Paxos avoids depending on a unique leader by using a
-distributed leader implementation that can make progress as long as a
-strict majority of the acceptors are correct, trading off resilience
-for a higher time complexity. Conservative ACA
-rounds can be implemented like with Classic Paxos, replacing the
-leader by the distributed leader of Multi-Coordinated Paxos.
-
-In Chain Replication, acceptors are arranged in a chain whose head is
-the leader of the round. As in Classic Paxos, abort histories and
-proposed commands are sent to the leader. However, instead of the
-leader directly sending its local history to all the other acceptors,
-the local history of the leader is
-forwarded along the chain from one acceptor to the other until it
-reaches the last acceptor, which sends the history to the learners. 
-By spreading the load more evenly on the all the servers, Chain Replication
-can achieve higher throughput that Classic Paxos.
-Conservative ACA rounds can be implemented like with Classic Paxos,
-replacing the broadcast of the leader by the propagation of the leader's history
-along the chain of servers.
-
-Ring Paxos optimizes Chain Replication by having the leader broadcast
-commands to the acceptors using IP multicast and only sending hashes
-of the commands along the chain. Ring Paxos round can be modified to implement
-conservative ACA rounds in a similar way as for Chain Replication.
-
-\subsection{Fast Paxos}
-
-Fast Paxos is composed of two types of rounds: classic rounds, like in Classic
-Paxos, and fast rounds.
-The fast rounds can be transformed into
-implementations of fast ACA rounds. Instead of sending their
-commands to a leader, as in Classic Paxos, the proposers of fast
-rounds directly broadcast their commands to all the acceptors. As long
-as commands are received in the same order by a learn quorum of
-acceptors or are received in different orders but commute, commands can be learned in two message-propagation delays instead of three in Classic Paxos.
-However, if non-commuting commands are received in different orders by
-several acceptors, the local history of the acceptors may not form a
-chain anymore and the rounds must abort. The fast rounds of Fast
-Paxos implement fast ACA rounds.
-
-Fast Paxos ACA rounds are specified using the same set of variables as for
-Classic Paxos, initialized to the same values.
-The next-state relation is composed of the following types of
-transitions, where the definitions of learn quorums, abort quorums,
-and of the $SafeAborts\left( R \right)$ operator are the ones of fast
-ACA rounds.
-\begin{compactitem}
-    \item A $Propose\left( p,c \right)$ transition, for a proposer $p$
-        and a command $c$, not guarded, inserts $c$ in the set
-        $proposed$ and broadcasts the message $\aseq{\quo{prop},c}$ to
-        all the acceptors.
-    \item An $Init\left( s, h \right)$ transition, for a switcher $s$
-        and a history $h$, not guarded, inserts $h$ in the set $inits$ and
-        broadcasts the message $\aseq{\quo{init},h}$ to all the acceptors.
-    \item A $WakeUp\left( a \right)$ transition, for an acceptor $a$,
-        is enabled when $status\left[ a \right]$ is ``idle'' and $a$
-        can receive an $\aseq{\quo{init},h}$ message. Its effect is to
-        receive the message, set $hist\left[ a \right]$ to $h$, and to
-        set $status\left[ a \right]$ to ``ready''.
-    \item An $Accept\left( a \right)$ transition, for an acceptor $a$,
-        is enabled when $status\left[ a \right]$ is ``ready'' and $a$
-        can receive a $\aseq{\quo{prop},c}$ message. The effect of the
-        action is to receive the message, to update $hist\left[ a
-        \right]$ to $hist\left[ a \right]\star c$, and to broadcast
-        the message $\aseq{\quo{ack},hist\left[ a \right]\star c}$ to
-        the learners.
-    \item The $RcvAck\left( l \right)$ transitions, for a learner $l$,
-        the $Learn\left( l, h \right)$ transitions, for a learner $l$
-        and a history $h$, the $Stop\left( s \right)$ transitions, for
-        a switcher $s$, $Stop\left( a \right)$ transitions, for an
-        acceptor $a$, the $RcvStopped\left( s \right)$ transitions,
-        for a switcher $s$, and the $Abort\left( s, h \right)$
-        transitions, for a switcher $s$ and a history $h$, which are
-        all the same as in Classic Paxos ACA rounds.
-\end{compactitem}
-
-As for Classic Paxos ACA rounds, Fast Paxos ACA rounds implement ACA
-under the refinement mapping consisting projecting the state of the
-round onto the variables $proposed$, $inits$, $aborts$, and, for every
-learner $l$, $learned\left[ l \right]$.
-
-\newpage
-
-\printbibliography
-
-\begin{appendix}
-
-  \label{sec:isaproofs}
-
-%  \includepdf[pages={1},pagecommand={\section{Mechanical Proof of the
-%  Composition Theorem in Isabelle/HOL}\label{sec:isaproofs}},scale=1]{../IsabellePDFS/podctheories.pdf}
-%  \includepdf[pages={2-},pagecommand={},scale=1]{../IsabellePDFS/podctheories.pdf}
-
-\end{appendix}
-
-\end{document}
-
-    We present a new abstraction, Abortable Generalized Consensus, which enables the
-    correct-by-construction composition of heterogeneous State-Machine
-    Replication algorithms.
-
-    integrating state of the art techniques to optimize executions.  ACA is the most complete approach to date enabling the
-    correct-by-construction composition of heterogeneous State-Machine
-    Replication algorithms. We provide two specifications satisfying
-    ACA and we illustrate their power by refining them to
-    obtain a versatile adaptive algorithm that combines five existing
-    algorithms into a single one. ACA is completely formalized and its
-    underlying composition theorem has been mechanically checked by an interactive
-    proof assistant.
-}
-
-\vspace{2cm}
-
-\begin{center}
-This paper is a regular submission.
-
-\vspace{1cm}
-This paper is a student paper.
-\end{center}
-
-\newpage
-
-\section{Introduction}
-
-The State-Machine Replication technique (abbreviated SMR) allows
-building reliable services on top of unreliable hardware. The basic idea is to 
-replicate a service over several servers, each executing the same
-sequence of (deterministic) commands. 
-Making SMR algorithms efficient has been a long-standing
-challenge and a multitude of algorithms have been proposed. Examples
-include Fast Paxos \cite{Lamport06FastPaxos}, Disk Paxos
-\cite{GafniLamport03DiskPaxos}, Chain Replication
-\cite{RenesseSchneider04ChainReplicationSupportingHighThroughputAvailability},
-Ring Paxos
-\cite{MarandiETAL10RingPaxosHighthroughputAtomicBroadcastProtocol},
-Egalitarian Paxos
-\cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments},
-Multi-Coordinated Paxos
-\cite{CamargosSchmidtPedone07MulticoordinatedPaxos}, Cheap Paxos
-\cite{LamportMassa04CheapPaxos}, Vertical Paxos
-\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication},
-Paxos-MIC
-\cite{HurfinMoiseNarzul11AdaptiveFastPaxosMakingQuickEverlasting},
-Mencius
-\cite{MaoJunqueiraMarzullo08MenciusBuildingEfficientReplicatedStateMachine},
-and Fast Mencius \cite{WeiETAL13FastMenciusMenciusLowCommitLatency}.
-
-Each of the existing algorithms is best suited to a particular range
-of environments, characterized, e.g., by client load, network
-topology, likelihood of faults, communication latency and bandwidth, processing
-power of nodes, availability of disk storage and memory.
-However, the environment faced by real systems evolves over time:
-hardware is upgraded, new data-centers are built, the popularity of a
-service changes, etc.  Consistent performance therefore comes with
-adaptation to change.  Unfortunately, existing SMR algorithms cannot easily be
-stopped and replaced by another algorithm on the fly. Moreover,
-devising a new algorithm that can change behavior dynamically is hard:
-dynamically switching from one SMR algorithm to another requires
-synchronizing notoriously intricate algorithms.
-
-We propose a new abstraction, \emph{Abortable Chain Agreement}
-(abbreviated ACA), which subsumes SMR and whose implementations,
-called \emph{ACA rounds}, can, by construction, pass the baton from
-one to another on the fly and efficiently. Moreover, optimizing a replicated
-service consisting of several ACA rounds becomes just a matter of devising a new
-ACA round, which can then be plugged-in on the fly: any new ACA round
-is compatible with every existing ACA round by construction.
-
-Abortable Chain Agreement is the \emph{most complete framework to date} enabling
-correct by construction composition of heterogeneous rounds.  
-ACA is completely formalized and its underlying composition theorem has been
-mechanically checked by an interactive proof assistant.  
-Compared to ACA,
-existing approaches
-\cite{BarNoyETAL87ShiftingGearsChangingAlgorithmsFlyTo,GuerraouiETAL10Next700BftProtocols,GuerraouiKuncakLosa12SpeculativeLinearizability}
-are limited. In particular, they do not allow the disorderly execution of
-commuting and read-only commands, an important practical optimization
-which is at the root of recent work on SMR like Generalized Consensus
-\cite{Lamport05GeneralizeConsensus} and Generalized Lattice Agreement
-\cite{FalerioETAL12GeneralizedLatticeAgreement}.
-
-\begin{comment}
-On top of the traditional roles of proposer, acceptor, and learners
-\cite{lamport2001paxos} used in consensus algorithms, ACA introduces
-the notion of \emph{switchers} and of \emph{scheduling policy}. The switchers
-have the task of aborting a running ACA round and initializing a new
-ACA round in a safe and live manner.  The \emph{scheduling policy}
-decides when to abort an ACA round and choses the next ACA round to
-execute. It must ensure that all switchers agree on the next ACA round
-to initialize.
-\end{comment}
-
-We observe that, because of the FLP impossibility result
-\cite{FischerLynchPaterson83ImpossibilityDistributedConsensusOneFaultyProcess},
-most existing SMR algorithms are structured in a sequence of abortable
-rounds which can easily be transformed into ACA rounds. To simplify
-this task even further, we propose two specifications, \emph{conservative
-ACA rounds} and \emph{fast ACA rounds}, which, as we show, can be concretized
-to obtain ACA versions of the rounds of five existing algorithms: 
-Classic Paxos, Fast Paxos, Chain Replication, Ring Paxos,
-and Multi-Coordinated Paxos. ACA guarantees that the composition of the five ACA
-rounds is correct by construction, forming a versatile adaptive replication
-algorithm.
-
-The rest of the paper is structured as follows.
-\Cref{sec:related} surveys recent related work. \Cref{sec:ca} presents Chain
-Agreement, which embodies the requirements for implementing SMR. \Cref{sec:aca}
-presents Abortable Chain Agreement and its optimizations to treat commuting
-and read-only commands efficiently. \Cref{sec:rounds} presents our specification
-of conservative and fast ACA rounds, which implement ACA. In
-\cref{sec:examples}, we show that the rounds of  
-Classic Paxos, Fast Paxos, Chain Replication, Ring Paxos,
-and Multi-Coordinated Paxos can easily be transformed into ACA implementations
-by using our specifications of conservative and fast ACA rounds.
-Finally, in \cref{sec:isaproofs}, we provide a complete formalization of ACA
-and a mechanical proof of the \emph{composition theorem} (\cref{thm:comp}),
-checked by the interactive proof assistant Isabelle/HOL \cite{NipkowPaulsonWenzel02IsabelleHOL}.
-
-\begin{comment}
-Abortable Chain Agreement builds on several recent advances in SMR
-algorithms: ACA allows acceptors to only eventually reach agreement,
-like in Generalized Paxos \cite{Lamport05GeneralizeConsensus},
-boosting performance when most commands commute; ACA allows fast
-execution of read commands like in Generalized Lattice Agreement
-\cite{FalerioETAL12GeneralizedLatticeAgreement}; ACA implementations
-can tolerate $f$ faults with only $f+1$ replicas and replace crashed replicas
-upon changing round, as in Vertical Paxos  \cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}.  
-\end{comment}
-
-\section{Related Work}
-\label{sec:related}
-
-SMR is often implemented using a sequence of instances of Consensus.
-There are numerous examples of consensus algorithms in the literature.
-However, Abortable Chain Agreement is based on Generalized Consensus. 
-In Generalized Consensus \cite{Lamport05GeneralizeConsensus}, the learners of a
-SMR algorithm repeatedly learn new values belonging to a \emph{C-struct set}. A
-C-struct can be seen as a data structure encoding  sequences of commands modulo
-the ordering of commands that commute. The advantage of Generalized
-Consensus over Consensus is that
-two servers which receive commuting commands in different orders can still reach
-agreement on C-structs without further synchronization. 
-Our Abortable Chain Agreement can also use C-structs, with the same advantages
-as Generalized Consensus. For simplicity, the presentation in this
-paper uses the more concrete notion of \emph{command history} (see
-\cref{sec:histories}).
-
-Generalized Lattice Agreement \cite{FalerioETAL12GeneralizedLatticeAgreement} allows more efficient implementations of SMR when
-commands are partitioned into read-only commands and write-only commands.
-Abortable Chain Agreement supports the same optimizations as Generalized Lattice Agreement.
-Notably, read-only commands are treated specially so as to optimize their
-executions (see \cref{sec:readonly}).
-
-Vertical Paxos algorithms 
-\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication} are SMR algorithms
-whose set of replica servers can be reconfigured without stopping to process new
-commands. Crashed servers can therefore be replaced by new servers, ensuring the
-long-term availability of a replicated service.
-Vertical Paxos algorithms use an external reconfiguration master, itself
-implemented as a replicated state machine, which determines the servers that should
-be removed from or added to the configuration.
-ACA shares some aspects of Vertical Paxos algorithms: the task of the
-reconfiguration master is similar to the task of the scheduling policy
-component of ACA (see \cref{sec:aca}).
-However, in ACA, changing round can be used not only to replace crashed servers
-but also to change the algorithm that the servers are running. 
-Changing algorithm is not possible in Vertical Paxos.
-In both Vertical Paxos and ACA, a reliable reconfiguration master, resp. a
-reliable scheduling policy, allows tolerating $f$ crashes with $f+1$ servers.
-
-The Abstract Framework \cite{GuerraouiETAL10Next700BftProtocols} proposes to
-build Byzantine Fault-Tolerant algorithms as the succession of abortable rounds
-called Abstract instances. Abstract instances must satisfy the Abstract
-correctness properties. By construction, the composition of any number of
-Abstract instances is a correct SMR implementation. However, the Abstract
-Framework uses totally ordered sequence of commands, making difficult the
-optimization of the execution of commuting and read-only commands that are
-possible with ACA.  Moreover, the Abstract Framework does not identify the
-crucial role of the scheduling policy for the long-term resilience of a service.
-Abortable Chain Agreement combines the ideas of the Abstract Framework,
-Generalized Paxos, Generalized Lattice Agreement, and Vertical Paxos in a
-unified abstraction.
-
-Finally, the Speculative Linearizability framework
-\cite{GuerraouiKuncakLosa12SpeculativeLinearizability} was a first
-attempt by the authors of the present paper at generalizing the
-Abstract Framework and mechanically proving its properties with the
-Isabelle/HOL interactive
-proofs assistant.  Like the Abstract Framework, the
-Speculative Linearizability framework is based on totally ordered
-sequences of commands, with the same drawback for the execution of
-commuting and read-only commands. Moreover, Abortable Chain Agreement is both
-conceptually simpler and subsumes the Speculative Linearizability
-property.  Like Speculative Linearizability, Abortable Chain Agreement
-and its main property, the composition theorem (\cref{thm:comp}), are
-formalized in Isabelle/HOL (see \cref{sec:isaproofs}).
-
-Abortable Chain Agreement is the most complete framework to date for building
-adaptive replication algorithms out of heterogeneous rounds, making state of the
-art optimization techniques available to algorithm designers, which was not
-possible before.
-
-\section{Chain Agreement}
-\label{sec:ca}
-
-We consider a set of asynchronous servers communicating by message passing in an
-asynchronous network. Servers are not necessarily sequential processes. We
-describe a behavior of the system using an infinite sequence of states, where a
-state is a function from variable names to values (i.e. a valuation
-of variables).  We specify sets of allowed behaviors using an initial predicate,
-a next-state relation (using unprimed and primed variables) and a fairness
-condition, all in the style of TLA+ \cite{Lamport02SpecifyingSystems}. However,
-for succinctness, our specification are rather informal, eluding non-essential
-details.
-
-Our specifications all contain by default a state variable $crashed$ and
-$Crash\left( x \right)$ transition which is always enabled. A server $x$ faithfully
-executes the algorithm that is assigned to it until its identifier is added to
-the variable $crashed$ by a $Crash\left( x \right)$ transitions.  When the identifier of a
-server is in the set $crashed$, all its actions are disabled.  In an execution,
-a server that is never added to the set $crashed$ is said correct.
-
-We consider a computing service exposing a set of commands $C$. The
-service has a sequential specification consisting of an initial state
-and a deterministic transition relation in which every command
-atomically updates the state of the service and produces an output. We
-assume that duplicate commands have no effect on the state of the
-service and return the same output as their first occurrence.
-
-In a traditional SMR algorithm, the servers can be seen has playing one or
-several of three roles: \emph{proposers}, which propose new commands to execute,
-\emph{acceptors}, which accept and order commands in a sequence, and
-\emph{learners}, which learn about the growing sequence of commands to
-execute \cite{lamport2001paxos}. Often, but not necessarily, every replica server plays the
-three roles of proposer, acceptor, and learner: clients of the service
-send their commands to some replica servers; replica servers submit
-the received commands to the SMR protocol in the role of proposers;
-replica servers respond to commands by executing the commands learned
-in the role of learners; replica servers participate in command
-ordering in the role of acceptors.  An SMR algorithm guarantees that
-the replicas learn the same sequences of commands, possibly up to the
-reordering of commuting commands. Therefore, every learner can respond to client
-requests using its last learned sequence of commands, and clients will receive
-the same responses as the ones produces by a centralized, non-replicated,
-service.
-
-Chain Agreement precisely specifies the allowed behaviors of the roles of \emph{proposers}
-and \emph{learners}, abstracting over the behavior of acceptors.
-
-\begin{comment}
-We suppose that each server can play a number of \emph{roles}.
-Chain Agreement specifies the allowed behaviors the roles of \emph{proposers}
-and \emph{learners}. Proposers receive commands from the clients of the service
-and propose them for execution by the replicated service. Each learner learns about
-a growing sequence of commands. Thanks to the sequential specification of the
-service, a sequence of commands determines an output for each the included commands. 
-In a traditional SMR algorithm, servers can also play the role of
-\emph{acceptors}, whose task is to collectively order commands and let the
-learners know about the latest sequence of ordered commands.
-
-The point of Chain Agreement is to implement a replicated service by having every learner learn about a growing sequence of commands, which
-represent an execution of the sequential specification of the service.
-Crucially, at all times, the sequences of commands learned by the
-different learners must form a chain in the prefix order on sequences.
-Moreover, for every command $c$ proposed by a correct proposer, every
-correct learner must eventually learn a sequence containing $c$.  Once
-a learner learns about a sequence containing a command $c$, it can
-determine the output produced by $c$, using the sequential
-specification of the service, and respond to the client that issued
-the command $c$.
-\end{comment}
-
-To specify Chain Agreement, we need the notion of ordering of sequences and of
-maximum sequence.  The prefix order on sequences is denoted by the operator
-$\leq$ ($s_1 \leq s_2$ means that $s_1$ is a prefix of $s_2$). 
-%Using the operator $\leq$, we define $\geq$, $<$, and $>$ as expected. 
-Note that $\leq$ is a partial order on sequences. The maximum over a set of
-sequences $S$ is denoted by $Max\left( S \right)$ and $Max\left( \aset{}
-\right)$ is the empty sequence.
-
-We specify Chain Agreement using the state variables $crashed$, $proposed$, containing the proposed commands,
-and, for every learner $l$,  $learned\left[ l \right]$, a set of
-sequences of commands. Initially, $crashed$, $proposed$ and, for every learner $l$,
-$learned\left[ l \right]$ are empty sets. 
-Remember that any transition involving a server $x$ is always disabled when $x\in crashed$.
-
-The next-state relation consists of three types of transitions:
-$Crash\left( x \right)$ transitions; $Propose\left( l, c
-\right)$ transitions, for a learner $l$ and a command $c$, in which
-$c$ is inserted in the set $proposed$; $Learn\left( l,cs \right)$
-transitions, for a learner $l$ and a sequence of commands $cs$, in
-which $cs$ is inserted in $learned\left[ l \right]$. 
-A $Learn\left( l, cs \right)$ transition is guarded by the following
-conditions:
-\begin{compactitem}
-    \item[\textbf{Agreement}:] The elements of the set $\aset{learned'\left[ l
-      \right] : l \in Learners}$, the learned sequences after the transition,
-      form a chain in the partial order $\leq$.
-    \item[\textbf{Validity}:] Every learned sequence contains only
-        proposed commands.  
-    \item[\textbf{Irrevocability}:] The new learned sequence
-        $cs$ is a strict extension of $Max\left(learned\left[ l
-        \right]\right)$. 
-        %\\ $cs > Max\left(learned\left[ l \right]\right)$.
-\end{compactitem}
-Finally, we add, for every learner $l$, one weak fairness condition
-\cite{Lamport02SpecifyingSystems} on the transitions $Learn\left( l,
-cs \right)$. The weak fairness conditions ensure that every correct
-learner eventually learns about every proposed command.
-
-Thanks to the properties of Chain Agreement, the clients, observing
-their input commands and the received responses, see a
-\emph{linearizable} execution
-\cite{HerlihyWing90LinearizabilityCorrectnessConditionConcurrentObjects}.
-Moreover, the service remains available as long as at least one
-proposer and one learner are correct.
-
-\section{Abortable Chain Agreement}
-\label{sec:aca}
-
-We now consider implementing Chain Agreement as a sequence of \emph{Abortable Chain
-Agreement rounds}, where each round makes progress ordering requests or aborts
-its execution and passes the baton to the next round in the sequence.
-An algorithm composed of a sequence of ACA rounds is called an \emph{ACA
-algorithm}.
-
-On top of the traditional roles of proposer, acceptor, and learner, ACA
-introduces the new role of \emph{switcher} and a new component, the
-\emph{scheduling policy}. In ACA, the assignment of the roles of proposer,
-acceptor, learner, and switcher can change in each rounds, allowing for
-reconfiguration. 
-In normal circumstances, the proposers, acceptors,
-and learners of an ACA round work together to implement Chain Agreement.
-However, if the environment becomes unfavorable to a running ACA round, the
-scheduling policy can instruct the round to abort its execution and pass the
-baton, using the switchers, to a new round.
-
-A switcher of an aborting round determines, from the state of the
-acceptors, an \emph{abort sequence} that it transfers to one or several
-switchers of the new round.  Upon receiving an abort sequence, a switcher of the
-new round initializes the acceptors of the new round, which can then
-proceed implementing SMR, until the new round itself aborts and the
-process is repeated. 
-We assume that in the first round, which has no previous round, every 
-switchers receives the empty abort sequence.
-
-The scheduling policy determines when to abort an ACA round and which new ACA
-round should take over. The scheduling policy must inform the switchers about the
-identity of the next round and must guarantee that all switchers receive the
-same information.  In this paper we abstract over the scheduling policy, assuming
-that a round that aborts passes the baton a nondeterministically chosen new
-round. A reliable scheduling-policy component can realistically be implemented
-as a replicated state machine running on a large number of servers otherwise
-dedicated to other tasks
-\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}. We will see
- in \cref{sec:rounds} that, as in Vertical Paxos, a reliable scheduling policy component allows
-tolerating $f$ faults with $f+1$ replicas, while maintaining the long-term
-resilience of the service by replacing faulty servers upon round change.
-
-The clients of an ACA algorithm can send their commands to the proposers of any round (in
-practice, clients send their commands to proposers of the largest
-active round) and servers can use sequences learned in any round
-to determine outputs.
-
-ACA specifies the allowed behaviors of the proposers, learners, and switchers
-\emph{of a single round}. Similarly to
-Chain Agreement, we use the state variables $crashed$, $proposed$,
-and, for every learner $l$, $learned\left[ l \right]$.  Additionally,
-we use the state variables $inits$, containing the abort sequences
-received from the previous round, and $aborts$, containing the abort
-sequences transfered to the next round.
-Two consecutive ACA rounds are composed by letting the $aborts$ variable of the
-first round be the $inits$ variable of the second round.
-
-Initially, $crashed$, $proposed$, for every
-learner $l$, $learned\left[ l \right]$, $inits$, and $aborts$ are all empty sets.
-The next-state relation consists of five types of transitions. The
-first three, $Crash\left( x \right)$, $Propose\left( l, c
-\right)$, and $Learn\left( l, cs \right)$, for a learner, proposer, or
-switcher $x$, for a learner $l$, and for a sequence of commands
-$cs$, have the same effect as in Chain Agreement.
-ACA introduces two new transitions, $Init\left(
-s, cs \right)$ and $Abort\left( s, cs \right)$, for a switcher $s$ and a sequence of
-commands $cs$. The transition $Init\left( s, cs \right)$ adds $cs$ to the set $inits$, and models a switcher receiving the
-abort sequence $cs$ from the preceding round.
-The transition $Abort\left( s, cs \right)$ adds $cs$ to the set $aborts$ and models a switcher transferring the abort sequence $cs$ to a switcher of the next round.
-All the transition are guarded by the following conditions:
-\begin{compactitem}
-    \item[\textbf{Agreement}:] The elements of the set $\aset{learned'\left[ l
-      \right] : l \in Learners}$ form a chain in the partial order.
-    \item[\textbf{Validity}:] Every abort sequence or learned sequence
-        is of the form $h \star cs$, where $cs$ contains only commands
-        of the set $proposed$ and $h$ is the greatest lower bound of a
-        nonempty subset of $inits$.
-    \item[\textbf{Initialization}:] If $inits$ is empty, then no
-        sequence can be learned and no switcher can abort.
-    \item[\textbf{Irrevocability}:] In the case of a $Learn\left( l,cs
-        \right)$ transition, the new learned sequence $cs$ is a strict
-        extension of $Max\left(learned\left[ l \right]\right)$.
-        %, $cs
-        %\geq Max\left(learned\left[ l \right]\right)$.
-    \item[\textbf{Safe Abort}:] Every abort sequence is larger in the
-        partial order $\leq$ than every learned sequence.
-\end{compactitem}
-Note that Agreement and Irrevocability are the same as in Chain
-Agreement.
-
-Finally, for every learner $l$, we add a weak fairness condition on the
-following state transition: 
-\begin{equation}
-    \exists  s \in Switchers, cs \in Sequence\left( C \right) : Learn\left( l, cs \right) \vee Abort\left( s,cs \right)
-\end{equation}
-The weak fairness conditions ensure that when a correct learner has not
-learned a sequence containing all commands proposed in the round, then
-it eventually learns a new command or, if there is at least one correct switcher, at least one switch value is transfered to the next round. 
-Therefore, if a client has submitted a command to a proposer, eventually,
-either the client gets a response from the current round, or
-the next round is initialized with an abort sequence and the
-command can be resubmitted to the next round.
-
-\subsection{The Composition Theorem}
-
-Consider the succession of $n$ rounds numbered $1$ to $n$.  To
-distinguish the variables of each round, we annotate the local
-variables of each round with the number of the round. For example, the
-variables of the seventh round are written $proposed_7$, for every
-learner $l$, $learned_7\left[ l \right]$, $inits_7$, and $aborts_7$.
-
-Moreover, we assume that a command, which is proposed to a round and that
-is not learned before the first abort sequence is transfered to the
-next round, is proposed anew to the next round.
-
-Define the state variable $proposed$ as the union over all rounds $i$
-of the $proposed_i$ variables and, for every learner $l$, 
-define the $learned\left[ l \right]$ variable as the union over all
-rounds $i$ of the $learned_i\left[ l \right]$ variables.
-
-\begin{thm}[Composition Theorem]
-    \label{thm:comp}
-    In every execution of a succession of ACA rounds, the sequence of valuations
-    of the variable $proposed$ and, for every learner $l$, of the variables
-    $learned\left[ l \right]$ satisfies Chain Agreement.
-\end{thm}
-\Cref{thm:comp} can be shown by induction on the number of rounds,
-using the following lemma.
-\begin{lem}
-  \label{lem:comp}
-    Consider two rounds numbered $i$ and $i+1$. In every execution of a
-    succession of ACA rounds, the sequence of valuations of the variables $inits
-    = inits_i$, $propose = proposed_i \cup propose_{i+1}$, for every learner
-    $l$, $learned\left[ l \right] = learned_i\left[ l \right] \cup
-    learned_{i+1}\left[ l \right]$and $aborts = aborts_{i+1}$ satisfies ACA.
-\end{lem}
-\begin{proof}
-  A mechanically-checked proof in Isabelle/HOL appears in \cref{sec:isaproofs}.
-  The proofs uses a version of the specification of ACA that has a more
-  executable form that presented in this section. Moreover,
-  instead of sequences of commands, the Isabelle/HOL specification uses the more
-  general notion of command history, presented in the next subsection.
-  The Isabelle/HOL formalization of ACA uses the theory of I/O automata, which
-  is also formalized in the Isabelle/HOL. The proof uses a refinement
-  mapping from the composition of two consecutive ACA instances to a single ACA
-  instance.
-\end{proof}
-
-\Cref{thm:comp} guarantees that the successive composition of any
-number of ACA rounds implements Chain Agreement.
-Therefore, with ACA, one can easily build replicated services that
-dynamically adapt to the current operating conditions of the system: a
-replicated service has several types of rounds at its disposal, each being
-tailored to particular conditions. Each time the conditions change,
-the current ACA round aborts and passes the baton to a more 
-appropriate type of round, as determined by the scheduling policy. Moreover, a new type
-of ACA round can always be devised after the system is deployed and added on the fly to
-the existing types of rounds.
-
-We now describe two important practical optimizations of ACA.
-
-\subsection{Histories}
-\label{sec:histories}
-
-ACA can be implemented more efficiently by substituting \emph{command histories} for sequences of commands, as in Generalized Consensus \cite{Lamport05GeneralizeConsensus}.
-
-We say that two commands $c_1$ and $c_2$ \emph{commute} when, in every sequential
-execution of the service, reversing the order of $c_1$ and $c_2$ does
-not change any of the outputs, including those given to $c_1$ and
-$c_2$.
-
-A command history, abbreviated history, is a data structure that
-represents a set of sequences which are the same modulo the reordering
-of commuting commands, and in which duplicate commands do not appear.
-All the sequences represented by a given history contain the same
-commands and determine the same outputs for each command.  Therefore,
-to implement a replicated service, it is sufficient that learners
-learn command histories instead of sequences.  For a precise
-definition of command histories, we refer the reader to
-\cite{Lamport05GeneralizeConsensus}. Our Isabelle/HOL formalization
-(\cref{sec:isaproofs}) abstracts
-over the concrete representation of histories, assuming only their necessary
-properties. Command histories were first
-introduced in trace theory \cite{Mazurkiewicz84Semantics}.
-
-A command history can be built by appending a sequence of commands
-$cs$ to a history $h$, obtain the history $h \star cs$.  The major
-advantage that histories have over sequences is that if a sequence
-$cs_1$ is obtained by reordering the commuting commands of $cs_2$,
-then $h \star cs_1 = h \star cs_2$. Therefore, in  implementations of
-ACA, servers can agree on a history even though
-they chose different orders for some commuting commands.
-
-Command histories share the four properties of sequences that are
-necessary and sufficient to substitute histories for sequences in the
-definition of Chain Agreement and ACA.
-\begin{compactitem}
-    \item[\textbf{Distinctness}:] No duplicate command appears in a history. 
-    \item[\textbf{Ordering}:] Command histories form a partial order:
-        $h_1 \leq h_2$ when there exists a sequence $s$ such that $h_2
-        = h_1 \star s$.
-    \item[\textbf{GLB}:] Every finite set of command histories has a
-        \emph{greatest lower bound}.
-%    \item[\textbf{LUB}:] Every finite set of compatible command
-%        histories has a \emph{least upper bound}, where a set of
-%        command histories is compatible when they have a common upper
-%        bound.
-    \item[\textbf{Consistency}:] If $h_1 \leq h_2 \leq h_3$ and $h_3 =
-        h_1 \star s$ then there exists $s'$ and $s''$ where every
-        request of $s'$ and $s''$ appears in $s$ and such that $h_2 =
-        h_1 \star s'$ and $h_3 = h_2 \star s''$.
-\end{compactitem}
-The optimized version of ACA is obtained by
-simply substituting histories for sequences in the definition of ACA.
-The Ordering, GLB, and Consistency properties of histories ensure that
-\cref{thm:comp} still holds: the Isabelle/HOL proof of \cref{thm:comp} (see
-\cref{sec:isaproofs}) relies only on the four properties above.
-
-\subsection{Optimizing the Execution of Read-Only Commands}
-\label{sec:readonly}
-
-Chain Agreement can be further optimized by treating 
-read-only commands specially, as in Generalized Lattice Agreement
-\cite{FalerioETAL12GeneralizedLatticeAgreement}.
-
-We consider the case in which a set of servers each play the four
-roles of proposer, acceptor, learner, and switcher.
-Clients sends their commands to any server, which proposes them as a
-proposer, and records the command as pending.
-%When a server learns a history that contains one of its pending
-%commands, it uses the history to determine the response to the pending
-%command and forwards the response to the appropriate client.
-
-As observed in \cite{FalerioETAL12GeneralizedLatticeAgreement}, read-only
-commands can be executed faster by directly reading from the last learned
-history, instead of proposing the read commands and waiting for it to appear in
-a future learned history.  Because learned histories form a chain, all the reads
-return consistent values and this approach yields a \emph{serializable}
-implementation of the service.
-
-However linearizability is not satisfied if reads are served directly
-from the last learned history: a lagging replica server could respond
-to a read using a history which does not contain all completed update
-commands. To make the executions linearizable without resorting to
-proposing read-only commands as normal commands, we can employ the same
-technique as in \cite{FalerioETAL12GeneralizedLatticeAgreement}: upon
-receiving a read command, a server proposes a special no-op command
-that commutes with all other commands; the server then responds to the
-read command using the first learned history containing the no-op
-command. This ensures that all the commands that had completed before
-the read command was issued are present in the history.  Moreover,
-since no-op commands commute with all other commands, they can be
-treated more efficiently than regular read commands by the acceptors,
-like, for example, in a fast round of Fast Paxos.
-
-\section{Fast and Conservative Rounds}
-\label{sec:rounds}
-
-Most existing SMR algorithms can easily be obtained as the
-composition of one or more ACA rounds.
-
-The FLP impossibility result \cite{FischerLynchPaterson83ImpossibilityDistributedConsensusOneFaultyProcess}
-implies that no algorithm can implement SMR to the letter in practice.
-Instead, SMR algorithms guarantee safety (i.e. replica execute the
-same sequence of commands) but may not be able to make progress (i.e.
-order new commands). To make progress as likely as possible, most SMR
-algorithms are structured in consecutive rounds (also called ballots,
-views, or epochs). In each round, the algorithm tries to make progress
-but cannot be guaranteed to succeed. When the algorithm detects that
-progress is no more possible, a new round is started.  The new round
-is initialized so as to preserve safety, usually by one of the
-acceptors. For example, in Paxos, the leader of a newly started ballot
-numbered $i$ determines a value that is \emph{safe at $i$} and
-proposes it as first value to the other acceptors of the ballot $i$.
-In existing SMR algorithms, new rounds are initialized ad-hoc, which
-prevents mixing rounds of two different algorithms.  Instead,
-ACA is an abstraction of rounds which establishes a common interface and contract for switching from
-one round to the next, enabling switching from any ACA round to any other.
-
-To simplify the task of expressing existing algorithms as the composition of one
-or more ACA rounds, we propose two specifications, \emph{conservative ACA
-rounds} and
-\emph{fast ACA rounds}, which can easily be concretized to obtain executable rounds.
-We show that ACA versions of Classic Paxos, Multi-Coordinated Paxos, Chain
-Replication, and Ring Paxos can be obtained by concretizing the specification of
-conservative ACA rounds, and that ACA versions of the fast rounds of Fast Paxos can be obtained by concretizing the specification of fast ACA rounds.
-
-Fast and conservative ACA rounds are two specifications, at an
-intermediate level of abstraction, which satisfy ACA. Both
-specifications specify how the state of the acceptors should be
-updated but abstract over the particular communication topology used
-to implement the state accesses: the specifications use a global state
-which is accessible by all servers in all roles. One can refine fast
-or conservative rounds by implementing the state accesses and coordination
-using the network, obtaining a concrete algorithm.
-
-Conservative and fast ACA rounds are specified using the state variables $inits$, $proposed$, for every learner $l$, $learned\left[ l \right]$
-and $aborts$, as in ACA, and, for every acceptor $a$, two new variables
-$status\left[ a \right]$, which can be either ``idle'', ``ready'', or
-``stopped'' and $hist\left[ a \right]$, the local history of the
-acceptor $a$, which can be either a history or the special value $None$.
-Initially, $inits$, $aborts$, and, for every learner
-$l$, $learned\left[ l \right]$ are empty sets, and, for every acceptor $a$,
-$status\left[ a \right]$  is ``idle'' and $hist\left[ a \right]$ is the special
-value $None$.
-
-The next-state relations of conservative and fast rounds are composed
-of the following types of transitions (and the default $Crash\left( x
-\right)$ transition).
-\begin{compactitem}
-    \item A $Propose\left( p,c \right)$ transition, for a proposer $p$
-        and a command $c$, not guarded, inserts $c$ in the set
-        $proposed$. 
-    \item An $Init\left( s, h \right)$ transition, for a switcher $s$
-        and a history $h$, not guarded, inserts $h$ in the set $inits$.
-    \item A $WakeUp\left( a \right)$ transition, for an acceptor $a$,
-        is enabled when $status\left[ a \right]$ is ``idle''. Its
-        effect is to set $hist\left[ a \right]$ to a history found in
-        $inits$ and to set $status\left[ a \right]$ to ``ready''.
-    \item An $Accept\left( a \right)$ transition, for an acceptor $a$,
-        is enabled when $status\left[ a \right]$ is ``ready'', and
-        updates $hist\left[ a \right]$ to $hist\left[ a \right]\star
-        c$, where $c$ is a proposed command. In conservative rounds,
-        the transition is additionally guarded by the condition that
-        the resulting set of local acceptor histories,
-        $\aset{hist'\left[ a \right] : a \in Acceptors}$, forms a chain
-        in the partial order $\leq$ (this is typically implemented with a
-        leader).
-    \item A $Learn\left( l, h \right)$ transition, for a learner $l$
-      and a history $h$, enabled when $h > learned\left[ l \right]$ and is
-      smaller than the GLB of the histories of a learn quorum of acceptors. The effect of
-        the transition is to set $learned\left[ l \right]$ to $h$.
-    \item A $Stop\left( a \right)$ transition, for an acceptor $a$,
-        enabled when $status\left[ a \right]$ is ``ready'', sets
-        $status\left[ a \right]$ to ``stopped'' (note that
-        $status\left[ a \right] = $ ``stopped'' prevents any further
-        ``accept'' transition of acceptor $a$).
-%    \item REALLY NEEDED? A $TimeOut\left( a \right)$ transition, for an acceptor $a$,
-%        is enabled when $status\left[ a \right]$ is ``idle'' and sets 
-%        $status\left[ a \right]$ to ``stopped'' and $hist\left[ a
-%        \right]$ to the special value $None$. 
-    \item An $Abort\left( s, h \right)$ transition, for a switcher $s$
-        and a history $h$, is enabled when there exists a abort quorum
-        $R$ such that, for every member $a$ of $R$, $status\left[ a
-        \right]$ is ``stopped'' and $h \in SafeAborts\left( R
-        \right)$. Its effect is to insert  $h$ in the set $aborts$. 
-\end{compactitem}
-
-On top of the additional guard on the $Accept$ transition present only
-in conservative rounds, conservative and fast ACA rounds differ in
-their definition of learn quorums, abort quorums, and in their
-definition of the $SafeAborts\left( R \right)$ operator.
-To concisely define read and learn quorums, we define 
-\begin{equation}
-    Q\left( f \right) = \aset{as \subseteq Acceptors :
-        Card\left(as\right) \geq \left\lfloor f\times Card\left(
-    Acceptors \right)\right\rfloor + 1}.
-\end{equation}
-
-In conservative ACA rounds, abort and learn quorums are defined such
-that the intersection between two learn quorums is nonempty and the
-intersection between a learn quorum and an abort quorum is nonempty.
-The definition leads to two interesting cases: $LearnQ = AbortQ =
-Q\left( 1/2 \right)$; $LearnQ = \aset{Acceptors}$ and $AbortQ =
-\aset{\aset{a} : a \in Acceptors}$. The operator $SafeAborts\left( R
-\right)$ is defined as the singleton set containing the maximum
-element over the local histories of the acceptors in $R$ which are not
-$None$. Note that the maximum exists because conservative rounds
-guarantee that the histories of the acceptors form a chain.
-
-In fast ACA rounds, abort and learn quorums are defined such that
-the intersection between any two learn quorums is nonempty and the
-intersection between any two learn quorums and an abort quorum is
-nonempty.  The definition leads to three interesting cases: $LearnQ =
-AbortQ = Q\left( 2/3 \right)$; $LearnQ = Q\left( 1/2 \right)$ and
-$AbortQ = Q\left( 3/4 \right)$; $LearnQ = \aset{Acceptors}$ and
-$AbortQ = \aset{\aset{a : a \in Acceptors}}$.
-The operator $SafeAborts\left( R \right)$ is defined as the set of
-maximal elements over the set $\aset{GLB\left( \aset{hist\left[ a
-\right] : a \in R \cap L \wedge hist\left[ a \right]\neq None} \right)
-: L \in LearnQ}$.
-
-The cases in which abort quorums are singleton sets, in both
-conservative and fast rounds, allow tolerating
-$f$ faults with $f+1$ servers: a round can abort as long as one
-acceptor is correct and a new round, possibly using new correct
-servers, can take over. Note that changing round cannot happen
-without the scheduling policy instructing the switchers about the
-identity of the next round, and thus tolerating $f$ faults with
-$f+1$ servers relies on the assumption of a reliable scheduling policy.
-
-Define, for every learner $l$, the history variable $\overline{learned}\left[ l \right]$ which is initialized to the empty set and which is
-updated each time a new local history is accepted by the learner $l$ by
-inserting the new local history in $\overline{learned}\left[ l \right]$ (see \cite{AbadiLamport91ExistenceRefinementMappings} for the
-definition of a history variable).
-\begin{thm}
-  \label{thm:rounds}
-    Both conservative and fast ACA rounds implement ACA under the refinement
-    mapping that consists in substituting $\overline{learned\left[ l \right]}$
-    for $learned\left[ l \right]$.
-\end{thm}
-
-\section{Examples of Concrete ACA Algorithms}
-\label{sec:examples}
-
-In this section we show how to modify the rounds of Classic Paxos,
-Multi-Coordinated Paxos, Chain Replication, Ring Paxos, and Fast Paxos to obtain
-ACA rounds. \Cref{thm:comp} then guarantee that all those different types of
-rounds can be used in the same algorithm to yield an implementation of Chain
-Agreement.
-
-\subsection{Classic Paxos}
-
-The rounds of the Classic Paxos algorithm can be modified so as to
-obtain an implementation of conservative ACA rounds.  We define
-Classic Paxos ACA rounds by specializing the specification of
-conservative ACA rounds, giving a concrete protocol for disseminating
-the proposed commands to the acceptors and enforcing that the learned
-histories form a chain.
-
-Each Classic Paxos round has a unique leader among the acceptors which
-centrally orders the proposed commands, which are accepted by the
-other acceptors in the order determined by the leader.  This protocol
-ensures that every local acceptor history is a prefix of the local
-history of the leader, therefore guaranteeing that the local acceptor histories
-forming a chain.
-
-Acceptor broadcast each newly accepted history to the learners, which
-learn a history $h$ when they receive enough extensions of $h$ from the
-acceptors.  When instructed by the scheduling policy to abort, a
-switcher sends a stop message to the acceptors, which irrevocably stop
-updating their local history and send back to the switcher an
-acknowledgement containing their local history. A switcher
-determines an abort history once it has received acknowledgments from
-an abort quorum of acceptors $R$, using the operator $SafeAborts\left(
-R \right)$.
-
-The specification of Classic Paxos ACA rounds uses the variables of
-the specification of conservative ACA rounds, namely the variables
-$proposed$, $inits$, $aborts$, the arrays $learned\left[ l \right]$,
-$status\left[ a \right]$, and $hist\left[ a \right]$, and two
-additional arrays $accepted\left[ l \right]\left[ a \right]$ and
-$stop\left[ s \right]\left[ a \right]$, where $l$ is a learner, $a$ is
-an acceptor, and $s$ is a switcher.
-
-There is also a network component that we do not describe explicitly.
-However, the network allows any servers to communicate by sending and
-receiving messages.
-
-Initially, the variables $proposed$, $inits$, $aborts$, and, for every
-learner $l$, $learned\left[ l \right]$ are empty sets. For every
-acceptor $a$, $status\left[ a \right]$ is ``idle'' and $hist\left[ a
-\right]$ is the special value $None$. For every learner $l$ and acceptor $a$,
-$accepted\left[ l \right]\left[ a \right]$ is the special value
-$None$. For every switcher $s$ and acceptor $a$, $stopped\left[ s
-\right]\left[ a \right]$ is set to $false$.  There are initially no
-messages in the network.
-
-The next-state relation is composed of the following types of
-transitions, where the definitions of learn quorums, abort quorums,
-and of the $SafeAborts\left( R \right)$ operator are the ones of conservative
-ACA rounds.
-\begin{compactitem}
-    \item A $Propose\left( p,c \right)$ transition, for a proposer $p$
-        and a command $c$, not guarded, inserts $c$ in the set
-        $proposed$ and sends the message $\aseq{\quo{prop},c}$ to
-        the leader.
-    \item An $Init\left( s, h \right)$ transition, for a switcher $s$
-        and a history $h$, not guarded, inserts $h$ in the set $inits$ and
-        sends the message $\aseq{\quo{init},h}$ to the leader.
-    \item A $WakeUp\left( leader \right)$ transition, enabled when
-        $status\left[ leader \right]$ is ``idle'' and the leader can
-        receive a $\aseq{\quo{init},h}$ message. Its effect is to
-        receive the message, set $hist\left[ leader \right]$ to $h$,
-        to set $status\left[ leader \right]$ to ``ready'', and to
-        broadcast the message $\aseq{\quo{leader-accept},h}$ to all
-        other acceptors.
-    \item An $Accept\left( leader \right)$ transition, enabled when
-        $status\left[ leader \right]$ is ``ready'' and the leader can
-        receive a $\aseq{\quo{prop},c}$ message. The effect of the
-        action is to receive the message, to update $hist\left[ leader
-        \right]$ to $hist\left[ leader \right]\star c$, and to
-        broadcast the message $\aseq{\quo{ack},hist\left[ leader
-        \right]\star c}$ to the learners, and to broadcast the message
-        $\aseq{\quo{leader-accept},hist\left[ leader \right]\star c}$
-        to all the other acceptors.
-    \item An $Accept\left( a \right)$ transition, for an acceptor $a$
-        which is not the leader, is enabled when $status\left[ a
-        \right]$ is ``ready'' and $a$ can receive a
-        $\aseq{\quo{leader-accept},h}$ message where $h > hist\left[ a
-        \right]$. The effect of the action is to receive the message,
-        to update $hist\left[ a \right]$ to $h$, and to broadcast the
-        message $\aseq{\quo{ack},hist\left[ a \right]\star c}$ to the
-        learners.
-    \item A $RcvAck\left( l \right)$ transition, for a learner $l$, is
-        enabled when $l$ can receive a message $\aseq{\quo{ack},h}$
-        from an acceptor $a$. Its effect is to receive the message and
-        to set $accepted\left[ l \right]\left[ a \right]$ to $h$.
-    \item A $Learn\left( l, h \right)$ transition, for a learner $l$
-        and a history $h$, is enabled when $h > Max\left(
-        learned\left[ l \right] \right)$ and there is a learn quorum
-        $Q$ of acceptors such that $h = GLB\left( \aset{accepted\left[
-        l \right]\left[ a \right] : a \in Q }\right)$. Its effect is
-        to insert $h$ in $learned\left[ l \right]$.
-    \item A $Stop\left( s \right)$ transition, for a switcher $s$,
-        always enabled, broadcasts a $\aseq{\quo{stop}}$ message to all
-        the acceptors.  This transition models a switcher being
-        instructed by the scheduling policy to abort the current round.
-    \item A $Stop\left( a \right)$ transition, for an acceptor $a$,
-        enabled when $a$ can receive a $\aseq{\quo{stop}}$ message
-        from a switcher $s$. Its effect is to receive the message, set
-        $status\left[ a \right]$ to ``stopped'', and to send the
-        message $\aseq{\quo{stopped},hist\left[ a \right]}$ to the
-        switcher $s$.
-    \item A $RcvStopped\left( s \right)$ transition, for a switcher
-        $s$, enabled when $s$ can receive a $\aseq{\quo{stopped},h}$
-        message from an acceptor $a$. Its effect is to receive the
-        message and to set $stopped\left[ s \right]\left[ a \right]$
-        to $h$.
-    \item An $Abort\left( s, h \right)$ transition, for a switcher $s$
-        and a history $h$, is enabled when there exists an abort quorum
-        $R$ such that, for every member $a$ of $R$, $stopped\left[ s
-        \right]\left[ a \right]$ is not $false$ and $h \in
-        SafeAborts\left( R \right)$, where $SafeAborts\left( R
-        \right)$ is computed using the array $stopped\left[ a
-        \right]$. Its effect is to insert  $h$ in the set $aborts$. 
-\end{compactitem}
-Compared to the rounds of the original Classic Paxos, Classic Paxos ACA
-rounds add an extra level of indirection when changing round: in the
-original rounds, the leader of the new round directly queries the
-acceptors of the preceeding round, whereas in ACA rounds, the
-switchers query the acceptors and then send their abort histories to
-the leader. However, changing round is an infrequent event, therefore
-this difference should not have much practical impact on performance.
-\begin{thm}
-  Classic Paxos ACA rounds satisfy Abortable Chain Agreement.
-\end{thm}
-\begin{proof}
-  A Classic Paxos ACA round implements the specification of Abortable
-  Chain Agreement under the refinement mapping consisting in
-  projecting the state of the round onto the variables $proposed$,
-  $inits$, $aborts$, and, for every learner $l$, $learned\left[ l
-  \right]$.
-\end{proof}
-
-\subsection{Multi-Coordinated Paxos, Chain Replication, and Ring Paxos.}
-
-Multi-Coordinated Paxos, Chain Replication, and Ring Paxos rounds have a similar
-structure as Classic Paxos rounds but differ in the way the leader is
-implemented and in the way that the histories of the leader are disseminated to
-the other acceptors.
-
-Multi-Coordinated Paxos avoids depending on a unique leader by using a
-distributed leader implementation that can make progress as long as a
-strict majority of the acceptors are correct, trading off resilience
-for a higher time complexity. Conservative ACA
-rounds can be implemented like with Classic Paxos, replacing the
-leader by the distributed leader of Multi-Coordinated Paxos.
-
-In Chain Replication, acceptors are arranged in a chain whose head is
-the leader of the round. As in Classic Paxos, abort histories and
-proposed commands are sent to the leader. However, instead of the
-leader directly sending its local history to all the other acceptors,
-the local history of the leader is
-forwarded along the chain from one acceptor to the other until it
-reaches the last acceptor, which sends the history to the learners. 
-By spreading the load more evenly on the all the servers, Chain Replication
-can achieve higher throughput that Classic Paxos.
-Conservative ACA rounds can be implemented like with Classic Paxos,
-replacing the broadcast of the leader by the propagation of the leader's history
-along the chain of servers.
-
-Ring Paxos optimizes Chain Replication by having the leader broadcast
-commands to the acceptors using IP multicast and only sending hashes
-of the commands along the chain. Ring Paxos round can be modified to implement
-conservative ACA rounds in a similar way as for Chain Replication.
-
-\subsection{Fast Paxos}
-
-Fast Paxos is composed of two types of rounds: classic rounds, like in Classic
-Paxos, and fast rounds.
-The fast rounds can be transformed into
-implementations of fast ACA rounds. Instead of sending their
-commands to a leader, as in Classic Paxos, the proposers of fast
-rounds directly broadcast their commands to all the acceptors. As long
-as commands are received in the same order by a learn quorum of
-acceptors or are received in different orders but commute, commands can be learned in two message-propagation delays instead of three in Classic Paxos.
-However, if non-commuting commands are received in different orders by
-several acceptors, the local history of the acceptors may not form a
-chain anymore and the rounds must abort. The fast rounds of Fast
-Paxos implement fast ACA rounds.
-
-Fast Paxos ACA rounds are specified using the same set of variables as for
-Classic Paxos, initialized to the same values.
-The next-state relation is composed of the following types of
-transitions, where the definitions of learn quorums, abort quorums,
-and of the $SafeAborts\left( R \right)$ operator are the ones of fast
-ACA rounds.
-\begin{compactitem}
-    \item A $Propose\left( p,c \right)$ transition, for a proposer $p$
-        and a command $c$, not guarded, inserts $c$ in the set
-        $proposed$ and broadcasts the message $\aseq{\quo{prop},c}$ to
-        all the acceptors.
-    \item An $Init\left( s, h \right)$ transition, for a switcher $s$
-        and a history $h$, not guarded, inserts $h$ in the set $inits$ and
-        broadcasts the message $\aseq{\quo{init},h}$ to all the acceptors.
-    \item A $WakeUp\left( a \right)$ transition, for an acceptor $a$,
-        is enabled when $status\left[ a \right]$ is ``idle'' and $a$
-        can receive an $\aseq{\quo{init},h}$ message. Its effect is to
-        receive the message, set $hist\left[ a \right]$ to $h$, and to
-        set $status\left[ a \right]$ to ``ready''.
-    \item An $Accept\left( a \right)$ transition, for an acceptor $a$,
-        is enabled when $status\left[ a \right]$ is ``ready'' and $a$
-        can receive a $\aseq{\quo{prop},c}$ message. The effect of the
-        action is to receive the message, to update $hist\left[ a
-        \right]$ to $hist\left[ a \right]\star c$, and to broadcast
-        the message $\aseq{\quo{ack},hist\left[ a \right]\star c}$ to
-        the learners.
-    \item The $RcvAck\left( l \right)$ transitions, for a learner $l$,
-        the $Learn\left( l, h \right)$ transitions, for a learner $l$
-        and a history $h$, the $Stop\left( s \right)$ transitions, for
-        a switcher $s$, $Stop\left( a \right)$ transitions, for an
-        acceptor $a$, the $RcvStopped\left( s \right)$ transitions,
-        for a switcher $s$, and the $Abort\left( s, h \right)$
-        transitions, for a switcher $s$ and a history $h$, which are
-        all the same as in Classic Paxos ACA rounds.
-\end{compactitem}
-
-As for Classic Paxos ACA rounds, Fast Paxos ACA rounds implement ACA
-under the refinement mapping consisting projecting the state of the
-round onto the variables $proposed$, $inits$, $aborts$, and, for every
-learner $l$, $learned\left[ l \right]$.
-
-\newpage
-
-\printbibliography
-
-\begin{appendix}
-
-  \label{sec:isaproofs}
-
-%  \includepdf[pages={1},pagecommand={\section{Mechanical Proof of the
-%  Composition Theorem in Isabelle/HOL}\label{sec:isaproofs}},scale=1]{../IsabellePDFS/podctheories.pdf}
-%  \includepdf[pages={2-},pagecommand={},scale=1]{../IsabellePDFS/podctheories.pdf}
-
-\end{appendix}
-
-\end{document}

### 2015-05-09 paper/abstract2.tex
Work on the abstract

\abstract{
  State-machine replication enables building reliable computing services on top
  of several unreliable hardware units. SMR is use in clusters of machines and
  will become increasingly important as the individual cores of many-core
  processors become more unreliable but more numerous.

    To achieve consistent performance, a distributed computing service must
    dynamically adapt to unpredictable crashes, communication failures, 
    shared resources availability, load and access patterns.
    Most existing algorithms are not adaptive: outside of their sweet spot, they have poor performance.
    We focus on State-Machine Replication (SMR) algorithms, a basic building block of reliable
    distributed computing.
    We present a framework that allows changing SMR algorithm on the fly while
    preserving the semantics of SMR. Once an SMR algorithm is cast into our
    framework and proved composable, it can be composed with any other
    composable SMR algorithm in any order.

    The key novelty of our framework is its practial applicability, having state
    of the art optimization techniques built-in: it supports executing commutative commands 
    without synchronization, optimizing the execution of read-only commands,
    minimizing the size of crucial data structures.

    We present a detailed, mechanically-proved formalization of our framework
    and we show how several SMR algorithms fit into our framework.
    We classify SMR algorithms in two categories corresponding to two abstract
    specifications that the algorithms refine.

    The framework is fully formalized and its main properties
    mechanically proven with Isabelle/HOL. The concrete
    algorithms we present are formalized in TLA+ and shown to
    fit in our framework using the TLC model-checker.
  }

### 2015-05-09 paper/paper.tex
Work on the abstract

@@
 
-\abstract{
-    To achieve consistent performance, a distributed computing service must
-    dynamically adapt to unpredictable crashes, communication failures, and
-    changes in the availability of shared resources.
-    Unfortunately, most existing algorithms are not adaptive: outside of their
-    sweet spot, they have poor performance.
-
-    We present a new abstraction, Abortable Generalized Consensus, which enables the
-    correct-by-construction composition of heterogeneous State-Machine
-    Replication algorithms.
-
-    integrating state of the art techniques to optimize executions.  ACA is the most complete approach to date enabling the
-    correct-by-construction composition of heterogeneous State-Machine
-    Replication algorithms. We provide two specifications satisfying
-    ACA and we illustrate their power by refining them to
-    obtain a versatile adaptive algorithm that combines five existing
-    algorithms into a single one. ACA is completely formalized and its
-    underlying composition theorem has been mechanically checked by an interactive
-    proof assistant.
-}
+\input{abstract2}
 

### 2015-05-09 paper/abstract2.tex
Work on the abstract

@@
 
-    The key novelty of our framework is its practial applicability, having state
-    of the art optimization techniques built-in: it supports executing commutative commands 
+    The key novelty of our framework is its practical applicability, having state
+    of the art optimization and reconfiguration techniques built-in: tolerating k fault with k+1
+    replicas using an external reconfiguration master, executing commutative commands 
     without synchronization, optimizing the execution of read-only commands,
@@
 
-    We present a detailed, mechanically-proved formalization of our framework
-    and we show how several SMR algorithms fit into our framework.
-    We classify SMR algorithms in two categories corresponding to two abstract
-    specifications that the algorithms refine.
+    Moreover, our framework allows a novel characterization of SMR algorithms in two categories corresponding to two abstract
+    specifications that the algorithms refine. The two categories represent a
+    fundamental three-dimensional trade-off relating resilience to crashes,
+    resilience to contention, and processing latency. Cluster idea?
+
+    We present a detailed, mechanically-proved formalization of our framework in
+    Isabelle/HOL, and we show how several SMR algorithms fit into our framework.
 

### 2015-05-09 paper/abstract2.tex
Work on the abstract

@@
     The key novelty of our framework is its practical applicability, having state
-    of the art optimization and reconfiguration techniques built-in: tolerating k fault with k+1
+    of the art optimization and reconfiguration techniques built-in:
+    primary-backup replication, tolerating k fault with k+1
     replicas using an external reconfiguration master, executing commutative commands 
@@
     minimizing the size of crucial data structures.
+
+    \begin{itemize}
+      \item Checkpointing
+      \item Primary-backup replication with expeditive reads, k-fault tolerance with
+        k+1 replicas.
+      \item Latency of two messages when contention happens only on commutative
+        commands.
+    \end{itemize}
 

### 2015-05-10 paper/abstract3.tex
work on the abstract

\abstract{
  State-Machine Replication (SMR) is a basic building block of reliable distributed
  systems that allows to implement a reliable service out of unreliable
  components. Practical implementations of SMR that can be used in cloud
  computing would have a direct impact on the reliability of many computing 
  services. However, devising new practical SMR algorithms is hard, often
  justify a publication in a top conference.

  The problem is that SMR algorithms are not modular. Thus all the reasonging,
  proofs, and test need to be done again from scratch when devising a new algorithm. With the number of
  different situations that a practical SMR algorithm must face, a modular approach
  allowing the reuse of existing techniques is especially important.

  We present a framework that enables composing SMR algorithms.
  Crucially, with our framework, an algorithm  is  proved  composable  once  and
  for all in isolation. Our composition theorem, mechanically proved with Isabelle/HOL,
  guarantees that  the  composition of  any  number of composable  algorithms  is a
  correct SMR algorithm.

  The key novelty of our framework is its practical  applicability,  with  state
  of  the  art  optimization  and  reconfiguration  techniques  being  built-in:
  tolerating $k$ faults with  $k+1$  replicas,  executing  commutative  commands
  without synchronization, optimizing the execution of read-only  commands,  and
  checkpointing.

  To demonstrate our framework  we  present  two  abstract
  composable algorithms  modeling  concrete algorithms  based  on  two  types  of  quorum
  systems used in practice. We demonstrate that concrete algorithms can easily be
  obtained by refinement from the abstract algorithms and we provide TLA+ specifications 
  ready to be model-checked with the TLC model-checker.
}

### 2015-05-10 paper/abstract3.tex
work on the abstract

@@
 \abstract{
-  State-Machine Replication (SMR) is a basic building block of reliable distributed
-  systems that allows to implement a reliable service out of unreliable
-  components. Practical implementations of SMR that can be used in cloud
-  computing would have a direct impact on the reliability of many computing 
-  services. However, devising new practical SMR algorithms is hard, often
-  justify a publication in a top conference.
+  State-Machine Replication (SMR) is a basic building block of reliable
+distributed systems. Improving SMR algorithms used by cloud-computing providers
+can have a big impact on the reliability of the many computing services they
+support. However, devising a new practical SMR algorithm is hard, often
+justifying a publication in a top conference.
 
-  The problem is that SMR algorithms are not modular. Thus all the reasonging,
-  proofs, and test need to be done again from scratch when devising a new algorithm. With the number of
-  different situations that a practical SMR algorithm must face, a modular approach
-  allowing the reuse of existing techniques is especially important.
+The problem is that SMR algorithms are not modular. Thus all the analysis,
+proofs, and tests need to be done again from scratch when devising a new
+algorithm. A modular approach allowing the reuse of existing techniques is
+especially important because of the many different situations that a practical
+SMR algorithm must face.
 
-  We present a framework that enables composing SMR algorithms.
-  Crucially, with our framework, an algorithm  is  proved  composable  once  and
-  for all in isolation. Our composition theorem, mechanically proved with Isabelle/HOL,
-  guarantees that  the  composition of  any  number of composable  algorithms  is a
-  correct SMR algorithm.
+We present a framework that enables devising and composing modular SMR
+algorithms. Crucially, with our framework, an algorithm is proved composable
+once and for all in isolation. Our composition theorem, mechanically proved
+with Isabelle/HOL, guarantees that the composition of any number of composable
+algorithms is a correct SMR algorithm.
 
@@
 of the art optimization and reconfiguration techniques being built-in:
-  tolerating $k$ faults with  $k+1$  replicas,  executing  commutative  commands
+tolerating k faults with k + 1 replicas, executing commutative commands
 without synchronization, optimizing the execution of read-only commands, and
@@
 
-  To demonstrate our framework  we  present  two  abstract
-  composable algorithms  modeling  concrete algorithms  based  on  two  types  of  quorum
-  systems used in practice. We demonstrate that concrete algorithms can easily be
-  obtained by refinement from the abstract algorithms and we provide TLA+ specifications 
-  ready to be model-checked with the TLC model-checker.
+To demonstrate our framework we present two abstract composable algorithms
+characterizing two popular types of quorum systems. We demonstrate that concrete
+algorithms can easily be obtained by refinement from the abstract algorithms, as
+evidenced by TLA+ specifications ready to be checked by the TLC model checker.
 }

### 2015-05-11 paper/paper.tex
Work on the abstract
Work on the introduction

@@
 
-\input{abstract2}
+\input{abstract3}
 
@@
 challenge and a multitude of algorithms have been proposed. Examples
-include Fast Paxos \cite{Lamport06FastPaxos}, Disk Paxos
+include Paxos \cite{lamport2001paxos}, Fast Paxos \cite{Lamport06FastPaxos}, Disk Paxos
 \cite{GafniLamport03DiskPaxos}, Chain Replication
@@
 \cite{MaoJunqueiraMarzullo08MenciusBuildingEfficientReplicatedStateMachine},
-and Fast Mencius \cite{WeiETAL13FastMenciusMenciusLowCommitLatency}.
+Fast Mencius \cite{WeiETAL13FastMenciusMenciusLowCommitLatency}, the Chandra-Toueg \cite{ChandraToueg96UnreliableFailureDetectorsReliableDistributedSystems}, the Ben-Or algorithm \cite{BenOr83AnotherAdvantageFreeChoiceCompletelyAsynchronous}, Raft \cite{OngaroOusterhout14SearchUnderstandableConsensusAlgorithm}.
+
+With the wide use of cloud computing, improvements in the performance and efficiency of SMR algorithms have the potential to impact most online services: advanced SMR algorithms can be implemented by cloud computing service providers and benefit all their customer's applications.
+
+Devising new SMR algorithms, optimized with particular situations or metrics in mind, is hard: today, several decades after the first SMR algorithms \cite{Lamport98ParttimeParliament,BirmanJoseph87ReliableCommunicationPresenceFailures,OkiLiskov88ViewstampedReplicationGeneralPrimaryCopy,DworkLynchStockmeyer84ConsensusPresencePartialSynchronyPreliminaryVersion} were published, devising new SMR algorithms is still hard enough to justify a publication in a top conference \cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments,OngaroOusterhout14SearchUnderstandableConsensusAlgorithm}.
+
+The problem is that it is very hard to reuse existing algorithms and ideas because SMR algorithms are not modular.
+When devising a new algorithm, all the careful analysis, tests, and proofs have to be done again from scratch.
+
+We propose a framework that simplifies devising new SMR algorithms by decomposing them into independent and composable modules.
+Crucially, in our framework, a SMR module can be analyzed, tested, and proved correct  once and for all. 
+Our composition theorem guarantees that composing any number of correct modules can be composed to form a correct SMR algorithm.
+
+A composite SMR algorithm built with our framework works by running the SMR module most appropriate at any given time, dynamically changing SMR module when needed.
+Thus algorithms designers who want to create a new SMR algorithm, optimized for particular conditions, can focus only on the particular conditions of interest and reuse existing modules to handle the other cases.
+All the performance analysis, tests, and proofs of the existing SMR modules remain valid and only the new SMR needs to be analyzed.
+Our framework thus enables incremental development of SMR algorithms, where each new development can build on the previous ones instead of starting again from scratch.
+
+An important practical aspect of SMR is fast reconfiguration to replace faulty components, without which the stability of a system is inevitably compromised.
+
+To be usable in  practice, SMR implementations must have high performance, be it throughput, latency, or some other metric depending on the setting, and resort to many optimizations which a framework like ours must support to be practical.
+Two key practical aspects of SMR are the ability to reconfigure the system to replace faulty components and the ability to process commutative commands without synchronization.
+There lies our key contribution: 
+Our framework is based on Generalized Consensus, a formulation of the State-Machine Replication problem that allows processing commutative commands without synchronization.
+
+At the heart of our framework lies a simple abstraction of the familiar notion of rounds (also called epochs, ballots, or views in the literature) used in most SMR algorithms.
+To illustrate our approach, let us consider Fast Paxos, one of the first examples of SMR algorithm which can change mode dynamically.
+Fast Paxos is of round-based SMR algorithm that can change mode.
+Each round is categorized as a fast or a normal round.
+Fast rounds process request faster than normal rounds but may fail to make progress in case of contention, at which point a normal round can take over the execution and make progress, later passing the baton back to a fast round when contention becomes low again.
+As in the original Paxos algorithm, normal rounds can also succeed each other in order to change
+leader.
+Each kind of round maintains different state invariants, thus a normal round must know which past rounds are fast or normal rounds in order to take over the execution safely.
+When adding more kinds of rounds such an approach will quickly become unmanageable, as each type of rounds must have one case for each possible other kind of round, making $n^2$ cases to consider in total.
+
+Our notion of Composable Generalized Consensus defines a common interface and a common correctness property for a round, abstracting over the particular round implementations and making all rounds compatible with each other.
+Moreover, our interface for rounds enable a straightforward crucial optimization seen in practical systems: introducing an external reconfiguration master which manages the addition of new replicas and the removal of failed ones.
+Interestingly, as explained in \cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}, in a real system it is often possible to implement a reliable reconfiguration master which allows to tolerate $f$ faults with $f+1$ replicas instead of the traditional $2f+1$ replicas .
+
+The reconfiguration master can also make the choice of the new type of round to execute, providing the system's reconfiguration policy.
+
+************************
 

### 2015-05-12 paper/paper.tex
Note about lattice agreement
Notes plus fixed typos

@@
 Moreover, our interface for rounds enable a straightforward crucial optimization seen in practical systems: introducing an external reconfiguration master which manages the addition of new replicas and the removal of failed ones.
-Interestingly, as explained in \cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}, in a real system it is often possible to implement a reliable reconfiguration master which allows to tolerate $f$ faults with $f+1$ replicas instead of the traditional $2f+1$ replicas .
+Interestingly, as explained in \cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}, in a real system it is often possible to implement a reliable reconfiguration master which allows to tolerate $f$ faults with $f+1$ replicas instead of the traditional $2f+1$ replicas.
 
@@
 \end{document}
-
     We present a new abstraction, Abortable Generalized Consensus, which enables the
@@
   general notion of command history, presented in the next subsection.
-  The Isabelle/HOL formalization of ACA uses the theory of I/O automata, which
-  is also formalized in the Isabelle/HOL. The proof uses a refinement
+  The Isabelle/HOL formalization of ACA uses the theory of I/O-automata, which
+  is also formalized in the Isabelle/HOL development. The proof uses a refinement
   mapping from the composition of two consecutive ACA instances to a single ACA
@@
 history of the leader, therefore guaranteeing that the local acceptor histories
-forming a chain.
-
-Acceptor broadcast each newly accepted history to the learners, which
+form a chain.
+
+Each acceptor broadcasts each newly accepted history to the learners, which
 learn a history $h$ when they receive enough extensions of $h$ from the
@@
 original rounds, the leader of the new round directly queries the
-acceptors of the preceeding round, whereas in ACA rounds, the
+acceptors of the preceeding rounds, whereas in ACA rounds, the
 switchers query the acceptors and then send their abort histories to

### 2015-05-14 paper/intro.tex
Moved the intro and roadmap to their own files.

--- 2015-05-12 paper/paper.tex
@@
-\input{header.tex}
-\input{defs.tex}
-
-\addbibresource{./managed.bib}
-
-\title{Changing Flights In The Air\\ (Abortable Chain Agreement)}
-
-\author{
-  Rachid Guerraoui\\
-  \texttt{rachid.guerraoui@epfl.ch}
-  \and
-  Giuliano Losa\\
-  \texttt{giuliano.losa@epfl.ch}
-}
-
-\date{}
-
-\begin{document}
-
-\maketitle
-
-\input{abstract3}
-
-\vspace{2cm}
-
-\begin{center}
-This paper is a regular submission.
-
-\vspace{1cm}
-This paper is a student paper.
-\end{center}
-
-\newpage
-
 \section{Introduction}
@@
 The reconfiguration master can also make the choice of the new type of round to execute, providing the system's reconfiguration policy.
-
-************************
-
-Each of the existing algorithms is best suited to a particular range
-of environments, characterized, e.g., by client load, network
-topology, likelihood of faults, communication latency and bandwidth, processing
-power of nodes, availability of disk storage and memory.
-However, the environment faced by real systems evolves over time:
-hardware is upgraded, new data-centers are built, the popularity of a
-service changes, etc.  Consistent performance therefore comes with
-adaptation to change.  Unfortunately, existing SMR algorithms cannot easily be
-stopped and replaced by another algorithm on the fly. Moreover,
-devising a new algorithm that can change behavior dynamically is hard:
-dynamically switching from one SMR algorithm to another requires
-synchronizing notoriously intricate algorithms.
-
-We propose a new abstraction, \emph{Abortable Chain Agreement}
-(abbreviated ACA), which subsumes SMR and whose implementations,
-called \emph{ACA rounds}, can, by construction, pass the baton from
-one to another on the fly and efficiently. Moreover, optimizing a replicated
-service consisting of several ACA rounds becomes just a matter of devising a new
-ACA round, which can then be plugged-in on the fly: any new ACA round
-is compatible with every existing ACA round by construction.
-
-Abortable Chain Agreement is the \emph{most complete framework to date} enabling
-correct by construction composition of heterogeneous rounds.  
-ACA is completely formalized and its underlying composition theorem has been
-mechanically checked by an interactive proof assistant.  
-Compared to ACA,
-existing approaches
-\cite{BarNoyETAL87ShiftingGearsChangingAlgorithmsFlyTo,GuerraouiETAL10Next700BftProtocols,GuerraouiKuncakLosa12SpeculativeLinearizability}
-are limited. In particular, they do not allow the disorderly execution of
-commuting and read-only commands, an important practical optimization
-which is at the root of recent work on SMR like Generalized Consensus
-\cite{Lamport05GeneralizeConsensus} and Generalized Lattice Agreement
-\cite{FalerioETAL12GeneralizedLatticeAgreement}.
-
-\begin{comment}
-On top of the traditional roles of proposer, acceptor, and learners
-\cite{lamport2001paxos} used in consensus algorithms, ACA introduces
-the notion of \emph{switchers} and of \emph{scheduling policy}. The switchers
-have the task of aborting a running ACA round and initializing a new
-ACA round in a safe and live manner.  The \emph{scheduling policy}
-decides when to abort an ACA round and choses the next ACA round to
-execute. It must ensure that all switchers agree on the next ACA round
-to initialize.
-\end{comment}
-
-We observe that, because of the FLP impossibility result
-\cite{FischerLynchPaterson83ImpossibilityDistributedConsensusOneFaultyProcess},
-most existing SMR algorithms are structured in a sequence of abortable
-rounds which can easily be transformed into ACA rounds. To simplify
-this task even further, we propose two specifications, \emph{conservative
-ACA rounds} and \emph{fast ACA rounds}, which, as we show, can be concretized
-to obtain ACA versions of the rounds of five existing algorithms: 
-Classic Paxos, Fast Paxos, Chain Replication, Ring Paxos,
-and Multi-Coordinated Paxos. ACA guarantees that the composition of the five ACA
-rounds is correct by construction, forming a versatile adaptive replication
-algorithm.
-
-The rest of the paper is structured as follows.
-\Cref{sec:related} surveys recent related work. \Cref{sec:ca} presents Chain
-Agreement, which embodies the requirements for implementing SMR. \Cref{sec:aca}
-presents Abortable Chain Agreement and its optimizations to treat commuting
-and read-only commands efficiently. \Cref{sec:rounds} presents our specification
-of conservative and fast ACA rounds, which implement ACA. In
-\cref{sec:examples}, we show that the rounds of  
-Classic Paxos, Fast Paxos, Chain Replication, Ring Paxos,
-and Multi-Coordinated Paxos can easily be transformed into ACA implementations
-by using our specifications of conservative and fast ACA rounds.
-Finally, in \cref{sec:isaproofs}, we provide a complete formalization of ACA
-and a mechanical proof of the \emph{composition theorem} (\cref{thm:comp}),
-checked by the interactive proof assistant Isabelle/HOL \cite{NipkowPaulsonWenzel02IsabelleHOL}.
-
-\begin{comment}
-Abortable Chain Agreement builds on several recent advances in SMR
-algorithms: ACA allows acceptors to only eventually reach agreement,
-like in Generalized Paxos \cite{Lamport05GeneralizeConsensus},
-boosting performance when most commands commute; ACA allows fast
-execution of read commands like in Generalized Lattice Agreement
-\cite{FalerioETAL12GeneralizedLatticeAgreement}; ACA implementations
-can tolerate $f$ faults with only $f+1$ replicas and replace crashed replicas
-upon changing round, as in Vertical Paxos  \cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}.  
-\end{comment}
-
-\section{Related Work}
-\label{sec:related}
-
-SMR is often implemented using a sequence of instances of Consensus.
-There are numerous examples of consensus algorithms in the literature.
-However, Abortable Chain Agreement is based on Generalized Consensus. 
-In Generalized Consensus \cite{Lamport05GeneralizeConsensus}, the learners of a
-SMR algorithm repeatedly learn new values belonging to a \emph{C-struct set}. A
-C-struct can be seen as a data structure encoding  sequences of commands modulo
-the ordering of commands that commute. The advantage of Generalized
-Consensus over Consensus is that
-two servers which receive commuting commands in different orders can still reach
-agreement on C-structs without further synchronization. 
-Our Abortable Chain Agreement can also use C-structs, with the same advantages
-as Generalized Consensus. For simplicity, the presentation in this
-paper uses the more concrete notion of \emph{command history} (see
-\cref{sec:histories}).
-
-Generalized Lattice Agreement \cite{FalerioETAL12GeneralizedLatticeAgreement} allows more efficient implementations of SMR when
-commands are partitioned into read-only commands and write-only commands.
-Abortable Chain Agreement supports the same optimizations as Generalized Lattice Agreement.
-Notably, read-only commands are treated specially so as to optimize their
-executions (see \cref{sec:readonly}).
-
-Vertical Paxos algorithms 
-\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication} are SMR algorithms
-whose set of replica servers can be reconfigured without stopping to process new
-commands. Crashed servers can therefore be replaced by new servers, ensuring the
-long-term availability of a replicated service.
-Vertical Paxos algorithms use an external reconfiguration master, itself
-implemented as a replicated state machine, which determines the servers that should
-be removed from or added to the configuration.
-ACA shares some aspects of Vertical Paxos algorithms: the task of the
-reconfiguration master is similar to the task of the scheduling policy
-component of ACA (see \cref{sec:aca}).
-However, in ACA, changing round can be used not only to replace crashed servers
-but also to change the algorithm that the servers are running. 
-Changing algorithm is not possible in Vertical Paxos.
-In both Vertical Paxos and ACA, a reliable reconfiguration master, resp. a
-reliable scheduling policy, allows tolerating $f$ crashes with $f+1$ servers.
-
-The Abstract Framework \cite{GuerraouiETAL10Next700BftProtocols} proposes to
-build Byzantine Fault-Tolerant algorithms as the succession of abortable rounds
-called Abstract instances. Abstract instances must satisfy the Abstract
-correctness properties. By construction, the composition of any number of
-Abstract instances is a correct SMR implementation. However, the Abstract
-Framework uses totally ordered sequence of commands, making difficult the
-optimization of the execution of commuting and read-only commands that are
-possible with ACA.  Moreover, the Abstract Framework does not identify the
-crucial role of the scheduling policy for the long-term resilience of a service.
-Abortable Chain Agreement combines the ideas of the Abstract Framework,
-Generalized Paxos, Generalized Lattice Agreement, and Vertical Paxos in a
-unified abstraction.
-
-Finally, the Speculative Linearizability framework
-\cite{GuerraouiKuncakLosa12SpeculativeLinearizability} was a first
-attempt by the authors of the present paper at generalizing the
-Abstract Framework and mechanically proving its properties with the
-Isabelle/HOL interactive
-proofs assistant.  Like the Abstract Framework, the
-Speculative Linearizability framework is based on totally ordered
-sequences of commands, with the same drawback for the execution of
-commuting and read-only commands. Moreover, Abortable Chain Agreement is both
-conceptually simpler and subsumes the Speculative Linearizability
-property.  Like Speculative Linearizability, Abortable Chain Agreement
-and its main property, the composition theorem (\cref{thm:comp}), are
-formalized in Isabelle/HOL (see \cref{sec:isaproofs}).
-
-Abortable Chain Agreement is the most complete framework to date for building
-adaptive replication algorithms out of heterogeneous rounds, making state of the
-art optimization techniques available to algorithm designers, which was not
-possible before.
-
-\section{Chain Agreement}
-\label{sec:ca}
-
-We consider a set of asynchronous servers communicating by message passing in an
-asynchronous network. Servers are not necessarily sequential processes. We
-describe a behavior of the system using an infinite sequence of states, where a
-state is a function from variable names to values (i.e. a valuation
-of variables).  We specify sets of allowed behaviors using an initial predicate,
-a next-state relation (using unprimed and primed variables) and a fairness
-condition, all in the style of TLA+ \cite{Lamport02SpecifyingSystems}. However,
-for succinctness, our specification are rather informal, eluding non-essential
-details.
-
-Our specifications all contain by default a state variable $crashed$ and
-$Crash\left( x \right)$ transition which is always enabled. A server $x$ faithfully
-executes the algorithm that is assigned to it until its identifier is added to
-the variable $crashed$ by a $Crash\left( x \right)$ transitions.  When the identifier of a
-server is in the set $crashed$, all its actions are disabled.  In an execution,
-a server that is never added to the set $crashed$ is said correct.
-
-We consider a computing service exposing a set of commands $C$. The
-service has a sequential specification consisting of an initial state
-and a deterministic transition relation in which every command
-atomically updates the state of the service and produces an output. We
-assume that duplicate commands have no effect on the state of the
-service and return the same output as their first occurrence.
-
-In a traditional SMR algorithm, the servers can be seen has playing one or
-several of three roles: \emph{proposers}, which propose new commands to execute,
-\emph{acceptors}, which accept and order commands in a sequence, and
-\emph{learners}, which learn about the growing sequence of commands to
-execute \cite{lamport2001paxos}. Often, but not necessarily, every replica server plays the
-three roles of proposer, acceptor, and learner: clients of the service
-send their commands to some replica servers; replica servers submit
-the received commands to the SMR protocol in the role of proposers;
-replica servers respond to commands by executing the commands learned
-in the role of learners; replica servers participate in command
-ordering in the role of acceptors.  An SMR algorithm guarantees that
-the replicas learn the same sequences of commands, possibly up to the
-reordering of commuting commands. Therefore, every learner can respond to client
-requests using its last learned sequence of commands, and clients will receive
-the same responses as the ones produces by a centralized, non-replicated,
-service.
-
-Chain Agreement precisely specifies the allowed behaviors of the roles of \emph{proposers}
-and \emph{learners}, abstracting over the behavior of acceptors.
-
-\begin{comment}
-We suppose that each server can play a number of \emph{roles}.
-Chain Agreement specifies the allowed behaviors the roles of \emph{proposers}
-and \emph{learners}. Proposers receive commands from the clients of the service
-and propose them for execution by the replicated service. Each learner learns about
-a growing sequence of commands. Thanks to the sequential specification of the
-service, a sequence of commands determines an output for each the included commands. 
-In a traditional SMR algorithm, servers can also play the role of
-\emph{acceptors}, whose task is to collectively order commands and let the
-learners know about the latest sequence of ordered commands.
-
-The point of Chain Agreement is to implement a replicated service by having every learner learn about a growing sequence of commands, which
-represent an execution of the sequential specification of the service.
-Crucially, at all times, the sequences of commands learned by the
-different learners must form a chain in the prefix order on sequences.
-Moreover, for every command $c$ proposed by a correct proposer, every
-correct learner must eventually learn a sequence containing $c$.  Once
-a learner learns about a sequence containing a command $c$, it can
-determine the output produced by $c$, using the sequential
-specification of the service, and respond to the client that issued
-the command $c$.
-\end{comment}
-
-To specify Chain Agreement, we need the notion of ordering of sequences and of
-maximum sequence.  The prefix order on sequences is denoted by the operator
-$\leq$ ($s_1 \leq s_2$ means that $s_1$ is a prefix of $s_2$). 
-%Using the operator $\leq$, we define $\geq$, $<$, and $>$ as expected. 
-Note that $\leq$ is a partial order on sequences. The maximum over a set of
-sequences $S$ is denoted by $Max\left( S \right)$ and $Max\left( \aset{}
-\right)$ is the empty sequence.
-
-We specify Chain Agreement using the state variables $crashed$, $proposed$, containing the proposed commands,
-and, for every learner $l$,  $learned\left[ l \right]$, a set of
-sequences of commands. Initially, $crashed$, $proposed$ and, for every learner $l$,
-$learned\left[ l \right]$ are empty sets. 
-Remember that any transition involving a server $x$ is always disabled when $x\in crashed$.
-
-The next-state relation consists of three types of transitions:
-$Crash\left( x \right)$ transitions; $Propose\left( l, c
-\right)$ transitions, for a learner $l$ and a command $c$, in which
-$c$ is inserted in the set $proposed$; $Learn\left( l,cs \right)$
-transitions, for a learner $l$ and a sequence of commands $cs$, in
-which $cs$ is inserted in $learned\left[ l \right]$. 
-A $Learn\left( l, cs \right)$ transition is guarded by the following
-conditions:
-\begin{compactitem}
-    \item[\textbf{Agreement}:] The elements of the set $\aset{learned'\left[ l
-      \right] : l \in Learners}$, the learned sequences after the transition,
-      form a chain in the partial order $\leq$.
-    \item[\textbf{Validity}:] Every learned sequence contains only
-        proposed commands.  
-    \item[\textbf{Irrevocability}:] The new learned sequence
-        $cs$ is a strict extension of $Max\left(learned\left[ l
-        \right]\right)$. 
-        %\\ $cs > Max\left(learned\left[ l \right]\right)$.
-\end{compactitem}
-Finally, we add, for every learner $l$, one weak fairness condition
-\cite{Lamport02SpecifyingSystems} on the transitions $Learn\left( l,
-cs \right)$. The weak fairness conditions ensure that every correct
-learner eventually learns about every proposed command.
-
-Thanks to the properties of Chain Agreement, the clients, observing
-their input commands and the received responses, see a
-\emph{linearizable} execution
-\cite{HerlihyWing90LinearizabilityCorrectnessConditionConcurrentObjects}.
-Moreover, the service remains available as long as at least one
-proposer and one learner are correct.
-
-\section{Abortable Chain Agreement}
-\label{sec:aca}
-
-We now consider implementing Chain Agreement as a sequence of \emph{Abortable Chain
-Agreement rounds}, where each round makes progress ordering requests or aborts
-its execution and passes the baton to the next round in the sequence.
-An algorithm composed of a sequence of ACA rounds is called an \emph{ACA
-algorithm}.
-
-On top of the traditional roles of proposer, acceptor, and learner, ACA
-introduces the new role of \emph{switcher} and a new component, the
-\emph{scheduling policy}. In ACA, the assignment of the roles of proposer,
-acceptor, learner, and switcher can change in each rounds, allowing for
-reconfiguration. 
-In normal circumstances, the proposers, acceptors,
-and learners of an ACA round work together to implement Chain Agreement.
-However, if the environment becomes unfavorable to a running ACA round, the
-scheduling policy can instruct the round to abort its execution and pass the
-baton, using the switchers, to a new round.
-
-A switcher of an aborting round determines, from the state of the
-acceptors, an \emph{abort sequence} that it transfers to one or several
-switchers of the new round.  Upon receiving an abort sequence, a switcher of the
-new round initializes the acceptors of the new round, which can then
-proceed implementing SMR, until the new round itself aborts and the
-process is repeated. 
-We assume that in the first round, which has no previous round, every 
-switchers receives the empty abort sequence.
-
-The scheduling policy determines when to abort an ACA round and which new ACA
-round should take over. The scheduling policy must inform the switchers about the
-identity of the next round and must guarantee that all switchers receive the
-same information.  In this paper we abstract over the scheduling policy, assuming
-that a round that aborts passes the baton a nondeterministically chosen new
-round. A reliable scheduling-policy component can realistically be implemented
-as a replicated state machine running on a large number of servers otherwise
-dedicated to other tasks
-\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}. We will see
- in \cref{sec:rounds} that, as in Vertical Paxos, a reliable scheduling policy component allows
-tolerating $f$ faults with $f+1$ replicas, while maintaining the long-term
-resilience of the service by replacing faulty servers upon round change.
-
-The clients of an ACA algorithm can send their commands to the proposers of any round (in
-practice, clients send their commands to proposers of the largest
-active round) and servers can use sequences learned in any round
-to determine outputs.
-
-ACA specifies the allowed behaviors of the proposers, learners, and switchers
-\emph{of a single round}. Similarly to
-Chain Agreement, we use the state variables $crashed$, $proposed$,
-and, for every learner $l$, $learned\left[ l \right]$.  Additionally,
-we use the state variables $inits$, containing the abort sequences
-received from the previous round, and $aborts$, containing the abort
-sequences transfered to the next round.
-Two consecutive ACA rounds are composed by letting the $aborts$ variable of the
-first round be the $inits$ variable of the second round.
-
-Initially, $crashed$, $proposed$, for every
-learner $l$, $learned\left[ l \right]$, $inits$, and $aborts$ are all empty sets.
-The next-state relation consists of five types of transitions. The
-first three, $Crash\left( x \right)$, $Propose\left( l, c
-\right)$, and $Learn\left( l, cs \right)$, for a learner, proposer, or
-switcher $x$, for a learner $l$, and for a sequence of commands
-$cs$, have the same effect as in Chain Agreement.
-ACA introduces two new transitions, $Init\left(
-s, cs \right)$ and $Abort\left( s, cs \right)$, for a switcher $s$ and a sequence of
-commands $cs$. The transition $Init\left( s, cs \right)$ adds $cs$ to the set $inits$, and models a switcher receiving the
-abort sequence $cs$ from the preceding round.
-The transition $Abort\left( s, cs \right)$ adds $cs$ to the set $aborts$ and models a switcher transferring the abort sequence $cs$ to a switcher of the next round.
-All the transition are guarded by the following conditions:
-\begin{compactitem}
-    \item[\textbf{Agreement}:] The elements of the set $\aset{learned'\left[ l
-      \right] : l \in Learners}$ form a chain in the partial order.
-    \item[\textbf{Validity}:] Every abort sequence or learned sequence
-        is of the form $h \star cs$, where $cs$ contains only commands
-        of the set $proposed$ and $h$ is the greatest lower bound of a
-        nonempty subset of $inits$.
-    \item[\textbf{Initialization}:] If $inits$ is empty, then no
-        sequence can be learned and no switcher can abort.
-    \item[\textbf{Irrevocability}:] In the case of a $Learn\left( l,cs
-        \right)$ transition, the new learned sequence $cs$ is a strict
-        extension of $Max\left(learned\left[ l \right]\right)$.
-        %, $cs
-        %\geq Max\left(learned\left[ l \right]\right)$.
-    \item[\textbf{Safe Abort}:] Every abort sequence is larger in the
-        partial order $\leq$ than every learned sequence.
-\end{compactitem}
-Note that Agreement and Irrevocability are the same as in Chain
-Agreement.
-
-Finally, for every learner $l$, we add a weak fairness condition on the
-following state transition: 
-\begin{equation}
-    \exists  s \in Switchers, cs \in Sequence\left( C \right) : Learn\left( l, cs \right) \vee Abort\left( s,cs \right)
-\end{equation}
-The weak fairness conditions ensure that when a correct learner has not
-learned a sequence containing all commands proposed in the round, then
-it eventually learns a new command or, if there is at least one correct switcher, at least one switch value is transfered to the next round. 
-Therefore, if a client has submitted a command to a proposer, eventually,
-either the client gets a response from the current round, or
-the next round is initialized with an abort sequence and the
-command can be resubmitted to the next round.
-
-\subsection{The Composition Theorem}
-
-Consider the succession of $n$ rounds numbered $1$ to $n$.  To
-distinguish the variables of each round, we annotate the local
-variables of each round with the number of the round. For example, the
-variables of the seventh round are written $proposed_7$, for every
-learner $l$, $learned_7\left[ l \right]$, $inits_7$, and $aborts_7$.
-
-Moreover, we assume that a command, which is proposed to a round and that
-is not learned before the first abort sequence is transfered to the
-next round, is proposed anew to the next round.
-
-Define the state variable $proposed$ as the union over all rounds $i$
-of the $proposed_i$ variables and, for every learner $l$, 
-define the $learned\left[ l \right]$ variable as the union over all
-rounds $i$ of the $learned_i\left[ l \right]$ variables.
-
-\begin{thm}[Composition Theorem]
-    \label{thm:comp}
-    In every execution of a succession of ACA rounds, the sequence of valuations
-    of the variable $proposed$ and, for every learner $l$, of the variables
-    $learned\left[ l \right]$ satisfies Chain Agreement.
-\end{thm}
-\Cref{thm:comp} can be shown by induction on the number of rounds,
-using the following lemma.
-\begin{lem}
-  \label{lem:comp}
-    Consider two rounds numbered $i$ and $i+1$. In every execution of a
-    succession of ACA rounds, the sequence of valuations of the variables $inits
-    = inits_i$, $propose = proposed_i \cup propose_{i+1}$, for every learner
-    $l$, $learned\left[ l \right] = learned_i\left[ l \right] \cup
-    learned_{i+1}\left[ l \right]$and $aborts = aborts_{i+1}$ satisfies ACA.
-\end{lem}
-\begin{proof}
-  A mechanically-checked proof in Isabelle/HOL appears in \cref{sec:isaproofs}.
-  The proofs uses a version of the specification of ACA that has a more
-  executable form that presented in this section. Moreover,
-  instead of sequences of commands, the Isabelle/HOL specification uses the more
-  general notion of command history, presented in the next subsection.
-  The Isabelle/HOL formalization of ACA uses the theory of I/O automata, which
-  is also formalized in the Isabelle/HOL. The proof uses a refinement
-  mapping from the composition of two consecutive ACA instances to a single ACA
-  instance.
-\end{proof}
-
-\Cref{thm:comp} guarantees that the successive composition of any
-number of ACA rounds implements Chain Agreement.
-Therefore, with ACA, one can easily build replicated services that
-dynamically adapt to the current operating conditions of the system: a
-replicated service has several types of rounds at its disposal, each being
-tailored to particular conditions. Each time the conditions change,
-the current ACA round aborts and passes the baton to a more 
-appropriate type of round, as determined by the scheduling policy. Moreover, a new type
-of ACA round can always be devised after the system is deployed and added on the fly to
-the existing types of rounds.
-
-We now describe two important practical optimizations of ACA.
-
-\subsection{Histories}
-\label{sec:histories}
-
-ACA can be implemented more efficiently by substituting \emph{command histories} for sequences of commands, as in Generalized Consensus \cite{Lamport05GeneralizeConsensus}.
-
-We say that two commands $c_1$ and $c_2$ \emph{commute} when, in every sequential
-execution of the service, reversing the order of $c_1$ and $c_2$ does
-not change any of the outputs, including those given to $c_1$ and
-$c_2$.
-
-A command history, abbreviated history, is a data structure that
-represents a set of sequences which are the same modulo the reordering
-of commuting commands, and in which duplicate commands do not appear.
-All the sequences represented by a given history contain the same
-commands and determine the same outputs for each command.  Therefore,
-to implement a replicated service, it is sufficient that learners
-learn command histories instead of sequences.  For a precise
-definition of command histories, we refer the reader to
-\cite{Lamport05GeneralizeConsensus}. Our Isabelle/HOL formalization
-(\cref{sec:isaproofs}) abstracts
-over the concrete representation of histories, assuming only their necessary
-properties. Command histories were first
-introduced in trace theory \cite{Mazurkiewicz84Semantics}.
-
-A command history can be built by appending a sequence of commands
-$cs$ to a history $h$, obtain the history $h \star cs$.  The major
-advantage that histories have over sequences is that if a sequence
-$cs_1$ is obtained by reordering the commuting commands of $cs_2$,
-then $h \star cs_1 = h \star cs_2$. Therefore, in  implementations of
-ACA, servers can agree on a history even though
-they chose different orders for some commuting commands.
-
-Command histories share the four properties of sequences that are
-necessary and sufficient to substitute histories for sequences in the
-definition of Chain Agreement and ACA.
-\begin{compactitem}
-    \item[\textbf{Distinctness}:] No duplicate command appears in a history. 
-    \item[\textbf{Ordering}:] Command histories form a partial order:
-        $h_1 \leq h_2$ when there exists a sequence $s$ such that $h_2
-        = h_1 \star s$.
-    \item[\textbf{GLB}:] Every finite set of command histories has a
-        \emph{greatest lower bound}.
-%    \item[\textbf{LUB}:] Every finite set of compatible command
-%        histories has a \emph{least upper bound}, where a set of
-%        command histories is compatible when they have a common upper
-%        bound.
-    \item[\textbf{Consistency}:] If $h_1 \leq h_2 \leq h_3$ and $h_3 =
-        h_1 \star s$ then there exists $s'$ and $s''$ where every
-        request of $s'$ and $s''$ appears in $s$ and such that $h_2 =
-        h_1 \star s'$ and $h_3 = h_2 \star s''$.
-\end{compactitem}
-The optimized version of ACA is obtained by
-simply substituting histories for sequences in the definition of ACA.
-The Ordering, GLB, and Consistency properties of histories ensure that
-\cref{thm:comp} still holds: the Isabelle/HOL proof of \cref{thm:comp} (see
-\cref{sec:isaproofs}) relies only on the four properties above.
-
-\subsection{Optimizing the Execution of Read-Only Commands}
-\label{sec:readonly}
-
-Chain Agreement can be further optimized by treating 
-read-only commands specially, as in Generalized Lattice Agreement
-\cite{FalerioETAL12GeneralizedLatticeAgreement}.
-
-We consider the case in which a set of servers each play the four
-roles of proposer, acceptor, learner, and switcher.
-Clients sends their commands to any server, which proposes them as a
-proposer, and records the command as pending.
-%When a server learns a history that contains one of its pending
-%commands, it uses the history to determine the response to the pending
-%command and forwards the response to the appropriate client.
-
-As observed in \cite{FalerioETAL12GeneralizedLatticeAgreement}, read-only
-commands can be executed faster by directly reading from the last learned
-history, instead of proposing the read commands and waiting for it to appear in
-a future learned history.  Because learned histories form a chain, all the reads
-return consistent values and this approach yields a \emph{serializable}
-implementation of the service.
-
-However linearizability is not satisfied if reads are served directly
-from the last learned history: a lagging replica server could respond
-to a read using a history which does not contain all completed update
-commands. To make the executions linearizable without resorting to
-proposing read-only commands as normal commands, we can employ the same
-technique as in \cite{FalerioETAL12GeneralizedLatticeAgreement}: upon
-receiving a read command, a server proposes a special no-op command
-that commutes with all other commands; the server then responds to the
-read command using the first learned history containing the no-op
-command. This ensures that all the commands that had completed before
-the read command was issued are present in the history.  Moreover,
-since no-op commands commute with all other commands, they can be
-treated more efficiently than regular read commands by the acceptors,
-like, for example, in a fast round of Fast Paxos.
-
-\section{Fast and Conservative Rounds}
-\label{sec:rounds}
-
-Most existing SMR algorithms can easily be obtained as the
-composition of one or more ACA rounds.
-
-The FLP impossibility result \cite{FischerLynchPaterson83ImpossibilityDistributedConsensusOneFaultyProcess}
-implies that no algorithm can implement SMR to the letter in practice.
-Instead, SMR algorithms guarantee safety (i.e. replica execute the
-same sequence of commands) but may not be able to make progress (i.e.
-order new commands). To make progress as likely as possible, most SMR
-algorithms are structured in consecutive rounds (also called ballots,
-views, or epochs). In each round, the algorithm tries to make progress
-but cannot be guaranteed to succeed. When the algorithm detects that
-progress is no more possible, a new round is started.  The new round
-is initialized so as to preserve safety, usually by one of the
-acceptors. For example, in Paxos, the leader of a newly started ballot
-numbered $i$ determines a value that is \emph{safe at $i$} and
-proposes it as first value to the other acceptors of the ballot $i$.
-In existing SMR algorithms, new rounds are initialized ad-hoc, which
-prevents mixing rounds of two different algorithms.  Instead,
-ACA is an abstraction of rounds which establishes a common interface and contract for switching from
-one round to the next, enabling switching from any ACA round to any other.
-
-To simplify the task of expressing existing algorithms as the composition of one
-or more ACA rounds, we propose two specifications, \emph{conservative ACA
-rounds} and
-\emph{fast ACA rounds}, which can easily be concretized to obtain executable rounds.
-We show that ACA versions of Classic Paxos, Multi-Coordinated Paxos, Chain
-Replication, and Ring Paxos can be obtained by concretizing the specification of
-conservative ACA rounds, and that ACA versions of the fast rounds of Fast Paxos can be obtained by concretizing the specification of fast ACA rounds.
-
-Fast and conservative ACA rounds are two specifications, at an
-intermediate level of abstraction, which satisfy ACA. Both
-specifications specify how the state of the acceptors should be
-updated but abstract over the particular communication topology used
-to implement the state accesses: the specifications use a global state
-which is accessible by all servers in all roles. One can refine fast
-or conservative rounds by implementing the state accesses and coordination
-using the network, obtaining a concrete algorithm.
-
-Conservative and fast ACA rounds are specified using the state variables $inits$, $proposed$, for every learner $l$, $learned\left[ l \right]$
-and $aborts$, as in ACA, and, for every acceptor $a$, two new variables
-$status\left[ a \right]$, which can be either ``idle'', ``ready'', or
-``stopped'' and $hist\left[ a \right]$, the local history of the
-acceptor $a$, which can be either a history or the special value $None$.
-Initially, $inits$, $aborts$, and, for every learner
-$l$, $learned\left[ l \right]$ are empty sets, and, for every acceptor $a$,
-$status\left[ a \right]$  is ``idle'' and $hist\left[ a \right]$ is the special
-value $None$.
-
-The next-state relations of conservative and fast rounds are composed
-of the following types of transitions (and the default $Crash\left( x
-\right)$ transition).
-\begin{compactitem}
-    \item A $Propose\left( p,c \right)$ transition, for a proposer $p$
-        and a command $c$, not guarded, inserts $c$ in the set
-        $proposed$. 
-    \item An $Init\left( s, h \right)$ transition, for a switcher $s$
-        and a history $h$, not guarded, inserts $h$ in the set $inits$.
-    \item A $WakeUp\left( a \right)$ transition, for an acceptor $a$,
-        is enabled when $status\left[ a \right]$ is ``idle''. Its
-        effect is to set $hist\left[ a \right]$ to a history found in
-        $inits$ and to set $status\left[ a \right]$ to ``ready''.
-    \item An $Accept\left( a \right)$ transition, for an acceptor $a$,
-        is enabled when $status\left[ a \right]$ is ``ready'', and
-        updates $hist\left[ a \right]$ to $hist\left[ a \right]\star
-        c$, where $c$ is a proposed command. In conservative rounds,
-        the transition is additionally guarded by the condition that
-        the resulting set of local acceptor histories,
-        $\aset{hist'\left[ a \right] : a \in Acceptors}$, forms a chain
-        in the partial order $\leq$ (this is typically implemented with a
-        leader).
-    \item A $Learn\left( l, h \right)$ transition, for a learner $l$
-      and a history $h$, enabled when $h > learned\left[ l \right]$ and is
-      smaller than the GLB of the histories of a learn quorum of acceptors. The effect of
-        the transition is to set $learned\left[ l \right]$ to $h$.
-    \item A $Stop\left( a \right)$ transition, for an acceptor $a$,
-        enabled when $status\left[ a \right]$ is ``ready'', sets
-        $status\left[ a \right]$ to ``stopped'' (note that
-        $status\left[ a \right] = $ ``stopped'' prevents any further
-        ``accept'' transition of acceptor $a$).
-%    \item REALLY NEEDED? A $TimeOut\left( a \right)$ transition, for an acceptor $a$,
-%        is enabled when $status\left[ a \right]$ is ``idle'' and sets 
-%        $status\left[ a \right]$ to ``stopped'' and $hist\left[ a
-%        \right]$ to the special value $None$. 
-    \item An $Abort\left( s, h \right)$ transition, for a switcher $s$
-        and a history $h$, is enabled when there exists a abort quorum
-        $R$ such that, for every member $a$ of $R$, $status\left[ a
-        \right]$ is ``stopped'' and $h \in SafeAborts\left( R
-        \right)$. Its effect is to insert  $h$ in the set $aborts$. 
-\end{compactitem}
-
-On top of the additional guard on the $Accept$ transition present only
-in conservative rounds, conservative and fast ACA rounds differ in
-their definition of learn quorums, abort quorums, and in their
-definition of the $SafeAborts\left( R \right)$ operator.
-To concisely define read and learn quorums, we define 
-\begin{equation}
-    Q\left( f \right) = \aset{as \subseteq Acceptors :
-        Card\left(as\right) \geq \left\lfloor f\times Card\left(
-    Acceptors \right)\right\rfloor + 1}.
-\end{equation}
-
-In conservative ACA rounds, abort and learn quorums are defined such
-that the intersection between two learn quorums is nonempty and the
-intersection between a learn quorum and an abort quorum is nonempty.
-The definition leads to two interesting cases: $LearnQ = AbortQ =
-Q\left( 1/2 \right)$; $LearnQ = \aset{Acceptors}$ and $AbortQ =
-\aset{\aset{a} : a \in Acceptors}$. The operator $SafeAborts\left( R
-\right)$ is defined as the singleton set containing the maximum
-element over the local histories of the acceptors in $R$ which are not
-$None$. Note that the maximum exists because conservative rounds
-guarantee that the histories of the acceptors form a chain.
-
-In fast ACA rounds, abort and learn quorums are defined such that
-the intersection between any two learn quorums is nonempty and the
-intersection between any two learn quorums and an abort quorum is
-nonempty.  The definition leads to three interesting cases: $LearnQ =
-AbortQ = Q\left( 2/3 \right)$; $LearnQ = Q\left( 1/2 \right)$ and
-$AbortQ = Q\left( 3/4 \right)$; $LearnQ = \aset{Acceptors}$ and
-$AbortQ = \aset{\aset{a : a \in Acceptors}}$.
-The operator $SafeAborts\left( R \right)$ is defined as the set of
-maximal elements over the set $\aset{GLB\left( \aset{hist\left[ a
-\right] : a \in R \cap L \wedge hist\left[ a \right]\neq None} \right)
-: L \in LearnQ}$.
-
-The cases in which abort quorums are singleton sets, in both
-conservative and fast rounds, allow tolerating
-$f$ faults with $f+1$ servers: a round can abort as long as one
-acceptor is correct and a new round, possibly using new correct
-servers, can take over. Note that changing round cannot happen
-without the scheduling policy instructing the switchers about the
-identity of the next round, and thus tolerating $f$ faults with
-$f+1$ servers relies on the assumption of a reliable scheduling policy.
-
-Define, for every learner $l$, the history variable $\overline{learned}\left[ l \right]$ which is initialized to the empty set and which is
-updated each time a new local history is accepted by the learner $l$ by
-inserting the new local history in $\overline{learned}\left[ l \right]$ (see \cite{AbadiLamport91ExistenceRefinementMappings} for the
-definition of a history variable).
-\begin{thm}
-  \label{thm:rounds}
-    Both conservative and fast ACA rounds implement ACA under the refinement
-    mapping that consists in substituting $\overline{learned\left[ l \right]}$
-    for $learned\left[ l \right]$.
-\end{thm}
-
-\section{Examples of Concrete ACA Algorithms}
-\label{sec:examples}
-
-In this section we show how to modify the rounds of Classic Paxos,
-Multi-Coordinated Paxos, Chain Replication, Ring Paxos, and Fast Paxos to obtain
-ACA rounds. \Cref{thm:comp} then guarantee that all those different types of
-rounds can be used in the same algorithm to yield an implementation of Chain
-Agreement.
-
-\subsection{Classic Paxos}
-
-The rounds of the Classic Paxos algorithm can be modified so as to
-obtain an implementation of conservative ACA rounds.  We define
-Classic Paxos ACA rounds by specializing the specification of
-conservative ACA rounds, giving a concrete protocol for disseminating
-the proposed commands to the acceptors and enforcing that the learned
-histories form a chain.
-
-Each Classic Paxos round has a unique leader among the acceptors which
-centrally orders the proposed commands, which are accepted by the
-other acceptors in the order determined by the leader.  This protocol
-ensures that every local acceptor history is a prefix of the local
-history of the leader, therefore guaranteeing that the local acceptor histories
-forming a chain.
-
-Acceptor broadcast each newly accepted history to the learners, which
-learn a history $h$ when they receive enough extensions of $h$ from the
-acceptors.  When instructed by the scheduling policy to abort, a
-switcher sends a stop message to the acceptors, which irrevocably stop
-updating their local history and send back to the switcher an
-acknowledgement containing their local history. A switcher
-determines an abort history once it has received acknowledgments from
-an abort quorum of acceptors $R$, using the operator $SafeAborts\left(
-R \right)$.
-
-The specification of Classic Paxos ACA rounds uses the variables of
-the specification of conservative ACA rounds, namely the variables
-$proposed$, $inits$, $aborts$, the arrays $learned\left[ l \right]$,
-$status\left[ a \right]$, and $hist\left[ a \right]$, and two
-additional arrays $accepted\left[ l \right]\left[ a \right]$ and
-$stop\left[ s \right]\left[ a \right]$, where $l$ is a learner, $a$ is
-an acceptor, and $s$ is a switcher.
-
-There is also a network component that we do not describe explicitly.
-However, the network allows any servers to communicate by sending and
-receiving messages.
-
-Initially, the variables $proposed$, $inits$, $aborts$, and, for every
-learner $l$, $learned\left[ l \right]$ are empty sets. For every
-acceptor $a$, $status\left[ a \right]$ is ``idle'' and $hist\left[ a
-\right]$ is the special value $None$. For every learner $l$ and acceptor $a$,
-$accepted\left[ l \right]\left[ a \right]$ is the special value
-$None$. For every switcher $s$ and acceptor $a$, $stopped\left[ s
-\right]\left[ a \right]$ is set to $false$.  There are initially no
-messages in the network.
-
-The next-state relation is composed of the following types of
-transitions, where the definitions of learn quorums, abort quorums,
-and of the $SafeAborts\left( R \right)$ operator are the ones of conservative
-ACA rounds.
-\begin{compactitem}
-    \item A $Propose\left( p,c \right)$ transition, for a proposer $p$
-        and a command $c$, not guarded, inserts $c$ in the set
-        $proposed$ and sends the message $\aseq{\quo{prop},c}$ to
-        the leader.
-    \item An $Init\left( s, h \right)$ transition, for a switcher $s$
-        and a history $h$, not guarded, inserts $h$ in the set $inits$ and
-        sends the message $\aseq{\quo{init},h}$ to the leader.
-    \item A $WakeUp\left( leader \right)$ transition, enabled when
-        $status\left[ leader \right]$ is ``idle'' and the leader can
-        receive a $\aseq{\quo{init},h}$ message. Its effect is to
-        receive the message, set $hist\left[ leader \right]$ to $h$,
-        to set $status\left[ leader \right]$ to ``ready'', and to
-        broadcast the message $\aseq{\quo{leader-accept},h}$ to all
-        other acceptors.
-    \item An $Accept\left( leader \right)$ transition, enabled when
-        $status\left[ leader \right]$ is ``ready'' and the leader can
-        receive a $\aseq{\quo{prop},c}$ message. The effect of the
-        action is to receive the message, to update $hist\left[ leader
-        \right]$ to $hist\left[ leader \right]\star c$, and to
-        broadcast the message $\aseq{\quo{ack},hist\left[ leader
-        \right]\star c}$ to the learners, and to broadcast the message
-        $\aseq{\quo{leader-accept},hist\left[ leader \right]\star c}$
-        to all the other acceptors.
-    \item An $Accept\left( a \right)$ transition, for an acceptor $a$
-        which is not the leader, is enabled when $status\left[ a
-        \right]$ is ``ready'' and $a$ can receive a
-        $\aseq{\quo{leader-accept},h}$ message where $h > hist\left[ a
-        \right]$. The effect of the action is to receive the message,
-        to update $hist\left[ a \right]$ to $h$, and to broadcast the
-        message $\aseq{\quo{ack},hist\left[ a \right]\star c}$ to the
-        learners.
-    \item A $RcvAck\left( l \right)$ transition, for a learner $l$, is
-        enabled when $l$ can receive a message $\aseq{\quo{ack},h}$
-        from an acceptor $a$. Its effect is to receive the message and
-        to set $accepted\left[ l \right]\left[ a \right]$ to $h$.
-    \item A $Learn\left( l, h \right)$ transition, for a learner $l$
-        and a history $h$, is enabled when $h > Max\left(
-        learned\left[ l \right] \right)$ and there is a learn quorum
-        $Q$ of acceptors such that $h = GLB\left( \aset{accepted\left[
-        l \right]\left[ a \right] : a \in Q }\right)$. Its effect is
-        to insert $h$ in $learned\left[ l \right]$.
-    \item A $Stop\left( s \right)$ transition, for a switcher $s$,
-        always enabled, broadcasts a $\aseq{\quo{stop}}$ message to all
-        the acceptors.  This transition models a switcher being
-        instructed by the scheduling policy to abort the current round.
-    \item A $Stop\left( a \right)$ transition, for an acceptor $a$,
-        enabled when $a$ can receive a $\aseq{\quo{stop}}$ message
-        from a switcher $s$. Its effect is to receive the message, set
-        $status\left[ a \right]$ to ``stopped'', and to send the
-        message $\aseq{\quo{stopped},hist\left[ a \right]}$ to the
-        switcher $s$.
-    \item A $RcvStopped\left( s \right)$ transition, for a switcher
-        $s$, enabled when $s$ can receive a $\aseq{\quo{stopped},h}$
-        message from an acceptor $a$. Its effect is to receive the
-        message and to set $stopped\left[ s \right]\left[ a \right]$
-        to $h$.
-    \item An $Abort\left( s, h \right)$ transition, for a switcher $s$
-        and a history $h$, is enabled when there exists an abort quorum
-        $R$ such that, for every member $a$ of $R$, $stopped\left[ s
-        \right]\left[ a \right]$ is not $false$ and $h \in
-        SafeAborts\left( R \right)$, where $SafeAborts\left( R
-        \right)$ is computed using the array $stopped\left[ a
-        \right]$. Its effect is to insert  $h$ in the set $aborts$. 
-\end{compactitem}
-Compared to the rounds of the original Classic Paxos, Classic Paxos ACA
-rounds add an extra level of indirection when changing round: in the
-original rounds, the leader of the new round directly queries the
-acceptors of the preceeding round, whereas in ACA rounds, the
-switchers query the acceptors and then send their abort histories to
-the leader. However, changing round is an infrequent event, therefore
-this difference should not have much practical impact on performance.
-\begin{thm}
-  Classic Paxos ACA rounds satisfy Abortable Chain Agreement.
-\end{thm}
-\begin{proof}
-  A Classic Paxos ACA round implements the specification of Abortable
-  Chain Agreement under the refinement mapping consisting in
-  projecting the state of the round onto the variables $proposed$,
-  $inits$, $aborts$, and, for every learner $l$, $learned\left[ l
-  \right]$.
-\end{proof}
-
-\subsection{Multi-Coordinated Paxos, Chain Replication, and Ring Paxos.}
-
-Multi-Coordinated Paxos, Chain Replication, and Ring Paxos rounds have a similar
-structure as Classic Paxos rounds but differ in the way the leader is
-implemented and in the way that the histories of the leader are disseminated to
-the other acceptors.
-
-Multi-Coordinated Paxos avoids depending on a unique leader by using a
-distributed leader implementation that can make progress as long as a
-strict majority of the acceptors are correct, trading off resilience
-for a higher time complexity. Conservative ACA
-rounds can be implemented like with Classic Paxos, replacing the
-leader by the distributed leader of Multi-Coordinated Paxos.
-
-In Chain Replication, acceptors are arranged in a chain whose head is
-the leader of the round. As in Classic Paxos, abort histories and
-proposed commands are sent to the leader. However, instead of the
-leader directly sending its local history to all the other acceptors,
-the local history of the leader is
-forwarded along the chain from one acceptor to the other until it
-reaches the last acceptor, which sends the history to the learners. 
-By spreading the load more evenly on the all the servers, Chain Replication
-can achieve higher throughput that Classic Paxos.
-Conservative ACA rounds can be implemented like with Classic Paxos,
-replacing the broadcast of the leader by the propagation of the leader's history
-along the chain of servers.
-
-Ring Paxos optimizes Chain Replication by having the leader broadcast
-commands to the acceptors using IP multicast and only sending hashes
-of the commands along the chain. Ring Paxos round can be modified to implement
-conservative ACA rounds in a similar way as for Chain Replication.
-
-\subsection{Fast Paxos}
-
-Fast Paxos is composed of two types of rounds: classic rounds, like in Classic
-Paxos, and fast rounds.
-The fast rounds can be transformed into
-implementations of fast ACA rounds. Instead of sending their
-commands to a leader, as in Classic Paxos, the proposers of fast
-rounds directly broadcast their commands to all the acceptors. As long
-as commands are received in the same order by a learn quorum of
-acceptors or are received in different orders but commute, commands can be learned in two message-propagation delays instead of three in Classic Paxos.
-However, if non-commuting commands are received in different orders by
-several acceptors, the local history of the acceptors may not form a
-chain anymore and the rounds must abort. The fast rounds of Fast
-Paxos implement fast ACA rounds.
-
-Fast Paxos ACA rounds are specified using the same set of variables as for
-Classic Paxos, initialized to the same values.
-The next-state relation is composed of the following types of
-transitions, where the definitions of learn quorums, abort quorums,
-and of the $SafeAborts\left( R \right)$ operator are the ones of fast
-ACA rounds.
-\begin{compactitem}
-    \item A $Propose\left( p,c \right)$ transition, for a proposer $p$
-        and a command $c$, not guarded, inserts $c$ in the set
-        $proposed$ and broadcasts the message $\aseq{\quo{prop},c}$ to
-        all the acceptors.
-    \item An $Init\left( s, h \right)$ transition, for a switcher $s$
-        and a history $h$, not guarded, inserts $h$ in the set $inits$ and
-        broadcasts the message $\aseq{\quo{init},h}$ to all the acceptors.
-    \item A $WakeUp\left( a \right)$ transition, for an acceptor $a$,
-        is enabled when $status\left[ a \right]$ is ``idle'' and $a$
-        can receive an $\aseq{\quo{init},h}$ message. Its effect is to
-        receive the message, set $hist\left[ a \right]$ to $h$, and to
-        set $status\left[ a \right]$ to ``ready''.
-    \item An $Accept\left( a \right)$ transition, for an acceptor $a$,
-        is enabled when $status\left[ a \right]$ is ``ready'' and $a$
-        can receive a $\aseq{\quo{prop},c}$ message. The effect of the
-        action is to receive the message, to update $hist\left[ a
-        \right]$ to $hist\left[ a \right]\star c$, and to broadcast
-        the message $\aseq{\quo{ack},hist\left[ a \right]\star c}$ to
-        the learners.
-    \item The $RcvAck\left( l \right)$ transitions, for a learner $l$,
-        the $Learn\left( l, h \right)$ transitions, for a learner $l$
-        and a history $h$, the $Stop\left( s \right)$ transitions, for
-        a switcher $s$, $Stop\left( a \right)$ transitions, for an
-        acceptor $a$, the $RcvStopped\left( s \right)$ transitions,
-        for a switcher $s$, and the $Abort\left( s, h \right)$
-        transitions, for a switcher $s$ and a history $h$, which are
-        all the same as in Classic Paxos ACA rounds.
-\end{compactitem}
-
-As for Classic Paxos ACA rounds, Fast Paxos ACA rounds implement ACA
-under the refinement mapping consisting projecting the state of the
-round onto the variables $proposed$, $inits$, $aborts$, and, for every
-learner $l$, $learned\left[ l \right]$.
-
-\newpage
-
-\printbibliography
-
-\begin{appendix}
-
-  \label{sec:isaproofs}
-
-%  \includepdf[pages={1},pagecommand={\section{Mechanical Proof of the
-%  Composition Theorem in Isabelle/HOL}\label{sec:isaproofs}},scale=1]{../IsabellePDFS/podctheories.pdf}
-%  \includepdf[pages={2-},pagecommand={},scale=1]{../IsabellePDFS/podctheories.pdf}
-
-\end{appendix}
-
-\end{document}
-    We present a new abstraction, Abortable Generalized Consensus, which enables the
-    correct-by-construction composition of heterogeneous State-Machine
-    Replication algorithms.
-
-    integrating state of the art techniques to optimize executions.  ACA is the most complete approach to date enabling the
-    correct-by-construction composition of heterogeneous State-Machine
-    Replication algorithms. We provide two specifications satisfying
-    ACA and we illustrate their power by refining them to
-    obtain a versatile adaptive algorithm that combines five existing
-    algorithms into a single one. ACA is completely formalized and its
-    underlying composition theorem has been mechanically checked by an interactive
-    proof assistant.
-}
-
-\vspace{2cm}
-
-\begin{center}
-This paper is a regular submission.
-
-\vspace{1cm}
-This paper is a student paper.
-\end{center}
-
-\newpage
-
-\section{Introduction}
-
-The State-Machine Replication technique (abbreviated SMR) allows
-building reliable services on top of unreliable hardware. The basic idea is to 
-replicate a service over several servers, each executing the same
-sequence of (deterministic) commands. 
-Making SMR algorithms efficient has been a long-standing
-challenge and a multitude of algorithms have been proposed. Examples
-include Fast Paxos \cite{Lamport06FastPaxos}, Disk Paxos
-\cite{GafniLamport03DiskPaxos}, Chain Replication
-\cite{RenesseSchneider04ChainReplicationSupportingHighThroughputAvailability},
-Ring Paxos
-\cite{MarandiETAL10RingPaxosHighthroughputAtomicBroadcastProtocol},
-Egalitarian Paxos
-\cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments},
-Multi-Coordinated Paxos
-\cite{CamargosSchmidtPedone07MulticoordinatedPaxos}, Cheap Paxos
-\cite{LamportMassa04CheapPaxos}, Vertical Paxos
-\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication},
-Paxos-MIC
-\cite{HurfinMoiseNarzul11AdaptiveFastPaxosMakingQuickEverlasting},
-Mencius
-\cite{MaoJunqueiraMarzullo08MenciusBuildingEfficientReplicatedStateMachine},
-and Fast Mencius \cite{WeiETAL13FastMenciusMenciusLowCommitLatency}.
-
-Each of the existing algorithms is best suited to a particular range
-of environments, characterized, e.g., by client load, network
-topology, likelihood of faults, communication latency and bandwidth, processing
-power of nodes, availability of disk storage and memory.
-However, the environment faced by real systems evolves over time:
-hardware is upgraded, new data-centers are built, the popularity of a
-service changes, etc.  Consistent performance therefore comes with
-adaptation to change.  Unfortunately, existing SMR algorithms cannot easily be
-stopped and replaced by another algorithm on the fly. Moreover,
-devising a new algorithm that can change behavior dynamically is hard:
-dynamically switching from one SMR algorithm to another requires
-synchronizing notoriously intricate algorithms.
-
-We propose a new abstraction, \emph{Abortable Chain Agreement}
-(abbreviated ACA), which subsumes SMR and whose implementations,
-called \emph{ACA rounds}, can, by construction, pass the baton from
-one to another on the fly and efficiently. Moreover, optimizing a replicated
-service consisting of several ACA rounds becomes just a matter of devising a new
-ACA round, which can then be plugged-in on the fly: any new ACA round
-is compatible with every existing ACA round by construction.
-
-Abortable Chain Agreement is the \emph{most complete framework to date} enabling
-correct by construction composition of heterogeneous rounds.  
-ACA is completely formalized and its underlying composition theorem has been
-mechanically checked by an interactive proof assistant.  
-Compared to ACA,
-existing approaches
-\cite{BarNoyETAL87ShiftingGearsChangingAlgorithmsFlyTo,GuerraouiETAL10Next700BftProtocols,GuerraouiKuncakLosa12SpeculativeLinearizability}
-are limited. In particular, they do not allow the disorderly execution of
-commuting and read-only commands, an important practical optimization
-which is at the root of recent work on SMR like Generalized Consensus
-\cite{Lamport05GeneralizeConsensus} and Generalized Lattice Agreement
-\cite{FalerioETAL12GeneralizedLatticeAgreement}.
-
-\begin{comment}
-On top of the traditional roles of proposer, acceptor, and learners
-\cite{lamport2001paxos} used in consensus algorithms, ACA introduces
-the notion of \emph{switchers} and of \emph{scheduling policy}. The switchers
-have the task of aborting a running ACA round and initializing a new
-ACA round in a safe and live manner.  The \emph{scheduling policy}
-decides when to abort an ACA round and choses the next ACA round to
-execute. It must ensure that all switchers agree on the next ACA round
-to initialize.
-\end{comment}
-
-We observe that, because of the FLP impossibility result
-\cite{FischerLynchPaterson83ImpossibilityDistributedConsensusOneFaultyProcess},
-most existing SMR algorithms are structured in a sequence of abortable
-rounds which can easily be transformed into ACA rounds. To simplify
-this task even further, we propose two specifications, \emph{conservative
-ACA rounds} and \emph{fast ACA rounds}, which, as we show, can be concretized
-to obtain ACA versions of the rounds of five existing algorithms: 
-Classic Paxos, Fast Paxos, Chain Replication, Ring Paxos,
-and Multi-Coordinated Paxos. ACA guarantees that the composition of the five ACA
-rounds is correct by construction, forming a versatile adaptive replication
-algorithm.
-
-The rest of the paper is structured as follows.
-\Cref{sec:related} surveys recent related work. \Cref{sec:ca} presents Chain
-Agreement, which embodies the requirements for implementing SMR. \Cref{sec:aca}
-presents Abortable Chain Agreement and its optimizations to treat commuting
-and read-only commands efficiently. \Cref{sec:rounds} presents our specification
-of conservative and fast ACA rounds, which implement ACA. In
-\cref{sec:examples}, we show that the rounds of  
-Classic Paxos, Fast Paxos, Chain Replication, Ring Paxos,
-and Multi-Coordinated Paxos can easily be transformed into ACA implementations
-by using our specifications of conservative and fast ACA rounds.
-Finally, in \cref{sec:isaproofs}, we provide a complete formalization of ACA
-and a mechanical proof of the \emph{composition theorem} (\cref{thm:comp}),
-checked by the interactive proof assistant Isabelle/HOL \cite{NipkowPaulsonWenzel02IsabelleHOL}.
-
-\begin{comment}
-Abortable Chain Agreement builds on several recent advances in SMR
-algorithms: ACA allows acceptors to only eventually reach agreement,
-like in Generalized Paxos \cite{Lamport05GeneralizeConsensus},
-boosting performance when most commands commute; ACA allows fast
-execution of read commands like in Generalized Lattice Agreement
-\cite{FalerioETAL12GeneralizedLatticeAgreement}; ACA implementations
-can tolerate $f$ faults with only $f+1$ replicas and replace crashed replicas
-upon changing round, as in Vertical Paxos  \cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}.  
-\end{comment}
-
-\section{Related Work}
-\label{sec:related}
-
-SMR is often implemented using a sequence of instances of Consensus.
-There are numerous examples of consensus algorithms in the literature.
-However, Abortable Chain Agreement is based on Generalized Consensus. 
-In Generalized Consensus \cite{Lamport05GeneralizeConsensus}, the learners of a
-SMR algorithm repeatedly learn new values belonging to a \emph{C-struct set}. A
-C-struct can be seen as a data structure encoding  sequences of commands modulo
-the ordering of commands that commute. The advantage of Generalized
-Consensus over Consensus is that
-two servers which receive commuting commands in different orders can still reach
-agreement on C-structs without further synchronization. 
-Our Abortable Chain Agreement can also use C-structs, with the same advantages
-as Generalized Consensus. For simplicity, the presentation in this
-paper uses the more concrete notion of \emph{command history} (see
-\cref{sec:histories}).
-
-Generalized Lattice Agreement \cite{FalerioETAL12GeneralizedLatticeAgreement} allows more efficient implementations of SMR when
-commands are partitioned into read-only commands and write-only commands.
-Abortable Chain Agreement supports the same optimizations as Generalized Lattice Agreement.
-Notably, read-only commands are treated specially so as to optimize their
-executions (see \cref{sec:readonly}).
-
-Vertical Paxos algorithms 
-\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication} are SMR algorithms
-whose set of replica servers can be reconfigured without stopping to process new
-commands. Crashed servers can therefore be replaced by new servers, ensuring the
-long-term availability of a replicated service.
-Vertical Paxos algorithms use an external reconfiguration master, itself
-implemented as a replicated state machine, which determines the servers that should
-be removed from or added to the configuration.
-ACA shares some aspects of Vertical Paxos algorithms: the task of the
-reconfiguration master is similar to the task of the scheduling policy
-component of ACA (see \cref{sec:aca}).
-However, in ACA, changing round can be used not only to replace crashed servers
-but also to change the algorithm that the servers are running. 
-Changing algorithm is not possible in Vertical Paxos.
-In both Vertical Paxos and ACA, a reliable reconfiguration master, resp. a
-reliable scheduling policy, allows tolerating $f$ crashes with $f+1$ servers.
-
-The Abstract Framework \cite{GuerraouiETAL10Next700BftProtocols} proposes to
-build Byzantine Fault-Tolerant algorithms as the succession of abortable rounds
-called Abstract instances. Abstract instances must satisfy the Abstract
-correctness properties. By construction, the composition of any number of
-Abstract instances is a correct SMR implementation. However, the Abstract
-Framework uses totally ordered sequence of commands, making difficult the
-optimization of the execution of commuting and read-only commands that are
-possible with ACA.  Moreover, the Abstract Framework does not identify the
-crucial role of the scheduling policy for the long-term resilience of a service.
-Abortable Chain Agreement combines the ideas of the Abstract Framework,
-Generalized Paxos, Generalized Lattice Agreement, and Vertical Paxos in a
-unified abstraction.
-
-Finally, the Speculative Linearizability framework
-\cite{GuerraouiKuncakLosa12SpeculativeLinearizability} was a first
-attempt by the authors of the present paper at generalizing the
-Abstract Framework and mechanically proving its properties with the
-Isabelle/HOL interactive
-proofs assistant.  Like the Abstract Framework, the
-Speculative Linearizability framework is based on totally ordered
-sequences of commands, with the same drawback for the execution of
-commuting and read-only commands. Moreover, Abortable Chain Agreement is both
-conceptually simpler and subsumes the Speculative Linearizability
-property.  Like Speculative Linearizability, Abortable Chain Agreement
-and its main property, the composition theorem (\cref{thm:comp}), are
-formalized in Isabelle/HOL (see \cref{sec:isaproofs}).
-
-Abortable Chain Agreement is the most complete framework to date for building
-adaptive replication algorithms out of heterogeneous rounds, making state of the
-art optimization techniques available to algorithm designers, which was not
-possible before.
-
-\section{Chain Agreement}
-\label{sec:ca}
-
-We consider a set of asynchronous servers communicating by message passing in an
-asynchronous network. Servers are not necessarily sequential processes. We
-describe a behavior of the system using an infinite sequence of states, where a
-state is a function from variable names to values (i.e. a valuation
-of variables).  We specify sets of allowed behaviors using an initial predicate,
-a next-state relation (using unprimed and primed variables) and a fairness
-condition, all in the style of TLA+ \cite{Lamport02SpecifyingSystems}. However,
-for succinctness, our specification are rather informal, eluding non-essential
-details.
-
-Our specifications all contain by default a state variable $crashed$ and
-$Crash\left( x \right)$ transition which is always enabled. A server $x$ faithfully
-executes the algorithm that is assigned to it until its identifier is added to
-the variable $crashed$ by a $Crash\left( x \right)$ transitions.  When the identifier of a
-server is in the set $crashed$, all its actions are disabled.  In an execution,
-a server that is never added to the set $crashed$ is said correct.
-
-We consider a computing service exposing a set of commands $C$. The
-service has a sequential specification consisting of an initial state
-and a deterministic transition relation in which every command
-atomically updates the state of the service and produces an output. We
-assume that duplicate commands have no effect on the state of the
-service and return the same output as their first occurrence.
-
-In a traditional SMR algorithm, the servers can be seen has playing one or
-several of three roles: \emph{proposers}, which propose new commands to execute,
-\emph{acceptors}, which accept and order commands in a sequence, and
-\emph{learners}, which learn about the growing sequence of commands to
-execute \cite{lamport2001paxos}. Often, but not necessarily, every replica server plays the
-three roles of proposer, acceptor, and learner: clients of the service
-send their commands to some replica servers; replica servers submit
-the received commands to the SMR protocol in the role of proposers;
-replica servers respond to commands by executing the commands learned
-in the role of learners; replica servers participate in command
-ordering in the role of acceptors.  An SMR algorithm guarantees that
-the replicas learn the same sequences of commands, possibly up to the
-reordering of commuting commands. Therefore, every learner can respond to client
-requests using its last learned sequence of commands, and clients will receive
-the same responses as the ones produces by a centralized, non-replicated,
-service.
-
-Chain Agreement precisely specifies the allowed behaviors of the roles of \emph{proposers}
-and \emph{learners}, abstracting over the behavior of acceptors.
-
-\begin{comment}
-We suppose that each server can play a number of \emph{roles}.
-Chain Agreement specifies the allowed behaviors the roles of \emph{proposers}
-and \emph{learners}. Proposers receive commands from the clients of the service
-and propose them for execution by the replicated service. Each learner learns about
-a growing sequence of commands. Thanks to the sequential specification of the
-service, a sequence of commands determines an output for each the included commands. 
-In a traditional SMR algorithm, servers can also play the role of
-\emph{acceptors}, whose task is to collectively order commands and let the
-learners know about the latest sequence of ordered commands.
-
-The point of Chain Agreement is to implement a replicated service by having every learner learn about a growing sequence of commands, which
-represent an execution of the sequential specification of the service.
-Crucially, at all times, the sequences of commands learned by the
-different learners must form a chain in the prefix order on sequences.
-Moreover, for every command $c$ proposed by a correct proposer, every
-correct learner must eventually learn a sequence containing $c$.  Once
-a learner learns about a sequence containing a command $c$, it can
-determine the output produced by $c$, using the sequential
-specification of the service, and respond to the client that issued
-the command $c$.
-\end{comment}
-
-To specify Chain Agreement, we need the notion of ordering of sequences and of
-maximum sequence.  The prefix order on sequences is denoted by the operator
-$\leq$ ($s_1 \leq s_2$ means that $s_1$ is a prefix of $s_2$). 
-%Using the operator $\leq$, we define $\geq$, $<$, and $>$ as expected. 
-Note that $\leq$ is a partial order on sequences. The maximum over a set of
-sequences $S$ is denoted by $Max\left( S \right)$ and $Max\left( \aset{}
-\right)$ is the empty sequence.
-
-We specify Chain Agreement using the state variables $crashed$, $proposed$, containing the proposed commands,
-and, for every learner $l$,  $learned\left[ l \right]$, a set of
-sequences of commands. Initially, $crashed$, $proposed$ and, for every learner $l$,
-$learned\left[ l \right]$ are empty sets. 
-Remember that any transition involving a server $x$ is always disabled when $x\in crashed$.
-
-The next-state relation consists of three types of transitions:
-$Crash\left( x \right)$ transitions; $Propose\left( l, c
-\right)$ transitions, for a learner $l$ and a command $c$, in which
-$c$ is inserted in the set $proposed$; $Learn\left( l,cs \right)$
-transitions, for a learner $l$ and a sequence of commands $cs$, in
-which $cs$ is inserted in $learned\left[ l \right]$. 
-A $Learn\left( l, cs \right)$ transition is guarded by the following
-conditions:
-\begin{compactitem}
-    \item[\textbf{Agreement}:] The elements of the set $\aset{learned'\left[ l
-      \right] : l \in Learners}$, the learned sequences after the transition,
-      form a chain in the partial order $\leq$.
-    \item[\textbf{Validity}:] Every learned sequence contains only
-        proposed commands.  
-    \item[\textbf{Irrevocability}:] The new learned sequence
-        $cs$ is a strict extension of $Max\left(learned\left[ l
-        \right]\right)$. 
-        %\\ $cs > Max\left(learned\left[ l \right]\right)$.
-\end{compactitem}
-Finally, we add, for every learner $l$, one weak fairness condition
-\cite{Lamport02SpecifyingSystems} on the transitions $Learn\left( l,
-cs \right)$. The weak fairness conditions ensure that every correct
-learner eventually learns about every proposed command.
-
-Thanks to the properties of Chain Agreement, the clients, observing
-their input commands and the received responses, see a
-\emph{linearizable} execution
-\cite{HerlihyWing90LinearizabilityCorrectnessConditionConcurrentObjects}.
-Moreover, the service remains available as long as at least one
-proposer and one learner are correct.
-
-\section{Abortable Chain Agreement}
-\label{sec:aca}
-
-We now consider implementing Chain Agreement as a sequence of \emph{Abortable Chain
-Agreement rounds}, where each round makes progress ordering requests or aborts
-its execution and passes the baton to the next round in the sequence.
-An algorithm composed of a sequence of ACA rounds is called an \emph{ACA
-algorithm}.
-
-On top of the traditional roles of proposer, acceptor, and learner, ACA
-introduces the new role of \emph{switcher} and a new component, the
-\emph{scheduling policy}. In ACA, the assignment of the roles of proposer,
-acceptor, learner, and switcher can change in each rounds, allowing for
-reconfiguration. 
-In normal circumstances, the proposers, acceptors,
-and learners of an ACA round work together to implement Chain Agreement.
-However, if the environment becomes unfavorable to a running ACA round, the
-scheduling policy can instruct the round to abort its execution and pass the
-baton, using the switchers, to a new round.
-
-A switcher of an aborting round determines, from the state of the
-acceptors, an \emph{abort sequence} that it transfers to one or several
-switchers of the new round.  Upon receiving an abort sequence, a switcher of the
-new round initializes the acceptors of the new round, which can then
-proceed implementing SMR, until the new round itself aborts and the
-process is repeated. 
-We assume that in the first round, which has no previous round, every 
-switchers receives the empty abort sequence.
-
-The scheduling policy determines when to abort an ACA round and which new ACA
-round should take over. The scheduling policy must inform the switchers about the
-identity of the next round and must guarantee that all switchers receive the
-same information.  In this paper we abstract over the scheduling policy, assuming
-that a round that aborts passes the baton a nondeterministically chosen new
-round. A reliable scheduling-policy component can realistically be implemented
-as a replicated state machine running on a large number of servers otherwise
-dedicated to other tasks
-\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}. We will see
- in \cref{sec:rounds} that, as in Vertical Paxos, a reliable scheduling policy component allows
-tolerating $f$ faults with $f+1$ replicas, while maintaining the long-term
-resilience of the service by replacing faulty servers upon round change.
-
-The clients of an ACA algorithm can send their commands to the proposers of any round (in
-practice, clients send their commands to proposers of the largest
-active round) and servers can use sequences learned in any round
-to determine outputs.
-
-ACA specifies the allowed behaviors of the proposers, learners, and switchers
-\emph{of a single round}. Similarly to
-Chain Agreement, we use the state variables $crashed$, $proposed$,
-and, for every learner $l$, $learned\left[ l \right]$.  Additionally,
-we use the state variables $inits$, containing the abort sequences
-received from the previous round, and $aborts$, containing the abort
-sequences transfered to the next round.
-Two consecutive ACA rounds are composed by letting the $aborts$ variable of the
-first round be the $inits$ variable of the second round.
-
-Initially, $crashed$, $proposed$, for every
-learner $l$, $learned\left[ l \right]$, $inits$, and $aborts$ are all empty sets.
-The next-state relation consists of five types of transitions. The
-first three, $Crash\left( x \right)$, $Propose\left( l, c
-\right)$, and $Learn\left( l, cs \right)$, for a learner, proposer, or
-switcher $x$, for a learner $l$, and for a sequence of commands
-$cs$, have the same effect as in Chain Agreement.
-ACA introduces two new transitions, $Init\left(
-s, cs \right)$ and $Abort\left( s, cs \right)$, for a switcher $s$ and a sequence of
-commands $cs$. The transition $Init\left( s, cs \right)$ adds $cs$ to the set $inits$, and models a switcher receiving the
-abort sequence $cs$ from the preceding round.
-The transition $Abort\left( s, cs \right)$ adds $cs$ to the set $aborts$ and models a switcher transferring the abort sequence $cs$ to a switcher of the next round.
-All the transition are guarded by the following conditions:
-\begin{compactitem}
-    \item[\textbf{Agreement}:] The elements of the set $\aset{learned'\left[ l
-      \right] : l \in Learners}$ form a chain in the partial order.
-    \item[\textbf{Validity}:] Every abort sequence or learned sequence
-        is of the form $h \star cs$, where $cs$ contains only commands
-        of the set $proposed$ and $h$ is the greatest lower bound of a
-        nonempty subset of $inits$.
-    \item[\textbf{Initialization}:] If $inits$ is empty, then no
-        sequence can be learned and no switcher can abort.
-    \item[\textbf{Irrevocability}:] In the case of a $Learn\left( l,cs
-        \right)$ transition, the new learned sequence $cs$ is a strict
-        extension of $Max\left(learned\left[ l \right]\right)$.
-        %, $cs
-        %\geq Max\left(learned\left[ l \right]\right)$.
-    \item[\textbf{Safe Abort}:] Every abort sequence is larger in the
-        partial order $\leq$ than every learned sequence.
-\end{compactitem}
-Note that Agreement and Irrevocability are the same as in Chain
-Agreement.
-
-Finally, for every learner $l$, we add a weak fairness condition on the
-following state transition: 
-\begin{equation}
-    \exists  s \in Switchers, cs \in Sequence\left( C \right) : Learn\left( l, cs \right) \vee Abort\left( s,cs \right)
-\end{equation}
-The weak fairness conditions ensure that when a correct learner has not
-learned a sequence containing all commands proposed in the round, then
-it eventually learns a new command or, if there is at least one correct switcher, at least one switch value is transfered to the next round. 
-Therefore, if a client has submitted a command to a proposer, eventually,
-either the client gets a response from the current round, or
-the next round is initialized with an abort sequence and the
-command can be resubmitted to the next round.
-
-\subsection{The Composition Theorem}
-
-Consider the succession of $n$ rounds numbered $1$ to $n$.  To
-distinguish the variables of each round, we annotate the local
-variables of each round with the number of the round. For example, the
-variables of the seventh round are written $proposed_7$, for every
-learner $l$, $learned_7\left[ l \right]$, $inits_7$, and $aborts_7$.
-
-Moreover, we assume that a command, which is proposed to a round and that
-is not learned before the first abort sequence is transfered to the
-next round, is proposed anew to the next round.
-
-Define the state variable $proposed$ as the union over all rounds $i$
-of the $proposed_i$ variables and, for every learner $l$, 
-define the $learned\left[ l \right]$ variable as the union over all
-rounds $i$ of the $learned_i\left[ l \right]$ variables.
-
-\begin{thm}[Composition Theorem]
-    \label{thm:comp}
-    In every execution of a succession of ACA rounds, the sequence of valuations
-    of the variable $proposed$ and, for every learner $l$, of the variables
-    $learned\left[ l \right]$ satisfies Chain Agreement.
-\end{thm}
-\Cref{thm:comp} can be shown by induction on the number of rounds,
-using the following lemma.
-\begin{lem}
-  \label{lem:comp}
-    Consider two rounds numbered $i$ and $i+1$. In every execution of a
-    succession of ACA rounds, the sequence of valuations of the variables $inits
-    = inits_i$, $propose = proposed_i \cup propose_{i+1}$, for every learner
-    $l$, $learned\left[ l \right] = learned_i\left[ l \right] \cup
-    learned_{i+1}\left[ l \right]$and $aborts = aborts_{i+1}$ satisfies ACA.
-\end{lem}
-\begin{proof}
-  A mechanically-checked proof in Isabelle/HOL appears in \cref{sec:isaproofs}.
-  The proofs uses a version of the specification of ACA that has a more
-  executable form that presented in this section. Moreover,
-  instead of sequences of commands, the Isabelle/HOL specification uses the more
-  general notion of command history, presented in the next subsection.
-  The Isabelle/HOL formalization of ACA uses the theory of I/O-automata, which
-  is also formalized in the Isabelle/HOL development. The proof uses a refinement
-  mapping from the composition of two consecutive ACA instances to a single ACA
-  instance.
-\end{proof}
-
-\Cref{thm:comp} guarantees that the successive composition of any
-number of ACA rounds implements Chain Agreement.
-Therefore, with ACA, one can easily build replicated services that
-dynamically adapt to the current operating conditions of the system: a
-replicated service has several types of rounds at its disposal, each being
-tailored to particular conditions. Each time the conditions change,
-the current ACA round aborts and passes the baton to a more 
-appropriate type of round, as determined by the scheduling policy. Moreover, a new type
-of ACA round can always be devised after the system is deployed and added on the fly to
-the existing types of rounds.
-
-We now describe two important practical optimizations of ACA.
-
-\subsection{Histories}
-\label{sec:histories}
-
-ACA can be implemented more efficiently by substituting \emph{command histories} for sequences of commands, as in Generalized Consensus \cite{Lamport05GeneralizeConsensus}.
-
-We say that two commands $c_1$ and $c_2$ \emph{commute} when, in every sequential
-execution of the service, reversing the order of $c_1$ and $c_2$ does
-not change any of the outputs, including those given to $c_1$ and
-$c_2$.
-
-A command history, abbreviated history, is a data structure that
-represents a set of sequences which are the same modulo the reordering
-of commuting commands, and in which duplicate commands do not appear.
-All the sequences represented by a given history contain the same
-commands and determine the same outputs for each command.  Therefore,
-to implement a replicated service, it is sufficient that learners
-learn command histories instead of sequences.  For a precise
-definition of command histories, we refer the reader to
-\cite{Lamport05GeneralizeConsensus}. Our Isabelle/HOL formalization
-(\cref{sec:isaproofs}) abstracts
-over the concrete representation of histories, assuming only their necessary
-properties. Command histories were first
-introduced in trace theory \cite{Mazurkiewicz84Semantics}.
-
-A command history can be built by appending a sequence of commands
-$cs$ to a history $h$, obtain the history $h \star cs$.  The major
-advantage that histories have over sequences is that if a sequence
-$cs_1$ is obtained by reordering the commuting commands of $cs_2$,
-then $h \star cs_1 = h \star cs_2$. Therefore, in  implementations of
-ACA, servers can agree on a history even though
-they chose different orders for some commuting commands.
-
-Command histories share the four properties of sequences that are
-necessary and sufficient to substitute histories for sequences in the
-definition of Chain Agreement and ACA.
-\begin{compactitem}
-    \item[\textbf{Distinctness}:] No duplicate command appears in a history. 
-    \item[\textbf{Ordering}:] Command histories form a partial order:
-        $h_1 \leq h_2$ when there exists a sequence $s$ such that $h_2
-        = h_1 \star s$.
-    \item[\textbf{GLB}:] Every finite set of command histories has a
-        \emph{greatest lower bound}.
-%    \item[\textbf{LUB}:] Every finite set of compatible command
-%        histories has a \emph{least upper bound}, where a set of
-%        command histories is compatible when they have a common upper
-%        bound.
-    \item[\textbf{Consistency}:] If $h_1 \leq h_2 \leq h_3$ and $h_3 =
-        h_1 \star s$ then there exists $s'$ and $s''$ where every
-        request of $s'$ and $s''$ appears in $s$ and such that $h_2 =
-        h_1 \star s'$ and $h_3 = h_2 \star s''$.
-\end{compactitem}
-The optimized version of ACA is obtained by
-simply substituting histories for sequences in the definition of ACA.
-The Ordering, GLB, and Consistency properties of histories ensure that
-\cref{thm:comp} still holds: the Isabelle/HOL proof of \cref{thm:comp} (see
-\cref{sec:isaproofs}) relies only on the four properties above.
-
-\subsection{Optimizing the Execution of Read-Only Commands}
-\label{sec:readonly}
-
-Chain Agreement can be further optimized by treating 
-read-only commands specially, as in Generalized Lattice Agreement
-\cite{FalerioETAL12GeneralizedLatticeAgreement}.
-
-We consider the case in which a set of servers each play the four
-roles of proposer, acceptor, learner, and switcher.
-Clients sends their commands to any server, which proposes them as a
-proposer, and records the command as pending.
-%When a server learns a history that contains one of its pending
-%commands, it uses the history to determine the response to the pending
-%command and forwards the response to the appropriate client.
-
-As observed in \cite{FalerioETAL12GeneralizedLatticeAgreement}, read-only
-commands can be executed faster by directly reading from the last learned
-history, instead of proposing the read commands and waiting for it to appear in
-a future learned history.  Because learned histories form a chain, all the reads
-return consistent values and this approach yields a \emph{serializable}
-implementation of the service.
-
-However linearizability is not satisfied if reads are served directly
-from the last learned history: a lagging replica server could respond
-to a read using a history which does not contain all completed update
-commands. To make the executions linearizable without resorting to
-proposing read-only commands as normal commands, we can employ the same
-technique as in \cite{FalerioETAL12GeneralizedLatticeAgreement}: upon
-receiving a read command, a server proposes a special no-op command
-that commutes with all other commands; the server then responds to the
-read command using the first learned history containing the no-op
-command. This ensures that all the commands that had completed before
-the read command was issued are present in the history.  Moreover,
-since no-op commands commute with all other commands, they can be
-treated more efficiently than regular read commands by the acceptors,
-like, for example, in a fast round of Fast Paxos.
-
-\section{Fast and Conservative Rounds}
-\label{sec:rounds}
-
-Most existing SMR algorithms can easily be obtained as the
-composition of one or more ACA rounds.
-
-The FLP impossibility result \cite{FischerLynchPaterson83ImpossibilityDistributedConsensusOneFaultyProcess}
-implies that no algorithm can implement SMR to the letter in practice.
-Instead, SMR algorithms guarantee safety (i.e. replica execute the
-same sequence of commands) but may not be able to make progress (i.e.
-order new commands). To make progress as likely as possible, most SMR
-algorithms are structured in consecutive rounds (also called ballots,
-views, or epochs). In each round, the algorithm tries to make progress
-but cannot be guaranteed to succeed. When the algorithm detects that
-progress is no more possible, a new round is started.  The new round
-is initialized so as to preserve safety, usually by one of the
-acceptors. For example, in Paxos, the leader of a newly started ballot
-numbered $i$ determines a value that is \emph{safe at $i$} and
-proposes it as first value to the other acceptors of the ballot $i$.
-In existing SMR algorithms, new rounds are initialized ad-hoc, which
-prevents mixing rounds of two different algorithms.  Instead,
-ACA is an abstraction of rounds which establishes a common interface and contract for switching from
-one round to the next, enabling switching from any ACA round to any other.
-
-To simplify the task of expressing existing algorithms as the composition of one
-or more ACA rounds, we propose two specifications, \emph{conservative ACA
-rounds} and
-\emph{fast ACA rounds}, which can easily be concretized to obtain executable rounds.
-We show that ACA versions of Classic Paxos, Multi-Coordinated Paxos, Chain
-Replication, and Ring Paxos can be obtained by concretizing the specification of
-conservative ACA rounds, and that ACA versions of the fast rounds of Fast Paxos can be obtained by concretizing the specification of fast ACA rounds.
-
-Fast and conservative ACA rounds are two specifications, at an
-intermediate level of abstraction, which satisfy ACA. Both
-specifications specify how the state of the acceptors should be
-updated but abstract over the particular communication topology used
-to implement the state accesses: the specifications use a global state
-which is accessible by all servers in all roles. One can refine fast
-or conservative rounds by implementing the state accesses and coordination
-using the network, obtaining a concrete algorithm.
-
-Conservative and fast ACA rounds are specified using the state variables $inits$, $proposed$, for every learner $l$, $learned\left[ l \right]$
-and $aborts$, as in ACA, and, for every acceptor $a$, two new variables
-$status\left[ a \right]$, which can be either ``idle'', ``ready'', or
-``stopped'' and $hist\left[ a \right]$, the local history of the
-acceptor $a$, which can be either a history or the special value $None$.
-Initially, $inits$, $aborts$, and, for every learner
-$l$, $learned\left[ l \right]$ are empty sets, and, for every acceptor $a$,
-$status\left[ a \right]$  is ``idle'' and $hist\left[ a \right]$ is the special
-value $None$.
-
-The next-state relations of conservative and fast rounds are composed
-of the following types of transitions (and the default $Crash\left( x
-\right)$ transition).
-\begin{compactitem}
-    \item A $Propose\left( p,c \right)$ transition, for a proposer $p$
-        and a command $c$, not guarded, inserts $c$ in the set
-        $proposed$. 
-    \item An $Init\left( s, h \right)$ transition, for a switcher $s$
-        and a history $h$, not guarded, inserts $h$ in the set $inits$.
-    \item A $WakeUp\left( a \right)$ transition, for an acceptor $a$,
-        is enabled when $status\left[ a \right]$ is ``idle''. Its
-        effect is to set $hist\left[ a \right]$ to a history found in
-        $inits$ and to set $status\left[ a \right]$ to ``ready''.
-    \item An $Accept\left( a \right)$ transition, for an acceptor $a$,
-        is enabled when $status\left[ a \right]$ is ``ready'', and
-        updates $hist\left[ a \right]$ to $hist\left[ a \right]\star
-        c$, where $c$ is a proposed command. In conservative rounds,
-        the transition is additionally guarded by the condition that
-        the resulting set of local acceptor histories,
-        $\aset{hist'\left[ a \right] : a \in Acceptors}$, forms a chain
-        in the partial order $\leq$ (this is typically implemented with a
-        leader).
-    \item A $Learn\left( l, h \right)$ transition, for a learner $l$
-      and a history $h$, enabled when $h > learned\left[ l \right]$ and is
-      smaller than the GLB of the histories of a learn quorum of acceptors. The effect of
-        the transition is to set $learned\left[ l \right]$ to $h$.
-    \item A $Stop\left( a \right)$ transition, for an acceptor $a$,
-        enabled when $status\left[ a \right]$ is ``ready'', sets
-        $status\left[ a \right]$ to ``stopped'' (note that
-        $status\left[ a \right] = $ ``stopped'' prevents any further
-        ``accept'' transition of acceptor $a$).
-%    \item REALLY NEEDED? A $TimeOut\left( a \right)$ transition, for an acceptor $a$,
-%        is enabled when $status\left[ a \right]$ is ``idle'' and sets 
-%        $status\left[ a \right]$ to ``stopped'' and $hist\left[ a
-%        \right]$ to the special value $None$. 
-    \item An $Abort\left( s, h \right)$ transition, for a switcher $s$
-        and a history $h$, is enabled when there exists a abort quorum
-        $R$ such that, for every member $a$ of $R$, $status\left[ a
-        \right]$ is ``stopped'' and $h \in SafeAborts\left( R
-        \right)$. Its effect is to insert  $h$ in the set $aborts$. 
-\end{compactitem}
-
-On top of the additional guard on the $Accept$ transition present only
-in conservative rounds, conservative and fast ACA rounds differ in
-their definition of learn quorums, abort quorums, and in their
-definition of the $SafeAborts\left( R \right)$ operator.
-To concisely define read and learn quorums, we define 
-\begin{equation}
-    Q\left( f \right) = \aset{as \subseteq Acceptors :
-        Card\left(as\right) \geq \left\lfloor f\times Card\left(
-    Acceptors \right)\right\rfloor + 1}.
-\end{equation}
-
-In conservative ACA rounds, abort and learn quorums are defined such
-that the intersection between two learn quorums is nonempty and the
-intersection between a learn quorum and an abort quorum is nonempty.
-The definition leads to two interesting cases: $LearnQ = AbortQ =
-Q\left( 1/2 \right)$; $LearnQ = \aset{Acceptors}$ and $AbortQ =
-\aset{\aset{a} : a \in Acceptors}$. The operator $SafeAborts\left( R
-\right)$ is defined as the singleton set containing the maximum
-element over the local histories of the acceptors in $R$ which are not
-$None$. Note that the maximum exists because conservative rounds
-guarantee that the histories of the acceptors form a chain.
-
-In fast ACA rounds, abort and learn quorums are defined such that
-the intersection between any two learn quorums is nonempty and the
-intersection between any two learn quorums and an abort quorum is
-nonempty.  The definition leads to three interesting cases: $LearnQ =
-AbortQ = Q\left( 2/3 \right)$; $LearnQ = Q\left( 1/2 \right)$ and
-$AbortQ = Q\left( 3/4 \right)$; $LearnQ = \aset{Acceptors}$ and
-$AbortQ = \aset{\aset{a : a \in Acceptors}}$.
-The operator $SafeAborts\left( R \right)$ is defined as the set of
-maximal elements over the set $\aset{GLB\left( \aset{hist\left[ a
-\right] : a \in R \cap L \wedge hist\left[ a \right]\neq None} \right)
-: L \in LearnQ}$.
-
-The cases in which abort quorums are singleton sets, in both
-conservative and fast rounds, allow tolerating
-$f$ faults with $f+1$ servers: a round can abort as long as one
-acceptor is correct and a new round, possibly using new correct
-servers, can take over. Note that changing round cannot happen
-without the scheduling policy instructing the switchers about the
-identity of the next round, and thus tolerating $f$ faults with
-$f+1$ servers relies on the assumption of a reliable scheduling policy.
-
-Define, for every learner $l$, the history variable $\overline{learned}\left[ l \right]$ which is initialized to the empty set and which is
-updated each time a new local history is accepted by the learner $l$ by
-inserting the new local history in $\overline{learned}\left[ l \right]$ (see \cite{AbadiLamport91ExistenceRefinementMappings} for the
-definition of a history variable).
-\begin{thm}
-  \label{thm:rounds}
-    Both conservative and fast ACA rounds implement ACA under the refinement
-    mapping that consists in substituting $\overline{learned\left[ l \right]}$
-    for $learned\left[ l \right]$.
-\end{thm}
-
-\section{Examples of Concrete ACA Algorithms}
-\label{sec:examples}
-
-In this section we show how to modify the rounds of Classic Paxos,
-Multi-Coordinated Paxos, Chain Replication, Ring Paxos, and Fast Paxos to obtain
-ACA rounds. \Cref{thm:comp} then guarantee that all those different types of
-rounds can be used in the same algorithm to yield an implementation of Chain
-Agreement.
-
-\subsection{Classic Paxos}
-
-The rounds of the Classic Paxos algorithm can be modified so as to
-obtain an implementation of conservative ACA rounds.  We define
-Classic Paxos ACA rounds by specializing the specification of
-conservative ACA rounds, giving a concrete protocol for disseminating
-the proposed commands to the acceptors and enforcing that the learned
-histories form a chain.
-
-Each Classic Paxos round has a unique leader among the acceptors which
-centrally orders the proposed commands, which are accepted by the
-other acceptors in the order determined by the leader.  This protocol
-ensures that every local acceptor history is a prefix of the local
-history of the leader, therefore guaranteeing that the local acceptor histories
-form a chain.
-
-Each acceptor broadcasts each newly accepted history to the learners, which
-learn a history $h$ when they receive enough extensions of $h$ from the
-acceptors.  When instructed by the scheduling policy to abort, a
-switcher sends a stop message to the acceptors, which irrevocably stop
-updating their local history and send back to the switcher an
-acknowledgement containing their local history. A switcher
-determines an abort history once it has received acknowledgments from
-an abort quorum of acceptors $R$, using the operator $SafeAborts\left(
-R \right)$.
-
-The specification of Classic Paxos ACA rounds uses the variables of
-the specification of conservative ACA rounds, namely the variables
-$proposed$, $inits$, $aborts$, the arrays $learned\left[ l \right]$,
-$status\left[ a \right]$, and $hist\left[ a \right]$, and two
-additional arrays $accepted\left[ l \right]\left[ a \right]$ and
-$stop\left[ s \right]\left[ a \right]$, where $l$ is a learner, $a$ is
-an acceptor, and $s$ is a switcher.
-
-There is also a network component that we do not describe explicitly.
-However, the network allows any servers to communicate by sending and
-receiving messages.
-
-Initially, the variables $proposed$, $inits$, $aborts$, and, for every
-learner $l$, $learned\left[ l \right]$ are empty sets. For every
-acceptor $a$, $status\left[ a \right]$ is ``idle'' and $hist\left[ a
-\right]$ is the special value $None$. For every learner $l$ and acceptor $a$,
-$accepted\left[ l \right]\left[ a \right]$ is the special value
-$None$. For every switcher $s$ and acceptor $a$, $stopped\left[ s
-\right]\left[ a \right]$ is set to $false$.  There are initially no
-messages in the network.
-
-The next-state relation is composed of the following types of
-transitions, where the definitions of learn quorums, abort quorums,
-and of the $SafeAborts\left( R \right)$ operator are the ones of conservative
-ACA rounds.
-\begin{compactitem}
-    \item A $Propose\left( p,c \right)$ transition, for a proposer $p$
-        and a command $c$, not guarded, inserts $c$ in the set
-        $proposed$ and sends the message $\aseq{\quo{prop},c}$ to
-        the leader.
-    \item An $Init\left( s, h \right)$ transition, for a switcher $s$
-        and a history $h$, not guarded, inserts $h$ in the set $inits$ and
-        sends the message $\aseq{\quo{init},h}$ to the leader.
-    \item A $WakeUp\left( leader \right)$ transition, enabled when
-        $status\left[ leader \right]$ is ``idle'' and the leader can
-        receive a $\aseq{\quo{init},h}$ message. Its effect is to
-        receive the message, set $hist\left[ leader \right]$ to $h$,
-        to set $status\left[ leader \right]$ to ``ready'', and to
-        broadcast the message $\aseq{\quo{leader-accept},h}$ to all
-        other acceptors.
-    \item An $Accept\left( leader \right)$ transition, enabled when
-        $status\left[ leader \right]$ is ``ready'' and the leader can
-        receive a $\aseq{\quo{prop},c}$ message. The effect of the
-        action is to receive the message, to update $hist\left[ leader
-        \right]$ to $hist\left[ leader \right]\star c$, and to
-        broadcast the message $\aseq{\quo{ack},hist\left[ leader
-        \right]\star c}$ to the learners, and to broadcast the message
-        $\aseq{\quo{leader-accept},hist\left[ leader \right]\star c}$
-        to all the other acceptors.
-    \item An $Accept\left( a \right)$ transition, for an acceptor $a$
-        which is not the leader, is enabled when $status\left[ a
-        \right]$ is ``ready'' and $a$ can receive a
-        $\aseq{\quo{leader-accept},h}$ message where $h > hist\left[ a
-        \right]$. The effect of the action is to receive the message,
-        to update $hist\left[ a \right]$ to $h$, and to broadcast the
-        message $\aseq{\quo{ack},hist\left[ a \right]\star c}$ to the
-        learners.
-    \item A $RcvAck\left( l \right)$ transition, for a learner $l$, is
-        enabled when $l$ can receive a message $\aseq{\quo{ack},h}$
-        from an acceptor $a$. Its effect is to receive the message and
-        to set $accepted\left[ l \right]\left[ a \right]$ to $h$.
-    \item A $Learn\left( l, h \right)$ transition, for a learner $l$
-        and a history $h$, is enabled when $h > Max\left(
-        learned\left[ l \right] \right)$ and there is a learn quorum
-        $Q$ of acceptors such that $h = GLB\left( \aset{accepted\left[
-        l \right]\left[ a \right] : a \in Q }\right)$. Its effect is
-        to insert $h$ in $learned\left[ l \right]$.
-    \item A $Stop\left( s \right)$ transition, for a switcher $s$,
-        always enabled, broadcasts a $\aseq{\quo{stop}}$ message to all
-        the acceptors.  This transition models a switcher being
-        instructed by the scheduling policy to abort the current round.
-    \item A $Stop\left( a \right)$ transition, for an acceptor $a$,
-        enabled when $a$ can receive a $\aseq{\quo{stop}}$ message
-        from a switcher $s$. Its effect is to receive the message, set
-        $status\left[ a \right]$ to ``stopped'', and to send the
-        message $\aseq{\quo{stopped},hist\left[ a \right]}$ to the
-        switcher $s$.
-    \item A $RcvStopped\left( s \right)$ transition, for a switcher
-        $s$, enabled when $s$ can receive a $\aseq{\quo{stopped},h}$
-        message from an acceptor $a$. Its effect is to receive the
-        message and to set $stopped\left[ s \right]\left[ a \right]$
-        to $h$.
-    \item An $Abort\left( s, h \right)$ transition, for a switcher $s$
-        and a history $h$, is enabled when there exists an abort quorum
-        $R$ such that, for every member $a$ of $R$, $stopped\left[ s
-        \right]\left[ a \right]$ is not $false$ and $h \in
-        SafeAborts\left( R \right)$, where $SafeAborts\left( R
-        \right)$ is computed using the array $stopped\left[ a
-        \right]$. Its effect is to insert  $h$ in the set $aborts$. 
-\end{compactitem}
-Compared to the rounds of the original Classic Paxos, Classic Paxos ACA
-rounds add an extra level of indirection when changing round: in the
-original rounds, the leader of the new round directly queries the
-acceptors of the preceeding rounds, whereas in ACA rounds, the
-switchers query the acceptors and then send their abort histories to
-the leader. However, changing round is an infrequent event, therefore
-this difference should not have much practical impact on performance.
-\begin{thm}
-  Classic Paxos ACA rounds satisfy Abortable Chain Agreement.
-\end{thm}
-\begin{proof}
-  A Classic Paxos ACA round implements the specification of Abortable
-  Chain Agreement under the refinement mapping consisting in
-  projecting the state of the round onto the variables $proposed$,
-  $inits$, $aborts$, and, for every learner $l$, $learned\left[ l
-  \right]$.
-\end{proof}
-
-\subsection{Multi-Coordinated Paxos, Chain Replication, and Ring Paxos.}
-
-Multi-Coordinated Paxos, Chain Replication, and Ring Paxos rounds have a similar
-structure as Classic Paxos rounds but differ in the way the leader is
-implemented and in the way that the histories of the leader are disseminated to
-the other acceptors.
-
-Multi-Coordinated Paxos avoids depending on a unique leader by using a
-distributed leader implementation that can make progress as long as a
-strict majority of the acceptors are correct, trading off resilience
-for a higher time complexity. Conservative ACA
-rounds can be implemented like with Classic Paxos, replacing the
-leader by the distributed leader of Multi-Coordinated Paxos.
-
-In Chain Replication, acceptors are arranged in a chain whose head is
-the leader of the round. As in Classic Paxos, abort histories and
-proposed commands are sent to the leader. However, instead of the
-leader directly sending its local history to all the other acceptors,
-the local history of the leader is
-forwarded along the chain from one acceptor to the other until it
-reaches the last acceptor, which sends the history to the learners. 
-By spreading the load more evenly on the all the servers, Chain Replication
-can achieve higher throughput that Classic Paxos.
-Conservative ACA rounds can be implemented like with Classic Paxos,
-replacing the broadcast of the leader by the propagation of the leader's history
-along the chain of servers.
-
-Ring Paxos optimizes Chain Replication by having the leader broadcast
-commands to the acceptors using IP multicast and only sending hashes
-of the commands along the chain. Ring Paxos round can be modified to implement
-conservative ACA rounds in a similar way as for Chain Replication.
-
-\subsection{Fast Paxos}
-
-Fast Paxos is composed of two types of rounds: classic rounds, like in Classic
-Paxos, and fast rounds.
-The fast rounds can be transformed into
-implementations of fast ACA rounds. Instead of sending their
-commands to a leader, as in Classic Paxos, the proposers of fast
-rounds directly broadcast their commands to all the acceptors. As long
-as commands are received in the same order by a learn quorum of
-acceptors or are received in different orders but commute, commands can be learned in two message-propagation delays instead of three in Classic Paxos.
-However, if non-commuting commands are received in different orders by
-several acceptors, the local history of the acceptors may not form a
-chain anymore and the rounds must abort. The fast rounds of Fast
-Paxos implement fast ACA rounds.
-
-Fast Paxos ACA rounds are specified using the same set of variables as for
-Classic Paxos, initialized to the same values.
-The next-state relation is composed of the following types of
-transitions, where the definitions of learn quorums, abort quorums,
-and of the $SafeAborts\left( R \right)$ operator are the ones of fast
-ACA rounds.
-\begin{compactitem}
-    \item A $Propose\left( p,c \right)$ transition, for a proposer $p$
-        and a command $c$, not guarded, inserts $c$ in the set
-        $proposed$ and broadcasts the message $\aseq{\quo{prop},c}$ to
-        all the acceptors.
-    \item An $Init\left( s, h \right)$ transition, for a switcher $s$
-        and a history $h$, not guarded, inserts $h$ in the set $inits$ and
-        broadcasts the message $\aseq{\quo{init},h}$ to all the acceptors.
-    \item A $WakeUp\left( a \right)$ transition, for an acceptor $a$,
-        is enabled when $status\left[ a \right]$ is ``idle'' and $a$
-        can receive an $\aseq{\quo{init},h}$ message. Its effect is to
-        receive the message, set $hist\left[ a \right]$ to $h$, and to
-        set $status\left[ a \right]$ to ``ready''.
-    \item An $Accept\left( a \right)$ transition, for an acceptor $a$,
-        is enabled when $status\left[ a \right]$ is ``ready'' and $a$
-        can receive a $\aseq{\quo{prop},c}$ message. The effect of the
-        action is to receive the message, to update $hist\left[ a
-        \right]$ to $hist\left[ a \right]\star c$, and to broadcast
-        the message $\aseq{\quo{ack},hist\left[ a \right]\star c}$ to
-        the learners.
-    \item The $RcvAck\left( l \right)$ transitions, for a learner $l$,
-        the $Learn\left( l, h \right)$ transitions, for a learner $l$
-        and a history $h$, the $Stop\left( s \right)$ transitions, for
-        a switcher $s$, $Stop\left( a \right)$ transitions, for an
-        acceptor $a$, the $RcvStopped\left( s \right)$ transitions,
-        for a switcher $s$, and the $Abort\left( s, h \right)$
-        transitions, for a switcher $s$ and a history $h$, which are
-        all the same as in Classic Paxos ACA rounds.
-\end{compactitem}
-
-As for Classic Paxos ACA rounds, Fast Paxos ACA rounds implement ACA
-under the refinement mapping consisting projecting the state of the
-round onto the variables $proposed$, $inits$, $aborts$, and, for every
-learner $l$, $learned\left[ l \right]$.
-
-\newpage
-
-\printbibliography
-
-\begin{appendix}
-
-  \label{sec:isaproofs}
-
-%  \includepdf[pages={1},pagecommand={\section{Mechanical Proof of the
-%  Composition Theorem in Isabelle/HOL}\label{sec:isaproofs}},scale=1]{../IsabellePDFS/podctheories.pdf}
-%  \includepdf[pages={2-},pagecommand={},scale=1]{../IsabellePDFS/podctheories.pdf}
-
-\end{appendix}
-
-\end{document}

### 2015-05-14 paper/paper.tex
Moved the intro and roadmap to their own files.

--- 2015-05-09 paper/paper.tex
@@
 
-\section{Introduction}
-
-The State-Machine Replication technique (abbreviated SMR) allows
-building reliable services on top of unreliable hardware. The basic idea is to 
-replicate a service over several servers, each executing the same
-sequence of (deterministic) commands. 
-Making SMR algorithms efficient has been a long-standing
-challenge and a multitude of algorithms have been proposed. Examples
-include Fast Paxos \cite{Lamport06FastPaxos}, Disk Paxos
-\cite{GafniLamport03DiskPaxos}, Chain Replication
-\cite{RenesseSchneider04ChainReplicationSupportingHighThroughputAvailability},
-Ring Paxos
-\cite{MarandiETAL10RingPaxosHighthroughputAtomicBroadcastProtocol},
-Egalitarian Paxos
-\cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments},
-Multi-Coordinated Paxos
-\cite{CamargosSchmidtPedone07MulticoordinatedPaxos}, Cheap Paxos
-\cite{LamportMassa04CheapPaxos}, Vertical Paxos
-\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication},
-Paxos-MIC
-\cite{HurfinMoiseNarzul11AdaptiveFastPaxosMakingQuickEverlasting},
-Mencius
-\cite{MaoJunqueiraMarzullo08MenciusBuildingEfficientReplicatedStateMachine},
-and Fast Mencius \cite{WeiETAL13FastMenciusMenciusLowCommitLatency}.
-
-Each of the existing algorithms is best suited to a particular range
-of environments, characterized, e.g., by client load, network
-topology, likelihood of faults, communication latency and bandwidth, processing
-power of nodes, availability of disk storage and memory.
-However, the environment faced by real systems evolves over time:
-hardware is upgraded, new data-centers are built, the popularity of a
-service changes, etc.  Consistent performance therefore comes with
-adaptation to change.  Unfortunately, existing SMR algorithms cannot easily be
-stopped and replaced by another algorithm on the fly. Moreover,
-devising a new algorithm that can change behavior dynamically is hard:
-dynamically switching from one SMR algorithm to another requires
-synchronizing notoriously intricate algorithms.
-
-We propose a new abstraction, \emph{Abortable Chain Agreement}
-(abbreviated ACA), which subsumes SMR and whose implementations,
-called \emph{ACA rounds}, can, by construction, pass the baton from
-one to another on the fly and efficiently. Moreover, optimizing a replicated
-service consisting of several ACA rounds becomes just a matter of devising a new
-ACA round, which can then be plugged-in on the fly: any new ACA round
-is compatible with every existing ACA round by construction.
-
-Abortable Chain Agreement is the \emph{most complete framework to date} enabling
-correct by construction composition of heterogeneous rounds.  
-ACA is completely formalized and its underlying composition theorem has been
-mechanically checked by an interactive proof assistant.  
-Compared to ACA,
-existing approaches
-\cite{BarNoyETAL87ShiftingGearsChangingAlgorithmsFlyTo,GuerraouiETAL10Next700BftProtocols,GuerraouiKuncakLosa12SpeculativeLinearizability}
-are limited. In particular, they do not allow the disorderly execution of
-commuting and read-only commands, an important practical optimization
-which is at the root of recent work on SMR like Generalized Consensus
-\cite{Lamport05GeneralizeConsensus} and Generalized Lattice Agreement
-\cite{FalerioETAL12GeneralizedLatticeAgreement}.
-
-\begin{comment}
-On top of the traditional roles of proposer, acceptor, and learners
-\cite{lamport2001paxos} used in consensus algorithms, ACA introduces
-the notion of \emph{switchers} and of \emph{scheduling policy}. The switchers
-have the task of aborting a running ACA round and initializing a new
-ACA round in a safe and live manner.  The \emph{scheduling policy}
-decides when to abort an ACA round and choses the next ACA round to
-execute. It must ensure that all switchers agree on the next ACA round
-to initialize.
-\end{comment}
-
-We observe that, because of the FLP impossibility result
-\cite{FischerLynchPaterson83ImpossibilityDistributedConsensusOneFaultyProcess},
-most existing SMR algorithms are structured in a sequence of abortable
-rounds which can easily be transformed into ACA rounds. To simplify
-this task even further, we propose two specifications, \emph{conservative
-ACA rounds} and \emph{fast ACA rounds}, which, as we show, can be concretized
-to obtain ACA versions of the rounds of five existing algorithms: 
-Classic Paxos, Fast Paxos, Chain Replication, Ring Paxos,
-and Multi-Coordinated Paxos. ACA guarantees that the composition of the five ACA
-rounds is correct by construction, forming a versatile adaptive replication
-algorithm.
-
-The rest of the paper is structured as follows.
-\Cref{sec:related} surveys recent related work. \Cref{sec:ca} presents Chain
-Agreement, which embodies the requirements for implementing SMR. \Cref{sec:aca}
-presents Abortable Chain Agreement and its optimizations to treat commuting
-and read-only commands efficiently. \Cref{sec:rounds} presents our specification
-of conservative and fast ACA rounds, which implement ACA. In
-\cref{sec:examples}, we show that the rounds of  
-Classic Paxos, Fast Paxos, Chain Replication, Ring Paxos,
-and Multi-Coordinated Paxos can easily be transformed into ACA implementations
-by using our specifications of conservative and fast ACA rounds.
-Finally, in \cref{sec:isaproofs}, we provide a complete formalization of ACA
-and a mechanical proof of the \emph{composition theorem} (\cref{thm:comp}),
-checked by the interactive proof assistant Isabelle/HOL \cite{NipkowPaulsonWenzel02IsabelleHOL}.
+\input{intro}
+\input{roadmap}
 

### 2015-05-14 paper/roadmap.tex
Moved the intro and roadmap to their own files.

--- 2015-05-14 paper/paper.tex
@@
-\input{header.tex}
-\input{defs.tex}
-
-\addbibresource{./managed.bib}
-
-\title{Changing Flights In The Air\\ (Abortable Chain Agreement)}
-
-\author{
-  Rachid Guerraoui\\
-  \texttt{rachid.guerraoui@epfl.ch}
-  \and
-  Giuliano Losa\\
-  \texttt{giuliano.losa@epfl.ch}
-}
-
-\date{}
-
-\begin{document}
-
-\maketitle
-
-\input{abstract3}
-
-\vspace{2cm}
-
-\begin{center}
-This paper is a regular submission.
-
-\vspace{1cm}
-This paper is a student paper.
-\end{center}
-
-\newpage
-
-\input{intro}
-\input{roadmap}
-
-\begin{comment}
-Abortable Chain Agreement builds on several recent advances in SMR
-algorithms: ACA allows acceptors to only eventually reach agreement,
-like in Generalized Paxos \cite{Lamport05GeneralizeConsensus},
-boosting performance when most commands commute; ACA allows fast
-execution of read commands like in Generalized Lattice Agreement
-\cite{FalerioETAL12GeneralizedLatticeAgreement}; ACA implementations
-can tolerate $f$ faults with only $f+1$ replicas and replace crashed replicas
-upon changing round, as in Vertical Paxos  \cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}.  
-\end{comment}
-
-\section{Related Work}
-\label{sec:related}
-
-SMR is often implemented using a sequence of instances of Consensus.
-There are numerous examples of consensus algorithms in the literature.
-However, Abortable Chain Agreement is based on Generalized Consensus. 
-In Generalized Consensus \cite{Lamport05GeneralizeConsensus}, the learners of a
-SMR algorithm repeatedly learn new values belonging to a \emph{C-struct set}. A
-C-struct can be seen as a data structure encoding  sequences of commands modulo
-the ordering of commands that commute. The advantage of Generalized
-Consensus over Consensus is that
-two servers which receive commuting commands in different orders can still reach
-agreement on C-structs without further synchronization. 
-Our Abortable Chain Agreement can also use C-structs, with the same advantages
-as Generalized Consensus. For simplicity, the presentation in this
-paper uses the more concrete notion of \emph{command history} (see
-\cref{sec:histories}).
-
-Generalized Lattice Agreement \cite{FalerioETAL12GeneralizedLatticeAgreement} allows more efficient implementations of SMR when
-commands are partitioned into read-only commands and write-only commands.
-Abortable Chain Agreement supports the same optimizations as Generalized Lattice Agreement.
-Notably, read-only commands are treated specially so as to optimize their
-executions (see \cref{sec:readonly}).
-
-Vertical Paxos algorithms 
-\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication} are SMR algorithms
-whose set of replica servers can be reconfigured without stopping to process new
-commands. Crashed servers can therefore be replaced by new servers, ensuring the
-long-term availability of a replicated service.
-Vertical Paxos algorithms use an external reconfiguration master, itself
-implemented as a replicated state machine, which determines the servers that should
-be removed from or added to the configuration.
-ACA shares some aspects of Vertical Paxos algorithms: the task of the
-reconfiguration master is similar to the task of the scheduling policy
-component of ACA (see \cref{sec:aca}).
-However, in ACA, changing round can be used not only to replace crashed servers
-but also to change the algorithm that the servers are running. 
-Changing algorithm is not possible in Vertical Paxos.
-In both Vertical Paxos and ACA, a reliable reconfiguration master, resp. a
-reliable scheduling policy, allows tolerating $f$ crashes with $f+1$ servers.
-
-The Abstract Framework \cite{GuerraouiETAL10Next700BftProtocols} proposes to
-build Byzantine Fault-Tolerant algorithms as the succession of abortable rounds
-called Abstract instances. Abstract instances must satisfy the Abstract
-correctness properties. By construction, the composition of any number of
-Abstract instances is a correct SMR implementation. However, the Abstract
-Framework uses totally ordered sequence of commands, making difficult the
-optimization of the execution of commuting and read-only commands that are
-possible with ACA.  Moreover, the Abstract Framework does not identify the
-crucial role of the scheduling policy for the long-term resilience of a service.
-Abortable Chain Agreement combines the ideas of the Abstract Framework,
-Generalized Paxos, Generalized Lattice Agreement, and Vertical Paxos in a
-unified abstraction.
-
-Finally, the Speculative Linearizability framework
-\cite{GuerraouiKuncakLosa12SpeculativeLinearizability} was a first
-attempt by the authors of the present paper at generalizing the
-Abstract Framework and mechanically proving its properties with the
-Isabelle/HOL interactive
-proofs assistant.  Like the Abstract Framework, the
-Speculative Linearizability framework is based on totally ordered
-sequences of commands, with the same drawback for the execution of
-commuting and read-only commands. Moreover, Abortable Chain Agreement is both
-conceptually simpler and subsumes the Speculative Linearizability
-property.  Like Speculative Linearizability, Abortable Chain Agreement
-and its main property, the composition theorem (\cref{thm:comp}), are
-formalized in Isabelle/HOL (see \cref{sec:isaproofs}).
-
-Abortable Chain Agreement is the most complete framework to date for building
-adaptive replication algorithms out of heterogeneous rounds, making state of the
-art optimization techniques available to algorithm designers, which was not
-possible before.
-
-\section{Chain Agreement}
-\label{sec:ca}
-
-We consider a set of asynchronous servers communicating by message passing in an
-asynchronous network. Servers are not necessarily sequential processes. We
-describe a behavior of the system using an infinite sequence of states, where a
-state is a function from variable names to values (i.e. a valuation
-of variables).  We specify sets of allowed behaviors using an initial predicate,
-a next-state relation (using unprimed and primed variables) and a fairness
-condition, all in the style of TLA+ \cite{Lamport02SpecifyingSystems}. However,
-for succinctness, our specification are rather informal, eluding non-essential
-details.
-
-Our specifications all contain by default a state variable $crashed$ and
-$Crash\left( x \right)$ transition which is always enabled. A server $x$ faithfully
-executes the algorithm that is assigned to it until its identifier is added to
-the variable $crashed$ by a $Crash\left( x \right)$ transitions.  When the identifier of a
-server is in the set $crashed$, all its actions are disabled.  In an execution,
-a server that is never added to the set $crashed$ is said correct.
-
-We consider a computing service exposing a set of commands $C$. The
-service has a sequential specification consisting of an initial state
-and a deterministic transition relation in which every command
-atomically updates the state of the service and produces an output. We
-assume that duplicate commands have no effect on the state of the
-service and return the same output as their first occurrence.
-
-In a traditional SMR algorithm, the servers can be seen has playing one or
-several of three roles: \emph{proposers}, which propose new commands to execute,
-\emph{acceptors}, which accept and order commands in a sequence, and
-\emph{learners}, which learn about the growing sequence of commands to
-execute \cite{lamport2001paxos}. Often, but not necessarily, every replica server plays the
-three roles of proposer, acceptor, and learner: clients of the service
-send their commands to some replica servers; replica servers submit
-the received commands to the SMR protocol in the role of proposers;
-replica servers respond to commands by executing the commands learned
-in the role of learners; replica servers participate in command
-ordering in the role of acceptors.  An SMR algorithm guarantees that
-the replicas learn the same sequences of commands, possibly up to the
-reordering of commuting commands. Therefore, every learner can respond to client
-requests using its last learned sequence of commands, and clients will receive
-the same responses as the ones produces by a centralized, non-replicated,
-service.
-
-Chain Agreement precisely specifies the allowed behaviors of the roles of \emph{proposers}
-and \emph{learners}, abstracting over the behavior of acceptors.
-
-\begin{comment}
-We suppose that each server can play a number of \emph{roles}.
-Chain Agreement specifies the allowed behaviors the roles of \emph{proposers}
-and \emph{learners}. Proposers receive commands from the clients of the service
-and propose them for execution by the replicated service. Each learner learns about
-a growing sequence of commands. Thanks to the sequential specification of the
-service, a sequence of commands determines an output for each the included commands. 
-In a traditional SMR algorithm, servers can also play the role of
-\emph{acceptors}, whose task is to collectively order commands and let the
-learners know about the latest sequence of ordered commands.
-
-The point of Chain Agreement is to implement a replicated service by having every learner learn about a growing sequence of commands, which
-represent an execution of the sequential specification of the service.
-Crucially, at all times, the sequences of commands learned by the
-different learners must form a chain in the prefix order on sequences.
-Moreover, for every command $c$ proposed by a correct proposer, every
-correct learner must eventually learn a sequence containing $c$.  Once
-a learner learns about a sequence containing a command $c$, it can
-determine the output produced by $c$, using the sequential
-specification of the service, and respond to the client that issued
-the command $c$.
-\end{comment}
-
-To specify Chain Agreement, we need the notion of ordering of sequences and of
-maximum sequence.  The prefix order on sequences is denoted by the operator
-$\leq$ ($s_1 \leq s_2$ means that $s_1$ is a prefix of $s_2$). 
-%Using the operator $\leq$, we define $\geq$, $<$, and $>$ as expected. 
-Note that $\leq$ is a partial order on sequences. The maximum over a set of
-sequences $S$ is denoted by $Max\left( S \right)$ and $Max\left( \aset{}
-\right)$ is the empty sequence.
-
-We specify Chain Agreement using the state variables $crashed$, $proposed$, containing the proposed commands,
-and, for every learner $l$,  $learned\left[ l \right]$, a set of
-sequences of commands. Initially, $crashed$, $proposed$ and, for every learner $l$,
-$learned\left[ l \right]$ are empty sets. 
-Remember that any transition involving a server $x$ is always disabled when $x\in crashed$.
-
-The next-state relation consists of three types of transitions:
-$Crash\left( x \right)$ transitions; $Propose\left( l, c
-\right)$ transitions, for a learner $l$ and a command $c$, in which
-$c$ is inserted in the set $proposed$; $Learn\left( l,cs \right)$
-transitions, for a learner $l$ and a sequence of commands $cs$, in
-which $cs$ is inserted in $learned\left[ l \right]$. 
-A $Learn\left( l, cs \right)$ transition is guarded by the following
-conditions:
-\begin{compactitem}
-    \item[\textbf{Agreement}:] The elements of the set $\aset{learned'\left[ l
-      \right] : l \in Learners}$, the learned sequences after the transition,
-      form a chain in the partial order $\leq$.
-    \item[\textbf{Validity}:] Every learned sequence contains only
-        proposed commands.  
-    \item[\textbf{Irrevocability}:] The new learned sequence
-        $cs$ is a strict extension of $Max\left(learned\left[ l
-        \right]\right)$. 
-        %\\ $cs > Max\left(learned\left[ l \right]\right)$.
-\end{compactitem}
-Finally, we add, for every learner $l$, one weak fairness condition
-\cite{Lamport02SpecifyingSystems} on the transitions $Learn\left( l,
-cs \right)$. The weak fairness conditions ensure that every correct
-learner eventually learns about every proposed command.
-
-Thanks to the properties of Chain Agreement, the clients, observing
-their input commands and the received responses, see a
-\emph{linearizable} execution
-\cite{HerlihyWing90LinearizabilityCorrectnessConditionConcurrentObjects}.
-Moreover, the service remains available as long as at least one
-proposer and one learner are correct.
-
-\section{Abortable Chain Agreement}
-\label{sec:aca}
-
-We now consider implementing Chain Agreement as a sequence of \emph{Abortable Chain
-Agreement rounds}, where each round makes progress ordering requests or aborts
-its execution and passes the baton to the next round in the sequence.
-An algorithm composed of a sequence of ACA rounds is called an \emph{ACA
-algorithm}.
-
-On top of the traditional roles of proposer, acceptor, and learner, ACA
-introduces the new role of \emph{switcher} and a new component, the
-\emph{scheduling policy}. In ACA, the assignment of the roles of proposer,
-acceptor, learner, and switcher can change in each rounds, allowing for
-reconfiguration. 
-In normal circumstances, the proposers, acceptors,
-and learners of an ACA round work together to implement Chain Agreement.
-However, if the environment becomes unfavorable to a running ACA round, the
-scheduling policy can instruct the round to abort its execution and pass the
-baton, using the switchers, to a new round.
-
-A switcher of an aborting round determines, from the state of the
-acceptors, an \emph{abort sequence} that it transfers to one or several
-switchers of the new round.  Upon receiving an abort sequence, a switcher of the
-new round initializes the acceptors of the new round, which can then
-proceed implementing SMR, until the new round itself aborts and the
-process is repeated. 
-We assume that in the first round, which has no previous round, every 
-switchers receives the empty abort sequence.
-
-The scheduling policy determines when to abort an ACA round and which new ACA
-round should take over. The scheduling policy must inform the switchers about the
-identity of the next round and must guarantee that all switchers receive the
-same information.  In this paper we abstract over the scheduling policy, assuming
-that a round that aborts passes the baton a nondeterministically chosen new
-round. A reliable scheduling-policy component can realistically be implemented
-as a replicated state machine running on a large number of servers otherwise
-dedicated to other tasks
-\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}. We will see
- in \cref{sec:rounds} that, as in Vertical Paxos, a reliable scheduling policy component allows
-tolerating $f$ faults with $f+1$ replicas, while maintaining the long-term
-resilience of the service by replacing faulty servers upon round change.
-
-The clients of an ACA algorithm can send their commands to the proposers of any round (in
-practice, clients send their commands to proposers of the largest
-active round) and servers can use sequences learned in any round
-to determine outputs.
-
-ACA specifies the allowed behaviors of the proposers, learners, and switchers
-\emph{of a single round}. Similarly to
-Chain Agreement, we use the state variables $crashed$, $proposed$,
-and, for every learner $l$, $learned\left[ l \right]$.  Additionally,
-we use the state variables $inits$, containing the abort sequences
-received from the previous round, and $aborts$, containing the abort
-sequences transfered to the next round.
-Two consecutive ACA rounds are composed by letting the $aborts$ variable of the
-first round be the $inits$ variable of the second round.
-
-Initially, $crashed$, $proposed$, for every
-learner $l$, $learned\left[ l \right]$, $inits$, and $aborts$ are all empty sets.
-The next-state relation consists of five types of transitions. The
-first three, $Crash\left( x \right)$, $Propose\left( l, c
-\right)$, and $Learn\left( l, cs \right)$, for a learner, proposer, or
-switcher $x$, for a learner $l$, and for a sequence of commands
-$cs$, have the same effect as in Chain Agreement.
-ACA introduces two new transitions, $Init\left(
-s, cs \right)$ and $Abort\left( s, cs \right)$, for a switcher $s$ and a sequence of
-commands $cs$. The transition $Init\left( s, cs \right)$ adds $cs$ to the set $inits$, and models a switcher receiving the
-abort sequence $cs$ from the preceding round.
-The transition $Abort\left( s, cs \right)$ adds $cs$ to the set $aborts$ and models a switcher transferring the abort sequence $cs$ to a switcher of the next round.
-All the transition are guarded by the following conditions:
-\begin{compactitem}
-    \item[\textbf{Agreement}:] The elements of the set $\aset{learned'\left[ l
-      \right] : l \in Learners}$ form a chain in the partial order.
-    \item[\textbf{Validity}:] Every abort sequence or learned sequence
-        is of the form $h \star cs$, where $cs$ contains only commands
-        of the set $proposed$ and $h$ is the greatest lower bound of a
-        nonempty subset of $inits$.
-    \item[\textbf{Initialization}:] If $inits$ is empty, then no
-        sequence can be learned and no switcher can abort.
-    \item[\textbf{Irrevocability}:] In the case of a $Learn\left( l,cs
-        \right)$ transition, the new learned sequence $cs$ is a strict
-        extension of $Max\left(learned\left[ l \right]\right)$.
-        %, $cs
-        %\geq Max\left(learned\left[ l \right]\right)$.
-    \item[\textbf{Safe Abort}:] Every abort sequence is larger in the
-        partial order $\leq$ than every learned sequence.
-\end{compactitem}
-Note that Agreement and Irrevocability are the same as in Chain
-Agreement.
-
-Finally, for every learner $l$, we add a weak fairness condition on the
-following state transition: 
-\begin{equation}
-    \exists  s \in Switchers, cs \in Sequence\left( C \right) : Learn\left( l, cs \right) \vee Abort\left( s,cs \right)
-\end{equation}
-The weak fairness conditions ensure that when a correct learner has not
-learned a sequence containing all commands proposed in the round, then
-it eventually learns a new command or, if there is at least one correct switcher, at least one switch value is transfered to the next round. 
-Therefore, if a client has submitted a command to a proposer, eventually,
-either the client gets a response from the current round, or
-the next round is initialized with an abort sequence and the
-command can be resubmitted to the next round.
-
-\subsection{The Composition Theorem}
-
-Consider the succession of $n$ rounds numbered $1$ to $n$.  To
-distinguish the variables of each round, we annotate the local
-variables of each round with the number of the round. For example, the
-variables of the seventh round are written $proposed_7$, for every
-learner $l$, $learned_7\left[ l \right]$, $inits_7$, and $aborts_7$.
-
-Moreover, we assume that a command, which is proposed to a round and that
-is not learned before the first abort sequence is transfered to the
-next round, is proposed anew to the next round.
-
-Define the state variable $proposed$ as the union over all rounds $i$
-of the $proposed_i$ variables and, for every learner $l$, 
-define the $learned\left[ l \right]$ variable as the union over all
-rounds $i$ of the $learned_i\left[ l \right]$ variables.
-
-\begin{thm}[Composition Theorem]
-    \label{thm:comp}
-    In every execution of a succession of ACA rounds, the sequence of valuations
-    of the variable $proposed$ and, for every learner $l$, of the variables
-    $learned\left[ l \right]$ satisfies Chain Agreement.
-\end{thm}
-\Cref{thm:comp} can be shown by induction on the number of rounds,
-using the following lemma.
-\begin{lem}
-  \label{lem:comp}
-    Consider two rounds numbered $i$ and $i+1$. In every execution of a
-    succession of ACA rounds, the sequence of valuations of the variables $inits
-    = inits_i$, $propose = proposed_i \cup propose_{i+1}$, for every learner
-    $l$, $learned\left[ l \right] = learned_i\left[ l \right] \cup
-    learned_{i+1}\left[ l \right]$and $aborts = aborts_{i+1}$ satisfies ACA.
-\end{lem}
-\begin{proof}
-  A mechanically-checked proof in Isabelle/HOL appears in \cref{sec:isaproofs}.
-  The proofs uses a version of the specification of ACA that has a more
-  executable form that presented in this section. Moreover,
-  instead of sequences of commands, the Isabelle/HOL specification uses the more
-  general notion of command history, presented in the next subsection.
-  The Isabelle/HOL formalization of ACA uses the theory of I/O automata, which
-  is also formalized in the Isabelle/HOL. The proof uses a refinement
-  mapping from the composition of two consecutive ACA instances to a single ACA
-  instance.
-\end{proof}
-
-\Cref{thm:comp} guarantees that the successive composition of any
-number of ACA rounds implements Chain Agreement.
-Therefore, with ACA, one can easily build replicated services that
-dynamically adapt to the current operating conditions of the system: a
-replicated service has several types of rounds at its disposal, each being
-tailored to particular conditions. Each time the conditions change,
-the current ACA round aborts and passes the baton to a more 
-appropriate type of round, as determined by the scheduling policy. Moreover, a new type
-of ACA round can always be devised after the system is deployed and added on the fly to
-the existing types of rounds.
-
-We now describe two important practical optimizations of ACA.
-
-\subsection{Histories}
-\label{sec:histories}
-
-ACA can be implemented more efficiently by substituting \emph{command histories} for sequences of commands, as in Generalized Consensus \cite{Lamport05GeneralizeConsensus}.
-
-We say that two commands $c_1$ and $c_2$ \emph{commute} when, in every sequential
-execution of the service, reversing the order of $c_1$ and $c_2$ does
-not change any of the outputs, including those given to $c_1$ and
-$c_2$.
-
-A command history, abbreviated history, is a data structure that
-represents a set of sequences which are the same modulo the reordering
-of commuting commands, and in which duplicate commands do not appear.
-All the sequences represented by a given history contain the same
-commands and determine the same outputs for each command.  Therefore,
-to implement a replicated service, it is sufficient that learners
-learn command histories instead of sequences.  For a precise
-definition of command histories, we refer the reader to
-\cite{Lamport05GeneralizeConsensus}. Our Isabelle/HOL formalization
-(\cref{sec:isaproofs}) abstracts
-over the concrete representation of histories, assuming only their necessary
-properties. Command histories were first
-introduced in trace theory \cite{Mazurkiewicz84Semantics}.
-
-A command history can be built by appending a sequence of commands
-$cs$ to a history $h$, obtain the history $h \star cs$.  The major
-advantage that histories have over sequences is that if a sequence
-$cs_1$ is obtained by reordering the commuting commands of $cs_2$,
-then $h \star cs_1 = h \star cs_2$. Therefore, in  implementations of
-ACA, servers can agree on a history even though
-they chose different orders for some commuting commands.
-
-Command histories share the four properties of sequences that are
-necessary and sufficient to substitute histories for sequences in the
-definition of Chain Agreement and ACA.
-\begin{compactitem}
-    \item[\textbf{Distinctness}:] No duplicate command appears in a history. 
-    \item[\textbf{Ordering}:] Command histories form a partial order:
-        $h_1 \leq h_2$ when there exists a sequence $s$ such that $h_2
-        = h_1 \star s$.
-    \item[\textbf{GLB}:] Every finite set of command histories has a
-        \emph{greatest lower bound}.
-%    \item[\textbf{LUB}:] Every finite set of compatible command
-%        histories has a \emph{least upper bound}, where a set of
-%        command histories is compatible when they have a common upper
-%        bound.
-    \item[\textbf{Consistency}:] If $h_1 \leq h_2 \leq h_3$ and $h_3 =
-        h_1 \star s$ then there exists $s'$ and $s''$ where every
-        request of $s'$ and $s''$ appears in $s$ and such that $h_2 =
-        h_1 \star s'$ and $h_3 = h_2 \star s''$.
-\end{compactitem}
-The optimized version of ACA is obtained by
-simply substituting histories for sequences in the definition of ACA.
-The Ordering, GLB, and Consistency properties of histories ensure that
-\cref{thm:comp} still holds: the Isabelle/HOL proof of \cref{thm:comp} (see
-\cref{sec:isaproofs}) relies only on the four properties above.
-
-\subsection{Optimizing the Execution of Read-Only Commands}
-\label{sec:readonly}
-
-Chain Agreement can be further optimized by treating 
-read-only commands specially, as in Generalized Lattice Agreement
-\cite{FalerioETAL12GeneralizedLatticeAgreement}.
-
-We consider the case in which a set of servers each play the four
-roles of proposer, acceptor, learner, and switcher.
-Clients sends their commands to any server, which proposes them as a
-proposer, and records the command as pending.
-%When a server learns a history that contains one of its pending
-%commands, it uses the history to determine the response to the pending
-%command and forwards the response to the appropriate client.
-
-As observed in \cite{FalerioETAL12GeneralizedLatticeAgreement}, read-only
-commands can be executed faster by directly reading from the last learned
-history, instead of proposing the read commands and waiting for it to appear in
-a future learned history.  Because learned histories form a chain, all the reads
-return consistent values and this approach yields a \emph{serializable}
-implementation of the service.
-
-However linearizability is not satisfied if reads are served directly
-from the last learned history: a lagging replica server could respond
-to a read using a history which does not contain all completed update
-commands. To make the executions linearizable without resorting to
-proposing read-only commands as normal commands, we can employ the same
-technique as in \cite{FalerioETAL12GeneralizedLatticeAgreement}: upon
-receiving a read command, a server proposes a special no-op command
-that commutes with all other commands; the server then responds to the
-read command using the first learned history containing the no-op
-command. This ensures that all the commands that had completed before
-the read command was issued are present in the history.  Moreover,
-since no-op commands commute with all other commands, they can be
-treated more efficiently than regular read commands by the acceptors,
-like, for example, in a fast round of Fast Paxos.
-
-\section{Fast and Conservative Rounds}
-\label{sec:rounds}
-
-Most existing SMR algorithms can easily be obtained as the
-composition of one or more ACA rounds.
-
-The FLP impossibility result \cite{FischerLynchPaterson83ImpossibilityDistributedConsensusOneFaultyProcess}
-implies that no algorithm can implement SMR to the letter in practice.
-Instead, SMR algorithms guarantee safety (i.e. replica execute the
-same sequence of commands) but may not be able to make progress (i.e.
-order new commands). To make progress as likely as possible, most SMR
-algorithms are structured in consecutive rounds (also called ballots,
-views, or epochs). In each round, the algorithm tries to make progress
-but cannot be guaranteed to succeed. When the algorithm detects that
-progress is no more possible, a new round is started.  The new round
-is initialized so as to preserve safety, usually by one of the
-acceptors. For example, in Paxos, the leader of a newly started ballot
-numbered $i$ determines a value that is \emph{safe at $i$} and
-proposes it as first value to the other acceptors of the ballot $i$.
-In existing SMR algorithms, new rounds are initialized ad-hoc, which
-prevents mixing rounds of two different algorithms.  Instead,
-ACA is an abstraction of rounds which establishes a common interface and contract for switching from
-one round to the next, enabling switching from any ACA round to any other.
-
-To simplify the task of expressing existing algorithms as the composition of one
-or more ACA rounds, we propose two specifications, \emph{conservative ACA
-rounds} and
-\emph{fast ACA rounds}, which can easily be concretized to obtain executable rounds.
-We show that ACA versions of Classic Paxos, Multi-Coordinated Paxos, Chain
-Replication, and Ring Paxos can be obtained by concretizing the specification of
-conservative ACA rounds, and that ACA versions of the fast rounds of Fast Paxos can be obtained by concretizing the specification of fast ACA rounds.
-
-Fast and conservative ACA rounds are two specifications, at an
-intermediate level of abstraction, which satisfy ACA. Both
-specifications specify how the state of the acceptors should be
-updated but abstract over the particular communication topology used
-to implement the state accesses: the specifications use a global state
-which is accessible by all servers in all roles. One can refine fast
-or conservative rounds by implementing the state accesses and coordination
-using the network, obtaining a concrete algorithm.
-
-Conservative and fast ACA rounds are specified using the state variables $inits$, $proposed$, for every learner $l$, $learned\left[ l \right]$
-and $aborts$, as in ACA, and, for every acceptor $a$, two new variables
-$status\left[ a \right]$, which can be either ``idle'', ``ready'', or
-``stopped'' and $hist\left[ a \right]$, the local history of the
-acceptor $a$, which can be either a history or the special value $None$.
-Initially, $inits$, $aborts$, and, for every learner
-$l$, $learned\left[ l \right]$ are empty sets, and, for every acceptor $a$,
-$status\left[ a \right]$  is ``idle'' and $hist\left[ a \right]$ is the special
-value $None$.
-
-The next-state relations of conservative and fast rounds are composed
-of the following types of transitions (and the default $Crash\left( x
-\right)$ transition).
-\begin{compactitem}
-    \item A $Propose\left( p,c \right)$ transition, for a proposer $p$
-        and a command $c$, not guarded, inserts $c$ in the set
-        $proposed$. 
-    \item An $Init\left( s, h \right)$ transition, for a switcher $s$
-        and a history $h$, not guarded, inserts $h$ in the set $inits$.
-    \item A $WakeUp\left( a \right)$ transition, for an acceptor $a$,
-        is enabled when $status\left[ a \right]$ is ``idle''. Its
-        effect is to set $hist\left[ a \right]$ to a history found in
-        $inits$ and to set $status\left[ a \right]$ to ``ready''.
-    \item An $Accept\left( a \right)$ transition, for an acceptor $a$,
-        is enabled when $status\left[ a \right]$ is ``ready'', and
-        updates $hist\left[ a \right]$ to $hist\left[ a \right]\star
-        c$, where $c$ is a proposed command. In conservative rounds,
-        the transition is additionally guarded by the condition that
-        the resulting set of local acceptor histories,
-        $\aset{hist'\left[ a \right] : a \in Acceptors}$, forms a chain
-        in the partial order $\leq$ (this is typically implemented with a
-        leader).
-    \item A $Learn\left( l, h \right)$ transition, for a learner $l$
-      and a history $h$, enabled when $h > learned\left[ l \right]$ and is
-      smaller than the GLB of the histories of a learn quorum of acceptors. The effect of
-        the transition is to set $learned\left[ l \right]$ to $h$.
-    \item A $Stop\left( a \right)$ transition, for an acceptor $a$,
-        enabled when $status\left[ a \right]$ is ``ready'', sets
-        $status\left[ a \right]$ to ``stopped'' (note that
-        $status\left[ a \right] = $ ``stopped'' prevents any further
-        ``accept'' transition of acceptor $a$).
-%    \item REALLY NEEDED? A $TimeOut\left( a \right)$ transition, for an acceptor $a$,
-%        is enabled when $status\left[ a \right]$ is ``idle'' and sets 
-%        $status\left[ a \right]$ to ``stopped'' and $hist\left[ a
-%        \right]$ to the special value $None$. 
-    \item An $Abort\left( s, h \right)$ transition, for a switcher $s$
-        and a history $h$, is enabled when there exists a abort quorum
-        $R$ such that, for every member $a$ of $R$, $status\left[ a
-        \right]$ is ``stopped'' and $h \in SafeAborts\left( R
-        \right)$. Its effect is to insert  $h$ in the set $aborts$. 
-\end{compactitem}
-
-On top of the additional guard on the $Accept$ transition present only
-in conservative rounds, conservative and fast ACA rounds differ in
-their definition of learn quorums, abort quorums, and in their
-definition of the $SafeAborts\left( R \right)$ operator.
-To concisely define read and learn quorums, we define 
-\begin{equation}
-    Q\left( f \right) = \aset{as \subseteq Acceptors :
-        Card\left(as\right) \geq \left\lfloor f\times Card\left(
-    Acceptors \right)\right\rfloor + 1}.
-\end{equation}
-
-In conservative ACA rounds, abort and learn quorums are defined such
-that the intersection between two learn quorums is nonempty and the
-intersection between a learn quorum and an abort quorum is nonempty.
-The definition leads to two interesting cases: $LearnQ = AbortQ =
-Q\left( 1/2 \right)$; $LearnQ = \aset{Acceptors}$ and $AbortQ =
-\aset{\aset{a} : a \in Acceptors}$. The operator $SafeAborts\left( R
-\right)$ is defined as the singleton set containing the maximum
-element over the local histories of the acceptors in $R$ which are not
-$None$. Note that the maximum exists because conservative rounds
-guarantee that the histories of the acceptors form a chain.
-
-In fast ACA rounds, abort and learn quorums are defined such that
-the intersection between any two learn quorums is nonempty and the
-intersection between any two learn quorums and an abort quorum is
-nonempty.  The definition leads to three interesting cases: $LearnQ =
-AbortQ = Q\left( 2/3 \right)$; $LearnQ = Q\left( 1/2 \right)$ and
-$AbortQ = Q\left( 3/4 \right)$; $LearnQ = \aset{Acceptors}$ and
-$AbortQ = \aset{\aset{a : a \in Acceptors}}$.
-The operator $SafeAborts\left( R \right)$ is defined as the set of
-maximal elements over the set $\aset{GLB\left( \aset{hist\left[ a
-\right] : a \in R \cap L \wedge hist\left[ a \right]\neq None} \right)
-: L \in LearnQ}$.
-
-The cases in which abort quorums are singleton sets, in both
-conservative and fast rounds, allow tolerating
-$f$ faults with $f+1$ servers: a round can abort as long as one
-acceptor is correct and a new round, possibly using new correct
-servers, can take over. Note that changing round cannot happen
-without the scheduling policy instructing the switchers about the
-identity of the next round, and thus tolerating $f$ faults with
-$f+1$ servers relies on the assumption of a reliable scheduling policy.
-
-Define, for every learner $l$, the history variable $\overline{learned}\left[ l \right]$ which is initialized to the empty set and which is
-updated each time a new local history is accepted by the learner $l$ by
-inserting the new local history in $\overline{learned}\left[ l \right]$ (see \cite{AbadiLamport91ExistenceRefinementMappings} for the
-definition of a history variable).
-\begin{thm}
-  \label{thm:rounds}
-    Both conservative and fast ACA rounds implement ACA under the refinement
-    mapping that consists in substituting $\overline{learned\left[ l \right]}$
-    for $learned\left[ l \right]$.
-\end{thm}
-
-\section{Examples of Concrete ACA Algorithms}
-\label{sec:examples}
-
-In this section we show how to modify the rounds of Classic Paxos,
-Multi-Coordinated Paxos, Chain Replication, Ring Paxos, and Fast Paxos to obtain
-ACA rounds. \Cref{thm:comp} then guarantee that all those different types of
-rounds can be used in the same algorithm to yield an implementation of Chain
-Agreement.
-
-\subsection{Classic Paxos}
-
-The rounds of the Classic Paxos algorithm can be modified so as to
-obtain an implementation of conservative ACA rounds.  We define
-Classic Paxos ACA rounds by specializing the specification of
-conservative ACA rounds, giving a concrete protocol for disseminating
-the proposed commands to the acceptors and enforcing that the learned
-histories form a chain.
-
-Each Classic Paxos round has a unique leader among the acceptors which
-centrally orders the proposed commands, which are accepted by the
-other acceptors in the order determined by the leader.  This protocol
-ensures that every local acceptor history is a prefix of the local
-history of the leader, therefore guaranteeing that the local acceptor histories
-forming a chain.
-
-Acceptor broadcast each newly accepted history to the learners, which
-learn a history $h$ when they receive enough extensions of $h$ from the
-acceptors.  When instructed by the scheduling policy to abort, a
-switcher sends a stop message to the acceptors, which irrevocably stop
-updating their local history and send back to the switcher an
-acknowledgement containing their local history. A switcher
-determines an abort history once it has received acknowledgments from
-an abort quorum of acceptors $R$, using the operator $SafeAborts\left(
-R \right)$.
-
-The specification of Classic Paxos ACA rounds uses the variables of
-the specification of conservative ACA rounds, namely the variables
-$proposed$, $inits$, $aborts$, the arrays $learned\left[ l \right]$,
-$status\left[ a \right]$, and $hist\left[ a \right]$, and two
-additional arrays $accepted\left[ l \right]\left[ a \right]$ and
-$stop\left[ s \right]\left[ a \right]$, where $l$ is a learner, $a$ is
-an acceptor, and $s$ is a switcher.
-
-There is also a network component that we do not describe explicitly.
-However, the network allows any servers to communicate by sending and
-receiving messages.
-
-Initially, the variables $proposed$, $inits$, $aborts$, and, for every
-learner $l$, $learned\left[ l \right]$ are empty sets. For every
-acceptor $a$, $status\left[ a \right]$ is ``idle'' and $hist\left[ a
-\right]$ is the special value $None$. For every learner $l$ and acceptor $a$,
-$accepted\left[ l \right]\left[ a \right]$ is the special value
-$None$. For every switcher $s$ and acceptor $a$, $stopped\left[ s
-\right]\left[ a \right]$ is set to $false$.  There are initially no
-messages in the network.
-
-The next-state relation is composed of the following types of
-transitions, where the definitions of learn quorums, abort quorums,
-and of the $SafeAborts\left( R \right)$ operator are the ones of conservative
-ACA rounds.
-\begin{compactitem}
-    \item A $Propose\left( p,c \right)$ transition, for a proposer $p$
-        and a command $c$, not guarded, inserts $c$ in the set
-        $proposed$ and sends the message $\aseq{\quo{prop},c}$ to
-        the leader.
-    \item An $Init\left( s, h \right)$ transition, for a switcher $s$
-        and a history $h$, not guarded, inserts $h$ in the set $inits$ and
-        sends the message $\aseq{\quo{init},h}$ to the leader.
-    \item A $WakeUp\left( leader \right)$ transition, enabled when
-        $status\left[ leader \right]$ is ``idle'' and the leader can
-        receive a $\aseq{\quo{init},h}$ message. Its effect is to
-        receive the message, set $hist\left[ leader \right]$ to $h$,
-        to set $status\left[ leader \right]$ to ``ready'', and to
-        broadcast the message $\aseq{\quo{leader-accept},h}$ to all
-        other acceptors.
-    \item An $Accept\left( leader \right)$ transition, enabled when
-        $status\left[ leader \right]$ is ``ready'' and the leader can
-        receive a $\aseq{\quo{prop},c}$ message. The effect of the
-        action is to receive the message, to update $hist\left[ leader
-        \right]$ to $hist\left[ leader \right]\star c$, and to
-        broadcast the message $\aseq{\quo{ack},hist\left[ leader
-        \right]\star c}$ to the learners, and to broadcast the message
-        $\aseq{\quo{leader-accept},hist\left[ leader \right]\star c}$
-        to all the other acceptors.
-    \item An $Accept\left( a \right)$ transition, for an acceptor $a$
-        which is not the leader, is enabled when $status\left[ a
-        \right]$ is ``ready'' and $a$ can receive a
-        $\aseq{\quo{leader-accept},h}$ message where $h > hist\left[ a
-        \right]$. The effect of the action is to receive the message,
-        to update $hist\left[ a \right]$ to $h$, and to broadcast the
-        message $\aseq{\quo{ack},hist\left[ a \right]\star c}$ to the
-        learners.
-    \item A $RcvAck\left( l \right)$ transition, for a learner $l$, is
-        enabled when $l$ can receive a message $\aseq{\quo{ack},h}$
-        from an acceptor $a$. Its effect is to receive the message and
-        to set $accepted\left[ l \right]\left[ a \right]$ to $h$.
-    \item A $Learn\left( l, h \right)$ transition, for a learner $l$
-        and a history $h$, is enabled when $h > Max\left(
-        learned\left[ l \right] \right)$ and there is a learn quorum
-        $Q$ of acceptors such that $h = GLB\left( \aset{accepted\left[
-        l \right]\left[ a \right] : a \in Q }\right)$. Its effect is
-        to insert $h$ in $learned\left[ l \right]$.
-    \item A $Stop\left( s \right)$ transition, for a switcher $s$,
-        always enabled, broadcasts a $\aseq{\quo{stop}}$ message to all
-        the acceptors.  This transition models a switcher being
-        instructed by the scheduling policy to abort the current round.
-    \item A $Stop\left( a \right)$ transition, for an acceptor $a$,
-        enabled when $a$ can receive a $\aseq{\quo{stop}}$ message
-        from a switcher $s$. Its effect is to receive the message, set
-        $status\left[ a \right]$ to ``stopped'', and to send the
-        message $\aseq{\quo{stopped},hist\left[ a \right]}$ to the
-        switcher $s$.
-    \item A $RcvStopped\left( s \right)$ transition, for a switcher
-        $s$, enabled when $s$ can receive a $\aseq{\quo{stopped},h}$
-        message from an acceptor $a$. Its effect is to receive the
-        message and to set $stopped\left[ s \right]\left[ a \right]$
-        to $h$.
-    \item An $Abort\left( s, h \right)$ transition, for a switcher $s$
-        and a history $h$, is enabled when there exists an abort quorum
-        $R$ such that, for every member $a$ of $R$, $stopped\left[ s
-        \right]\left[ a \right]$ is not $false$ and $h \in
-        SafeAborts\left( R \right)$, where $SafeAborts\left( R
-        \right)$ is computed using the array $stopped\left[ a
-        \right]$. Its effect is to insert  $h$ in the set $aborts$. 
-\end{compactitem}
-Compared to the rounds of the original Classic Paxos, Classic Paxos ACA
-rounds add an extra level of indirection when changing round: in the
-original rounds, the leader of the new round directly queries the
-acceptors of the preceeding round, whereas in ACA rounds, the
-switchers query the acceptors and then send their abort histories to
-the leader. However, changing round is an infrequent event, therefore
-this difference should not have much practical impact on performance.
-\begin{thm}
-  Classic Paxos ACA rounds satisfy Abortable Chain Agreement.
-\end{thm}
-\begin{proof}
-  A Classic Paxos ACA round implements the specification of Abortable
-  Chain Agreement under the refinement mapping consisting in
-  projecting the state of the round onto the variables $proposed$,
-  $inits$, $aborts$, and, for every learner $l$, $learned\left[ l
-  \right]$.
-\end{proof}
-
-\subsection{Multi-Coordinated Paxos, Chain Replication, and Ring Paxos.}
-
-Multi-Coordinated Paxos, Chain Replication, and Ring Paxos rounds have a similar
-structure as Classic Paxos rounds but differ in the way the leader is
-implemented and in the way that the histories of the leader are disseminated to
-the other acceptors.
-
-Multi-Coordinated Paxos avoids depending on a unique leader by using a
-distributed leader implementation that can make progress as long as a
-strict majority of the acceptors are correct, trading off resilience
-for a higher time complexity. Conservative ACA
-rounds can be implemented like with Classic Paxos, replacing the
-leader by the distributed leader of Multi-Coordinated Paxos.
-
-In Chain Replication, acceptors are arranged in a chain whose head is
-the leader of the round. As in Classic Paxos, abort histories and
-proposed commands are sent to the leader. However, instead of the
-leader directly sending its local history to all the other acceptors,
-the local history of the leader is
-forwarded along the chain from one acceptor to the other until it
-reaches the last acceptor, which sends the history to the learners. 
-By spreading the load more evenly on the all the servers, Chain Replication
-can achieve higher throughput that Classic Paxos.
-Conservative ACA rounds can be implemented like with Classic Paxos,
-replacing the broadcast of the leader by the propagation of the leader's history
-along the chain of servers.
-
-Ring Paxos optimizes Chain Replication by having the leader broadcast
-commands to the acceptors using IP multicast and only sending hashes
-of the commands along the chain. Ring Paxos round can be modified to implement
-conservative ACA rounds in a similar way as for Chain Replication.
-
-\subsection{Fast Paxos}
-
-Fast Paxos is composed of two types of rounds: classic rounds, like in Classic
-Paxos, and fast rounds.
-The fast rounds can be transformed into
-implementations of fast ACA rounds. Instead of sending their
-commands to a leader, as in Classic Paxos, the proposers of fast
-rounds directly broadcast their commands to all the acceptors. As long
-as commands are received in the same order by a learn quorum of
-acceptors or are received in different orders but commute, commands can be learned in two message-propagation delays instead of three in Classic Paxos.
-However, if non-commuting commands are received in different orders by
-several acceptors, the local history of the acceptors may not form a
-chain anymore and the rounds must abort. The fast rounds of Fast
-Paxos implement fast ACA rounds.
-
-Fast Paxos ACA rounds are specified using the same set of variables as for
-Classic Paxos, initialized to the same values.
-The next-state relation is composed of the following types of
-transitions, where the definitions of learn quorums, abort quorums,
-and of the $SafeAborts\left( R \right)$ operator are the ones of fast
-ACA rounds.
-\begin{compactitem}
-    \item A $Propose\left( p,c \right)$ transition, for a proposer $p$
-        and a command $c$, not guarded, inserts $c$ in the set
-        $proposed$ and broadcasts the message $\aseq{\quo{prop},c}$ to
-        all the acceptors.
-    \item An $Init\left( s, h \right)$ transition, for a switcher $s$
-        and a history $h$, not guarded, inserts $h$ in the set $inits$ and
-        broadcasts the message $\aseq{\quo{init},h}$ to all the acceptors.
-    \item A $WakeUp\left( a \right)$ transition, for an acceptor $a$,
-        is enabled when $status\left[ a \right]$ is ``idle'' and $a$
-        can receive an $\aseq{\quo{init},h}$ message. Its effect is to
-        receive the message, set $hist\left[ a \right]$ to $h$, and to
-        set $status\left[ a \right]$ to ``ready''.
-    \item An $Accept\left( a \right)$ transition, for an acceptor $a$,
-        is enabled when $status\left[ a \right]$ is ``ready'' and $a$
-        can receive a $\aseq{\quo{prop},c}$ message. The effect of the
-        action is to receive the message, to update $hist\left[ a
-        \right]$ to $hist\left[ a \right]\star c$, and to broadcast
-        the message $\aseq{\quo{ack},hist\left[ a \right]\star c}$ to
-        the learners.
-    \item The $RcvAck\left( l \right)$ transitions, for a learner $l$,
-        the $Learn\left( l, h \right)$ transitions, for a learner $l$
-        and a history $h$, the $Stop\left( s \right)$ transitions, for
-        a switcher $s$, $Stop\left( a \right)$ transitions, for an
-        acceptor $a$, the $RcvStopped\left( s \right)$ transitions,
-        for a switcher $s$, and the $Abort\left( s, h \right)$
-        transitions, for a switcher $s$ and a history $h$, which are
-        all the same as in Classic Paxos ACA rounds.
-\end{compactitem}
-
-As for Classic Paxos ACA rounds, Fast Paxos ACA rounds implement ACA
-under the refinement mapping consisting projecting the state of the
-round onto the variables $proposed$, $inits$, $aborts$, and, for every
-learner $l$, $learned\left[ l \right]$.
-
-\newpage
-
-\printbibliography
-
-\begin{appendix}
-
-  \label{sec:isaproofs}
-
-%  \includepdf[pages={1},pagecommand={\section{Mechanical Proof of the
-%  Composition Theorem in Isabelle/HOL}\label{sec:isaproofs}},scale=1]{../IsabellePDFS/podctheories.pdf}
-%  \includepdf[pages={2-},pagecommand={},scale=1]{../IsabellePDFS/podctheories.pdf}
-
-\end{appendix}
-
-\end{document}
-    We present a new abstraction, Abortable Generalized Consensus, which enables the
-    correct-by-construction composition of heterogeneous State-Machine
-    Replication algorithms.
-
-    integrating state of the art techniques to optimize executions.  ACA is the most complete approach to date enabling the
-    correct-by-construction composition of heterogeneous State-Machine
-    Replication algorithms. We provide two specifications satisfying
-    ACA and we illustrate their power by refining them to
-    obtain a versatile adaptive algorithm that combines five existing
-    algorithms into a single one. ACA is completely formalized and its
-    underlying composition theorem has been mechanically checked by an interactive
-    proof assistant.
-}
-
-\vspace{2cm}
-
-\begin{center}
-This paper is a regular submission.
-
-\vspace{1cm}
-This paper is a student paper.
-\end{center}
-
-\newpage
-
-\section{Introduction}
-
-The State-Machine Replication technique (abbreviated SMR) allows
-building reliable services on top of unreliable hardware. The basic idea is to 
-replicate a service over several servers, each executing the same
-sequence of (deterministic) commands. 
-Making SMR algorithms efficient has been a long-standing
-challenge and a multitude of algorithms have been proposed. Examples
-include Fast Paxos \cite{Lamport06FastPaxos}, Disk Paxos
-\cite{GafniLamport03DiskPaxos}, Chain Replication
-\cite{RenesseSchneider04ChainReplicationSupportingHighThroughputAvailability},
-Ring Paxos
-\cite{MarandiETAL10RingPaxosHighthroughputAtomicBroadcastProtocol},
-Egalitarian Paxos
-\cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments},
-Multi-Coordinated Paxos
-\cite{CamargosSchmidtPedone07MulticoordinatedPaxos}, Cheap Paxos
-\cite{LamportMassa04CheapPaxos}, Vertical Paxos
-\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication},
-Paxos-MIC
-\cite{HurfinMoiseNarzul11AdaptiveFastPaxosMakingQuickEverlasting},
-Mencius
-\cite{MaoJunqueiraMarzullo08MenciusBuildingEfficientReplicatedStateMachine},
-and Fast Mencius \cite{WeiETAL13FastMenciusMenciusLowCommitLatency}.
-
-Each of the existing algorithms is best suited to a particular range
-of environments, characterized, e.g., by client load, network
-topology, likelihood of faults, communication latency and bandwidth, processing
-power of nodes, availability of disk storage and memory.
-However, the environment faced by real systems evolves over time:
-hardware is upgraded, new data-centers are built, the popularity of a
-service changes, etc.  Consistent performance therefore comes with
-adaptation to change.  Unfortunately, existing SMR algorithms cannot easily be
-stopped and replaced by another algorithm on the fly. Moreover,
-devising a new algorithm that can change behavior dynamically is hard:
-dynamically switching from one SMR algorithm to another requires
-synchronizing notoriously intricate algorithms.
-
-We propose a new abstraction, \emph{Abortable Chain Agreement}
-(abbreviated ACA), which subsumes SMR and whose implementations,
-called \emph{ACA rounds}, can, by construction, pass the baton from
-one to another on the fly and efficiently. Moreover, optimizing a replicated
-service consisting of several ACA rounds becomes just a matter of devising a new
-ACA round, which can then be plugged-in on the fly: any new ACA round
-is compatible with every existing ACA round by construction.
-
-Abortable Chain Agreement is the \emph{most complete framework to date} enabling
-correct by construction composition of heterogeneous rounds.  
-ACA is completely formalized and its underlying composition theorem has been
-mechanically checked by an interactive proof assistant.  
-Compared to ACA,
-existing approaches
-\cite{BarNoyETAL87ShiftingGearsChangingAlgorithmsFlyTo,GuerraouiETAL10Next700BftProtocols,GuerraouiKuncakLosa12SpeculativeLinearizability}
-are limited. In particular, they do not allow the disorderly execution of
-commuting and read-only commands, an important practical optimization
-which is at the root of recent work on SMR like Generalized Consensus
-\cite{Lamport05GeneralizeConsensus} and Generalized Lattice Agreement
-\cite{FalerioETAL12GeneralizedLatticeAgreement}.
-
-\begin{comment}
-On top of the traditional roles of proposer, acceptor, and learners
-\cite{lamport2001paxos} used in consensus algorithms, ACA introduces
-the notion of \emph{switchers} and of \emph{scheduling policy}. The switchers
-have the task of aborting a running ACA round and initializing a new
-ACA round in a safe and live manner.  The \emph{scheduling policy}
-decides when to abort an ACA round and choses the next ACA round to
-execute. It must ensure that all switchers agree on the next ACA round
-to initialize.
-\end{comment}
-
-We observe that, because of the FLP impossibility result
-\cite{FischerLynchPaterson83ImpossibilityDistributedConsensusOneFaultyProcess},
-most existing SMR algorithms are structured in a sequence of abortable
-rounds which can easily be transformed into ACA rounds. To simplify
-this task even further, we propose two specifications, \emph{conservative
-ACA rounds} and \emph{fast ACA rounds}, which, as we show, can be concretized
-to obtain ACA versions of the rounds of five existing algorithms: 
-Classic Paxos, Fast Paxos, Chain Replication, Ring Paxos,
-and Multi-Coordinated Paxos. ACA guarantees that the composition of the five ACA
-rounds is correct by construction, forming a versatile adaptive replication
-algorithm.
-
 The rest of the paper is structured as follows.
@@
 checked by the interactive proof assistant Isabelle/HOL \cite{NipkowPaulsonWenzel02IsabelleHOL}.
-
-\begin{comment}
-Abortable Chain Agreement builds on several recent advances in SMR
-algorithms: ACA allows acceptors to only eventually reach agreement,
-like in Generalized Paxos \cite{Lamport05GeneralizeConsensus},
-boosting performance when most commands commute; ACA allows fast
-execution of read commands like in Generalized Lattice Agreement
-\cite{FalerioETAL12GeneralizedLatticeAgreement}; ACA implementations
-can tolerate $f$ faults with only $f+1$ replicas and replace crashed replicas
-upon changing round, as in Vertical Paxos  \cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}.  
-\end{comment}
-
-\section{Related Work}
-\label{sec:related}
-
-SMR is often implemented using a sequence of instances of Consensus.
-There are numerous examples of consensus algorithms in the literature.
-However, Abortable Chain Agreement is based on Generalized Consensus. 
-In Generalized Consensus \cite{Lamport05GeneralizeConsensus}, the learners of a
-SMR algorithm repeatedly learn new values belonging to a \emph{C-struct set}. A
-C-struct can be seen as a data structure encoding  sequences of commands modulo
-the ordering of commands that commute. The advantage of Generalized
-Consensus over Consensus is that
-two servers which receive commuting commands in different orders can still reach
-agreement on C-structs without further synchronization. 
-Our Abortable Chain Agreement can also use C-structs, with the same advantages
-as Generalized Consensus. For simplicity, the presentation in this
-paper uses the more concrete notion of \emph{command history} (see
-\cref{sec:histories}).
-
-Generalized Lattice Agreement \cite{FalerioETAL12GeneralizedLatticeAgreement} allows more efficient implementations of SMR when
-commands are partitioned into read-only commands and write-only commands.
-Abortable Chain Agreement supports the same optimizations as Generalized Lattice Agreement.
-Notably, read-only commands are treated specially so as to optimize their
-executions (see \cref{sec:readonly}).
-
-Vertical Paxos algorithms 
-\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication} are SMR algorithms
-whose set of replica servers can be reconfigured without stopping to process new
-commands. Crashed servers can therefore be replaced by new servers, ensuring the
-long-term availability of a replicated service.
-Vertical Paxos algorithms use an external reconfiguration master, itself
-implemented as a replicated state machine, which determines the servers that should
-be removed from or added to the configuration.
-ACA shares some aspects of Vertical Paxos algorithms: the task of the
-reconfiguration master is similar to the task of the scheduling policy
-component of ACA (see \cref{sec:aca}).
-However, in ACA, changing round can be used not only to replace crashed servers
-but also to change the algorithm that the servers are running. 
-Changing algorithm is not possible in Vertical Paxos.
-In both Vertical Paxos and ACA, a reliable reconfiguration master, resp. a
-reliable scheduling policy, allows tolerating $f$ crashes with $f+1$ servers.
-
-The Abstract Framework \cite{GuerraouiETAL10Next700BftProtocols} proposes to
-build Byzantine Fault-Tolerant algorithms as the succession of abortable rounds
-called Abstract instances. Abstract instances must satisfy the Abstract
-correctness properties. By construction, the composition of any number of
-Abstract instances is a correct SMR implementation. However, the Abstract
-Framework uses totally ordered sequence of commands, making difficult the
-optimization of the execution of commuting and read-only commands that are
-possible with ACA.  Moreover, the Abstract Framework does not identify the
-crucial role of the scheduling policy for the long-term resilience of a service.
-Abortable Chain Agreement combines the ideas of the Abstract Framework,
-Generalized Paxos, Generalized Lattice Agreement, and Vertical Paxos in a
-unified abstraction.
-
-Finally, the Speculative Linearizability framework
-\cite{GuerraouiKuncakLosa12SpeculativeLinearizability} was a first
-attempt by the authors of the present paper at generalizing the
-Abstract Framework and mechanically proving its properties with the
-Isabelle/HOL interactive
-proofs assistant.  Like the Abstract Framework, the
-Speculative Linearizability framework is based on totally ordered
-sequences of commands, with the same drawback for the execution of
-commuting and read-only commands. Moreover, Abortable Chain Agreement is both
-conceptually simpler and subsumes the Speculative Linearizability
-property.  Like Speculative Linearizability, Abortable Chain Agreement
-and its main property, the composition theorem (\cref{thm:comp}), are
-formalized in Isabelle/HOL (see \cref{sec:isaproofs}).
-
-Abortable Chain Agreement is the most complete framework to date for building
-adaptive replication algorithms out of heterogeneous rounds, making state of the
-art optimization techniques available to algorithm designers, which was not
-possible before.
-
-\section{Chain Agreement}
-\label{sec:ca}
-
-We consider a set of asynchronous servers communicating by message passing in an
-asynchronous network. Servers are not necessarily sequential processes. We
-describe a behavior of the system using an infinite sequence of states, where a
-state is a function from variable names to values (i.e. a valuation
-of variables).  We specify sets of allowed behaviors using an initial predicate,
-a next-state relation (using unprimed and primed variables) and a fairness
-condition, all in the style of TLA+ \cite{Lamport02SpecifyingSystems}. However,
-for succinctness, our specification are rather informal, eluding non-essential
-details.
-
-Our specifications all contain by default a state variable $crashed$ and
-$Crash\left( x \right)$ transition which is always enabled. A server $x$ faithfully
-executes the algorithm that is assigned to it until its identifier is added to
-the variable $crashed$ by a $Crash\left( x \right)$ transitions.  When the identifier of a
-server is in the set $crashed$, all its actions are disabled.  In an execution,
-a server that is never added to the set $crashed$ is said correct.
-
-We consider a computing service exposing a set of commands $C$. The
-service has a sequential specification consisting of an initial state
-and a deterministic transition relation in which every command
-atomically updates the state of the service and produces an output. We
-assume that duplicate commands have no effect on the state of the
-service and return the same output as their first occurrence.
-
-In a traditional SMR algorithm, the servers can be seen has playing one or
-several of three roles: \emph{proposers}, which propose new commands to execute,
-\emph{acceptors}, which accept and order commands in a sequence, and
-\emph{learners}, which learn about the growing sequence of commands to
-execute \cite{lamport2001paxos}. Often, but not necessarily, every replica server plays the
-three roles of proposer, acceptor, and learner: clients of the service
-send their commands to some replica servers; replica servers submit
-the received commands to the SMR protocol in the role of proposers;
-replica servers respond to commands by executing the commands learned
-in the role of learners; replica servers participate in command
-ordering in the role of acceptors.  An SMR algorithm guarantees that
-the replicas learn the same sequences of commands, possibly up to the
-reordering of commuting commands. Therefore, every learner can respond to client
-requests using its last learned sequence of commands, and clients will receive
-the same responses as the ones produces by a centralized, non-replicated,
-service.
-
-Chain Agreement precisely specifies the allowed behaviors of the roles of \emph{proposers}
-and \emph{learners}, abstracting over the behavior of acceptors.
-
-\begin{comment}
-We suppose that each server can play a number of \emph{roles}.
-Chain Agreement specifies the allowed behaviors the roles of \emph{proposers}
-and \emph{learners}. Proposers receive commands from the clients of the service
-and propose them for execution by the replicated service. Each learner learns about
-a growing sequence of commands. Thanks to the sequential specification of the
-service, a sequence of commands determines an output for each the included commands. 
-In a traditional SMR algorithm, servers can also play the role of
-\emph{acceptors}, whose task is to collectively order commands and let the
-learners know about the latest sequence of ordered commands.
-
-The point of Chain Agreement is to implement a replicated service by having every learner learn about a growing sequence of commands, which
-represent an execution of the sequential specification of the service.
-Crucially, at all times, the sequences of commands learned by the
-different learners must form a chain in the prefix order on sequences.
-Moreover, for every command $c$ proposed by a correct proposer, every
-correct learner must eventually learn a sequence containing $c$.  Once
-a learner learns about a sequence containing a command $c$, it can
-determine the output produced by $c$, using the sequential
-specification of the service, and respond to the client that issued
-the command $c$.
-\end{comment}
-
-To specify Chain Agreement, we need the notion of ordering of sequences and of
-maximum sequence.  The prefix order on sequences is denoted by the operator
-$\leq$ ($s_1 \leq s_2$ means that $s_1$ is a prefix of $s_2$). 
-%Using the operator $\leq$, we define $\geq$, $<$, and $>$ as expected. 
-Note that $\leq$ is a partial order on sequences. The maximum over a set of
-sequences $S$ is denoted by $Max\left( S \right)$ and $Max\left( \aset{}
-\right)$ is the empty sequence.
-
-We specify Chain Agreement using the state variables $crashed$, $proposed$, containing the proposed commands,
-and, for every learner $l$,  $learned\left[ l \right]$, a set of
-sequences of commands. Initially, $crashed$, $proposed$ and, for every learner $l$,
-$learned\left[ l \right]$ are empty sets. 
-Remember that any transition involving a server $x$ is always disabled when $x\in crashed$.
-
-The next-state relation consists of three types of transitions:
-$Crash\left( x \right)$ transitions; $Propose\left( l, c
-\right)$ transitions, for a learner $l$ and a command $c$, in which
-$c$ is inserted in the set $proposed$; $Learn\left( l,cs \right)$
-transitions, for a learner $l$ and a sequence of commands $cs$, in
-which $cs$ is inserted in $learned\left[ l \right]$. 
-A $Learn\left( l, cs \right)$ transition is guarded by the following
-conditions:
-\begin{compactitem}
-    \item[\textbf{Agreement}:] The elements of the set $\aset{learned'\left[ l
-      \right] : l \in Learners}$, the learned sequences after the transition,
-      form a chain in the partial order $\leq$.
-    \item[\textbf{Validity}:] Every learned sequence contains only
-        proposed commands.  
-    \item[\textbf{Irrevocability}:] The new learned sequence
-        $cs$ is a strict extension of $Max\left(learned\left[ l
-        \right]\right)$. 
-        %\\ $cs > Max\left(learned\left[ l \right]\right)$.
-\end{compactitem}
-Finally, we add, for every learner $l$, one weak fairness condition
-\cite{Lamport02SpecifyingSystems} on the transitions $Learn\left( l,
-cs \right)$. The weak fairness conditions ensure that every correct
-learner eventually learns about every proposed command.
-
-Thanks to the properties of Chain Agreement, the clients, observing
-their input commands and the received responses, see a
-\emph{linearizable} execution
-\cite{HerlihyWing90LinearizabilityCorrectnessConditionConcurrentObjects}.
-Moreover, the service remains available as long as at least one
-proposer and one learner are correct.
-
-\section{Abortable Chain Agreement}
-\label{sec:aca}
-
-We now consider implementing Chain Agreement as a sequence of \emph{Abortable Chain
-Agreement rounds}, where each round makes progress ordering requests or aborts
-its execution and passes the baton to the next round in the sequence.
-An algorithm composed of a sequence of ACA rounds is called an \emph{ACA
-algorithm}.
-
-On top of the traditional roles of proposer, acceptor, and learner, ACA
-introduces the new role of \emph{switcher} and a new component, the
-\emph{scheduling policy}. In ACA, the assignment of the roles of proposer,
-acceptor, learner, and switcher can change in each rounds, allowing for
-reconfiguration. 
-In normal circumstances, the proposers, acceptors,
-and learners of an ACA round work together to implement Chain Agreement.
-However, if the environment becomes unfavorable to a running ACA round, the
-scheduling policy can instruct the round to abort its execution and pass the
-baton, using the switchers, to a new round.
-
-A switcher of an aborting round determines, from the state of the
-acceptors, an \emph{abort sequence} that it transfers to one or several
-switchers of the new round.  Upon receiving an abort sequence, a switcher of the
-new round initializes the acceptors of the new round, which can then
-proceed implementing SMR, until the new round itself aborts and the
-process is repeated. 
-We assume that in the first round, which has no previous round, every 
-switchers receives the empty abort sequence.
-
-The scheduling policy determines when to abort an ACA round and which new ACA
-round should take over. The scheduling policy must inform the switchers about the
-identity of the next round and must guarantee that all switchers receive the
-same information.  In this paper we abstract over the scheduling policy, assuming
-that a round that aborts passes the baton a nondeterministically chosen new
-round. A reliable scheduling-policy component can realistically be implemented
-as a replicated state machine running on a large number of servers otherwise
-dedicated to other tasks
-\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}. We will see
- in \cref{sec:rounds} that, as in Vertical Paxos, a reliable scheduling policy component allows
-tolerating $f$ faults with $f+1$ replicas, while maintaining the long-term
-resilience of the service by replacing faulty servers upon round change.
-
-The clients of an ACA algorithm can send their commands to the proposers of any round (in
-practice, clients send their commands to proposers of the largest
-active round) and servers can use sequences learned in any round
-to determine outputs.
-
-ACA specifies the allowed behaviors of the proposers, learners, and switchers
-\emph{of a single round}. Similarly to
-Chain Agreement, we use the state variables $crashed$, $proposed$,
-and, for every learner $l$, $learned\left[ l \right]$.  Additionally,
-we use the state variables $inits$, containing the abort sequences
-received from the previous round, and $aborts$, containing the abort
-sequences transfered to the next round.
-Two consecutive ACA rounds are composed by letting the $aborts$ variable of the
-first round be the $inits$ variable of the second round.
-
-Initially, $crashed$, $proposed$, for every
-learner $l$, $learned\left[ l \right]$, $inits$, and $aborts$ are all empty sets.
-The next-state relation consists of five types of transitions. The
-first three, $Crash\left( x \right)$, $Propose\left( l, c
-\right)$, and $Learn\left( l, cs \right)$, for a learner, proposer, or
-switcher $x$, for a learner $l$, and for a sequence of commands
-$cs$, have the same effect as in Chain Agreement.
-ACA introduces two new transitions, $Init\left(
-s, cs \right)$ and $Abort\left( s, cs \right)$, for a switcher $s$ and a sequence of
-commands $cs$. The transition $Init\left( s, cs \right)$ adds $cs$ to the set $inits$, and models a switcher receiving the
-abort sequence $cs$ from the preceding round.
-The transition $Abort\left( s, cs \right)$ adds $cs$ to the set $aborts$ and models a switcher transferring the abort sequence $cs$ to a switcher of the next round.
-All the transition are guarded by the following conditions:
-\begin{compactitem}
-    \item[\textbf{Agreement}:] The elements of the set $\aset{learned'\left[ l
-      \right] : l \in Learners}$ form a chain in the partial order.
-    \item[\textbf{Validity}:] Every abort sequence or learned sequence
-        is of the form $h \star cs$, where $cs$ contains only commands
-        of the set $proposed$ and $h$ is the greatest lower bound of a
-        nonempty subset of $inits$.
-    \item[\textbf{Initialization}:] If $inits$ is empty, then no
-        sequence can be learned and no switcher can abort.
-    \item[\textbf{Irrevocability}:] In the case of a $Learn\left( l,cs
-        \right)$ transition, the new learned sequence $cs$ is a strict
-        extension of $Max\left(learned\left[ l \right]\right)$.
-        %, $cs
-        %\geq Max\left(learned\left[ l \right]\right)$.
-    \item[\textbf{Safe Abort}:] Every abort sequence is larger in the
-        partial order $\leq$ than every learned sequence.
-\end{compactitem}
-Note that Agreement and Irrevocability are the same as in Chain
-Agreement.
-
-Finally, for every learner $l$, we add a weak fairness condition on the
-following state transition: 
-\begin{equation}
-    \exists  s \in Switchers, cs \in Sequence\left( C \right) : Learn\left( l, cs \right) \vee Abort\left( s,cs \right)
-\end{equation}
-The weak fairness conditions ensure that when a correct learner has not
-learned a sequence containing all commands proposed in the round, then
-it eventually learns a new command or, if there is at least one correct switcher, at least one switch value is transfered to the next round. 
-Therefore, if a client has submitted a command to a proposer, eventually,
-either the client gets a response from the current round, or
-the next round is initialized with an abort sequence and the
-command can be resubmitted to the next round.
-
-\subsection{The Composition Theorem}
-
-Consider the succession of $n$ rounds numbered $1$ to $n$.  To
-distinguish the variables of each round, we annotate the local
-variables of each round with the number of the round. For example, the
-variables of the seventh round are written $proposed_7$, for every
-learner $l$, $learned_7\left[ l \right]$, $inits_7$, and $aborts_7$.
-
-Moreover, we assume that a command, which is proposed to a round and that
-is not learned before the first abort sequence is transfered to the
-next round, is proposed anew to the next round.
-
-Define the state variable $proposed$ as the union over all rounds $i$
-of the $proposed_i$ variables and, for every learner $l$, 
-define the $learned\left[ l \right]$ variable as the union over all
-rounds $i$ of the $learned_i\left[ l \right]$ variables.
-
-\begin{thm}[Composition Theorem]
-    \label{thm:comp}
-    In every execution of a succession of ACA rounds, the sequence of valuations
-    of the variable $proposed$ and, for every learner $l$, of the variables
-    $learned\left[ l \right]$ satisfies Chain Agreement.
-\end{thm}
-\Cref{thm:comp} can be shown by induction on the number of rounds,
-using the following lemma.
-\begin{lem}
-  \label{lem:comp}
-    Consider two rounds numbered $i$ and $i+1$. In every execution of a
-    succession of ACA rounds, the sequence of valuations of the variables $inits
-    = inits_i$, $propose = proposed_i \cup propose_{i+1}$, for every learner
-    $l$, $learned\left[ l \right] = learned_i\left[ l \right] \cup
-    learned_{i+1}\left[ l \right]$and $aborts = aborts_{i+1}$ satisfies ACA.
-\end{lem}
-\begin{proof}
-  A mechanically-checked proof in Isabelle/HOL appears in \cref{sec:isaproofs}.
-  The proofs uses a version of the specification of ACA that has a more
-  executable form that presented in this section. Moreover,
-  instead of sequences of commands, the Isabelle/HOL specification uses the more
-  general notion of command history, presented in the next subsection.
-  The Isabelle/HOL formalization of ACA uses the theory of I/O-automata, which
-  is also formalized in the Isabelle/HOL development. The proof uses a refinement
-  mapping from the composition of two consecutive ACA instances to a single ACA
-  instance.
-\end{proof}
-
-\Cref{thm:comp} guarantees that the successive composition of any
-number of ACA rounds implements Chain Agreement.
-Therefore, with ACA, one can easily build replicated services that
-dynamically adapt to the current operating conditions of the system: a
-replicated service has several types of rounds at its disposal, each being
-tailored to particular conditions. Each time the conditions change,
-the current ACA round aborts and passes the baton to a more 
-appropriate type of round, as determined by the scheduling policy. Moreover, a new type
-of ACA round can always be devised after the system is deployed and added on the fly to
-the existing types of rounds.
-
-We now describe two important practical optimizations of ACA.
-
-\subsection{Histories}
-\label{sec:histories}
-
-ACA can be implemented more efficiently by substituting \emph{command histories} for sequences of commands, as in Generalized Consensus \cite{Lamport05GeneralizeConsensus}.
-
-We say that two commands $c_1$ and $c_2$ \emph{commute} when, in every sequential
-execution of the service, reversing the order of $c_1$ and $c_2$ does
-not change any of the outputs, including those given to $c_1$ and
-$c_2$.
-
-A command history, abbreviated history, is a data structure that
-represents a set of sequences which are the same modulo the reordering
-of commuting commands, and in which duplicate commands do not appear.
-All the sequences represented by a given history contain the same
-commands and determine the same outputs for each command.  Therefore,
-to implement a replicated service, it is sufficient that learners
-learn command histories instead of sequences.  For a precise
-definition of command histories, we refer the reader to
-\cite{Lamport05GeneralizeConsensus}. Our Isabelle/HOL formalization
-(\cref{sec:isaproofs}) abstracts
-over the concrete representation of histories, assuming only their necessary
-properties. Command histories were first
-introduced in trace theory \cite{Mazurkiewicz84Semantics}.
-
-A command history can be built by appending a sequence of commands
-$cs$ to a history $h$, obtain the history $h \star cs$.  The major
-advantage that histories have over sequences is that if a sequence
-$cs_1$ is obtained by reordering the commuting commands of $cs_2$,
-then $h \star cs_1 = h \star cs_2$. Therefore, in  implementations of
-ACA, servers can agree on a history even though
-they chose different orders for some commuting commands.
-
-Command histories share the four properties of sequences that are
-necessary and sufficient to substitute histories for sequences in the
-definition of Chain Agreement and ACA.
-\begin{compactitem}
-    \item[\textbf{Distinctness}:] No duplicate command appears in a history. 
-    \item[\textbf{Ordering}:] Command histories form a partial order:
-        $h_1 \leq h_2$ when there exists a sequence $s$ such that $h_2
-        = h_1 \star s$.
-    \item[\textbf{GLB}:] Every finite set of command histories has a
-        \emph{greatest lower bound}.
-%    \item[\textbf{LUB}:] Every finite set of compatible command
-%        histories has a \emph{least upper bound}, where a set of
-%        command histories is compatible when they have a common upper
-%        bound.
-    \item[\textbf{Consistency}:] If $h_1 \leq h_2 \leq h_3$ and $h_3 =
-        h_1 \star s$ then there exists $s'$ and $s''$ where every
-        request of $s'$ and $s''$ appears in $s$ and such that $h_2 =
-        h_1 \star s'$ and $h_3 = h_2 \star s''$.
-\end{compactitem}
-The optimized version of ACA is obtained by
-simply substituting histories for sequences in the definition of ACA.
-The Ordering, GLB, and Consistency properties of histories ensure that
-\cref{thm:comp} still holds: the Isabelle/HOL proof of \cref{thm:comp} (see
-\cref{sec:isaproofs}) relies only on the four properties above.
-
-\subsection{Optimizing the Execution of Read-Only Commands}
-\label{sec:readonly}
-
-Chain Agreement can be further optimized by treating 
-read-only commands specially, as in Generalized Lattice Agreement
-\cite{FalerioETAL12GeneralizedLatticeAgreement}.
-
-We consider the case in which a set of servers each play the four
-roles of proposer, acceptor, learner, and switcher.
-Clients sends their commands to any server, which proposes them as a
-proposer, and records the command as pending.
-%When a server learns a history that contains one of its pending
-%commands, it uses the history to determine the response to the pending
-%command and forwards the response to the appropriate client.
-
-As observed in \cite{FalerioETAL12GeneralizedLatticeAgreement}, read-only
-commands can be executed faster by directly reading from the last learned
-history, instead of proposing the read commands and waiting for it to appear in
-a future learned history.  Because learned histories form a chain, all the reads
-return consistent values and this approach yields a \emph{serializable}
-implementation of the service.
-
-However linearizability is not satisfied if reads are served directly
-from the last learned history: a lagging replica server could respond
-to a read using a history which does not contain all completed update
-commands. To make the executions linearizable without resorting to
-proposing read-only commands as normal commands, we can employ the same
-technique as in \cite{FalerioETAL12GeneralizedLatticeAgreement}: upon
-receiving a read command, a server proposes a special no-op command
-that commutes with all other commands; the server then responds to the
-read command using the first learned history containing the no-op
-command. This ensures that all the commands that had completed before
-the read command was issued are present in the history.  Moreover,
-since no-op commands commute with all other commands, they can be
-treated more efficiently than regular read commands by the acceptors,
-like, for example, in a fast round of Fast Paxos.
-
-\section{Fast and Conservative Rounds}
-\label{sec:rounds}
-
-Most existing SMR algorithms can easily be obtained as the
-composition of one or more ACA rounds.
-
-The FLP impossibility result \cite{FischerLynchPaterson83ImpossibilityDistributedConsensusOneFaultyProcess}
-implies that no algorithm can implement SMR to the letter in practice.
-Instead, SMR algorithms guarantee safety (i.e. replica execute the
-same sequence of commands) but may not be able to make progress (i.e.
-order new commands). To make progress as likely as possible, most SMR
-algorithms are structured in consecutive rounds (also called ballots,
-views, or epochs). In each round, the algorithm tries to make progress
-but cannot be guaranteed to succeed. When the algorithm detects that
-progress is no more possible, a new round is started.  The new round
-is initialized so as to preserve safety, usually by one of the
-acceptors. For example, in Paxos, the leader of a newly started ballot
-numbered $i$ determines a value that is \emph{safe at $i$} and
-proposes it as first value to the other acceptors of the ballot $i$.
-In existing SMR algorithms, new rounds are initialized ad-hoc, which
-prevents mixing rounds of two different algorithms.  Instead,
-ACA is an abstraction of rounds which establishes a common interface and contract for switching from
-one round to the next, enabling switching from any ACA round to any other.
-
-To simplify the task of expressing existing algorithms as the composition of one
-or more ACA rounds, we propose two specifications, \emph{conservative ACA
-rounds} and
-\emph{fast ACA rounds}, which can easily be concretized to obtain executable rounds.
-We show that ACA versions of Classic Paxos, Multi-Coordinated Paxos, Chain
-Replication, and Ring Paxos can be obtained by concretizing the specification of
-conservative ACA rounds, and that ACA versions of the fast rounds of Fast Paxos can be obtained by concretizing the specification of fast ACA rounds.
-
-Fast and conservative ACA rounds are two specifications, at an
-intermediate level of abstraction, which satisfy ACA. Both
-specifications specify how the state of the acceptors should be
-updated but abstract over the particular communication topology used
-to implement the state accesses: the specifications use a global state
-which is accessible by all servers in all roles. One can refine fast
-or conservative rounds by implementing the state accesses and coordination
-using the network, obtaining a concrete algorithm.
-
-Conservative and fast ACA rounds are specified using the state variables $inits$, $proposed$, for every learner $l$, $learned\left[ l \right]$
-and $aborts$, as in ACA, and, for every acceptor $a$, two new variables
-$status\left[ a \right]$, which can be either ``idle'', ``ready'', or
-``stopped'' and $hist\left[ a \right]$, the local history of the
-acceptor $a$, which can be either a history or the special value $None$.
-Initially, $inits$, $aborts$, and, for every learner
-$l$, $learned\left[ l \right]$ are empty sets, and, for every acceptor $a$,
-$status\left[ a \right]$  is ``idle'' and $hist\left[ a \right]$ is the special
-value $None$.
-
-The next-state relations of conservative and fast rounds are composed
-of the following types of transitions (and the default $Crash\left( x
-\right)$ transition).
-\begin{compactitem}
-    \item A $Propose\left( p,c \right)$ transition, for a proposer $p$
-        and a command $c$, not guarded, inserts $c$ in the set
-        $proposed$. 
-    \item An $Init\left( s, h \right)$ transition, for a switcher $s$
-        and a history $h$, not guarded, inserts $h$ in the set $inits$.
-    \item A $WakeUp\left( a \right)$ transition, for an acceptor $a$,
-        is enabled when $status\left[ a \right]$ is ``idle''. Its
-        effect is to set $hist\left[ a \right]$ to a history found in
-        $inits$ and to set $status\left[ a \right]$ to ``ready''.
-    \item An $Accept\left( a \right)$ transition, for an acceptor $a$,
-        is enabled when $status\left[ a \right]$ is ``ready'', and
-        updates $hist\left[ a \right]$ to $hist\left[ a \right]\star
-        c$, where $c$ is a proposed command. In conservative rounds,
-        the transition is additionally guarded by the condition that
-        the resulting set of local acceptor histories,
-        $\aset{hist'\left[ a \right] : a \in Acceptors}$, forms a chain
-        in the partial order $\leq$ (this is typically implemented with a
-        leader).
-    \item A $Learn\left( l, h \right)$ transition, for a learner $l$
-      and a history $h$, enabled when $h > learned\left[ l \right]$ and is
-      smaller than the GLB of the histories of a learn quorum of acceptors. The effect of
-        the transition is to set $learned\left[ l \right]$ to $h$.
-    \item A $Stop\left( a \right)$ transition, for an acceptor $a$,
-        enabled when $status\left[ a \right]$ is ``ready'', sets
-        $status\left[ a \right]$ to ``stopped'' (note that
-        $status\left[ a \right] = $ ``stopped'' prevents any further
-        ``accept'' transition of acceptor $a$).
-%    \item REALLY NEEDED? A $TimeOut\left( a \right)$ transition, for an acceptor $a$,
-%        is enabled when $status\left[ a \right]$ is ``idle'' and sets 
-%        $status\left[ a \right]$ to ``stopped'' and $hist\left[ a
-%        \right]$ to the special value $None$. 
-    \item An $Abort\left( s, h \right)$ transition, for a switcher $s$
-        and a history $h$, is enabled when there exists a abort quorum
-        $R$ such that, for every member $a$ of $R$, $status\left[ a
-        \right]$ is ``stopped'' and $h \in SafeAborts\left( R
-        \right)$. Its effect is to insert  $h$ in the set $aborts$. 
-\end{compactitem}
-
-On top of the additional guard on the $Accept$ transition present only
-in conservative rounds, conservative and fast ACA rounds differ in
-their definition of learn quorums, abort quorums, and in their
-definition of the $SafeAborts\left( R \right)$ operator.
-To concisely define read and learn quorums, we define 
-\begin{equation}
-    Q\left( f \right) = \aset{as \subseteq Acceptors :
-        Card\left(as\right) \geq \left\lfloor f\times Card\left(
-    Acceptors \right)\right\rfloor + 1}.
-\end{equation}
-
-In conservative ACA rounds, abort and learn quorums are defined such
-that the intersection between two learn quorums is nonempty and the
-intersection between a learn quorum and an abort quorum is nonempty.
-The definition leads to two interesting cases: $LearnQ = AbortQ =
-Q\left( 1/2 \right)$; $LearnQ = \aset{Acceptors}$ and $AbortQ =
-\aset{\aset{a} : a \in Acceptors}$. The operator $SafeAborts\left( R
-\right)$ is defined as the singleton set containing the maximum
-element over the local histories of the acceptors in $R$ which are not
-$None$. Note that the maximum exists because conservative rounds
-guarantee that the histories of the acceptors form a chain.
-
-In fast ACA rounds, abort and learn quorums are defined such that
-the intersection between any two learn quorums is nonempty and the
-intersection between any two learn quorums and an abort quorum is
-nonempty.  The definition leads to three interesting cases: $LearnQ =
-AbortQ = Q\left( 2/3 \right)$; $LearnQ = Q\left( 1/2 \right)$ and
-$AbortQ = Q\left( 3/4 \right)$; $LearnQ = \aset{Acceptors}$ and
-$AbortQ = \aset{\aset{a : a \in Acceptors}}$.
-The operator $SafeAborts\left( R \right)$ is defined as the set of
-maximal elements over the set $\aset{GLB\left( \aset{hist\left[ a
-\right] : a \in R \cap L \wedge hist\left[ a \right]\neq None} \right)
-: L \in LearnQ}$.
-
-The cases in which abort quorums are singleton sets, in both
-conservative and fast rounds, allow tolerating
-$f$ faults with $f+1$ servers: a round can abort as long as one
-acceptor is correct and a new round, possibly using new correct
-servers, can take over. Note that changing round cannot happen
-without the scheduling policy instructing the switchers about the
-identity of the next round, and thus tolerating $f$ faults with
-$f+1$ servers relies on the assumption of a reliable scheduling policy.
-
-Define, for every learner $l$, the history variable $\overline{learned}\left[ l \right]$ which is initialized to the empty set and which is
-updated each time a new local history is accepted by the learner $l$ by
-inserting the new local history in $\overline{learned}\left[ l \right]$ (see \cite{AbadiLamport91ExistenceRefinementMappings} for the
-definition of a history variable).
-\begin{thm}
-  \label{thm:rounds}
-    Both conservative and fast ACA rounds implement ACA under the refinement
-    mapping that consists in substituting $\overline{learned\left[ l \right]}$
-    for $learned\left[ l \right]$.
-\end{thm}
-
-\section{Examples of Concrete ACA Algorithms}
-\label{sec:examples}
-
-In this section we show how to modify the rounds of Classic Paxos,
-Multi-Coordinated Paxos, Chain Replication, Ring Paxos, and Fast Paxos to obtain
-ACA rounds. \Cref{thm:comp} then guarantee that all those different types of
-rounds can be used in the same algorithm to yield an implementation of Chain
-Agreement.
-
-\subsection{Classic Paxos}
-
-The rounds of the Classic Paxos algorithm can be modified so as to
-obtain an implementation of conservative ACA rounds.  We define
-Classic Paxos ACA rounds by specializing the specification of
-conservative ACA rounds, giving a concrete protocol for disseminating
-the proposed commands to the acceptors and enforcing that the learned
-histories form a chain.
-
-Each Classic Paxos round has a unique leader among the acceptors which
-centrally orders the proposed commands, which are accepted by the
-other acceptors in the order determined by the leader.  This protocol
-ensures that every local acceptor history is a prefix of the local
-history of the leader, therefore guaranteeing that the local acceptor histories
-form a chain.
-
-Each acceptor broadcasts each newly accepted history to the learners, which
-learn a history $h$ when they receive enough extensions of $h$ from the
-acceptors.  When instructed by the scheduling policy to abort, a
-switcher sends a stop message to the acceptors, which irrevocably stop
-updating their local history and send back to the switcher an
-acknowledgement containing their local history. A switcher
-determines an abort history once it has received acknowledgments from
-an abort quorum of acceptors $R$, using the operator $SafeAborts\left(
-R \right)$.
-
-The specification of Classic Paxos ACA rounds uses the variables of
-the specification of conservative ACA rounds, namely the variables
-$proposed$, $inits$, $aborts$, the arrays $learned\left[ l \right]$,
-$status\left[ a \right]$, and $hist\left[ a \right]$, and two
-additional arrays $accepted\left[ l \right]\left[ a \right]$ and
-$stop\left[ s \right]\left[ a \right]$, where $l$ is a learner, $a$ is
-an acceptor, and $s$ is a switcher.
-
-There is also a network component that we do not describe explicitly.
-However, the network allows any servers to communicate by sending and
-receiving messages.
-
-Initially, the variables $proposed$, $inits$, $aborts$, and, for every
-learner $l$, $learned\left[ l \right]$ are empty sets. For every
-acceptor $a$, $status\left[ a \right]$ is ``idle'' and $hist\left[ a
-\right]$ is the special value $None$. For every learner $l$ and acceptor $a$,
-$accepted\left[ l \right]\left[ a \right]$ is the special value
-$None$. For every switcher $s$ and acceptor $a$, $stopped\left[ s
-\right]\left[ a \right]$ is set to $false$.  There are initially no
-messages in the network.
-
-The next-state relation is composed of the following types of
-transitions, where the definitions of learn quorums, abort quorums,
-and of the $SafeAborts\left( R \right)$ operator are the ones of conservative
-ACA rounds.
-\begin{compactitem}
-    \item A $Propose\left( p,c \right)$ transition, for a proposer $p$
-        and a command $c$, not guarded, inserts $c$ in the set
-        $proposed$ and sends the message $\aseq{\quo{prop},c}$ to
-        the leader.
-    \item An $Init\left( s, h \right)$ transition, for a switcher $s$
-        and a history $h$, not guarded, inserts $h$ in the set $inits$ and
-        sends the message $\aseq{\quo{init},h}$ to the leader.
-    \item A $WakeUp\left( leader \right)$ transition, enabled when
-        $status\left[ leader \right]$ is ``idle'' and the leader can
-        receive a $\aseq{\quo{init},h}$ message. Its effect is to
-        receive the message, set $hist\left[ leader \right]$ to $h$,
-        to set $status\left[ leader \right]$ to ``ready'', and to
-        broadcast the message $\aseq{\quo{leader-accept},h}$ to all
-        other acceptors.
-    \item An $Accept\left( leader \right)$ transition, enabled when
-        $status\left[ leader \right]$ is ``ready'' and the leader can
-        receive a $\aseq{\quo{prop},c}$ message. The effect of the
-        action is to receive the message, to update $hist\left[ leader
-        \right]$ to $hist\left[ leader \right]\star c$, and to
-        broadcast the message $\aseq{\quo{ack},hist\left[ leader
-        \right]\star c}$ to the learners, and to broadcast the message
-        $\aseq{\quo{leader-accept},hist\left[ leader \right]\star c}$
-        to all the other acceptors.
-    \item An $Accept\left( a \right)$ transition, for an acceptor $a$
-        which is not the leader, is enabled when $status\left[ a
-        \right]$ is ``ready'' and $a$ can receive a
-        $\aseq{\quo{leader-accept},h}$ message where $h > hist\left[ a
-        \right]$. The effect of the action is to receive the message,
-        to update $hist\left[ a \right]$ to $h$, and to broadcast the
-        message $\aseq{\quo{ack},hist\left[ a \right]\star c}$ to the
-        learners.
-    \item A $RcvAck\left( l \right)$ transition, for a learner $l$, is
-        enabled when $l$ can receive a message $\aseq{\quo{ack},h}$
-        from an acceptor $a$. Its effect is to receive the message and
-        to set $accepted\left[ l \right]\left[ a \right]$ to $h$.
-    \item A $Learn\left( l, h \right)$ transition, for a learner $l$
-        and a history $h$, is enabled when $h > Max\left(
-        learned\left[ l \right] \right)$ and there is a learn quorum
-        $Q$ of acceptors such that $h = GLB\left( \aset{accepted\left[
-        l \right]\left[ a \right] : a \in Q }\right)$. Its effect is
-        to insert $h$ in $learned\left[ l \right]$.
-    \item A $Stop\left( s \right)$ transition, for a switcher $s$,
-        always enabled, broadcasts a $\aseq{\quo{stop}}$ message to all
-        the acceptors.  This transition models a switcher being
-        instructed by the scheduling policy to abort the current round.
-    \item A $Stop\left( a \right)$ transition, for an acceptor $a$,
-        enabled when $a$ can receive a $\aseq{\quo{stop}}$ message
-        from a switcher $s$. Its effect is to receive the message, set
-        $status\left[ a \right]$ to ``stopped'', and to send the
-        message $\aseq{\quo{stopped},hist\left[ a \right]}$ to the
-        switcher $s$.
-    \item A $RcvStopped\left( s \right)$ transition, for a switcher
-        $s$, enabled when $s$ can receive a $\aseq{\quo{stopped},h}$
-        message from an acceptor $a$. Its effect is to receive the
-        message and to set $stopped\left[ s \right]\left[ a \right]$
-        to $h$.
-    \item An $Abort\left( s, h \right)$ transition, for a switcher $s$
-        and a history $h$, is enabled when there exists an abort quorum
-        $R$ such that, for every member $a$ of $R$, $stopped\left[ s
-        \right]\left[ a \right]$ is not $false$ and $h \in
-        SafeAborts\left( R \right)$, where $SafeAborts\left( R
-        \right)$ is computed using the array $stopped\left[ a
-        \right]$. Its effect is to insert  $h$ in the set $aborts$. 
-\end{compactitem}
-Compared to the rounds of the original Classic Paxos, Classic Paxos ACA
-rounds add an extra level of indirection when changing round: in the
-original rounds, the leader of the new round directly queries the
-acceptors of the preceeding rounds, whereas in ACA rounds, the
-switchers query the acceptors and then send their abort histories to
-the leader. However, changing round is an infrequent event, therefore
-this difference should not have much practical impact on performance.
-\begin{thm}
-  Classic Paxos ACA rounds satisfy Abortable Chain Agreement.
-\end{thm}
-\begin{proof}
-  A Classic Paxos ACA round implements the specification of Abortable
-  Chain Agreement under the refinement mapping consisting in
-  projecting the state of the round onto the variables $proposed$,
-  $inits$, $aborts$, and, for every learner $l$, $learned\left[ l
-  \right]$.
-\end{proof}
-
-\subsection{Multi-Coordinated Paxos, Chain Replication, and Ring Paxos.}
-
-Multi-Coordinated Paxos, Chain Replication, and Ring Paxos rounds have a similar
-structure as Classic Paxos rounds but differ in the way the leader is
-implemented and in the way that the histories of the leader are disseminated to
-the other acceptors.
-
-Multi-Coordinated Paxos avoids depending on a unique leader by using a
-distributed leader implementation that can make progress as long as a
-strict majority of the acceptors are correct, trading off resilience
-for a higher time complexity. Conservative ACA
-rounds can be implemented like with Classic Paxos, replacing the
-leader by the distributed leader of Multi-Coordinated Paxos.
-
-In Chain Replication, acceptors are arranged in a chain whose head is
-the leader of the round. As in Classic Paxos, abort histories and
-proposed commands are sent to the leader. However, instead of the
-leader directly sending its local history to all the other acceptors,
-the local history of the leader is
-forwarded along the chain from one acceptor to the other until it
-reaches the last acceptor, which sends the history to the learners. 
-By spreading the load more evenly on the all the servers, Chain Replication
-can achieve higher throughput that Classic Paxos.
-Conservative ACA rounds can be implemented like with Classic Paxos,
-replacing the broadcast of the leader by the propagation of the leader's history
-along the chain of servers.
-
-Ring Paxos optimizes Chain Replication by having the leader broadcast
-commands to the acceptors using IP multicast and only sending hashes
-of the commands along the chain. Ring Paxos round can be modified to implement
-conservative ACA rounds in a similar way as for Chain Replication.
-
-\subsection{Fast Paxos}
-
-Fast Paxos is composed of two types of rounds: classic rounds, like in Classic
-Paxos, and fast rounds.
-The fast rounds can be transformed into
-implementations of fast ACA rounds. Instead of sending their
-commands to a leader, as in Classic Paxos, the proposers of fast
-rounds directly broadcast their commands to all the acceptors. As long
-as commands are received in the same order by a learn quorum of
-acceptors or are received in different orders but commute, commands can be learned in two message-propagation delays instead of three in Classic Paxos.
-However, if non-commuting commands are received in different orders by
-several acceptors, the local history of the acceptors may not form a
-chain anymore and the rounds must abort. The fast rounds of Fast
-Paxos implement fast ACA rounds.
-
-Fast Paxos ACA rounds are specified using the same set of variables as for
-Classic Paxos, initialized to the same values.
-The next-state relation is composed of the following types of
-transitions, where the definitions of learn quorums, abort quorums,
-and of the $SafeAborts\left( R \right)$ operator are the ones of fast
-ACA rounds.
-\begin{compactitem}
-    \item A $Propose\left( p,c \right)$ transition, for a proposer $p$
-        and a command $c$, not guarded, inserts $c$ in the set
-        $proposed$ and broadcasts the message $\aseq{\quo{prop},c}$ to
-        all the acceptors.
-    \item An $Init\left( s, h \right)$ transition, for a switcher $s$
-        and a history $h$, not guarded, inserts $h$ in the set $inits$ and
-        broadcasts the message $\aseq{\quo{init},h}$ to all the acceptors.
-    \item A $WakeUp\left( a \right)$ transition, for an acceptor $a$,
-        is enabled when $status\left[ a \right]$ is ``idle'' and $a$
-        can receive an $\aseq{\quo{init},h}$ message. Its effect is to
-        receive the message, set $hist\left[ a \right]$ to $h$, and to
-        set $status\left[ a \right]$ to ``ready''.
-    \item An $Accept\left( a \right)$ transition, for an acceptor $a$,
-        is enabled when $status\left[ a \right]$ is ``ready'' and $a$
-        can receive a $\aseq{\quo{prop},c}$ message. The effect of the
-        action is to receive the message, to update $hist\left[ a
-        \right]$ to $hist\left[ a \right]\star c$, and to broadcast
-        the message $\aseq{\quo{ack},hist\left[ a \right]\star c}$ to
-        the learners.
-    \item The $RcvAck\left( l \right)$ transitions, for a learner $l$,
-        the $Learn\left( l, h \right)$ transitions, for a learner $l$
-        and a history $h$, the $Stop\left( s \right)$ transitions, for
-        a switcher $s$, $Stop\left( a \right)$ transitions, for an
-        acceptor $a$, the $RcvStopped\left( s \right)$ transitions,
-        for a switcher $s$, and the $Abort\left( s, h \right)$
-        transitions, for a switcher $s$ and a history $h$, which are
-        all the same as in Classic Paxos ACA rounds.
-\end{compactitem}
-
-As for Classic Paxos ACA rounds, Fast Paxos ACA rounds implement ACA
-under the refinement mapping consisting projecting the state of the
-round onto the variables $proposed$, $inits$, $aborts$, and, for every
-learner $l$, $learned\left[ l \right]$.
-
-\newpage
-
-\printbibliography
-
-\begin{appendix}
-
-  \label{sec:isaproofs}
-
-%  \includepdf[pages={1},pagecommand={\section{Mechanical Proof of the
-%  Composition Theorem in Isabelle/HOL}\label{sec:isaproofs}},scale=1]{../IsabellePDFS/podctheories.pdf}
-%  \includepdf[pages={2-},pagecommand={},scale=1]{../IsabellePDFS/podctheories.pdf}
-
-\end{appendix}
-
-\end{document}

### 2015-05-14 paper/intro.tex
Work on the introduction.

@@
 \cite{MaoJunqueiraMarzullo08MenciusBuildingEfficientReplicatedStateMachine},
-Fast Mencius \cite{WeiETAL13FastMenciusMenciusLowCommitLatency}, the Chandra-Toueg \cite{ChandraToueg96UnreliableFailureDetectorsReliableDistributedSystems}, the Ben-Or algorithm \cite{BenOr83AnotherAdvantageFreeChoiceCompletelyAsynchronous}, Raft \cite{OngaroOusterhout14SearchUnderstandableConsensusAlgorithm}.
+Fast Mencius \cite{WeiETAL13FastMenciusMenciusLowCommitLatency}, the Chandra-Toueg \cite{ChandraToueg96UnreliableFailureDetectorsReliableDistributedSystems}, the Ben-Or algorithm \cite{BenOr83AnotherAdvantageFreeChoiceCompletelyAsynchronous}, and Raft \cite{OngaroOusterhout14SearchUnderstandableConsensusAlgorithm}.
 
@@
 
-Devising new SMR algorithms, optimized with particular situations or metrics in mind, is hard: today, several decades after the first SMR algorithms \cite{Lamport98ParttimeParliament,BirmanJoseph87ReliableCommunicationPresenceFailures,OkiLiskov88ViewstampedReplicationGeneralPrimaryCopy,DworkLynchStockmeyer84ConsensusPresencePartialSynchronyPreliminaryVersion} were published, devising new SMR algorithms is still hard enough to justify a publication in a top conference \cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments,OngaroOusterhout14SearchUnderstandableConsensusAlgorithm}.
+However, devising new SMR algorithms, optimized with particular situations or metrics in mind, is hard: today, several decades after the first SMR algorithms \cite{Lamport98ParttimeParliament,BirmanJoseph87ReliableCommunicationPresenceFailures,OkiLiskov88ViewstampedReplicationGeneralPrimaryCopy,DworkLynchStockmeyer84ConsensusPresencePartialSynchronyPreliminaryVersion} were published, devising new SMR algorithms is still hard enough to justify a publication in a top conference \cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments,OngaroOusterhout14SearchUnderstandableConsensusAlgorithm}.
 
@@
 
-We propose a framework that simplifies devising new SMR algorithms by decomposing them into independent and composable modules.
-Crucially, in our framework, a SMR module can be analyzed, tested, and proved correct  once and for all. 
-Our composition theorem guarantees that composing any number of correct modules can be composed to form a correct SMR algorithm.
+We propose a framework to make existing SMR algorithms modular and add new optimizations incrementally, without modification to the existing. 
+An algorithm in our framework is composed of independent modules that can run, abort their execution, and pass the baton to any other module.
+Our framework has the following features:
+\begin{itemize}
+  \item Effortless re-use of existing modules: A SMR algorithm expressed in our framework can can be analyzed, tested, and proved correct in isolation and once and for all.
+Our composition theorem guarantees that the composition of any number of correct modules is a correct SMR algorithm.
+  \item Simplified algorithms thanks to a clear separation of concerns: one can devise a new specialized SMR algorithm focusing only on the situation for which it is specialized.
+A composable SMR algorithm is not required to make progress in all situations: if it does not, the system can switch to another SMR algorithm.
 
-A composite SMR algorithm built with our framework works by running the SMR module most appropriate at any given time, dynamically changing SMR module when needed.
+  \item Seamless combination of different SMR algorithms: one can combine different SMR algorithms, optimized for different situations and metrics, without any changes and with guaranteed correctness.
+  \item Practical
+    \begin{itemize}
+  \item Tolerating $k$ fault with $k+1$ replicas and a reliable reconfiguration master. 
+  \item Built-in support for crucial practical optimizations such as executing commuting commands without synchronization and expediting the execution of read-only commands.
+    \end{itemize}
+\end{itemize}
+
+A composite SMR algorithm built with our framework works by running the SMR module deemed the most appropriate at any given time, dynamically changing SMR module when needed.
 Thus algorithms designers who want to create a new SMR algorithm, optimized for particular conditions, can focus only on the particular conditions of interest and reuse existing modules to handle the other cases.
@@
 To illustrate our approach, let us consider Fast Paxos, one of the first examples of SMR algorithm which can change mode dynamically.
-Fast Paxos is of round-based SMR algorithm that can change mode.
-Each round is categorized as a fast or a normal round.
-Fast rounds process request faster than normal rounds but may fail to make progress in case of contention, at which point a normal round can take over the execution and make progress, later passing the baton back to a fast round when contention becomes low again.
-As in the original Paxos algorithm, normal rounds can also succeed each other in order to change
-leader.
-Each kind of round maintains different state invariants, thus a normal round must know which past rounds are fast or normal rounds in order to take over the execution safely.
+Fast Paxos is a round-based SMR algorithm that has two types of rounds: fast rounds and normal rounds.
+Fast rounds process requests faster than normal rounds in contention-free executions but fall back to a slower scheme under contention.
+Normal rounds are the same as in the classic Paxos.
+Any process can initiate a round change at any point during execution: this allows switching between fast and normal rounds as well as, similarly to classic Paxos, changing the leader from one normal round to
+another.
+Each kind of round maintains different invariants relating the state of each process and the the values that might have been chosen in the previous rounds, thus a normal round must know which past rounds are fast or normal rounds in order to take over the execution safely.
 When adding more kinds of rounds such an approach will quickly become unmanageable, as each type of rounds must have one case for each possible other kind of round, making $n^2$ cases to consider in total.

### 2015-05-14 paper/intro.tex
Work on the introduction.

@@
 
-However, devising new SMR algorithms, optimized with particular situations or metrics in mind, is hard: today, several decades after the first SMR algorithms \cite{Lamport98ParttimeParliament,BirmanJoseph87ReliableCommunicationPresenceFailures,OkiLiskov88ViewstampedReplicationGeneralPrimaryCopy,DworkLynchStockmeyer84ConsensusPresencePartialSynchronyPreliminaryVersion} were published, devising new SMR algorithms is still hard enough to justify a publication in a top conference \cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments,OngaroOusterhout14SearchUnderstandableConsensusAlgorithm}.
+However, devising new SMR algorithms, optimized with particular situations or metrics in mind, is hard: today, more than two decades after the first SMR algorithms \cite{Lamport98ParttimeParliament,BirmanJoseph87ReliableCommunicationPresenceFailures,OkiLiskov88ViewstampedReplicationGeneralPrimaryCopy,DworkLynchStockmeyer84ConsensusPresencePartialSynchronyPreliminaryVersion} were published, devising new SMR algorithms is still hard enough to justify a publication in a top conference \cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments,OngaroOusterhout14SearchUnderstandableConsensusAlgorithm}.
 
@@
 An algorithm in our framework is composed of independent modules that can run, abort their execution, and pass the baton to any other module.
-Our framework has the following features:
-\begin{itemize}
-  \item Effortless re-use of existing modules: A SMR algorithm expressed in our framework can can be analyzed, tested, and proved correct in isolation and once and for all.
-Our composition theorem guarantees that the composition of any number of correct modules is a correct SMR algorithm.
-  \item Simplified algorithms thanks to a clear separation of concerns: one can devise a new specialized SMR algorithm focusing only on the situation for which it is specialized.
-A composable SMR algorithm is not required to make progress in all situations: if it does not, the system can switch to another SMR algorithm.
 
-  \item Seamless combination of different SMR algorithms: one can combine different SMR algorithms, optimized for different situations and metrics, without any changes and with guaranteed correctness.
-  \item Practical
-    \begin{itemize}
-  \item Tolerating $k$ fault with $k+1$ replicas and a reliable reconfiguration master. 
-  \item Built-in support for crucial practical optimizations such as executing commuting commands without synchronization and expediting the execution of read-only commands.
-    \end{itemize}
-\end{itemize}
+Crucially, when adding a new module to an existing algorithm, all the performance analysis, tests, and proofs of the existing modules remain valid and only the new SMR module needs to be analyzed: Our composition theorem guarantees that the composition of any number of correct modules is a correct SMR algorithm, thus one can combine different SMR algorithms, optimized for different situations and metrics, without any changes and with guaranteed correctness.
+Our framework therefore enables incremental development of SMR algorithms, where each new development can build on the previous ones instead of starting again from scratch.
 
+Moreover, our framework enables a clear separation of concerns  which simplifies the designing individual modules.
 A composite SMR algorithm built with our framework works by running the SMR module deemed the most appropriate at any given time, dynamically changing SMR module when needed.
-Thus algorithms designers who want to create a new SMR algorithm, optimized for particular conditions, can focus only on the particular conditions of interest and reuse existing modules to handle the other cases.
-All the performance analysis, tests, and proofs of the existing SMR modules remain valid and only the new SMR needs to be analyzed.
-Our framework thus enables incremental development of SMR algorithms, where each new development can build on the previous ones instead of starting again from scratch.
+Therefore a composable SMR algorithm is not required to make progress in all situations: if it does not, the system can switch to another SMR algorithm.
+Thus algorithms designers who want to create a new SMR algorithm, optimized for particular conditions, can focus only on the particular conditions of interest, knowing that the other existing modules will handle the other cases.
 
-An important practical aspect of SMR is fast reconfiguration to replace faulty components, without which the stability of a system is inevitably compromised.
+The key novelty of our framework lies in its practical applicability.
+In practice, a SMR algorithm must support fast reconfiguration to replace faulty components, without which the long-term stability of a system is inevitably compromised.
+Our framework clearly separates the execution of commands from the reconfiguration decisions (including changing SMR module), allowing the use of an external reconfiguration master.
+This enables tolerating $k$ failures with $k+1$ replicas, a very
+important improvement over the traditional $2k+1$ needed replicas.
+Moreover, in practice, commands sent to a SMR algorithm often commute.
+Our framework is based on Generalized Consensus, a formulation of the State-Machine Replication problem that allows processing commutative commands without synchronization.
 
-To be usable in  practice, SMR implementations must have high performance, be it throughput, latency, or some other metric depending on the setting, and resort to many optimizations which a framework like ours must support to be practical.
-Two key practical aspects of SMR are the ability to reconfigure the system to replace faulty components and the ability to process commutative commands without synchronization.
-There lies our key contribution: 
-Our framework is based on Generalized Consensus, a formulation of the State-Machine Replication problem that allows processing commutative commands without synchronization.
+TODO: expand on the preceding paragraph.
+
+Our framework is fully formalized and its composition theorem proved in Isabelle/HOL.
+Moreover, to quickly check concrete modules, we also provide a TLA+ specification of our framework which allows model-checking algorithms with the TLC model-checker.
+
+TODO: the two abstract specs.
 
@@
 another.
-Each kind of round maintains different invariants relating the state of each process and the the values that might have been chosen in the previous rounds, thus a normal round must know which past rounds are fast or normal rounds in order to take over the execution safely.
+Each kind of round maintains different invariants relating the state of each process and the values that might have been chosen in the previous rounds, thus a normal round must know which past rounds are fast or normal rounds in order to take over the execution safely.
 When adding more kinds of rounds such an approach will quickly become unmanageable, as each type of rounds must have one case for each possible other kind of round, making $n^2$ cases to consider in total.
@@
 Our notion of Composable Generalized Consensus defines a common interface and a common correctness property for a round, abstracting over the particular round implementations and making all rounds compatible with each other.
-Moreover, our interface for rounds enable a straightforward crucial optimization seen in practical systems: introducing an external reconfiguration master which manages the addition of new replicas and the removal of failed ones.
+Moreover, our interface for rounds enable a crucial optimization seen in practical systems: introducing an external reconfiguration master which manages the addition of new replicas and the removal of failed ones.
 Interestingly, as explained in \cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}, in a real system it is often possible to implement a reliable reconfiguration master which allows to tolerate $f$ faults with $f+1$ replicas instead of the traditional $2f+1$ replicas.
-
 The reconfiguration master can also make the choice of the new type of round to execute, providing the system's reconfiguration policy.

### 2015-05-14 paper/intro.tex
Work on the introduction

@@
 
-The key novelty of our framework lies in its practical applicability.
+The key novelty of our framework lies in its practical applicability, allowing the execution of commutative commands without synchronization and reducing the replication overhead from $2*k+1$ replicas to $k+1$ replicas, to tolerate $k$ replica crashes, using a reliable reconfiguration master.
+
+Allowing the execution of commuting commands to proceed without synchronization is an important practical optimization, as commands may often commute in practice, for example when operating on disjoint parts of a data structure.
+To enable the synchronization-free execution of commuting commands, we have based our framework on Generalized Consensus \cite{Lamport05GeneralizeConsensus}.
+Traditional SMR implementations are implemented by having all the replicas execute the same sequence of commands, using a separate instance of a consensus algorithm to agree on the command to execute at each index of the command sequence.
+Using separate instances of a consensus algorithm for each index in the sequence of commands
+makes it hard to optimize the execution of commuting command because commutativity is not a property of a single command.
+Instead of restricting agreement on a single value, as in the consensus problem, in the  Generalized Consensus problem the set of replica servers must agree on the state of a growing data structure called a command-structure set.
+A command-structure set is a data structure with an append operation such that the order in which two commuting commands are appended does not change the resulting state.
+Thanks to this property, two replicas executing commuting commands in a different order will
+not end up in a different state and can continue executing without further synchronization.
+This property can be leveraged to reduce the latency of processing one command to two message delays instead of three in message passing systems.
+As we will show, it also enables the processing of read-only commands in two message delays, although read-only commands do not commute with write commands.
+
 In practice, a SMR algorithm must support fast reconfiguration to replace faulty components, without which the long-term stability of a system is inevitably compromised.
 Our framework clearly separates the execution of commands from the reconfiguration decisions (including changing SMR module), allowing the use of an external reconfiguration master.
-This enables tolerating $k$ failures with $k+1$ replicas, a very
-important improvement over the traditional $2k+1$ needed replicas.
-Moreover, in practice, commands sent to a SMR algorithm often commute.
-Our framework is based on Generalized Consensus, a formulation of the State-Machine Replication problem that allows processing commutative commands without synchronization.
+This enables tolerating $k$ failures with $k+1$ replicas, an important improvement over the traditional $2k+1$ needed replicas.
 
-TODO: expand on the preceding paragraph.
+To demonstrate the practical applicability of our framework we provide composable versions of the generalized variations of popular algorithms like the classic Paxos, Fast Paxos, and multi-coordinated Paxos.
+We have checked the composability of these algorithms with the TLC model-checker and we provide all the specifications needed to replicate these experiments.
 
-Our framework is fully formalized and its composition theorem proved in Isabelle/HOL.
-Moreover, to quickly check concrete modules, we also provide a TLA+ specification of our framework which allows model-checking algorithms with the TLC model-checker.
-
-TODO: the two abstract specs.
+Our framework is also fully formalized and its composition theorem proved in the foundational interactive theorem prover Isabelle/HOL, certifying the correctness of the composition theorem at the deepest level.
 

### 2015-05-14 paper/intro.tex
Work on the introduction.

--- 2015-05-14 paper/intro.tex
@@
 replicate a service over several servers, each executing the same
-sequence of (deterministic) commands. 
+sequence of deterministic commands.
+In such a setup, any replica can be used to determine the result of a command.
+Therefore, if a replica crashes, clients can be transparently redirected to another one and will have the illusion of accessing a centralized service at all times.
+
 Making SMR algorithms efficient has been a long-standing
@@
 \cite{MaoJunqueiraMarzullo08MenciusBuildingEfficientReplicatedStateMachine},
-Fast Mencius \cite{WeiETAL13FastMenciusMenciusLowCommitLatency}, the Chandra-Toueg \cite{ChandraToueg96UnreliableFailureDetectorsReliableDistributedSystems}, the Ben-Or algorithm \cite{BenOr83AnotherAdvantageFreeChoiceCompletelyAsynchronous}, and Raft \cite{OngaroOusterhout14SearchUnderstandableConsensusAlgorithm}.
+Fast Mencius \cite{WeiETAL13FastMenciusMenciusLowCommitLatency}, the Chandra-Toueg algorithm \cite{ChandraToueg96UnreliableFailureDetectorsReliableDistributedSystems}, the Ben-Or algorithm \cite{BenOr83AnotherAdvantageFreeChoiceCompletelyAsynchronous}, and Raft \cite{OngaroOusterhout14SearchUnderstandableConsensusAlgorithm}.
 
@@
 
-However, devising new SMR algorithms, optimized with particular situations or metrics in mind, is hard: today, more than two decades after the first SMR algorithms \cite{Lamport98ParttimeParliament,BirmanJoseph87ReliableCommunicationPresenceFailures,OkiLiskov88ViewstampedReplicationGeneralPrimaryCopy,DworkLynchStockmeyer84ConsensusPresencePartialSynchronyPreliminaryVersion} were published, devising new SMR algorithms is still hard enough to justify a publication in a top conference \cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments,OngaroOusterhout14SearchUnderstandableConsensusAlgorithm}.
+However, devising new SMR algorithms, optimized with particular situations or metrics in mind, is hard: today, more than two decades after the first SMR algorithms \cite{Lamport98ParttimeParliament,BirmanJoseph87ReliableCommunicationPresenceFailures,OkiLiskov88ViewstampedReplicationGeneralPrimaryCopy,DworkLynchStockmeyer84ConsensusPresencePartialSynchronyPreliminaryVersion} were proposed, devising new SMR algorithms is still hard enough to justify a publication in a top conference \cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments,OngaroOusterhout14SearchUnderstandableConsensusAlgorithm}.
 
-The problem is that it is very hard to reuse existing algorithms and ideas because SMR algorithms are not modular.
+The problem is that it is hard to reuse existing algorithms and ideas because SMR algorithms are not modular.
 When devising a new algorithm, all the careful analysis, tests, and proofs have to be done again from scratch.
 
-We propose a framework to make existing SMR algorithms modular and add new optimizations incrementally, without modification to the existing. 
-An algorithm in our framework is composed of independent modules that can run, abort their execution, and pass the baton to any other module.
+We propose a framework to make existing Generalized Consensus algorithms modular and to enable adding new optimizations incrementally, without modification to the existing.
+Generalized consensus, abbreviated GC, is a precise formulation of State-Machine Replication that subsumes consensus.
+An algorithm in our framework is composed of independent modules that can run, abort their execution, and pass the baton to any other module on the fly.
 
-Crucially, when adding a new module to an existing algorithm, all the performance analysis, tests, and proofs of the existing modules remain valid and only the new SMR module needs to be analyzed: Our composition theorem guarantees that the composition of any number of correct modules is a correct SMR algorithm, thus one can combine different SMR algorithms, optimized for different situations and metrics, without any changes and with guaranteed correctness.
-Our framework therefore enables incremental development of SMR algorithms, where each new development can build on the previous ones instead of starting again from scratch.
+Crucially, when adding a new module to an existing algorithm, all the performance analysis, tests, and proofs of the existing modules remain valid and only the new GC module needs to be analyzed: Our composition theorem guarantees that the composition of any number of correct modules is a correct GC algorithm, thus one can combine different GC algorithms, optimized for different situations and metrics, without any changes and with guaranteed correctness.
+Our framework therefore enables incremental development of GC algorithms, where each new development can build on the previous ones instead of starting again from scratch.
 
-Moreover, our framework enables a clear separation of concerns  which simplifies the designing individual modules.
-A composite SMR algorithm built with our framework works by running the SMR module deemed the most appropriate at any given time, dynamically changing SMR module when needed.
-Therefore a composable SMR algorithm is not required to make progress in all situations: if it does not, the system can switch to another SMR algorithm.
-Thus algorithms designers who want to create a new SMR algorithm, optimized for particular conditions, can focus only on the particular conditions of interest, knowing that the other existing modules will handle the other cases.
+Moreover, our framework enables a clear separation of concerns  which simplifies the design of new individual modules.
+A composite GC algorithm built with our framework works by running the GC module deemed the most appropriate at any given time, dynamically changing GC module when needed.
+Therefore a composable GC algorithm is not required to make progress in all situations: if it does not make progress, the system can switch to another GC module.
+Thus algorithms designers who want to create a new GC algorithm, optimized for particular conditions, can focus only on the particular conditions of interest, knowing that the other existing modules will handle the other cases.
 
-The key novelty of our framework lies in its practical applicability.
+The key novelty of our framework lies in its practical applicability, allowing the execution of commutative commands without synchronization and reducing the replication overhead from $2*k+1$ replicas to $k+1$ replicas, to tolerate $k$ replica crashes, using a reliable external reconfiguration master.
+
+Allowing the execution of commuting commands to proceed without synchronization is an important practical optimization, as commands may often commute in practice, for example when operating on disjoint parts of a data structure.
+To enable the synchronization-free execution of commuting commands, we have based our framework on Generalized Consensus \cite{Lamport05GeneralizeConsensus}.
+Traditional SMR implementations use a separate instance of a consensus algorithm to agree on the command to execute at each index of the command sequence on which the replicas must agree.
+Using separate consensus instances makes it hard to optimize the execution of commuting command because commutativity is not a property of a single command.
+Instead of restricting agreement to a single value, the  Generalized Consensus problem requires that set of replica servers agree on the state of a growing data structure called a command-structure set.
+A command-structure set is a data structure with an append operation such that the order in which two commuting commands are appended does not change the resulting state.
+Thanks to this property, two replicas executing commuting commands in a different order will
+not end up in a different state and can continue executing without further synchronization.
+In message-passing systems, this property can be leveraged to reduce the latency of processing one command to two message delays instead of three.
+As we will show, it also enables the processing of read-only commands in two message delays, although read-only commands do not commute with write commands.
+
 In practice, a SMR algorithm must support fast reconfiguration to replace faulty components, without which the long-term stability of a system is inevitably compromised. 
-Our framework clearly separates the execution of commands from the reconfiguration decisions (including changing SMR module), allowing the use of an external reconfiguration master.
-This enables tolerating $k$ failures with $k+1$ replicas, a very
-important improvement over the traditional $2k+1$ needed replicas.
-Moreover, in practice, commands sent to a SMR algorithm often commute.
-Our framework is based on Generalized Consensus, a formulation of the State-Machine Replication problem that allows processing commutative commands without synchronization.
+Our framework abstracts over how the reconfiguration decisions are made, providing an interface allowing to plug an external reconfiguration master implementing the reconfiguration policy of the system.
+The master decides which GC module to run next and which replica servers will participate, removing failed replicas and adding new ones.
+As we will explain, this enables tolerating $k$ failures with $k+1$ replicas, an important improvement over the traditional $2k+1$ needed replicas.
 
-TODO: expand on the preceding paragraph.
+To demonstrate the practical applicability of our framework we provide composable versions of the generalized variations of popular consensus algorithms: classic Paxos, Fast Paxos, Chain Replication, and multi-coordinated Paxos.
+We have checked the correctness of these algorithms with the TLC model-checker and we provide all the specifications needed to replicate these experiments.
 
-Our framework is fully formalized and its composition theorem proved in Isabelle/HOL.
-Moreover, to quickly check concrete modules, we also provide a TLA+ specification of our framework which allows model-checking algorithms with the TLC model-checker.
+Our framework is also fully formalized and its composition theorem proved in Isabelle/HOL, a foundational interactive theorem prover, thereby certifying the correctness of the composition theorem at the deepest level.
 
-TODO: the two abstract specs.
-
-At the heart of our framework lies a simple abstraction of the familiar notion of rounds (also called epochs, ballots, or views in the literature) used in most SMR algorithms.
-To illustrate our approach, let us consider Fast Paxos, one of the first examples of SMR algorithm which can change mode dynamically.
+At the heart of our framework lies an abstraction, called Composable Generalized Consensus, of the familiar notion of round (also called epoch, ballot, or view in the literature) used in most SMR algorithms.
+As an example, let us consider Fast Paxos, one of the first SMR algorithm which can change strategy dynamically.
 Fast Paxos is a round-based SMR algorithm that has two types of rounds: fast rounds and normal rounds.
 Fast rounds process requests faster than normal rounds in contention-free executions but fall back to a slower scheme under contention.
-Normal rounds are the same as in the classic Paxos.
-Any process can initiate a round change at any point during execution: this allows switching between fast and normal rounds as well as, similarly to classic Paxos, changing the leader from one normal round to
-another.
-Each kind of round maintains different invariants relating the state of each process and the values that might have been chosen in the previous rounds, thus a normal round must know which past rounds are fast or normal rounds in order to take over the execution safely.
-When adding more kinds of rounds such an approach will quickly become unmanageable, as each type of rounds must have one case for each possible other kind of round, making $n^2$ cases to consider in total.
+Normal rounds are the same as in the classic Paxos and are slower than fast rounds in contention-free executions.
+Any process can initiate a round change at any point during execution: this allows switching between fast and normal rounds as well as changing the
+leader from one normal round to the next, similarly to classic Paxos.
+Each kind of round maintains different invariants relating the state of each process and the values that might have been chosen in the previous rounds, thus a given round must know which past rounds are fast or normal rounds in order to take over the execution safely.
+When adding more kinds of rounds such an approach will quickly become unmanageable, as each type of rounds must have one case for each possible other kind of
+round, making $O(n^2)$ cases to consider in total.
 
-Our notion of Composable Generalized Consensus defines a common interface and a common correctness property for a round, abstracting over the particular round implementations and making all rounds compatible with each other.
-Moreover, our interface for rounds enable a crucial optimization seen in practical systems: introducing an external reconfiguration master which manages the addition of new replicas and the removal of failed ones.
+Our notion of Composable Generalized Consensus defines a common interface and a common correctness property for a round, abstracting over the particular round implementations and making all types of rounds compatible with each other.
+Moreover, our interface for rounds enables a crucial optimization seen in practical systems: introducing an external reconfiguration master which manages the addition of new replicas and the removal of failed ones.
 Interestingly, as explained in \cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}, in a real system it is often possible to implement a reliable reconfiguration master which allows to tolerate $f$ faults with $f+1$ replicas instead of the traditional $2f+1$ replicas.

### 2015-05-14 paper/intro.tex
Work on the intro, added /web/

@@
 
-The State-Machine Replication technique (abbreviated SMR) allows
-building reliable services on top of unreliable hardware. The basic idea is to 
-replicate a service over several servers, each executing the same
-sequence of deterministic commands.
-In such a setup, any replica can be used to determine the result of a command.
-Therefore, if a replica crashes, clients can be transparently redirected to another one and will have the illusion of accessing a centralized service at all times.
+The State-Machine Replication technique (abbreviated SMR) allows building
+reliable services on top of unreliable hardware. The basic idea is to
+replicate a service over several servers, each executing the same sequence of
+deterministic commands. In such a setup, any replica can be used to determine
+the result of a command. Therefore, if a replica crashes, clients can be
+transparently redirected to another one and will have the illusion of accessing
+a centralized service at all times.
 
-Making SMR algorithms efficient has been a long-standing
-challenge and a multitude of algorithms have been proposed. Examples
-include Paxos \cite{lamport2001paxos}, Fast Paxos \cite{Lamport06FastPaxos}, Disk Paxos
-\cite{GafniLamport03DiskPaxos}, Chain Replication
+Making SMR algorithms efficient has been a long-standing challenge
+and a multitude of algorithms have been proposed. Examples include
+Paxos \cite{lamport2001paxos}, Fast Paxos \cite{Lamport06FastPaxos},
+Disk Paxos \cite{GafniLamport03DiskPaxos}, Chain Replication
 \cite{RenesseSchneider04ChainReplicationSupportingHighThroughputAvailability},
-Ring Paxos
-\cite{MarandiETAL10RingPaxosHighthroughputAtomicBroadcastProtocol},
+Ring Paxos \cite{MarandiETAL10RingPaxosHighthroughputAtomicBroadcastProtocol},
 Egalitarian Paxos
 \cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments},
-Multi-Coordinated Paxos
-\cite{CamargosSchmidtPedone07MulticoordinatedPaxos}, Cheap Paxos
-\cite{LamportMassa04CheapPaxos}, Vertical Paxos
-\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication},
-Paxos-MIC
-\cite{HurfinMoiseNarzul11AdaptiveFastPaxosMakingQuickEverlasting},
-Mencius
+Multi-Coordinated Paxos \cite{CamargosSchmidtPedone07MulticoordinatedPaxos},
+Cheap Paxos \cite{LamportMassa04CheapPaxos}, Vertical Paxos
+\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}, Paxos-MIC
+\cite{HurfinMoiseNarzul11AdaptiveFastPaxosMakingQuickEverlasting}, Mencius
 \cite{MaoJunqueiraMarzullo08MenciusBuildingEfficientReplicatedStateMachine},
-Fast Mencius \cite{WeiETAL13FastMenciusMenciusLowCommitLatency}, the Chandra-Toueg algorithm \cite{ChandraToueg96UnreliableFailureDetectorsReliableDistributedSystems}, the Ben-Or algorithm \cite{BenOr83AnotherAdvantageFreeChoiceCompletelyAsynchronous}, and Raft \cite{OngaroOusterhout14SearchUnderstandableConsensusAlgorithm}.
+Fast Mencius \cite{WeiETAL13FastMenciusMenciusLowCommitLatency},
+the Chandra-Toueg algorithm
+\cite{ChandraToueg96UnreliableFailureDetectorsReliableDistributedSystems}, the
+Ben-Or algorithm \cite{BenOr83AnotherAdvantageFreeChoiceCompletelyAsynchronous},
+and Raft \cite{OngaroOusterhout14SearchUnderstandableConsensusAlgorithm}.
 
-With the wide use of cloud computing, improvements in the performance and efficiency of SMR algorithms have the potential to impact most online services: advanced SMR algorithms can be implemented by cloud computing service providers and benefit all their customer's applications.
+With the wide use of cloud computing, improvements in the performance and
+efficiency of SMR algorithms have the potential to impact most online services:
+advanced SMR algorithms can be implemented by cloud computing service providers
+and benefit all their customer's applications.
 
-However, devising new SMR algorithms, optimized with particular situations or metrics in mind, is hard: today, more than two decades after the first SMR algorithms \cite{Lamport98ParttimeParliament,BirmanJoseph87ReliableCommunicationPresenceFailures,OkiLiskov88ViewstampedReplicationGeneralPrimaryCopy,DworkLynchStockmeyer84ConsensusPresencePartialSynchronyPreliminaryVersion} were proposed, devising new SMR algorithms is still hard enough to justify a publication in a top conference \cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments,OngaroOusterhout14SearchUnderstandableConsensusAlgorithm}.
+However, devising new SMR algorithms, optimized with
+particular situations or metrics in mind, is hard: today,
+more than two decades after the first SMR algorithms
+\cite{Lamport98ParttimeParliament,BirmanJoseph87ReliableCommunicationPresenceFailures,OkiLiskov88ViewstampedReplicationGeneralPrimaryCopy,DworkLynchStockmeyer84ConsensusPresencePartialSynchronyPreliminaryVersion} were proposed, devising new
+SMR algorithms is still complex enough to justify a publication in a top conference
+\cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments,OngaroOusterhout14SearchUnderstandableConsensusAlgorithm}.
 
-The problem is that it is hard to reuse existing algorithms and ideas because SMR algorithms are not modular.
-When devising a new algorithm, all the careful analysis, tests, and proofs have to be done again from scratch.
+The problem is that it is difficult to reuse existing algorithms and ideas because
+SMR algorithms are not modular. When devising a new algorithm, all the careful
+analysis, tests, and proofs have to be done again from scratch.
 
-We propose a framework to make existing Generalized Consensus algorithms modular and to enable adding new optimizations incrementally, without modification to the existing.
-Generalized consensus, abbreviated GC, is a precise formulation of State-Machine Replication that subsumes consensus.
-An algorithm in our framework is composed of independent modules that can run, abort their execution, and pass the baton to any other module on the fly.
+We propose a framework to make existing Generalized Consensus algorithms modular
+and to enable adding new optimizations incrementally, without modification to
+the existing. Generalized consensus, abbreviated GC, is a precise formulation of
+State-Machine Replication that subsumes consensus. An algorithm in our framework
+is composed of independent modules that can run, abort their execution, and pass
+the baton to any other module on the fly.
 
-Crucially, when adding a new module to an existing algorithm, all the performance analysis, tests, and proofs of the existing modules remain valid and only the new GC module needs to be analyzed: Our composition theorem guarantees that the composition of any number of correct modules is a correct GC algorithm, thus one can combine different GC algorithms, optimized for different situations and metrics, without any changes and with guaranteed correctness.
-Our framework therefore enables incremental development of GC algorithms, where each new development can build on the previous ones instead of starting again from scratch.
+Crucially, when adding a new module to an existing algorithm, all the
+performance analysis, tests, and proofs of the existing modules remain valid and
+only the new GC module needs to be analyzed: Our composition theorem guarantees
+that the composition of any number of correct modules is a correct GC algorithm,
+thus one can combine different GC algorithms, optimized for different situations
+and metrics, without any changes and with guaranteed correctness. Our framework
+therefore enables incremental development of GC algorithms, where each new
+development can build on the previous ones instead of starting again from
+scratch.
 
-Moreover, our framework enables a clear separation of concerns  which simplifies the design of new individual modules.
-A composite GC algorithm built with our framework works by running the GC module deemed the most appropriate at any given time, dynamically changing GC module when needed.
-Therefore a composable GC algorithm is not required to make progress in all situations: if it does not make progress, the system can switch to another GC module.
-Thus algorithms designers who want to create a new GC algorithm, optimized for particular conditions, can focus only on the particular conditions of interest, knowing that the other existing modules will handle the other cases.
+Moreover, our framework enables a clear separation of concerns which simplifies
+the design of new individual modules. A composite GC algorithm built with our
+framework works by running the GC module deemed the most appropriate at any
+given time, dynamically changing GC module when needed. Therefore a composable
+GC algorithm is not required to make progress in all situations: if it does
+not make progress, the system can switch to another GC module. Thus algorithms
+designers who want to create a new GC algorithm, optimized for particular
+conditions, can focus only on the particular conditions of interest, knowing
+that the other existing modules will handle the other cases.
 
-The key novelty of our framework lies in its practical applicability, allowing the execution of commutative commands without synchronization and reducing the replication overhead from $2*k+1$ replicas to $k+1$ replicas, to tolerate $k$ replica crashes, using a reliable external reconfiguration master.
+The key novelty of our framework lies in its practical applicability, allowing
+the execution of commutative commands without synchronization and reducing the
+replication overhead from $2*k+1$ replicas to $k+1$ replicas, to tolerate $k$
+replica crashes, using a reliable external reconfiguration master.
 
-Allowing the execution of commuting commands to proceed without synchronization is an important practical optimization, as commands may often commute in practice, for example when operating on disjoint parts of a data structure.
-To enable the synchronization-free execution of commuting commands, we have based our framework on Generalized Consensus \cite{Lamport05GeneralizeConsensus}.
-Traditional SMR implementations use a separate instance of a consensus algorithm to agree on the command to execute at each index of the command sequence on which the replicas must agree.
-Using separate consensus instances makes it hard to optimize the execution of commuting command because commutativity is not a property of a single command.
-Instead of restricting agreement to a single value, the  Generalized Consensus problem requires that set of replica servers agree on the state of a growing data structure called a command-structure set.
-A command-structure set is a data structure with an append operation such that the order in which two commuting commands are appended does not change the resulting state.
-Thanks to this property, two replicas executing commuting commands in a different order will
-not end up in a different state and can continue executing without further synchronization.
-In message-passing systems, this property can be leveraged to reduce the latency of processing one command to two message delays instead of three.
-As we will show, it also enables the processing of read-only commands in two message delays, although read-only commands do not commute with write commands.
+Allowing the execution of commuting commands to proceed without
+synchronization is an important practical optimization, as commands may
+often commute in practice, for example when operating on disjoint parts
+of a data structure. To enable the synchronization-free execution of
+commuting commands, we have based our framework on Generalized Consensus
+\cite{Lamport05GeneralizeConsensus}. Traditional SMR implementations use a
+separate instance of a consensus algorithm to agree on the command to execute
+at each index of the command sequence on which the replicas must agree. Using
+separate consensus instances makes it hard to optimize the execution of
+commuting command because commutativity is not a property of a single command.
+Instead of restricting agreement to a single value, the Generalized Consensus
+problem requires that set of replica servers agree on the state of a growing
+data structure called a command-structure set. A command-structure set is
+a data structure with an append operation such that the order in which two
+commuting commands are appended does not change the resulting state. Thanks to
+this property, two replicas executing commuting commands in a different order
+will not end up in a different state and can continue executing without further
+synchronization. In message-passing systems, this property can be leveraged to
+reduce the latency of processing one command to two message delays instead of
+three. As we will show, it also enables the processing of read-only commands
+in two message delays, although read-only commands do not commute with write
+commands.
 
-In practice, a SMR algorithm must support fast reconfiguration to replace faulty components, without which the long-term stability of a system is inevitably compromised. 
-Our framework abstracts over how the reconfiguration decisions are made, providing an interface allowing to plug an external reconfiguration master implementing the reconfiguration policy of the system.
-The master decides which GC module to run next and which replica servers will participate, removing failed replicas and adding new ones.
-As we will explain, this enables tolerating $k$ failures with $k+1$ replicas, an important improvement over the traditional $2k+1$ needed replicas.
+In practice, a SMR algorithm must support fast reconfiguration to replace faulty
+components, without which the long-term stability of a system is inevitably
+compromised. Our framework abstracts over how the reconfiguration decisions
+are made, providing an interface allowing to plug an external reconfiguration
+master implementing the reconfiguration policy of the system. The master decides
+which GC module to run next and which replica servers will participate, removing
+failed replicas and adding new ones. As we will explain, this enables tolerating
+$k$ failures with $k+1$ replicas, an important improvement over the traditional
+$2k+1$ needed replicas.
 
-To demonstrate the practical applicability of our framework we provide composable versions of the generalized variations of popular consensus algorithms: classic Paxos, Fast Paxos, Chain Replication, and multi-coordinated Paxos.
-We have checked the correctness of these algorithms with the TLC model-checker and we provide all the specifications needed to replicate these experiments.
+To demonstrate the practical applicability of our framework we provide
+composable versions of the generalized variations of popular consensus
+algorithms: classic Paxos, Fast Paxos, Chain Replication, and multi-coordinated
+Paxos. We have checked the correctness of these algorithms with the TLC
+model-checker and we provide all the specifications needed to replicate these
+experiments. Our framework is also fully formalized and its composition theorem
+proved in Isabelle/HOL, a foundational interactive theorem prover, thereby
+certifying the correctness of the composition theorem at the deepest level.
 
-Our framework is also fully formalized and its composition theorem proved in Isabelle/HOL, a foundational interactive theorem prover, thereby certifying the correctness of the composition theorem at the deepest level.
+At the heart of our framework lies an abstraction, called Composable Generalized
+Consensus, of the familiar notion of round (also called epoch, ballot, or
+view in the literature) used in most SMR algorithms. As an example, let us
+consider Fast Paxos, one of the first SMR algorithm which can change strategy
+dynamically. Fast Paxos is a round-based SMR algorithm that has two types of
+rounds: fast rounds and normal rounds. Fast rounds process requests faster
+than normal rounds in contention-free executions but fall back to a slower
+scheme under contention. Normal rounds are the same as in the classic Paxos
+and are slower than fast rounds in contention-free executions. Any process can
+initiate a round change at any point during execution: this allows switching
+between fast and normal rounds as well as changing the leader from one normal
+round to the next, similarly to classic Paxos. Each kind of round maintains
+different invariants relating the state of each process and the values that
+might have been chosen in the previous rounds, thus a given round must know
+which past rounds are fast or normal rounds in order to take over the execution
+safely. When adding more kinds of rounds such an approach will quickly become
+unmanageable, as each type of rounds must have one case for each possible other
+kind of round, making $O(n^2)$ cases to consider in total.
 
-At the heart of our framework lies an abstraction, called Composable Generalized Consensus, of the familiar notion of round (also called epoch, ballot, or view in the literature) used in most SMR algorithms.
-As an example, let us consider Fast Paxos, one of the first SMR algorithm which can change strategy dynamically.
-Fast Paxos is a round-based SMR algorithm that has two types of rounds: fast rounds and normal rounds.
-Fast rounds process requests faster than normal rounds in contention-free executions but fall back to a slower scheme under contention.
-Normal rounds are the same as in the classic Paxos and are slower than fast rounds in contention-free executions.
-Any process can initiate a round change at any point during execution: this allows switching between fast and normal rounds as well as changing the
-leader from one normal round to the next, similarly to classic Paxos.
-Each kind of round maintains different invariants relating the state of each process and the values that might have been chosen in the previous rounds, thus a given round must know which past rounds are fast or normal rounds in order to take over the execution safely.
-When adding more kinds of rounds such an approach will quickly become unmanageable, as each type of rounds must have one case for each possible other kind of
-round, making $O(n^2)$ cases to consider in total.
-
-Our notion of Composable Generalized Consensus defines a common interface and a common correctness property for a round, abstracting over the particular round implementations and making all types of rounds compatible with each other.
-Moreover, our interface for rounds enables a crucial optimization seen in practical systems: introducing an external reconfiguration master which manages the addition of new replicas and the removal of failed ones.
-Interestingly, as explained in \cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}, in a real system it is often possible to implement a reliable reconfiguration master which allows to tolerate $f$ faults with $f+1$ replicas instead of the traditional $2f+1$ replicas.
-The reconfiguration master can also make the choice of the new type of round to execute, providing the system's reconfiguration policy.
+Our notion of Composable Generalized Consensus defines a common interface and a
+common correctness property for a round, abstracting over the particular round
+implementations and making all types of rounds compatible with each other.
+Moreover, our interface for rounds enables a crucial optimization seen in
+practical systems: introducing an external reconfiguration master which manages
+the addition of new replicas and the removal of failed ones. Interestingly, as
+explained in \cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication},
+in a real system it is often possible to implement a reliable reconfiguration
+master which allows to tolerate $f$ faults with $f+1$ replicas instead of the
+traditional $2f+1$ replicas. The reconfiguration master can also make the choice
+of the new type of round to execute, providing the system's reconfiguration
+policy.

### 2015-05-14 paper/paper.tex
Work on the introduction.
Work on the intro, added /web/

@@
 
-\addbibresource{./managed.bib}
+\addbibresource{managed.bib}
 
@@
   Giuliano Losa\\
-  \texttt{giuliano.losa@epfl.ch}
+  \texttt{giuliano@losa.fr}
 }
@@
 \vspace{2cm}
-
 \begin{center}
-This paper is a regular submission.
-
-\vspace{1cm}
-This paper is a student paper.
+This is a regular paper.
 \end{center}
@@
 \input{intro}
+
 \input{roadmap}
@@
 \end{document}
-    We present a new abstraction, Abortable Generalized Consensus, which enables the
-    correct-by-construction composition of heterogeneous State-Machine
-    Replication algorithms.
-
-    integrating state of the art techniques to optimize executions.  ACA is the most complete approach to date enabling the
-    correct-by-construction composition of heterogeneous State-Machine
-    Replication algorithms. We provide two specifications satisfying
-    ACA and we illustrate their power by refining them to
-    obtain a versatile adaptive algorithm that combines five existing
-    algorithms into a single one. ACA is completely formalized and its
-    underlying composition theorem has been mechanically checked by an interactive
-    proof assistant.
-}
-
-\vspace{2cm}
-
-\begin{center}
-This paper is a regular submission.
-
-\vspace{1cm}
-This paper is a student paper.
-\end{center}
-
-\newpage
-
-\section{Introduction}
-
-The State-Machine Replication technique (abbreviated SMR) allows
-building reliable services on top of unreliable hardware. The basic idea is to 
-replicate a service over several servers, each executing the same
-sequence of (deterministic) commands. 
-Making SMR algorithms efficient has been a long-standing
-challenge and a multitude of algorithms have been proposed. Examples
-include Fast Paxos \cite{Lamport06FastPaxos}, Disk Paxos
-\cite{GafniLamport03DiskPaxos}, Chain Replication
-\cite{RenesseSchneider04ChainReplicationSupportingHighThroughputAvailability},
-Ring Paxos
-\cite{MarandiETAL10RingPaxosHighthroughputAtomicBroadcastProtocol},
-Egalitarian Paxos
-\cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments},
-Multi-Coordinated Paxos
-\cite{CamargosSchmidtPedone07MulticoordinatedPaxos}, Cheap Paxos
-\cite{LamportMassa04CheapPaxos}, Vertical Paxos
-\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication},
-Paxos-MIC
-\cite{HurfinMoiseNarzul11AdaptiveFastPaxosMakingQuickEverlasting},
-Mencius
-\cite{MaoJunqueiraMarzullo08MenciusBuildingEfficientReplicatedStateMachine},
-and Fast Mencius \cite{WeiETAL13FastMenciusMenciusLowCommitLatency}.
-
-Each of the existing algorithms is best suited to a particular range
-of environments, characterized, e.g., by client load, network
-topology, likelihood of faults, communication latency and bandwidth, processing
-power of nodes, availability of disk storage and memory.
-However, the environment faced by real systems evolves over time:
-hardware is upgraded, new data-centers are built, the popularity of a
-service changes, etc.  Consistent performance therefore comes with
-adaptation to change.  Unfortunately, existing SMR algorithms cannot easily be
-stopped and replaced by another algorithm on the fly. Moreover,
-devising a new algorithm that can change behavior dynamically is hard:
-dynamically switching from one SMR algorithm to another requires
-synchronizing notoriously intricate algorithms.
-
-We propose a new abstraction, \emph{Abortable Chain Agreement}
-(abbreviated ACA), which subsumes SMR and whose implementations,
-called \emph{ACA rounds}, can, by construction, pass the baton from
-one to another on the fly and efficiently. Moreover, optimizing a replicated
-service consisting of several ACA rounds becomes just a matter of devising a new
-ACA round, which can then be plugged-in on the fly: any new ACA round
-is compatible with every existing ACA round by construction.
-
-Abortable Chain Agreement is the \emph{most complete framework to date} enabling
-correct by construction composition of heterogeneous rounds.  
-ACA is completely formalized and its underlying composition theorem has been
-mechanically checked by an interactive proof assistant.  
-Compared to ACA,
-existing approaches
-\cite{BarNoyETAL87ShiftingGearsChangingAlgorithmsFlyTo,GuerraouiETAL10Next700BftProtocols,GuerraouiKuncakLosa12SpeculativeLinearizability}
-are limited. In particular, they do not allow the disorderly execution of
-commuting and read-only commands, an important practical optimization
-which is at the root of recent work on SMR like Generalized Consensus
-\cite{Lamport05GeneralizeConsensus} and Generalized Lattice Agreement
-\cite{FalerioETAL12GeneralizedLatticeAgreement}.
-
-\begin{comment}
-On top of the traditional roles of proposer, acceptor, and learners
-\cite{lamport2001paxos} used in consensus algorithms, ACA introduces
-the notion of \emph{switchers} and of \emph{scheduling policy}. The switchers
-have the task of aborting a running ACA round and initializing a new
-ACA round in a safe and live manner.  The \emph{scheduling policy}
-decides when to abort an ACA round and choses the next ACA round to
-execute. It must ensure that all switchers agree on the next ACA round
-to initialize.
-\end{comment}
-
-We observe that, because of the FLP impossibility result
-\cite{FischerLynchPaterson83ImpossibilityDistributedConsensusOneFaultyProcess},
-most existing SMR algorithms are structured in a sequence of abortable
-rounds which can easily be transformed into ACA rounds. To simplify
-this task even further, we propose two specifications, \emph{conservative
-ACA rounds} and \emph{fast ACA rounds}, which, as we show, can be concretized
-to obtain ACA versions of the rounds of five existing algorithms: 
-Classic Paxos, Fast Paxos, Chain Replication, Ring Paxos,
-and Multi-Coordinated Paxos. ACA guarantees that the composition of the five ACA
-rounds is correct by construction, forming a versatile adaptive replication
-algorithm.
-
-The rest of the paper is structured as follows.
-\Cref{sec:related} surveys recent related work. \Cref{sec:ca} presents Chain
-Agreement, which embodies the requirements for implementing SMR. \Cref{sec:aca}
-presents Abortable Chain Agreement and its optimizations to treat commuting
-and read-only commands efficiently. \Cref{sec:rounds} presents our specification
-of conservative and fast ACA rounds, which implement ACA. In
-\cref{sec:examples}, we show that the rounds of  
-Classic Paxos, Fast Paxos, Chain Replication, Ring Paxos,
-and Multi-Coordinated Paxos can easily be transformed into ACA implementations
-by using our specifications of conservative and fast ACA rounds.
-Finally, in \cref{sec:isaproofs}, we provide a complete formalization of ACA
-and a mechanical proof of the \emph{composition theorem} (\cref{thm:comp}),
-checked by the interactive proof assistant Isabelle/HOL \cite{NipkowPaulsonWenzel02IsabelleHOL}.
-
-\begin{comment}
-Abortable Chain Agreement builds on several recent advances in SMR
-algorithms: ACA allows acceptors to only eventually reach agreement,
-like in Generalized Paxos \cite{Lamport05GeneralizeConsensus},
-boosting performance when most commands commute; ACA allows fast
-execution of read commands like in Generalized Lattice Agreement
-\cite{FalerioETAL12GeneralizedLatticeAgreement}; ACA implementations
-can tolerate $f$ faults with only $f+1$ replicas and replace crashed replicas
-upon changing round, as in Vertical Paxos  \cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}.  
-\end{comment}
-
-\section{Related Work}
-\label{sec:related}
-
-SMR is often implemented using a sequence of instances of Consensus.
-There are numerous examples of consensus algorithms in the literature.
-However, Abortable Chain Agreement is based on Generalized Consensus. 
-In Generalized Consensus \cite{Lamport05GeneralizeConsensus}, the learners of a
-SMR algorithm repeatedly learn new values belonging to a \emph{C-struct set}. A
-C-struct can be seen as a data structure encoding  sequences of commands modulo
-the ordering of commands that commute. The advantage of Generalized
-Consensus over Consensus is that
-two servers which receive commuting commands in different orders can still reach
-agreement on C-structs without further synchronization. 
-Our Abortable Chain Agreement can also use C-structs, with the same advantages
-as Generalized Consensus. For simplicity, the presentation in this
-paper uses the more concrete notion of \emph{command history} (see
-\cref{sec:histories}).
-
-Generalized Lattice Agreement \cite{FalerioETAL12GeneralizedLatticeAgreement} allows more efficient implementations of SMR when
-commands are partitioned into read-only commands and write-only commands.
-Abortable Chain Agreement supports the same optimizations as Generalized Lattice Agreement.
-Notably, read-only commands are treated specially so as to optimize their
-executions (see \cref{sec:readonly}).
-
-Vertical Paxos algorithms 
-\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication} are SMR algorithms
-whose set of replica servers can be reconfigured without stopping to process new
-commands. Crashed servers can therefore be replaced by new servers, ensuring the
-long-term availability of a replicated service.
-Vertical Paxos algorithms use an external reconfiguration master, itself
-implemented as a replicated state machine, which determines the servers that should
-be removed from or added to the configuration.
-ACA shares some aspects of Vertical Paxos algorithms: the task of the
-reconfiguration master is similar to the task of the scheduling policy
-component of ACA (see \cref{sec:aca}).
-However, in ACA, changing round can be used not only to replace crashed servers
-but also to change the algorithm that the servers are running. 
-Changing algorithm is not possible in Vertical Paxos.
-In both Vertical Paxos and ACA, a reliable reconfiguration master, resp. a
-reliable scheduling policy, allows tolerating $f$ crashes with $f+1$ servers.
-
-The Abstract Framework \cite{GuerraouiETAL10Next700BftProtocols} proposes to
-build Byzantine Fault-Tolerant algorithms as the succession of abortable rounds
-called Abstract instances. Abstract instances must satisfy the Abstract
-correctness properties. By construction, the composition of any number of
-Abstract instances is a correct SMR implementation. However, the Abstract
-Framework uses totally ordered sequence of commands, making difficult the
-optimization of the execution of commuting and read-only commands that are
-possible with ACA.  Moreover, the Abstract Framework does not identify the
-crucial role of the scheduling policy for the long-term resilience of a service.
-Abortable Chain Agreement combines the ideas of the Abstract Framework,
-Generalized Paxos, Generalized Lattice Agreement, and Vertical Paxos in a
-unified abstraction.
-
-Finally, the Speculative Linearizability framework
-\cite{GuerraouiKuncakLosa12SpeculativeLinearizability} was a first
-attempt by the authors of the present paper at generalizing the
-Abstract Framework and mechanically proving its properties with the
-Isabelle/HOL interactive
-proofs assistant.  Like the Abstract Framework, the
-Speculative Linearizability framework is based on totally ordered
-sequences of commands, with the same drawback for the execution of
-commuting and read-only commands. Moreover, Abortable Chain Agreement is both
-conceptually simpler and subsumes the Speculative Linearizability
-property.  Like Speculative Linearizability, Abortable Chain Agreement
-and its main property, the composition theorem (\cref{thm:comp}), are
-formalized in Isabelle/HOL (see \cref{sec:isaproofs}).
-
-Abortable Chain Agreement is the most complete framework to date for building
-adaptive replication algorithms out of heterogeneous rounds, making state of the
-art optimization techniques available to algorithm designers, which was not
-possible before.
-
-\section{Chain Agreement}
-\label{sec:ca}
-
-We consider a set of asynchronous servers communicating by message passing in an
-asynchronous network. Servers are not necessarily sequential processes. We
-describe a behavior of the system using an infinite sequence of states, where a
-state is a function from variable names to values (i.e. a valuation
-of variables).  We specify sets of allowed behaviors using an initial predicate,
-a next-state relation (using unprimed and primed variables) and a fairness
-condition, all in the style of TLA+ \cite{Lamport02SpecifyingSystems}. However,
-for succinctness, our specification are rather informal, eluding non-essential
-details.
-
-Our specifications all contain by default a state variable $crashed$ and
-$Crash\left( x \right)$ transition which is always enabled. A server $x$ faithfully
-executes the algorithm that is assigned to it until its identifier is added to
-the variable $crashed$ by a $Crash\left( x \right)$ transitions.  When the identifier of a
-server is in the set $crashed$, all its actions are disabled.  In an execution,
-a server that is never added to the set $crashed$ is said correct.
-
-We consider a computing service exposing a set of commands $C$. The
-service has a sequential specification consisting of an initial state
-and a deterministic transition relation in which every command
-atomically updates the state of the service and produces an output. We
-assume that duplicate commands have no effect on the state of the
-service and return the same output as their first occurrence.
-
-In a traditional SMR algorithm, the servers can be seen has playing one or
-several of three roles: \emph{proposers}, which propose new commands to execute,
-\emph{acceptors}, which accept and order commands in a sequence, and
-\emph{learners}, which learn about the growing sequence of commands to
-execute \cite{lamport2001paxos}. Often, but not necessarily, every replica server plays the
-three roles of proposer, acceptor, and learner: clients of the service
-send their commands to some replica servers; replica servers submit
-the received commands to the SMR protocol in the role of proposers;
-replica servers respond to commands by executing the commands learned
-in the role of learners; replica servers participate in command
-ordering in the role of acceptors.  An SMR algorithm guarantees that
-the replicas learn the same sequences of commands, possibly up to the
-reordering of commuting commands. Therefore, every learner can respond to client
-requests using its last learned sequence of commands, and clients will receive
-the same responses as the ones produces by a centralized, non-replicated,
-service.
-
-Chain Agreement precisely specifies the allowed behaviors of the roles of \emph{proposers}
-and \emph{learners}, abstracting over the behavior of acceptors.
-
-\begin{comment}
-We suppose that each server can play a number of \emph{roles}.
-Chain Agreement specifies the allowed behaviors the roles of \emph{proposers}
-and \emph{learners}. Proposers receive commands from the clients of the service
-and propose them for execution by the replicated service. Each learner learns about
-a growing sequence of commands. Thanks to the sequential specification of the
-service, a sequence of commands determines an output for each the included commands. 
-In a traditional SMR algorithm, servers can also play the role of
-\emph{acceptors}, whose task is to collectively order commands and let the
-learners know about the latest sequence of ordered commands.
-
-The point of Chain Agreement is to implement a replicated service by having every learner learn about a growing sequence of commands, which
-represent an execution of the sequential specification of the service.
-Crucially, at all times, the sequences of commands learned by the
-different learners must form a chain in the prefix order on sequences.
-Moreover, for every command $c$ proposed by a correct proposer, every
-correct learner must eventually learn a sequence containing $c$.  Once
-a learner learns about a sequence containing a command $c$, it can
-determine the output produced by $c$, using the sequential
-specification of the service, and respond to the client that issued
-the command $c$.
-\end{comment}
-
-To specify Chain Agreement, we need the notion of ordering of sequences and of
-maximum sequence.  The prefix order on sequences is denoted by the operator
-$\leq$ ($s_1 \leq s_2$ means that $s_1$ is a prefix of $s_2$). 
-%Using the operator $\leq$, we define $\geq$, $<$, and $>$ as expected. 
-Note that $\leq$ is a partial order on sequences. The maximum over a set of
-sequences $S$ is denoted by $Max\left( S \right)$ and $Max\left( \aset{}
-\right)$ is the empty sequence.
-
-We specify Chain Agreement using the state variables $crashed$, $proposed$, containing the proposed commands,
-and, for every learner $l$,  $learned\left[ l \right]$, a set of
-sequences of commands. Initially, $crashed$, $proposed$ and, for every learner $l$,
-$learned\left[ l \right]$ are empty sets. 
-Remember that any transition involving a server $x$ is always disabled when $x\in crashed$.
-
-The next-state relation consists of three types of transitions:
-$Crash\left( x \right)$ transitions; $Propose\left( l, c
-\right)$ transitions, for a learner $l$ and a command $c$, in which
-$c$ is inserted in the set $proposed$; $Learn\left( l,cs \right)$
-transitions, for a learner $l$ and a sequence of commands $cs$, in
-which $cs$ is inserted in $learned\left[ l \right]$. 
-A $Learn\left( l, cs \right)$ transition is guarded by the following
-conditions:
-\begin{compactitem}
-    \item[\textbf{Agreement}:] The elements of the set $\aset{learned'\left[ l
-      \right] : l \in Learners}$, the learned sequences after the transition,
-      form a chain in the partial order $\leq$.
-    \item[\textbf{Validity}:] Every learned sequence contains only
-        proposed commands.  
-    \item[\textbf{Irrevocability}:] The new learned sequence
-        $cs$ is a strict extension of $Max\left(learned\left[ l
-        \right]\right)$. 
-        %\\ $cs > Max\left(learned\left[ l \right]\right)$.
-\end{compactitem}
-Finally, we add, for every learner $l$, one weak fairness condition
-\cite{Lamport02SpecifyingSystems} on the transitions $Learn\left( l,
-cs \right)$. The weak fairness conditions ensure that every correct
-learner eventually learns about every proposed command.
-
-Thanks to the properties of Chain Agreement, the clients, observing
-their input commands and the received responses, see a
-\emph{linearizable} execution
-\cite{HerlihyWing90LinearizabilityCorrectnessConditionConcurrentObjects}.
-Moreover, the service remains available as long as at least one
-proposer and one learner are correct.
-
-\section{Abortable Chain Agreement}
-\label{sec:aca}
-
-We now consider implementing Chain Agreement as a sequence of \emph{Abortable Chain
-Agreement rounds}, where each round makes progress ordering requests or aborts
-its execution and passes the baton to the next round in the sequence.
-An algorithm composed of a sequence of ACA rounds is called an \emph{ACA
-algorithm}.
-
-On top of the traditional roles of proposer, acceptor, and learner, ACA
-introduces the new role of \emph{switcher} and a new component, the
-\emph{scheduling policy}. In ACA, the assignment of the roles of proposer,
-acceptor, learner, and switcher can change in each rounds, allowing for
-reconfiguration. 
-In normal circumstances, the proposers, acceptors,
-and learners of an ACA round work together to implement Chain Agreement.
-However, if the environment becomes unfavorable to a running ACA round, the
-scheduling policy can instruct the round to abort its execution and pass the
-baton, using the switchers, to a new round.
-
-A switcher of an aborting round determines, from the state of the
-acceptors, an \emph{abort sequence} that it transfers to one or several
-switchers of the new round.  Upon receiving an abort sequence, a switcher of the
-new round initializes the acceptors of the new round, which can then
-proceed implementing SMR, until the new round itself aborts and the
-process is repeated. 
-We assume that in the first round, which has no previous round, every 
-switchers receives the empty abort sequence.
-
-The scheduling policy determines when to abort an ACA round and which new ACA
-round should take over. The scheduling policy must inform the switchers about the
-identity of the next round and must guarantee that all switchers receive the
-same information.  In this paper we abstract over the scheduling policy, assuming
-that a round that aborts passes the baton a nondeterministically chosen new
-round. A reliable scheduling-policy component can realistically be implemented
-as a replicated state machine running on a large number of servers otherwise
-dedicated to other tasks
-\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}. We will see
- in \cref{sec:rounds} that, as in Vertical Paxos, a reliable scheduling policy component allows
-tolerating $f$ faults with $f+1$ replicas, while maintaining the long-term
-resilience of the service by replacing faulty servers upon round change.
-
-The clients of an ACA algorithm can send their commands to the proposers of any round (in
-practice, clients send their commands to proposers of the largest
-active round) and servers can use sequences learned in any round
-to determine outputs.
-
-ACA specifies the allowed behaviors of the proposers, learners, and switchers
-\emph{of a single round}. Similarly to
-Chain Agreement, we use the state variables $crashed$, $proposed$,
-and, for every learner $l$, $learned\left[ l \right]$.  Additionally,
-we use the state variables $inits$, containing the abort sequences
-received from the previous round, and $aborts$, containing the abort
-sequences transfered to the next round.
-Two consecutive ACA rounds are composed by letting the $aborts$ variable of the
-first round be the $inits$ variable of the second round.
-
-Initially, $crashed$, $proposed$, for every
-learner $l$, $learned\left[ l \right]$, $inits$, and $aborts$ are all empty sets.
-The next-state relation consists of five types of transitions. The
-first three, $Crash\left( x \right)$, $Propose\left( l, c
-\right)$, and $Learn\left( l, cs \right)$, for a learner, proposer, or
-switcher $x$, for a learner $l$, and for a sequence of commands
-$cs$, have the same effect as in Chain Agreement.
-ACA introduces two new transitions, $Init\left(
-s, cs \right)$ and $Abort\left( s, cs \right)$, for a switcher $s$ and a sequence of
-commands $cs$. The transition $Init\left( s, cs \right)$ adds $cs$ to the set $inits$, and models a switcher receiving the
-abort sequence $cs$ from the preceding round.
-The transition $Abort\left( s, cs \right)$ adds $cs$ to the set $aborts$ and models a switcher transferring the abort sequence $cs$ to a switcher of the next round.
-All the transition are guarded by the following conditions:
-\begin{compactitem}
-    \item[\textbf{Agreement}:] The elements of the set $\aset{learned'\left[ l
-      \right] : l \in Learners}$ form a chain in the partial order.
-    \item[\textbf{Validity}:] Every abort sequence or learned sequence
-        is of the form $h \star cs$, where $cs$ contains only commands
-        of the set $proposed$ and $h$ is the greatest lower bound of a
-        nonempty subset of $inits$.
-    \item[\textbf{Initialization}:] If $inits$ is empty, then no
-        sequence can be learned and no switcher can abort.
-    \item[\textbf{Irrevocability}:] In the case of a $Learn\left( l,cs
-        \right)$ transition, the new learned sequence $cs$ is a strict
-        extension of $Max\left(learned\left[ l \right]\right)$.
-        %, $cs
-        %\geq Max\left(learned\left[ l \right]\right)$.
-    \item[\textbf{Safe Abort}:] Every abort sequence is larger in the
-        partial order $\leq$ than every learned sequence.
-\end{compactitem}
-Note that Agreement and Irrevocability are the same as in Chain
-Agreement.
-
-Finally, for every learner $l$, we add a weak fairness condition on the
-following state transition: 
-\begin{equation}
-    \exists  s \in Switchers, cs \in Sequence\left( C \right) : Learn\left( l, cs \right) \vee Abort\left( s,cs \right)
-\end{equation}
-The weak fairness conditions ensure that when a correct learner has not
-learned a sequence containing all commands proposed in the round, then
-it eventually learns a new command or, if there is at least one correct switcher, at least one switch value is transfered to the next round. 
-Therefore, if a client has submitted a command to a proposer, eventually,
-either the client gets a response from the current round, or
-the next round is initialized with an abort sequence and the
-command can be resubmitted to the next round.
-
-\subsection{The Composition Theorem}
-
-Consider the succession of $n$ rounds numbered $1$ to $n$.  To
-distinguish the variables of each round, we annotate the local
-variables of each round with the number of the round. For example, the
-variables of the seventh round are written $proposed_7$, for every
-learner $l$, $learned_7\left[ l \right]$, $inits_7$, and $aborts_7$.
-
-Moreover, we assume that a command, which is proposed to a round and that
-is not learned before the first abort sequence is transfered to the
-next round, is proposed anew to the next round.
-
-Define the state variable $proposed$ as the union over all rounds $i$
-of the $proposed_i$ variables and, for every learner $l$, 
-define the $learned\left[ l \right]$ variable as the union over all
-rounds $i$ of the $learned_i\left[ l \right]$ variables.
-
-\begin{thm}[Composition Theorem]
-    \label{thm:comp}
-    In every execution of a succession of ACA rounds, the sequence of valuations
-    of the variable $proposed$ and, for every learner $l$, of the variables
-    $learned\left[ l \right]$ satisfies Chain Agreement.
-\end{thm}
-\Cref{thm:comp} can be shown by induction on the number of rounds,
-using the following lemma.
-\begin{lem}
-  \label{lem:comp}
-    Consider two rounds numbered $i$ and $i+1$. In every execution of a
-    succession of ACA rounds, the sequence of valuations of the variables $inits
-    = inits_i$, $propose = proposed_i \cup propose_{i+1}$, for every learner
-    $l$, $learned\left[ l \right] = learned_i\left[ l \right] \cup
-    learned_{i+1}\left[ l \right]$and $aborts = aborts_{i+1}$ satisfies ACA.
-\end{lem}
-\begin{proof}
-  A mechanically-checked proof in Isabelle/HOL appears in \cref{sec:isaproofs}.
-  The proofs uses a version of the specification of ACA that has a more
-  executable form that presented in this section. Moreover,
-  instead of sequences of commands, the Isabelle/HOL specification uses the more
-  general notion of command history, presented in the next subsection.
-  The Isabelle/HOL formalization of ACA uses the theory of I/O-automata, which
-  is also formalized in the Isabelle/HOL development. The proof uses a refinement
-  mapping from the composition of two consecutive ACA instances to a single ACA
-  instance.
-\end{proof}
-
-\Cref{thm:comp} guarantees that the successive composition of any
-number of ACA rounds implements Chain Agreement.
-Therefore, with ACA, one can easily build replicated services that
-dynamically adapt to the current operating conditions of the system: a
-replicated service has several types of rounds at its disposal, each being
-tailored to particular conditions. Each time the conditions change,
-the current ACA round aborts and passes the baton to a more 
-appropriate type of round, as determined by the scheduling policy. Moreover, a new type
-of ACA round can always be devised after the system is deployed and added on the fly to
-the existing types of rounds.
-
-We now describe two important practical optimizations of ACA.
-
-\subsection{Histories}
-\label{sec:histories}
-
-ACA can be implemented more efficiently by substituting \emph{command histories} for sequences of commands, as in Generalized Consensus \cite{Lamport05GeneralizeConsensus}.
-
-We say that two commands $c_1$ and $c_2$ \emph{commute} when, in every sequential
-execution of the service, reversing the order of $c_1$ and $c_2$ does
-not change any of the outputs, including those given to $c_1$ and
-$c_2$.
-
-A command history, abbreviated history, is a data structure that
-represents a set of sequences which are the same modulo the reordering
-of commuting commands, and in which duplicate commands do not appear.
-All the sequences represented by a given history contain the same
-commands and determine the same outputs for each command.  Therefore,
-to implement a replicated service, it is sufficient that learners
-learn command histories instead of sequences.  For a precise
-definition of command histories, we refer the reader to
-\cite{Lamport05GeneralizeConsensus}. Our Isabelle/HOL formalization
-(\cref{sec:isaproofs}) abstracts
-over the concrete representation of histories, assuming only their necessary
-properties. Command histories were first
-introduced in trace theory \cite{Mazurkiewicz84Semantics}.
-
-A command history can be built by appending a sequence of commands
-$cs$ to a history $h$, obtain the history $h \star cs$.  The major
-advantage that histories have over sequences is that if a sequence
-$cs_1$ is obtained by reordering the commuting commands of $cs_2$,
-then $h \star cs_1 = h \star cs_2$. Therefore, in  implementations of
-ACA, servers can agree on a history even though
-they chose different orders for some commuting commands.
-
-Command histories share the four properties of sequences that are
-necessary and sufficient to substitute histories for sequences in the
-definition of Chain Agreement and ACA.
-\begin{compactitem}
-    \item[\textbf{Distinctness}:] No duplicate command appears in a history. 
-    \item[\textbf{Ordering}:] Command histories form a partial order:
-        $h_1 \leq h_2$ when there exists a sequence $s$ such that $h_2
-        = h_1 \star s$.
-    \item[\textbf{GLB}:] Every finite set of command histories has a
-        \emph{greatest lower bound}.
-%    \item[\textbf{LUB}:] Every finite set of compatible command
-%        histories has a \emph{least upper bound}, where a set of
-%        command histories is compatible when they have a common upper
-%        bound.
-    \item[\textbf{Consistency}:] If $h_1 \leq h_2 \leq h_3$ and $h_3 =
-        h_1 \star s$ then there exists $s'$ and $s''$ where every
-        request of $s'$ and $s''$ appears in $s$ and such that $h_2 =
-        h_1 \star s'$ and $h_3 = h_2 \star s''$.
-\end{compactitem}
-The optimized version of ACA is obtained by
-simply substituting histories for sequences in the definition of ACA.
-The Ordering, GLB, and Consistency properties of histories ensure that
-\cref{thm:comp} still holds: the Isabelle/HOL proof of \cref{thm:comp} (see
-\cref{sec:isaproofs}) relies only on the four properties above.
-
-\subsection{Optimizing the Execution of Read-Only Commands}
-\label{sec:readonly}
-
-Chain Agreement can be further optimized by treating 
-read-only commands specially, as in Generalized Lattice Agreement
-\cite{FalerioETAL12GeneralizedLatticeAgreement}.
-
-We consider the case in which a set of servers each play the four
-roles of proposer, acceptor, learner, and switcher.
-Clients sends their commands to any server, which proposes them as a
-proposer, and records the command as pending.
-%When a server learns a history that contains one of its pending
-%commands, it uses the history to determine the response to the pending
-%command and forwards the response to the appropriate client.
-
-As observed in \cite{FalerioETAL12GeneralizedLatticeAgreement}, read-only
-commands can be executed faster by directly reading from the last learned
-history, instead of proposing the read commands and waiting for it to appear in
-a future learned history.  Because learned histories form a chain, all the reads
-return consistent values and this approach yields a \emph{serializable}
-implementation of the service.
-
-However linearizability is not satisfied if reads are served directly
-from the last learned history: a lagging replica server could respond
-to a read using a history which does not contain all completed update
-commands. To make the executions linearizable without resorting to
-proposing read-only commands as normal commands, we can employ the same
-technique as in \cite{FalerioETAL12GeneralizedLatticeAgreement}: upon
-receiving a read command, a server proposes a special no-op command
-that commutes with all other commands; the server then responds to the
-read command using the first learned history containing the no-op
-command. This ensures that all the commands that had completed before
-the read command was issued are present in the history.  Moreover,
-since no-op commands commute with all other commands, they can be
-treated more efficiently than regular read commands by the acceptors,
-like, for example, in a fast round of Fast Paxos.
-
-\section{Fast and Conservative Rounds}
-\label{sec:rounds}
-
-Most existing SMR algorithms can easily be obtained as the
-composition of one or more ACA rounds.
-
-The FLP impossibility result \cite{FischerLynchPaterson83ImpossibilityDistributedConsensusOneFaultyProcess}
-implies that no algorithm can implement SMR to the letter in practice.
-Instead, SMR algorithms guarantee safety (i.e. replica execute the
-same sequence of commands) but may not be able to make progress (i.e.
-order new commands). To make progress as likely as possible, most SMR
-algorithms are structured in consecutive rounds (also called ballots,
-views, or epochs). In each round, the algorithm tries to make progress
-but cannot be guaranteed to succeed. When the algorithm detects that
-progress is no more possible, a new round is started.  The new round
-is initialized so as to preserve safety, usually by one of the
-acceptors. For example, in Paxos, the leader of a newly started ballot
-numbered $i$ determines a value that is \emph{safe at $i$} and
-proposes it as first value to the other acceptors of the ballot $i$.
-In existing SMR algorithms, new rounds are initialized ad-hoc, which
-prevents mixing rounds of two different algorithms.  Instead,
-ACA is an abstraction of rounds which establishes a common interface and contract for switching from
-one round to the next, enabling switching from any ACA round to any other.
-
-To simplify the task of expressing existing algorithms as the composition of one
-or more ACA rounds, we propose two specifications, \emph{conservative ACA
-rounds} and
-\emph{fast ACA rounds}, which can easily be concretized to obtain executable rounds.
-We show that ACA versions of Classic Paxos, Multi-Coordinated Paxos, Chain
-Replication, and Ring Paxos can be obtained by concretizing the specification of
-conservative ACA rounds, and that ACA versions of the fast rounds of Fast Paxos can be obtained by concretizing the specification of fast ACA rounds.
-
-Fast and conservative ACA rounds are two specifications, at an
-intermediate level of abstraction, which satisfy ACA. Both
-specifications specify how the state of the acceptors should be
-updated but abstract over the particular communication topology used
-to implement the state accesses: the specifications use a global state
-which is accessible by all servers in all roles. One can refine fast
-or conservative rounds by implementing the state accesses and coordination
-using the network, obtaining a concrete algorithm.
-
-Conservative and fast ACA rounds are specified using the state variables $inits$, $proposed$, for every learner $l$, $learned\left[ l \right]$
-and $aborts$, as in ACA, and, for every acceptor $a$, two new variables
-$status\left[ a \right]$, which can be either ``idle'', ``ready'', or
-``stopped'' and $hist\left[ a \right]$, the local history of the
-acceptor $a$, which can be either a history or the special value $None$.
-Initially, $inits$, $aborts$, and, for every learner
-$l$, $learned\left[ l \right]$ are empty sets, and, for every acceptor $a$,
-$status\left[ a \right]$  is ``idle'' and $hist\left[ a \right]$ is the special
-value $None$.
-
-The next-state relations of conservative and fast rounds are composed
-of the following types of transitions (and the default $Crash\left( x
-\right)$ transition).
-\begin{compactitem}
-    \item A $Propose\left( p,c \right)$ transition, for a proposer $p$
-        and a command $c$, not guarded, inserts $c$ in the set
-        $proposed$. 
-    \item An $Init\left( s, h \right)$ transition, for a switcher $s$
-        and a history $h$, not guarded, inserts $h$ in the set $inits$.
-    \item A $WakeUp\left( a \right)$ transition, for an acceptor $a$,
-        is enabled when $status\left[ a \right]$ is ``idle''. Its
-        effect is to set $hist\left[ a \right]$ to a history found in
-        $inits$ and to set $status\left[ a \right]$ to ``ready''.
-    \item An $Accept\left( a \right)$ transition, for an acceptor $a$,
-        is enabled when $status\left[ a \right]$ is ``ready'', and
-        updates $hist\left[ a \right]$ to $hist\left[ a \right]\star
-        c$, where $c$ is a proposed command. In conservative rounds,
-        the transition is additionally guarded by the condition that
-        the resulting set of local acceptor histories,
-        $\aset{hist'\left[ a \right] : a \in Acceptors}$, forms a chain
-        in the partial order $\leq$ (this is typically implemented with a
-        leader).
-    \item A $Learn\left( l, h \right)$ transition, for a learner $l$
-      and a history $h$, enabled when $h > learned\left[ l \right]$ and is
-      smaller than the GLB of the histories of a learn quorum of acceptors. The effect of
-        the transition is to set $learned\left[ l \right]$ to $h$.
-    \item A $Stop\left( a \right)$ transition, for an acceptor $a$,
-        enabled when $status\left[ a \right]$ is ``ready'', sets
-        $status\left[ a \right]$ to ``stopped'' (note that
-        $status\left[ a \right] = $ ``stopped'' prevents any further
-        ``accept'' transition of acceptor $a$).
-%    \item REALLY NEEDED? A $TimeOut\left( a \right)$ transition, for an acceptor $a$,
-%        is enabled when $status\left[ a \right]$ is ``idle'' and sets 
-%        $status\left[ a \right]$ to ``stopped'' and $hist\left[ a
-%        \right]$ to the special value $None$. 
-    \item An $Abort\left( s, h \right)$ transition, for a switcher $s$
-        and a history $h$, is enabled when there exists a abort quorum
-        $R$ such that, for every member $a$ of $R$, $status\left[ a
-        \right]$ is ``stopped'' and $h \in SafeAborts\left( R
-        \right)$. Its effect is to insert  $h$ in the set $aborts$. 
-\end{compactitem}
-
-On top of the additional guard on the $Accept$ transition present only
-in conservative rounds, conservative and fast ACA rounds differ in
-their definition of learn quorums, abort quorums, and in their
-definition of the $SafeAborts\left( R \right)$ operator.
-To concisely define read and learn quorums, we define 
-\begin{equation}
-    Q\left( f \right) = \aset{as \subseteq Acceptors :
-        Card\left(as\right) \geq \left\lfloor f\times Card\left(
-    Acceptors \right)\right\rfloor + 1}.
-\end{equation}
-
-In conservative ACA rounds, abort and learn quorums are defined such
-that the intersection between two learn quorums is nonempty and the
-intersection between a learn quorum and an abort quorum is nonempty.
-The definition leads to two interesting cases: $LearnQ = AbortQ =
-Q\left( 1/2 \right)$; $LearnQ = \aset{Acceptors}$ and $AbortQ =
-\aset{\aset{a} : a \in Acceptors}$. The operator $SafeAborts\left( R
-\right)$ is defined as the singleton set containing the maximum
-element over the local histories of the acceptors in $R$ which are not
-$None$. Note that the maximum exists because conservative rounds
-guarantee that the histories of the acceptors form a chain.
-
-In fast ACA rounds, abort and learn quorums are defined such that
-the intersection between any two learn quorums is nonempty and the
-intersection between any two learn quorums and an abort quorum is
-nonempty.  The definition leads to three interesting cases: $LearnQ =
-AbortQ = Q\left( 2/3 \right)$; $LearnQ = Q\left( 1/2 \right)$ and
-$AbortQ = Q\left( 3/4 \right)$; $LearnQ = \aset{Acceptors}$ and
-$AbortQ = \aset{\aset{a : a \in Acceptors}}$.
-The operator $SafeAborts\left( R \right)$ is defined as the set of
-maximal elements over the set $\aset{GLB\left( \aset{hist\left[ a
-\right] : a \in R \cap L \wedge hist\left[ a \right]\neq None} \right)
-: L \in LearnQ}$.
-
-The cases in which abort quorums are singleton sets, in both
-conservative and fast rounds, allow tolerating
-$f$ faults with $f+1$ servers: a round can abort as long as one
-acceptor is correct and a new round, possibly using new correct
-servers, can take over. Note that changing round cannot happen
-without the scheduling policy instructing the switchers about the
-identity of the next round, and thus tolerating $f$ faults with
-$f+1$ servers relies on the assumption of a reliable scheduling policy.
-
-Define, for every learner $l$, the history variable $\overline{learned}\left[ l \right]$ which is initialized to the empty set and which is
-updated each time a new local history is accepted by the learner $l$ by
-inserting the new local history in $\overline{learned}\left[ l \right]$ (see \cite{AbadiLamport91ExistenceRefinementMappings} for the
-definition of a history variable).
-\begin{thm}
-  \label{thm:rounds}
-    Both conservative and fast ACA rounds implement ACA under the refinement
-    mapping that consists in substituting $\overline{learned\left[ l \right]}$
-    for $learned\left[ l \right]$.
-\end{thm}
-
-\section{Examples of Concrete ACA Algorithms}
-\label{sec:examples}
-
-In this section we show how to modify the rounds of Classic Paxos,
-Multi-Coordinated Paxos, Chain Replication, Ring Paxos, and Fast Paxos to obtain
-ACA rounds. \Cref{thm:comp} then guarantee that all those different types of
-rounds can be used in the same algorithm to yield an implementation of Chain
-Agreement.
-
-\subsection{Classic Paxos}
-
-The rounds of the Classic Paxos algorithm can be modified so as to
-obtain an implementation of conservative ACA rounds.  We define
-Classic Paxos ACA rounds by specializing the specification of
-conservative ACA rounds, giving a concrete protocol for disseminating
-the proposed commands to the acceptors and enforcing that the learned
-histories form a chain.
-
-Each Classic Paxos round has a unique leader among the acceptors which
-centrally orders the proposed commands, which are accepted by the
-other acceptors in the order determined by the leader.  This protocol
-ensures that every local acceptor history is a prefix of the local
-history of the leader, therefore guaranteeing that the local acceptor histories
-form a chain.
-
-Each acceptor broadcasts each newly accepted history to the learners, which
-learn a history $h$ when they receive enough extensions of $h$ from the
-acceptors.  When instructed by the scheduling policy to abort, a
-switcher sends a stop message to the acceptors, which irrevocably stop
-updating their local history and send back to the switcher an
-acknowledgement containing their local history. A switcher
-determines an abort history once it has received acknowledgments from
-an abort quorum of acceptors $R$, using the operator $SafeAborts\left(
-R \right)$.
-
-The specification of Classic Paxos ACA rounds uses the variables of
-the specification of conservative ACA rounds, namely the variables
-$proposed$, $inits$, $aborts$, the arrays $learned\left[ l \right]$,
-$status\left[ a \right]$, and $hist\left[ a \right]$, and two
-additional arrays $accepted\left[ l \right]\left[ a \right]$ and
-$stop\left[ s \right]\left[ a \right]$, where $l$ is a learner, $a$ is
-an acceptor, and $s$ is a switcher.
-
-There is also a network component that we do not describe explicitly.
-However, the network allows any servers to communicate by sending and
-receiving messages.
-
-Initially, the variables $proposed$, $inits$, $aborts$, and, for every
-learner $l$, $learned\left[ l \right]$ are empty sets. For every
-acceptor $a$, $status\left[ a \right]$ is ``idle'' and $hist\left[ a
-\right]$ is the special value $None$. For every learner $l$ and acceptor $a$,
-$accepted\left[ l \right]\left[ a \right]$ is the special value
-$None$. For every switcher $s$ and acceptor $a$, $stopped\left[ s
-\right]\left[ a \right]$ is set to $false$.  There are initially no
-messages in the network.
-
-The next-state relation is composed of the following types of
-transitions, where the definitions of learn quorums, abort quorums,
-and of the $SafeAborts\left( R \right)$ operator are the ones of conservative
-ACA rounds.
-\begin{compactitem}
-    \item A $Propose\left( p,c \right)$ transition, for a proposer $p$
-        and a command $c$, not guarded, inserts $c$ in the set
-        $proposed$ and sends the message $\aseq{\quo{prop},c}$ to
-        the leader.
-    \item An $Init\left( s, h \right)$ transition, for a switcher $s$
-        and a history $h$, not guarded, inserts $h$ in the set $inits$ and
-        sends the message $\aseq{\quo{init},h}$ to the leader.
-    \item A $WakeUp\left( leader \right)$ transition, enabled when
-        $status\left[ leader \right]$ is ``idle'' and the leader can
-        receive a $\aseq{\quo{init},h}$ message. Its effect is to
-        receive the message, set $hist\left[ leader \right]$ to $h$,
-        to set $status\left[ leader \right]$ to ``ready'', and to
-        broadcast the message $\aseq{\quo{leader-accept},h}$ to all
-        other acceptors.
-    \item An $Accept\left( leader \right)$ transition, enabled when
-        $status\left[ leader \right]$ is ``ready'' and the leader can
-        receive a $\aseq{\quo{prop},c}$ message. The effect of the
-        action is to receive the message, to update $hist\left[ leader
-        \right]$ to $hist\left[ leader \right]\star c$, and to
-        broadcast the message $\aseq{\quo{ack},hist\left[ leader
-        \right]\star c}$ to the learners, and to broadcast the message
-        $\aseq{\quo{leader-accept},hist\left[ leader \right]\star c}$
-        to all the other acceptors.
-    \item An $Accept\left( a \right)$ transition, for an acceptor $a$
-        which is not the leader, is enabled when $status\left[ a
-        \right]$ is ``ready'' and $a$ can receive a
-        $\aseq{\quo{leader-accept},h}$ message where $h > hist\left[ a
-        \right]$. The effect of the action is to receive the message,
-        to update $hist\left[ a \right]$ to $h$, and to broadcast the
-        message $\aseq{\quo{ack},hist\left[ a \right]\star c}$ to the
-        learners.
-    \item A $RcvAck\left( l \right)$ transition, for a learner $l$, is
-        enabled when $l$ can receive a message $\aseq{\quo{ack},h}$
-        from an acceptor $a$. Its effect is to receive the message and
-        to set $accepted\left[ l \right]\left[ a \right]$ to $h$.
-    \item A $Learn\left( l, h \right)$ transition, for a learner $l$
-        and a history $h$, is enabled when $h > Max\left(
-        learned\left[ l \right] \right)$ and there is a learn quorum
-        $Q$ of acceptors such that $h = GLB\left( \aset{accepted\left[
-        l \right]\left[ a \right] : a \in Q }\right)$. Its effect is
-        to insert $h$ in $learned\left[ l \right]$.
-    \item A $Stop\left( s \right)$ transition, for a switcher $s$,
-        always enabled, broadcasts a $\aseq{\quo{stop}}$ message to all
-        the acceptors.  This transition models a switcher being
-        instructed by the scheduling policy to abort the current round.
-    \item A $Stop\left( a \right)$ transition, for an acceptor $a$,
-        enabled when $a$ can receive a $\aseq{\quo{stop}}$ message
-        from a switcher $s$. Its effect is to receive the message, set
-        $status\left[ a \right]$ to ``stopped'', and to send the
-        message $\aseq{\quo{stopped},hist\left[ a \right]}$ to the
-        switcher $s$.
-    \item A $RcvStopped\left( s \right)$ transition, for a switcher
-        $s$, enabled when $s$ can receive a $\aseq{\quo{stopped},h}$
-        message from an acceptor $a$. Its effect is to receive the
-        message and to set $stopped\left[ s \right]\left[ a \right]$
-        to $h$.
-    \item An $Abort\left( s, h \right)$ transition, for a switcher $s$
-        and a history $h$, is enabled when there exists an abort quorum
-        $R$ such that, for every member $a$ of $R$, $stopped\left[ s
-        \right]\left[ a \right]$ is not $false$ and $h \in
-        SafeAborts\left( R \right)$, where $SafeAborts\left( R
-        \right)$ is computed using the array $stopped\left[ a
-        \right]$. Its effect is to insert  $h$ in the set $aborts$. 
-\end{compactitem}
-Compared to the rounds of the original Classic Paxos, Classic Paxos ACA
-rounds add an extra level of indirection when changing round: in the
-original rounds, the leader of the new round directly queries the
-acceptors of the preceeding rounds, whereas in ACA rounds, the
-switchers query the acceptors and then send their abort histories to
-the leader. However, changing round is an infrequent event, therefore
-this difference should not have much practical impact on performance.
-\begin{thm}
-  Classic Paxos ACA rounds satisfy Abortable Chain Agreement.
-\end{thm}
-\begin{proof}
-  A Classic Paxos ACA round implements the specification of Abortable
-  Chain Agreement under the refinement mapping consisting in
-  projecting the state of the round onto the variables $proposed$,
-  $inits$, $aborts$, and, for every learner $l$, $learned\left[ l
-  \right]$.
-\end{proof}
-
-\subsection{Multi-Coordinated Paxos, Chain Replication, and Ring Paxos.}
-
-Multi-Coordinated Paxos, Chain Replication, and Ring Paxos rounds have a similar
-structure as Classic Paxos rounds but differ in the way the leader is
-implemented and in the way that the histories of the leader are disseminated to
-the other acceptors.
-
-Multi-Coordinated Paxos avoids depending on a unique leader by using a
-distributed leader implementation that can make progress as long as a
-strict majority of the acceptors are correct, trading off resilience
-for a higher time complexity. Conservative ACA
-rounds can be implemented like with Classic Paxos, replacing the
-leader by the distributed leader of Multi-Coordinated Paxos.
-
-In Chain Replication, acceptors are arranged in a chain whose head is
-the leader of the round. As in Classic Paxos, abort histories and
-proposed commands are sent to the leader. However, instead of the
-leader directly sending its local history to all the other acceptors,
-the local history of the leader is
-forwarded along the chain from one acceptor to the other until it
-reaches the last acceptor, which sends the history to the learners. 
-By spreading the load more evenly on the all the servers, Chain Replication
-can achieve higher throughput that Classic Paxos.
-Conservative ACA rounds can be implemented like with Classic Paxos,
-replacing the broadcast of the leader by the propagation of the leader's history
-along the chain of servers.
-
-Ring Paxos optimizes Chain Replication by having the leader broadcast
-commands to the acceptors using IP multicast and only sending hashes
-of the commands along the chain. Ring Paxos round can be modified to implement
-conservative ACA rounds in a similar way as for Chain Replication.
-
-\subsection{Fast Paxos}
-
-Fast Paxos is composed of two types of rounds: classic rounds, like in Classic
-Paxos, and fast rounds.
-The fast rounds can be transformed into
-implementations of fast ACA rounds. Instead of sending their
-commands to a leader, as in Classic Paxos, the proposers of fast
-rounds directly broadcast their commands to all the acceptors. As long
-as commands are received in the same order by a learn quorum of
-acceptors or are received in different orders but commute, commands can be learned in two message-propagation delays instead of three in Classic Paxos.
-However, if non-commuting commands are received in different orders by
-several acceptors, the local history of the acceptors may not form a
-chain anymore and the rounds must abort. The fast rounds of Fast
-Paxos implement fast ACA rounds.
-
-Fast Paxos ACA rounds are specified using the same set of variables as for
-Classic Paxos, initialized to the same values.
-The next-state relation is composed of the following types of
-transitions, where the definitions of learn quorums, abort quorums,
-and of the $SafeAborts\left( R \right)$ operator are the ones of fast
-ACA rounds.
-\begin{compactitem}
-    \item A $Propose\left( p,c \right)$ transition, for a proposer $p$
-        and a command $c$, not guarded, inserts $c$ in the set
-        $proposed$ and broadcasts the message $\aseq{\quo{prop},c}$ to
-        all the acceptors.
-    \item An $Init\left( s, h \right)$ transition, for a switcher $s$
-        and a history $h$, not guarded, inserts $h$ in the set $inits$ and
-        broadcasts the message $\aseq{\quo{init},h}$ to all the acceptors.
-    \item A $WakeUp\left( a \right)$ transition, for an acceptor $a$,
-        is enabled when $status\left[ a \right]$ is ``idle'' and $a$
-        can receive an $\aseq{\quo{init},h}$ message. Its effect is to
-        receive the message, set $hist\left[ a \right]$ to $h$, and to
-        set $status\left[ a \right]$ to ``ready''.
-    \item An $Accept\left( a \right)$ transition, for an acceptor $a$,
-        is enabled when $status\left[ a \right]$ is ``ready'' and $a$
-        can receive a $\aseq{\quo{prop},c}$ message. The effect of the
-        action is to receive the message, to update $hist\left[ a
-        \right]$ to $hist\left[ a \right]\star c$, and to broadcast
-        the message $\aseq{\quo{ack},hist\left[ a \right]\star c}$ to
-        the learners.
-    \item The $RcvAck\left( l \right)$ transitions, for a learner $l$,
-        the $Learn\left( l, h \right)$ transitions, for a learner $l$
-        and a history $h$, the $Stop\left( s \right)$ transitions, for
-        a switcher $s$, $Stop\left( a \right)$ transitions, for an
-        acceptor $a$, the $RcvStopped\left( s \right)$ transitions,
-        for a switcher $s$, and the $Abort\left( s, h \right)$
-        transitions, for a switcher $s$ and a history $h$, which are
-        all the same as in Classic Paxos ACA rounds.
-\end{compactitem}
-
-As for Classic Paxos ACA rounds, Fast Paxos ACA rounds implement ACA
-under the refinement mapping consisting projecting the state of the
-round onto the variables $proposed$, $inits$, $aborts$, and, for every
-learner $l$, $learned\left[ l \right]$.
-
-\newpage
-
-\printbibliography
-
-\begin{appendix}
-
-  \label{sec:isaproofs}
-
-%  \includepdf[pages={1},pagecommand={\section{Mechanical Proof of the
-%  Composition Theorem in Isabelle/HOL}\label{sec:isaproofs}},scale=1]{../IsabellePDFS/podctheories.pdf}
-%  \includepdf[pages={2-},pagecommand={},scale=1]{../IsabellePDFS/podctheories.pdf}
-
-\end{appendix}
-
-\end{document}

### 2015-05-14 paper/paper.tex
Work on the presentation of CGC (section 3)

@@
 
-\title{Changing Flights In The Air\\ (Abortable Chain Agreement)}
+\title{Composable Generalized Consensus}
 
@@
 \input{roadmap}
-
-\begin{comment}
-Abortable Chain Agreement builds on several recent advances in SMR
-algorithms: ACA allows acceptors to only eventually reach agreement,
-like in Generalized Paxos \cite{Lamport05GeneralizeConsensus},
-boosting performance when most commands commute; ACA allows fast
-execution of read commands like in Generalized Lattice Agreement
-\cite{FalerioETAL12GeneralizedLatticeAgreement}; ACA implementations
-can tolerate $f$ faults with only $f+1$ replicas and replace crashed replicas
-upon changing round, as in Vertical Paxos  \cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}.  
-\end{comment}
 
@@
 
-\section{Chain Agreement}
-\label{sec:ca}
+\section{Composable Generalized Consensus}
+\label{sec:cgc}
 
-We consider a set of asynchronous servers communicating by message passing in an
-asynchronous network. Servers are not necessarily sequential processes. We
-describe a behavior of the system using an infinite sequence of states, where a
-state is a function from variable names to values (i.e. a valuation
-of variables).  We specify sets of allowed behaviors using an initial predicate,
-a next-state relation (using unprimed and primed variables) and a fairness
-condition, all in the style of TLA+ \cite{Lamport02SpecifyingSystems}. However,
-for succinctness, our specification are rather informal, eluding non-essential
-details.
+We consider a set of asynchronous servers communicating by message passing in
+an asynchronous network. Servers are not necessarily sequential processes.
+We describe a behavior of the system using an infinite sequence of states,
+where a state is a function from variable names to values (i.e. a valuation of
+variables). We specify sets of allowed behaviors using an initial predicate
+and a next-state relation expressed using unprimed and primed variables in the
+style of TLA+ \cite{Lamport02SpecifyingSystems}. However, for succinctness, our
+specification are rather informal, eluding non-essential details.
 
 Our specifications all contain by default a state variable $crashed$ and
-$Crash\left( x \right)$ transition which is always enabled. A server $x$ faithfully
+ a $Crash\left( x \right)$ transition which is always enabled. A server $x$ faithfully
 executes the algorithm that is assigned to it until its identifier is added to
@@
 assume that duplicate commands have no effect on the state of the
-service and return the same output as their first occurrence.
+service and return the same output as did their first occurrence.
 
+The goal of SMR is to provide a reliable implementation of the computing service despite server crashes.
+To do so, an SMR algorithm replicates the execution of the service on several servers. 
 In a traditional SMR algorithm, the servers can be seen has playing one or
 several of three roles: \emph{proposers}, which propose new commands to execute,
-\emph{acceptors}, which accept and order commands in a sequence, and
+\emph{acceptors}, whose role is to order commands, and
 \emph{learners}, which learn about the growing sequence of commands to
-execute \cite{lamport2001paxos}. Often, but not necessarily, every replica server plays the
+execute \cite{lamport2001paxos} by querying the acceptors. Often, but not necessarily, every replica server plays the
 three roles of proposer, acceptor, and learner: clients of the service
@@
 in the role of learners; replica servers participate in command
-ordering in the role of acceptors.  An SMR algorithm guarantees that
+ordering in the role of acceptors.  A SMR algorithm guarantees that
 the replicas learn the same sequences of commands, possibly up to the
@@
 
-Chain Agreement precisely specifies the allowed behaviors of the roles of \emph{proposers}
+\subsection{Histories}
+
+To simplify the specification of SMR algorithms optimizing the execution of commuting commands 
+we will use the notion of a command history, or history for short, in our specification of Composable Generalized Consensus.
+
+A command history, abbreviated history, is a data structure that
+represents a set of sequences which are the same modulo the reordering
+of commuting commands and addition or removal of duplicate commands.
+We say that two commands $c_1$ and $c_2$ \emph{commute} when, in every
+sequential execution of the service in which $c_1$ and $c_2$ are adjacent,
+reversing the order of $c_1$ and $c_2$ does not change any of the outputs,
+including those given to $c_1$ and $c_2$.
+
+All the sequences represented by a given history contain the same commands
+and determine the same outputs for each command. Therefore, to implement a
+replicated service, it is sufficient that learners learn histories instead of
+sequences. For a thourough discussion of command histories we refer the reader
+to \cite{Lamport05GeneralizeConsensus}. Our Isabelle/HOL formalization \cite{}
+abstracts over the concrete representation of histories, assuming only their
+necessary properties. Command histories were first introduced in trace theory
+\cite{Mazurkiewicz84Semantics}.
+
+Command histories can be built from the empty history using the $\star$
+operator, which appends a sequence of commands $cs$ to a history $h$, obtaining
+the history $h \star cs$. We can then define define an ordering relation among
+histories: $h_1 \leq h_2$ when there exists a sequence $cs$ such that $h_2 =
+h_1 \star cs$. 
+The maximum over a set of
+histories $H$ is denoted by $Max\left( H \right)$ and $Max\left( \aset{}
+\right)$ is the empty history.
+Histories have the property that every finite set $H$ of 
+histories has a \emph{greatest lower bound}, noted $GLB\left(H\right)$.
+
+The major advantage that histories have over sequences is that if a sequence
+$cs_1$ is obtained by reordering the commuting commands of $cs_2$, then $h \star
+cs_1 = h \star cs_2$. Therefore, in implementations of CGC, servers can agree on
+a history even though they chose different orders for some commuting commands.
+
+Generalized Consensus precisely specifies the allowed behaviors of the roles of \emph{proposers}
 and \emph{learners}, abstracting over the behavior of acceptors.
-
-\begin{comment}
-We suppose that each server can play a number of \emph{roles}.
-Chain Agreement specifies the allowed behaviors the roles of \emph{proposers}
-and \emph{learners}. Proposers receive commands from the clients of the service
-and propose them for execution by the replicated service. Each learner learns about
-a growing sequence of commands. Thanks to the sequential specification of the
-service, a sequence of commands determines an output for each the included commands. 
-In a traditional SMR algorithm, servers can also play the role of
-\emph{acceptors}, whose task is to collectively order commands and let the
-learners know about the latest sequence of ordered commands.
-
-The point of Chain Agreement is to implement a replicated service by having every learner learn about a growing sequence of commands, which
-represent an execution of the sequential specification of the service.
-Crucially, at all times, the sequences of commands learned by the
-different learners must form a chain in the prefix order on sequences.
-Moreover, for every command $c$ proposed by a correct proposer, every
-correct learner must eventually learn a sequence containing $c$.  Once
-a learner learns about a sequence containing a command $c$, it can
-determine the output produced by $c$, using the sequential
-specification of the service, and respond to the client that issued
-the command $c$.
-\end{comment}
 
@@
 its execution and passes the baton to the next round in the sequence.
-An algorithm composed of a sequence of ACA rounds is called an \emph{ACA
+An algorithm composed of a sequence of CGC instances is called an \emph{ACA
 algorithm}.
@@
 In normal circumstances, the proposers, acceptors,
-and learners of an ACA round work together to implement Chain Agreement.
-However, if the environment becomes unfavorable to a running ACA round, the
+and learners of a CGC instance work together to implement Chain Agreement.
+However, if the environment becomes unfavorable to a running CGC instance, the
 scheduling policy can instruct the round to abort its execution and pass the
@@
 
-The scheduling policy determines when to abort an ACA round and which new ACA
+The scheduling policy determines when to abort a CGC instance and which new ACA
 round should take over. The scheduling policy must inform the switchers about the
@@
 same information.  In this paper we abstract over the scheduling policy, assuming
-that a round that aborts passes the baton a nondeterministically chosen new
+that a round that aborts passes the baton to a nondeterministically chosen new
 round. A reliable scheduling-policy component can realistically be implemented
@@
 sequences transfered to the next round.
-Two consecutive ACA rounds are composed by letting the $aborts$ variable of the
+Two consecutive CGC instances are composed by letting the $aborts$ variable of the
 first round be the $inits$ variable of the second round.
@@
     \label{thm:comp}
-    In every execution of a succession of ACA rounds, the sequence of valuations
+    In every execution of a succession of CGC instances, the sequence of valuations
     of the variable $proposed$ and, for every learner $l$, of the variables
@@
     Consider two rounds numbered $i$ and $i+1$. In every execution of a
-    succession of ACA rounds, the sequence of valuations of the variables $inits
+    succession of CGC instances, the sequence of valuations of the variables $inits
     = inits_i$, $propose = proposed_i \cup propose_{i+1}$, for every learner
@@
 \Cref{thm:comp} guarantees that the successive composition of any
-number of ACA rounds implements Chain Agreement.
+number of CGC instances implements Chain Agreement.
 Therefore, with ACA, one can easily build replicated services that
@@
 tailored to particular conditions. Each time the conditions change,
-the current ACA round aborts and passes the baton to a more 
+the current CGC instance aborts and passes the baton to a more 
 appropriate type of round, as determined by the scheduling policy. Moreover, a new type
-of ACA round can always be devised after the system is deployed and added on the fly to
+of CGC instance can always be devised after the system is deployed and added on the fly to
 the existing types of rounds.
@@
 Most existing SMR algorithms can easily be obtained as the
-composition of one or more ACA rounds.
+composition of one or more CGC instances.
 
@@
 ACA is an abstraction of rounds which establishes a common interface and contract for switching from
-one round to the next, enabling switching from any ACA round to any other.
+one round to the next, enabling switching from any CGC instance to any other.
 
 To simplify the task of expressing existing algorithms as the composition of one
-or more ACA rounds, we propose two specifications, \emph{conservative ACA
+or more CGC instances, we propose two specifications, \emph{conservative ACA
 rounds} and
-\emph{fast ACA rounds}, which can easily be concretized to obtain executable rounds.
+\emph{fast CGC instances}, which can easily be concretized to obtain executable rounds.
 We show that ACA versions of Classic Paxos, Multi-Coordinated Paxos, Chain
 Replication, and Ring Paxos can be obtained by concretizing the specification of
-conservative ACA rounds, and that ACA versions of the fast rounds of Fast Paxos can be obtained by concretizing the specification of fast ACA rounds.
+conservative CGC instances, and that ACA versions of the fast rounds of Fast Paxos can be obtained by concretizing the specification of fast CGC instances.
 
-Fast and conservative ACA rounds are two specifications, at an
+Fast and conservative CGC instances are two specifications, at an
 intermediate level of abstraction, which satisfy ACA. Both
@@
 
-Conservative and fast ACA rounds are specified using the state variables $inits$, $proposed$, for every learner $l$, $learned\left[ l \right]$
+Conservative and fast CGC instances are specified using the state variables $inits$, $proposed$, for every learner $l$, $learned\left[ l \right]$
 and $aborts$, as in ACA, and, for every acceptor $a$, two new variables
@@
 On top of the additional guard on the $Accept$ transition present only
-in conservative rounds, conservative and fast ACA rounds differ in
+in conservative rounds, conservative and fast CGC instances differ in
 their definition of learn quorums, abort quorums, and in their
@@
 
-In conservative ACA rounds, abort and learn quorums are defined such
+In conservative CGC instances, abort and learn quorums are defined such
 that the intersection between two learn quorums is nonempty and the
@@
 
-In fast ACA rounds, abort and learn quorums are defined such that
+In fast CGC instances, abort and learn quorums are defined such that
 the intersection between any two learn quorums is nonempty and the
@@
   \label{thm:rounds}
-    Both conservative and fast ACA rounds implement ACA under the refinement
+    Both conservative and fast CGC instances implement ACA under the refinement
     mapping that consists in substituting $\overline{learned\left[ l \right]}$
@@
 Multi-Coordinated Paxos, Chain Replication, Ring Paxos, and Fast Paxos to obtain
-ACA rounds. \Cref{thm:comp} then guarantee that all those different types of
+CGC instances. \Cref{thm:comp} then guarantee that all those different types of
 rounds can be used in the same algorithm to yield an implementation of Chain
@@
 The rounds of the Classic Paxos algorithm can be modified so as to
-obtain an implementation of conservative ACA rounds.  We define
-Classic Paxos ACA rounds by specializing the specification of
-conservative ACA rounds, giving a concrete protocol for disseminating
+obtain an implementation of conservative CGC instances.  We define
+Classic Paxos CGC instances by specializing the specification of
+conservative CGC instances, giving a concrete protocol for disseminating
 the proposed commands to the acceptors and enforcing that the learned
@@
 
-The specification of Classic Paxos ACA rounds uses the variables of
-the specification of conservative ACA rounds, namely the variables
+The specification of Classic Paxos CGC instances uses the variables of
+the specification of conservative CGC instances, namely the variables
 $proposed$, $inits$, $aborts$, the arrays $learned\left[ l \right]$,
@@
 and of the $SafeAborts\left( R \right)$ operator are the ones of conservative
-ACA rounds.
+CGC instances.
 \begin{compactitem}
@@
 original rounds, the leader of the new round directly queries the
-acceptors of the preceeding round, whereas in ACA rounds, the
+acceptors of the preceeding round, whereas in CGC instances, the
 switchers query the acceptors and then send their abort histories to
@@
 \begin{thm}
-  Classic Paxos ACA rounds satisfy Abortable Chain Agreement.
+  Classic Paxos CGC instances satisfy Abortable Chain Agreement.
 \end{thm}
 \begin{proof}
-  A Classic Paxos ACA round implements the specification of Abortable
+  A Classic Paxos CGC instance implements the specification of Abortable
   Chain Agreement under the refinement mapping consisting in
@@
 can achieve higher throughput that Classic Paxos.
-Conservative ACA rounds can be implemented like with Classic Paxos,
+Conservative CGC instances can be implemented like with Classic Paxos,
 replacing the broadcast of the leader by the propagation of the leader's history
@@
 of the commands along the chain. Ring Paxos round can be modified to implement
-conservative ACA rounds in a similar way as for Chain Replication.
+conservative CGC instances in a similar way as for Chain Replication.
 
@@
 The fast rounds can be transformed into
-implementations of fast ACA rounds. Instead of sending their
+implementations of fast CGC instances. Instead of sending their
 commands to a leader, as in Classic Paxos, the proposers of fast
@@
 chain anymore and the rounds must abort. The fast rounds of Fast
-Paxos implement fast ACA rounds.
+Paxos implement fast CGC instances.
 
-Fast Paxos ACA rounds are specified using the same set of variables as for
+Fast Paxos CGC instances are specified using the same set of variables as for
 Classic Paxos, initialized to the same values.
@@
 and of the $SafeAborts\left( R \right)$ operator are the ones of fast
-ACA rounds.
+CGC instances.
 \begin{compactitem}
@@
         transitions, for a switcher $s$ and a history $h$, which are
-        all the same as in Classic Paxos ACA rounds.
+        all the same as in Classic Paxos CGC instances.
 \end{compactitem}
 
-As for Classic Paxos ACA rounds, Fast Paxos ACA rounds implement ACA
+As for Classic Paxos CGC instances, Fast Paxos CGC instances implement ACA
 under the refinement mapping consisting projecting the state of the

### 2015-05-14 paper/paper.tex
Work on the specification of CGC. Noticed that there is a problem: if learned histories are required to form a chain then the algorithms optimizing for the execution of commuting commands do not implement the spec. Fixed it in the paper but there clearly is no time to fix the TLA+ specs and even less the Isabelle theories.

@@
 histories has a \emph{greatest lower bound}, noted $GLB\left(H\right)$.
+Finally, we say that a set of histories $H$ is compatible when all the histories in $H$ have a common upper bound $h$ such
+that $h$ can be built using only commands found in the histories in $H$.
 
@@
 
-To specify Chain Agreement, we need the notion of ordering of sequences and of
-maximum sequence.  The prefix order on sequences is denoted by the operator
-$\leq$ ($s_1 \leq s_2$ means that $s_1$ is a prefix of $s_2$). 
-%Using the operator $\leq$, we define $\geq$, $<$, and $>$ as expected. 
-Note that $\leq$ is a partial order on sequences. The maximum over a set of
-sequences $S$ is denoted by $Max\left( S \right)$ and $Max\left( \aset{}
-\right)$ is the empty sequence.
-
-We specify Chain Agreement using the state variables $crashed$, $proposed$, containing the proposed commands,
+We specify Generalized Consensus using the state variables $crashed$, $proposed$, containing the proposed commands,
 and, for every learner $l$,  $learned\left[ l \right]$, a set of
@@
 \begin{compactitem}
-    \item[\textbf{Agreement}:] The elements of the set $\aset{learned'\left[ l
-      \right] : l \in Learners}$, the learned sequences after the transition,
-      form a chain in the partial order $\leq$.
+    \item[\textbf{Agreement}:] The set of learned histories after the transition, $\aset{learned'\left[ l
+      \right] : l \in Learners}$, is compatible.
     \item[\textbf{Validity}:] Every learned sequence contains only
@@
 \end{compactitem}
-Finally, we add, for every learner $l$, one weak fairness condition
-\cite{Lamport02SpecifyingSystems} on the transitions $Learn\left( l,
-cs \right)$. The weak fairness conditions ensure that every correct
-learner eventually learns about every proposed command.
 
-Thanks to the properties of Chain Agreement, the clients, observing
+Thanks to the properties of Generalized Consensus, the clients, observing
 their input commands and the received responses, see a
@@
 We now describe two important practical optimizations of ACA.
-
-\subsection{Histories}
-\label{sec:histories}
-
-ACA can be implemented more efficiently by substituting \emph{command histories} for sequences of commands, as in Generalized Consensus \cite{Lamport05GeneralizeConsensus}.
-
-We say that two commands $c_1$ and $c_2$ \emph{commute} when, in every sequential
-execution of the service, reversing the order of $c_1$ and $c_2$ does
-not change any of the outputs, including those given to $c_1$ and
-$c_2$.
-
-A command history, abbreviated history, is a data structure that
-represents a set of sequences which are the same modulo the reordering
-of commuting commands, and in which duplicate commands do not appear.
-All the sequences represented by a given history contain the same
-commands and determine the same outputs for each command.  Therefore,
-to implement a replicated service, it is sufficient that learners
-learn command histories instead of sequences.  For a precise
-definition of command histories, we refer the reader to
-\cite{Lamport05GeneralizeConsensus}. Our Isabelle/HOL formalization
-(\cref{sec:isaproofs}) abstracts
-over the concrete representation of histories, assuming only their necessary
-properties. Command histories were first
-introduced in trace theory \cite{Mazurkiewicz84Semantics}.
-
-A command history can be built by appending a sequence of commands
-$cs$ to a history $h$, obtain the history $h \star cs$.  The major
-advantage that histories have over sequences is that if a sequence
-$cs_1$ is obtained by reordering the commuting commands of $cs_2$,
-then $h \star cs_1 = h \star cs_2$. Therefore, in  implementations of
-ACA, servers can agree on a history even though
-they chose different orders for some commuting commands.
-
-Command histories share the four properties of sequences that are
-necessary and sufficient to substitute histories for sequences in the
-definition of Chain Agreement and ACA.
-\begin{compactitem}
-    \item[\textbf{Distinctness}:] No duplicate command appears in a history. 
-    \item[\textbf{Ordering}:] Command histories form a partial order:
-        $h_1 \leq h_2$ when there exists a sequence $s$ such that $h_2
-        = h_1 \star s$.
-    \item[\textbf{GLB}:] Every finite set of command histories has a
-        \emph{greatest lower bound}.
-%    \item[\textbf{LUB}:] Every finite set of compatible command
-%        histories has a \emph{least upper bound}, where a set of
-%        command histories is compatible when they have a common upper
-%        bound.
-    \item[\textbf{Consistency}:] If $h_1 \leq h_2 \leq h_3$ and $h_3 =
-        h_1 \star s$ then there exists $s'$ and $s''$ where every
-        request of $s'$ and $s''$ appears in $s$ and such that $h_2 =
-        h_1 \star s'$ and $h_3 = h_2 \star s''$.
-\end{compactitem}
-The optimized version of ACA is obtained by
-simply substituting histories for sequences in the definition of ACA.
-The Ordering, GLB, and Consistency properties of histories ensure that
-\cref{thm:comp} still holds: the Isabelle/HOL proof of \cref{thm:comp} (see
-\cref{sec:isaproofs}) relies only on the four properties above.
 

### 2015-05-14 paper/paper.tex
A few changes

@@
 
-\section{Composable Generalized Consensus}
-\label{sec:cgc}
+\section{Generalized Consensus}
+\label{sec:gc}
 
@@
 the history $h \star cs$. We can then define define an ordering relation among
-histories: $h_1 \leq h_2$ when there exists a sequence $cs$ such that $h_2 =
-h_1 \star cs$. 
-The maximum over a set of
-histories $H$ is denoted by $Max\left( H \right)$ and $Max\left( \aset{}
-\right)$ is the empty history.
-Histories have the property that every finite set $H$ of 
-histories has a \emph{greatest lower bound}, noted $GLB\left(H\right)$.
-Finally, we say that a set of histories $H$ is compatible when all the histories in $H$ have a common upper bound $h$ such
-that $h$ can be built using only commands found in the histories in $H$.
+histories: $h_1 \leq h_2$ when there exists a sequence $cs$ such that $h_2 = h_1
+\star cs$. Histories have the property that every finite set $H$ of histories
+has a \emph{greatest lower bound}, noted $GLB\left(H\right)$. Finally, we say
+that a set of histories $H$ is compatible when all the histories in $H$ have a
+common upper bound $h$ such that $h$ can be built using only commands found in
+the histories in $H$.
+%The maximum over a set of histories $H$ is denoted by $Max\left( H \right)$ and $Max\left( \aset{} \right)$ is the empty history.
 
@@
 a history even though they chose different orders for some commuting commands.
+
+\subsection{Generalized Consensus}
 
@@
     \item[\textbf{Irrevocability}:] The new learned sequence
-        $cs$ is a strict extension of $Max\left(learned\left[ l
-        \right]\right)$. 
+        $cs$ is an extension of $learned\left[ l
+        \right]$. 
         %\\ $cs > Max\left(learned\left[ l \right]\right)$.
@@
 \cite{HerlihyWing90LinearizabilityCorrectnessConditionConcurrentObjects}.
-Moreover, the service remains available as long as at least one
-proposer and one learner are correct.
 

### 2015-05-15 paper/paper.tex
Rework of the definition of CGC

@@
 
+\section{Modular Implementations of Generalized Consensus}
+
 \section{Abortable Chain Agreement}
@@
 \begin{compactitem}
-    \item[\textbf{Agreement}:] The elements of the set $\aset{learned'\left[ l
-      \right] : l \in Learners}$ form a chain in the partial order.
-    \item[\textbf{Validity}:] Every abort sequence or learned sequence
+    \item[\textbf{Agreement}:] The set $\aset{learned'\left[ l
+      \right] : l \in Learners}$ is compatible.
+    \item[\textbf{Validity}:] Every abort history or learned history 
         is of the form $h \star cs$, where $cs$ contains only commands
-        of the set $proposed$ and $h$ is the greatest lower bound of a
-        nonempty subset of $inits$.
+        of the set $proposed$ and $h$ is a history belonging to $inits$.
     \item[\textbf{Initialization}:] If $inits$ is empty, then no
         sequence can be learned and no switcher can abort.
-    \item[\textbf{Irrevocability}:] In the case of a $Learn\left( l,cs
-        \right)$ transition, the new learned sequence $cs$ is a strict
-        extension of $Max\left(learned\left[ l \right]\right)$.
-        %, $cs
-        %\geq Max\left(learned\left[ l \right]\right)$.
-    \item[\textbf{Safe Abort}:] Every abort sequence is larger in the
-        partial order $\leq$ than every learned sequence.
+    \item[\textbf{Irrevocability}:] In the case of a $Learn\left( l,h
+        \right)$ transition, the new learned history $h$ is an 
+        extension of $learned\left[ l \right]$.
+    \item[\textbf{Safe Abort}:] Every abort sequence is an extension in the 
+        partial order $\leq$ of a learned history.
 \end{compactitem}

### 2015-05-15 paper/paper.tex
Work on Fast GC in section 4.

@@
 \section{Modular Implementations of Generalized Consensus}
+
+Before formally introducing our framework, let us see how we can build a modular implementation similar to Fast Paxos by first designing a fast but fragile modular GC algorithm, then combining it unchanged with a more resilient module, obtaining an algorithm similar to Fast Paxos.
+
+\subsection{Fast Generalized Consensus}
+
+In our first algorithm, Fast GC, learners each maintain a variable $hist\left[ l \right]$ which contain the currently accepted history of learner $l$.
+For every learner $l$, $accepted\left[ l \right]$ is initialized to the empty history.
+A proposer broadcasts its proposal to all the learners which unconditionally accept it by setting appending it to their $accepted\left[ l \right]$
+variable and broadcasting to all the acceptors the new value of their accepted history.
+Each acceptor $l$ maintains a variable $learned\left[ l \right]$, initially the empty history, containing its last learned history. A learner $l$
+learns a new history $h$ when $h \geq learned\left[ l \right]$ and $h$ is a lower bound of the set consisting of the last accepted history received
+from each acceptor.
+
+More precisely, the Fast-GC algorithm is described by the sequence of states obtained by the transition relation obtain as the disjunction of the
+following transitions. We do not model the network explicitly using state variables but instead we just say that servers send, broadcast, or receive messages.
+
+\begin{compactitem}
+
+\item A $Propose\left( p,c \right)$ transition, for a proposer $p$ and a command $c$, not guarded, broadcasts the message $\aseq{\quo{prop},c}$ to all
+the acceptors.
+
+\item An $Accept\left( a \right)$ transition, for an acceptor $a$, enabled when $a$ can receive an $\aseq{\quo{prop},c}$ message from a proposer,
+updates $accepted\left[ a \right]$ to $accepted\left[ a \right]\star \aseq{c}$ and broadcasts the message $\aseq{\quo{accepted}, a, accepted'\left[ a \right]}$ to all the learners
+($accepted'\left[ a \right]$ is the new value of the variable $accepted\left[ a \right]$).
+
+\item A $Learn\left( l, h \right)$ transition, for a learner $l$ and a history $h$, enabled when $h \geq learned\left[ l \right]$ and $h$ is a lower
+bound of the set $\aset{LastReceived\left( a \right) : a \in Acceptors}$, where $LastReceived\left( a \right)$ is the last accepted history received
+from the acceptor $a$.
+
+\end{compactitem}
+
+The algorithm Fast GC cannot make progress if two acceptors $a_1$ and $a_2$ have two histories which are not compatible because, since one can only append to the histories $accepted\left[ l \right]$, all the possible future lower bounds of the set $\aset{LastReceived\left( a \right) : a \in Acceptors}$ are bounded above by $GLB\left( \aset{accepted\left[ a_1 \right], accepted\left[ a_2 \right]} \right)$.
+This situation occurs when there is contention on non-commuting commands: for example, when two different proposers send two commands $c_1$ and $c_2$ that do not commute, an acceptors $a_1$ accepts $c_1$ before $c_2$, and an acceptor $a_2 \neq a_1$ accepts $c_2$ before $c_1$.
+
+To overcome this deadlock, we will make some modifications to the algorithm that will allow us to start a fresh instance $FGC_2$, called a round, of the Fast GC algorithm.
+
+Basically, we introduce the new role of switcher, that any server can play additionally from being a proposer, acceptor, or learner.
+A switcher will extract a maximal accepted history from the first round and try to initialize all the acceptors of the second round with the extracted history.
+If we are luckier than in the first round, there may be no contention between switchers and every acceptor $a$ will start accepting new proposals in the second round based on the same initial history $accepted\left[ a \right]$.
+
+We write $FGC_1$ for the first, deadlocked, round of the Fast GC algorithm.
+The rounds $FCG_1$ and $FCG_2$ each have their own copies of the arrays $accepted$ and $learned$, noted $accepted_1$, $accepted_2$, $learend_1$, and $learned_2$ when the round is not clear from the context.
+We assume that network messages are tagged with a round number and that a message tagged with $i\in\aset{1,2}$ can only be received in $FCG_i$, creating two isolated logical networks.
+
+Switchers have the role of initializing the new round of Fast GC so that it can resume the execution and overcome the deadlock.
+
+We modify the FCG algorithm by introducing three new transitions and modifying the $Accept\left( a \right)$ transition as follows.
+We add the new variables $status\left[ a \right]$ for every acceptor $a$, which are initialized to $\quo{idle}$.
+The purpose of $status\left[ a \right]$ is to prevent the acceptor $a$ from accepting commands until its status changes from $\quo{idle}$ to $\quo{ready}$.
+We also add the new variables $inits\left[ s \right]$ and $aborts\left[ s \right]$, for every switcher $s$, initialized to the empty history in $FGC_1$ and to $\quo{none}$ in $FGC_2$.
+The purpose of the arrays $aborts$ and $inits$ is to transfer state between $FGC_1$ and $FGC_2$.
+The algorithms $FGC_1$ and $FGC_2$ each have their own copy of the arrays $status$, $inits$, and $aborts$, and we adopt the same convention as above consisting in adding a subscript denoted the round to a variable when the round is not clear from the context.
+
+\begin{compactitem}
+  \item We modify the $Accept\left( a \right)$ transition by having the acceptor $a$ additionally broadcast its new accepted value to the switchers.
+Switchers can therefore detect that the algorithm is deadlocked when there are two broadcast messages $\aseq{\quo{accepted}, a_1, h_1}$ and $\aseq{\quo{accepted}, a_2, h_2}$ in the network such that $h_1$ and $h_2$ are incompatible.
+
+\item The $Abort\left( s \right)$ transition, for a switcher $s$, is enabled when the switcher has detected a deadlock. Its effect is to 
+   set the variable $aborts\left[ s \right]$ to the maximum of the histories received from $a$, where $a$ is any acceptor. 
+   The history $aborts'\left[ s \right]$ is then called an abort history.
+
+ \item The $Init\left( s \right)$ transition, for a switcher $s$, always enabled, and its effect is to to pick a history $h$, set $inits\left[ s \right]$ to $h$, and to broadcast the message $\aseq{\quo{init}, h}$ to the acceptors.
+In $FGC_2$, composed with $FGC_1$, this history is constrained to be an abort history of $FGC_1$.
+   The history $inits\left[ s \right]$ is then called an init history.
+
+ \item The $WakeUp\left( a \right)$ transition, for an acceptor $a$, is enabled when $status\left[ a \right]$ is $\quo{idle}$ and there is at least one switcher $s$ such that $inits\left[ s \right] \neq \quo{none}$.
+Its effect is to set set $status\left[ a \right]$ to $\quo{ready}$ and $accepted\left[ a \right]$ to $inits\left[ s \right]$.
+
+\end{compactitem}
+
+The two rounds $FGC_1$ and $FGC_2$ are composed by requiring that $aborts_1 = inits_2$ at all times.
+This model the switchers passing the histories $aborts\left[ s \right]$ from $FGC_1$ to $FGC_2$, where they broadcast to the acceptors and used in the $WakeUp\left( a \right)$ transitions to initialize the state of the acceptors.
+This process can be repeated as often as needed, with new rounds $FGC_3$, $FGC_4$, etc., to try to make progress.
+
+The initial $FGC$ algorithm also enters a deadlock whenever one acceptor crashes because the learners wait for the lower bound over all accepted histories of all acceptors to increase.
+However, the round-change mechanism is very resilient to faults: a single live acceptor with a single live switcher can abort the execution and change round.
+To allow a new round to make progress after some acceptors crashed in a previous round, we introduce a component called the switching policy.
+With the Fast GC rounds, the task of the switching policy is to update the set of acceptors from one round to the next, ensuring that failed acceptors are removed or replaced by live ones.
+The switching policy must be queried by the switchers upon changing round to determine the new set of acceptors.
+This information can then be piggybacked to the
+acceptors and learners, or the acceptors and learns can directly query the switching policy too. 
+The set of switchers can be changed from round to round in the same fashion.
+Moreover, in an algorithm composed of several different types of rounds, the switching policy will also be responsible for choosing the type of round to execute next.
+
+\section{Classic Generalized Consensus}
+
+The Fast GC algorithm can process a proposal in two message delays when there is no contention on non-commuting commands. However, if there is
+repeated contention on non-commuting commands, the algorithm may not make progress despite changing round, as the acceptors of the each new round
+may be initialized with incompatible histories. The Fast Paxos algorithm \cite{Lamport06FastPaxos} is able to make progress under high contention by
+designating a leader server that will totally order the proposals.
+However, to make our Fast GC algorithm work like Fast Paxos, we would have to modify it substantially.
+If we had spent effort testing Fast GC implementations and proving it correct, this effort would be wasted and we would have to redo all the tests and proofs from scratch.
+
+Instead, we will now show how to obtain an algorithm with performance characteristics close to Fast Paxos, employing a leader to deal with contention, without modifying at all our existing Fast GC algorithm.
 

### 2015-05-15 paper/paper.tex
Work on presenting the Simple-Leader algorithm.

@@
 
-\subsection{Fast Generalized Consensus}
+\subsection{Fast Generalized Consensus (Aggrement By All?)}
 
@@
 Each acceptor $l$ maintains a variable $learned\left[ l \right]$, initially the empty history, containing its last learned history. A learner $l$
-learns a new history $h$ when $h \geq learned\left[ l \right]$ and $h$ is a lower bound of the set consisting of the last accepted history received
+learns a new history $h$ when $h \geq learned\left[ l \right]$ and $h$ is a lower bound of the set consisting of the maximal accepted history received
 from each acceptor.
@@
 \item A $Learn\left( l, h \right)$ transition, for a learner $l$ and a history $h$, enabled when $h \geq learned\left[ l \right]$ and $h$ is a lower
-bound of the set $\aset{LastReceived\left( a \right) : a \in Acceptors}$, where $LastReceived\left( a \right)$ is the last accepted history received
+bound of the set $\aset{MaxReceivedFrom\left( a \right) : a \in Acceptors}$, where $MaxReceivedFrom\left( a \right)$ is the maximal accepted history received
 from the acceptor $a$.
@@
 
-The algorithm Fast GC cannot make progress if two acceptors $a_1$ and $a_2$ have two histories which are not compatible because, since one can only append to the histories $accepted\left[ l \right]$, all the possible future lower bounds of the set $\aset{LastReceived\left( a \right) : a \in Acceptors}$ are bounded above by $GLB\left( \aset{accepted\left[ a_1 \right], accepted\left[ a_2 \right]} \right)$.
+The algorithm Fast GC cannot learn new histories if two acceptors $a_1$ and $a_2$ have two histories which are not compatible because, since one can only append to the histories $accepted\left[ l \right]$, all the possible future lower bounds of the set $\aset{MaxReceivedFrom\left( a \right) : a \in Acceptors}$ are bounded above by $GLB\left( \aset{accepted\left[ a_1 \right], accepted\left[ a_2 \right]} \right)$.
 This situation occurs when there is contention on non-commuting commands: for example, when two different proposers send two commands $c_1$ and $c_2$ that do not commute, an acceptors $a_1$ accepts $c_1$ before $c_2$, and an acceptor $a_2 \neq a_1$ accepts $c_2$ before $c_1$.
@@
 
-We write $FGC_1$ for the first, deadlocked, round of the Fast GC algorithm.
+We write $FGC_1$ for the first round of the Fast GC algorithm.
 The rounds $FCG_1$ and $FCG_2$ each have their own copies of the arrays $accepted$ and $learned$, noted $accepted_1$, $accepted_2$, $learend_1$, and $learned_2$ when the round is not clear from the context.
@@
 
-We modify the FCG algorithm by introducing three new transitions and modifying the $Accept\left( a \right)$ transition as follows.
+We modify the FCG algorithm by introducing new transitions and modifying the $Accept\left( a \right)$ transition as follows.
 We add the new variables $status\left[ a \right]$ for every acceptor $a$, which are initialized to $\quo{idle}$.
-The purpose of $status\left[ a \right]$ is to prevent the acceptor $a$ from accepting commands until its status changes from $\quo{idle}$ to $\quo{ready}$.
+The purpose of $status\left[ a \right]$ is to prevent the acceptor $a$ from accepting commands until its status changes from $\quo{idle}$ to $\quo{ready}$ and to stop it when its status goes to $\quo{stopped}$.
 We also add the new variables $inits\left[ s \right]$ and $aborts\left[ s \right]$, for every switcher $s$, initialized to the empty history in $FGC_1$ and to $\quo{none}$ in $FGC_2$.
 The purpose of the arrays $aborts$ and $inits$ is to transfer state between $FGC_1$ and $FGC_2$.
+The arrays $aborts$ and $inits$ form what we call the \emph{inter-round} interface.
 The algorithms $FGC_1$ and $FGC_2$ each have their own copy of the arrays $status$, $inits$, and $aborts$, and we adopt the same convention as above consisting in adding a subscript denoted the round to a variable when the round is not clear from the context.
@@
 \begin{compactitem}
-  \item We modify the $Accept\left( a \right)$ transition by having the acceptor $a$ additionally broadcast its new accepted value to the switchers.
+\item We modify the $Accept\left( a \right)$ transition by enabling it only if $status\left[ a \right] = \quo{ready}$ and by having the acceptor $a$ additionally broadcast its new accepted value to the switchers.
 Switchers can therefore detect that the algorithm is deadlocked when there are two broadcast messages $\aseq{\quo{accepted}, a_1, h_1}$ and $\aseq{\quo{accepted}, a_2, h_2}$ in the network such that $h_1$ and $h_2$ are incompatible.
 
-\item The $Abort\left( s \right)$ transition, for a switcher $s$, is enabled when the switcher has detected a deadlock. Its effect is to 
-   set the variable $aborts\left[ s \right]$ to the maximum of the histories received from $a$, where $a$ is any acceptor. 
-   The history $aborts'\left[ s \right]$ is then called an abort history.
+\item The $Stop\left( s \right)$ transition, for a switcher $s$, is enabled when the switcher has detected a deadlock or has waited for too long without receiving any message from the acceptors.
+Its effect is to broadcast a $\aseq{\quo{stop}, s}$ message to the acceptors.
 
- \item The $Init\left( s \right)$ transition, for a switcher $s$, always enabled, and its effect is to to pick a history $h$, set $inits\left[ s \right]$ to $h$, and to broadcast the message $\aseq{\quo{init}, h}$ to the acceptors.
+\item The $Stop\left( a \right)$ transition, for an acceptor $a$, is enabled when the acceptor $a$ can receive a $\aseq{\quo{stop}, s}$ message from a switcher $s$.
+  Its effect is to update $status\left[ a \right]$ to $\quo{stopped}$ and to send the message $\aseq{\quo{stopped},a,accepted\left[ a \right]}$ to the switcher $s$.
+
+\item The $Abort\left( s \right)$ transition, for a switcher $s$, is enabled when $aborts\left[ s \right] = \quo{none}$ and the switcher has received
+at least on $\aseq{\quo{stopped}, a, h}$ message. Its effect is to set the variable $aborts\left[ s \right]$ to the history $h$ received from $a$. The
+history $aborts'\left[ s \right]$ is then called an abort history.
+Because the switcher has received a $\aseq{\quo{stopped},a,h}$ message from the acceptor $a$, no history greater than $h$ can ever be learned anymore.
+Therefore, an abort history can be seen as a safe starting point for a new round to resume execution.
+
+\item The $Init\left( s \right)$ transition, for a switcher $s$, is enabled when $inits\left[ s \right]=\quo{none}$, and its effect is to to pick a history $h$, set $inits\left[ s \right]$ to $h$, and to broadcast the message $\aseq{\quo{init}, h}$ to the acceptors.
 In $FGC_2$, composed with $FGC_1$, this history is constrained to be an abort history of $FGC_1$.
@@
 
-The two rounds $FGC_1$ and $FGC_2$ are composed by requiring that $aborts_1 = inits_2$ at all times.
-This model the switchers passing the histories $aborts\left[ s \right]$ from $FGC_1$ to $FGC_2$, where they broadcast to the acceptors and used in the $WakeUp\left( a \right)$ transitions to initialize the state of the acceptors.
-This process can be repeated as often as needed, with new rounds $FGC_3$, $FGC_4$, etc., to try to make progress.
+The two rounds $FGC_1$ and $FGC_2$ are composed as follows
+\begin{itemize}
+  \item Requiring that a transition of the composition be a transition of either specification but not both, except for the $Abort\left( s \right)$ and $Init\left( s \right)$ transitions, which are allowed to take place at the same time.
+  \item Requiring that $aborts_1 = inits_2$ at all times.
+\end{itemize}
+This models two independent rounds being linked only by the switchers passing the histories $aborts\left[ s \right]$ from $FGC_1$ to $FGC_2$: switchers gather abort histories in $FGC_1$, then they broadcast them as init histories to the acceptors in $FGC_2$.
+The acceptors then initialize their state to an init histories in the $WakeUp\left( a \right)$ transition.
+This process can be repeated as often as needed, with new rounds $FGC_3$, $FGC_4$, etc., to try to make progress learning new histories.
 
 The initial $FGC$ algorithm also enters a deadlock whenever one acceptor crashes because the learners wait for the lower bound over all accepted histories of all acceptors to increase.
-However, the round-change mechanism is very resilient to faults: a single live acceptor with a single live switcher can abort the execution and change round.
+However, the round-change mechanism is very resilient to faults and can abort anyway thanks to the timeout triggering the $Stop\left( s \right)$ transition: a single live acceptor with a single live switcher can abort the execution and change round.
 To allow a new round to make progress after some acceptors crashed in a previous round, we introduce a component called the switching policy.
@@
 
-\section{Classic Generalized Consensus}
+\section{The Simple-Leader Algorithm}
 
@@
 repeated contention on non-commuting commands, the algorithm may not make progress despite changing round, as the acceptors of the each new round
-may be initialized with incompatible histories. The Fast Paxos algorithm \cite{Lamport06FastPaxos} is able to make progress under high contention by
+may be initialized with incompatible histories. The Fast Paxos algorithm \cite{Lamport06FastPaxos} is able to learn new histories under high contention by
 designating a leader server that will totally order the proposals.
@@
 
-Instead, we will now show how to obtain an algorithm with performance characteristics close to Fast Paxos, employing a leader to deal with contention, without modifying at all our existing Fast GC algorithm.
+Instead, we will now show how to obtain an algorithm with performance characteristics close to those of Fast Paxos, employing a leader to deal with contention, without modifying at all our existing Fast GC algorithm.
+To do so, we define a new kind of rounds, the Simple-Leader rounds, which uses the same inter-round interface as the Agreement-By-All algorithm. 
+The inter-round interface will allow use to combine Simple-Leader rounds and Agreement-By-All rounds in any order, obtaining an algorithm which can adapt its strategy to its environment, selecting the Agreement-By-All algorithm under low contention and switching to the more expensive Simple-Leader algorithm when there is contention.
+
+In the Simple-Leader algorithm, the proposers send their proposals to a distinguished acceptor called the leader. 
+The leader totally orders the proposals it receives and forwards them to all the acceptors which accept them in the order defined by the leader.
+As in the Agreement-By-All algorithm, acceptors broadcast their accepted histories to the learners, which learn a new history when it is a lower bound of 
+the set of $\aset{MaxReceivedFrom\left( a \right) : a \in Acceptors}$.
+Thanks to its leader, a Simple-Leader round continues to learn new histories even when proposals conflict because the leader ensures that the accepted histories of the acceptors remain consistent.
+
+The Simple-Leader algorithm treats all proposals equally and does not take advantage of commutativity.
+When the leader does not crash, each proposal is processed in three message delays.
+In comparison, in the absence of failures, the Agreement-By-All algorithm processes non-conflicting proposals in two message delays but can take arbitrarily long to process conflicting proposals.
+
+To initialize a Simple-Leader round, the switchers send their init histories to the leader, which choose a unique one that it forwards to the acceptors, guaranteeing that all the acceptors are initialized with the same history even when conflicting init histories are sent concurrently.
+
+The Simple-Leader algorithm can not learn new histories if its leader, or any other acceptor, crashes.
+To recover from this deadlock we pass the baton to a new round, as in the Agreement-By-All algorithm, with a similar same mechanism: a switcher that times-out waiting for new messages from the acceptors broadcast a stop message, gathers the accepted history from at least one of the acceptors, and aborts.
+
+Let us specify more precisely the Simple-Leader algorithm.
+
+The specification of Simple-Leader algorithm uses the sets   
+ $inits$, $aborts$, the arrays of variables $learned\left[ l \right]$,
+$status\left[ a \right]$, and $accepted\left[ a \right]$, and two
+additional arrays $accepted\left[ l \right]\left[ a \right]$ and
+$stop\left[ s \right]\left[ a \right]$, where $l$ is a learner, $a$ is
+an acceptor, and $s$ is a switcher.
+
+There is also a network component that we do not describe explicitly.
+However, the network allows any servers to communicate by sending and
+receiving messages.
+
+Initially, the variables $inits$, $aborts$, and, for every
+learner $l$, $learned\left[ l \right]$ are empty sets. For every
+acceptor $a$, $status\left[ a \right]$ is ``idle'' and $hist\left[ a
+\right]$ is the special value $None$. For every learner $l$ and acceptor $a$,
+$accepted\left[ l \right]\left[ a \right]$ is the special value
+$None$. For every switcher $s$ and acceptor $a$, $stopped\left[ s
+\right]\left[ a \right]$ is set to $false$.  There are initially no
+messages in the network.
+
+The next-state relation is composed of the following types of
+transitions, where the definitions of learn quorums, abort quorums,
+and of the $SafeAborts\left( R \right)$ operator are the ones of conservative
+CGC instances.
+\begin{compactitem}
+    \item A $Propose\left( p,c \right)$ transition, for a proposer $p$
+        and a command $c$, not guarded, inserts $c$ in the set
+        $proposed$ and sends the message $\aseq{\quo{prop},c}$ to
+        the leader.
+    \item An $Init\left( s, h \right)$ transition, for a switcher $s$
+        and a history $h$, not guarded, inserts $h$ in the set $inits$ and
+        sends the message $\aseq{\quo{init},h}$ to the leader.
+    \item A $WakeUp\left( leader \right)$ transition, enabled when
+        $status\left[ leader \right]$ is ``idle'' and the leader can
+        receive a $\aseq{\quo{init},h}$ message. Its effect is to
+        receive the message, set $hist\left[ leader \right]$ to $h$,
+        to set $status\left[ leader \right]$ to ``ready'', and to
+        broadcast the message $\aseq{\quo{leader-accept},h}$ to all
+        other acceptors.
+    \item An $Accept\left( leader \right)$ transition, enabled when
+        $status\left[ leader \right]$ is ``ready'' and the leader can
+        receive a $\aseq{\quo{prop},c}$ message. The effect of the
+        action is to receive the message, to update $hist\left[ leader
+        \right]$ to $hist\left[ leader \right]\star c$, and to
+        broadcast the message $\aseq{\quo{ack},hist\left[ leader
+        \right]\star c}$ to the learners, and to broadcast the message
+        $\aseq{\quo{leader-accept},hist\left[ leader \right]\star c}$
+        to all the other acceptors.
+    \item An $Accept\left( a \right)$ transition, for an acceptor $a$
+        which is not the leader, is enabled when $status\left[ a
+        \right]$ is ``ready'' and $a$ can receive a
+        $\aseq{\quo{leader-accept},h}$ message where $h > hist\left[ a
+        \right]$. The effect of the action is to receive the message,
+        to update $hist\left[ a \right]$ to $h$, and to broadcast the
+        message $\aseq{\quo{ack},hist\left[ a \right]\star c}$ to the
+        learners.
+    \item A $RcvAck\left( l \right)$ transition, for a learner $l$, is
+        enabled when $l$ can receive a message $\aseq{\quo{ack},h}$
+        from an acceptor $a$. Its effect is to receive the message and
+        to set $accepted\left[ l \right]\left[ a \right]$ to $h$.
+    \item A $Learn\left( l, h \right)$ transition, for a learner $l$
+        and a history $h$, is enabled when $h > Max\left(
+        learned\left[ l \right] \right)$ and there is a learn quorum
+        $Q$ of acceptors such that $h = GLB\left( \aset{accepted\left[
+        l \right]\left[ a \right] : a \in Q }\right)$. Its effect is
+        to insert $h$ in $learned\left[ l \right]$.
+    \item A $Stop\left( s \right)$ transition, for a switcher $s$,
+        always enabled, broadcasts a $\aseq{\quo{stop}}$ message to all
+        the acceptors.  This transition models a switcher being
+        instructed by the scheduling policy to abort the current round.
+    \item A $Stop\left( a \right)$ transition, for an acceptor $a$,
+        enabled when $a$ can receive a $\aseq{\quo{stop}}$ message
+        from a switcher $s$. Its effect is to receive the message, set
+        $status\left[ a \right]$ to ``stopped'', and to send the
+        message $\aseq{\quo{stopped},hist\left[ a \right]}$ to the
+        switcher $s$.
+    \item A $RcvStopped\left( s \right)$ transition, for a switcher
+        $s$, enabled when $s$ can receive a $\aseq{\quo{stopped},h}$
+        message from an acceptor $a$. Its effect is to receive the
+        message and to set $stopped\left[ s \right]\left[ a \right]$
+        to $h$.
+    \item An $Abort\left( s, h \right)$ transition, for a switcher $s$
+        and a history $h$, is enabled when there exists an abort quorum
+        $R$ such that, for every member $a$ of $R$, $stopped\left[ s
+        \right]\left[ a \right]$ is not $false$ and $h \in
+        SafeAborts\left( R \right)$, where $SafeAborts\left( R
+        \right)$ is computed using the array $stopped\left[ a
+        \right]$. Its effect is to insert  $h$ in the set $aborts$. 
+\end{compactitem}
+
+Thanks to the inter-round interface, Simple-Leader rounds and Agreement-By-All rounds can be composed in any order.
+To compose a Simple-Leader round $SL_1$ with an Agreement-By-All round $ABA_2$, we compose their specifications as before, adding the constraint that the set $aborts_1$ be equal to the set $inits_2$ at all times.
+
+Informally, the composition is correct because abort histories are always safe in the sense that any of the abort histories could be the next learned history.
+However not all can be learned and the initialization mechanism of the rounds ensures that at most one is chosen.
+In the next section we present Composable Generalized Consensus, a specification of the behavior of rounds at the level of proposals, learned histories, init histories, and abort histories.
+Our composition theorem ensures that any set of rounds individually satisfying Composable Generalized Consensus can be composed in any order to yield an algorithm satisfying Composable Generalized Consensus.
 
@@
 commands to the acceptors using IP multicast and only sending hashes
-of the commands along the chain. Ring Paxos round can be modified to implement
+of the commands along the chain. Ring Paxos rounds can be modified to implement
 conservative CGC instances in a similar way as for Chain Replication.

### 2015-05-15 paper/paper.tex
Work on the description of the Broadcast and Simple-Leader algorithms.

--- 2015-05-15 paper/paper.tex
@@
 There are numerous examples of consensus algorithms in the literature.
-However, Abortable Chain Agreement is based on Generalized Consensus. 
+However, Composable Generalized Consensus is based on Generalized Consensus. 
 In Generalized Consensus \cite{Lamport05GeneralizeConsensus}, the learners of a
@@
 agreement on C-structs without further synchronization. 
-Our Abortable Chain Agreement can also use C-structs, with the same advantages
+Our Composable Generalized Consensus can also use C-structs, with the same advantages
 as Generalized Consensus. For simplicity, the presentation in this
@@
 commands are partitioned into read-only commands and write-only commands.
-Abortable Chain Agreement supports the same optimizations as Generalized Lattice Agreement.
+Composable Generalized Consensus supports the same optimizations as Generalized Lattice Agreement.
 Notably, read-only commands are treated specially so as to optimize their
@@
 be removed from or added to the configuration.
-ACA shares some aspects of Vertical Paxos algorithms: the task of the
+CGC shares some aspects of Vertical Paxos algorithms: the task of the
 reconfiguration master is similar to the task of the scheduling policy
-component of ACA (see \cref{sec:aca}).
-However, in ACA, changing round can be used not only to replace crashed servers
+component of CGC (see \cref{sec:aca}).
+However, in CGC, changing round can be used not only to replace crashed servers
 but also to change the algorithm that the servers are running. 
 Changing algorithm is not possible in Vertical Paxos.
-In both Vertical Paxos and ACA, a reliable reconfiguration master, resp. a
+In both Vertical Paxos and CGC, a reliable reconfiguration master, resp. a
 reliable scheduling policy, allows tolerating $f$ crashes with $f+1$ servers.
@@
 optimization of the execution of commuting and read-only commands that are
-possible with ACA.  Moreover, the Abstract Framework does not identify the
+possible with CGC.  Moreover, the Abstract Framework does not identify the
 crucial role of the scheduling policy for the long-term resilience of a service.
-Abortable Chain Agreement combines the ideas of the Abstract Framework,
+Composable Generalized Consensus combines the ideas of the Abstract Framework,
 Generalized Paxos, Generalized Lattice Agreement, and Vertical Paxos in a
@@
 sequences of commands, with the same drawback for the execution of
-commuting and read-only commands. Moreover, Abortable Chain Agreement is both
+commuting and read-only commands. Moreover, Composable Generalized Consensus is both
 conceptually simpler and subsumes the Speculative Linearizability
-property.  Like Speculative Linearizability, Abortable Chain Agreement
+property.  Like Speculative Linearizability, Composable Generalized Consensus
 and its main property, the composition theorem (\cref{thm:comp}), are
@@
 
-Abortable Chain Agreement is the most complete framework to date for building
+Composable Generalized Consensus is the most complete framework to date for building
 adaptive replication algorithms out of heterogeneous rounds, making state of the
@@
 
-\section{Abortable Chain Agreement}
+Before formally introducing our framework, let us see how we can build a modular implementation similar to Fast Paxos by first designing a fast but fragile modular GC algorithm, then combining it unchanged with a more resilient module, obtaining an algorithm similar to Fast Paxos.
+
+\subsection{The Broadcast Algorithm}
+
+In our first algorithm, Broadcast, learners each maintain a variable $hist\left[ l \right]$ which contain the currently accepted history of learner $l$.
+For every learner $l$, $accepted\left[ l \right]$ is initialized to the empty history.
+A proposer broadcasts its proposal to all the learners which unconditionally accept it by appending the proposal to their $accepted\left[ l \right]$
+history and by broadcasting to all the acceptors their new accepted history, $accepted'\left[ l \right]$.
+
+Each acceptor $l$ maintains a variable $learned\left[ l \right]$, initially the empty history, containing its last learned history. A learner $l$
+learns a new history $h$ when $h \geq learned\left[ l \right]$ and $h$ is a lower bound of the set consisting of the maximal accepted history received
+from each acceptor.
+
+More precisely, the Broadcast algorithm is described by the transition relation obtained as the disjunction of the
+following transitions. We do not model the network explicitly using state variables but instead we just say that servers send, broadcast, or receive messages.
+
+\begin{compactitem}
+
+\item A $Propose\left( p,c \right)$ transition, for a proposer $p$ and a command $c$, not guarded, broadcasts the message $\aseq{\quo{prop},c}$ to all
+the acceptors.
+
+\item An $Accept\left( a \right)$ transition, for an acceptor $a$, enabled when $a$ can receive a $\aseq{\quo{prop},c}$ message from a proposer,
+updates $accepted\left[ a \right]$ to $accepted\left[ a \right]\star \aseq{c}$ and broadcasts the message $\aseq{\quo{accepted}, a, accepted'\left[ a \right]}$ to all the learners
+($accepted'\left[ a \right]$ is the new value of the variable $accepted\left[ a \right]$).
+
+\item A $Learn\left( l, h \right)$ transition, for a learner $l$ and a history $h$, enabled when $h \geq learned\left[ l \right]$ and $h$ is a lower
+bound of the set $\aset{MaxReceivedFrom\left( a \right) : a \in Acceptors}$, where $MaxReceivedFrom\left( a \right)$ is the maximal accepted history received
+from the acceptor $a$.
+
+\end{compactitem}
+
+The algorithm Broadcast cannot learn new histories if two acceptors $a_1$ and $a_2$ have two histories which are not compatible because, since an acceptor can only append new proposals to its history $accepted\left[ l \right]$, all the possible future lower bounds of the set $\aset{MaxReceivedFrom\left( a \right) : a \in Acceptors}$ are bounded above by $GLB\left( \aset{accepted\left[ a_1 \right], accepted\left[ a_2 \right]} \right)$.
+This situation occurs when non-commuting commands contend: for example, when two different proposers send two commands $c_1$ and $c_2$ that do not commute, the acceptors $a_1$ accepts $c_1$ before $c_2$, and the acceptor $a_2 \neq a_1$ accepts $c_2$ before $c_1$.
+
+To overcome this deadlock, we will make some modifications to the algorithm that will allow us to start a fresh instance $B_2$, called a round, of the Broadcast algorithm.
+To allow switching to a fresh new instance of the Broadcast algorithm, we introduce the new role of switcher, that any server can play additionally from being a proposer, acceptor, or learner.
+A switcher will extract an accepted history from the first round and try to initialize all the acceptors of the second round with the extracted history.
+If we are luckier than in the first round, there may be no contention between switchers and every acceptor $a$ will start accepting new proposals in the second round based on the same initial history.
+
+We write $B_1$ for the first round of the Broadcast algorithm.
+The rounds $B_1$ and $B_2$ each have their own copies of the arrays $accepted$ and $learned$, noted $accepted_1$, $accepted_2$, $learned_1$, and $learned_2$ when the round is not clear from the context.
+We assume that network messages are tagged with a round number and that a message tagged with $i\in\aset{1,2}$ can only be received in $B_i$, creating two isolated logical networks.
+
+In our example with the deadlock, switchers have the role of initializing the new round of $B_2$ so that it can resume the execution and overcome the deadlock.
+We modify the Broadcast algorithm by introducing new transitions and modifying the $Accept\left( a \right)$ transition as follows.
+We add the new variables $status\left[ a \right]$ for every acceptor $a$, which are initialized to $\quo{idle}$.
+The purpose of $status\left[ a \right]$ is to prevent the acceptor $a$ from accepting commands until its status changes from $\quo{idle}$ to $\quo{ready}$ and to stop it when its status goes to $\quo{stopped}$.
+We also add the new arrays of variables $inits\left[ s \right]$ and $aborts\left[ s \right]$, for every switcher $s$, initialized to the empty history in $B_1$ and to $\quo{none}$ in $B_2$.
+The purpose of the arrays of variables $aborts$ and $inits$ is to transfer state between $B_1$ and $B_2$.
+The arrays of variables $aborts$ and $inits$ form what we call the \emph{inter-round} interface.
+As for the arrays $accepted$ and $learned$, the rounds $B_1$ and $B_2$ each have their own copy of the arrays of variables $status$, $inits$, and $aborts$, and we adopt add subscripts denoting the round when the round is not clear from the context.
+
+\begin{compactitem}
+\item We modify the $Accept\left( a \right)$ transition by enabling it only if $status\left[ a \right] = \quo{ready}$ and by having the acceptor $a$ additionally broadcast its new accepted value to the switchers.
+Switchers detect that the algorithm is deadlocked when they have received two messages $\aseq{\quo{accepted}, a_1, h_1}$ and $\aseq{\quo{accepted}, a_2, h_2}$ such that $h_1$ and $h_2$ are incompatible.
+
+\item The $Stop\left( s \right)$ transition, for a switcher $s$, is enabled when the switcher has detected a deadlock or has waited for too long without receiving any message from the acceptors.
+Its effect is to broadcast a $\aseq{\quo{stop}, s}$ message to the acceptors.
+
+\item The $Stop\left( a \right)$ transition, for an acceptor $a$, is enabled when the acceptor $a$ can receive a $\aseq{\quo{stop}, s}$ message from a switcher $s$.
+  Its effect is to update $status\left[ a \right]$ to $\quo{stopped}$ and to send the message $\aseq{\quo{stopped},a,accepted\left[ a \right]}$ to the switcher $s$.
+
+\item The $Abort\left( s \right)$ transition, for a switcher $s$, is enabled when $aborts\left[ s \right] = \quo{none}$ and the switcher has received
+  at least one $\aseq{\quo{stopped}, a, h}$ message.
+Its effect is to set the variable $aborts\left[ s \right]$ to the history $h$ received in the $\quo{stopped}$ message.
+The history $aborts'\left[ s \right]$ is then called an abort history.
+Because the switcher has received a $\aseq{\quo{stopped},a,h}$ message from the acceptor $a$, the acceptor $a$ no longer accepts new proposals, therefore no history greater than $h$ can ever be learned anymore.
+Therefore, an abort history can be seen as a safe starting point for a new round to resume execution.
+
+\item The $Init\left( s \right)$ transition, for a switcher $s$, is enabled when $inits\left[ s \right]=\quo{none}$, and its effect is to pick a history $h$, set $inits\left[ s \right]$ to $h$, and to broadcast the message $\aseq{\quo{init}, h}$ to the acceptors.
+The history $inits\left[ s \right]$ is then called an init history.
+As described below, in $B_2$ composed with $B_1$, this history is constrained to be an abort history of $B_1$.
+
+ \item The $WakeUp\left( a \right)$ transition, for an acceptor $a$, is enabled when $status\left[ a \right]$ is $\quo{idle}$ and there is at least one switcher $s$ such that $inits\left[ s \right] \neq \quo{none}$.
+Its effect is to set set $status\left[ a \right]$ to $\quo{ready}$ and $accepted\left[ a \right]$ to $inits\left[ s \right]$.
+
+\end{compactitem}
+
+The two rounds $B_1$ and $B_2$ are composed as follows:
+\begin{itemize}
+  \item We Require that a transition of the composition be a transition of either specification but not both, except for the $Abort\left( s \right)$ and $Init\left( s \right)$ transitions, which are allowed to take place at the same time.
+  \item We require that $aborts_1 = inits_2$ at all times.
+\end{itemize}
+The composition models two independent rounds being linked only by the switchers passing the abort histories from $B_1$ to $B_2$, where they become init histories: switchers gather abort histories in $B_1$, then they broadcast them as init histories to the acceptors in $B_2$.
+The acceptors then initialize their state to a received init history in the $WakeUp\left( a \right)$ transition.
+This process can be repeated as often as needed, with new rounds $B_3$, $B_4$, etc., to try to make progress learning new histories.
+
+The Broadcast algorithm can no more learn new histories whenever one acceptor crashes because the learners wait for the lower bound of the accepted histories of all acceptors to increase.
+However, the round-change mechanism is very resilient to faults and can abort anyway thanks to the timeout triggering the $Stop\left( s \right)$ transition: a single live acceptor with a single live switcher can abort the execution and change round.
+
+To allow a new round to make progress after some acceptors crashed in a previous round, we introduce a component called the switching policy.
+With the Broadcast rounds, the task of the switching policy is to update the set of acceptors from one round to the next, ensuring that failed acceptors are removed or replaced by live ones.
+The switching policy must be queried by the switchers upon changing round to determine the new set of acceptors.
+This information can then be piggybacked to the
+acceptors and learners, or the acceptors and learners can directly query the switching policy. 
+The set of switchers can be changed from round to round in the same fashion.
+Moreover, in an algorithm composed of several different types of rounds, the switching policy will also be responsible for choosing the type of round to execute next.
+
+\section{The Simple-Leader Algorithm}
+
+The Broadcast algorithm can process a proposal in two message delays when there is no contention on non-commuting commands. However, if there is
+repeated contention on non-commuting commands, the algorithm may not make progress despite changing round, as the acceptors of the each new round
+may be initialized with incompatible histories. The Fast Paxos algorithm \cite{Lamport06FastPaxos} is able to learn new histories under high contention by
+designating a leader server that will totally order the proposals.
+However, to make our Broadcast algorithm work like Fast Paxos, we would have to modify it substantially.
+If we had spent effort testing Broadcast implementations and proving it correct, this effort would be wasted and we would have to redo all the tests and proofs from scratch.
+
+Instead, we will now show how to obtain an algorithm with performance characteristics close to those of Fast Paxos, employing a leader to deal with contention, without modifying at all our existing Broadcast algorithm.
+To do so, we define a new kind of rounds, the Simple-Leader rounds, which uses the same inter-round interface as the Broadcast algorithm. 
+The inter-round interface will allow use to combine Simple-Leader rounds and Broadcast rounds in any order, obtaining an algorithm which can adapt its strategy to its environment, selecting the Broadcast algorithm under low contention and switching to the more expensive Simple-Leader algorithm when there is contention.
+
+In the Simple-Leader algorithm, the proposers send their proposals to a distinguished acceptor called the leader. 
+The leader totally orders the proposals it receives and forwards them to all the acceptors which accept them in the order defined by the leader.
+As in the Broadcast algorithm, acceptors broadcast their accepted histories to the learners, which learn a new history when it is a lower bound of 
+the set of $\aset{MaxReceivedFrom\left( a \right) : a \in Acceptors}$.
+Thanks to its leader, a Simple-Leader round continues to learn new histories even when proposals conflict because the leader ensures that the accepted histories of the acceptors remain consistent.
+
+The Simple-Leader algorithm treats all proposals equally and does not take advantage of commutativity.
+When the leader does not crash, each proposal is processed in three message delays.
+In comparison, in the absence of failures, the Broadcast algorithm processes non-conflicting proposals in two message delays but can take arbitrarily long to process conflicting proposals.
+
+To initialize a Simple-Leader round, the switchers send their init histories to the leader, which choose a unique one that it forwards to the acceptors, guaranteeing that all the acceptors are initialized with the same history even when conflicting init histories are sent concurrently.
+
+The Simple-Leader algorithm can not learn new histories if its leader, or any other acceptor, crashes.
+To recover from this deadlock we pass the baton to a new round, as in the Broadcast algorithm, with a similar same mechanism: a switcher that times-out waiting for new messages from the acceptors broadcast a stop message, gathers the accepted history from at least one of the acceptors, and aborts.
+
+Let us specify more precisely the Simple-Leader algorithm.
+
+The specification of Simple-Leader algorithm uses the sets   
+ $inits$, $aborts$, the arrays of variables $learned\left[ l \right]$,
+$status\left[ a \right]$, and $accepted\left[ a \right]$, and two
+additional arrays $accepted\left[ l \right]\left[ a \right]$ and
+$stop\left[ s \right]\left[ a \right]$, where $l$ is a learner, $a$ is
+an acceptor, and $s$ is a switcher.
+
+There is also a network component that we do not describe explicitly.
+However, the network allows any servers to communicate by sending and
+receiving messages.
+
+Initially, the variables $inits$, $aborts$, and, for every
+learner $l$, $learned\left[ l \right]$ are empty sets. For every
+acceptor $a$, $status\left[ a \right]$ is ``idle'' and $hist\left[ a
+\right]$ is the special value $None$. For every learner $l$ and acceptor $a$,
+$accepted\left[ l \right]\left[ a \right]$ is the special value
+$None$. For every switcher $s$ and acceptor $a$, $stopped\left[ s
+\right]\left[ a \right]$ is set to $false$.  There are initially no
+messages in the network.
+
+The next-state relation is composed of the following types of
+transitions, where the definitions of learn quorums, abort quorums,
+and of the $SafeAborts\left( R \right)$ operator are the ones of conservative
+CGC instances.
+\begin{compactitem}
+    \item A $Propose\left( p,c \right)$ transition, for a proposer $p$
+        and a command $c$, not guarded, inserts $c$ in the set
+        $proposed$ and sends the message $\aseq{\quo{prop},c}$ to
+        the leader.
+    \item An $Init\left( s, h \right)$ transition, for a switcher $s$
+        and a history $h$, not guarded, inserts $h$ in the set $inits$ and
+        sends the message $\aseq{\quo{init},h}$ to the leader.
+    \item A $WakeUp\left( leader \right)$ transition, enabled when
+        $status\left[ leader \right]$ is ``idle'' and the leader can
+        receive a $\aseq{\quo{init},h}$ message. Its effect is to
+        receive the message, set $hist\left[ leader \right]$ to $h$,
+        to set $status\left[ leader \right]$ to ``ready'', and to
+        broadcast the message $\aseq{\quo{leader-accept},h}$ to all
+        other acceptors.
+    \item An $Accept\left( leader \right)$ transition, enabled when
+        $status\left[ leader \right]$ is ``ready'' and the leader can
+        receive a $\aseq{\quo{prop},c}$ message. The effect of the
+        action is to receive the message, to update $hist\left[ leader
+        \right]$ to $hist\left[ leader \right]\star c$, and to
+        broadcast the message $\aseq{\quo{ack},hist\left[ leader
+        \right]\star c}$ to the learners, and to broadcast the message
+        $\aseq{\quo{leader-accept},hist\left[ leader \right]\star c}$
+        to all the other acceptors.
+    \item An $Accept\left( a \right)$ transition, for an acceptor $a$
+        which is not the leader, is enabled when $status\left[ a
+        \right]$ is ``ready'' and $a$ can receive a
+        $\aseq{\quo{leader-accept},h}$ message where $h > hist\left[ a
+        \right]$. The effect of the action is to receive the message,
+        to update $hist\left[ a \right]$ to $h$, and to broadcast the
+        message $\aseq{\quo{ack},hist\left[ a \right]\star c}$ to the
+        learners.
+    \item A $RcvAck\left( l \right)$ transition, for a learner $l$, is
+        enabled when $l$ can receive a message $\aseq{\quo{ack},h}$
+        from an acceptor $a$. Its effect is to receive the message and
+        to set $accepted\left[ l \right]\left[ a \right]$ to $h$.
+    \item A $Learn\left( l, h \right)$ transition, for a learner $l$
+        and a history $h$, is enabled when $h > Max\left(
+        learned\left[ l \right] \right)$ and there is a learn quorum
+        $Q$ of acceptors such that $h = GLB\left( \aset{accepted\left[
+        l \right]\left[ a \right] : a \in Q }\right)$. Its effect is
+        to insert $h$ in $learned\left[ l \right]$.
+    \item A $Stop\left( s \right)$ transition, for a switcher $s$,
+        always enabled, broadcasts a $\aseq{\quo{stop}}$ message to all
+        the acceptors.  This transition models a switcher being
+        instructed by the scheduling policy to abort the current round.
+    \item A $Stop\left( a \right)$ transition, for an acceptor $a$,
+        enabled when $a$ can receive a $\aseq{\quo{stop}}$ message
+        from a switcher $s$. Its effect is to receive the message, set
+        $status\left[ a \right]$ to ``stopped'', and to send the
+        message $\aseq{\quo{stopped},hist\left[ a \right]}$ to the
+        switcher $s$.
+    \item A $RcvStopped\left( s \right)$ transition, for a switcher
+        $s$, enabled when $s$ can receive a $\aseq{\quo{stopped},h}$
+        message from an acceptor $a$. Its effect is to receive the
+        message and to set $stopped\left[ s \right]\left[ a \right]$
+        to $h$.
+    \item An $Abort\left( s, h \right)$ transition, for a switcher $s$
+        and a history $h$, is enabled when there exists an abort quorum
+        $R$ such that, for every member $a$ of $R$, $stopped\left[ s
+        \right]\left[ a \right]$ is not $false$ and $h \in
+        SafeAborts\left( R \right)$, where $SafeAborts\left( R
+        \right)$ is computed using the array $stopped\left[ a
+        \right]$. Its effect is to insert  $h$ in the set $aborts$. 
+\end{compactitem}
+
+Thanks to the inter-round interface, Simple-Leader rounds and Broadcast rounds can be composed in any order.
+To compose a Simple-Leader round $SL_1$ with an Broadcast round $ABA_2$, we compose their specifications as before, adding the constraint that the set $aborts_1$ be equal to the set $inits_2$ at all times.
+
+Informally, the composition is correct because abort histories are always safe in the sense that any of the abort histories could be the next learned history.
+However not all can be learned and the initialization mechanism of the rounds ensures that at most one is chosen.
+In the next section we present Composable Generalized Consensus, a specification of the behavior of rounds at the level of proposals, learned histories, init histories, and abort histories.
+Our composition theorem ensures that any set of rounds individually satisfying Composable Generalized Consensus can be composed in any order to yield an algorithm satisfying Composable Generalized Consensus.
+
+\section{Composable Generalized Consensus}
 \label{sec:aca}
 
-We now consider implementing Chain Agreement as a sequence of \emph{Abortable Chain
-Agreement rounds}, where each round makes progress ordering requests or aborts
-its execution and passes the baton to the next round in the sequence.
-An algorithm composed of a sequence of CGC instances is called an \emph{ACA
-algorithm}.
+We now consider implementing Generalized Consensus as a sequence of \emph{Composable Generalized Consensus rounds}, where each round makes progress
+ordering requests or aborts its execution and passes the baton to the next round in the sequence. An algorithm composed of a sequence of CGC instances
+is called a \emph{CGC algorithm}.
 
-On top of the traditional roles of proposer, acceptor, and learner, ACA
+On top of the traditional roles of proposer, acceptor, and learner, CGC
 introduces the new role of \emph{switcher} and a new component, the
-\emph{scheduling policy}. In ACA, the assignment of the roles of proposer,
+\emph{scheduling policy}. In CGC, the assignment of the roles of proposer,
 acceptor, learner, and switcher can change in each rounds, allowing for
@@
 In normal circumstances, the proposers, acceptors,
-and learners of a CGC instance work together to implement Chain Agreement.
+and learners of a CGC instance work together to implement Generalized Consensus.
 However, if the environment becomes unfavorable to a running CGC instance, the
@@
 
-The scheduling policy determines when to abort a CGC instance and which new ACA
+The scheduling policy determines when to abort a CGC instance and which new CGC
 round should take over. The scheduling policy must inform the switchers about the
@@
 
-The clients of an ACA algorithm can send their commands to the proposers of any round (in
+The clients of a CGC algorithm can send their commands to the proposers of any round (in
 practice, clients send their commands to proposers of the largest
@@
 
-ACA specifies the allowed behaviors of the proposers, learners, and switchers
+CGC specifies the allowed behaviors of the proposers, learners, and switchers
 \emph{of a single round}. Similarly to
-Chain Agreement, we use the state variables $crashed$, $proposed$,
+Generalized Consensus, we use the state variables $crashed$, $proposed$,
 and, for every learner $l$, $learned\left[ l \right]$.  Additionally,
@@
 switcher $x$, for a learner $l$, and for a sequence of commands
-$cs$, have the same effect as in Chain Agreement.
-ACA introduces two new transitions, $Init\left(
+$cs$, have the same effect as in Generalized Consensus.
+CGC introduces two new transitions, $Init\left(
 s, cs \right)$ and $Abort\left( s, cs \right)$, for a switcher $s$ and a sequence of
@@
       \right] : l \in Learners}$ is compatible.
-    \item[\textbf{Validity}:] Every abort history or learned history 
+    \item[\textbf{Validity}:] There is a history $h$ belonging to $inits$ such that every abort history or learned history 
         is of the form $h \star cs$, where $cs$ contains only commands
-        of the set $proposed$ and $h$ is a history belonging to $inits$.
+        of the set $proposed$.
     \item[\textbf{Initialization}:] If $inits$ is empty, then no
@@
 \end{compactitem}
-Note that Agreement and Irrevocability are the same as in Chain
-Agreement.
+Note that Agreement and Irrevocability are the same as in Generalized Consensus.
 
-Finally, for every learner $l$, we add a weak fairness condition on the
-following state transition: 
-\begin{equation}
-    \exists  s \in Switchers, cs \in Sequence\left( C \right) : Learn\left( l, cs \right) \vee Abort\left( s,cs \right)
-\end{equation}
-The weak fairness conditions ensure that when a correct learner has not
-learned a sequence containing all commands proposed in the round, then
-it eventually learns a new command or, if there is at least one correct switcher, at least one switch value is transfered to the next round. 
-Therefore, if a client has submitted a command to a proposer, eventually,
+If a client has submitted a command to a proposer, eventually,
 either the client gets a response from the current round, or
@@
 \subsection{The Composition Theorem}
+
+TODO: frame in terms of composability.
 
@@
     of the variable $proposed$ and, for every learner $l$, of the variables
-    $learned\left[ l \right]$ satisfies Chain Agreement.
+    $learned\left[ l \right]$ satisfies Generalized Consensus.
 \end{thm}
@@
     $l$, $learned\left[ l \right] = learned_i\left[ l \right] \cup
-    learned_{i+1}\left[ l \right]$and $aborts = aborts_{i+1}$ satisfies ACA.
+    learned_{i+1}\left[ l \right]$and $aborts = aborts_{i+1}$ satisfies CGC.
 \end{lem}
@@
   A mechanically-checked proof in Isabelle/HOL appears in \cref{sec:isaproofs}.
-  The proofs uses a version of the specification of ACA that has a more
+  The proofs uses a version of the specification of CGC that has a more
   executable form that presented in this section. Moreover,
@@
   general notion of command history, presented in the next subsection.
-  The Isabelle/HOL formalization of ACA uses the theory of I/O automata, which
+  The Isabelle/HOL formalization of CGC uses the theory of I/O automata, which
   is also formalized in the Isabelle/HOL. The proof uses a refinement
-  mapping from the composition of two consecutive ACA instances to a single ACA
+  mapping from the composition of two consecutive CGC instances to a single CGC
   instance.
@@
 \Cref{thm:comp} guarantees that the successive composition of any
-number of CGC instances implements Chain Agreement.
-Therefore, with ACA, one can easily build replicated services that
+number of CGC instances implements Generalized Consensus.
+Therefore, with CGC, one can easily build replicated services that
 dynamically adapt to the current operating conditions of the system: a
@@
 
-We now describe two important practical optimizations of ACA.
+We now describe two important practical optimizations of CGC.
 
@@
 
-Chain Agreement can be further optimized by treating 
+Generalized Consensus can be further optimized by treating 
 read-only commands specially, as in Generalized Lattice Agreement
@@
 prevents mixing rounds of two different algorithms.  Instead,
-ACA is an abstraction of rounds which establishes a common interface and contract for switching from
+CGC is an abstraction of rounds which establishes a common interface and contract for switching from
 one round to the next, enabling switching from any CGC instance to any other.
@@
 To simplify the task of expressing existing algorithms as the composition of one
-or more CGC instances, we propose two specifications, \emph{conservative ACA
+or more CGC instances, we propose two specifications, \emph{conservative CGC
 rounds} and
 \emph{fast CGC instances}, which can easily be concretized to obtain executable rounds.
-We show that ACA versions of Classic Paxos, Multi-Coordinated Paxos, Chain
+We show that CGC versions of Classic Paxos, Multi-Coordinated Paxos, Chain
 Replication, and Ring Paxos can be obtained by concretizing the specification of
-conservative CGC instances, and that ACA versions of the fast rounds of Fast Paxos can be obtained by concretizing the specification of fast CGC instances.
+conservative CGC instances, and that CGC versions of the fast rounds of Fast Paxos can be obtained by concretizing the specification of fast CGC instances.
 
 Fast and conservative CGC instances are two specifications, at an
-intermediate level of abstraction, which satisfy ACA. Both
+intermediate level of abstraction, which satisfy CGC. Both
 specifications specify how the state of the acceptors should be
@@
 Conservative and fast CGC instances are specified using the state variables $inits$, $proposed$, for every learner $l$, $learned\left[ l \right]$
-and $aborts$, as in ACA, and, for every acceptor $a$, two new variables
+and $aborts$, as in CGC, and, for every acceptor $a$, two new variables
 $status\left[ a \right]$, which can be either ``idle'', ``ready'', or
@@
   \label{thm:rounds}
-    Both conservative and fast CGC instances implement ACA under the refinement
+    Both conservative and fast CGC instances implement CGC under the refinement
     mapping that consists in substituting $\overline{learned\left[ l \right]}$
@@
 
-\section{Examples of Concrete ACA Algorithms}
+\begin{comment}
+\section{Examples of Concrete CGC Algorithms}
 \label{sec:examples}
@@
 \end{compactitem}
-Compared to the rounds of the original Classic Paxos, Classic Paxos ACA
+Compared to the rounds of the original Classic Paxos, Classic Paxos CGC
 rounds add an extra level of indirection when changing round: in the
@@
 \begin{thm}
-  Classic Paxos CGC instances satisfy Abortable Chain Agreement.
+  Classic Paxos CGC instances satisfy Composable Generalized Consensus.
 \end{thm}
@@
   A Classic Paxos CGC instance implements the specification of Abortable
-  Chain Agreement under the refinement mapping consisting in
+  Generalized Consensus under the refinement mapping consisting in
   projecting the state of the round onto the variables $proposed$,
@@
 
+\end{comment}
 \subsection{Multi-Coordinated Paxos, Chain Replication, and Ring Paxos.}
@@
 strict majority of the acceptors are correct, trading off resilience
-for a higher time complexity. Conservative ACA
+for a higher time complexity. Conservative CGC
 rounds can be implemented like with Classic Paxos, replacing the
@@
 commands to the acceptors using IP multicast and only sending hashes
-of the commands along the chain. Ring Paxos round can be modified to implement
+of the commands along the chain. Ring Paxos rounds can be modified to implement
 conservative CGC instances in a similar way as for Chain Replication.
 
+\begin{comment}
 \subsection{Fast Paxos}
@@
 
-As for Classic Paxos CGC instances, Fast Paxos CGC instances implement ACA
+As for Classic Paxos CGC instances, Fast Paxos CGC instances implement CGC
 under the refinement mapping consisting projecting the state of the
@@
 learner $l$, $learned\left[ l \right]$.
+
+\end{comment}
 

### 2015-05-15 paper/paper.tex
Fixed typos and terminology

@@
 
-Command histories can be built from the empty history using the $\star$
+Command histories can be built from the empty history, noted $\bot$, using the $\star$
 operator, which appends a sequence of commands $cs$ to a history $h$, obtaining
@@
 \star cs$. Histories have the property that every finite set $H$ of histories
-has a \emph{greatest lower bound}, noted $GLB\left(H\right)$. Finally, we say
+has a \emph{greatest lower bound}, noted $GLB\left(H\right)$. 
+We say that a history $h$ contains the command $c$ when there exists a sequence of commands 
+$cs$ containing $c$ such that $\bot \star cs = h$.
+Finally, we say
 that a set of histories $H$ is compatible when all the histories in $H$ have a
@@
 $cs_1$ is obtained by reordering the commuting commands of $cs_2$, then $h \star
-cs_1 = h \star cs_2$. Therefore, in implementations of CGC, servers can agree on
+cs_1 = h \star cs_2$. Therefore, in SMR implementations, servers can agree on
 a history even though they chose different orders for some commuting commands.
@@
 
-We specify Generalized Consensus using the state variables $crashed$, $proposed$, containing the proposed commands,
-and, for every learner $l$,  $learned\left[ l \right]$, a set of
-sequences of commands. Initially, $crashed$, $proposed$ and, for every learner $l$,
-$learned\left[ l \right]$ are empty sets. 
+We specify Generalized Consensus using the variables $crashed$, the variable $proposed$, containing the proposed commands,
+and, for every learner $l$,  a variable $learned\left[ l \right]$, holding a history. Initially, $crashed$ and $proposed$ are empty sets and, for every learner $l$,
+$learned\left[ l \right]$ is the empty history. 
 Remember that any transition involving a server $x$ is always disabled when $x\in crashed$.
@@
 \right)$ transitions, for a learner $l$ and a command $c$, in which
-$c$ is inserted in the set $proposed$; $Learn\left( l,cs \right)$
-transitions, for a learner $l$ and a sequence of commands $cs$, in
-which $cs$ is inserted in $learned\left[ l \right]$. 
-A $Learn\left( l, cs \right)$ transition is guarded by the following
-conditions:
+$c$ is inserted in the set $proposed$; $Learn\left( l,h \right)$
+transitions, for a learner $l$ and a history $h$, in
+which $learned\left[ l \right]$ is update to $h$. 
+A $Learn\left( l, h \right)$ transition is guarded by the following conditions:
 \begin{compactitem}
@@
       \right] : l \in Learners}$, is compatible.
-    \item[\textbf{Validity}:] Every learned sequence contains only
+    \item[\textbf{Validity}:] Every learned history contains only
         proposed commands.  
-    \item[\textbf{Irrevocability}:] The new learned sequence
-        $cs$ is an extension of $learned\left[ l
+    \item[\textbf{Irrevocability}:] The new learned history
+      $learned'\left[ l \right]$ is an extension of $learned\left[ l
         \right]$. 
@@
 
-Thanks to the properties of Generalized Consensus, the clients, observing
-their input commands and the received responses, see a
-\emph{linearizable} execution
+When clients of a replicated service submit their commands to the proposers and learners compute the response to give to clients using their local $learned\left[ l \right]$ history, the three properties above imply that the clients see a \emph{linearizable} execution
 \cite{HerlihyWing90LinearizabilityCorrectnessConditionConcurrentObjects}.
@@
 
-\subsection{The Broadcast Algorithm}
-
-In our first algorithm, Broadcast, learners each maintain a variable $hist\left[ l \right]$ which contain the currently accepted history of learner $l$.
+\subsection{The Broadcast-Decide Algorithm}
+
+In our first algorithm, Broadcast-Decide, learners each maintain a variable $hist\left[ l \right]$ which contain the currently accepted history of learner $l$.
 For every learner $l$, $accepted\left[ l \right]$ is initialized to the empty history.
@@
 
-More precisely, the Broadcast algorithm is described by the transition relation obtained as the disjunction of the
+More precisely, the Broadcast-Decide algorithm is described by the transition relation obtained as the disjunction of the
 following transitions. We do not model the network explicitly using state variables but instead we just say that servers send, broadcast, or receive messages.
@@
 
-The algorithm Broadcast cannot learn new histories if two acceptors $a_1$ and $a_2$ have two histories which are not compatible because, since an acceptor can only append new proposals to its history $accepted\left[ l \right]$, all the possible future lower bounds of the set $\aset{MaxReceivedFrom\left( a \right) : a \in Acceptors}$ are bounded above by $GLB\left( \aset{accepted\left[ a_1 \right], accepted\left[ a_2 \right]} \right)$.
+The algorithm Broadcast-Decide cannot learn new histories if two acceptors $a_1$ and $a_2$ have two histories which are not compatible because, since an acceptor can only append new proposals to its history $accepted\left[ l \right]$, all the possible future lower bounds of the set $\aset{MaxReceivedFrom\left( a \right) : a \in Acceptors}$ are bounded above by $GLB\left( \aset{accepted\left[ a_1 \right], accepted\left[ a_2 \right]} \right)$.
 This situation occurs when non-commuting commands contend: for example, when two different proposers send two commands $c_1$ and $c_2$ that do not commute, the acceptors $a_1$ accepts $c_1$ before $c_2$, and the acceptor $a_2 \neq a_1$ accepts $c_2$ before $c_1$.
 
-To overcome this deadlock, we will make some modifications to the algorithm that will allow us to start a fresh instance $B_2$, called a round, of the Broadcast algorithm.
-To allow switching to a fresh new instance of the Broadcast algorithm, we introduce the new role of switcher, that any server can play additionally from being a proposer, acceptor, or learner.
+To overcome this deadlock, we will make some modifications to the algorithm that will allow us to start a fresh instance $B_2$, called a round, of the Broadcast-Decide algorithm.
+To allow switching to a fresh new instance of the Broadcast-Decide algorithm, we introduce the new role of switcher, that any server can play additionally from being a proposer, acceptor, or learner.
 A switcher will extract an accepted history from the first round and try to initialize all the acceptors of the second round with the extracted history.
@@
 
-We write $B_1$ for the first round of the Broadcast algorithm.
+We write $B_1$ for the first round of the Broadcast-Decide algorithm.
 The rounds $B_1$ and $B_2$ each have their own copies of the arrays $accepted$ and $learned$, noted $accepted_1$, $accepted_2$, $learned_1$, and $learned_2$ when the round is not clear from the context.
@@
 In our example with the deadlock, switchers have the role of initializing the new round of $B_2$ so that it can resume the execution and overcome the deadlock.
-We modify the Broadcast algorithm by introducing new transitions and modifying the $Accept\left( a \right)$ transition as follows.
+We modify the Broadcast-Decide algorithm by introducing new transitions and modifying the $Accept\left( a \right)$ transition as follows.
 We add the new variables $status\left[ a \right]$ for every acceptor $a$, which are initialized to $\quo{idle}$.
@@
 
-The Broadcast algorithm can no more learn new histories whenever one acceptor crashes because the learners wait for the lower bound of the accepted histories of all acceptors to increase.
+The Broadcast-Decide algorithm can no more learn new histories whenever one acceptor crashes because the learners wait for the lower bound of the accepted histories of all acceptors to increase.
 However, the round-change mechanism is very resilient to faults and can abort anyway thanks to the timeout triggering the $Stop\left( s \right)$ transition: a single live acceptor with a single live switcher can abort the execution and change round.
@@
 To allow a new round to make progress after some acceptors crashed in a previous round, we introduce a component called the switching policy.
-With the Broadcast rounds, the task of the switching policy is to update the set of acceptors from one round to the next, ensuring that failed acceptors are removed or replaced by live ones.
+With the Broadcast-Decide rounds, the task of the switching policy is to update the set of acceptors from one round to the next, ensuring that failed acceptors are removed or replaced by live ones.
 The switching policy must be queried by the switchers upon changing round to determine the new set of acceptors.
@@
 
-The Broadcast algorithm can process a proposal in two message delays when there is no contention on non-commuting commands. However, if there is
+The Broadcast-Decide algorithm can process a proposal in two message delays when there is no contention on non-commuting commands. However, if there is
 repeated contention on non-commuting commands, the algorithm may not make progress despite changing round, as the acceptors of the each new round
@@
 designating a leader server that will totally order the proposals.
-However, to make our Broadcast algorithm work like Fast Paxos, we would have to modify it substantially.
-If we had spent effort testing Broadcast implementations and proving it correct, this effort would be wasted and we would have to redo all the tests and proofs from scratch.
-
-Instead, we will now show how to obtain an algorithm with performance characteristics close to those of Fast Paxos, employing a leader to deal with contention, without modifying at all our existing Broadcast algorithm.
-To do so, we define a new kind of rounds, the Simple-Leader rounds, which uses the same inter-round interface as the Broadcast algorithm. 
-The inter-round interface will allow use to combine Simple-Leader rounds and Broadcast rounds in any order, obtaining an algorithm which can adapt its strategy to its environment, selecting the Broadcast algorithm under low contention and switching to the more expensive Simple-Leader algorithm when there is contention.
+However, to make our Broadcast-Decide algorithm work like Fast Paxos, we would have to modify it substantially.
+If we had spent effort testing Broadcast-Decide implementations and proving it correct, this effort would be wasted and we would have to redo all the tests and proofs from scratch.
+
+Instead, we will now show how to obtain an algorithm with performance characteristics close to those of Fast Paxos, employing a leader to deal with contention, without modifying at all our existing Broadcast-Decide algorithm.
+To do so, we define a new kind of rounds, the Simple-Leader rounds, which uses the same inter-round interface as the Broadcast-Decide algorithm. 
+The inter-round interface will allow use to combine Simple-Leader rounds and Broadcast-Decide rounds in any order, obtaining an algorithm which can adapt its strategy to its environment, selecting the Broadcast-Decide algorithm under low contention and switching to the more expensive Simple-Leader algorithm when there is contention.
 
@@
 The leader totally orders the proposals it receives and forwards them to all the acceptors which accept them in the order defined by the leader.
-As in the Broadcast algorithm, acceptors broadcast their accepted histories to the learners, which learn a new history when it is a lower bound of 
+As in the Broadcast-Decide algorithm, acceptors broadcast their accepted histories to the learners, which learn a new history when it is a lower bound of 
 the set of $\aset{MaxReceivedFrom\left( a \right) : a \in Acceptors}$.
@@
 When the leader does not crash, each proposal is processed in three message delays.
-In comparison, in the absence of failures, the Broadcast algorithm processes non-conflicting proposals in two message delays but can take arbitrarily long to process conflicting proposals.
+In comparison, in the absence of failures, the Broadcast-Decide algorithm processes non-conflicting proposals in two message delays but can take arbitrarily long to process conflicting proposals.
 
@@
 The Simple-Leader algorithm can not learn new histories if its leader, or any other acceptor, crashes.
-To recover from this deadlock we pass the baton to a new round, as in the Broadcast algorithm, with a similar same mechanism: a switcher that times-out waiting for new messages from the acceptors broadcast a stop message, gathers the accepted history from at least one of the acceptors, and aborts.
+To recover from this deadlock we pass the baton to a new round, as in the Broadcast-Decide algorithm, with a similar same mechanism: a switcher that times-out waiting for new messages from the acceptors broadcast a stop message, gathers the accepted history from at least one of the acceptors, and aborts.
+
+****************************************
 
@@
 
-Thanks to the inter-round interface, Simple-Leader rounds and Broadcast rounds can be composed in any order.
-To compose a Simple-Leader round $SL_1$ with an Broadcast round $ABA_2$, we compose their specifications as before, adding the constraint that the set $aborts_1$ be equal to the set $inits_2$ at all times.
+Thanks to the inter-round interface, Simple-Leader rounds and Broadcast-Decide rounds can be composed in any order.
+To compose a Simple-Leader round $SL_1$ with an Broadcast-Decide round $B_2$, we compose their specifications as before, adding the constraint that the set $aborts_1$ be equal to the set $inits_2$ at all times.
 

### 2015-05-15 paper/paper.tex
Work on the description of the Simple-Leader algorithm.

@@
 from the acceptor $a$.
+ Its effect is to set $learned\left[ l \right]$ to $h$.
 
@@
 
-\item The $Abort\left( s \right)$ transition, for a switcher $s$, is enabled when $aborts\left[ s \right] = \quo{none}$ and the switcher has received
+\item The $Abort\left( s, h\right)$ transition, for a switcher $s$, is enabled when $aborts\left[ s \right] = \quo{none}$ and the switcher has received
   at least one $\aseq{\quo{stopped}, a, h}$ message.
@@
 
-\item The $Init\left( s \right)$ transition, for a switcher $s$, is enabled when $inits\left[ s \right]=\quo{none}$, and its effect is to pick a history $h$, set $inits\left[ s \right]$ to $h$, and to broadcast the message $\aseq{\quo{init}, h}$ to the acceptors.
+\item The $Init\left( s,h \right)$ transition, for a switcher $s$, is enabled when $inits\left[ s \right]=\quo{none}$ and $h$ is a history, and its effect is to set $inits\left[ s \right]$ to $h$, and to broadcast the message $\aseq{\quo{init}, h}$ to the acceptors.
 The history $inits\left[ s \right]$ is then called an init history.
@@
 
-The two rounds $B_1$ and $B_2$ are composed as follows:
+The composition $B_1\times B_2$ of the two rounds $B_1$ and $B_2$ is obtained as follows:
 \begin{itemize}
@@
 
-****************************************
-
 Let us specify more precisely the Simple-Leader algorithm.
 
-The specification of Simple-Leader algorithm uses the sets   
- $inits$, $aborts$, the arrays of variables $learned\left[ l \right]$,
-$status\left[ a \right]$, and $accepted\left[ a \right]$, and two
-additional arrays $accepted\left[ l \right]\left[ a \right]$ and
-$stop\left[ s \right]\left[ a \right]$, where $l$ is a learner, $a$ is
-an acceptor, and $s$ is a switcher.
-
-There is also a network component that we do not describe explicitly.
-However, the network allows any servers to communicate by sending and
-receiving messages.
-
-Initially, the variables $inits$, $aborts$, and, for every
-learner $l$, $learned\left[ l \right]$ are empty sets. For every
-acceptor $a$, $status\left[ a \right]$ is ``idle'' and $hist\left[ a
-\right]$ is the special value $None$. For every learner $l$ and acceptor $a$,
-$accepted\left[ l \right]\left[ a \right]$ is the special value
-$None$. For every switcher $s$ and acceptor $a$, $stopped\left[ s
-\right]\left[ a \right]$ is set to $false$.  There are initially no
-messages in the network.
-
-The next-state relation is composed of the following types of
-transitions, where the definitions of learn quorums, abort quorums,
-and of the $SafeAborts\left( R \right)$ operator are the ones of conservative
-CGC instances.
+The specification of Simple-Leader algorithm uses the same variables as the Broadcast-Decide algorithm and the same initial state. 
+%that is the arrays of variables    
+%$inits\left[ s \right]$, $aborts\left[ s \right]$, $learned\left[ l \right]$,
+%$status\left[ a \right]$, and $accepted\left[ a \right]$, where $l$ is a learner, $a$ is
+%an acceptor, and $s$ is a switcher.
+
+%Initially, the variables $inits$, $aborts$, and, for every
+%learner $l$, $learned\left[ l \right]$ are empty sets. For every
+%acceptor $a$, $status\left[ a \right]$ is ``idle'' and $hist\left[ a
+%\right]$ is the special value $None$.
+
+The next-state relation is the disjunction of the following types of
+transitions. 
 \begin{compactitem}
 \item A $Propose\left( p,c \right)$ transition, for a proposer $p$
-        and a command $c$, not guarded, inserts $c$ in the set
-        $proposed$ and sends the message $\aseq{\quo{prop},c}$ to
+  and a command $c$, not guarded, sends the message $\aseq{\quo{prop},c}$ to
   the leader.
 \item An $Init\left( s, h \right)$ transition, for a switcher $s$
-        and a history $h$, not guarded, inserts $h$ in the set $inits$ and
-        sends the message $\aseq{\quo{init},h}$ to the leader.
+  and a history $h$, is enabled when $inits\left[ s \right] = \quo{none}$, and its effect is to set $inits\left[ s \right]$ to $h$ and
+  to sends the message $\aseq{\quo{init},h}$ to the leader.
 \item A $WakeUp\left( leader \right)$ transition, enabled when
-        $status\left[ leader \right]$ is ``idle'' and the leader can
-        receive a $\aseq{\quo{init},h}$ message. Its effect is to
-        receive the message, set $hist\left[ leader \right]$ to $h$,
+  $status\left[ leader \right]$ is ``idle'' and 
+  a message $\aseq{\quo{init},h}$ was sent to the leader. Its effect is to
+  receive the message, set $accepted\left[ leader \right]$ to $h$,
   to set $status\left[ leader \right]$ to ``ready'', and to
@@
 \item An $Accept\left( leader \right)$ transition, enabled when
-        $status\left[ leader \right]$ is ``ready'' and the leader can
-        receive a $\aseq{\quo{prop},c}$ message. The effect of the
-        action is to receive the message, to update $hist\left[ leader
-        \right]$ to $hist\left[ leader \right]\star c$, and to
-        broadcast the message $\aseq{\quo{ack},hist\left[ leader
-        \right]\star c}$ to the learners, and to broadcast the message
-        $\aseq{\quo{leader-accept},hist\left[ leader \right]\star c}$
+  $status\left[ leader \right]$ is ``ready'' and a $\aseq{\quo{prop},c}$ message has been sent to the leader. The effect of the
+  action is to receive the message, to update $accepted\left[ leader
+  \right]$ to $accepted\left[ leader \right]\star c$, and to
+  broadcast the message $\aseq{\quo{accepted},accepted'\left[ leader
+  \right]}$ to the learners, and to broadcast the message
+  $\aseq{\quo{leader-accept},accepted'\left[ leader \right]}$
   to all the other acceptors.
@@
   which is not the leader, is enabled when $status\left[ a
-        \right]$ is ``ready'' and $a$ can receive a
-        $\aseq{\quo{leader-accept},h}$ message where $h > hist\left[ a
+  \right]$ is ``ready'' and the leader has sent a
+  $\aseq{\quo{leader-accept},h}$ message to $a$, where $h > accepted\left[ a
   \right]$. The effect of the action is to receive the message,
-        to update $hist\left[ a \right]$ to $h$, and to broadcast the
-        message $\aseq{\quo{ack},hist\left[ a \right]\star c}$ to the
+  to update $accepted\left[ a \right]$ to $h$, and to broadcast the
+  message $\aseq{\quo{accepted},hist\left[ a \right]\star c}$ to the
   learners.
-    \item A $RcvAck\left( l \right)$ transition, for a learner $l$, is
-        enabled when $l$ can receive a message $\aseq{\quo{ack},h}$
-        from an acceptor $a$. Its effect is to receive the message and
-        to set $accepted\left[ l \right]\left[ a \right]$ to $h$.
 \item A $Learn\left( l, h \right)$ transition, for a learner $l$
-        and a history $h$, is enabled when $h > Max\left(
-        learned\left[ l \right] \right)$ and there is a learn quorum
-        $Q$ of acceptors such that $h = GLB\left( \aset{accepted\left[
-        l \right]\left[ a \right] : a \in Q }\right)$. Its effect is
-        to insert $h$ in $learned\left[ l \right]$.
+  and a history $h$, is enabled when $h \geq learned\left[ l \right]$ and $h$ is a lower
+  bound of the set $\aset{MaxReceivedFrom\left( a \right) : a \in Acceptors}$, where $MaxReceivedFrom\left( a \right)$ is the maximal accepted history received
+  from the acceptor $a$.  Its effect is to set $learned\left[ l \right]$ to $h$.
 \item A $Stop\left( s \right)$ transition, for a switcher $s$,
-        always enabled, broadcasts a $\aseq{\quo{stop}}$ message to all
-        the acceptors.  This transition models a switcher being
-        instructed by the scheduling policy to abort the current round.
+  enabled when the switcher $s$ has waited for too long without receiving any messages from the acceptors, broadcasts a $\aseq{\quo{stop}, s}$ message to all
+  the acceptors.  
 \item A $Stop\left( a \right)$ transition, for an acceptor $a$,
-        enabled when $a$ can receive a $\aseq{\quo{stop}}$ message
+  enabled when $a$ can receive a $\aseq{\quo{stop},s}$ message
   from a switcher $s$. Its effect is to receive the message, set
   $status\left[ a \right]$ to ``stopped'', and to send the
-        message $\aseq{\quo{stopped},hist\left[ a \right]}$ to the
+  message $\aseq{\quo{stopped},a,hist\left[ a \right]}$ to the
   switcher $s$.
-    \item A $RcvStopped\left( s \right)$ transition, for a switcher
-        $s$, enabled when $s$ can receive a $\aseq{\quo{stopped},h}$
-        message from an acceptor $a$. Its effect is to receive the
-        message and to set $stopped\left[ s \right]\left[ a \right]$
-        to $h$.
-    \item An $Abort\left( s, h \right)$ transition, for a switcher $s$
-        and a history $h$, is enabled when there exists an abort quorum
-        $R$ such that, for every member $a$ of $R$, $stopped\left[ s
-        \right]\left[ a \right]$ is not $false$ and $h \in
-        SafeAborts\left( R \right)$, where $SafeAborts\left( R
-        \right)$ is computed using the array $stopped\left[ a
-        \right]$. Its effect is to insert  $h$ in the set $aborts$. 
+\item An $Abort\left( s, h \right)$ transition, for a switcher $s$ and a history $h$, is enabled when $aborts\left[ s \right] = \quo{none}$ and
+  the switcher has received at least one $\aseq{\quo{stopped}, a, h}$ message. Its effect is to set the variable $aborts\left[ s \right]$ to the
+  history $h$ received in the $\quo{stopped}$ message.
 \end{compactitem}
 
-Thanks to the inter-round interface, Simple-Leader rounds and Broadcast-Decide rounds can be composed in any order.
-To compose a Simple-Leader round $SL_1$ with an Broadcast-Decide round $B_2$, we compose their specifications as before, adding the constraint that the set $aborts_1$ be equal to the set $inits_2$ at all times.
+Thanks to the inter-round interface, which is the same for Simple-Leader rounds and Broadcast-Decide rounds, Simple-Leader rounds and Broadcast-Decide rounds can be repeatedly composed in a sequence.
 

### 2015-05-15 paper/paper.tex
Working on the presentation of CGC

@@
 
-Generalized Lattice Agreement \cite{FalerioETAL12GeneralizedLatticeAgreement} allows more efficient implementations of SMR when
-commands are partitioned into read-only commands and write-only commands.
-Composable Generalized Consensus supports the same optimizations as Generalized Lattice Agreement.
-Notably, read-only commands are treated specially so as to optimize their
-executions (see \cref{sec:readonly}).
+%Generalized Lattice Agreement \cite{FalerioETAL12GeneralizedLatticeAgreement} allows more efficient implementations of SMR when
+%commands are partitioned into read-only commands and write-only commands.
+%Composable Generalized Consensus supports the same optimizations as Generalized Lattice Agreement.
+%Notably, read-only commands are treated specially so as to optimize their
+%executions (see \cref{sec:readonly}).
 
@@
 specification are rather informal, eluding non-essential details.
-
-Our specifications all contain by default a state variable $crashed$ and
- a $Crash\left( x \right)$ transition which is always enabled. A server $x$ faithfully
-executes the algorithm that is assigned to it until its identifier is added to
-the variable $crashed$ by a $Crash\left( x \right)$ transitions.  When the identifier of a
-server is in the set $crashed$, all its actions are disabled.  In an execution,
-a server that is never added to the set $crashed$ is said correct.
 
@@
 
-Before formally introducing our framework, let us see how we can build a modular implementation similar to Fast Paxos by first designing a fast but fragile modular GC algorithm, then combining it unchanged with a more resilient module, obtaining an algorithm similar to Fast Paxos.
+In this section we introduce our framework by example, building a modular Generalized Consensus algorithm inspired from Fast Paxos by first designing a fast but fragile modular algorithm, then combining it unchanged with a more resilient algorithm.
 
@@
 The history $aborts'\left[ s \right]$ is then called an abort history.
-Because the switcher has received a $\aseq{\quo{stopped},a,h}$ message from the acceptor $a$, the acceptor $a$ no longer accepts new proposals, therefore no history greater than $h$ can ever be learned anymore.
-Therefore, an abort history can be seen as a safe starting point for a new round to resume execution.
 
@@
 This process can be repeated as often as needed, with new rounds $B_3$, $B_4$, etc., to try to make progress learning new histories.
+
+Crucially, the composition of any number of rounds implements Generalized Consensus.
+To understand why, observe that in an $Abort\left( s,h \right)$ transition, the switcher $s$ has received a $\aseq{\quo{stopped},a,h}$ message from the acceptor $a$, which means that the acceptor $a$ no longer accepts new proposals and that $h$ is the current value of the $learned\left[ a \right]$ variable.
+Therefore no history greater than $h$ can ever be learned anymore.
+An abort history can thus be seen as a safe starting point for a new round to resume execution.
 
@@
 
-Informally, the composition is correct because abort histories are always safe in the sense that any of the abort histories could be the next learned history.
-However not all can be learned and the initialization mechanism of the rounds ensures that at most one is chosen.
-In the next section we present Composable Generalized Consensus, a specification of the behavior of rounds at the level of proposals, learned histories, init histories, and abort histories.
+Observe that a Simple-Leader round maintains a similar invariant, relating abort histories to learn histories, than the BroadCast-Decide algorithm:
+the learned histories are necessarily smaller than the abort histories in the partial order on histories. In both types of rounds, an abort history is
+a safe starting point for initializing the variables $accepted\left[ a \right]$ of the acceptors. Moreover, both types of rounds guarantee that only a
+single history will be used as a basis for the learners to learn new histories.
+
+In the next section we present Composable Generalized Consensus, a specification of the behavior of rounds at the level of proposals, learned histories, init histories, and abort histories, which allows making precise the informal reasoning above.
 Our composition theorem ensures that any set of rounds individually satisfying Composable Generalized Consensus can be composed in any order to yield an algorithm satisfying Composable Generalized Consensus.
+Therefore we can devise many other types of rounds and compose them unchanged with our two BroadCast-Decide and Simple-Leader rounds.
 
@@
 
-We now consider implementing Generalized Consensus as a sequence of \emph{Composable Generalized Consensus rounds}, where each round makes progress
-ordering requests or aborts its execution and passes the baton to the next round in the sequence. An algorithm composed of a sequence of CGC instances
-is called a \emph{CGC algorithm}.
+In this section we present Composable Generalized Consensus, our specification for composable Generalized Consensus rounds.
 
-On top of the traditional roles of proposer, acceptor, and learner, CGC
-introduces the new role of \emph{switcher} and a new component, the
-\emph{scheduling policy}. In CGC, the assignment of the roles of proposer,
+As described in the preceding section, the servers can have any of the four roles of proposer, acceptor, learner, and switcher. 
+The assignment of the roles of proposer,
 acceptor, learner, and switcher can change in each rounds, allowing for
 reconfiguration.
-In normal circumstances, the proposers, acceptors,
-and learners of a CGC instance work together to implement Generalized Consensus.
-However, if the environment becomes unfavorable to a running CGC instance, the
-scheduling policy can instruct the round to abort its execution and pass the
-baton, using the switchers, to a new round.
 
-A switcher of an aborting round determines, from the state of the
-acceptors, an \emph{abort sequence} that it transfers to one or several
-switchers of the new round.  Upon receiving an abort sequence, a switcher of the
-new round initializes the acceptors of the new round, which can then
-proceed implementing SMR, until the new round itself aborts and the
-process is repeated. 
 We assume that in the first round, which has no previous round, every 
@@
 
+\begin{comment}
 The scheduling policy determines when to abort a CGC instance and which new CGC
@@
 resilience of the service by replacing faulty servers upon round change.
+\end{comment}
 
@@
 sequences transfered to the next round.
-Two consecutive CGC instances are composed by letting the $aborts$ variable of the
+Two consecutive CGC rounds are composed by letting the $aborts$ variable of the
 first round be the $inits$ variable of the second round.
 
-Initially, $crashed$, $proposed$, for every
-learner $l$, $learned\left[ l \right]$, $inits$, and $aborts$ are all empty sets.
-The next-state relation consists of five types of transitions. The
-first three, $Crash\left( x \right)$, $Propose\left( l, c
-\right)$, and $Learn\left( l, cs \right)$, for a learner, proposer, or
-switcher $x$, for a learner $l$, and for a sequence of commands
-$cs$, have the same effect as in Generalized Consensus.
-CGC introduces two new transitions, $Init\left(
-s, cs \right)$ and $Abort\left( s, cs \right)$, for a switcher $s$ and a sequence of
-commands $cs$. The transition $Init\left( s, cs \right)$ adds $cs$ to the set $inits$, and models a switcher receiving the
-abort sequence $cs$ from the preceding round.
-The transition $Abort\left( s, cs \right)$ adds $cs$ to the set $aborts$ and models a switcher transferring the abort sequence $cs$ to a switcher of the next round.
-All the transition are guarded by the following conditions:
+Initially, $crashed$, $proposed$, for every learner $l$, $learned\left[ l \right]$, $inits$, and $aborts$ are all empty sets. The next-state relation
+consists of four types of transitions. The first two, $Propose\left( l, c \right)$, and $Learn\left( l, h \right)$, for a learner $l$, and for a
+history $h$, have the same effect as in Generalized Consensus. CGC introduces two new transitions, $Init\left( s, h \right)$ and $Abort\left( s, h
+\right)$, for a switcher $s$ and a history $h$. The transition $Init\left( s,h \right)$ adds $h$ to the set $inits$, and models a switcher receiving
+the abort sequence $h$ from the preceding round. The transition $Abort\left( s, h \right)$ adds $cs$ to the set $aborts$ and models a switcher
+transferring the abort sequence $cs$ to a switcher of the next round. All the transition are guarded by the following conditions:
 \begin{compactitem}

### 2015-05-15 paper/paper.tex
Formatting a various fixes

@@
 
-\section{The Simple-Leader Algorithm}
+\section{The Leader Algorithm}
 
@@
 Instead, we will now show how to obtain an algorithm with performance characteristics close to those of Fast Paxos, employing a leader to deal with contention, without modifying at all our existing Broadcast-Decide algorithm.
-To do so, we define a new kind of rounds, the Simple-Leader rounds, which uses the same inter-round interface as the Broadcast-Decide algorithm. 
-The inter-round interface will allow use to combine Simple-Leader rounds and Broadcast-Decide rounds in any order, obtaining an algorithm which can adapt its strategy to its environment, selecting the Broadcast-Decide algorithm under low contention and switching to the more expensive Simple-Leader algorithm when there is contention.
+To do so, we define a new kind of rounds, the Leader rounds, which uses the same inter-round interface as the Broadcast-Decide algorithm. 
+The inter-round interface will allow use to combine Leader rounds and Broadcast-Decide rounds in any order, obtaining an algorithm which can adapt its strategy to its environment, selecting the Broadcast-Decide algorithm under low contention and switching to the more expensive Leader algorithm when there is contention.
 
-In the Simple-Leader algorithm, the proposers send their proposals to a distinguished acceptor called the leader. 
+In the Leader algorithm, the proposers send their proposals to a distinguished acceptor called the leader. 
 The leader totally orders the proposals it receives and forwards them to all the acceptors which accept them in the order defined by the leader.
@@
 the set of $\aset{MaxReceivedFrom\left( a \right) : a \in Acceptors}$.
-Thanks to its leader, a Simple-Leader round continues to learn new histories even when proposals conflict because the leader ensures that the accepted histories of the acceptors remain consistent.
+Thanks to its leader, a Leader round continues to learn new histories even when proposals conflict because the leader ensures that the accepted histories of the acceptors remain consistent.
 
-The Simple-Leader algorithm treats all proposals equally and does not take advantage of commutativity.
+The Leader algorithm treats all proposals equally and does not take advantage of commutativity.
 When the leader does not crash, each proposal is processed in three message delays.
@@
 
-To initialize a Simple-Leader round, the switchers send their init histories to the leader, which choose a unique one that it forwards to the acceptors, guaranteeing that all the acceptors are initialized with the same history even when conflicting init histories are sent concurrently.
+To initialize a Leader round, the switchers send their init histories to the leader, which choose a unique one that it forwards to the acceptors, guaranteeing that all the acceptors are initialized with the same history even when conflicting init histories are sent concurrently.
 
-The Simple-Leader algorithm can not learn new histories if its leader, or any other acceptor, crashes.
+The Leader algorithm can not learn new histories if its leader, or any other acceptor, crashes.
 To recover from this deadlock we pass the baton to a new round, as in the Broadcast-Decide algorithm, with a similar same mechanism: a switcher that times-out waiting for new messages from the acceptors broadcast a stop message, gathers the accepted history from at least one of the acceptors, and aborts.
 
-Let us specify more precisely the Simple-Leader algorithm.
+Let us specify more precisely the Leader algorithm.
 
-The specification of Simple-Leader algorithm uses the same variables as the Broadcast-Decide algorithm and the same initial state. 
+The specification of Leader algorithm uses the same variables as the Broadcast-Decide algorithm and the same initial state. 
 %that is the arrays of variables    
@@
 
-Thanks to the inter-round interface, which is the same for Simple-Leader rounds and Broadcast-Decide rounds, Simple-Leader rounds and Broadcast-Decide rounds can be repeatedly composed in a sequence.
+Thanks to the inter-round interface, which is the same for Leader rounds and Broadcast-Decide rounds, Leader rounds and Broadcast-Decide rounds can be repeatedly composed in a sequence.
 
-Observe that a Simple-Leader round maintains a similar invariant, relating abort histories to learn histories, than the BroadCast-Decide algorithm:
+Observe that a Leader round maintains a similar invariant, relating abort histories to learn histories, than the BroadCast-Decide algorithm:
 the learned histories are necessarily smaller than the abort histories in the partial order on histories. In both types of rounds, an abort history is
@@
 single history will be used as a basis for the learners to learn new histories.
+These common invariants make the two kinds of rounds compatible, so that we can compose them and obtain an algorithm implementing Generalized Consensus.
 
@@
 Our composition theorem ensures that any set of rounds individually satisfying Composable Generalized Consensus can be composed in any order to yield an algorithm satisfying Composable Generalized Consensus.
-Therefore we can devise many other types of rounds and compose them unchanged with our two BroadCast-Decide and Simple-Leader rounds.
+Therefore we can devise many other types of rounds and compose them unchanged with our two BroadCast-Decide and Leader rounds.
 
@@
 As described in the preceding section, the servers can have any of the four roles of proposer, acceptor, learner, and switcher. 
-The assignment of the roles of proposer,
-acceptor, learner, and switcher can change in each rounds, allowing for
+The assignment of the roles of proposer, acceptor, learner, and switcher can change in each rounds, allowing for
 reconfiguration.
@@
 We assume that in the first round, which has no previous round, every 
-switchers receives the empty abort sequence.
+switchers receives the empty abort history.
 
@@
 and, for every learner $l$, $learned\left[ l \right]$.  Additionally,
-we use the state variables $inits$, containing the abort sequences
+we use the state variables $inits$, containing the init histories 
 received from the previous round, and $aborts$, containing the abort
-sequences transfered to the next round.
+histories transfered to the next round.
 Two consecutive CGC rounds are composed by letting the $aborts$ variable of the
-first round be the $inits$ variable of the second round.
+first round be equal to the $inits$ variable of the second round.
 
-Initially, $crashed$, $proposed$, for every learner $l$, $learned\left[ l \right]$, $inits$, and $aborts$ are all empty sets. The next-state relation
+Initially, $proposed$, for every learner $l$, $learned\left[ l \right]$, $inits$, and $aborts$ are all empty sets. The next-state relation
 consists of four types of transitions. The first two, $Propose\left( l, c \right)$, and $Learn\left( l, h \right)$, for a learner $l$, and for a
@@
 \right)$, for a switcher $s$ and a history $h$. The transition $Init\left( s,h \right)$ adds $h$ to the set $inits$, and models a switcher receiving
-the abort sequence $h$ from the preceding round. The transition $Abort\left( s, h \right)$ adds $cs$ to the set $aborts$ and models a switcher
-transferring the abort sequence $cs$ to a switcher of the next round. All the transition are guarded by the following conditions:
+the abort history $h$ from the preceding round. The transition $Abort\left( s, h \right)$ adds $cs$ to the set $aborts$ and models a switcher
+transferring the abort history $cs$ to a switcher of the next round. All the transition are guarded by the following conditions:
 \begin{compactitem}
@@
     \item[\textbf{Initialization}:] If $inits$ is empty, then no
-        sequence can be learned and no switcher can abort.
+        history can be learned and no switcher can abort.
     \item[\textbf{Irrevocability}:] In the case of a $Learn\left( l,h
@@
         extension of $learned\left[ l \right]$.
-    \item[\textbf{Safe Abort}:] Every abort sequence is an extension in the 
+    \item[\textbf{Safe Abort}:] Every abort history is an extension in the 
         partial order $\leq$ of a learned history.
@@
 either the client gets a response from the current round, or
-the next round is initialized with an abort sequence and the
+the next round is initialized with an abort history and the
 command can be resubmitted to the next round.
@@
 Moreover, we assume that a command, which is proposed to a round and that
-is not learned before the first abort sequence is transfered to the
+is not learned before the first abort history is transfered to the
 next round, is proposed anew to the next round.
@@
 
+\begin{comment}
 \begin{thm}[Composition Theorem]
     \label{thm:comp}
-    In every execution of a succession of CGC instances, the sequence of valuations
+    In every execution of a succession of CGC instances, the history of valuations
     of the variable $proposed$ and, for every learner $l$, of the variables
@@
     Consider two rounds numbered $i$ and $i+1$. In every execution of a
-    succession of CGC instances, the sequence of valuations of the variables $inits
+    succession of CGC instances, the history of valuations of the variables $inits
     = inits_i$, $propose = proposed_i \cup propose_{i+1}$, for every learner
@@
   executable form that presented in this section. Moreover,
-  instead of sequences of commands, the Isabelle/HOL specification uses the more
+  instead of history of commands, the Isabelle/HOL specification uses the more
   general notion of command history, presented in the next subsection.
@@
 
+\end{comment}
+
 \section{Fast and Conservative Rounds}
@@
 
-Most existing SMR algorithms can easily be obtained as the
-composition of one or more CGC instances.
-
-The FLP impossibility result \cite{FischerLynchPaterson83ImpossibilityDistributedConsensusOneFaultyProcess}
-implies that no algorithm can implement SMR to the letter in practice.
-Instead, SMR algorithms guarantee safety (i.e. replica execute the
-same sequence of commands) but may not be able to make progress (i.e.
-order new commands). To make progress as likely as possible, most SMR
-algorithms are structured in consecutive rounds (also called ballots,
-views, or epochs). In each round, the algorithm tries to make progress
-but cannot be guaranteed to succeed. When the algorithm detects that
-progress is no more possible, a new round is started.  The new round
-is initialized so as to preserve safety, usually by one of the
-acceptors. For example, in Paxos, the leader of a newly started ballot
-numbered $i$ determines a value that is \emph{safe at $i$} and
-proposes it as first value to the other acceptors of the ballot $i$.
-In existing SMR algorithms, new rounds are initialized ad-hoc, which
-prevents mixing rounds of two different algorithms.  Instead,
-CGC is an abstraction of rounds which establishes a common interface and contract for switching from
-one round to the next, enabling switching from any CGC instance to any other.
+Most existing SMR algorithms are structured in rounds that can be modified to obtain several CGC rounds.
+However, checking that the obtained rounds are correct may be quite difficult in the case of algorithms using complex quorum
+systems. In particular, safely aborting can become complex.
 
@@
 or more CGC instances, we propose two specifications, \emph{conservative CGC
-rounds} and
-\emph{fast CGC instances}, which can easily be concretized to obtain executable rounds.
+rounds} and \emph{fast CGC rounds}, which can easily be concretized to obtain executable rounds.
 We show that CGC versions of Classic Paxos, Multi-Coordinated Paxos, Chain
@@
 The next-state relations of conservative and fast rounds are composed
-of the following types of transitions (and the default $Crash\left( x
-\right)$ transition).
+of the following types of transitions:
 \begin{compactitem}

### 2015-05-15 paper/intro.tex
Work on the intro, emphasis on the two intermediate specs.

@@
 
-The key novelty of our framework lies in its practical applicability, allowing
-the execution of commutative commands without synchronization and reducing the
-replication overhead from $2*k+1$ replicas to $k+1$ replicas, to tolerate $k$
-replica crashes, using a reliable external reconfiguration master.
+The key novelty of our framework lies in its practical applicability.
+Most SMR algorithms are variations on a basic structure consisting of one of two kinds of 
+quorum systems \cite{GuerraouiVukolic10RefinedQuorumSystems}. 
+The first kind, that we call resilient quorums, are used in leader-based algorithms, while the second
+kind of quorums, that we call fast quorums, are used in leader-less algorithms and can order commuting commands without synchronization.
+We provide two specifications, at an intermediate level of abstraction, that capture the essence of composing algorithms using these two kinds of quorums.
+Our two specifications allow to easily check that new modules based on resilient or fast quorums are correct, covering a large spectrum of practical algorithms.
 

### 2015-05-15 paper/abstract3.tex
work on the abstract
Modified the abstract

@@
 
-The key novelty of our framework is its practical applicability, with state
-of the art optimization and reconfiguration techniques being built-in:
-tolerating k faults with k + 1 replicas, executing commutative commands
-without synchronization, optimizing the execution of read-only commands, and
-checkpointing.
+The key novelty of our framework is its practical applicability. 
+Our framework is based on Generalized Paxos, therefore providing buit-in support
+for algorithms optimizing the execution of commuting commands.
+Moreover, we propose two specifications 
+capturing the essence of modular SMR algorithms based on quorum systems.
+We demonstrate that concrete
+algorithms can easily be obtained by refinement from our abstract specifications, as
+evidenced by TLA+ specifications ready to be checked by the TLC model checker.
 
-To demonstrate our framework we present two abstract composable algorithms
-characterizing two popular types of quorum systems. We demonstrate that concrete
-algorithms can easily be obtained by refinement from the abstract algorithms, as
-evidenced by TLA+ specifications ready to be checked by the TLC model checker.
 }

### 2015-05-20 paper/abstract3.tex
Submitted version

@@
 
-The key novelty of our framework is its practical applicability. 
-Our framework is based on Generalized Paxos, therefore providing buit-in support
-for algorithms optimizing the execution of commuting commands.
-Moreover, we propose two specifications 
-capturing the essence of modular SMR algorithms based on quorum systems.
-We demonstrate that concrete
-algorithms can easily be obtained by refinement from our abstract specifications, as
-evidenced by TLA+ specifications ready to be checked by the TLC model checker.
-
+The key novelty of our framework is its practical applicability: we show that
+a large class of SMR algorithms can be expressed in our framework by proposing
+two specifications capturing the essence of modular SMR algorithms based
+on Generalized Consensus and using quorum systems for synchronization. We
+demonstrate that concrete algorithms can easily be obtained by refinement from
+our abstract specifications, as evidenced by TLA+ specifications ready to be
+checked by the TLC model checker.
 }

### 2015-05-20 paper/intro.tex
Submitted version

@@
 
-The key novelty of our framework lies in its practical applicability.
-Most SMR algorithms are variations on a basic structure consisting of one of two kinds of 
+The key novelty of our framework lies in its practical applicability:
+\begin{itemize}
+  \item Our framework is based on Generalized Consensus, a generalization of the consensus problem 
+    that allows optimizing the execution of commuting commands.
+  \item Many SMR algorithms are variations on a basic structure consisting of one of two kinds of 
     quorum systems \cite{GuerraouiVukolic10RefinedQuorumSystems}. 
@@
     Our two specifications allow to easily check that new modules based on resilient or fast quorums are correct, covering a large spectrum of practical algorithms. 
+\end{itemize}
 
@@
 reduce the latency of processing one command to two message delays instead of
-three. As we will show, it also enables the processing of read-only commands
-in two message delays, although read-only commands do not commute with write
-commands.
+three. 
+%As we will see, fast execution of commutative requests also enables the processing of read-only commands
+%in two message delays, although read-only commands do not commute with write
+%commands.
 
-In practice, a SMR algorithm must support fast reconfiguration to replace faulty
-components, without which the long-term stability of a system is inevitably
-compromised. Our framework abstracts over how the reconfiguration decisions
-are made, providing an interface allowing to plug an external reconfiguration
-master implementing the reconfiguration policy of the system. The master decides
-which GC module to run next and which replica servers will participate, removing
-failed replicas and adding new ones. As we will explain, this enables tolerating
-$k$ failures with $k+1$ replicas, an important improvement over the traditional
-$2k+1$ needed replicas.
+In practice, a SMR algorithm must support fast reconfiguration to replace
+faulty components, without which the long-term stability of a system is
+inevitably compromised. Our framework abstracts over how the reconfiguration
+decisions are made, providing an interface allowing to plug in an external
+reconfiguration master implementing the reconfiguration policy of the
+system. The master decides which GC module to run next and which replica
+servers will participate, removing failed replicas and adding new ones. The
+reconfiguration master enables tolerating $k$ failures with $k+1$ replicas
+\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}, an important
+improvement over the traditional $2k+1$ needed replicas.
 
@@
 composable versions of the generalized variations of popular consensus
-algorithms: classic Paxos, Fast Paxos, Chain Replication, and multi-coordinated
-Paxos. We have checked the correctness of these algorithms with the TLC
+algorithms: Classic Paxos, Fast Paxos, Chain Replication, and multi-coordinated
+Paxos. Thanks to our specifications of algorithms based on quorum-systems, 
+we have checked the correctness of our version of Fast Paxos and Chain Replication with the TLC
 model-checker and we provide all the specifications needed to replicate these
@@
 certifying the correctness of the composition theorem at the deepest level.
+The TLA+ and Isabelle/HOL development can be found at \url{http://losa.fr/cgc}.
 
 At the heart of our framework lies an abstraction, called Composable Generalized
-Consensus, of the familiar notion of round (also called epoch, ballot, or
+Consensus (CGC), of the familiar notion of round (also called epoch, ballot, or
 view in the literature) used in most SMR algorithms. As an example, let us
@@
 implementations and making all types of rounds compatible with each other.
+
 Moreover, our interface for rounds enables a crucial optimization seen in

### 2015-05-20 paper/paper.tex
Submitted version

@@
 
-\input{roadmap}
+%\input{roadmap}
 
@@
 
+\begin{comment}
 SMR is often implemented using a sequence of instances of Consensus.
@@
 Our Composable Generalized Consensus can also use C-structs, with the same advantages
-as Generalized Consensus. For simplicity, the presentation in this
-paper uses the more concrete notion of \emph{command history} (see
-\cref{sec:histories}).
+as Generalized Consensus. 
+\end{comment}
 
-%Generalized Lattice Agreement \cite{FalerioETAL12GeneralizedLatticeAgreement} allows more efficient implementations of SMR when
-%commands are partitioned into read-only commands and write-only commands.
-%Composable Generalized Consensus supports the same optimizations as Generalized Lattice Agreement.
-%Notably, read-only commands are treated specially so as to optimize their
-%executions (see \cref{sec:readonly}).
+Composable Generalized Consensus combines the ideas of the Abstract Framework \cite{GuerraouiETAL10Next700BftProtocols}, Generalized Conensus  \cite{Lamport05GeneralizeConsensus}, and Vertical Paxos \cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication} in a
+unified abstraction.
 
-Vertical Paxos algorithms 
-\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication} are SMR algorithms
-whose set of replica servers can be reconfigured without stopping to process new
-commands. Crashed servers can therefore be replaced by new servers, ensuring the
-long-term availability of a replicated service.
-Vertical Paxos algorithms use an external reconfiguration master, itself
-implemented as a replicated state machine, which determines the servers that should
-be removed from or added to the configuration.
-CGC shares some aspects of Vertical Paxos algorithms: the task of the
-reconfiguration master is similar to the task of the scheduling policy
-component of CGC (see \cref{sec:aca}).
-However, in CGC, changing round can be used not only to replace crashed servers
-but also to change the algorithm that the servers are running. 
-Changing algorithm is not possible in Vertical Paxos.
-In both Vertical Paxos and CGC, a reliable reconfiguration master, resp. a
-reliable scheduling policy, allows tolerating $f$ crashes with $f+1$ servers.
+Generalized Consensus is an agreement problem relaxing the ordering constraints
+imposed on commuting commands and Generalized Paxos is an algorithm, similar
+in structure to Fast Paxos, which takes advantage of the relaxed ordering
+constraint to process commuting commands in two message delays, when three are
+needed in the Classic Paxos algorithm.
 
-The Abstract Framework \cite{GuerraouiETAL10Next700BftProtocols} proposes to
+Vertical Paxos algorithms are SMR algorithms whose set of replica servers can
+be reconfigured without stopping to process new commands. Crashed servers can
+therefore be replaced by new servers, ensuring the long-term availability of a
+replicated service. Vertical Paxos algorithms use an external reconfiguration
+master, itself implemented as a replicated state machine, which determines
+the servers that should be removed from or added to the configuration. The
+role of the reconfiguration master is similar to the role of the scheduling
+policy in our framework. In contrast to our work, Vertical Paxos does not define
+an interface and correctness condition for encapsulating the reconfigurable
+instances of SMR algorithms and therefore does not allow the reuse of existing
+algorithms and the incremental development of SMR optimizations.
+
+Our work builds on the Abstract Framework , which proposes to
 build Byzantine Fault-Tolerant algorithms as the succession of abortable rounds
@@
 correctness properties. By construction, the composition of any number of
-Abstract instances is a correct SMR implementation. However, the Abstract
-Framework uses totally ordered sequence of commands, making difficult the
-optimization of the execution of commuting and read-only commands that are
-possible with CGC.  Moreover, the Abstract Framework does not identify the
-crucial role of the scheduling policy for the long-term resilience of a service.
-Composable Generalized Consensus combines the ideas of the Abstract Framework,
-Generalized Paxos, Generalized Lattice Agreement, and Vertical Paxos in a
-unified abstraction.
+Abstract instances is a correct SMR implementation. The Abstract
+Framework is similar to our framework but uses totally ordered sequence of commands, making it difficult apply it to 
+practical algorithms that optimize the execution of commuting and read-only commands.
 
@@
 Abstract Framework and mechanically proving its properties with the
-Isabelle/HOL interactive
-proofs assistant.  Like the Abstract Framework, the
+Isabelle/HOL interactive proofs assistant.  Like the Abstract Framework, the
 Speculative Linearizability framework is based on totally ordered
@@
 conceptually simpler and subsumes the Speculative Linearizability
-property.  Like Speculative Linearizability, Composable Generalized Consensus
-and its main property, the composition theorem (\cref{thm:comp}), are
-formalized in Isabelle/HOL (see \cref{sec:isaproofs}).
+property. Composable Generalized Consensus
+and its main property, the composition theorem, are also formalized in Isabelle/HOL.
 
+\begin{comment}
 Composable Generalized Consensus is the most complete framework to date for building
@@
 possible before.
+\end{comment}
 
@@
 an asynchronous network. Servers are not necessarily sequential processes.
-We describe a behavior of the system using an infinite sequence of states,
+We describe a behavior of the system using a sequence of states,
 where a state is a function from variable names to values (i.e. a valuation of
 variables). We specify sets of allowed behaviors using an initial predicate
-and a next-state relation expressed using unprimed and primed variables in the
+and a next-state relation expressed using unprimed variables and primed variables (next-state variables) in the
 style of TLA+ \cite{Lamport02SpecifyingSystems}. However, for succinctness, our
@@
 The goal of SMR is to provide a reliable implementation of the computing service despite server crashes.
-To do so, an SMR algorithm replicates the execution of the service on several servers. 
-In a traditional SMR algorithm, the servers can be seen has playing one or
+To do so, a SMR algorithm replicates the execution of the service on several servers. 
+Traditionally, the servers play one or
 several of three roles: \emph{proposers}, which propose new commands to execute,
@@
 \emph{learners}, which learn about the growing sequence of commands to
-execute \cite{lamport2001paxos} by querying the acceptors. Often, but not necessarily, every replica server plays the
+execute by querying the acceptors \cite{lamport2001paxos}. Often, but not necessarily, every replica server plays the
 three roles of proposer, acceptor, and learner: clients of the service
@@
 
-A command history, abbreviated history, is a data structure that
+A history is a data structure that
 represents a set of sequences which are the same modulo the reordering
@@
 
-All the sequences represented by a given history contain the same commands
+By definition, all the sequences represented by a given history contain the same commands
 and determine the same outputs for each command. Therefore, to implement a
 replicated service, it is sufficient that learners learn histories instead of
-sequences. For a thourough discussion of command histories we refer the reader
-to \cite{Lamport05GeneralizeConsensus}. Our Isabelle/HOL formalization \cite{}
-abstracts over the concrete representation of histories, assuming only their
+sequences. For a thorough discussion of command histories we refer the reader
+to \cite{Lamport05GeneralizeConsensus}. 
+Our Isabelle/HOL formalization abstracts over the concrete representation of histories, assuming only their
 necessary properties. Command histories were first introduced in trace theory
@@
 
-Command histories can be built from the empty history, noted $\bot$, using the $\star$
+Command histories are built from the empty history, noted $\bot$, using the $\star$
 operator, which appends a sequence of commands $cs$ to a history $h$, obtaining
-the history $h \star cs$. We can then define define an ordering relation among
+the history $h \star cs$. We define a partial order relation among
 histories: $h_1 \leq h_2$ when there exists a sequence $cs$ such that $h_2 = h_1
@@
 that a set of histories $H$ is compatible when all the histories in $H$ have a
-common upper bound $h$ such that $h$ can be built using only commands found in
+common upper bound $h$ such that $h$ can be built using only commands contained in
 the histories in $H$.
@@
 Generalized Consensus precisely specifies the allowed behaviors of the roles of \emph{proposers}
-and \emph{learners}, abstracting over the behavior of acceptors.
+and \emph{learners}, abstracting over the behavior of acceptors. Acceptors are used to implement Generalized Consensus but do not appear in its interface.
 
-We specify Generalized Consensus using the variables $crashed$, the variable $proposed$, containing the proposed commands,
-and, for every learner $l$,  a variable $learned\left[ l \right]$, holding a history. Initially, $crashed$ and $proposed$ are empty sets and, for every learner $l$,
-$learned\left[ l \right]$ is the empty history. 
-Remember that any transition involving a server $x$ is always disabled when $x\in crashed$.
+We specify Generalized Consensus using the variables $proposed$, containing the proposed commands so far,
+and, for every learner $l$,  a variable $learned\left[ l \right]$, a set containing the histories learned by $l$ so far. Initially, $proposed = \aset{}$ and, for every learner $l$,
+$learned\left[ l \right] = \aset{}$.
 
-The next-state relation consists of three types of transitions:
-$Crash\left( x \right)$ transitions; $Propose\left( l, c
+The next-state relation consists of two types of transitions:
+$Propose\left( l, c
 \right)$ transitions, for a learner $l$ and a command $c$, in which
@@
 transitions, for a learner $l$ and a history $h$, in
-which $learned\left[ l \right]$ is update to $h$. 
+which $h$ is inserted in the set $learned\left[ l \right]$. 
 A $Learn\left( l, h \right)$ transition is guarded by the following conditions:
 \begin{compactitem}
-    \item[\textbf{Agreement}:] The set of learned histories after the transition, $\aset{learned'\left[ l
+    \item[\textbf{Agreement}:] The set of learned histories after the transition, $\aset{\bigcup learned'\left[ l
       \right] : l \in Learners}$, is compatible.
@@
     \item[\textbf{Irrevocability}:] The new learned history
-      $learned'\left[ l \right]$ is an extension of $learned\left[ l
-        \right]$. 
+      $h$ is an extension of every history in $learned\left[ l \right]$. 
         %\\ $cs > Max\left(learned\left[ l \right]\right)$.
@@
 
-When clients of a replicated service submit their commands to the proposers and learners compute the response to give to clients using their local $learned\left[ l \right]$ history, the three properties above imply that the clients see a \emph{linearizable} execution
+When clients of a replicated service submit their commands to the proposers and learners compute the response to give to clients using the maximum over the histories in $learned\left[ l \right]$, the three properties above imply that the clients see a \emph{linearizable} execution
 \cite{HerlihyWing90LinearizabilityCorrectnessConditionConcurrentObjects}.
@@
 
-In this section we introduce our framework by example, building a modular Generalized Consensus algorithm inspired from Fast Paxos by first designing a fast but fragile modular algorithm, then combining it unchanged with a more resilient algorithm.
+In this section we introduce our framework by example, building a modular Generalized Consensus algorithm inspired from Fast Paxos and Generalized Paxos by first designing a fast but fragile modular algorithm, then combining it unchanged with a more resilient algorithm.
 
@@
 
-In our first algorithm, Broadcast-Decide, learners each maintain a variable $hist\left[ l \right]$ which contain the currently accepted history of learner $l$.
+In our first algorithm, Broadcast-Decide, learners each maintain a variable $accepted\left[ l \right]$ which contains the currently accepted history of learner $l$.
 For every learner $l$, $accepted\left[ l \right]$ is initialized to the empty history.
@@
 
-Each acceptor $l$ maintains a variable $learned\left[ l \right]$, initially the empty history, containing its last learned history. A learner $l$
-learns a new history $h$ when $h \geq learned\left[ l \right]$ and $h$ is a lower bound of the set consisting of the maximal accepted history received
+Each learner $l$ maintains a variable $learned\left[ l \right]$, initially the empty history, containing its last learned history. A learner $l$
+learns a new history $h$ when $h \geq learned\left[ l \right]$ and $h$ is a lower bound of the set consisting of the largest history received
 from each acceptor.
@@
 
-The algorithm Broadcast-Decide cannot learn new histories if two acceptors $a_1$ and $a_2$ have two histories which are not compatible because, since an acceptor can only append new proposals to its history $accepted\left[ l \right]$, all the possible future lower bounds of the set $\aset{MaxReceivedFrom\left( a \right) : a \in Acceptors}$ are bounded above by $GLB\left( \aset{accepted\left[ a_1 \right], accepted\left[ a_2 \right]} \right)$.
+The algorithm Broadcast-Decide satisfies the specification of Generalized Consensus, however it cannot learn new histories if two acceptors $a_1$ and $a_2$ have two histories which are not compatible: since an acceptor can only append new proposals to its accepted history, all the possible future lower bounds of the set $\aset{MaxReceivedFrom\left( a \right) : a \in Acceptors}$ are bounded above by $GLB\left( \aset{accepted\left[ a_1 \right], accepted\left[ a_2 \right]} \right)$.
 This situation occurs when non-commuting commands contend: for example, when two different proposers send two commands $c_1$ and $c_2$ that do not commute, the acceptors $a_1$ accepts $c_1$ before $c_2$, and the acceptor $a_2 \neq a_1$ accepts $c_2$ before $c_1$.
@@
 To overcome this deadlock, we will make some modifications to the algorithm that will allow us to start a fresh instance $B_2$, called a round, of the Broadcast-Decide algorithm.
-To allow switching to a fresh new instance of the Broadcast-Decide algorithm, we introduce the new role of switcher, that any server can play additionally from being a proposer, acceptor, or learner.
+To allow switching to a fresh new round of the Broadcast-Decide algorithm, we introduce the new role of switcher, that any server can play additionally from being a proposer, acceptor, or learner.
 A switcher will extract an accepted history from the first round and try to initialize all the acceptors of the second round with the extracted history.
@@
 
-We write $B_1$ for the first round of the Broadcast-Decide algorithm.
-The rounds $B_1$ and $B_2$ each have their own copies of the arrays $accepted$ and $learned$, noted $accepted_1$, $accepted_2$, $learned_1$, and $learned_2$ when the round is not clear from the context.
+The first round $B_1$ and second round $B_2$ each have their own copies of the arrays $accepted$ and $learned$, noted $accepted_1$, $accepted_2$, $learned_1$, and $learned_2$, when the round is not clear from the context.
 We assume that network messages are tagged with a round number and that a message tagged with $i\in\aset{1,2}$ can only be received in $B_i$, creating two isolated logical networks.
@@
 We modify the Broadcast-Decide algorithm by introducing new transitions and modifying the $Accept\left( a \right)$ transition as follows.
-We add the new variables $status\left[ a \right]$ for every acceptor $a$, which are initialized to $\quo{idle}$.
+We add the new variables $status\left[ a \right]$ for every acceptor $a$, which are initialized to $\quo{idle}$, meaning that an acceptor has not been initialized.
 The purpose of $status\left[ a \right]$ is to prevent the acceptor $a$ from accepting commands until its status changes from $\quo{idle}$ to $\quo{ready}$ and to stop it when its status goes to $\quo{stopped}$.
-We also add the new arrays of variables $inits\left[ s \right]$ and $aborts\left[ s \right]$, for every switcher $s$, initialized to the empty history in $B_1$ and to $\quo{none}$ in $B_2$.
+We also add the new arrays of variables $inits\left[ s \right]$ and $aborts\left[ s \right]$, for every switcher $s$.
+The variables $inits\left[ s \right]$ are initialized to the empty history in $B_1$ and to $\quo{none}$ in $B_2$, while the variables $aborts\left[ s \right]$ are initialized to $\quo{none}$ in both $B_1$ and $B_2$.
 The purpose of the arrays of variables $aborts$ and $inits$ is to transfer state between $B_1$ and $B_2$.
 The arrays of variables $aborts$ and $inits$ form what we call the \emph{inter-round} interface.
-As for the arrays $accepted$ and $learned$, the rounds $B_1$ and $B_2$ each have their own copy of the arrays of variables $status$, $inits$, and $aborts$, and we adopt add subscripts denoting the round when the round is not clear from the context.
+The rounds $B_1$ and $B_2$ each have their own copy of the arrays of additional variables $status$, $inits$, and $aborts$, and we add subscripts denoting the round when the round is not clear from the context.
 
@@
 \item We modify the $Accept\left( a \right)$ transition by enabling it only if $status\left[ a \right] = \quo{ready}$ and by having the acceptor $a$ additionally broadcast its new accepted value to the switchers.
-Switchers detect that the algorithm is deadlocked when they have received two messages $\aseq{\quo{accepted}, a_1, h_1}$ and $\aseq{\quo{accepted}, a_2, h_2}$ such that $h_1$ and $h_2$ are incompatible.
+Switchers detect that the algorithm is deadlocked when they have received two messages $\aseq{\quo{accepted}, a_1, h_1}$ and $\aseq{\quo{accepted}, a_2, h_2}$ where  $h_1$ and $h_2$ are incompatible.
 
-\item The $Stop\left( s \right)$ transition, for a switcher $s$, is enabled when the switcher has detected a deadlock or has waited for too long without receiving any message from the acceptors.
+\item The $Stop\left( s \right)$ transition, for a switcher $s$, is enabled when the switcher has detected a deadlock or a crash (for example using a timeout).
 Its effect is to broadcast a $\aseq{\quo{stop}, s}$ message to the acceptors.
@@
 Its effect is to set the variable $aborts\left[ s \right]$ to the history $h$ received in the $\quo{stopped}$ message.
-The history $aborts'\left[ s \right]$ is then called an abort history.
+The history $aborts'\left[ s \right]$ is then called an \emph{abort history}.
 
 \item The $Init\left( s,h \right)$ transition, for a switcher $s$, is enabled when $inits\left[ s \right]=\quo{none}$ and $h$ is a history, and its effect is to set $inits\left[ s \right]$ to $h$, and to broadcast the message $\aseq{\quo{init}, h}$ to the acceptors.
-The history $inits\left[ s \right]$ is then called an init history.
+The history $inits\left[ s \right]$ is then called an \emph{init history}.
 As described below, in $B_2$ composed with $B_1$, this history is constrained to be an abort history of $B_1$.
 
- \item The $WakeUp\left( a \right)$ transition, for an acceptor $a$, is enabled when $status\left[ a \right]$ is $\quo{idle}$ and there is at least one switcher $s$ such that $inits\left[ s \right] \neq \quo{none}$.
-Its effect is to set set $status\left[ a \right]$ to $\quo{ready}$ and $accepted\left[ a \right]$ to $inits\left[ s \right]$.
+ \item The $WakeUp\left( a \right)$ transition, for an acceptor $a$, is enabled when $status\left[ a \right]$ is $\quo{idle}$ and $a$ can receive a $\aseq{\quo{init}, h}$ message from a switcher.
+Its effect is to set set $status\left[ a \right]$ to $\quo{ready}$ and $accepted\left[ a \right]$ to $h$.
 
@@
 \end{itemize}
-The composition models two independent rounds being linked only by the switchers passing the abort histories from $B_1$ to $B_2$, where they become init histories: switchers gather abort histories in $B_1$, then they broadcast them as init histories to the acceptors in $B_2$.
+The composition models two independent rounds being linked only by the switchers passing the abort histories of $B_1$ to $B_2$, where they become init histories: switchers gather abort histories in $B_1$ thanks to the $Stop$ transitions, then they broadcast them as init histories to the acceptors in $B_2$.
 The acceptors then initialize their state to a received init history in the $WakeUp\left( a \right)$ transition.
-This process can be repeated as often as needed, with new rounds $B_3$, $B_4$, etc., to try to make progress learning new histories.
+This process can be repeated as often as needed to switch to new rounds $B_3$, $B_4$, $B_5$, etc., each time a round cannot learn new histories anymore.
 
@@
 To understand why, observe that in an $Abort\left( s,h \right)$ transition, the switcher $s$ has received a $\aseq{\quo{stopped},a,h}$ message from the acceptor $a$, which means that the acceptor $a$ no longer accepts new proposals and that $h$ is the current value of the $learned\left[ a \right]$ variable.
-Therefore no history greater than $h$ can ever be learned anymore.
-An abort history can thus be seen as a safe starting point for a new round to resume execution.
+Therefore no history greater than $h$ can ever be learned anymore in this round.
+An abort history is thus a safe
+starting point for a new round to resume execution.
+Since a new round of the Broadcast-Decide algorithm guarantees that a single init history will be used as a prefix a new learned histories, the properties of Generalized Consensus are preserved.
 
-The Broadcast-Decide algorithm can no more learn new histories whenever one acceptor crashes because the learners wait for the lower bound of the accepted histories of all acceptors to increase.
+The Broadcast-Decide algorithm can no more learn new histories whenever one acceptor crashes because the crashed acceptor prevents the lower bound of the accepted histories to ever increase.
 However, the round-change mechanism is very resilient to faults and can abort anyway thanks to the timeout triggering the $Stop\left( s \right)$ transition: a single live acceptor with a single live switcher can abort the execution and change round.
@@
 
-\section{The Leader Algorithm}
+\subsection{The Leader Algorithm}
 
-The Broadcast-Decide algorithm can process a proposal in two message delays when there is no contention on non-commuting commands. However, if there is
-repeated contention on non-commuting commands, the algorithm may not make progress despite changing round, as the acceptors of the each new round
+The Broadcast-Decide algorithm can process a proposal in two message delays when there are no conflicts, i.e.
+no contention on non-commuting commands.
+However, if there are repeated conflicts, the algorithm may not make progress despite changing round, as the acceptors of the each new round
 may be initialized with incompatible histories. The Fast Paxos algorithm \cite{Lamport06FastPaxos} is able to learn new histories under high contention by
@@
 To do so, we define a new kind of rounds, the Leader rounds, which uses the same inter-round interface as the Broadcast-Decide algorithm. 
-The inter-round interface will allow use to combine Leader rounds and Broadcast-Decide rounds in any order, obtaining an algorithm which can adapt its strategy to its environment, selecting the Broadcast-Decide algorithm under low contention and switching to the more expensive Leader algorithm when there is contention.
+The inter-round interface will allow use to combine Leader rounds and Broadcast-Decide rounds in any order, obtaining an algorithm which can adapt its strategy to its environment, selecting the Broadcast-Decide algorithm under low contention and switching to the slower Leader algorithm when there is contention.
 
@@
 As in the Broadcast-Decide algorithm, acceptors broadcast their accepted histories to the learners, which learn a new history when it is a lower bound of 
-the set of $\aset{MaxReceivedFrom\left( a \right) : a \in Acceptors}$.
+the set $\aset{MaxReceivedFrom\left( a \right) : a \in Acceptors}$.
 Thanks to its leader, a Leader round continues to learn new histories even when proposals conflict because the leader ensures that the accepted histories of the acceptors remain consistent.
@@
 The Leader algorithm treats all proposals equally and does not take advantage of commutativity.
-When the leader does not crash, each proposal is processed in three message delays.
-In comparison, in the absence of failures, the Broadcast-Decide algorithm processes non-conflicting proposals in two message delays but can take arbitrarily long to process conflicting proposals.
+When there are no server crashes, each proposal is processed in three message delays.
+In comparison, in the absence of crashes, the Broadcast-Decide algorithm processes non-conflicting proposals in two message delays but can take arbitrarily long to process conflicting proposals.
 
-To initialize a Leader round, the switchers send their init histories to the leader, which choose a unique one that it forwards to the acceptors, guaranteeing that all the acceptors are initialized with the same history even when conflicting init histories are sent concurrently.
+To initialize a Leader round, the switchers send their init histories to the leader, which chooses a unique one that it forwards to the acceptors, guaranteeing that all the acceptors are initialized with the same history even when conflicting init histories are sent concurrently.
 
 The Leader algorithm can not learn new histories if its leader, or any other acceptor, crashes.
-To recover from this deadlock we pass the baton to a new round, as in the Broadcast-Decide algorithm, with a similar same mechanism: a switcher that times-out waiting for new messages from the acceptors broadcast a stop message, gathers the accepted history from at least one of the acceptors, and aborts.
+To recover from this deadlock we pass the baton to a new round, as in the Broadcast-Decide algorithm, with a similar mechanism: a switcher that times-out waiting for new messages from the acceptors broadcasts a stop message, waits for receiving the accepted history of an acceptor, and finally aborts using the received history as abort history.
 
 Let us specify more precisely the Leader algorithm.
-
-The specification of Leader algorithm uses the same variables as the Broadcast-Decide algorithm and the same initial state. 
-%that is the arrays of variables    
-%$inits\left[ s \right]$, $aborts\left[ s \right]$, $learned\left[ l \right]$,
-%$status\left[ a \right]$, and $accepted\left[ a \right]$, where $l$ is a learner, $a$ is
-%an acceptor, and $s$ is a switcher.
-
-%Initially, the variables $inits$, $aborts$, and, for every
-%learner $l$, $learned\left[ l \right]$ are empty sets. For every
-%acceptor $a$, $status\left[ a \right]$ is ``idle'' and $hist\left[ a
-%\right]$ is the special value $None$.
-
+The specification uses the same variables as the Broadcast-Decide algorithm and starts in the same initial state. 
 The next-state relation is the disjunction of the following types of
@@
   action is to receive the message, to update $accepted\left[ leader
-  \right]$ to $accepted\left[ leader \right]\star c$, and to
+  \right]$ to $accepted\left[ leader \right]\star c$,  to broadcast the message
+  $\aseq{\quo{leader-accept},accepted'\left[ leader \right]}$
+  to all the other acceptors, and to
   broadcast the message $\aseq{\quo{accepted},accepted'\left[ leader
-  \right]}$ to the learners, and to broadcast the message
-  $\aseq{\quo{leader-accept},accepted'\left[ leader \right]}$
-  to all the other acceptors.
+  \right]}$ to the learners.
 \item An $Accept\left( a \right)$ transition, for an acceptor $a$
@@
   to update $accepted\left[ a \right]$ to $h$, and to broadcast the
-  message $\aseq{\quo{accepted},hist\left[ a \right]\star c}$ to the
+  message $\aseq{\quo{accepted},h}$ to the
   learners.
@@
 \item A $Stop\left( s \right)$ transition, for a switcher $s$,
-  enabled when the switcher $s$ has waited for too long without receiving any messages from the acceptors, broadcasts a $\aseq{\quo{stop}, s}$ message to all
+  enabled when the switcher $s$ has detected a crash (for example using a timeout), broadcasts a $\aseq{\quo{stop}, s}$ message to all
   the acceptors. 
@@
   $status\left[ a \right]$ to ``stopped'', and to send the
-  message $\aseq{\quo{stopped},a,hist\left[ a \right]}$ to the
+  message $\aseq{\quo{stopped},a,accepted\left[ a \right]}$ to the
   switcher $s$. 
@@
 
-Thanks to the inter-round interface, which is the same for Leader rounds and Broadcast-Decide rounds, Leader rounds and Broadcast-Decide rounds can be repeatedly composed in a sequence.
+Thanks to the inter-round interface, which is the same for Leader rounds and Broadcast-Decide rounds, new Leader rounds and Broadcast-Decide rounds can be repeatedly composed in a sequence.
 
@@
 
-In the next section we present Composable Generalized Consensus, a specification of the behavior of rounds at the level of proposals, learned histories, init histories, and abort histories, which allows making precise the informal reasoning above.
+In the next section we present Composable Generalized Consensus, a specification of the behavior of rounds at the level of proposals, learned histories, init histories, and abort histories, which allows making the informal reasoning above precise.
 Our composition theorem ensures that any set of rounds individually satisfying Composable Generalized Consensus can be composed in any order to yield an algorithm satisfying Composable Generalized Consensus.
@@
 
-In this section we present Composable Generalized Consensus, our specification for composable Generalized Consensus rounds.
+We consider a set of servers playing one or multiple of the four roles of proposer, acceptor, learner, and switcher. 
+Composable Generalized Consensus specifies the allowed behaviors of the proposers, learners, and switchers
+\emph{of a single round}. 
+To specify Composable Generalized Consensus, we use the variable $proposed$, a set of proposals, the array of set of histories $learned\left[ l \right]$, where $l$ ranges of learners, the variable $inits$, a set containing the init histories received from the previous
+round, and the variable $aborts$, a set containing the abort histories transfered to the next round.
+Initially all variables are empty sets.
 
-As described in the preceding section, the servers can have any of the four roles of proposer, acceptor, learner, and switcher. 
-The assignment of the roles of proposer, acceptor, learner, and switcher can change in each rounds, allowing for
-reconfiguration.
+The next-state relation
+consists of four types of transitions. The first two, $Propose\left( l, c \right)$, and $Learn\left( l, h \right)$, for a learner $l$ and a
+history $h$, have the same effect as in the specification of Generalized Consensus: 
+$Propose\left( l,c \right)$ adds $c$ to the set $proposed$ and $Learn\left( l,h \right)$ adds $h$ to the set $learned\left[ l \right]$.
 
-We assume that in the first round, which has no previous round, every 
-switchers receives the empty abort history.
-
-\begin{comment}
-The scheduling policy determines when to abort a CGC instance and which new CGC
-round should take over. The scheduling policy must inform the switchers about the
-identity of the next round and must guarantee that all switchers receive the
-same information.  In this paper we abstract over the scheduling policy, assuming
-that a round that aborts passes the baton to a nondeterministically chosen new
-round. A reliable scheduling-policy component can realistically be implemented
-as a replicated state machine running on a large number of servers otherwise
-dedicated to other tasks
-\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}. We will see
- in \cref{sec:rounds} that, as in Vertical Paxos, a reliable scheduling policy component allows
-tolerating $f$ faults with $f+1$ replicas, while maintaining the long-term
-resilience of the service by replacing faulty servers upon round change.
-\end{comment}
-
-The clients of a CGC algorithm can send their commands to the proposers of any round (in
-practice, clients send their commands to proposers of the largest
-active round) and servers can use sequences learned in any round
-to determine outputs.
-
-CGC specifies the allowed behaviors of the proposers, learners, and switchers
-\emph{of a single round}. Similarly to
-Generalized Consensus, we use the state variables $crashed$, $proposed$,
-and, for every learner $l$, $learned\left[ l \right]$.  Additionally,
-we use the state variables $inits$, containing the init histories 
-received from the previous round, and $aborts$, containing the abort
-histories transfered to the next round.
-Two consecutive CGC rounds are composed by letting the $aborts$ variable of the
-first round be equal to the $inits$ variable of the second round.
-
-Initially, $proposed$, for every learner $l$, $learned\left[ l \right]$, $inits$, and $aborts$ are all empty sets. The next-state relation
-consists of four types of transitions. The first two, $Propose\left( l, c \right)$, and $Learn\left( l, h \right)$, for a learner $l$, and for a
-history $h$, have the same effect as in Generalized Consensus. CGC introduces two new transitions, $Init\left( s, h \right)$ and $Abort\left( s, h
-\right)$, for a switcher $s$ and a history $h$. The transition $Init\left( s,h \right)$ adds $h$ to the set $inits$, and models a switcher receiving
-the abort history $h$ from the preceding round. The transition $Abort\left( s, h \right)$ adds $cs$ to the set $aborts$ and models a switcher
-transferring the abort history $cs$ to a switcher of the next round. All the transition are guarded by the following conditions:
+Composable Generalized Consensus introduces two new transitions: $Init\left( s, h \right)$ and $Abort\left( s, h
+\right)$, for a switcher $s$ and a history $h$. The transition $Init\left( s,h \right)$ inserts $h$ in the set $inits$, and models the switcher $s$ receiving
+the abort history $h$ from the preceding round. The transition $Abort\left( s, h \right)$ inserts $h$ in the set $aborts$ and models the switcher $s$ passing the abort history $h$ to the next round.  
+All the transition are guarded by the following conditions:
 \begin{compactitem}
-    \item[\textbf{Agreement}:] The set $\aset{learned'\left[ l
+    \item[\textbf{Agreement}:] The set $\aset{\bigcup learned'\left[ l
       \right] : l \in Learners}$ is compatible.
@@
         \right)$ transition, the new learned history $h$ is an 
-        extension of $learned\left[ l \right]$.
-    \item[\textbf{Safe Abort}:] Every abort history is an extension in the 
-        partial order $\leq$ of a learned history.
+        extension of every history in $learned\left[ l \right]$.
+    \item[\textbf{Safe Abort}:] For every abort history $h$ there is a learner $l$ such that $h$ is an extension of the maximum learned history of $l$.
 \end{compactitem}
-Note that Agreement and Irrevocability are the same as in Generalized Consensus.
 
-If a client has submitted a command to a proposer, eventually,
-either the client gets a response from the current round, or
-the next round is initialized with an abort history and the
-command can be resubmitted to the next round.
+Note that if a round is initialized with empty abort histories, then it implements Generalized Consensus.
+
+We compose two rounds of Composable Generalized Consensus as we did in the preceding section.
+We write $proposed_2$, $learned_2$, $inits_2$, and $aborts_2$ for the variables of the first round and  $proposed_2$, $learned_2$, $inits_2$, and $aborts_2$
+for the variables of the second round.
+The transition relation of the composition is the disjunction of the following actions, conjoined with the condition that $aborts_1=inits_2$:
+\begin{itemize}
+  \item A $Propose$ or $Learn$ action of round 1, leaving the variables of round 2 unchanged.
+  \item A $Propose$ or $Learn$ action of round 2, leaving the variables of round 1 unchanged.
+  \item An $Init$ action of round 1, leaving the variables of round 2 unchanged.
+  \item An $Abort$ action of round 2, leaving the variables of round 1 unchanged.
+  \item A joint action consisting of an $Abort$ action of round 1 together with an $Init$ action of round 2.
+\end{itemize}
 
@@
 
-TODO: frame in terms of composability.
+\begin{thm}[Composition Theorem]
+    \label{thm:comp}
+Consider the composition of two rounds as defined above.
+Define the state variable $proposed$ as $proposed_1 \cup proposed_2$. For every learner $l$, define 
+$learned\left[ l \right]$ as the maximum of $learned_1\left[ l \right]$ and $learned_2\left[ l \right]$.
+For every switcher $s$, define $inits\left[ s \right]$ as $inits_1\left[ s \right]$ and $aborts\left[ s \right]$ as $aborts_2\left[ s \right]$.
 
-Consider the succession of $n$ rounds numbered $1$ to $n$.  To
-distinguish the variables of each round, we annotate the local
-variables of each round with the number of the round. For example, the
-variables of the seventh round are written $proposed_7$, for every
-learner $l$, $learned_7\left[ l \right]$, $inits_7$, and $aborts_7$.
+Then the sequence of states of the variables $proposed$, $learned$, $inits$, and $aborts$ satisfy the specification of Composable Generalized Consensus.
+\end{thm}
 
-Moreover, we assume that a command, which is proposed to a round and that
-is not learned before the first abort history is transfered to the
-next round, is proposed anew to the next round.
+The composition theorem can be applied recursively to show that the composition of any number of rounds which individually and independently satisfy Composable Generalized Consensus itself satisfies Composable Generalized Consensus.
+Therefore SMR algorithms implemented as rounds in our framework can be reused unmodified and optimized by adding new types of rounds incrementally, building on existing work instead of having to redo all the tests and proofs from scratch every time one adds a new optimization to an existing algorithm.
 
-Define the state variable $proposed$ as the union over all rounds $i$
-of the $proposed_i$ variables and, for every learner $l$, 
-define the $learned\left[ l \right]$ variable as the union over all
-rounds $i$ of the $learned_i\left[ l \right]$ variables.
+However, it may not be clear that most SMR algorithms can be expressed as the composition of one or more types of rounds satisfying Composable Generalized Consensus.
+
+We provide a mechanically-checked proof of the composition theorem in Isabelle/HOL at \url{http://losa.fr/cgc}.
+The proofs uses a version of the specification of CGC that has a more
+executable form that presented in this section. 
+The Isabelle/HOL formalization of CGC uses the theory of I/O automata, which
+is also formalized in the Isabelle/HOL. The proof uses a refinement
+mapping from the composition of two consecutive CGC instances to a single CGC
+instance.
 
@@
 
-\section{Fast and Conservative Rounds}
+\section{Fast and Resilient Rounds}
 \label{sec:rounds}
 
-Most existing SMR algorithms are structured in rounds that can be modified to obtain several CGC rounds.
-However, checking that the obtained rounds are correct may be quite difficult in the case of algorithms using complex quorum
-systems. In particular, safely aborting can become complex.
+In this section we demonstrate how a large class of SMR algorithms, those based on quorum systems \cite{GuerraouiVukolic10RefinedQuorumSystems}, can be expressed as the composition of one or more Composable Generalized Consensus rounds.
+We propose two specifications, \emph{resilient CGC
+rounds} and \emph{fast CGC rounds}, which can be refined to obtain executable rounds.
+For example, CGC versions of Classic Paxos and Chain
+Replication can be obtained by refining the specification of
+resilient CGC round, and CGC versions of the fast rounds of Fast Paxos can be obtained by refining the specification of fast CGC rounds.
+We provide TLA+ specifications precisely demonstrating how to do so. The refinements have been checked with the TLC model-checker.
+The specifications are available at \url{http://losa.fr/cgc}.
 
-To simplify the task of expressing existing algorithms as the composition of one
-or more CGC instances, we propose two specifications, \emph{conservative CGC
-rounds} and \emph{fast CGC rounds}, which can easily be concretized to obtain executable rounds.
-We show that CGC versions of Classic Paxos, Multi-Coordinated Paxos, Chain
-Replication, and Ring Paxos can be obtained by concretizing the specification of
-conservative CGC instances, and that CGC versions of the fast rounds of Fast Paxos can be obtained by concretizing the specification of fast CGC instances.
-
-Fast and conservative CGC instances are two specifications, at an
+Fast and resilient CGC round are two specifications, at an
 intermediate level of abstraction, which satisfy CGC. Both
-specifications specify how the state of the acceptors should be
-updated but abstract over the particular communication topology used
+specifications describe how the state of the acceptors should be
+updated, using the notion of quorum, but abstract over the particular communication topology used
 to implement the state accesses: the specifications use a global state
 which is accessible by all servers in all roles. One can refine fast
-or conservative rounds by implementing the state accesses and coordination
+or resilient rounds by implementing the state accesses and coordination
 using the network, obtaining a concrete algorithm.
 
-Conservative and fast CGC instances are specified using the state variables $inits$, $proposed$, for every learner $l$, $learned\left[ l \right]$
-and $aborts$, as in CGC, and, for every acceptor $a$, two new variables
+Resilient and fast CGC round are specified using the variables $proposed$, $aborts$, $inits$, $learned\left[ l \right]$, for each learner $l$, 
+and, for every acceptor $a$, the variable
 $status\left[ a \right]$, which can be either ``idle'', ``ready'', or
-``stopped'' and $hist\left[ a \right]$, the local history of the
+``stopped'' and $hist\left[ a \right]$, the local accepted history of the
 acceptor $a$, which can be either a history or the special value $None$.
-Initially, $inits$, $aborts$, and, for every learner
-$l$, $learned\left[ l \right]$ are empty sets, and, for every acceptor $a$,
+Initially, $inits$ and $aborts$ are empty sets, and, for every learner
+$l$, $learned\left[ l \right]$ is the empty history, and, for every acceptor $a$,
 $status\left[ a \right]$  is ``idle'' and $hist\left[ a \right]$ is the special
@@
 
-The next-state relations of conservative and fast rounds are composed
-of the following types of transitions:
+The next-state relations of resilient and fast rounds are composed
+of the following types of transitions. We use the notion of \emph{learn quorum} and \emph{abort quorum}, defined shortly after.
 \begin{compactitem}
@@
         updates $hist\left[ a \right]$ to $hist\left[ a \right]\star
-        c$, where $c$ is a proposed command. In conservative rounds,
+        c$, where $c$ is a proposed command. In resilient rounds,
         the transition is additionally guarded by the condition that
@@
     \item A $Learn\left( l, h \right)$ transition, for a learner $l$
-      and a history $h$, enabled when $h > learned\left[ l \right]$ and is
+      and a history $h$, enabled when $h \geq learned\left[ l \right]$ and is
       smaller than the GLB of the histories of a learn quorum of acceptors. The effect of
@@
     \item An $Abort\left( s, h \right)$ transition, for a switcher $s$
-        and a history $h$, is enabled when there exists a abort quorum
+        and a history $h$, is enabled when there exists an abort quorum
         $R$ such that, for every member $a$ of $R$, $status\left[ a
@@
 On top of the additional guard on the $Accept$ transition present only
-in conservative rounds, conservative and fast CGC instances differ in
+in resilient rounds, resilient and fast CGC round differ in
 their definition of learn quorums, abort quorums, and in their
@@
 
-In conservative CGC instances, abort and learn quorums are defined such
+In resilient CGC round, abort and learn quorums are defined such
 that the intersection between two learn quorums is nonempty and the
@@
 element over the local histories of the acceptors in $R$ which are not
-$None$. Note that the maximum exists because conservative rounds
+$None$. Note that the maximum exists because resilient rounds
 guarantee that the histories of the acceptors form a chain.
 
-In fast CGC instances, abort and learn quorums are defined such that
+In fast CGC round, abort and learn quorums are defined such that
 the intersection between any two learn quorums is nonempty and the
@@
 The operator $SafeAborts\left( R \right)$ is defined as the set of
-maximal elements over the set $\aset{GLB\left( \aset{hist\left[ a
+maximal elements over the set 
+\begin{equation}
+\aset{GLB\left( \aset{hist\left[ a
 \right] : a \in R \cap L \wedge hist\left[ a \right]\neq None} \right)
-: L \in LearnQ}$.
+: L \in LearnQ}.
+\end{equation}
 
 The cases in which abort quorums are singleton sets, in both
-conservative and fast rounds, allow tolerating
+resilient and fast rounds, allow tolerating
 $f$ faults with $f+1$ servers: a round can abort as long as one
@@
 without the scheduling policy instructing the switchers about the
-identity of the next round, and thus tolerating $f$ faults with
+identity of the next round and of the new servers replacing the failed ones, and thus tolerating $f$ faults with
 $f+1$ servers relies on the assumption of a reliable scheduling policy.
@@
   \label{thm:rounds}
-    Both conservative and fast CGC instances implement CGC under the refinement
+    Both resilient and fast CGC round implement CGC under the refinement
     mapping that consists in substituting $\overline{learned\left[ l \right]}$
@@
 Multi-Coordinated Paxos, Chain Replication, Ring Paxos, and Fast Paxos to obtain
-CGC instances. \Cref{thm:comp} then guarantee that all those different types of
+CGC round. \Cref{thm:comp} then guarantee that all those different types of
 rounds can be used in the same algorithm to yield an implementation of Chain
@@
 The rounds of the Classic Paxos algorithm can be modified so as to
-obtain an implementation of conservative CGC instances.  We define
-Classic Paxos CGC instances by specializing the specification of
-conservative CGC instances, giving a concrete protocol for disseminating
+obtain an implementation of resilient CGC round.  We define
+Classic Paxos CGC round by specializing the specification of
+resilient CGC round, giving a concrete protocol for disseminating
 the proposed commands to the acceptors and enforcing that the learned
@@
 
-The specification of Classic Paxos CGC instances uses the variables of
-the specification of conservative CGC instances, namely the variables
+The specification of Classic Paxos CGC round uses the variables of
+the specification of resilient CGC round, namely the variables
 $proposed$, $inits$, $aborts$, the arrays $learned\left[ l \right]$,
@@
 transitions, where the definitions of learn quorums, abort quorums,
-and of the $SafeAborts\left( R \right)$ operator are the ones of conservative
-CGC instances.
+and of the $SafeAborts\left( R \right)$ operator are the ones of resilient
+CGC round.
 \begin{compactitem}
@@
 original rounds, the leader of the new round directly queries the
-acceptors of the preceeding round, whereas in CGC instances, the
+acceptors of the preceeding round, whereas in CGC round, the
 switchers query the acceptors and then send their abort histories to
@@
 \begin{thm}
-  Classic Paxos CGC instances satisfy Composable Generalized Consensus.
+  Classic Paxos CGC round satisfy Composable Generalized Consensus.
 \end{thm}
@@
 \end{comment}
-\subsection{Multi-Coordinated Paxos, Chain Replication, and Ring Paxos.}
 
-Multi-Coordinated Paxos, Chain Replication, and Ring Paxos rounds have a similar
-structure as Classic Paxos rounds but differ in the way the leader is
+\subsection{Refining Fast and Resilient Rounds}
+
+Classic Paxos \cite{lamport2001paxos}, Multi-Coordinated Paxos \cite{CamargosSchmidtPedone07MulticoordinatedPaxos}, Chain Replication \cite{RenesseSchneider04ChainReplicationSupportingHighThroughputAvailability}, and Ring Paxos \cite{MarandiETAL10RingPaxosHighthroughputAtomicBroadcastProtocol} refine Resilient Rounds, using a leader to ensure that the acceptor's accepted histories remain compatible.
+They differ in the way the leader is
 implemented and in the way that the histories of the leader are disseminated to
@@
 
+The rounds of Classic Paxos are similar to the Leader algorithm of section 4.2 but
+tolerates that a minority of servers crash by using majority quorums instead of
+waiting for histories from all the acceptors.
+
 Multi-Coordinated Paxos avoids depending on a unique leader by using a
-distributed leader implementation that can make progress as long as a
-strict majority of the acceptors are correct, trading off resilience
-for a higher time complexity. Conservative CGC
-rounds can be implemented like with Classic Paxos, replacing the
-leader by the distributed leader of Multi-Coordinated Paxos.
+distributed leader implementation that can make progress as long as a strict
+majority of the acceptors are correct, trading off resilience for a higher time
+complexity.
 
@@
 can achieve higher throughput that Classic Paxos.
-Conservative CGC instances can be implemented like with Classic Paxos,
-replacing the broadcast of the leader by the propagation of the leader's history
-along the chain of servers.
 
@@
 of the commands along the chain. Ring Paxos rounds can be modified to implement
-conservative CGC instances in a similar way as for Chain Replication.
+resilient CGC round in a similar way as for Chain Replication.
+
+The fast rounds of Fast Paxos \cite{Lamport06FastPaxos} and Generalized Paxos \cite{Lamport05GeneralizeConsensus} refine Fast Rounds.
+They allow the proposers to directly broadcast their proposal to the 
+acceptors and make progress as long as non-commuting proposals do not contend.
+They use the same quorums as our Fast Rounds to ensure that it is possible to recover from inconsistencies 
+in the acceptors histories.
+Their structure is similar to the one of our Broadcast-Decide round.
 

### 2015-06-03 paper/paper.tex
Fixed little errors in paper.tex

@@
 $AbortQ = Q\left( 3/4 \right)$; $LearnQ = \aset{Acceptors}$ and
-$AbortQ = \aset{\aset{a : a \in Acceptors}}$.
+$AbortQ = Q\left( 0 \right)$.
 The operator $SafeAborts\left( R \right)$ is defined as the set of

### 2015-07-05 paper/abstract3.tex
Modified the abstract, version registered for POPL.

@@
 can have a big impact on the reliability of the many computing services they
-support. However, devising a new practical SMR algorithm is hard, often
+support. However, improving upon existing SMR algorithms is hard, often
 justifying a publication in a top conference.
 
-The problem is that SMR algorithms are not modular. Thus all the analysis,
-proofs, and tests need to be done again from scratch when devising a new
-algorithm. A modular approach allowing the reuse of existing techniques is
-especially important because of the many different situations that a practical
-SMR algorithm must face.
+The problem is that SMR algorithms are not modular. 
+Thus one cannot optimize an existing algorithm for a specific situation without having to 
+ redo from scratch all the analysis, proofs, and tests that were done for the existing algorithm.
 
-We present a framework that enables devising and composing modular SMR
-algorithms. Crucially, with our framework, an algorithm is proved composable
-once and for all in isolation. Our composition theorem, mechanically proved
-with Isabelle/HOL, guarantees that the composition of any number of composable
-algorithms is a correct SMR algorithm.
+We present a framework that enables building SMR algorithms out of independent modules.
+A module is an SMR algorithm that can abort its execution and pass the baton to a new module, allowing a composite SMR algorithm to dynamically switch between modules.
+Crucially, each module can be proved correct once and for all in isolation and our composition theorem guarantees that 
+the composition of any number of correct modules is a correct SMR algorithm.
 
@@
 a large class of SMR algorithms can be expressed in our framework by proposing
-two specifications capturing the essence of modular SMR algorithms based
-on Generalized Consensus and using quorum systems for synchronization. We
+two specifications capturing the essence of modular SMR algorithms based on quorum systems. We
 demonstrate that concrete algorithms can easily be obtained by refinement from

### 2015-07-21 paper/intro.tex
A few fixes.

@@
 
+\begin{comment}
 Allowing the execution of commuting commands to proceed without
@@
 improvement over the traditional $2k+1$ needed replicas.
+\end{comment}
 

### 2015-07-21 paper/paper.tex
A few fixes.

@@
 
-\vspace{2cm}
-\begin{center}
-This is a regular paper.
-\end{center}
+%\vspace{2cm}
+%\begin{center}
+%This is a regular paper.
+%\end{center}
 
@@
 replicated service, it is sufficient that learners learn histories instead of
-sequences. For a thorough discussion of command histories we refer the reader
-to \cite{Lamport05GeneralizeConsensus}. 
-Our Isabelle/HOL formalization abstracts over the concrete representation of histories, assuming only their
-necessary properties. Command histories were first introduced in trace theory
-\cite{Mazurkiewicz84Semantics}.
+sequences.
 
@@
 a history even though they chose different orders for some commuting commands.
+
+For a thorough discussion of command histories we refer the reader
+to \cite{Lamport05GeneralizeConsensus}, where histories are called C-structs. 
+%Our Isabelle/HOL formalization abstracts over the concrete representation of histories, assuming only their
+%necessary properties. 
+Command histories were first introduced in trace theory by Mazurkiewicz 
+\cite{Mazurkiewicz84Semantics}.
 
@@
 \subsection{The Leader Algorithm} 
+\label{sec:leader}
 
@@
 To do so, we define a new kind of rounds, the Leader rounds, which uses the same inter-round interface as the Broadcast-Decide algorithm. 
-The inter-round interface will allow use to combine Leader rounds and Broadcast-Decide rounds in any order, obtaining an algorithm which can adapt its strategy to its environment, selecting the Broadcast-Decide algorithm under low contention and switching to the slower Leader algorithm when there is contention.
+The inter-round interface will allow us to combine Leader rounds and Broadcast-Decide rounds in any order, obtaining an algorithm which can adapt its strategy to its environment, selecting the Broadcast-Decide algorithm under low contention and switching to the slower Leader algorithm when there is contention.
 
@@
         extension of every history in $learned\left[ l \right]$.
-    \item[\textbf{Safe Abort}:] For every abort history $h$ there is a learner $l$ such that $h$ is an extension of the maximum learned history of $l$.
+    \item[\textbf{Safe Abort}:] Every abort history $h$ is an upper bound of the set of learned histories.
 \end{compactitem}
@@
 The Isabelle/HOL formalization of CGC uses the theory of I/O automata, which
-is also formalized in the Isabelle/HOL. The proof uses a refinement
+is also formalized in Isabelle/HOL. The proof uses a refinement
 mapping from the composition of two consecutive CGC instances to a single CGC
@@
 
-The rounds of Classic Paxos are similar to the Leader algorithm of section 4.2 but
-tolerates that a minority of servers crash by using majority quorums instead of
+The rounds of Classic Paxos are similar to the Leader algorithm of \cref{sec:leader} but
+tolerate that a minority of servers crash by using majority quorums instead of
 waiting for histories from all the acceptors.

### 2015-11-10 paper/intro.tex
Added two new references.

@@
 Paxos \cite{lamport2001paxos}, Fast Paxos \cite{Lamport06FastPaxos},
-Disk Paxos \cite{GafniLamport03DiskPaxos}, Chain Replication
+Disk Paxos \cite{GafniLamport03DiskPaxos}, Generalized Paxos \cite{Lamport05GeneralizeConsensus}, FGGC \cite{SutraShapiro11FastGenuineGeneralizedConsensus}, Chain Replication
 \cite{RenesseSchneider04ChainReplicationSupportingHighThroughputAvailability},

### 2016-01-01 paper/abstract3.tex
A new Isabelle formalization

@@
 We present a framework that enables building SMR algorithms out of independent modules.
-A module is an SMR algorithm that can abort its execution and pass the baton to a new module, allowing a composite SMR algorithm to dynamically switch between modules.
+A module is an SMR algorithm that can abort its execution and pass the baton to a new module, allowing to build composite SMR algorithms which can dynamically switch between modules.
 Crucially, each module can be proved correct once and for all in isolation and our composition theorem guarantees that 

### 2016-01-01 paper/paper.tex
A new Isabelle formalization

@@
 
-Composable Generalized Consensus combines the ideas of the Abstract Framework \cite{GuerraouiETAL10Next700BftProtocols}, Generalized Conensus  \cite{Lamport05GeneralizeConsensus}, and Vertical Paxos \cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication} in a
+Composable Generalized Consensus combines the ideas of the Abstract Framework \cite{GuerraouiETAL10Next700BftProtocols}, Generalized Consensus  \cite{Lamport05GeneralizeConsensus}, and Vertical Paxos \cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication} in a
 unified abstraction.
@@
 Generalized Consensus is an agreement problem relaxing the ordering constraints
-imposed on commuting commands and Generalized Paxos is an algorithm, similar
+imposed on commuting commands and Generalized Paxos \cite{Lamport05GeneralizeConsensus} is an algorithm, similar
 in structure to Fast Paxos, which takes advantage of the relaxed ordering
@@
 needed in the Classic Paxos algorithm.
+
+Generalized Paxos is also structured in a sequence of rounds, 
+where each round maintains the agreement property of Generalized Consensus.
+In the abstract presentation of Generalized Paxos, Lamport gives a generic and abstract algorithm (not a distributed algorithm, accesses directly the state of all acceptors and rounds) that a round must use to determine the values that are safe to be voted for in the round.
+If all round refine this abstract specification, then the resulting algorithm implements Generalized Consensus.
+Therefore any set of rounds refining Lamport's abstract specification can be combined to form a correct generalized consensus algorithm.
+However, no explicit interface for combining rounds is given and specifying a type of round in isolation from the others would require assume-guarantee reasoning.
+In contrast, our formulation uses encapsulared rounds and does not require assume-guarantee reasoning.
 

### 2016-01-05 paper/abstract3.tex
Some modifications to the abstract and beginning of the intro, doesn't compile anymore and I don't know why.

@@
 can have a big impact on the reliability of the many computing services they
-support. However, improving upon existing SMR algorithms is hard, often
-justifying a publication in a top conference.
+support. However, doing so is hard, often justifying a publication in a top conference.
 
@@
 A module is an SMR algorithm that can abort its execution and pass the baton to a new module, allowing to build composite SMR algorithms which can dynamically switch between modules. 
-Crucially, each module can be proved correct once and for all in isolation and our composition theorem guarantees that 
+Crucially, each module can be analyzed once and for all in isolation; our composition theorem guarantees that 
 the composition of any number of correct modules is a correct SMR algorithm.
 
-The key novelty of our framework is its practical applicability: we show that
+The key novelty of our framework is its wide applicability: we show that
 a large class of SMR algorithms can be expressed in our framework by proposing
@@
 demonstrate that concrete algorithms can easily be obtained by refinement from
-our abstract specifications, as evidenced by TLA+ specifications ready to be
-checked by the TLC model checker.
+our abstract specifications and we show that our framework enables formal specification and verification of composite SMR algorithms in TLA+ and Isabelle/HOL.
 }

### 2016-01-05 paper/intro.tex
Some modifications to the abstract and beginning of the intro, doesn't compile anymore and I don't know why.

--- 2015-07-21 paper/intro.tex
@@
 reliable services on top of unreliable hardware. The basic idea is to
-replicate a service over several servers, each executing the same sequence of
+replicate a service over several servers and enforce through a distributed protocol that each execute the same sequence of
 deterministic commands. In such a setup, any replica can be used to determine
@@
 and a multitude of algorithms have been proposed. Examples include
-Paxos \cite{lamport2001paxos}, Fast Paxos \cite{Lamport06FastPaxos},
-Disk Paxos \cite{GafniLamport03DiskPaxos}, Chain Replication
-\cite{RenesseSchneider04ChainReplicationSupportingHighThroughputAvailability},
-Ring Paxos \cite{MarandiETAL10RingPaxosHighthroughputAtomicBroadcastProtocol},
+Paxos~\cite{lamport2001paxos}, Fast Paxos~\cite{Lamport06FastPaxos},
+Disk Paxos~\cite{GafniLamport03DiskPaxos}, Generalized Paxos~\cite{Lamport05GeneralizeConsensus}, FGGC~\cite{SutraShapiro11FastGenuineGeneralizedConsensus}, Chain Replication~\cite{RenesseSchneider04ChainReplicationSupportingHighThroughputAvailability},
+Ring Paxos~\cite{MarandiETAL10RingPaxosHighthroughputAtomicBroadcastProtocol},
 Egalitarian Paxos
 \cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments},
-Multi-Coordinated Paxos \cite{CamargosSchmidtPedone07MulticoordinatedPaxos},
-Cheap Paxos \cite{LamportMassa04CheapPaxos}, Vertical Paxos
+Multi-Coordinated Paxos~\cite{CamargosSchmidtPedone07MulticoordinatedPaxos},
+Cheap Paxos~\cite{LamportMassa04CheapPaxos}, Vertical Paxos
 \cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}, Paxos-MIC
@@
 \cite{MaoJunqueiraMarzullo08MenciusBuildingEfficientReplicatedStateMachine},
-Fast Mencius \cite{WeiETAL13FastMenciusMenciusLowCommitLatency},
+Fast Mencius~\cite{WeiETAL13FastMenciusMenciusLowCommitLatency},
 the Chandra-Toueg algorithm
 \cite{ChandraToueg96UnreliableFailureDetectorsReliableDistributedSystems}, the
-Ben-Or algorithm \cite{BenOr83AnotherAdvantageFreeChoiceCompletelyAsynchronous},
-and Raft \cite{OngaroOusterhout14SearchUnderstandableConsensusAlgorithm}.
+Ben-Or algorithm~\cite{BenOr83AnotherAdvantageFreeChoiceCompletelyAsynchronous},
+and Raft~\cite{OngaroOusterhout14SearchUnderstandableConsensusAlgorithm}.
 
-With the wide use of cloud computing, improvements in the performance and
-efficiency of SMR algorithms have the potential to impact most online services:
+Improvements in the performance and
+efficiency of SMR algorithms have the potential to impact most online services, which widely rely on cloud computing:
 advanced SMR algorithms can be implemented by cloud computing service providers
@@
 particular situations or metrics in mind, is hard: today,
-more than two decades after the first SMR algorithms
-\cite{Lamport98ParttimeParliament,BirmanJoseph87ReliableCommunicationPresenceFailures,OkiLiskov88ViewstampedReplicationGeneralPrimaryCopy,DworkLynchStockmeyer84ConsensusPresencePartialSynchronyPreliminaryVersion} were proposed, devising new
-SMR algorithms is still complex enough to justify a publication in a top conference
-\cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments,OngaroOusterhout14SearchUnderstandableConsensusAlgorithm}.
+more than two decades after the first SMR algorithms~\cite{Lamport98ParttimeParliament,BirmanJoseph87ReliableCommunicationPresenceFailures,OkiLiskov88ViewstampedReplicationGeneralPrimaryCopy,DworkLynchStockmeyer84ConsensusPresencePartialSynchronyPreliminaryVersion} were proposed, devising new SMR algorithms is still complex enough to justify a publication in a top conference~\cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments,OngaroOusterhout14SearchUnderstandableConsensusAlgorithm}.
 
@@
 
-We propose a framework to make existing Generalized Consensus algorithms modular
+We propose a framework to make existing Generalized Consensus (TODO Generalized Consensus has not been introduced) algorithms modular
 and to enable adding new optimizations incrementally, without modification to
@@
   \item Many SMR algorithms are variations on a basic structure consisting of one of two kinds of 
-    quorum systems \cite{GuerraouiVukolic10RefinedQuorumSystems}. 
+    quorum systems~\cite{GuerraouiVukolic10RefinedQuorumSystems}. 
     The first kind, that we call resilient quorums, are used in leader-based algorithms, while the second
@@
 of a data structure. To enable the synchronization-free execution of
-commuting commands, we have based our framework on Generalized Consensus
-\cite{Lamport05GeneralizeConsensus}. Traditional SMR implementations use a
+commuting commands, we have based our framework on Generalized Consensus~\cite{Lamport05GeneralizeConsensus}. Traditional SMR implementations use a
 separate instance of a consensus algorithm to agree on the command to execute
@@
 servers will participate, removing failed replicas and adding new ones. The
-reconfiguration master enables tolerating $k$ failures with $k+1$ replicas
-\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}, an important
+reconfiguration master enables tolerating $k$ failures with $k+1$ replicas~\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}, an important
 improvement over the traditional $2k+1$ needed replicas.
@@
 the addition of new replicas and the removal of failed ones. Interestingly, as
-explained in \cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication},
+explained in~\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication},
 in a real system it is often possible to implement a reliable reconfiguration

### 2016-01-06 paper/abstract3.tex
Switched to the llncs template. Reworked the abstract and intro.

@@
-\abstract{
+\begin{abstract}
 State-Machine Replication (SMR) is a basic building block of reliable
@@
 can have a big impact on the reliability of the many computing services they
-support. However, doing so is hard, often justifying a publication in a top conference.
+support. However, doing so is hard, often justifying a publication in a top
+conference.
 
-The problem is that SMR algorithms are not modular. 
-Thus one cannot optimize an existing algorithm for a specific situation without having to 
- redo from scratch all the analysis, proofs, and tests that were done for the existing algorithm.
+The problem is that SMR algorithms are not modular. Thus one cannot optimize an
+existing algorithm for a specific situation without having to redo from scratch
+all the analysis, proofs, and tests that were done for the existing algorithm.
 
-We present a framework that enables building SMR algorithms out of independent modules. 
-A module is an SMR algorithm that can abort its execution and pass the baton to a new module, allowing to build composite SMR algorithms which can dynamically switch between modules. 
-Crucially, each module can be analyzed once and for all in isolation; our composition theorem guarantees that 
-the composition of any number of correct modules is a correct SMR algorithm.
+We present a framework that enables building SMR algorithms out of independent
+modules. A module is an SMR algorithm that can abort its execution and pass
+the baton to a new module, allowing to build composite SMR algorithms which
+can dynamically switch between modules. Crucially, each module can be analyzed
+once and for all in isolation; our composition theorem guarantees that the
+composition of any number of correct modules is a correct SMR algorithm.
 
-The key novelty of our framework is its wide applicability: we show that
-a large class of SMR algorithms can be expressed in our framework by proposing
-two specifications capturing the essence of modular SMR algorithms based on quorum systems. We
-demonstrate that concrete algorithms can easily be obtained by refinement from
-our abstract specifications and we show that our framework enables formal specification and verification of composite SMR algorithms in TLA+ and Isabelle/HOL.
-}
+The key novelty of our framework is its wide applicability: we show that a
+large class of SMR algorithms can be expressed in our framework by proposing
+two specifications capturing the essence of modular SMR algorithms based on
+quorum systems. We demonstrate that concrete algorithms can easily be obtained
+by refinement from our abstract specifications and we show that our framework
+enables formal specification and verification of composite SMR algorithms in
+TLA+ and Isabelle/HOL\@.
+\end{abstract}

### 2016-01-06 paper/abstract4.tex
Switched to the llncs template. Reworked the abstract and intro.

\begin{abstract}

State-Machine Replication (SMR) algorithms such as Paxos are notoriously hard to
understand, let alone to devise. Moreover, there is no practical way to reuse and optimize existing SMR
algorithms, therefore the cost of creating a new SMR algorithm is high.

To lower this cost, we present a framework for building SMR algorithms out of independent and reusable SMR modules. A module is an SMR
algorithm that can abort and pass the baton to another module, allowing to dynamically switch between modules. 
We propose Composable
Generalized Consensus (CGC), a correctness condition for individual modules, and
we prove that the composition of correct modules is a correct SMR algorithm.
Our framework consists of TLA$^+$ specifications of CGC and of two abstract
types of correct modules, fast rounds and conservative rounds, each providing a template to devise new SMR modules with the help of the TLA$^+$
tools. We show that several well-known SMR algorithms can be modified to
implement fast or conservative rounds, therefore enabling
their composition to obtain a single versatile SMR algorithm.

%With our specifications, algorithm designers can use the TLA$^+$ tools
%to prototype and debug new SMR algorithms. Moreover, we also provide a
%specification of CGC and a proof of the composition theorem in Isabelle/HOL\@.

\end{abstract}

### 2016-01-06 paper/intro.tex
Switched to the llncs template. Reworked the abstract and intro.

--- 2015-07-21 paper/intro.tex
@@
 The State-Machine Replication technique (abbreviated SMR) allows building
-reliable services on top of unreliable hardware. The basic idea is to
-replicate a service over several servers, each executing the same sequence of
-deterministic commands. In such a setup, any replica can be used to determine
-the result of a command. Therefore, if a replica crashes, clients can be
-transparently redirected to another one and will have the illusion of accessing
-a centralized service at all times.
+reliable computing services on top of unreliable hardware. The basic idea is
+to replicate a service over several servers and enforce, using a distributed
+protocol, that each server executes the same sequence of deterministic commands.
+In this fashion, a command will produce the same effect on all replicas.
+Therefore, if a replica crashes, clients of the service can be transparently redirected to
+another replica and will have the illusion of accessing a centralized service at
+all times. SMR is often formally specified using the Generalized Consensus
+specification~\cite{Lamport05GeneralizeConsensus}, abbreviated GC\@.
 
-Making SMR algorithms efficient has been a long-standing challenge
-and a multitude of algorithms have been proposed. Examples include
-Paxos \cite{lamport2001paxos}, Fast Paxos \cite{Lamport06FastPaxos},
-Disk Paxos \cite{GafniLamport03DiskPaxos}, Chain Replication
-\cite{RenesseSchneider04ChainReplicationSupportingHighThroughputAvailability},
-Ring Paxos \cite{MarandiETAL10RingPaxosHighthroughputAtomicBroadcastProtocol},
-Egalitarian Paxos
-\cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments},
-Multi-Coordinated Paxos \cite{CamargosSchmidtPedone07MulticoordinatedPaxos},
-Cheap Paxos \cite{LamportMassa04CheapPaxos}, Vertical Paxos
-\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}, Paxos-MIC
-\cite{HurfinMoiseNarzul11AdaptiveFastPaxosMakingQuickEverlasting}, Mencius
-\cite{MaoJunqueiraMarzullo08MenciusBuildingEfficientReplicatedStateMachine},
-Fast Mencius \cite{WeiETAL13FastMenciusMenciusLowCommitLatency},
-the Chandra-Toueg algorithm
-\cite{ChandraToueg96UnreliableFailureDetectorsReliableDistributedSystems}, the
-Ben-Or algorithm \cite{BenOr83AnotherAdvantageFreeChoiceCompletelyAsynchronous},
-and Raft \cite{OngaroOusterhout14SearchUnderstandableConsensusAlgorithm}.
+Because SMR is part of the basic infrastructure supporting cloud computing,
+improving the performance of SMR algorithms can result in performance gains,
+cost reduction, or improved resilience (because SMR may be used where it was
+too costly before) for many online services.
 
-With the wide use of cloud computing, improvements in the performance and
-efficiency of SMR algorithms have the potential to impact most online services:
-advanced SMR algorithms can be implemented by cloud computing service providers
-and benefit all their customer's applications.
+However, devising new SMR algorithms or optimizing existing ones for a
+particular use-case is hard: today, more than two decades after the first SMR
+algorithms~\cite{Lamport98ParttimeParliament,BirmanJoseph87ReliableCommunicationPresenceFailures,OkiLiskov88ViewstampedReplicationGeneralPrimaryCopy,DworkLynchStockmeyer84ConsensusPresencePartialSynchronyPreliminaryVersion}
+were proposed, devising a new SMR algorithm can justify a publication in a top
+conference~\cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments,OngaroOusterhout14SearchUnderstandableConsensusAlgorithm}.
 
-However, devising new SMR algorithms, optimized with
-particular situations or metrics in mind, is hard: today,
-more than two decades after the first SMR algorithms
-\cite{Lamport98ParttimeParliament,BirmanJoseph87ReliableCommunicationPresenceFailures,OkiLiskov88ViewstampedReplicationGeneralPrimaryCopy,DworkLynchStockmeyer84ConsensusPresencePartialSynchronyPreliminaryVersion} were proposed, devising new
-SMR algorithms is still complex enough to justify a publication in a top conference
-\cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments,OngaroOusterhout14SearchUnderstandableConsensusAlgorithm}.
+The problem is that it is difficult to reuse existing algorithms and ideas
+because SMR algorithms are not modular. When devising a new algorithm, any
+analysis, test, or correctness proof has to be done again from scratch.
 
-The problem is that it is difficult to reuse existing algorithms and ideas because
-SMR algorithms are not modular. When devising a new algorithm, all the careful
-analysis, tests, and proofs have to be done again from scratch.
+To simplify the task of devising new SMR algorithms we propose a framework for
+devising reusable SMR modules. A module in our framework is an SMR algorithm
+augmented with an \emph{abort interface} for aborting its execution and passing
+the baton to another module and with a matching \emph{init interface} for taking
+over the execution of an aborting module. Two modules are composed by passing
+the outputs produced by the abort interface of the first module, called abort
+values, as inputs to the init interface of the second module, called init values
+in the context of the second module.
 
-We propose a framework to make existing Generalized Consensus algorithms modular
-and to enable adding new optimizations incrementally, without modification to
-the existing. Generalized consensus, abbreviated GC, is a precise formulation of
-State-Machine Replication that subsumes consensus. An algorithm in our framework
-is composed of independent modules that can run, abort their execution, and pass
-the baton to any other module on the fly.
+A module must satisfy a correctness condition that we call Composable
+Generalized Consensus (abbreviated CGC). Any correct module is a safe
+Generalized Consensus~\cite{Lamport05GeneralizeConsensus} algorithm, but the progress guarantees may vary between
+modules: CGC only requires a module to make progress executing commands or to
+abort its execution.
+CGC also relates the init values, learned values (learned values represent the local state-machine state of individual replicas), and abort values in order for our composition theorem to hold.
+Generalized Consensus is a specification of SMR algorithms that takes into account the commutativity relation among commands and can off important performance advantages in practice.
 
-Crucially, when adding a new module to an existing algorithm, all the
-performance analysis, tests, and proofs of the existing modules remain valid and
-only the new GC module needs to be analyzed: Our composition theorem guarantees
-that the composition of any number of correct modules is a correct GC algorithm,
-thus one can combine different GC algorithms, optimized for different situations
-and metrics, without any changes and with guaranteed correctness. Our framework
-therefore enables incremental development of GC algorithms, where each new
-development can build on the previous ones instead of starting again from
-scratch.
+The composition theorem enables incremental development of SMR algorithms
+because it guarantees that any number of correct modules may be composed in
+a sequence to form a correct SMR algorithm. Therefore, our framework allows
+building \emph{dynamic SMR algorithms}, which dynamically change module when
+the currently running module aborts, choosing the module best suited to the
+current operating conditions. Thanks to the composition theorem, new modules can
+be added to a dynamic SMR algorithm without modifying the existing modules and
+with guaranteed correctness. Moreover, the relaxed progress requirements of CGC
+enable a separation of concerns that simplifies optimizing SMR algorithms for
+particular scenarios because there is no need to consider executing commands
+under a situation outside the scenario: the module can just abort in this case.
 
-Moreover, our framework enables a clear separation of concerns which simplifies
-the design of new individual modules. A composite GC algorithm built with our
-framework works by running the GC module deemed the most appropriate at any
-given time, dynamically changing GC module when needed. Therefore a composable
-GC algorithm is not required to make progress in all situations: if it does
-not make progress, the system can switch to another GC module. Thus algorithms
-designers who want to create a new GC algorithm, optimized for particular
-conditions, can focus only on the particular conditions of interest, knowing
-that the other existing modules will handle the other cases.
+In practice, we provide a TLA+ specification of CGC that the user of our framework can use to debug and prove new modules correct using the TLA$^+$ toolbox~\cite{TLATools}.
+Moreover, we also provide two TLA+ specifications simplifying the development of modules based on Refined Quorum Systems~\cite{GuerraouiVukolic10RefinedQuorumSystems}.
+These two specifications both implement CGC and provide a skeleton for a user to build her own module.
+The two specifications demonstrate how to update the local state of a replica after gathering the states or the init values of a write quorum of replicas, and how to
+produce an abort value after gathering the state of a read quorum of replicas. 
+The two specifications use different combinations of sizes of read and write quorums, which affects the invariants that the replicas must maintain and, in consequence, the latency with which a command can be processed but the module.
+Our two specifications abstract over how to disseminate the commands to the replicas (e.g.\ whether to use a leader as in Paxos, multiple leaders as in MultiCoordinatedPaxos, or whether to broadcast commands without using a leader as in Fast Paxos).
+The user can therefore specify a concrete command dissemination procedure and then model-check that the obtained algorithm refines one of the two specifications.  
+We demonstrate the practical use of our framework by presenting modules based on Paxos~\cite{Lamport98ParttimeParliament}, Fast Paxos\cite{Lamport06FastPaxos}, Chain Replication~\cite{RenesseSchneider04ChainReplicationSupportingHighThroughputAvailability}, a case of Vertical Paxos~\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}, and EPaxos~\cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments}
+(TODO, but would be nice).
 
-The key novelty of our framework lies in its practical applicability:
+Compared to related work, described in \cref{sec:related}, our key contributions are the following.
 \begin{itemize}
+    \item We provide tool support for designing CGC modules in the form of two TLA+~\cite{Lamport02SpecifyingSystems} specifications simplifying the design and model-checking of SMR algorithms based on Quorum Systems~\cite{GuerraouiVukolic10RefinedQuorumSystems} and demonstrate their use by designing CGC modules inspired by Paxos, Fast Paxos, and Primary-Backup replication.
     \item Our framework is based on Generalized Consensus, a generalization of the consensus problem 
-    that allows optimizing the execution of commuting commands.
-  \item Many SMR algorithms are variations on a basic structure consisting of one of two kinds of 
-    quorum systems \cite{GuerraouiVukolic10RefinedQuorumSystems}. 
-    The first kind, that we call resilient quorums, are used in leader-based algorithms, while the second
-    kind of quorums, that we call fast quorums, are used in leader-less algorithms and can order commuting commands without synchronization.
-    We provide two specifications, at an intermediate level of abstraction, that capture the essence of composing algorithms using these two kinds of quorums.
-    Our two specifications allow to easily check that new modules based on resilient or fast quorums are correct, covering a large spectrum of practical algorithms. 
+        that allows optimizing the execution of commuting commands, a common practical case.
 \end{itemize}
-
-\begin{comment}
-Allowing the execution of commuting commands to proceed without
-synchronization is an important practical optimization, as commands may
-often commute in practice, for example when operating on disjoint parts
-of a data structure. To enable the synchronization-free execution of
-commuting commands, we have based our framework on Generalized Consensus
-\cite{Lamport05GeneralizeConsensus}. Traditional SMR implementations use a
-separate instance of a consensus algorithm to agree on the command to execute
-at each index of the command sequence on which the replicas must agree. Using
-separate consensus instances makes it hard to optimize the execution of
-commuting command because commutativity is not a property of a single command.
-Instead of restricting agreement to a single value, the Generalized Consensus
-problem requires that set of replica servers agree on the state of a growing
-data structure called a command-structure set. A command-structure set is
-a data structure with an append operation such that the order in which two
-commuting commands are appended does not change the resulting state. Thanks to
-this property, two replicas executing commuting commands in a different order
-will not end up in a different state and can continue executing without further
-synchronization. In message-passing systems, this property can be leveraged to
-reduce the latency of processing one command to two message delays instead of
-three. 
-%As we will see, fast execution of commutative requests also enables the processing of read-only commands
-%in two message delays, although read-only commands do not commute with write
-%commands.
-
-In practice, a SMR algorithm must support fast reconfiguration to replace
-faulty components, without which the long-term stability of a system is
-inevitably compromised. Our framework abstracts over how the reconfiguration
-decisions are made, providing an interface allowing to plug in an external
-reconfiguration master implementing the reconfiguration policy of the
-system. The master decides which GC module to run next and which replica
-servers will participate, removing failed replicas and adding new ones. The
-reconfiguration master enables tolerating $k$ failures with $k+1$ replicas
-\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}, an important
-improvement over the traditional $2k+1$ needed replicas.
-\end{comment}
-
-To demonstrate the practical applicability of our framework we provide
-composable versions of the generalized variations of popular consensus
-algorithms: Classic Paxos, Fast Paxos, Chain Replication, and multi-coordinated
-Paxos. Thanks to our specifications of algorithms based on quorum-systems, 
-we have checked the correctness of our version of Fast Paxos and Chain Replication with the TLC
-model-checker and we provide all the specifications needed to replicate these
-experiments. Our framework is also fully formalized and its composition theorem
-proved in Isabelle/HOL, a foundational interactive theorem prover, thereby
-certifying the correctness of the composition theorem at the deepest level.
-The TLA+ and Isabelle/HOL development can be found at \url{http://losa.fr/cgc}.
-
-At the heart of our framework lies an abstraction, called Composable Generalized
-Consensus (CGC), of the familiar notion of round (also called epoch, ballot, or
-view in the literature) used in most SMR algorithms. As an example, let us
-consider Fast Paxos, one of the first SMR algorithm which can change strategy
-dynamically. Fast Paxos is a round-based SMR algorithm that has two types of
-rounds: fast rounds and normal rounds. Fast rounds process requests faster
-than normal rounds in contention-free executions but fall back to a slower
-scheme under contention. Normal rounds are the same as in the classic Paxos
-and are slower than fast rounds in contention-free executions. Any process can
-initiate a round change at any point during execution: this allows switching
-between fast and normal rounds as well as changing the leader from one normal
-round to the next, similarly to classic Paxos. Each kind of round maintains
-different invariants relating the state of each process and the values that
-might have been chosen in the previous rounds, thus a given round must know
-which past rounds are fast or normal rounds in order to take over the execution
-safely. When adding more kinds of rounds such an approach will quickly become
-unmanageable, as each type of rounds must have one case for each possible other
-kind of round, making $O(n^2)$ cases to consider in total.
-
-Our notion of Composable Generalized Consensus defines a common interface and a
-common correctness property for a round, abstracting over the particular round
-implementations and making all types of rounds compatible with each other.
-
-Moreover, our interface for rounds enables a crucial optimization seen in
-practical systems: introducing an external reconfiguration master which manages
-the addition of new replicas and the removal of failed ones. Interestingly, as
-explained in \cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication},
-in a real system it is often possible to implement a reliable reconfiguration
-master which allows to tolerate $f$ faults with $f+1$ replicas instead of the
-traditional $2f+1$ replicas. The reconfiguration master can also make the choice
-of the new type of round to execute, providing the system's reconfiguration
-policy.
+These two contributions set this paper apart from related work, notably the Abstract framework~\cite{GuerraouiETAL10Next700BftProtocols} and the Speculative Linearizability framework~\cite{GuerraouiKuncakLosa12SpeculativeLinearizability}, both described in \cref{sec:related}.
+Our TLA+ specifications can be found at \url{http://losa.fr/cgc}.
+Moreover, we have also formalized the CGC correctness condition and proved the composition theorem in Isabelle/HOL\@, and the Isabelle theories are available at the same URL\@.

### 2016-01-06 paper/paper.tex
Switched to the llncs template. Reworked the abstract and intro.

@@
-\input{header.tex}
+\documentclass{llncs}
+\usepackage[english]{babel}
+\usepackage{hyperref}
+\usepackage{cleveref}
+\usepackage{comment}
+\usepackage{datetime}
+\usepackage[url=true,doi=false,isbn=false,backend=biber]{biblatex}
+\usepackage{mathtools}
+\usepackage{paralist}
+\usepackage{mathtools}
+
+%\input{header.tex}
 \input{defs.tex}
@@
 
-\begin{comment}
-SMR is often implemented using a sequence of instances of Consensus.
-There are numerous examples of consensus algorithms in the literature.
-However, Composable Generalized Consensus is based on Generalized Consensus. 
-In Generalized Consensus \cite{Lamport05GeneralizeConsensus}, the learners of a
-SMR algorithm repeatedly learn new values belonging to a \emph{C-struct set}. A
-C-struct can be seen as a data structure encoding  sequences of commands modulo
-the ordering of commands that commute. The advantage of Generalized
-Consensus over Consensus is that
-two servers which receive commuting commands in different orders can still reach
-agreement on C-structs without further synchronization. 
-Our Composable Generalized Consensus can also use C-structs, with the same advantages
-as Generalized Consensus. 
-\end{comment}
+Our work is closely related to the Abstract
+framework~\cite{GuerraouiETAL10Next700BftProtocols}
+and the Speculative Linearizability
+framework~\cite{GuerraouiKuncakLosa12SpeculativeLinearizability}. The
+Abstract framework pioneered the idea of decomposing SMR algorithms in
+independent modules and experimentally showed that the resulting composite
+algorithms outperformed state-of-the-art monolithic algorithms. Compared
+to our work, the Abstract framework is not precisely formalized and lacks
+tools supporting the development of correct algorithms in the framework,
+which are direly needed given the notorious complexity of SMR algorithms.
+Moreover, the Abstract framework does not allow optimizing for the execution
+of commuting commands. However, the Abstract framework supports byzantine
+fault-tolerant~\cite{LamportShostakPease82ByzantineGeneralsProblem} SMR
+algorithms, which we do not.
 
-Composable Generalized Consensus combines the ideas of the Abstract Framework \cite{GuerraouiETAL10Next700BftProtocols}, Generalized Consensus  \cite{Lamport05GeneralizeConsensus}, and Vertical Paxos \cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication} in a
-unified abstraction.
+The Speculative Linearizability framework was a
+first attempts by the authors to formalize the Abstract framework, provide
+formal proof support in Isabelle/HOL~\cite{NipkowPaulsonWenzel02IsabelleHOL},
+and extend the Abstract framework to support algorithms that optimize
+the execution of commuting commands by requiring Speculatively
+Linearizable modules to refine Abstract modules (under a forward/backward
+simulation~\cite{LynchVaandrager95ForwardBackwardSimulationsIUntimedSystems})
+modulo an equivalence relation on histories of commands that reflects the
+commutativity relation among commands. However, the simulation technique led to
+complex correctness statements that were hard to scale beyond the examples of
+the paper, and no realistic protocol was built in the framework. The present
+work addresses the shortcoming of the Speculatively Linearizable framework by
+using the notion of command-structure set introduced for the definition of
+Generalized Consensus~\cite{Lamport05GeneralizeConsensus}. Moreover, we show how
+several well known and practical algorithms can be cast and optimized in our
+framework, providing a set of TLA+ specifications that ease the design and model
+checking of algorithms based on quorum systems.
+
+TODO\@: expand the following.
 
@@
 
-Our work builds on the Abstract Framework , which proposes to
-build Byzantine Fault-Tolerant algorithms as the succession of abortable rounds
-called Abstract instances. Abstract instances must satisfy the Abstract
-correctness properties. By construction, the composition of any number of
-Abstract instances is a correct SMR implementation. The Abstract
-Framework is similar to our framework but uses totally ordered sequence of commands, making it difficult apply it to 
-practical algorithms that optimize the execution of commuting and read-only commands.
-
-Finally, the Speculative Linearizability framework
-\cite{GuerraouiKuncakLosa12SpeculativeLinearizability} was a first
-attempt by the authors of the present paper at generalizing the
-Abstract Framework and mechanically proving its properties with the
-Isabelle/HOL interactive proofs assistant.  Like the Abstract Framework, the
-Speculative Linearizability framework is based on totally ordered
-sequences of commands, with the same drawback for the execution of
-commuting and read-only commands. Moreover, Composable Generalized Consensus is both
-conceptually simpler and subsumes the Speculative Linearizability
-property. Composable Generalized Consensus
-and its main property, the composition theorem, are also formalized in Isabelle/HOL.
-
-\begin{comment}
-Composable Generalized Consensus is the most complete framework to date for building
-adaptive replication algorithms out of heterogeneous rounds, making state of the
-art optimization techniques available to algorithm designers, which was not
-possible before.
-\end{comment}
-
 \section{Generalized Consensus}
@@
 We describe a behavior of the system using a sequence of states,
-where a state is a function from variable names to values (i.e. a valuation of
+where a state is a function from variable names to values (i.e.\ a valuation of
 variables). We specify sets of allowed behaviors using an initial predicate
@@
 
-The Broadcast-Decide algorithm can process a proposal in two message delays when there are no conflicts, i.e.
-no contention on non-commuting commands.
+The Broadcast-Decide algorithm can process a proposal in two message delays when there are no conflicts, i.e.\ no contention on non-commuting commands.
 However, if there are repeated conflicts, the algorithm may not make progress despite changing round, as the acceptors of the each new round
@@
 
-\begin{thm}[Composition Theorem]
+\begin{theorem}[Composition Theorem]
     \label{thm:comp}
@@
 Then the sequence of states of the variables $proposed$, $learned$, $inits$, and $aborts$ satisfy the specification of Composable Generalized Consensus.
-\end{thm}
+\end{theorem}
 
@@
 The Isabelle/HOL formalization of CGC uses the theory of I/O automata, which
-is also formalized in Isabelle/HOL. The proof uses a refinement
+is also formalized in Isabelle/HOL\@. The proof uses a refinement
 mapping from the composition of two consecutive CGC instances to a single CGC
@@
 \begin{comment}
-\begin{thm}[Composition Theorem]
+\begin{theorem}[Composition Theorem]
     \label{thm:comp}
@@
     $learned\left[ l \right]$ satisfies Generalized Consensus.
-\end{thm}
+\end{theorem}
 \Cref{thm:comp} can be shown by induction on the number of rounds,
@@
   The Isabelle/HOL formalization of CGC uses the theory of I/O automata, which
-  is also formalized in the Isabelle/HOL. The proof uses a refinement
+  is also formalized in the Isabelle/HOL\@. The proof uses a refinement
   mapping from the composition of two consecutive CGC instances to a single CGC
@@
 Fast and resilient CGC round are two specifications, at an
-intermediate level of abstraction, which satisfy CGC. Both
+intermediate level of abstraction, which satisfy CGC\@. Both
 specifications describe how the state of the acceptors should be
@@
 definition of a history variable).
-\begin{thm}
+\begin{theorem}
   \label{thm:rounds}
@@
     for $learned\left[ l \right]$.
-\end{thm}
+\end{theorem}
 
@@
 this difference should not have much practical impact on performance.
-\begin{thm}
+\begin{theorem}
   Classic Paxos CGC round satisfy Composable Generalized Consensus.
-\end{thm}
+\end{theorem}
 \begin{proof}

### 2016-01-06 paper/related.tex
Switched to the llncs template. Reworked the abstract and intro.

--- 2016-01-06 paper/paper.tex
@@
-\documentclass{llncs}
-\usepackage[english]{babel}
-\usepackage{hyperref}
-\usepackage{cleveref}
-\usepackage{comment}
-\usepackage{datetime}
-\usepackage[url=true,doi=false,isbn=false,backend=biber]{biblatex}
-\usepackage{mathtools}
-\usepackage{paralist}
-\usepackage{mathtools}
-
-%\input{header.tex}
-\input{defs.tex}
-
-\addbibresource{managed.bib}
-
-\title{Composable Generalized Consensus}
-
-\author{
-  Rachid Guerraoui\\
-  \texttt{rachid.guerraoui@epfl.ch}
-  \and
-  Giuliano Losa\\
-  \texttt{giuliano@losa.fr}
-}
-
-\date{}
-
-\begin{document}
-
-\maketitle
-
-\input{abstract3}
-
-%\vspace{2cm}
-%\begin{center}
-%This is a regular paper.
-%\end{center}
-
-\newpage
-
-\input{intro}
-
-%\input{roadmap}
-
 \section{Related Work}
 \label{sec:related}
+
+Generalized Consensus is a specification of SMR algorithms that takes into account the commutativity relation among commands and allows its implementation to optimize the execution of commuting commands.
+In cases in which many commands commute, which may be frequent in practice, Generalized Consensus algorithms can have a significant performance advantage over traditional consensus algorithms.
+
+Making SMR algorithms efficient has been a long-standing
+challenge and a multitude of algorithms have been proposed, each best suited to a particular use case.
+Examples include Paxos~\cite{lamport2001paxos}, Fast
+Paxos~\cite{Lamport06FastPaxos}, Disk Paxos~\cite{GafniLamport03DiskPaxos},
+Generalized Paxos~\cite{Lamport05GeneralizeConsensus},
+FGGC~\cite{SutraShapiro11FastGenuineGeneralizedConsensus}, Chain
+Replication~\cite{RenesseSchneider04ChainReplicationSupportingHighThroughputAvailability}, Ring
+Paxos~\cite{MarandiETAL10RingPaxosHighthroughputAtomicBroadcastProtocol},
+Egalitarian Paxos
+\cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments},
+Multi-Coordinated Paxos~\cite{CamargosSchmidtPedone07MulticoordinatedPaxos},
+Cheap Paxos~\cite{LamportMassa04CheapPaxos}, Vertical Paxos
+\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}, Paxos-MIC
+\cite{HurfinMoiseNarzul11AdaptiveFastPaxosMakingQuickEverlasting}, Mencius
+\cite{MaoJunqueiraMarzullo08MenciusBuildingEfficientReplicatedStateMachine},
+Fast Mencius~\cite{WeiETAL13FastMenciusMenciusLowCommitLatency},
+the Chandra-Toueg algorithm
+\cite{ChandraToueg96UnreliableFailureDetectorsReliableDistributedSystems}, the
+Ben-Or algorithm~\cite{BenOr83AnotherAdvantageFreeChoiceCompletelyAsynchronous},
+and Raft~\cite{OngaroOusterhout14SearchUnderstandableConsensusAlgorithm}.
 
@@
 Generalized Consensus is an agreement problem relaxing the ordering constraints
-imposed on commuting commands and Generalized Paxos \cite{Lamport05GeneralizeConsensus} is an algorithm, similar
+imposed on commuting commands and Generalized Paxos~\cite{Lamport05GeneralizeConsensus} is an algorithm, similar
 in structure to Fast Paxos, which takes advantage of the relaxed ordering
@@
 algorithms and the incremental development of SMR optimizations.
-
-\section{Generalized Consensus}
-\label{sec:gc}
-
-We consider a set of asynchronous servers communicating by message passing in
-an asynchronous network. Servers are not necessarily sequential processes.
-We describe a behavior of the system using a sequence of states,
-where a state is a function from variable names to values (i.e.\ a valuation of
-variables). We specify sets of allowed behaviors using an initial predicate
-and a next-state relation expressed using unprimed variables and primed variables (next-state variables) in the
-style of TLA+ \cite{Lamport02SpecifyingSystems}. However, for succinctness, our
-specification are rather informal, eluding non-essential details.
-
-We consider a computing service exposing a set of commands $C$. The
-service has a sequential specification consisting of an initial state
-and a deterministic transition relation in which every command
-atomically updates the state of the service and produces an output. We
-assume that duplicate commands have no effect on the state of the
-service and return the same output as did their first occurrence.
-
-The goal of SMR is to provide a reliable implementation of the computing service despite server crashes.
-To do so, a SMR algorithm replicates the execution of the service on several servers. 
-Traditionally, the servers play one or
-several of three roles: \emph{proposers}, which propose new commands to execute,
-\emph{acceptors}, whose role is to order commands, and
-\emph{learners}, which learn about the growing sequence of commands to
-execute by querying the acceptors \cite{lamport2001paxos}. Often, but not necessarily, every replica server plays the
-three roles of proposer, acceptor, and learner: clients of the service
-send their commands to some replica servers; replica servers submit
-the received commands to the SMR protocol in the role of proposers;
-replica servers respond to commands by executing the commands learned
-in the role of learners; replica servers participate in command
-ordering in the role of acceptors.  A SMR algorithm guarantees that
-the replicas learn the same sequences of commands, possibly up to the
-reordering of commuting commands. Therefore, every learner can respond to client
-requests using its last learned sequence of commands, and clients will receive
-the same responses as the ones produces by a centralized, non-replicated,
-service.
-
-\subsection{Histories}
-
-To simplify the specification of SMR algorithms optimizing the execution of commuting commands 
-we will use the notion of a command history, or history for short, in our specification of Composable Generalized Consensus.
-
-A history is a data structure that
-represents a set of sequences which are the same modulo the reordering
-of commuting commands and addition or removal of duplicate commands.
-We say that two commands $c_1$ and $c_2$ \emph{commute} when, in every
-sequential execution of the service in which $c_1$ and $c_2$ are adjacent,
-reversing the order of $c_1$ and $c_2$ does not change any of the outputs,
-including those given to $c_1$ and $c_2$.
-
-By definition, all the sequences represented by a given history contain the same commands
-and determine the same outputs for each command. Therefore, to implement a
-replicated service, it is sufficient that learners learn histories instead of
-sequences.
-
-Command histories are built from the empty history, noted $\bot$, using the $\star$
-operator, which appends a sequence of commands $cs$ to a history $h$, obtaining
-the history $h \star cs$. We define a partial order relation among
-histories: $h_1 \leq h_2$ when there exists a sequence $cs$ such that $h_2 = h_1
-\star cs$. Histories have the property that every finite set $H$ of histories
-has a \emph{greatest lower bound}, noted $GLB\left(H\right)$. 
-We say that a history $h$ contains the command $c$ when there exists a sequence of commands 
-$cs$ containing $c$ such that $\bot \star cs = h$.
-Finally, we say
-that a set of histories $H$ is compatible when all the histories in $H$ have a
-common upper bound $h$ such that $h$ can be built using only commands contained in
-the histories in $H$.
-%The maximum over a set of histories $H$ is denoted by $Max\left( H \right)$ and $Max\left( \aset{} \right)$ is the empty history.
-
-The major advantage that histories have over sequences is that if a sequence
-$cs_1$ is obtained by reordering the commuting commands of $cs_2$, then $h \star
-cs_1 = h \star cs_2$. Therefore, in SMR implementations, servers can agree on
-a history even though they chose different orders for some commuting commands.
-
-For a thorough discussion of command histories we refer the reader
-to \cite{Lamport05GeneralizeConsensus}, where histories are called C-structs. 
-%Our Isabelle/HOL formalization abstracts over the concrete representation of histories, assuming only their
-%necessary properties. 
-Command histories were first introduced in trace theory by Mazurkiewicz 
-\cite{Mazurkiewicz84Semantics}.
-
-\subsection{Generalized Consensus}
-
-Generalized Consensus precisely specifies the allowed behaviors of the roles of \emph{proposers}
-and \emph{learners}, abstracting over the behavior of acceptors. Acceptors are used to implement Generalized Consensus but do not appear in its interface.
-
-We specify Generalized Consensus using the variables $proposed$, containing the proposed commands so far,
-and, for every learner $l$,  a variable $learned\left[ l \right]$, a set containing the histories learned by $l$ so far. Initially, $proposed = \aset{}$ and, for every learner $l$,
-$learned\left[ l \right] = \aset{}$.
-
-The next-state relation consists of two types of transitions:
-$Propose\left( l, c
-\right)$ transitions, for a learner $l$ and a command $c$, in which
-$c$ is inserted in the set $proposed$; $Learn\left( l,h \right)$
-transitions, for a learner $l$ and a history $h$, in
-which $h$ is inserted in the set $learned\left[ l \right]$. 
-A $Learn\left( l, h \right)$ transition is guarded by the following conditions:
-\begin{compactitem}
-    \item[\textbf{Agreement}:] The set of learned histories after the transition, $\aset{\bigcup learned'\left[ l
-      \right] : l \in Learners}$, is compatible.
-    \item[\textbf{Validity}:] Every learned history contains only
-        proposed commands.  
-    \item[\textbf{Irrevocability}:] The new learned history
-      $h$ is an extension of every history in $learned\left[ l \right]$. 
-        %\\ $cs > Max\left(learned\left[ l \right]\right)$.
-\end{compactitem}
-
-When clients of a replicated service submit their commands to the proposers and learners compute the response to give to clients using the maximum over the histories in $learned\left[ l \right]$, the three properties above imply that the clients see a \emph{linearizable} execution
-\cite{HerlihyWing90LinearizabilityCorrectnessConditionConcurrentObjects}.
-
-\section{Modular Implementations of Generalized Consensus}
-
-In this section we introduce our framework by example, building a modular Generalized Consensus algorithm inspired from Fast Paxos and Generalized Paxos by first designing a fast but fragile modular algorithm, then combining it unchanged with a more resilient algorithm.
-
-\subsection{The Broadcast-Decide Algorithm}
-
-In our first algorithm, Broadcast-Decide, learners each maintain a variable $accepted\left[ l \right]$ which contains the currently accepted history of learner $l$.
-For every learner $l$, $accepted\left[ l \right]$ is initialized to the empty history.
-A proposer broadcasts its proposal to all the learners which unconditionally accept it by appending the proposal to their $accepted\left[ l \right]$
-history and by broadcasting to all the acceptors their new accepted history, $accepted'\left[ l \right]$.
-
-Each learner $l$ maintains a variable $learned\left[ l \right]$, initially the empty history, containing its last learned history. A learner $l$
-learns a new history $h$ when $h \geq learned\left[ l \right]$ and $h$ is a lower bound of the set consisting of the largest history received
-from each acceptor.
-
-More precisely, the Broadcast-Decide algorithm is described by the transition relation obtained as the disjunction of the
-following transitions. We do not model the network explicitly using state variables but instead we just say that servers send, broadcast, or receive messages.
-
-\begin{compactitem}
-
-\item A $Propose\left( p,c \right)$ transition, for a proposer $p$ and a command $c$, not guarded, broadcasts the message $\aseq{\quo{prop},c}$ to all
-the acceptors.
-
-\item An $Accept\left( a \right)$ transition, for an acceptor $a$, enabled when $a$ can receive a $\aseq{\quo{prop},c}$ message from a proposer,
-updates $accepted\left[ a \right]$ to $accepted\left[ a \right]\star \aseq{c}$ and broadcasts the message $\aseq{\quo{accepted}, a, accepted'\left[ a \right]}$ to all the learners
-($accepted'\left[ a \right]$ is the new value of the variable $accepted\left[ a \right]$).
-
-\item A $Learn\left( l, h \right)$ transition, for a learner $l$ and a history $h$, enabled when $h \geq learned\left[ l \right]$ and $h$ is a lower
-bound of the set $\aset{MaxReceivedFrom\left( a \right) : a \in Acceptors}$, where $MaxReceivedFrom\left( a \right)$ is the maximal accepted history received
-from the acceptor $a$.
- Its effect is to set $learned\left[ l \right]$ to $h$.
-
-\end{compactitem}
-
-The algorithm Broadcast-Decide satisfies the specification of Generalized Consensus, however it cannot learn new histories if two acceptors $a_1$ and $a_2$ have two histories which are not compatible: since an acceptor can only append new proposals to its accepted history, all the possible future lower bounds of the set $\aset{MaxReceivedFrom\left( a \right) : a \in Acceptors}$ are bounded above by $GLB\left( \aset{accepted\left[ a_1 \right], accepted\left[ a_2 \right]} \right)$.
-This situation occurs when non-commuting commands contend: for example, when two different proposers send two commands $c_1$ and $c_2$ that do not commute, the acceptors $a_1$ accepts $c_1$ before $c_2$, and the acceptor $a_2 \neq a_1$ accepts $c_2$ before $c_1$.
-
-To overcome this deadlock, we will make some modifications to the algorithm that will allow us to start a fresh instance $B_2$, called a round, of the Broadcast-Decide algorithm.
-To allow switching to a fresh new round of the Broadcast-Decide algorithm, we introduce the new role of switcher, that any server can play additionally from being a proposer, acceptor, or learner.
-A switcher will extract an accepted history from the first round and try to initialize all the acceptors of the second round with the extracted history.
-If we are luckier than in the first round, there may be no contention between switchers and every acceptor $a$ will start accepting new proposals in the second round based on the same initial history.
-
-The first round $B_1$ and second round $B_2$ each have their own copies of the arrays $accepted$ and $learned$, noted $accepted_1$, $accepted_2$, $learned_1$, and $learned_2$, when the round is not clear from the context.
-We assume that network messages are tagged with a round number and that a message tagged with $i\in\aset{1,2}$ can only be received in $B_i$, creating two isolated logical networks.
-
-In our example with the deadlock, switchers have the role of initializing the new round of $B_2$ so that it can resume the execution and overcome the deadlock.
-We modify the Broadcast-Decide algorithm by introducing new transitions and modifying the $Accept\left( a \right)$ transition as follows.
-We add the new variables $status\left[ a \right]$ for every acceptor $a$, which are initialized to $\quo{idle}$, meaning that an acceptor has not been initialized.
-The purpose of $status\left[ a \right]$ is to prevent the acceptor $a$ from accepting commands until its status changes from $\quo{idle}$ to $\quo{ready}$ and to stop it when its status goes to $\quo{stopped}$.
-We also add the new arrays of variables $inits\left[ s \right]$ and $aborts\left[ s \right]$, for every switcher $s$.
-The variables $inits\left[ s \right]$ are initialized to the empty history in $B_1$ and to $\quo{none}$ in $B_2$, while the variables $aborts\left[ s \right]$ are initialized to $\quo{none}$ in both $B_1$ and $B_2$.
-The purpose of the arrays of variables $aborts$ and $inits$ is to transfer state between $B_1$ and $B_2$.
-The arrays of variables $aborts$ and $inits$ form what we call the \emph{inter-round} interface.
-The rounds $B_1$ and $B_2$ each have their own copy of the arrays of additional variables $status$, $inits$, and $aborts$, and we add subscripts denoting the round when the round is not clear from the context.
-
-\begin{compactitem}
-\item We modify the $Accept\left( a \right)$ transition by enabling it only if $status\left[ a \right] = \quo{ready}$ and by having the acceptor $a$ additionally broadcast its new accepted value to the switchers.
-Switchers detect that the algorithm is deadlocked when they have received two messages $\aseq{\quo{accepted}, a_1, h_1}$ and $\aseq{\quo{accepted}, a_2, h_2}$ where  $h_1$ and $h_2$ are incompatible.
-
-\item The $Stop\left( s \right)$ transition, for a switcher $s$, is enabled when the switcher has detected a deadlock or a crash (for example using a timeout).
-Its effect is to broadcast a $\aseq{\quo{stop}, s}$ message to the acceptors.
-
-\item The $Stop\left( a \right)$ transition, for an acceptor $a$, is enabled when the acceptor $a$ can receive a $\aseq{\quo{stop}, s}$ message from a switcher $s$.
-  Its effect is to update $status\left[ a \right]$ to $\quo{stopped}$ and to send the message $\aseq{\quo{stopped},a,accepted\left[ a \right]}$ to the switcher $s$.
-
-\item The $Abort\left( s, h\right)$ transition, for a switcher $s$, is enabled when $aborts\left[ s \right] = \quo{none}$ and the switcher has received
-  at least one $\aseq{\quo{stopped}, a, h}$ message.
-Its effect is to set the variable $aborts\left[ s \right]$ to the history $h$ received in the $\quo{stopped}$ message.
-The history $aborts'\left[ s \right]$ is then called an \emph{abort history}.
-
-\item The $Init\left( s,h \right)$ transition, for a switcher $s$, is enabled when $inits\left[ s \right]=\quo{none}$ and $h$ is a history, and its effect is to set $inits\left[ s \right]$ to $h$, and to broadcast the message $\aseq{\quo{init}, h}$ to the acceptors.
-The history $inits\left[ s \right]$ is then called an \emph{init history}.
-As described below, in $B_2$ composed with $B_1$, this history is constrained to be an abort history of $B_1$.
-
- \item The $WakeUp\left( a \right)$ transition, for an acceptor $a$, is enabled when $status\left[ a \right]$ is $\quo{idle}$ and $a$ can receive a $\aseq{\quo{init}, h}$ message from a switcher.
-Its effect is to set set $status\left[ a \right]$ to $\quo{ready}$ and $accepted\left[ a \right]$ to $h$.
-
-\end{compactitem}
-
-The composition $B_1\times B_2$ of the two rounds $B_1$ and $B_2$ is obtained as follows:
-\begin{itemize}
-  \item We Require that a transition of the composition be a transition of either specification but not both, except for the $Abort\left( s \right)$ and $Init\left( s \right)$ transitions, which are allowed to take place at the same time.
-  \item We require that $aborts_1 = inits_2$ at all times.
-\end{itemize}
-The composition models two independent rounds being linked only by the switchers passing the abort histories of $B_1$ to $B_2$, where they become init histories: switchers gather abort histories in $B_1$ thanks to the $Stop$ transitions, then they broadcast them as init histories to the acceptors in $B_2$.
-The acceptors then initialize their state to a received init history in the $WakeUp\left( a \right)$ transition.
-This process can be repeated as often as needed to switch to new rounds $B_3$, $B_4$, $B_5$, etc., each time a round cannot learn new histories anymore.
-
-Crucially, the composition of any number of rounds implements Generalized Consensus.
-To understand why, observe that in an $Abort\left( s,h \right)$ transition, the switcher $s$ has received a $\aseq{\quo{stopped},a,h}$ message from the acceptor $a$, which means that the acceptor $a$ no longer accepts new proposals and that $h$ is the current value of the $learned\left[ a \right]$ variable.
-Therefore no history greater than $h$ can ever be learned anymore in this round.
-An abort history is thus a safe
-starting point for a new round to resume execution.
-Since a new round of the Broadcast-Decide algorithm guarantees that a single init history will be used as a prefix a new learned histories, the properties of Generalized Consensus are preserved.
-
-The Broadcast-Decide algorithm can no more learn new histories whenever one acceptor crashes because the crashed acceptor prevents the lower bound of the accepted histories to ever increase.
-However, the round-change mechanism is very resilient to faults and can abort anyway thanks to the timeout triggering the $Stop\left( s \right)$ transition: a single live acceptor with a single live switcher can abort the execution and change round.
-
-To allow a new round to make progress after some acceptors crashed in a previous round, we introduce a component called the switching policy.
-With the Broadcast-Decide rounds, the task of the switching policy is to update the set of acceptors from one round to the next, ensuring that failed acceptors are removed or replaced by live ones.
-The switching policy must be queried by the switchers upon changing round to determine the new set of acceptors.
-This information can then be piggybacked to the
-acceptors and learners, or the acceptors and learners can directly query the switching policy. 
-The set of switchers can be changed from round to round in the same fashion.
-Moreover, in an algorithm composed of several different types of rounds, the switching policy will also be responsible for choosing the type of round to execute next.
-
-\subsection{The Leader Algorithm} 
-\label{sec:leader}
-
-The Broadcast-Decide algorithm can process a proposal in two message delays when there are no conflicts, i.e.\ no contention on non-commuting commands.
-However, if there are repeated conflicts, the algorithm may not make progress despite changing round, as the acceptors of the each new round
-may be initialized with incompatible histories. The Fast Paxos algorithm \cite{Lamport06FastPaxos} is able to learn new histories under high contention by
-designating a leader server that will totally order the proposals.
-However, to make our Broadcast-Decide algorithm work like Fast Paxos, we would have to modify it substantially.
-If we had spent effort testing Broadcast-Decide implementations and proving it correct, this effort would be wasted and we would have to redo all the tests and proofs from scratch.
-
-Instead, we will now show how to obtain an algorithm with performance characteristics close to those of Fast Paxos, employing a leader to deal with contention, without modifying at all our existing Broadcast-Decide algorithm.
-To do so, we define a new kind of rounds, the Leader rounds, which uses the same inter-round interface as the Broadcast-Decide algorithm. 
-The inter-round interface will allow us to combine Leader rounds and Broadcast-Decide rounds in any order, obtaining an algorithm which can adapt its strategy to its environment, selecting the Broadcast-Decide algorithm under low contention and switching to the slower Leader algorithm when there is contention.
-
-In the Leader algorithm, the proposers send their proposals to a distinguished acceptor called the leader. 
-The leader totally orders the proposals it receives and forwards them to all the acceptors which accept them in the order defined by the leader.
-As in the Broadcast-Decide algorithm, acceptors broadcast their accepted histories to the learners, which learn a new history when it is a lower bound of 
-the set $\aset{MaxReceivedFrom\left( a \right) : a \in Acceptors}$.
-Thanks to its leader, a Leader round continues to learn new histories even when proposals conflict because the leader ensures that the accepted histories of the acceptors remain consistent.
-
-The Leader algorithm treats all proposals equally and does not take advantage of commutativity.
-When there are no server crashes, each proposal is processed in three message delays.
-In comparison, in the absence of crashes, the Broadcast-Decide algorithm processes non-conflicting proposals in two message delays but can take arbitrarily long to process conflicting proposals.
-
-To initialize a Leader round, the switchers send their init histories to the leader, which chooses a unique one that it forwards to the acceptors, guaranteeing that all the acceptors are initialized with the same history even when conflicting init histories are sent concurrently.
-
-The Leader algorithm can not learn new histories if its leader, or any other acceptor, crashes.
-To recover from this deadlock we pass the baton to a new round, as in the Broadcast-Decide algorithm, with a similar mechanism: a switcher that times-out waiting for new messages from the acceptors broadcasts a stop message, waits for receiving the accepted history of an acceptor, and finally aborts using the received history as abort history.
-
-Let us specify more precisely the Leader algorithm.
-The specification uses the same variables as the Broadcast-Decide algorithm and starts in the same initial state. 
-The next-state relation is the disjunction of the following types of
-transitions. 
-\begin{compactitem}
-\item A $Propose\left( p,c \right)$ transition, for a proposer $p$
-  and a command $c$, not guarded, sends the message $\aseq{\quo{prop},c}$ to
-  the leader.
-\item An $Init\left( s, h \right)$ transition, for a switcher $s$
-  and a history $h$, is enabled when $inits\left[ s \right] = \quo{none}$, and its effect is to set $inits\left[ s \right]$ to $h$ and
-  to sends the message $\aseq{\quo{init},h}$ to the leader.
-\item A $WakeUp\left( leader \right)$ transition, enabled when
-  $status\left[ leader \right]$ is ``idle'' and 
-  a message $\aseq{\quo{init},h}$ was sent to the leader. Its effect is to
-  receive the message, set $accepted\left[ leader \right]$ to $h$,
-  to set $status\left[ leader \right]$ to ``ready'', and to
-  broadcast the message $\aseq{\quo{leader-accept},h}$ to all
-  other acceptors.
-\item An $Accept\left( leader \right)$ transition, enabled when
-  $status\left[ leader \right]$ is ``ready'' and a $\aseq{\quo{prop},c}$ message has been sent to the leader. The effect of the
-  action is to receive the message, to update $accepted\left[ leader
-  \right]$ to $accepted\left[ leader \right]\star c$,  to broadcast the message
-  $\aseq{\quo{leader-accept},accepted'\left[ leader \right]}$
-  to all the other acceptors, and to
-  broadcast the message $\aseq{\quo{accepted},accepted'\left[ leader
-  \right]}$ to the learners.
-\item An $Accept\left( a \right)$ transition, for an acceptor $a$
-  which is not the leader, is enabled when $status\left[ a
-  \right]$ is ``ready'' and the leader has sent a
-  $\aseq{\quo{leader-accept},h}$ message to $a$, where $h > accepted\left[ a
-  \right]$. The effect of the action is to receive the message,
-  to update $accepted\left[ a \right]$ to $h$, and to broadcast the
-  message $\aseq{\quo{accepted},h}$ to the
-  learners.
-\item A $Learn\left( l, h \right)$ transition, for a learner $l$
-  and a history $h$, is enabled when $h \geq learned\left[ l \right]$ and $h$ is a lower
-  bound of the set $\aset{MaxReceivedFrom\left( a \right) : a \in Acceptors}$, where $MaxReceivedFrom\left( a \right)$ is the maximal accepted history received
-  from the acceptor $a$.  Its effect is to set $learned\left[ l \right]$ to $h$.
-\item A $Stop\left( s \right)$ transition, for a switcher $s$,
-  enabled when the switcher $s$ has detected a crash (for example using a timeout), broadcasts a $\aseq{\quo{stop}, s}$ message to all
-  the acceptors. 
-\item A $Stop\left( a \right)$ transition, for an acceptor $a$,
-  enabled when $a$ can receive a $\aseq{\quo{stop},s}$ message
-  from a switcher $s$. Its effect is to receive the message, set
-  $status\left[ a \right]$ to ``stopped'', and to send the
-  message $\aseq{\quo{stopped},a,accepted\left[ a \right]}$ to the
-  switcher $s$. 
-\item An $Abort\left( s, h \right)$ transition, for a switcher $s$ and a history $h$, is enabled when $aborts\left[ s \right] = \quo{none}$ and
-  the switcher has received at least one $\aseq{\quo{stopped}, a, h}$ message. Its effect is to set the variable $aborts\left[ s \right]$ to the
-  history $h$ received in the $\quo{stopped}$ message.
-\end{compactitem}
-
-Thanks to the inter-round interface, which is the same for Leader rounds and Broadcast-Decide rounds, new Leader rounds and Broadcast-Decide rounds can be repeatedly composed in a sequence.
-
-Observe that a Leader round maintains a similar invariant, relating abort histories to learn histories, than the BroadCast-Decide algorithm:
-the learned histories are necessarily smaller than the abort histories in the partial order on histories. In both types of rounds, an abort history is
-a safe starting point for initializing the variables $accepted\left[ a \right]$ of the acceptors. Moreover, both types of rounds guarantee that only a
-single history will be used as a basis for the learners to learn new histories.
-These common invariants make the two kinds of rounds compatible, so that we can compose them and obtain an algorithm implementing Generalized Consensus.
-
-In the next section we present Composable Generalized Consensus, a specification of the behavior of rounds at the level of proposals, learned histories, init histories, and abort histories, which allows making the informal reasoning above precise.
-Our composition theorem ensures that any set of rounds individually satisfying Composable Generalized Consensus can be composed in any order to yield an algorithm satisfying Composable Generalized Consensus.
-Therefore we can devise many other types of rounds and compose them unchanged with our two BroadCast-Decide and Leader rounds.
-
-\section{Composable Generalized Consensus}
-\label{sec:aca}
-
-We consider a set of servers playing one or multiple of the four roles of proposer, acceptor, learner, and switcher. 
-Composable Generalized Consensus specifies the allowed behaviors of the proposers, learners, and switchers
-\emph{of a single round}. 
-To specify Composable Generalized Consensus, we use the variable $proposed$, a set of proposals, the array of set of histories $learned\left[ l \right]$, where $l$ ranges of learners, the variable $inits$, a set containing the init histories received from the previous
-round, and the variable $aborts$, a set containing the abort histories transfered to the next round.
-Initially all variables are empty sets.
-
-The next-state relation
-consists of four types of transitions. The first two, $Propose\left( l, c \right)$, and $Learn\left( l, h \right)$, for a learner $l$ and a
-history $h$, have the same effect as in the specification of Generalized Consensus: 
-$Propose\left( l,c \right)$ adds $c$ to the set $proposed$ and $Learn\left( l,h \right)$ adds $h$ to the set $learned\left[ l \right]$.
-
-Composable Generalized Consensus introduces two new transitions: $Init\left( s, h \right)$ and $Abort\left( s, h
-\right)$, for a switcher $s$ and a history $h$. The transition $Init\left( s,h \right)$ inserts $h$ in the set $inits$, and models the switcher $s$ receiving
-the abort history $h$ from the preceding round. The transition $Abort\left( s, h \right)$ inserts $h$ in the set $aborts$ and models the switcher $s$ passing the abort history $h$ to the next round.  
-All the transition are guarded by the following conditions:
-\begin{compactitem}
-    \item[\textbf{Agreement}:] The set $\aset{\bigcup learned'\left[ l
-      \right] : l \in Learners}$ is compatible.
-    \item[\textbf{Validity}:] There is a history $h$ belonging to $inits$ such that every abort history or learned history 
-        is of the form $h \star cs$, where $cs$ contains only commands
-        of the set $proposed$.
-    \item[\textbf{Initialization}:] If $inits$ is empty, then no
-        history can be learned and no switcher can abort.
-    \item[\textbf{Irrevocability}:] In the case of a $Learn\left( l,h
-        \right)$ transition, the new learned history $h$ is an 
-        extension of every history in $learned\left[ l \right]$.
-    \item[\textbf{Safe Abort}:] Every abort history $h$ is an upper bound of the set of learned histories.
-\end{compactitem}
-
-Note that if a round is initialized with empty abort histories, then it implements Generalized Consensus.
-
-We compose two rounds of Composable Generalized Consensus as we did in the preceding section.
-We write $proposed_2$, $learned_2$, $inits_2$, and $aborts_2$ for the variables of the first round and  $proposed_2$, $learned_2$, $inits_2$, and $aborts_2$
-for the variables of the second round.
-The transition relation of the composition is the disjunction of the following actions, conjoined with the condition that $aborts_1=inits_2$:
-\begin{itemize}
-  \item A $Propose$ or $Learn$ action of round 1, leaving the variables of round 2 unchanged.
-  \item A $Propose$ or $Learn$ action of round 2, leaving the variables of round 1 unchanged.
-  \item An $Init$ action of round 1, leaving the variables of round 2 unchanged.
-  \item An $Abort$ action of round 2, leaving the variables of round 1 unchanged.
-  \item A joint action consisting of an $Abort$ action of round 1 together with an $Init$ action of round 2.
-\end{itemize}
-
-\subsection{The Composition Theorem}
-
-\begin{theorem}[Composition Theorem]
-    \label{thm:comp}
-Consider the composition of two rounds as defined above.
-Define the state variable $proposed$ as $proposed_1 \cup proposed_2$. For every learner $l$, define 
-$learned\left[ l \right]$ as the maximum of $learned_1\left[ l \right]$ and $learned_2\left[ l \right]$.
-For every switcher $s$, define $inits\left[ s \right]$ as $inits_1\left[ s \right]$ and $aborts\left[ s \right]$ as $aborts_2\left[ s \right]$.
-
-Then the sequence of states of the variables $proposed$, $learned$, $inits$, and $aborts$ satisfy the specification of Composable Generalized Consensus.
-\end{theorem}
-
-The composition theorem can be applied recursively to show that the composition of any number of rounds which individually and independently satisfy Composable Generalized Consensus itself satisfies Composable Generalized Consensus.
-Therefore SMR algorithms implemented as rounds in our framework can be reused unmodified and optimized by adding new types of rounds incrementally, building on existing work instead of having to redo all the tests and proofs from scratch every time one adds a new optimization to an existing algorithm.
-
-However, it may not be clear that most SMR algorithms can be expressed as the composition of one or more types of rounds satisfying Composable Generalized Consensus.
-
-We provide a mechanically-checked proof of the composition theorem in Isabelle/HOL at \url{http://losa.fr/cgc}.
-The proofs uses a version of the specification of CGC that has a more
-executable form that presented in this section. 
-The Isabelle/HOL formalization of CGC uses the theory of I/O automata, which
-is also formalized in Isabelle/HOL\@. The proof uses a refinement
-mapping from the composition of two consecutive CGC instances to a single CGC
-instance.
-
-\begin{comment}
-\begin{theorem}[Composition Theorem]
-    \label{thm:comp}
-    In every execution of a succession of CGC instances, the history of valuations
-    of the variable $proposed$ and, for every learner $l$, of the variables
-    $learned\left[ l \right]$ satisfies Generalized Consensus.
-\end{theorem}
-\Cref{thm:comp} can be shown by induction on the number of rounds,
-using the following lemma.
-\begin{lem}
-  \label{lem:comp}
-    Consider two rounds numbered $i$ and $i+1$. In every execution of a
-    succession of CGC instances, the history of valuations of the variables $inits
-    = inits_i$, $propose = proposed_i \cup propose_{i+1}$, for every learner
-    $l$, $learned\left[ l \right] = learned_i\left[ l \right] \cup
-    learned_{i+1}\left[ l \right]$and $aborts = aborts_{i+1}$ satisfies CGC.
-\end{lem}
-\begin{proof}
-  A mechanically-checked proof in Isabelle/HOL appears in \cref{sec:isaproofs}.
-  The proofs uses a version of the specification of CGC that has a more
-  executable form that presented in this section. Moreover,
-  instead of history of commands, the Isabelle/HOL specification uses the more
-  general notion of command history, presented in the next subsection.
-  The Isabelle/HOL formalization of CGC uses the theory of I/O automata, which
-  is also formalized in the Isabelle/HOL\@. The proof uses a refinement
-  mapping from the composition of two consecutive CGC instances to a single CGC
-  instance.
-\end{proof}
-
-\Cref{thm:comp} guarantees that the successive composition of any
-number of CGC instances implements Generalized Consensus.
-Therefore, with CGC, one can easily build replicated services that
-dynamically adapt to the current operating conditions of the system: a
-replicated service has several types of rounds at its disposal, each being
-tailored to particular conditions. Each time the conditions change,
-the current CGC instance aborts and passes the baton to a more 
-appropriate type of round, as determined by the scheduling policy. Moreover, a new type
-of CGC instance can always be devised after the system is deployed and added on the fly to
-the existing types of rounds.
-
-We now describe two important practical optimizations of CGC.
-
-\subsection{Optimizing the Execution of Read-Only Commands}
-\label{sec:readonly}
-
-Generalized Consensus can be further optimized by treating 
-read-only commands specially, as in Generalized Lattice Agreement
-\cite{FalerioETAL12GeneralizedLatticeAgreement}.
-
-We consider the case in which a set of servers each play the four
-roles of proposer, acceptor, learner, and switcher.
-Clients sends their commands to any server, which proposes them as a
-proposer, and records the command as pending.
-%When a server learns a history that contains one of its pending
-%commands, it uses the history to determine the response to the pending
-%command and forwards the response to the appropriate client.
-
-As observed in \cite{FalerioETAL12GeneralizedLatticeAgreement}, read-only
-commands can be executed faster by directly reading from the last learned
-history, instead of proposing the read commands and waiting for it to appear in
-a future learned history.  Because learned histories form a chain, all the reads
-return consistent values and this approach yields a \emph{serializable}
-implementation of the service.
-
-However linearizability is not satisfied if reads are served directly
-from the last learned history: a lagging replica server could respond
-to a read using a history which does not contain all completed update
-commands. To make the executions linearizable without resorting to
-proposing read-only commands as normal commands, we can employ the same
-technique as in \cite{FalerioETAL12GeneralizedLatticeAgreement}: upon
-receiving a read command, a server proposes a special no-op command
-that commutes with all other commands; the server then responds to the
-read command using the first learned history containing the no-op
-command. This ensures that all the commands that had completed before
-the read command was issued are present in the history.  Moreover,
-since no-op commands commute with all other commands, they can be
-treated more efficiently than regular read commands by the acceptors,
-like, for example, in a fast round of Fast Paxos.
-
-\end{comment}
-
-\section{Fast and Resilient Rounds}
-\label{sec:rounds}
-
-In this section we demonstrate how a large class of SMR algorithms, those based on quorum systems \cite{GuerraouiVukolic10RefinedQuorumSystems}, can be expressed as the composition of one or more Composable Generalized Consensus rounds.
-We propose two specifications, \emph{resilient CGC
-rounds} and \emph{fast CGC rounds}, which can be refined to obtain executable rounds.
-For example, CGC versions of Classic Paxos and Chain
-Replication can be obtained by refining the specification of
-resilient CGC round, and CGC versions of the fast rounds of Fast Paxos can be obtained by refining the specification of fast CGC rounds.
-We provide TLA+ specifications precisely demonstrating how to do so. The refinements have been checked with the TLC model-checker.
-The specifications are available at \url{http://losa.fr/cgc}.
-
-Fast and resilient CGC round are two specifications, at an
-intermediate level of abstraction, which satisfy CGC\@. Both
-specifications describe how the state of the acceptors should be
-updated, using the notion of quorum, but abstract over the particular communication topology used
-to implement the state accesses: the specifications use a global state
-which is accessible by all servers in all roles. One can refine fast
-or resilient rounds by implementing the state accesses and coordination
-using the network, obtaining a concrete algorithm.
-
-Resilient and fast CGC round are specified using the variables $proposed$, $aborts$, $inits$, $learned\left[ l \right]$, for each learner $l$, 
-and, for every acceptor $a$, the variable
-$status\left[ a \right]$, which can be either ``idle'', ``ready'', or
-``stopped'' and $hist\left[ a \right]$, the local accepted history of the
-acceptor $a$, which can be either a history or the special value $None$.
-Initially, $inits$ and $aborts$ are empty sets, and, for every learner
-$l$, $learned\left[ l \right]$ is the empty history, and, for every acceptor $a$,
-$status\left[ a \right]$  is ``idle'' and $hist\left[ a \right]$ is the special
-value $None$.
-
-The next-state relations of resilient and fast rounds are composed
-of the following types of transitions. We use the notion of \emph{learn quorum} and \emph{abort quorum}, defined shortly after.
-\begin{compactitem}
-    \item A $Propose\left( p,c \right)$ transition, for a proposer $p$
-        and a command $c$, not guarded, inserts $c$ in the set
-        $proposed$. 
-    \item An $Init\left( s, h \right)$ transition, for a switcher $s$
-        and a history $h$, not guarded, inserts $h$ in the set $inits$.
-    \item A $WakeUp\left( a \right)$ transition, for an acceptor $a$,
-        is enabled when $status\left[ a \right]$ is ``idle''. Its
-        effect is to set $hist\left[ a \right]$ to a history found in
-        $inits$ and to set $status\left[ a \right]$ to ``ready''.
-    \item An $Accept\left( a \right)$ transition, for an acceptor $a$,
-        is enabled when $status\left[ a \right]$ is ``ready'', and
-        updates $hist\left[ a \right]$ to $hist\left[ a \right]\star
-        c$, where $c$ is a proposed command. In resilient rounds,
-        the transition is additionally guarded by the condition that
-        the resulting set of local acceptor histories,
-        $\aset{hist'\left[ a \right] : a \in Acceptors}$, forms a chain
-        in the partial order $\leq$ (this is typically implemented with a
-        leader).
-    \item A $Learn\left( l, h \right)$ transition, for a learner $l$
-      and a history $h$, enabled when $h \geq learned\left[ l \right]$ and is
-      smaller than the GLB of the histories of a learn quorum of acceptors. The effect of
-        the transition is to set $learned\left[ l \right]$ to $h$.
-    \item A $Stop\left( a \right)$ transition, for an acceptor $a$,
-        enabled when $status\left[ a \right]$ is ``ready'', sets
-        $status\left[ a \right]$ to ``stopped'' (note that
-        $status\left[ a \right] = $ ``stopped'' prevents any further
-        ``accept'' transition of acceptor $a$).
-%    \item REALLY NEEDED? A $TimeOut\left( a \right)$ transition, for an acceptor $a$,
-%        is enabled when $status\left[ a \right]$ is ``idle'' and sets 
-%        $status\left[ a \right]$ to ``stopped'' and $hist\left[ a
-%        \right]$ to the special value $None$. 
-    \item An $Abort\left( s, h \right)$ transition, for a switcher $s$
-        and a history $h$, is enabled when there exists an abort quorum
-        $R$ such that, for every member $a$ of $R$, $status\left[ a
-        \right]$ is ``stopped'' and $h \in SafeAborts\left( R
-        \right)$. Its effect is to insert  $h$ in the set $aborts$. 
-\end{compactitem}
-
-On top of the additional guard on the $Accept$ transition present only
-in resilient rounds, resilient and fast CGC round differ in
-their definition of learn quorums, abort quorums, and in their
-definition of the $SafeAborts\left( R \right)$ operator.
-To concisely define read and learn quorums, we define 
-\begin{equation}
-    Q\left( f \right) = \aset{as \subseteq Acceptors :
-        Card\left(as\right) \geq \left\lfloor f\times Card\left(
-    Acceptors \right)\right\rfloor + 1}.
-\end{equation}
-
-In resilient CGC round, abort and learn quorums are defined such
-that the intersection between two learn quorums is nonempty and the
-intersection between a learn quorum and an abort quorum is nonempty.
-The definition leads to two interesting cases: $LearnQ = AbortQ =
-Q\left( 1/2 \right)$; $LearnQ = \aset{Acceptors}$ and $AbortQ =
-\aset{\aset{a} : a \in Acceptors}$. The operator $SafeAborts\left( R
-\right)$ is defined as the singleton set containing the maximum
-element over the local histories of the acceptors in $R$ which are not
-$None$. Note that the maximum exists because resilient rounds
-guarantee that the histories of the acceptors form a chain.
-
-In fast CGC round, abort and learn quorums are defined such that
-the intersection between any two learn quorums is nonempty and the
-intersection between any two learn quorums and an abort quorum is
-nonempty.  The definition leads to three interesting cases: $LearnQ =
-AbortQ = Q\left( 2/3 \right)$; $LearnQ = Q\left( 1/2 \right)$ and
-$AbortQ = Q\left( 3/4 \right)$; $LearnQ = \aset{Acceptors}$ and
-$AbortQ = Q\left( 0 \right)$.
-The operator $SafeAborts\left( R \right)$ is defined as the set of
-maximal elements over the set 
-\begin{equation}
-\aset{GLB\left( \aset{hist\left[ a
-\right] : a \in R \cap L \wedge hist\left[ a \right]\neq None} \right)
-: L \in LearnQ}.
-\end{equation}
-
-The cases in which abort quorums are singleton sets, in both
-resilient and fast rounds, allow tolerating
-$f$ faults with $f+1$ servers: a round can abort as long as one
-acceptor is correct and a new round, possibly using new correct
-servers, can take over. Note that changing round cannot happen
-without the scheduling policy instructing the switchers about the
-identity of the next round and of the new servers replacing the failed ones, and thus tolerating $f$ faults with
-$f+1$ servers relies on the assumption of a reliable scheduling policy.
-
-Define, for every learner $l$, the history variable $\overline{learned}\left[ l \right]$ which is initialized to the empty set and which is
-updated each time a new local history is accepted by the learner $l$ by
-inserting the new local history in $\overline{learned}\left[ l \right]$ (see \cite{AbadiLamport91ExistenceRefinementMappings} for the
-definition of a history variable).
-\begin{theorem}
-  \label{thm:rounds}
-    Both resilient and fast CGC round implement CGC under the refinement
-    mapping that consists in substituting $\overline{learned\left[ l \right]}$
-    for $learned\left[ l \right]$.
-\end{theorem}
-
-\begin{comment}
-\section{Examples of Concrete CGC Algorithms}
-\label{sec:examples}
-
-In this section we show how to modify the rounds of Classic Paxos,
-Multi-Coordinated Paxos, Chain Replication, Ring Paxos, and Fast Paxos to obtain
-CGC round. \Cref{thm:comp} then guarantee that all those different types of
-rounds can be used in the same algorithm to yield an implementation of Chain
-Agreement.
-
-\subsection{Classic Paxos}
-
-The rounds of the Classic Paxos algorithm can be modified so as to
-obtain an implementation of resilient CGC round.  We define
-Classic Paxos CGC round by specializing the specification of
-resilient CGC round, giving a concrete protocol for disseminating
-the proposed commands to the acceptors and enforcing that the learned
-histories form a chain.
-
-Each Classic Paxos round has a unique leader among the acceptors which
-centrally orders the proposed commands, which are accepted by the
-other acceptors in the order determined by the leader.  This protocol
-ensures that every local acceptor history is a prefix of the local
-history of the leader, therefore guaranteeing that the local acceptor histories
-forming a chain.
-
-Acceptor broadcast each newly accepted history to the learners, which
-learn a history $h$ when they receive enough extensions of $h$ from the
-acceptors.  When instructed by the scheduling policy to abort, a
-switcher sends a stop message to the acceptors, which irrevocably stop
-updating their local history and send back to the switcher an
-acknowledgement containing their local history. A switcher
-determines an abort history once it has received acknowledgments from
-an abort quorum of acceptors $R$, using the operator $SafeAborts\left(
-R \right)$.
-
-The specification of Classic Paxos CGC round uses the variables of
-the specification of resilient CGC round, namely the variables
-$proposed$, $inits$, $aborts$, the arrays $learned\left[ l \right]$,
-$status\left[ a \right]$, and $hist\left[ a \right]$, and two
-additional arrays $accepted\left[ l \right]\left[ a \right]$ and
-$stop\left[ s \right]\left[ a \right]$, where $l$ is a learner, $a$ is
-an acceptor, and $s$ is a switcher.
-
-There is also a network component that we do not describe explicitly.
-However, the network allows any servers to communicate by sending and
-receiving messages.
-
-Initially, the variables $proposed$, $inits$, $aborts$, and, for every
-learner $l$, $learned\left[ l \right]$ are empty sets. For every
-acceptor $a$, $status\left[ a \right]$ is ``idle'' and $hist\left[ a
-\right]$ is the special value $None$. For every learner $l$ and acceptor $a$,
-$accepted\left[ l \right]\left[ a \right]$ is the special value
-$None$. For every switcher $s$ and acceptor $a$, $stopped\left[ s
-\right]\left[ a \right]$ is set to $false$.  There are initially no
-messages in the network.
-
-The next-state relation is composed of the following types of
-transitions, where the definitions of learn quorums, abort quorums,
-and of the $SafeAborts\left( R \right)$ operator are the ones of resilient
-CGC round.
-\begin{compactitem}
-    \item A $Propose\left( p,c \right)$ transition, for a proposer $p$
-        and a command $c$, not guarded, inserts $c$ in the set
-        $proposed$ and sends the message $\aseq{\quo{prop},c}$ to
-        the leader.
-    \item An $Init\left( s, h \right)$ transition, for a switcher $s$
-        and a history $h$, not guarded, inserts $h$ in the set $inits$ and
-        sends the message $\aseq{\quo{init},h}$ to the leader.
-    \item A $WakeUp\left( leader \right)$ transition, enabled when
-        $status\left[ leader \right]$ is ``idle'' and the leader can
-        receive a $\aseq{\quo{init},h}$ message. Its effect is to
-        receive the message, set $hist\left[ leader \right]$ to $h$,
-        to set $status\left[ leader \right]$ to ``ready'', and to
-        broadcast the message $\aseq{\quo{leader-accept},h}$ to all
-        other acceptors.
-    \item An $Accept\left( leader \right)$ transition, enabled when
-        $status\left[ leader \right]$ is ``ready'' and the leader can
-        receive a $\aseq{\quo{prop},c}$ message. The effect of the
-        action is to receive the message, to update $hist\left[ leader
-        \right]$ to $hist\left[ leader \right]\star c$, and to
-        broadcast the message $\aseq{\quo{ack},hist\left[ leader
-        \right]\star c}$ to the learners, and to broadcast the message
-        $\aseq{\quo{leader-accept},hist\left[ leader \right]\star c}$
-        to all the other acceptors.
-    \item An $Accept\left( a \right)$ transition, for an acceptor $a$
-        which is not the leader, is enabled when $status\left[ a
-        \right]$ is ``ready'' and $a$ can receive a
-        $\aseq{\quo{leader-accept},h}$ message where $h > hist\left[ a
-        \right]$. The effect of the action is to receive the message,
-        to update $hist\left[ a \right]$ to $h$, and to broadcast the
-        message $\aseq{\quo{ack},hist\left[ a \right]\star c}$ to the
-        learners.
-    \item A $RcvAck\left( l \right)$ transition, for a learner $l$, is
-        enabled when $l$ can receive a message $\aseq{\quo{ack},h}$
-        from an acceptor $a$. Its effect is to receive the message and
-        to set $accepted\left[ l \right]\left[ a \right]$ to $h$.
-    \item A $Learn\left( l, h \right)$ transition, for a learner $l$
-        and a history $h$, is enabled when $h > Max\left(
-        learned\left[ l \right] \right)$ and there is a learn quorum
-        $Q$ of acceptors such that $h = GLB\left( \aset{accepted\left[
-        l \right]\left[ a \right] : a \in Q }\right)$. Its effect is
-        to insert $h$ in $learned\left[ l \right]$.
-    \item A $Stop\left( s \right)$ transition, for a switcher $s$,
-        always enabled, broadcasts a $\aseq{\quo{stop}}$ message to all
-        the acceptors.  This transition models a switcher being
-        instructed by the scheduling policy to abort the current round.
-    \item A $Stop\left( a \right)$ transition, for an acceptor $a$,
-        enabled when $a$ can receive a $\aseq{\quo{stop}}$ message
-        from a switcher $s$. Its effect is to receive the message, set
-        $status\left[ a \right]$ to ``stopped'', and to send the
-        message $\aseq{\quo{stopped},hist\left[ a \right]}$ to the
-        switcher $s$.
-    \item A $RcvStopped\left( s \right)$ transition, for a switcher
-        $s$, enabled when $s$ can receive a $\aseq{\quo{stopped},h}$
-        message from an acceptor $a$. Its effect is to receive the
-        message and to set $stopped\left[ s \right]\left[ a \right]$
-        to $h$.
-    \item An $Abort\left( s, h \right)$ transition, for a switcher $s$
-        and a history $h$, is enabled when there exists an abort quorum
-        $R$ such that, for every member $a$ of $R$, $stopped\left[ s
-        \right]\left[ a \right]$ is not $false$ and $h \in
-        SafeAborts\left( R \right)$, where $SafeAborts\left( R
-        \right)$ is computed using the array $stopped\left[ a
-        \right]$. Its effect is to insert  $h$ in the set $aborts$. 
-\end{compactitem}
-Compared to the rounds of the original Classic Paxos, Classic Paxos CGC
-rounds add an extra level of indirection when changing round: in the
-original rounds, the leader of the new round directly queries the
-acceptors of the preceeding round, whereas in CGC round, the
-switchers query the acceptors and then send their abort histories to
-the leader. However, changing round is an infrequent event, therefore
-this difference should not have much practical impact on performance.
-\begin{theorem}
-  Classic Paxos CGC round satisfy Composable Generalized Consensus.
-\end{theorem}
-\begin{proof}
-  A Classic Paxos CGC instance implements the specification of Abortable
-  Generalized Consensus under the refinement mapping consisting in
-  projecting the state of the round onto the variables $proposed$,
-  $inits$, $aborts$, and, for every learner $l$, $learned\left[ l
-  \right]$.
-\end{proof}
-
-\end{comment}
-
-\subsection{Refining Fast and Resilient Rounds}
-
-Classic Paxos \cite{lamport2001paxos}, Multi-Coordinated Paxos \cite{CamargosSchmidtPedone07MulticoordinatedPaxos}, Chain Replication \cite{RenesseSchneider04ChainReplicationSupportingHighThroughputAvailability}, and Ring Paxos \cite{MarandiETAL10RingPaxosHighthroughputAtomicBroadcastProtocol} refine Resilient Rounds, using a leader to ensure that the acceptor's accepted histories remain compatible.
-They differ in the way the leader is
-implemented and in the way that the histories of the leader are disseminated to
-the other acceptors.
-
-The rounds of Classic Paxos are similar to the Leader algorithm of \cref{sec:leader} but
-tolerate that a minority of servers crash by using majority quorums instead of
-waiting for histories from all the acceptors.
-
-Multi-Coordinated Paxos avoids depending on a unique leader by using a
-distributed leader implementation that can make progress as long as a strict
-majority of the acceptors are correct, trading off resilience for a higher time
-complexity.
-
-In Chain Replication, acceptors are arranged in a chain whose head is
-the leader of the round. As in Classic Paxos, abort histories and
-proposed commands are sent to the leader. However, instead of the
-leader directly sending its local history to all the other acceptors,
-the local history of the leader is
-forwarded along the chain from one acceptor to the other until it
-reaches the last acceptor, which sends the history to the learners. 
-By spreading the load more evenly on the all the servers, Chain Replication
-can achieve higher throughput that Classic Paxos.
-
-Ring Paxos optimizes Chain Replication by having the leader broadcast
-commands to the acceptors using IP multicast and only sending hashes
-of the commands along the chain. Ring Paxos rounds can be modified to implement
-resilient CGC round in a similar way as for Chain Replication.
-
-The fast rounds of Fast Paxos \cite{Lamport06FastPaxos} and Generalized Paxos \cite{Lamport05GeneralizeConsensus} refine Fast Rounds.
-They allow the proposers to directly broadcast their proposal to the 
-acceptors and make progress as long as non-commuting proposals do not contend.
-They use the same quorums as our Fast Rounds to ensure that it is possible to recover from inconsistencies 
-in the acceptors histories.
-Their structure is similar to the one of our Broadcast-Decide round.
-
-\begin{comment}
-\subsection{Fast Paxos}
-
-Fast Paxos is composed of two types of rounds: classic rounds, like in Classic
-Paxos, and fast rounds.
-The fast rounds can be transformed into
-implementations of fast CGC instances. Instead of sending their
-commands to a leader, as in Classic Paxos, the proposers of fast
-rounds directly broadcast their commands to all the acceptors. As long
-as commands are received in the same order by a learn quorum of
-acceptors or are received in different orders but commute, commands can be learned in two message-propagation delays instead of three in Classic Paxos.
-However, if non-commuting commands are received in different orders by
-several acceptors, the local history of the acceptors may not form a
-chain anymore and the rounds must abort. The fast rounds of Fast
-Paxos implement fast CGC instances.
-
-Fast Paxos CGC instances are specified using the same set of variables as for
-Classic Paxos, initialized to the same values.
-The next-state relation is composed of the following types of
-transitions, where the definitions of learn quorums, abort quorums,
-and of the $SafeAborts\left( R \right)$ operator are the ones of fast
-CGC instances.
-\begin{compactitem}
-    \item A $Propose\left( p,c \right)$ transition, for a proposer $p$
-        and a command $c$, not guarded, inserts $c$ in the set
-        $proposed$ and broadcasts the message $\aseq{\quo{prop},c}$ to
-        all the acceptors.
-    \item An $Init\left( s, h \right)$ transition, for a switcher $s$
-        and a history $h$, not guarded, inserts $h$ in the set $inits$ and
-        broadcasts the message $\aseq{\quo{init},h}$ to all the acceptors.
-    \item A $WakeUp\left( a \right)$ transition, for an acceptor $a$,
-        is enabled when $status\left[ a \right]$ is ``idle'' and $a$
-        can receive an $\aseq{\quo{init},h}$ message. Its effect is to
-        receive the message, set $hist\left[ a \right]$ to $h$, and to
-        set $status\left[ a \right]$ to ``ready''.
-    \item An $Accept\left( a \right)$ transition, for an acceptor $a$,
-        is enabled when $status\left[ a \right]$ is ``ready'' and $a$
-        can receive a $\aseq{\quo{prop},c}$ message. The effect of the
-        action is to receive the message, to update $hist\left[ a
-        \right]$ to $hist\left[ a \right]\star c$, and to broadcast
-        the message $\aseq{\quo{ack},hist\left[ a \right]\star c}$ to
-        the learners.
-    \item The $RcvAck\left( l \right)$ transitions, for a learner $l$,
-        the $Learn\left( l, h \right)$ transitions, for a learner $l$
-        and a history $h$, the $Stop\left( s \right)$ transitions, for
-        a switcher $s$, $Stop\left( a \right)$ transitions, for an
-        acceptor $a$, the $RcvStopped\left( s \right)$ transitions,
-        for a switcher $s$, and the $Abort\left( s, h \right)$
-        transitions, for a switcher $s$ and a history $h$, which are
-        all the same as in Classic Paxos CGC instances.
-\end{compactitem}
-
-As for Classic Paxos CGC instances, Fast Paxos CGC instances implement CGC
-under the refinement mapping consisting projecting the state of the
-round onto the variables $proposed$, $inits$, $aborts$, and, for every
-learner $l$, $learned\left[ l \right]$.
-
-\end{comment}
-
-\newpage
-
-\printbibliography
-
-\begin{appendix}
-
-  \label{sec:isaproofs}
-
-%  \includepdf[pages={1},pagecommand={\section{Mechanical Proof of the
-%  Composition Theorem in Isabelle/HOL}\label{sec:isaproofs}},scale=1]{../IsabellePDFS/podctheories.pdf}
-%  \includepdf[pages={2-},pagecommand={},scale=1]{../IsabellePDFS/podctheories.pdf}
-
-\end{appendix}
-
-\end{document}

### 2016-01-07 paper/abstract4.tex
Work on the abstract.

@@
 State-Machine Replication (SMR) algorithms such as Paxos are notoriously hard to
-understand, let alone to devise. Moreover, there is no practical way to reuse and optimize existing SMR
-algorithms, therefore the cost of creating a new SMR algorithm is high.
+understand, let alone to devise. Moreover, there is no practical way to reuse
+and optimize existing SMR algorithms, therefore the cost of creating a new SMR
+algorithm is high.
 
-To lower this cost, we present a framework for building SMR algorithms out of independent and reusable SMR modules. A module is an SMR
-algorithm that can abort and pass the baton to another module, allowing to dynamically switch between modules. 
-We propose Composable
-Generalized Consensus (CGC), a correctness condition for individual modules, and
-we prove that the composition of correct modules is a correct SMR algorithm.
-Our framework consists of TLA$^+$ specifications of CGC and of two abstract
-types of correct modules, fast rounds and conservative rounds, each providing a template to devise new SMR modules with the help of the TLA$^+$
-tools. We show that several well-known SMR algorithms can be modified to
-implement fast or conservative rounds, therefore enabling
-their composition to obtain a single versatile SMR algorithm.
+To lower this cost, we present a framework for building SMR algorithms out of
+independent and reusable SMR modules. A module is an SMR algorithm that can
+abort and pass the baton to another module, allowing to dynamically switch
+between modules. Our composition theorem ensures that the composition of any
+number of modules satisfying Composable Generalized Consensus (CGC), our
+correctness condition for modules, is a correct SMR algorithm.
+
+To help design new modules in practice, we propose TLA$^+$ specifications
+of CGC and of two abstract types of modules, fast rounds and conservative
+rounds, each providing a template to devise new SMR modules with the help of the
+TLA$^+$ tools. We show that several well-known SMR algorithms can be modified to
+implement fast or conservative rounds, therefore enabling their composition to
+obtain a single versatile SMR algorithm.
 

### 2016-01-07 paper/abstract4.tex
work on the abstract: shortening of first paragraph (problem statement).

@@
 
-State-Machine Replication (SMR) algorithms such as Paxos are notoriously hard to
-understand, let alone to devise. Moreover, there is no practical way to reuse
-and optimize existing SMR algorithms, therefore the cost of creating a new SMR
-algorithm is high.
+%State-Machine Replication (SMR) algorithms such as Paxos are notoriously hard to
+%understand, let alone to devise. Moreover, there is no practical way to reuse
+%and optimize existing SMR algorithms, therefore the cost of creating a new SMR
+%algorithm is high.
 
-To lower this cost, we present a framework for building SMR algorithms out of
-independent and reusable SMR modules. A module is an SMR algorithm that can
-abort and pass the baton to another module, allowing to dynamically switch
-between modules. Our composition theorem ensures that the composition of any
-number of modules satisfying Composable Generalized Consensus (CGC), our
-correctness condition for modules, is a correct SMR algorithm.
+State-Machine Replication (SMR) algorithms are notoriously hard to understand
+and devising new and more performant SMR algorithms is costly. To lower this
+cost, we present a framework for building SMR algorithms out of independent
+and reusable modules. A module is an SMR algorithm that can abort and pass
+the baton to another module, allowing to dynamically switch between modules.
+Our composition theorem ensures that the composition of any number of modules
+satisfying Composable Generalized Consensus (CGC), our correctness condition for
+modules, is a correct SMR algorithm.
 
@@
 of CGC and of two abstract types of modules, fast rounds and conservative
-rounds, each providing a template to devise new SMR modules with the help of the
+rounds, each providing a template to devise new modules with the help of the
 TLA$^+$ tools. We show that several well-known SMR algorithms can be modified to

### 2016-01-07 paper/abstract4.tex
Working on the abstract: grammar and emphasis on the problem.

@@
 
-State-Machine Replication (SMR) algorithms are notoriously hard to understand
-and devising new and more performant SMR algorithms is costly. To lower this
-cost, we present a framework for building SMR algorithms out of independent
-and reusable modules. A module is an SMR algorithm that can abort and pass
-the baton to another module, allowing to dynamically switch between modules.
-Our composition theorem ensures that the composition of any number of modules
-satisfying Composable Generalized Consensus (CGC), our correctness condition for
-modules, is a correct SMR algorithm.
+State-Machine Replication (SMR) algorithms are notoriously hard to understand, and one can trust their correctness only by a costly formal specification and analysis.
+Moreover, it is often impossible to reuse the results of formal analyses because
+existing SMR algorithms are not modular.
+The design of new SMR algorithms is therefore expensive.
 
-To help design new modules in practice, we propose TLA$^+$ specifications
-of CGC and of two abstract types of modules, fast rounds and conservative
-rounds, each providing a template to devise new modules with the help of the
-TLA$^+$ tools. We show that several well-known SMR algorithms can be modified to
-implement fast or conservative rounds, therefore enabling their composition to
-obtain a single versatile SMR algorithm.
+To lower the cost of designing new SMR algorithms, we present a framework for composing independent and reusable SMR modules. 
+An SMR module is an SMR algorithm augmented with a composition interface. 
+In the composition of two modules, the first module starts executing and can abort its execution and pass the baton to the second module at any time.
+Our composition theorem guarantees that the result of composing two SMR modules is again an SMR module and a correct SMR algorithm.
+Module composition can be dynamic; therefore, our framework allows building SMR algorithms that choose and switch at runtime between a set of modules, and that can be augmented with new modules at any time.
+
+To help design new modules in practice, we propose several TLA$^+$ specifications that each
+provide a template to devise new modules with the help of the TLA$^+$ tools.
+Using our specifications, we demonstrate that several well-known SMR algorithms can be made modular and composed to form a versatile SMR algorithm.
 

### 2016-01-07 paper/abstract4.tex
Work on the abstract: shortening.

@@
 
-To lower the cost of designing new SMR algorithms, we present a framework for composing independent and reusable SMR modules. 
-An SMR module is an SMR algorithm augmented with a composition interface. 
+To lower the cost of designing new SMR algorithms, we present a framework for composing SMR modules whose analysis can be done in isolation and reused. 
+An SMR module is an SMR algorithm augmented with a composition interface and satisfying a specification called Composable Generalized Consensus. 
 In the composition of two modules, the first module starts executing and can abort its execution and pass the baton to the second module at any time.
-Our composition theorem guarantees that the result of composing two SMR modules is again an SMR module and a correct SMR algorithm.
+Our composition theorem guarantees that the composition of two SMR modules is itself an SMR module. %; therefore, composition can be applied repeatedly to a sequence of modules.
 Module composition can be dynamic; therefore, our framework allows building SMR algorithms that choose and switch at runtime between a set of modules, and that can be augmented with new modules at any time.
 
-To help design new modules in practice, we propose several TLA$^+$ specifications that each
+To help design new modules in practice, we propose TLA$^+$ specifications that each
 provide a template to devise new modules with the help of the TLA$^+$ tools.

### 2016-01-07 paper/abstract4.tex
work on the abstract.

@@
 Our composition theorem guarantees that the composition of two SMR modules is itself an SMR module. %; therefore, composition can be applied repeatedly to a sequence of modules.
-Module composition can be dynamic; therefore, our framework allows building SMR algorithms that choose and switch at runtime between a set of modules, and that can be augmented with new modules at any time.
+Module composition can be dynamic, allowing to build SMR algorithms that choose and switch at runtime between a set of modules, and that can be augmented with new modules at any time.
 
-To help design new modules in practice, we propose TLA$^+$ specifications that each
-provide a template to devise new modules with the help of the TLA$^+$ tools.
-Using our specifications, we demonstrate that several well-known SMR algorithms can be made modular and composed to form a versatile SMR algorithm.
+To help module designers in practice, we propose TLA$^+$ specifications 
+providing templates to devise SMR modules using the TLA$^+$ tools.
+Using our specifications, we build modular versions of well-known SMR algorithms, enabling their composition and reuse in new SMR algorithms.
 

### 2016-01-08 paper/intro.tex
Work on the intro, mainly.

@@
 The State-Machine Replication technique (abbreviated SMR) allows building
-reliable computing services on top of unreliable hardware. The basic idea is
-to replicate a service over several servers and enforce, using a distributed
-protocol, that each server executes the same sequence of deterministic commands.
-In this fashion, a command will produce the same effect on all replicas.
-Therefore, if a replica crashes, clients of the service can be transparently redirected to
-another replica and will have the illusion of accessing a centralized service at
-all times. SMR is often formally specified using the Generalized Consensus
-specification~\cite{Lamport05GeneralizeConsensus}, abbreviated GC\@.
+reliable computing services on top of unreliable hardware. The basic idea
+is to replicate a service over several replica servers and enforce, using
+a distributed algorithm, that each replica executes the same sequence of
+deterministic commands. In this fashion, a command will produce the same effect
+on all replicas. Therefore, if a replica crashes, clients of the service can
+be transparently redirected to another replica and will have the illusion
+of accessing a centralized service at all times. Formally, we will use the
+Generalized Consensus specification~\cite{Lamport05GeneralizeConsensus},
+abbreviated GC, as our correctness condition for SMR algorithms.
 
-Because SMR is part of the basic infrastructure supporting cloud computing,
+SMR is part of the basic infrastructure supporting cloud computing; therefore,
 improving the performance of SMR algorithms can result in performance gains,
@@
 
-However, devising new SMR algorithms or optimizing existing ones for a
-particular use-case is hard: today, more than two decades after the first SMR
-algorithms~\cite{Lamport98ParttimeParliament,BirmanJoseph87ReliableCommunicationPresenceFailures,OkiLiskov88ViewstampedReplicationGeneralPrimaryCopy,DworkLynchStockmeyer84ConsensusPresencePartialSynchronyPreliminaryVersion}
+However, devising new SMR algorithms is very costly because one has to employ
+formal verification techniques to ensure their correctness. SMR algorithms are
+notoriously difficult to understand and it is easy to overlook catastrophic bugs
+appearing not only in implementations but also in the high level algorithm.
+Moreover, concurrency, network behavior, and faults give rise to a number
+of possible interleavings of actions that is beyond the reach of analysis
+by testing method. Therefore, to gain confidence in the correctness of an
+SMR algorithm, one needs to resort to formal methods like model-checking or
+interactive theorem proving. However, this is costly: model-checking often
+requires manually building an abstraction of the algorithm to simplify the task
+of the model checker, and interactive theorem proving is very time consuming.
+Both require experts trained in the particular tool used for the task. As
+evidence of the difficulty of designing new SMR algorithms, note that today,
+more than two decades after the first SMR algorithms~\cite{Lamport98ParttimeParliament,BirmanJoseph87ReliableCommunicationPresenceFailures,OkiLiskov88ViewstampedReplicationGeneralPrimaryCopy,DworkLynchStockmeyer84ConsensusPresencePartialSynchronyPreliminaryVersion}
 were proposed, devising a new SMR algorithm can justify a publication in a top
@@
 
-The problem is that it is difficult to reuse existing algorithms and ideas
-because SMR algorithms are not modular. When devising a new algorithm, any
-analysis, test, or correctness proof has to be done again from scratch.
+Surprisingly, at a first glance, many SMR algorithms seem to share a common structure which would indicate that previous formal analysis could be reused. 
+That it because, in many cases, researchers obtained a new SMR algorithm by optimizing an existing SMR algorithm for a particular use case.
+For example, Fast Paxos~\cite{Lamport06FastPaxos} adds fast rounds to MultiPaxos~\cite{lamport2001paxos} to reduce the latency under low contention, 
+Disk Paxos~\cite{GafniLamport03DiskPaxos} optimizes Paxos for systems where disk nodes and processor nodes are distinct, 
+Chain Replication~\cite{RenesseSchneider04ChainReplicationSupportingHighThroughputAvailability} organizes the replicas of Paxos in a ring to distribute commands at a higher throughput but at the expense of latency,
+FGGC~\cite{SutraShapiro11FastGenuineGeneralizedConsensus} optimizes Generalized Paxos\cite{Lamport05GeneralizeConsensus} to reduce latency when the system contains a reliable and fast leader node, 
+Ring Paxos~\cite{MarandiETAL10RingPaxosHighthroughputAtomicBroadcastProtocol} optimizes the latency and throughput of Paxos when network multicast is available, 
+Multicoordinated Paxos reduces the performance impact of a leader fault in MultiPaxos by implementing a distributed leader, 
+Cheap Paxos~\cite{LamportMassa04CheapPaxos} optimizes Paxos to decrease the load on some replicas in case failures are sufficiently separated in time,
+Mencius~\cite{MaoJunqueiraMarzullo08MenciusBuildingEfficientReplicatedStateMachine} optimizes MultiPaxos for wide-area networks by rotating the role of leader among all replicas, etc.
+Close scrutiny reveals that most SMR algorithms can be seen as variations and optimizations around a few core ideas such as the round structure introduced by Paxos~\cite{Lamport98ParttimeParliament}, quorum systems~\cite{GuerraouiVukolic10RefinedQuorumSystems,Lamport06FastPaxos}, and agreement up to commutativity using command-structure sets~\cite{Lamport05GeneralizeConsensus}.
+
+Given their similarities, one may ask why is it not possible to reuse formal analysis of SMR algorithms.
+The problem is that SMR algorithms are not modular and that, because of their intricate structure, modifying a monolithic SMR algorithm may produce many unforeseen corner cases that invalidate previous analyses.
 
 To simplify the task of devising new SMR algorithms we propose a framework for
-devising reusable SMR modules. A module in our framework is an SMR algorithm
-augmented with an \emph{abort interface} for aborting its execution and passing
+building reusable SMR modules.  
+For example, in \cref{}, we augment MultiPaxos with our composition interface and prove that it satisfies Composable Generalized Consensus (CGC), our correctness condition for modules.
+We then build an SMR module whose structure is similar to the one of the fast rounds of Fast Paxos and prove that it satisfies CGC\@.
+Our Fast Paxos module has lower latency than our Paxos module when there is low contention among clients, but its performance drops below that of Paxos when contention increases. 
+Finally, we assemble unmodified our MultiPaxos SMR module and our Fast Paxos SMR module in an SMR algorithm that dynamically switch from one to the other depending on the contention level observed in the system, therefore getting the best of both.
+Crucially, we obtain the correctness of our composite SMR algorithm for free: our generic composition theorem and the correctness of our two SMR modules taken individually imply that the combination of the MultiPaxos SMR module and the Fast Paxos SMR module is a correct SMR algorithm.
+Later, in \cref{}, we present three other SMR modules that can be combined the MultiPaxos SMR module and the Fast Paxos SMR module in a composite SMR algorithm that can switch between 5 different kinds of modules, according to the system's operating conditions.
+
+A module in our framework is an SMR algorithm
+augmented with \emph{composition interface} and satisfying our correctness condition, called Composable Generalized Consensus (CGC).
+The composition interface consists of an \emph{abort interface} for aborting a module's execution and passing
 the baton to another module and with a matching \emph{init interface} for taking
@@
 in the context of the second module.
-
 A module must satisfy a correctness condition that we call Composable
 Generalized Consensus (abbreviated CGC). Any correct module is a safe
-Generalized Consensus~\cite{Lamport05GeneralizeConsensus} algorithm, but the progress guarantees may vary between
+Generalized Consensus~\cite{Lamport05GeneralizeConsensus} algorithm (i.e.\ a kind of SMR algorithm), but the progress guarantees may vary between
 modules: CGC only requires a module to make progress executing commands or to
@@
 CGC also relates the init values, learned values (learned values represent the local state-machine state of individual replicas), and abort values in order for our composition theorem to hold.
-Generalized Consensus is a specification of SMR algorithms that takes into account the commutativity relation among commands and can off important performance advantages in practice.
+%Note that Generalized Consensus is a specification of SMR algorithms that takes into account the commutativity relation among commands and can offer important performance advantages in practice.
 
-The composition theorem enables incremental development of SMR algorithms
+Our composition theorem enables incremental development of SMR algorithms
 because it guarantees that any number of correct modules may be composed in
@@
 produce an abort value after gathering the state of a read quorum of replicas.
+Each of the specifications achieves a particular consistency/latency trade-off~\cite{Abadi12ConsistencyTradeoffsModernDistributedDatabaseSystem}.
 The two specifications use different combinations of sizes of read and write quorums, which affects the invariants that the replicas must maintain and, in consequence, the latency with which a command can be processed but the module.

### 2016-01-09 paper/intro.tex
Small changes in the intro

@@
 
-Surprisingly, at a first glance, many SMR algorithms seem to share a common structure which would indicate that previous formal analysis could be reused. 
-That it because, in many cases, researchers obtained a new SMR algorithm by optimizing an existing SMR algorithm for a particular use case.
-For example, Fast Paxos~\cite{Lamport06FastPaxos} adds fast rounds to MultiPaxos~\cite{lamport2001paxos} to reduce the latency under low contention, 
-Disk Paxos~\cite{GafniLamport03DiskPaxos} optimizes Paxos for systems where disk nodes and processor nodes are distinct, 
-Chain Replication~\cite{RenesseSchneider04ChainReplicationSupportingHighThroughputAvailability} organizes the replicas of Paxos in a ring to distribute commands at a higher throughput but at the expense of latency,
-FGGC~\cite{SutraShapiro11FastGenuineGeneralizedConsensus} optimizes Generalized Paxos\cite{Lamport05GeneralizeConsensus} to reduce latency when the system contains a reliable and fast leader node, 
-Ring Paxos~\cite{MarandiETAL10RingPaxosHighthroughputAtomicBroadcastProtocol} optimizes the latency and throughput of Paxos when network multicast is available, 
-Multicoordinated Paxos reduces the performance impact of a leader fault in MultiPaxos by implementing a distributed leader, 
-Cheap Paxos~\cite{LamportMassa04CheapPaxos} optimizes Paxos to decrease the load on some replicas in case failures are sufficiently separated in time,
-Mencius~\cite{MaoJunqueiraMarzullo08MenciusBuildingEfficientReplicatedStateMachine} optimizes MultiPaxos for wide-area networks by rotating the role of leader among all replicas, etc.
-Close scrutiny reveals that most SMR algorithms can be seen as variations and optimizations around a few core ideas such as the round structure introduced by Paxos~\cite{Lamport98ParttimeParliament}, quorum systems~\cite{GuerraouiVukolic10RefinedQuorumSystems,Lamport06FastPaxos}, and agreement up to commutativity using command-structure sets~\cite{Lamport05GeneralizeConsensus}.
+Surprisingly, at a first glance, many SMR algorithms seem to share a
+common structure which would indicate that previous formal analysis could
+be reused. That is because, in many cases, researchers obtained a new SMR
+algorithm by optimizing an existing SMR algorithm for a particular use
+case. For example, Fast Paxos~\cite{Lamport06FastPaxos} adds fast rounds
+to MultiPaxos~\cite{lamport2001paxos} to reduce the latency under low
+contention, Disk Paxos~\cite{GafniLamport03DiskPaxos} optimizes Paxos
+for systems where disk nodes and processor nodes are distinct, Chain
+Replication~\cite{RenesseSchneider04ChainReplicationSupportingHighThroughputAvailability} 
+organizes the replicas of Paxos in a ring to distribute
+commands at a higher throughput but at the expense of latency,
+FGGC~\cite{SutraShapiro11FastGenuineGeneralizedConsensus} optimizes
+Generalized Paxos\cite{Lamport05GeneralizeConsensus} to reduce latency
+when the system contains a reliable and fast leader node, Ring
+Paxos~\cite{MarandiETAL10RingPaxosHighthroughputAtomicBroadcastProtocol}
+optimizes the latency and throughput of Paxos when network multicast is
+available, Multicoordinated Paxos reduces the performance impact of a
+leader fault in MultiPaxos by implementing a distributed leader, Cheap
+Paxos~\cite{LamportMassa04CheapPaxos} optimizes Paxos to decrease the load
+on some replicas in case failures are sufficiently separated in time,
+Mencius~\cite{MaoJunqueiraMarzullo08MenciusBuildingEfficientReplicatedStateMachine} 
+optimizes MultiPaxos for wide-area networks by rotating the role of leader
+among all replicas, etc. Close scrutiny reveals that most SMR algorithms can
+be seen as variations and optimizations around a few core ideas such as the
+round structure introduced by Paxos~\cite{Lamport98ParttimeParliament}, quorum
+systems~\cite{GuerraouiVukolic10RefinedQuorumSystems,Lamport06FastPaxos},
+and agreement up to commutativity using command-structure
+sets~\cite{Lamport05GeneralizeConsensus}.
 
-Given their similarities, one may ask why is it not possible to reuse formal analysis of SMR algorithms.
-The problem is that SMR algorithms are not modular and that, because of their intricate structure, modifying a monolithic SMR algorithm may produce many unforeseen corner cases that invalidate previous analyses.
+Given their similarities, one may ask why is it not possible to reuse formal
+analysis of SMR algorithms. The problem is that SMR algorithms are not modular
+and that, because of their intricate structure, modifying a monolithic SMR
+algorithm may produce many unforeseen corner cases that invalidate previous
+analyses.
 
-To simplify the task of devising new SMR algorithms we propose a framework for
-building reusable SMR modules.  
-For example, in \cref{}, we augment MultiPaxos with our composition interface and prove that it satisfies Composable Generalized Consensus (CGC), our correctness condition for modules.
-We then build an SMR module whose structure is similar to the one of the fast rounds of Fast Paxos and prove that it satisfies CGC\@.
-Our Fast Paxos module has lower latency than our Paxos module when there is low contention among clients, but its performance drops below that of Paxos when contention increases. 
-Finally, we assemble unmodified our MultiPaxos SMR module and our Fast Paxos SMR module in an SMR algorithm that dynamically switch from one to the other depending on the contention level observed in the system, therefore getting the best of both.
-Crucially, we obtain the correctness of our composite SMR algorithm for free: our generic composition theorem and the correctness of our two SMR modules taken individually imply that the combination of the MultiPaxos SMR module and the Fast Paxos SMR module is a correct SMR algorithm.
-Later, in \cref{}, we present three other SMR modules that can be combined the MultiPaxos SMR module and the Fast Paxos SMR module in a composite SMR algorithm that can switch between 5 different kinds of modules, according to the system's operating conditions.
+To simplify the task of devising new SMR algorithms we propose a framework
+for building reusable SMR modules. For example, in \cref{}, we augment
+MultiPaxos with our composition interface and prove that it satisfies Composable
+Generalized Consensus (CGC), our correctness condition for modules. We then
+build an SMR module whose structure is similar to the one of the fast rounds of
+Fast Paxos and prove that it satisfies CGC\@. Our Fast Paxos module has lower
+latency than our Paxos module when there is low contention among clients, but
+its performance drops below that of Paxos when contention increases. Finally,
+we assemble unmodified our MultiPaxos SMR module and our Fast Paxos SMR module
+in an SMR algorithm that dynamically switch from one to the other depending on
+the contention level observed in the system, therefore getting the best of both.
+Crucially, we obtain the correctness of our composite SMR algorithm for free:
+our generic composition theorem and the correctness of our two SMR modules taken
+individually imply that the combination of the MultiPaxos SMR module and the
+Fast Paxos SMR module is a correct SMR algorithm. Later, in \cref{}, we present
+three other SMR modules that can be combined the MultiPaxos SMR module and the
+Fast Paxos SMR module in a composite SMR algorithm that can switch between 5
+different kinds of SMR modules, according to the system's operating conditions.
 
-A module in our framework is an SMR algorithm
-augmented with \emph{composition interface} and satisfying our correctness condition, called Composable Generalized Consensus (CGC).
-The composition interface consists of an \emph{abort interface} for aborting a module's execution and passing
-the baton to another module and with a matching \emph{init interface} for taking
-over the execution of an aborting module. Two modules are composed by passing
-the outputs produced by the abort interface of the first module, called abort
-values, as inputs to the init interface of the second module, called init values
-in the context of the second module.
-A module must satisfy a correctness condition that we call Composable
-Generalized Consensus (abbreviated CGC). Any correct module is a safe
-Generalized Consensus~\cite{Lamport05GeneralizeConsensus} algorithm (i.e.\ a kind of SMR algorithm), but the progress guarantees may vary between
-modules: CGC only requires a module to make progress executing commands or to
-abort its execution.
-CGC also relates the init values, learned values (learned values represent the local state-machine state of individual replicas), and abort values in order for our composition theorem to hold.
+An SMR module is an SMR algorithm augmented with \emph{composition interface}
+and satisfying our correctness condition, called Composable Generalized
+Consensus (CGC). The composition interface consists of an \emph{abort interface}
+for aborting an SMR module's execution and passing the baton to another
+SMR module and with a matching \emph{init interface} for taking over the
+execution of an aborting SMR module. Two SMR modules are composed by passing
+the outputs produced by the abort interface of the first SMR module, called
+abort values, as inputs to the init interface of the second SMR module,
+called init values in the context of the second SMR module. A SMR module
+must satisfy a correctness condition that we call Composable Generalized
+Consensus (abbreviated CGC). Any correct SMR module is a safe Generalized
+Consensus~\cite{Lamport05GeneralizeConsensus} algorithm (i.e.\ a kind of SMR
+algorithm), but the progress guarantees may vary between SMR modules: CGC only
+requires a SMR module to make progress executing commands or to abort its
+execution. CGC also relates the init values, learned values (learned values
+represent the local state-machine state of individual replicas), and abort
+values in order for our composition theorem to hold.
 %Note that Generalized Consensus is a specification of SMR algorithms that takes into account the commutativity relation among commands and can offer important performance advantages in practice.
@@
 \begin{itemize}
+    \item We present the first framework to allowing to build Generalized Consensus algorithms our of independent modules.
     \item We provide tool support for designing CGC modules in the form of two TLA+~\cite{Lamport02SpecifyingSystems} specifications simplifying the design and model-checking of SMR algorithms based on Quorum Systems~\cite{GuerraouiVukolic10RefinedQuorumSystems} and demonstrate their use by designing CGC modules inspired by Paxos, Fast Paxos, and Primary-Backup replication.
-    \item Our framework is based on Generalized Consensus, a generalization of the consensus problem 
-        that allows optimizing the execution of commuting commands, a common practical case.
 \end{itemize}

### 2016-01-10 paper/cgc.tex
Working on presenting CGC.

\section{Composable Generalized Consensus}

We consider a computing service specified as a deterministic state machine $M$ whose transitions correspond to the commands that clients of the service may issue.
If $M$ is implemented on a single server then any fault happening on this server may become visible to the clients of the service.
To mask faults, one can replicate the implementation of $M$ on several replica servers (called replicas for short) and direct clients' commands to replicas which are known to be non-faulty.
However, the intent of replicatio is to mask faults while preserving the behavior of the service from the point of view of its clients.

State-Machine Replication~\cite{Schneider90ImplementingFaulttolerantServicesUsingStateMachine} aims at orchestrating the execution of the replicas to ensure that clients have the illusion of accessing a unique and centralized implementation of $M$ despite a number of faults, which must remain below a certain maximum depending on the algorithms used.
State-Machine Replication consists in replicating an implementation of $M$ over the replicas and ensuring with a distributed algorithm $A$ that all the replica execute the same sequence of commands, possibly up to the reordering of commuting commands.

Generalized Consensus~\cite{Lamport05GeneralizeConsensus} is a formal specification of the requirements that the distributed algorithm $A$ must meet in order for SMR to provide the illusion of a centralized fault-tolerant service.
Generalized Consensus uses the notion of a set of command-structures, or c-structs. C-structs are obtained by appending commands to the c-struct $\bot$. Given a c-struct $s$ and a command $c$, $s \bullet c$ is the c-struct obtained by appending $c$ to $s$. Given a sequence of commands $cs$, $s\star cs$ is the c-struct obtained by repeatedly appending the commands in $cs$ in the order where they appear in $cs$. By appropriately defining the operator $\bullet$ according to the
definition of $M$, C-structs can describe sequences of commands up the reordering of commuting
commands: it is possible to have $s \star \left[c_1,c_2\right] = s \star \left[c_2,c_1\right]$.
Moreover, c-structs must form a partial order under the operator $\preceq$ defined as $s_1 \preceq s_2$ if and only if there exists a sequence of comamnds $cs$ such that $s_2 = s_1 \star cs$; the greatest lower bound of a set of c-structs $S$ must always exists and be constructible from the the same commands as the c-structs in $S$; and a set of c-structs $S$ which has an upper bound, called a compatible set, must have a lowest upper bound which is constructible from the the same commands as the c-structs in $S$.

Generalized Consensus is formulated in terms of two types of processes: the set $P$ of proposers and the set $L$ of learners, but it is possible for each replica to play the two roles at the same time. Proposers propose commands and learners, which represent replicas, learn about c-structs.
We describe the execution of learners and proposers using the variable $prop$, which is a set containing all the commands that have been proposed, and the variable $learned$, which maps a learner to the latest c-structs it has learned.

Generalized Consensus requires that for any learner $l$, there is at all times a sequence of proposed commands $cs$ such that $learned\left[l\right] = \bot \bullet cs$ (Nontriviality); that for any learner $l$, $learned\left[ l \right]$ increase with time (Stability); that $\left\{ learned\left[ l \right] :  l \in L\right\}$ is always compatible (Agreement); and that for any proposed command $c$ and learner $l$, a c-struct of the form $s\bullet c$ is eventually learned by $l$ (Liveness).
We refer the reader to the work of Lamport~\cite{Lamport05GeneralizeConsensus} for a thourough discussion of Generalized Consensus.

Composable Generalized Consensus augments the Generalized Consensus specification with an abort interface that allows an SMR module to abort its execution, and with an init interface which allows an SMR module to start its execution where a preceeding aborted module left off.
To specify the composition interface, we introduced a third type of processes called the \emph{switchers}. Switchers can receive \emph{init values} from a preceeding aborting modules and can pass \emph{abort values} to pass the baton to a next SMR module when the current SMR module aborts.

We describe the behavior of switchers using the variable $from$, which contains all the init values received by any switcher, and the variable $to$, which contains all the abort values produced by the switchers.
The variable $prop$ is used as before, but the variable $learned$ now maps learneres to sets of c-structs and, for each learner $l$, $learned\left[ l \right]$ consists of all the c-structs learned so far by $l$.
We define the set of valid c-structs as the c-structs of the form $t\star cs$ where $cs$ is a sequence of proposed commands, and $t$ is the greatest lower bound of a non-empty subset of the init values received.

Composable Generalized Consensus requires that no value be learned unless at least one init value have been received (Initialization); that for any learner $l$, $learned\left[ l \right]$ is valid (Nontriviality); that $\left\{ learned\left[ l \right] :  l \in L\right\}$ is always compatible (Agreement); that every abort value is valid, that for any abort value $s_a$ and learned value $s_l$ we have $s_l \preceq s_a$ (Safe Abort); and that for any proposed command $c$ and learner $l$, a c-struct
of the form $s\bullet c$ is eventually learned by $l$ or a switcher eventually produces an abort value (Liveness).
A TLA$^+$ specification of CGC appears in \cref{fig:cgc}.

\begin{figure}
\begin{minipage}[t]{.49\textwidth}
    \centering
    \includegraphics[width=\textwidth,trim={129pt 397pt 225pt 123pt},clip]{../TLA/ComposableGC.pdf}
    \caption{TLA$⁺$ specification of CGC}\label{fig:cgc}%
\end{minipage}\hfill
\end{figure}

### 2016-01-10 paper/intro.tex
Working on presenting CGC.

@@
 
-An SMR module is an SMR algorithm augmented with \emph{composition interface}
+An SMR module is an SMR algorithm augmented with a \emph{composition interface}
 and satisfying our correctness condition, called Composable Generalized

### 2016-01-10 paper/related.tex
Work on the intro, mainly.
Working on presenting CGC.

@@
 
-Generalized Consensus is a specification of SMR algorithms that takes into account the commutativity relation among commands and allows its implementation to optimize the execution of commuting commands.
-In cases in which many commands commute, which may be frequent in practice, Generalized Consensus algorithms can have a significant performance advantage over traditional consensus algorithms.
-
-Making SMR algorithms efficient has been a long-standing
-challenge and a multitude of algorithms have been proposed, each best suited to a particular use case.
-Examples include Paxos~\cite{lamport2001paxos}, Fast
-Paxos~\cite{Lamport06FastPaxos}, Disk Paxos~\cite{GafniLamport03DiskPaxos},
-Generalized Paxos~\cite{Lamport05GeneralizeConsensus},
-FGGC~\cite{SutraShapiro11FastGenuineGeneralizedConsensus}, Chain
-Replication~\cite{RenesseSchneider04ChainReplicationSupportingHighThroughputAvailability}, Ring
-Paxos~\cite{MarandiETAL10RingPaxosHighthroughputAtomicBroadcastProtocol},
-Egalitarian Paxos
-\cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments},
-Multi-Coordinated Paxos~\cite{CamargosSchmidtPedone07MulticoordinatedPaxos},
-Cheap Paxos~\cite{LamportMassa04CheapPaxos}, Vertical Paxos
-\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}, Paxos-MIC
-\cite{HurfinMoiseNarzul11AdaptiveFastPaxosMakingQuickEverlasting}, Mencius
-\cite{MaoJunqueiraMarzullo08MenciusBuildingEfficientReplicatedStateMachine},
-Fast Mencius~\cite{WeiETAL13FastMenciusMenciusLowCommitLatency},
-the Chandra-Toueg algorithm
-\cite{ChandraToueg96UnreliableFailureDetectorsReliableDistributedSystems}, the
-Ben-Or algorithm~\cite{BenOr83AnotherAdvantageFreeChoiceCompletelyAsynchronous},
-and Raft~\cite{OngaroOusterhout14SearchUnderstandableConsensusAlgorithm}.
-
+% Abstract framework
 Our work is closely related to the Abstract
@@
 
+% Speclin
 The Speculative Linearizability framework was a
@@
 
-TODO\@: expand the following.
-
+% Generalized Paxos
 Generalized Consensus is an agreement problem relaxing the ordering constraints
@@
 
+% Vertical Paxos
 Vertical Paxos algorithms are SMR algorithms whose set of replica servers can

### 2016-01-11 paper/cgc.tex
Work on the cgc section, mostly.

@@
 To mask faults, one can replicate the implementation of $M$ on several replica servers (called replicas for short) and direct clients' commands to replicas which are known to be non-faulty.
-However, the intent of replicatio is to mask faults while preserving the behavior of the service from the point of view of its clients.
+However, the intent of replication is to mask faults while preserving the behavior of the service from the point of view of its clients.
 
@@
 State-Machine Replication consists in replicating an implementation of $M$ over the replicas and ensuring with a distributed algorithm $A$ that all the replica execute the same sequence of commands, possibly up to the reordering of commuting commands.
+
+\subsection{Generalized Consensus}
 
@@
 commands: it is possible to have $s \star \left[c_1,c_2\right] = s \star \left[c_2,c_1\right]$.
-Moreover, c-structs must form a partial order under the operator $\preceq$ defined as $s_1 \preceq s_2$ if and only if there exists a sequence of comamnds $cs$ such that $s_2 = s_1 \star cs$; the greatest lower bound of a set of c-structs $S$ must always exists and be constructible from the the same commands as the c-structs in $S$; and a set of c-structs $S$ which has an upper bound, called a compatible set, must have a lowest upper bound which is constructible from the the same commands as the c-structs in $S$.
+Moreover, c-structs must satisfy a number of properties: \setword{property 1}{prop:1}, c-structs must form a partial order under the operator $\preceq$ defined as: $s_1 \preceq s_2$ if and only if there exists a sequence of commands $cs$ such that $s_2 = s_1 \star cs$; \setword{property 2}{prop:2}, the greatest lower bound of a set of c-structs $S$ must always exists and be constructible from the same commands as the c-structs in $S$; and \setword{property 3}{prop:3}, a set of c-structs $S$ which has an upper bound, called a compatible set, must have a lowest upper bound which is constructible from the same commands as the c-structs in $S$.
 
@@
 
-Generalized Consensus requires that for any learner $l$, there is at all times a sequence of proposed commands $cs$ such that $learned\left[l\right] = \bot \bullet cs$ (Nontriviality); that for any learner $l$, $learned\left[ l \right]$ increase with time (Stability); that $\left\{ learned\left[ l \right] :  l \in L\right\}$ is always compatible (Agreement); and that for any proposed command $c$ and learner $l$, a c-struct of the form $s\bullet c$ is eventually learned by $l$ (Liveness).
-We refer the reader to the work of Lamport~\cite{Lamport05GeneralizeConsensus} for a thourough discussion of Generalized Consensus.
+Generalized Consensus requires the following properties: non-triviality, stating that for any learner $l$, there is at all times a sequence of proposed commands $cs$ such that $learned\left[l\right] = \bot \bullet cs$; stability, stating that for any learner $l$, $learned\left[ l \right]$ increase with time; agreement, stating that $\left\{ learned\left[ l \right] :  l \in L\right\}$ is always compatible; and liveness, stating that for any proposed command $c$ and learner $l$, a c-struct of the form $s\bullet c$ is eventually learned by $l$.
+Note that the FLP theorem~\cite{FischerLynchPaterson83ImpossibilityDistributedConsensusOneFaultyProcess} implies that the liveness property cannot be achieved in a networked system where processes may fail.
+Therefore, the liveness property guaranteed by Generalized Consensus algorithms is necessary weaker that the liveness property state above, maybe adding assumptions about faults and synchrony.
+We refer the reader to the work of Lamport~\cite{Lamport05GeneralizeConsensus} for a thorough discussion of Generalized Consensus.
 
-Composable Generalized Consensus augments the Generalized Consensus specification with an abort interface that allows an SMR module to abort its execution, and with an init interface which allows an SMR module to start its execution where a preceeding aborted module left off.
+\subsection{Composable Generalized Consensus}
+
+Composable Generalized Consensus augments the Generalized Consensus specification with an abort interface that allows an SMR module to abort its execution, and with an init interface which allows an SMR module to start its execution where a preceding aborted module left off.
 To specify the composition interface, we introduced a third type of processes called the \emph{switchers}.
-Switchers can receive \emph{init values} from a preceeding aborting modules and can pass \emph{abort values} to pass the baton to a next SMR module when the current SMR module aborts.
+Switchers can receive \emph{init values} from a preceding aborting modules and can pass \emph{abort values} to pass the baton to a next SMR module when the current SMR module aborts.
 
 We describe the behavior of switchers using the variable $from$, which contains all the init values received by any switcher, and the variable $to$, which contains all the abort values produced by the switchers.
-The variable $prop$ is used as before, but the variable $learned$ now maps learneres to sets of c-structs and, for each learner $l$, $learned\left[ l \right]$ consists of all the c-structs learned so far by $l$.
+The variable $prop$ is used as before, but the variable $learned$ now maps learners to sets of c-structs and, for each learner $l$, $learned\left[ l \right]$ consists of all the c-structs learned so far by $l$.
 We define the set of valid c-structs as the c-structs of the form $t\star cs$ where $cs$ is a sequence of proposed commands, and $t$ is the greatest lower bound of a non-empty subset of the init values received.
@@
 
+We say that an algorithm implementing CGC under some refinement mapping is a \emph{CGC instance}, and that two CGC instances $i_1$ and $i_2$ are compatible if they update disjoint variables except for the $to$ variable of $i_1$ and the $from$ variable of $i_2$, which must be identical.
+Note that in the execution of two compatible instances, any abort value produced by the first instance is an init value for the second instance.
+We say that a CGC instance is an initial instance when its set of init values is always a subset of $\left\{ \bot \right\}$.
+
+CGC enjoys two main properties: first, the composition of two CGC instances is itself a CGC instance when the first instance is an initial instance; second, an initial CGC instance satisfies the safety properties of Generalized Consensus under the restrictions explained below.
+We call the first property the \emph{composition theorem}.
+A formal statement of the composition theorem appears in \cref{fig:compthm}, and an informal proof appears in \cref{sec:compproof}. 
+Note that TLA$^+$ is a second order logic and can therefore not express the composition theorem as stated above because we cannot quantify of modules.
+Instead, we state the composition theorem for two CGC instances obtained by only renaming the variables of the CGC specification.
+Any CGC instance would trivially refine both of the two instances used in formalization of the composition theorem because a CGC instance refines the CGC specification.
+Therefore, we obtain the composition theorem by transitivity and monotonicity of refinement~\cite{AbadiLamport91ExistenceRefinementMappings}.
+
 \begin{figure}
@@
     \centering
-    \includegraphics[width=\textwidth,trim={129pt 397pt 225pt 123pt},clip]{../TLA/ComposableGC.pdf}
-    \caption{TLA$⁺$ specification of CGC}\label{fig:cgc}%
+    \includegraphics[width=\textwidth,trim={129pt 397pt 225pt 128pt},clip]{../TLA/ComposableGC.pdf}
+    \caption{TLA$⁺$ specification of CGC.}\label{fig:cgc}%
+\end{minipage}\hfill
+\begin{minipage}[t]{.49\textwidth}
+    \centering
+    \includegraphics[width=\textwidth,trim={129pt 373pt 220pt 128pt},clip]{../TLA/CompositionTheorem.pdf}
+    \caption{The composition theorem in TLA$^⁺$.}\label{fig:compthm}%
+\end{minipage}
+\end{figure}
+
+\begin{figure}
+\begin{minipage}[t]{.49\textwidth}
+    \centering
+    \includegraphics[width=\textwidth,trim={129pt 455 220pt 128pt},clip]{../TLA/InitialIsGC.pdf}
+    \caption{Initial instances implement Generalized Consensus.}\label{fig:cgc}%
 \end{minipage}\hfill
 \end{figure}
+
+\subsection{Proof of the Composition Theorem}%
+\label{sec:compproof}
+To prove the composition theorem, as expressed in \cref{fig:compthm}, we must show that $Init$ implies $CGC!Init$ and that every transition of $Spec \equiv Init \wedge \box\left[Next\right]_{vars}$ is a transition of $CGC!Spec$.
+Below we sketch a proof of the second condition.
+
+First note that the following invariants hold of the execution of $Spec$: invariant 1, stating that any learned c-struct of $CGC1$ is smaller than any abort c-struct of $CGC1$; invariant 2, stating that for any non-empty subset $S$ of the abort c-structs of $CGC1$, there is a sequence $cs$ of commands proposed in $CGC1$ such that $\operatorname{GLB}\left( S \right) = \bot\star cs$; and invariant 3, stating that for every learned c-struct or abort c-struct $s$ of $CGC2$, there is a non-empty subset $S$ of the init c-structs of $CGC2$ and a sequence of commands proposed in $CGC2$ such that $s  = \operatorname{GLB}\left( S \right)\star cs$.
+Invariant 1 is enforced by the first conjunct of the $Learn(l)$ transition of $CGC1$, and invariant 3 is enforced by the quantifier bound of the $Learn(l)$ and $Abort$ transition of $CGC2$, using the definition of $Valid$.
+To see why invariant 2 holds, note that the quantifier bound of the $Abort$ transition of $CGC1$, the fact that the only init c-struct of $CGC1$ is $\left\{ \bot \right\}$, and the definition of $Valid$ ensure that any abort c-struct of $CGC1$ is of the form $\bot\star cs$ where $cs$ is a sequence of proposed commands of $CGC1$.
+Then, we obtain invariant 2 using the~\ref{prop:2} of c-structs.
+
+We now prove that each transition of $Spec$ is a transition of $CGC!Spec$:
+The transition $CGC1!Propose$ inserts a command $c$ in the set $CCG1!prop$, and, since $CGC!prop = CGC1!prop \cup CGC2!prop$, the transition inserts $c$ in the set $CGC!prop$; therefore, $CGC1!Propose$ implies $CGC!Propose$.
+By an analogous reasoning, $CGC2!Propose$ implies $CGC!Propose$.
+The transition $CGC1!InitM$ inserts $\bot$ in the set $CGC1!from$ (because $CGC1$ is an initial instance), and, since $CGC!prop = CGC1!prop$, the transition also inserts $\bot$ in $CGC!prop$; therefore, $CGC1!InitM$ implies $CGC!InitM$.
+The $CGC1!Abort$ transition does not modify any variable of the module instance $CGC$; therefore, $CGC1!Abort$ is a stuttering transition of $CGC$.
+Similarly, the $CGC2!InitM$ transition does not modify any variable of the module instance $CGC$; therefore, $CGC2!InitM$ is a stuttering transition of $CGC$.
+TODO\@: complete.
+We have formalized informal proof above in Isabelle/HOL (and not TLAPS~\cite{CousineauETAL12TlaProofs}, because the authors are proficient in Isabelle/HOL but not in TLAPS); the proof is available at \url{http://losa.fr/cgc}.
+
+\section{Deploying SMR modules in practice}
+
+\subsection{Agreeing on the Next Module}

### 2016-01-11 paper/cgc.tex
working on the writeup of the cgc proof.

@@
 commands: it is possible to have $s \star \left[c_1,c_2\right] = s \star \left[c_2,c_1\right]$.
-Moreover, c-structs must satisfy a number of properties: \setword{property 1}{prop:1}, c-structs must form a partial order under the operator $\preceq$ defined as: $s_1 \preceq s_2$ if and only if there exists a sequence of commands $cs$ such that $s_2 = s_1 \star cs$; \setword{property 2}{prop:2}, the greatest lower bound of a set of c-structs $S$ must always exists and be constructible from the same commands as the c-structs in $S$; and \setword{property 3}{prop:3}, a set of c-structs $S$ which has an upper bound, called a compatible set, must have a lowest upper bound which is constructible from the same commands as the c-structs in $S$.
+Moreover, c-structs must satisfy a number of properties: \setword{property 1}{prop:1}, c-structs must form a partial order under the operator $\preceq$ defined as: $s_1 \preceq s_2$ if and only if there exists a sequence of commands $cs$ such that $s_2 = s_1 \star cs$; \setword{property 2}{prop:2}, the greatest lower bound of a set of c-structs $S$ must always exists and be constructible from the same commands as the c-structs in $S$; and \setword{property 3}{prop:3}, a set of c-structs $S$
+which has an upper bound, called a \emph{compatible} set, must have a lowest upper bound which is constructible from the same commands as the c-structs in $S$.
 
@@
 
-Generalized Consensus requires the following properties: non-triviality, stating that for any learner $l$, there is at all times a sequence of proposed commands $cs$ such that $learned\left[l\right] = \bot \bullet cs$; stability, stating that for any learner $l$, $learned\left[ l \right]$ increase with time; agreement, stating that $\left\{ learned\left[ l \right] :  l \in L\right\}$ is always compatible; and liveness, stating that for any proposed command $c$ and learner $l$, a c-struct of the form $s\bullet c$ is eventually learned by $l$.
+Generalized Consensus requires the following properties: \emph{non-triviality}, stating that for any learner $l$, there is at all times a sequence of proposed commands $cs$ such that $learned\left[l\right] = \bot \bullet cs$; \emph{stability}, stating that for any learner $l$, $learned\left[ l \right]$ increase with time; \emph{agreement}, stating that $\left\{ learned\left[ l \right] :  l \in L\right\}$ is always \emph{compatible}; and \emph{liveness}, stating that for any proposed command $c$ and learner $l$, a c-struct of the form $s\bullet c$ is eventually learned by $l$.
 Note that the FLP theorem~\cite{FischerLynchPaterson83ImpossibilityDistributedConsensusOneFaultyProcess} implies that the liveness property cannot be achieved in a networked system where processes may fail.
@@
 To specify the composition interface, we introduced a third type of processes called the \emph{switchers}.
-Switchers can receive \emph{init values} from a preceding aborting modules and can pass \emph{abort values} to pass the baton to a next SMR module when the current SMR module aborts.
+Switchers can receive \emph{init c-structs} from a preceding aborting modules and can pass \emph{abort c-structs} to pass the baton to a next SMR module when the current SMR module aborts.
 
-We describe the behavior of switchers using the variable $from$, which contains all the init values received by any switcher, and the variable $to$, which contains all the abort values produced by the switchers.
+We describe the behavior of switchers using the variable $from$, which contains all the init c-structs received by any switcher, and the variable $to$, which contains all the abort c-structs produced by the switchers.
 The variable $prop$ is used as before, but the variable $learned$ now maps learners to sets of c-structs and, for each learner $l$, $learned\left[ l \right]$ consists of all the c-structs learned so far by $l$.
-We define the set of valid c-structs as the c-structs of the form $t\star cs$ where $cs$ is a sequence of proposed commands, and $t$ is the greatest lower bound of a non-empty subset of the init values received.
+We define the set of valid c-structs as the c-structs of the form $t\star cs$ where $cs$ is a sequence of proposed commands, and $t$ is the greatest lower bound of a non-empty subset of the init c-structs received.
 
-Composable Generalized Consensus requires that no value be learned unless at least one init value have been received (Initialization); that for any learner $l$, $learned\left[ l \right]$ is valid (Nontriviality); that $\left\{ learned\left[ l \right] :  l \in L\right\}$ is always compatible (Agreement); that every abort value is valid, that for any abort value $s_a$ and learned value $s_l$ we have $s_l \preceq s_a$ (Safe Abort); and that for any proposed command $c$ and learner $l$, a c-struct
-of the form $s\bullet c$ is eventually learned by $l$ or a switcher eventually produces an abort value (Liveness).
-A TLA$^+$ specification of CGC appears in \cref{fig:cgc}.
+Composable Generalized Consensus requires the following properties: \emph{initialization}, stating that no value be learned unless at least one init c-struct have been received; \emph{non-triviality}, stating that for any learner $l$, $learned\left[ l \right]$ is valid; \emph{agreement}, stating that $\left\{ learned\left[ l \right] :  l \in L\right\}$ is always compatible; \emph{safe-abort}, stating that every abort c-struct is valid, that for any abort c-struct $s_a$ and learned value $s_l$ we have
+$s_l \preceq s_a$; and \emph{liveness}, stating that for any proposed command $c$ and learner $l$, a c-struct of the form $s\bullet c$ is eventually learned by $l$ or a switcher eventually produces an abort c-struct.
+A TLA$^+$ specification of the safety properties of CGC appears in \cref{fig:cgc}.
 
 We say that an algorithm implementing CGC under some refinement mapping is a \emph{CGC instance}, and that two CGC instances $i_1$ and $i_2$ are compatible if they update disjoint variables except for the $to$ variable of $i_1$ and the $from$ variable of $i_2$, which must be identical.
-Note that in the execution of two compatible instances, any abort value produced by the first instance is an init value for the second instance.
-We say that a CGC instance is an initial instance when its set of init values is always a subset of $\left\{ \bot \right\}$.
+Note that in the execution of two compatible instances, any abort c-struct produced by the first instance is an init c-struct for the second instance.
+We say that a CGC instance is an initial instance when its set of init c-structs is always a subset of $\left\{ \bot \right\}$.
 
@@
 
-First note that the following invariants hold of the execution of $Spec$: invariant 1, stating that any learned c-struct of $CGC1$ is smaller than any abort c-struct of $CGC1$; invariant 2, stating that for any non-empty subset $S$ of the abort c-structs of $CGC1$, there is a sequence $cs$ of commands proposed in $CGC1$ such that $\operatorname{GLB}\left( S \right) = \bot\star cs$; and invariant 3, stating that for every learned c-struct or abort c-struct $s$ of $CGC2$, there is a non-empty subset $S$ of the init c-structs of $CGC2$ and a sequence of commands proposed in $CGC2$ such that $s  = \operatorname{GLB}\left( S \right)\star cs$.
+First note that the following invariants hold of the execution of $Spec$: \emph{invariant 1}, stating that any learned c-struct of $CGC1$ is smaller than any abort c-struct of $CGC1$; \emph{invariant 2}, stating that for any non-empty subset $S$ of the abort c-structs of $CGC1$, there is a sequence $cs$ of commands proposed in $CGC1$ such that $\operatorname{GLB}\left( S \right) = \bot\star cs$; and \emph{invariant 3}, stating that for every learned c-struct or abort c-struct $s$ of $CGC2$, there is a non-empty subset $S$ of the init c-structs of $CGC2$ and a sequence of commands proposed in $CGC2$ such that $s  = \operatorname{GLB}\left( S \right)\star cs$.
 Invariant 1 is enforced by the first conjunct of the $Learn(l)$ transition of $CGC1$, and invariant 3 is enforced by the quantifier bound of the $Learn(l)$ and $Abort$ transition of $CGC2$, using the definition of $Valid$.
@@
 
-We now prove that each transition of $Spec$ is a transition of $CGC!Spec$:
-The transition $CGC1!Propose$ inserts a command $c$ in the set $CCG1!prop$, and, since $CGC!prop = CGC1!prop \cup CGC2!prop$, the transition inserts $c$ in the set $CGC!prop$; therefore, $CGC1!Propose$ implies $CGC!Propose$.
-By an analogous reasoning, $CGC2!Propose$ implies $CGC!Propose$.
-The transition $CGC1!InitM$ inserts $\bot$ in the set $CGC1!from$ (because $CGC1$ is an initial instance), and, since $CGC!prop = CGC1!prop$, the transition also inserts $\bot$ in $CGC!prop$; therefore, $CGC1!InitM$ implies $CGC!InitM$.
+We now prove that each transition of $Spec$ is a transition (i.e.\ implies a transition) of $CGC!Spec$, analysing the transition one by one.
+Note that the transition $CGC1!Propose$ inserts $c$ in the set $CGC!prop$, because $CGC!prop = CGC1!prop \cup CGC2!prop$; therefore, $CGC1!Propose$ implies $CGC!Propose$.
+Similarly, the $CGC2!Propose$ implies $CGC!Propose$.
+Note that the transition $CGC1!InitM$ inserts $\bot$ in the set $CGC!from$, because $CGC1$ is an initial instance and $CGC!from = CGC1!from$; therefore, $CGC1!InitM$ implies $CGC!InitM$.
 The $CGC1!Abort$ transition does not modify any variable of the module instance $CGC$; therefore, $CGC1!Abort$ is a stuttering transition of $CGC$.
-Similarly, the $CGC2!InitM$ transition does not modify any variable of the module instance $CGC$; therefore, $CGC2!InitM$ is a stuttering transition of $CGC$.
+Similarly, the $CGC2!InitM$ transition is a stuttering transition of $CGC$.
+
+Note that the $CGC1!Learn(l)$ transition inserts a c-struct $s$ in the set $CGC!learned\left[ l \right]$, because $CGC!leanred[l] = CGC1!learned\left[ l \right] \cup CGC2!learned\left[ l \right]$; therefore, to show that $CGC1!Learn\left( l \right)$ implies $CGC!Learn\left( l \right)$, we must show (1), that $s$ is smaller than any abort c-struct of $CGC$, (2), that $s$ is compatible with any learned c-struct of $CGC$, and (3), that $s$ is valid in $CGC$.
+To show (1), consider an abort c-struct $a$ of $CGC2$.
+By invariant 3, $a$ is of the form $\operatorname{GLB}\left( S \right)\star cs$ where $S$ is a subset of the init c-structs of $CGC2$.
+Therefore, $\operatorname{GLB}\left( S \right)\preceq a$.
+Moreover, $s$ is smaller than any init c-struct of $CGC2$, by invariant 1.
+Therefore, $\operatorname{GLB}\left( S \right)$ is bigger than $s$.
+By transitivity of $\preceq$, we get that $s$ is smaller than $a$.
+To show (2), consider a learned c-struct $s_l$ of $CGC$.
+There are two cases: (2a), $s_l$ is a learned c-struct of $CGC1$, and (2b), $s_l$ is a learned c-struct of $CGC2$.
+In case (2a), the second conjunct of the $CGC1!Learn(l)$ transition establishes that $s$ and $s_l$ are compatible.
+In case (2b), by reasoning as in (1), we establish that $s \preceq s_l$.
+Therefore, $s$ and $s_l$ are compatible.
+To show (3), observe that any c-struct valid in $CGC1$ is also valid in $CGC2$.
+
+Similarly, the $CGC2!Learn(l)$ transition inserts a valid c-struct $s$ in the set $CGC!learned\left[ l \right]$, and we must show the same 3 properties.
+Property (1) is established by the first conjunct of the $CGC2!Learn\left( l \right)$ transition, using the fact that $CGC!to = CGC2!to$.
+To show (2), consider a learned c-struct of $CGC$.
+There are two cases: (2a), $s_l$ is a learned c-struct of $CGC1$, and (2b), $s_l$ is a learned c-struct of $CGC2$.
+In case (2b), the second conjunct of the $CGC2!Learn(l)$ transition establishes that $s$ and $s_l$ are compatible.
+In case (2a), by reasoning similarly as in TODO use an invariant...
+To show (3)\ldots
 TODO\@: complete.

### 2016-01-11 paper/intro.tex
working on the writeup of the cgc proof.

@@
 the outputs produced by the abort interface of the first SMR module, called
-abort values, as inputs to the init interface of the second SMR module,
-called init values in the context of the second SMR module. A SMR module
+abort c-structs, as inputs to the init interface of the second SMR module,
+called init c-structs in the context of the second SMR module. A SMR module
 must satisfy a correctness condition that we call Composable Generalized
@@
 requires a SMR module to make progress executing commands or to abort its
-execution. CGC also relates the init values, learned values (learned values
+execution. CGC also relates the init c-structs, learned values (learned values
 represent the local state-machine state of individual replicas), and abort
@@
 These two specifications both implement CGC and provide a skeleton for a user to build her own module.
-The two specifications demonstrate how to update the local state of a replica after gathering the states or the init values of a write quorum of replicas, and how to
-produce an abort value after gathering the state of a read quorum of replicas.
+The two specifications demonstrate how to update the local state of a replica after gathering the states or the init c-structs of a write quorum of replicas, and how to
+produce an abort c-struct after gathering the state of a read quorum of replicas.
 Each of the specifications achieves a particular consistency/latency trade-off~\cite{Abadi12ConsistencyTradeoffsModernDistributedDatabaseSystem}.

### 2016-01-14 paper/cgc.tex
Work on the paper, with the goal of presenting it to colleagues.

@@
 However, the intent of replication is to mask faults while preserving the behavior of the service from the point of view of its clients.
+Therefore, replicas must be somehow synchronized to act as a centralized implementation.
 
@@
 
-Generalized Consensus~\cite{Lamport05GeneralizeConsensus} is a formal specification of the requirements that the distributed algorithm $A$ must meet in order for SMR to provide the illusion of a centralized fault-tolerant service.
-Generalized Consensus uses the notion of a set of command-structures, or c-structs.
+Composable Generalized Consensus is based on Generalized Consensus, and any CGC algorithm can be used as a Generalized Consensus algorithm.
+Therefore, we start by a review of generalized and present CGC in the next subsection.
+
+Generalized Consensus~\cite{Lamport05GeneralizeConsensus} is a formal specification of the requirements that the distributed algorithm $A$ described above must meet in order for SMR to provide the illusion of a centralized fault-tolerant service.
+Generalized Consensus uses the notion of command-structures, or c-structs.
 C-structs are obtained by appending commands to the c-struct $\bot$.
@@
 By appropriately defining the operator $\bullet$ according to the
-definition of $M$, C-structs can describe sequences of commands up the reordering of commuting
-commands: it is possible to have $s \star \left[c_1,c_2\right] = s \star \left[c_2,c_1\right]$.
-Moreover, c-structs must satisfy a number of properties: \setword{property 1}{prop:1}, c-structs must form a partial order under the operator $\preceq$ defined as: $s_1 \preceq s_2$ if and only if there exists a sequence of commands $cs$ such that $s_2 = s_1 \star cs$; \setword{property 2}{prop:2}, the greatest lower bound of a set of c-structs $S$ must always exists and be constructible from the same commands as the c-structs in $S$; and \setword{property 3}{prop:3}, a set of c-structs $S$
-which has an upper bound, called a \emph{compatible} set, must have a lowest upper bound which is constructible from the same commands as the c-structs in $S$.
+definition of the state-machine $M$, c-structs can stand for sequences of commands up the reordering of commuting
+commands; for example, we can have $s \star \left[c_1,c_2\right] = s \star \left[c_2,c_1\right]$ for some c-struct $s$ and two commands $c_1$ and $c_2$.
+
+C-structs must satisfy a number of properties: \setword{property 1}{prop:1}, c-structs must form a partial order under the operator $\preceq$ defined as: $s_1 \preceq s_2$ if and only if there exists a sequence of commands $cs$ such that $s_2 = s_1 \star cs$; \setword{property 2}{prop:2}, the greatest lower bound of a set of c-structs $S$ must always exist and be constructible from the same commands as the c-structs in $S$; and \setword{property 3}{prop:3}, a set of c-structs $S$
+which has an upper bound, called a \emph{compatible} set, must have a least upper bound which is constructible from the same commands as the c-structs in $S$.
 
@@
 Proposers propose commands and learners, which represent replicas, learn about c-structs.
-We describe the execution of learners and proposers using the variable $prop$, which is a set containing all the commands that have been proposed, and the variable $learned$, which maps a learner to the latest c-structs it has learned.
 
-Generalized Consensus requires the following properties: \emph{non-triviality}, stating that for any learner $l$, there is at all times a sequence of proposed commands $cs$ such that $learned\left[l\right] = \bot \bullet cs$; \emph{stability}, stating that for any learner $l$, $learned\left[ l \right]$ increase with time; \emph{agreement}, stating that $\left\{ learned\left[ l \right] :  l \in L\right\}$ is always \emph{compatible}; and \emph{liveness}, stating that for any proposed command $c$ and learner $l$, a c-struct of the form $s\bullet c$ is eventually learned by $l$.
+We specify the execution of learners and proposers in terms of a transition system consisting of the variable $prop$, which is a set containing all the commands that have been proposed, and the variable $learned$, which maps a learner to the latest c-structs it has learned.
+
+Generalized Consensus requires the following properties: \emph{non-triviality}, stating that for any learner $l$, there is at all times a sequence of proposed commands $cs$ such that $learned\left[l\right] = \bot \bullet cs$; \setword{\emph{stability}}{prop:stability}, stating that for any learner $l$, $learned\left[ l \right]$ increase with time; \emph{agreement}, stating that $\left\{ learned\left[ l \right] :  l \in L\right\}$ is always \emph{compatible}; and \emph{liveness}, stating that for any proposed command $c$ and learner $l$, a c-struct of the form $s\bullet c$ is eventually learned by $l$.
 Note that the FLP theorem~\cite{FischerLynchPaterson83ImpossibilityDistributedConsensusOneFaultyProcess} implies that the liveness property cannot be achieved in a networked system where processes may fail.
@@
 Composable Generalized Consensus augments the Generalized Consensus specification with an abort interface that allows an SMR module to abort its execution, and with an init interface which allows an SMR module to start its execution where a preceding aborted module left off.
-To specify the composition interface, we introduced a third type of processes called the \emph{switchers}.
-Switchers can receive \emph{init c-structs} from a preceding aborting modules and can pass \emph{abort c-structs} to pass the baton to a next SMR module when the current SMR module aborts.
+To specify the composition interface, we introduce a third type of processes called the \emph{switchers}.
+Switchers can receive \emph{init c-structs} from a preceding aborting module and can pass \emph{abort c-structs} to a next SMR module, which will treat them as init c-structs, when the current SMR module aborts.
 
-We describe the behavior of switchers using the variable $from$, which contains all the init c-structs received by any switcher, and the variable $to$, which contains all the abort c-structs produced by the switchers.
+We describe the behavior of switchers using the variable $from$, which contains all the init c-structs received by the switchers, and the variable $to$, which contains all the abort c-structs produced by the switchers.
 The variable $prop$ is used as before, but the variable $learned$ now maps learners to sets of c-structs and, for each learner $l$, $learned\left[ l \right]$ consists of all the c-structs learned so far by $l$.
@@
 
-Composable Generalized Consensus requires the following properties: \emph{initialization}, stating that no value be learned unless at least one init c-struct have been received; \emph{non-triviality}, stating that for any learner $l$, $learned\left[ l \right]$ is valid; \emph{agreement}, stating that $\left\{ learned\left[ l \right] :  l \in L\right\}$ is always compatible; \emph{safe-abort}, stating that every abort c-struct is valid, that for any abort c-struct $s_a$ and learned value $s_l$ we have
+Composable Generalized Consensus requires the following properties: \emph{initialization}, stating that no value be learned unless at least one init c-struct has been received; \emph{non-triviality}, stating that for any learner $l$, $learned\left[ l \right]$ is valid; \emph{agreement}, stating that $\left\{ learned\left[ l \right] :  l \in L\right\}$ is always compatible; \emph{safe-abort}, stating that every abort c-struct is valid, and that for any abort c-struct $s_a$ and learned value $s_l$ we have
 $s_l \preceq s_a$; and \emph{liveness}, stating that for any proposed command $c$ and learner $l$, a c-struct of the form $s\bullet c$ is eventually learned by $l$ or a switcher eventually produces an abort c-struct.
@@
 
-We say that an algorithm implementing CGC under some refinement mapping is a \emph{CGC instance}, and that two CGC instances $i_1$ and $i_2$ are compatible if they update disjoint variables except for the $to$ variable of $i_1$ and the $from$ variable of $i_2$, which must be identical.
-Note that in the execution of two compatible instances, any abort c-struct produced by the first instance is an init c-struct for the second instance.
+We say that an algorithm implementing CGC under some refinement mapping is a \emph{CGC instance}, and that two CGC instances $i_1$ and $i_2$ are compatible if they update disjoint variables except for the $to$ variable of $i_1$ and the $from$ variable of $i_2$ which must be identical, therefore identifying the abort c-struct of $i_1$ with the init c-structs of $i_2$.
+%Note that in the execution of two compatible instances, any abort c-struct produced by the first instance is an init c-struct for the second instance.
 We say that a CGC instance is an initial instance when its set of init c-structs is always a subset of $\left\{ \bot \right\}$.
 
-CGC enjoys two main properties: first, the composition of two CGC instances is itself a CGC instance when the first instance is an initial instance; second, an initial CGC instance satisfies the safety properties of Generalized Consensus under the restrictions explained below.
+CGC enjoys two main properties: first, the composition of two CGC instances is itself a CGC instance if the first instance is an initial instance; second, an initial CGC instance satisfies the safety properties of Generalized Consensus under the restrictions explained below.
 We call the first property the \emph{composition theorem}.
@@
 Instead, we state the composition theorem for two CGC instances obtained by only renaming the variables of the CGC specification.
-Any CGC instance would trivially refine both of the two instances used in formalization of the composition theorem because a CGC instance refines the CGC specification.
+Any CGC instance would trivially refine both of the two instances used in the formalization of the composition theorem.
 Therefore, we obtain the composition theorem by transitivity and monotonicity of refinement~\cite{AbadiLamport91ExistenceRefinementMappings}.
+To show that an initial instance implements Generalized Consensus, we must first wrap it in a small layer that prevents the learned c-struct of a learner from changing non-monotonically.
+This is because the Generalized Consensus specification requires the learned c-struct of a learner to grow monotonically (the \ref{prop:stability} property), whereas CGC allows the learned c-struct of a learner change non-monotonically.
+However, this is not a fundamental difference: when a learner
+learns a new c-struct in CGC, the additional layer simply updates the learned c-struct of a learner to the least upper-bound of its previous value and the value learned in CCG, thereby guaranteeing the stability property of Generalized Consensus.
 
@@
 \label{sec:compproof}
+\begin{comment}
 To prove the composition theorem, as expressed in \cref{fig:compthm}, we must show that $Init$ implies $CGC!Init$ and that every transition of $Spec \equiv Init \wedge \box\left[Next\right]_{vars}$ is a transition of $CGC!Spec$.
@@
 In case (2b), the second conjunct of the $CGC2!Learn(l)$ transition establishes that $s$ and $s_l$ are compatible.
-In case (2a), by reasoning similarly as in TODO use an invariant...
+In case (2a), by reasoning similarly as in TODO use an invariant.
 To show (3)\ldots
@@
 \subsection{Agreeing on the Next Module}
+
+\end{comment}

### 2016-01-14 paper/intro.tex
Work on the paper, with the goal of presenting it to colleagues.

@@
 
-Surprisingly, at a first glance, many SMR algorithms seem to share a
-common structure which would indicate that previous formal analysis could
-be reused. That is because, in many cases, researchers obtained a new SMR
-algorithm by optimizing an existing SMR algorithm for a particular use
+Surprisingly, at a first glance, many SMR algorithms seem to share a common
+structure which would indicate that previous formal analysis could be reused.
+This similarity in structure exists because, in many cases, researchers obtained
+a new SMR algorithm by optimizing an existing SMR algorithm for a particular use
 case. For example, Fast Paxos~\cite{Lamport06FastPaxos} adds fast rounds
@@
 systems~\cite{GuerraouiVukolic10RefinedQuorumSystems,Lamport06FastPaxos},
-and agreement up to commutativity using command-structure
-sets~\cite{Lamport05GeneralizeConsensus}.
+and agreement up to commutativity based on command-structures~\cite{Lamport05GeneralizeConsensus}.
 
-Given their similarities, one may ask why is it not possible to reuse formal
-analysis of SMR algorithms. The problem is that SMR algorithms are not modular
-and that, because of their intricate structure, modifying a monolithic SMR
-algorithm may produce many unforeseen corner cases that invalidate previous
-analyses.
+Given the similarities shared by many SMR algorithms, one may ask why is it
+not possible to reuse formal analyses to simplify the verification of new
+algorithms. The problem is that SMR algorithms are not modular and that, because
+of their intricate structure, modifying a monolithic SMR algorithm may produce
+many unforeseen corner cases that invalidate previous analyses.
 
-To simplify the task of devising new SMR algorithms we propose a framework
-for building reusable SMR modules. For example, in \cref{}, we augment
-MultiPaxos with our composition interface and prove that it satisfies Composable
-Generalized Consensus (CGC), our correctness condition for modules. We then
-build an SMR module whose structure is similar to the one of the fast rounds of
-Fast Paxos and prove that it satisfies CGC\@. Our Fast Paxos module has lower
-latency than our Paxos module when there is low contention among clients, but
-its performance drops below that of Paxos when contention increases. Finally,
-we assemble unmodified our MultiPaxos SMR module and our Fast Paxos SMR module
-in an SMR algorithm that dynamically switch from one to the other depending on
-the contention level observed in the system, therefore getting the best of both.
-Crucially, we obtain the correctness of our composite SMR algorithm for free:
-our generic composition theorem and the correctness of our two SMR modules taken
-individually imply that the combination of the MultiPaxos SMR module and the
-Fast Paxos SMR module is a correct SMR algorithm. Later, in \cref{}, we present
-three other SMR modules that can be combined the MultiPaxos SMR module and the
-Fast Paxos SMR module in a composite SMR algorithm that can switch between 5
-different kinds of SMR modules, according to the system's operating conditions.
+To simplify the task of devising new SMR algorithms we propose a framework for
+building self-contained and reusable SMR modules whose analysis is independent
+of the other modules they may be composed with. For example, in \cref{}, we
+augment MultiPaxos with our composition interface, and we prove that it satisfies
+our correctness condition for modules, called Composable Generalized Consensus (CGC).
+We then build an SMR module whose structure is similar to the one of the fast
+rounds of Fast Paxos, and we prove that it satisfies CGC\@. Finally, we assemble
+unmodified our MultiPaxos SMR module and our Fast Paxos SMR module in an SMR
+algorithm that dynamically switches from one to the other depending on the
+contention level observed in the system. This composite algorithm combines the
+advantages of both its SMR modules: in cases of low contention, our Fast Paxos
+module executes commands in two message delays, while MultiPaxos would need
+three, whereas the MultiPaxos module maintains its throughput in high contention
+cases, in which the performance of the Fast Paxos module would have been
+significantly reduced.
 
-An SMR module is an SMR algorithm augmented with a \emph{composition interface}
-and satisfying our correctness condition, called Composable Generalized
-Consensus (CGC). The composition interface consists of an \emph{abort interface}
+Crucially, we obtain the correctness of our composite
+SMR algorithm for free: our generic composition theorem and the correctness
+of our two SMR modules taken individually imply that the combination of the
+MultiPaxos SMR module and the Fast Paxos SMR module is a correct SMR algorithm.
+Later, in \cref{}, we present three other SMR modules that can be combined the
+MultiPaxos SMR module and the Fast Paxos SMR module in a composite SMR algorithm
+that can switch between 4 different kinds of SMR modules, according to the
+system's operating conditions.
+
+More generally, an SMR module is an SMR algorithm augmented with a \emph{composition interface}
+and satisfying a correctness condition that we call \emph{Composable Generalized
+Consensus}, abbreviated CGC\@. The composition interface consists of an \emph{abort interface}
 for aborting an SMR module's execution and passing the baton to another
@@
 the outputs produced by the abort interface of the first SMR module, called
-abort c-structs, as inputs to the init interface of the second SMR module,
-called init c-structs in the context of the second SMR module. A SMR module
-must satisfy a correctness condition that we call Composable Generalized
-Consensus (abbreviated CGC). Any correct SMR module is a safe Generalized
+\emph{abort c-structs}, as inputs to the init interface of the second SMR module,
+called \emph{init c-structs} in the context of the second SMR module. Any correct SMR module is a safe Generalized
 Consensus~\cite{Lamport05GeneralizeConsensus} algorithm (i.e.\ a kind of SMR
 algorithm), but the progress guarantees may vary between SMR modules: CGC only
-requires a SMR module to make progress executing commands or to abort its
-execution. CGC also relates the init c-structs, learned values (learned values
+requires an SMR module to make progress executing commands or to abort its
+execution. CGC also relates the init c-structs, learned values (learned c-structs 
 represent the local state-machine state of individual replicas), and abort
-values in order for our composition theorem to hold.
-%Note that Generalized Consensus is a specification of SMR algorithms that takes into account the commutativity relation among commands and can offer important performance advantages in practice.
+c-structs in order to guarantee that an SMR module consistently takes over the execution of its predecessor, as expressed in our composition theorem.
+Finally, our framework allows building \emph{dynamic SMR
+algorithms}, which dynamically change module when the currently running module
+aborts, choosing the module best suited to the current operating conditions.
 
 Our composition theorem enables incremental development of SMR algorithms
-because it guarantees that any number of correct modules may be composed in
-a sequence to form a correct SMR algorithm. Therefore, our framework allows
-building \emph{dynamic SMR algorithms}, which dynamically change module when
-the currently running module aborts, choosing the module best suited to the
-current operating conditions. Thanks to the composition theorem, new modules can
-be added to a dynamic SMR algorithm without modifying the existing modules and
-with guaranteed correctness. Moreover, the relaxed progress requirements of CGC
-enable a separation of concerns that simplifies optimizing SMR algorithms for
-particular scenarios because there is no need to consider executing commands
-under a situation outside the scenario: the module can just abort in this case.
+because it guarantees that any number of correct modules may be composed in a
+sequence to form a correct SMR algorithm. Thanks to the composition theorem,
+new modules can be added to a dynamic SMR algorithm without modifying the
+existing modules and with guaranteed correctness. Moreover, the relaxed progress
+requirements of CGC enable a separation of concerns that simplifies optimizing
+SMR algorithms for particular scenarios because there is no need to consider
+executing commands under a situation outside the scenario: the module can just
+abort in this case.
 
@@
 The user can therefore specify a concrete command dissemination procedure and then model-check that the obtained algorithm refines one of the two specifications.  
-We demonstrate the practical use of our framework by presenting modules based on Paxos~\cite{Lamport98ParttimeParliament}, Fast Paxos\cite{Lamport06FastPaxos}, Chain Replication~\cite{RenesseSchneider04ChainReplicationSupportingHighThroughputAvailability}, a case of Vertical Paxos~\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}, and EPaxos~\cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments}
-(TODO, but would be nice).
+We demonstrate the practical use of our framework by presenting modules based on Paxos~\cite{Lamport98ParttimeParliament}, Fast Paxos\cite{Lamport06FastPaxos}, Chain Replication~\cite{RenesseSchneider04ChainReplicationSupportingHighThroughputAvailability}, a case of Vertical Paxos~\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}.
 
@@
 \begin{itemize}
-    \item We present the first framework to allowing to build Generalized Consensus algorithms our of independent modules.
-    \item We provide tool support for designing CGC modules in the form of two TLA+~\cite{Lamport02SpecifyingSystems} specifications simplifying the design and model-checking of SMR algorithms based on Quorum Systems~\cite{GuerraouiVukolic10RefinedQuorumSystems} and demonstrate their use by designing CGC modules inspired by Paxos, Fast Paxos, and Primary-Backup replication.
+    \item We present the first framework allowing to build Generalized Consensus algorithms our of independent and reusable modules supporting modular formal analysis.
+    \item We provide tool support for designing CGC modules in the form of two TLA+~\cite{Lamport02SpecifyingSystems} specifications simplifying the design and model-checking of SMR algorithms based on Quorum Systems~\cite{GuerraouiVukolic10RefinedQuorumSystems} and demonstrate their use by designing four CGC modules inspired by MultiPaxos, Fast Paxos, Chain Replication, and Primary-Backup replication.
 \end{itemize}

### 2016-01-14 paper/related.tex
Work on the paper, with the goal of presenting it to colleagues.

@@
 simulation~\cite{LynchVaandrager95ForwardBackwardSimulationsIUntimedSystems})
-modulo an equivalence relation on histories of commands that reflects the
-commutativity relation among commands. However, the simulation technique led to
+modulo an equivalence relation on histories of commands. However, the simulation technique led to
 complex correctness statements that were hard to scale beyond the examples of
@@
 work addresses the shortcoming of the Speculatively Linearizable framework by
-using the notion of command-structure set introduced for the definition of
+using the notion of command-structures introduced for the definition of
 Generalized Consensus~\cite{Lamport05GeneralizeConsensus}. Moreover, we show how
 several well known and practical algorithms can be cast and optimized in our
-framework, providing a set of TLA+ specifications that ease the design and model
-checking of algorithms based on quorum systems.
+framework, providing a set of TLA+ specifications that simplify their design.
 
@@
 needed in the Classic Paxos algorithm.
-
 Generalized Paxos is also structured in a sequence of rounds, 
@@
 However, no explicit interface for combining rounds is given and specifying a type of round in isolation from the others would require assume-guarantee reasoning.
-In contrast, our formulation uses encapsulared rounds and does not require assume-guarantee reasoning.
+In contrast, our formulation uses encapsulated rounds and does not require assume-guarantee reasoning.
 
@@
 replicated service. Vertical Paxos algorithms use an external reconfiguration
-master, itself implemented as a replicated state machine, which determines
-the servers that should be removed from or added to the configuration. The
-role of the reconfiguration master is similar to the role of the scheduling
-policy in our framework. In contrast to our work, Vertical Paxos does not define
-an interface and correctness condition for encapsulating the reconfigurable
-instances of SMR algorithms and therefore does not allow the reuse of existing
-algorithms and the incremental development of SMR optimizations.
+master, itself implemented as a replicated state machine, which determines the
+servers that should be removed from or added to the configuration. In contrast
+to our work, Vertical Paxos does not define an interface and correctness
+condition for encapsulating the reconfigurable instances of SMR algorithms and
+therefore does not allow the reuse of existing algorithms and the incremental
+development of SMR optimizations.
